### 2021-11-26 main.tex
start thinking about challenge problems

% jfpguide.tex
% v1.0, released 19 Nov 2019
% Copyright 2019 Cambridge University Press

\documentclass{jfp}

\usepackage{graphicx}

% macros (move to its own file).

\begin{document}

\journaltitle{JFP}
\cpr{Cambridge University Press}
\doival{10.1017/xxxxx}

% \lefttitle{\LaTeX\ Supplement}
\righttitle{Journal of Functional Programming}

\totalpg{\pageref{lastpage01}}
\jnlDoiYr{2018}

\title{POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types}

\begin{authgrp}
\author{Anonymous Authors (even unknown for now)}
\affiliation{Universities\\
  Addresses\\
  (\email{anoymous@example.com})}
\end{authgrp}

\begin{abstract}
  The publication of POPLMark spearheaded the era of publishing formal
  proofs alongside with new research. This fostered an era of
  high-quality and high-assurance proofs at top venues for programming
  language research. This paper simultaneously acknowledges the impact
  of the original POPLMark challenge, and proposes a new benchmark
  intended to stimulate the research and dissemination of techniques
  that address the particular challenges of process calculi and
  behavioural type systems. In this work we propose three challenge
  problems concerning to process calculi and name passing, linearity
  in behavioural type systems, and coinductive reasoning for process
  algebras.
\end{abstract}

\maketitle[F]

\section{Introduction}

In recent years, at programming language venues there is the growing
expectation that the theoretical results of a paper are formalised in
a proof assistant. % rephrase
For this purpose, and to get the community on track the
POPLMark~\cite{Pierce:2005} challenge was proposed. Since then the
number of formal proofs accompanying papers has grown (add some POPL
statistics here). However, formal proofs are not equally spread in any
field, and the key contribution of challenges like POPLMark is to help
develop the necessary momentum and culture to be able to formalise
results. It has been a long time since the original challenge, and
many changes have happened. Particularly, POPLMark and any challenge
cannot address every conceivable technique. Challenges have to be
simple so that the techniques are not shadowed by technical details
and also feasible to implement without heroic efforts. In that vein,
it is apparent that new challenges are needed, work like POPLMark
Reloaded~\cite{Pientka:2019} strives to extend the scope of the
original challenge to proofs using logical relations. This technique
is crucial to many current developments.

In this work, we address another expansion of scope for such
challenges, going from the $\lambda$-calculus to process calculi like
the $\pi$-calculus and session type systems (and other behavioural
type systems). These systems require particular insights to implement.
We have identified three key aspects. First, the idea of scope
extrusion in name passing calculi. This generalises the idea of
binders in ways that techniques that were adequate for
$\lambda$-calculus binders are not so for name passing systems.
Second, session types in particular, and behavioural types in general
often depend on linearity in their type systems. The formalisation of
linear type systems posses a challenge given that it is not uniformly
well supported among the many proof assistants. Finally, processes in
concurrent systems often contain infinite behaviour that is modelled
with coinduction. This technique is used to reason about
(bi)similarity between processes, the equivalence of their traces, or
subtyping between infinite trees of possible behaviours. This
challenge concentrates on those issues.

In this work we propose a challenge where each part independently
exercises one of the three areas, however it is often the case that a
problem requires more than one of these techniques simultaneously. In
this work, we explore them in isolation, in order to facilitate the
implementation of smaller solutions. Concretely, we do not address the
idea of how to combine these techniques, leaving the discussion of
that to the future when we have collected enough entries to these
problems.

We do not claim in this work that these problems have never been
approached and formalised before. In fact, the contrary is true. All
of these problems have been addressed in several forms. This motivates
this challenge to enable an easier comparison between solutions (since
they all implement the same challenge), and in a setting that is
simple enough to not be distracting, yet complex enough that we can
compare strengths and weaknesses of each.

The rest of this paper is structured in the following way:
\cref{sec:rationale} expands on the rationale for the selected
challenge problems. Then \cref{sec:problems} presents the problems. We
discuss the solutions in \cref{sec:solutions}, in particular
\cref{sec:tutorial} walks through a tutorial solution. And finally
\cref{sec:conclusion} presents some conclusions and it contains a call
to action to invite the submission of new solutions.

\section{Rationale for choosing the problems}\label{sec:rationale}

This section is not (at least initially) meant to be in the final
manuscript but it is to document some reasoning for choosing
appropriate problems. The idea is to explore three areas, scope
extrusion in name passing calculi, linearity requirements in
behavioural types, and finally coinductive reasoning, in trace
equivalence, (bi)similarity, or subtyping in the presence of
recursion.

The idea is to have three small problems in the aforementioned
subjects that are independent of each other and that are easy to
understand and to implement. This is important so we have more
entries. They have to be small so it is not a lot to read, and
independent so people may choose to implement only one of them, or
work on them in any order.

Finally, it is important to mention in the rationale for problems that
the key aspects are:
\begin{itemize}
\item Scope extrusion
\item Linearity
\item Coinduction
\end{itemize}

But the exact presentation of the concrete problems should be somewhat
flexible. In the sense that the use of certain techniques or tools may
require changing the presentation of the system to fit the technique.
These changes should be allowed, and even encouraged, but the result
should be ``obviously'' the same system and there should be a
rationale for the changes in presentation. That way we could not only
see valid techniques but also what presentation of the rules to use
and the rationale to chose a particular approach.

\section{The challenge problems} \label{sec:problems}

\subsection{Name passing and scope extrusion}

\textbf{Problem idea}: A small proof about the untyped $\pi$-calculus.

\subsection{Linearity and behavioural type systems}

\textbf{Problem idea}: A small session type problem.

\subsection{Coinduction and reasoning about process algebras}

\textbf{Problem idea}: Something about traces, (bi)similarity, or
subtyping under non terminating problems in behavioural types.

\section{The solutions}\label{sec:solutions}

\subsection{A tutorial solution on paper}\label{sec:tutorial}

Probably the skeleton here and more details on an appendix.

\subsection{Flexibility in solving a challenge}

\section{Conclusion and related work} \label{sec:conclusion}

\bibliographystyle{JFPlike}
\bibliography{references}

\label{lastpage01}

\end{document}

### 2022-01-25 main.tex
add figure to start planning the concrete benchmark

@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\textwidth]{images/benchmark.jpeg}
+  \caption{The benchmark's big picture.}
+  \label{fig:bigpic}
+\end{figure}
+
+The big picture for the benchmark can be seen in \cref{fig:bigpic}
+where the three pillars (i.e.: concepts typically needed for process
+calculus proofs) appear in blue, with some ideas of what to prove. And
+a second step of combining the three aspects (for a total of six steps
+for the whole benchmark), and of course the $+1$-step for the
+combination of everything on the reader's next research papers.
+
 \subsection{Name passing and scope extrusion}

### 2022-06-08 main.tex
make the image smaller
change the format to a simpler article
rework on the draft

@@
-% jfpguide.tex
-% v1.0, released 19 Nov 2019
-% Copyright 2019 Cambridge University Press
-
-\documentclass{jfp}
+\documentclass{article}
 
 \usepackage{graphicx}
+\usepackage{cleveref}
+\usepackage{natbib}
 
-% macros (move to its own file).
+% macros (move to its own file when we have a macro).
 
@@
 
-\journaltitle{JFP}
-\cpr{Cambridge University Press}
-\doival{10.1017/xxxxx}
-
-% \lefttitle{\LaTeX\ Supplement}
-\righttitle{Journal of Functional Programming}
-
-\totalpg{\pageref{lastpage01}}
-\jnlDoiYr{2018}
-
 \title{POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types}
 
-\begin{authgrp}
-\author{Anonymous Authors (even unknown for now)}
-\affiliation{Universities\\
-  Addresses\\
-  (\email{anoymous@example.com})}
-\end{authgrp}
+\maketitle
 
@@
 
-\maketitle[F]
-
 \section{Introduction}
 
-In recent years, at programming language venues there is the growing
-expectation that the theoretical results of a paper are formalised in
-a proof assistant. % rephrase
-For this purpose, and to get the community on track the
-POPLMark~\cite{Pierce:2005} challenge was proposed. Since then the
-number of formal proofs accompanying papers has grown (add some POPL
-statistics here). However, formal proofs are not equally spread in any
-field, and the key contribution of challenges like POPLMark is to help
-develop the necessary momentum and culture to be able to formalise
-results. It has been a long time since the original challenge, and
-many changes have happened. Particularly, POPLMark and any challenge
-cannot address every conceivable technique. Challenges have to be
-simple so that the techniques are not shadowed by technical details
-and also feasible to implement without heroic efforts. In that vein,
-it is apparent that new challenges are needed, work like POPLMark
+In recent years, the programming language community has embraced
+machine checked proofs for theoretical results in papers. The
+POPLMark~\cite{Pierce:2005} challenge was proposed for this purpose,
+and to get the community on track. Since then the number of formal
+proofs accompanying papers has grown\footnote{add some POPL/PLDI/etc
+  statistics here}. However, formal proofs are not equally spread in
+the field. In this call to action, we propose to extend the key
+contribution of challenges like POPLMark is to help develop the
+necessary momentum and culture to be able to formalise results to the
+concurrent calculi community by proposing a challenge the exercises
+the key aspects of reasoning about process calculi.
+
+It has been a long time since the original challenge, and many changes
+have happened. Particularly, POPLMark and any challenge cannot address
+every conceivable technique. Challenges have to be simple so that the
+techniques are not shadowed by technical details and also feasible to
+implement without heroic efforts. In that vein, it is apparent that
+new challenges are needed, work like POPLMark
 Reloaded~\cite{Pientka:2019} strives to extend the scope of the
@@
 
-In this work, we address another expansion of scope for such
-challenges, going from the $\lambda$-calculus to process calculi like
-the $\pi$-calculus and session type systems (and other behavioural
-type systems). These systems require particular insights to implement.
-We have identified three key aspects. First, the idea of scope
-extrusion in name passing calculi. This generalises the idea of
+As mentioned before, in this work we address another expansion of
+scope for such challenges, going from the $\lambda$-calculus to
+process calculi such as the $\pi$-calculus and session type systems
+(and other behavioural type systems). These systems require particular
+insights to implement. So far, we have identified three key
+aspects\footnote{We want to keep the number low, but the exact nature
+  of the challenges is still open to discussion.}. First, the idea of
+scope extrusion in name passing calculi. This generalises the idea of
 binders in ways that techniques that were adequate for
-$\lambda$-calculus binders are not so for name passing systems.
-Second, session types in particular, and behavioural types in general
-often depend on linearity in their type systems. The formalisation of
-linear type systems posses a challenge given that it is not uniformly
-well supported among the many proof assistants. Finally, processes in
-concurrent systems often contain infinite behaviour that is modelled
-with coinduction. This technique is used to reason about
-(bi)similarity between processes, the equivalence of their traces, or
-subtyping between infinite trees of possible behaviours. This
-challenge concentrates on those issues.
+$\lambda$-calculus binders are not so convenient for name passing
+systems. Second, session types in particular, and behavioural types in
+general often depend on linearity in their type systems. The
+formalisation of linear type systems posses a challenge given that it
+is not uniformly well supported among the many proof assistants.
+Finally, processes in concurrent systems often contain infinite
+behaviour that is typically modelled with coinduction. This technique
+is used to reason about (bi)similarity between processes, the
+equivalence of their traces, or subtyping between infinite trees of
+possible behaviours. This challenge concentrates on those issues.
 
@@
 they all implement the same challenge), and in a setting that is
-simple enough to not be distracting, yet complex enough that we can
-compare strengths and weaknesses of each.
+\emph{simple enough to not be distracting, yet complex enough that we can
+compare strengths and weaknesses of each}.
 
@@
 manuscript but it is to document some reasoning for choosing
-appropriate problems. The idea is to explore three areas, scope
+appropriate problems. The idea is to explore three areas: scope
 extrusion in name passing calculi, linearity requirements in
-behavioural types, and finally coinductive reasoning, in trace
+behavioural types, and finally coinductive reasoning in trace
 equivalence, (bi)similarity, or subtyping in the presence of
@@
 independent so people may choose to implement only one of them, or
-work on them in any order.
+work on them in any order. Moreover, the specification should only
+need to be met in spirit. The idea is to provide a solution author
+with some space to show the best aspects of their solution even if
+that means adapting the problem definition while keeping the spirit of
+the challenge.
 
-Finally, it is important to mention in the rationale for problems that
+Finally, it will be important to mention in the rationale for problems that
 the key aspects are:
@@
 These changes should be allowed, and even encouraged, but the result
-should be ``obviously'' the same system and there should be a
+should be ``morally'' the same system and there should be a
 rationale for the changes in presentation. That way we could not only
@@
   \centering
-  \includegraphics[width=\textwidth]{images/benchmark.jpeg}
+  \includegraphics[width=0.75\textwidth]{images/benchmark.jpeg}
   \caption{The benchmark's big picture.}
@@
 
-\textbf{Problem idea}: A small session type problem.
+\textbf{Problem idea}: A small session type problem that exposes linear channels.
 
@@
 
+\dots
+
 \section{Conclusion and related work} \label{sec:conclusion}
 
-\bibliographystyle{JFPlike}
+\bibliographystyle{plainnat}
 \bibliography{references}

### 2022-06-08 main.tex
add link to the website

@@
 \usepackage{graphicx}
+\usepackage{natbib}
+\usepackage{hyperref}
 \usepackage{cleveref}
-\usepackage{natbib}
 
@@
 concurrent calculi community by proposing a challenge the exercises
-the key aspects of reasoning about process calculi.
+the key aspects of reasoning about process calculi. We propose this
+document, and a \href{https://concurrentbenchmark.github.io/}{website}
+to enable the community to collaborate on the project.
 
@@
 aspects\footnote{We want to keep the number low, but the exact nature
-  of the challenges is still open to discussion.}. First, the idea of
-scope extrusion in name passing calculi. This generalises the idea of
-binders in ways that techniques that were adequate for
-$\lambda$-calculus binders are not so convenient for name passing
-systems. Second, session types in particular, and behavioural types in
-general often depend on linearity in their type systems. The
-formalisation of linear type systems posses a challenge given that it
-is not uniformly well supported among the many proof assistants.
-Finally, processes in concurrent systems often contain infinite
-behaviour that is typically modelled with coinduction. This technique
-is used to reason about (bi)similarity between processes, the
-equivalence of their traces, or subtyping between infinite trees of
-possible behaviours. This challenge concentrates on those issues.
+  of the challenges is still open to discussion (as are most things in
+  this project).}. First, the idea of scope extrusion in name passing
+calculi. This generalises the idea of binders in ways that techniques
+that were adequate for $\lambda$-calculus binders are not so
+convenient for name passing systems. Second, session types in
+particular, and behavioural types in general often depend on linearity
+in their type systems. The formalisation of linear type systems posses
+a challenge given that it is not uniformly well supported among the
+many proof assistants. Finally, processes in concurrent systems often
+contain infinite behaviour that is typically modelled with
+coinduction. This technique is used to reason about (bi)similarity
+between processes, the equivalence of their traces, or subtyping
+between infinite trees of possible behaviours. This challenge
+concentrates on those issues.
 

### 2023-02-28 challenges/challenges.tex
Add basic description of challenge ideas.

\documentclass[a4paper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{bm}
\usepackage{graphicx}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

\begin{document}

\section{Challenge: Scope Extrusion}
The idea of this challenge is to formalize a proof that requires scope extrusion.
Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the recieving process.
This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.

The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
The objective of this challenge is to prove that barbed bisimulation for this calculus is an equivalence relation.

\section{Challenge: Coinduction}
The idea of this challenge is to formalize a proof that requires coinductive techniques.
Coinduction is a proof technique for infinite structures, which arise in this context due to systems with behaviours that continue indefinitely.
Coinduction is the dual of induction: whereas induction is useful for proving properties of least fixed points, coinduction is useful for proving properties of greatest fixed points.

The setting for this challenge is an untyped calculus of communicating systems with replication of processes, the syntax and semantics of which are presented below.
The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.

\section{Challenge: Linearity}
The setting for this challenge is a restricted \( \pi \)-calculus with a session type system.

The objective of this challenge is to prove type preservation.

\section{Bonus challenges}

\subsection{Linearity and Scope Extrusion}
The setting for this challenge is the \( \pi \)-calculus with a session type system.

The objective of this challenge is to prove type preservation.

\subsection{Scope Extrusion and Coinduction}
The setting for this challenge is a simply typed \( \pi \)-calculus.

The objective of this challenge is to prove bisimulation.

\subsection{Coinduction and Linearity}
The setting for this challenge is the Calculus of Communicating Systems with a session type system.

The objective of this challenge is to prove type preservation.

\section{Future work}
\begin{itemize}
\item Multiparty session types
\item Choreography
\item Encodings
\item Code extraction
\end{itemize}

\end{document}

### 2023-02-28 challenges/challenges.tex
Add description of Scope Extrusion challenge.

@@
 
+\newcommand{\Pend}{\bm{0}}
+\newcommand{\Ppar}[2]{#1 \mid #2}
+\newcommand{\Pres}[2]{(\bm{\nu} #1)~#2}
+\newcommand{\Pout}[3]{#1 ! #2 . #3}
+\newcommand{\Pin}[3]{#1 ? (#2) . #3}
+\newcommand{\Pchoice}[2]{#1 + #2}
+\newcommand{\Preplicate}[1]{{!}#1}
+
+\newcommand{\freenames}[1]{\textrm{fn}(#1)}
+\newcommand{\boundnames}[1]{\textrm{bn}(#1)}
+\newcommand{\names}[1]{\textrm{n}(#1)}
+
+\newcommand{\alphacon}[2]{#1 =_{\alpha} #2}
+\newcommand{\subst}[3]{#1\{#2/#3\}}
+
+\newcommand{\Aoutf}[2]{#1 ! #2}
+\newcommand{\Aoutb}[2]{#1 ! (#2)}
+\newcommand{\Ain}[2]{#1 ? #2}
+\newcommand{\Atau}{\tau}
+
+\newcommand{\transition}[3]{#1 \xrightarrow{#2} #3}
+
+\newcommand{\observable}[2]{#1\downarrow_{#2}}
+\newcommand{\obsin}[1]{#1?}
+\newcommand{\obsout}[1]{#1!}
+
+\newcommand\sbullet[1][.5]{\mathbin{\hbox{\scalebox{#1}{$\bullet$}}}}
+\newcommand{\sbbisim}[2]{#1 \overset{\sbullet}{\sim} #2}
+
 \begin{document}
@@
 The objective of this challenge is to prove that barbed bisimulation for this calculus is an equivalence relation.
+
+\subsection{Syntax}
+We assume the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
+The syntax is then:
+\begin{align*}
+  P,Q :=&&& \Pend \\
+  |&&& \Pout{x}{y}{P} \\
+  |&&& \Pin{x}{y}{P} \\
+  |&&& \Ppar{P}{Q} \\
+  |&&& \Pres{x}{P} \\
+  |&&& \Pchoice{P}{Q}
+\end{align*}
+The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
+The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
+The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
+The input operator thus binds the name \( y \) in \( P \).
+The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+The two components can proceed independently of each other, or they can interact via shared names.
+The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
+Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
+The restriction operator thus binds the name \( x \) in \( P \).
+Note that the scope of a restriction may change when processes interact.
+The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
+Note that there is no recursion or replication in the syntax, and thus no infinite behaviours can be expressed.
+
+We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
+We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
+We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
+
+Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
+As a convention, we will identify \( \alpha \)-convertible processes.
+
+As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
+This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
+
+\subsection{Semantics}
+The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
+The transitions are labelled by \emph{actions}, the syntax of which are as follows:
+\begin{align*}
+  \alpha := &&& \Aoutf{x}{y} \\
+  |&&& \Ain{x}{y} \\
+  |&&& \Aoutb{x}{y} \\
+  |&&& \Atau
+\end{align*}
+The \emph{free output action} \( \Aoutf{x}{y} \) is sending the name \( y \) via \( x \).
+The \emph{input action} \( \Ain{x}{y} \) is receiving the name \( y \) via \( x \).
+The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
+The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
+
+We will again use the notation \( \freenames{\alpha} \) to denote the set of names that occur free in the action \( \alpha \) and the notation \( \boundnames{\alpha} \) to denote the set of names that occur bound in the action \( \alpha \).
+In the free output action \( \Aoutf{x}{y} \) and the input action \( \Ain{x}{y} \), both \( x \) and \( y \) are free names.
+In the bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while \( y \) is a bound name.
+We will further use the notation \( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \) and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur in the action \( \alpha \).
+
+The transition relation is then defined by the following rules:
+\begin{mathpar}
+  \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
+  \and
+  \inferrule[In]{ }{\transition{\Pin{x}{z}{P}}{\Ain{x}{y}}{\subst{P}{y}{z}}}
+  \and
+  \inferrule[Sum-L]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{P}{Q}}{\alpha}{P'}}
+  \and
+  \inferrule[Sum-R]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{Q}{P}}{\alpha}{P'}}
+  \and
+  \inferrule[Par-L]{\transition{P}{\alpha}{P'} \\ \boundnames{\alpha} \cap \freenames{Q} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
+  \and
+  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'} \\ \boundnames{\alpha} \cap \freenames{P} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
+  \and
+  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{y}}{P'} \\ \transition{Q}{\Ain{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Comm-R]{\transition{P}{\Ain{x}{y}}{P'} \\ \transition{Q}{\Aoutf{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Close-L]{\transition{P}{\Aoutb{x}{z}}{P'} \\ \transition{Q}{\Ain{x}{z}}{Q'} \\ z \notin \freenames{Q}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
+  \and
+  \inferrule[Open]{\transition{P}{\Aoutf{x}{z}}{P'} \\ z \neq x}{\transition{\Pres{z}{P}}{\Aoutb{x}{z}}{P'}}
+  \and
+  \inferrule[Close-R]{\transition{P}{\Ain{x}{z}}{P'} \\ \transition{Q}{\Aoutb{x}{z}}{Q'} \\ z \notin \freenames{P}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
+  \and
+  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
+\end{mathpar}
+Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{In}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
+
+As a convention, we assume that bound names of any processes or actions are chosen to be different from the names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names.
+The convention has one exception, namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \), the name \( z \) (which occurs bound in \( P \) and the action \( \Aoutb{x}{z} \)) may occur free in \( Q \).
+Without this exception it would be impossible to express scope extrusion.
+
+\subsection{Bisimulation}
+Our notion of process equivalence relations builds on a notion of \emph{observables}.
+If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
+We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
+
+To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
+\begin{align}
+  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
+  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
+\end{align}
+
+Two processes \( P \) and \( Q \) are then \emph{strongly barbed bisimilar}, written \( \sbbisim{P}{Q} \), if:
+\begin{gather}
+  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
+  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{Q'}~\textrm{for some \( Q' \) with}~\sbbisim{P'}{Q'} \\
+  \observable{Q}{\mu}~\textrm{implies}~\observable{P}{\mu} \\
+  \transition{Q}{\Atau}{Q'}~\textrm{implies}~\transition{P}{\Atau}{P'}~\textrm{for some \( P' \) with}~\sbbisim{P'}{Q'}
+\end{gather}
+Note that, since our calculus does not have infinite behaviour, we do not specify that the relation is the largest of its kind, and the relation is thus not coinductively defined.
+
+\subsection{Challenge}
+The objective of this challenge is to prove the following theorem:
+\begin{theorem}
+  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
+\end{theorem}
 

### 2023-03-01 challenges/challenges.tex
Change heading to read Bisimilarity instead of Bisimulation.
Add description of Coinduction challenge.

@@
 
+\newcommand{\freevars}[1]{\textrm{fv}(#1)}
+\newcommand{\boundvars}[1]{\textrm{bv}(#1)}
+\newcommand{\vars}[1]{\textrm{vars}(#1)}
+
 \newcommand{\alphacon}[2]{#1 =_{\alpha} #2}
@@
 \newcommand{\sbbisim}[2]{#1 \overset{\sbullet}{\sim} #2}
+\newcommand{\nsbbisim}[2]{#1 \not\overset{\sbullet}{\sim} #2}
+
+\newcommand{\ctxhole}{[\cdot]}
+\newcommand{\applyctx}[2]{#1[#2]}
+\newcommand{\sbcong}[2]{#1 \simeq^c #2}
+
+\newcommand{\applysubst}[2]{#2#1}
 
@@
 
-\subsection{Bisimulation}
+\subsection{Bisimilarity}
 Our notion of process equivalence relations builds on a notion of \emph{observables}.
@@
 
+\subsection{Syntax}
+We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
+The syntax is then:
+\begin{align*}
+  v,w :=&&& a \\
+  |&&& l \\
+  P,Q :=&&& \Pend \\
+  |&&& \Pout{x}{v}{P} \\
+  |&&& \Pin{x}{l}{P} \\
+  |&&& \Ppar{P}{Q} \\
+  |&&& \Pres{x}{P} \\
+  |&&& \Preplicate{P}
+\end{align*}
+A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
+The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
+The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
+The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
+The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
+The input operator thus binds the variable name \( l \) in \( P \).
+The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+The two components can proceed independently of each other, or they can interact via shared names.
+The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
+Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
+The restriction operator thus binds the name \( x \) in \( P \).
+Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
+The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
+It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
+Replication makes it possible to express infinite behaviours.
+
+We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
+We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
+We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
+We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
+We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
+
+Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
+As a convention, we will identify \( \alpha \)-convertible processes.
+
+As a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
+This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
+
+\subsection{Semantics}
+The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
+The transitions are labelled by \emph{actions}, the syntax of which are as follows:
+\begin{align*}
+  \alpha := &&& \Aoutf{x}{a} \\
+  |&&& \Ain{x}{a} \\
+  |&&& \Atau
+\end{align*}
+The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value \( a \) via \( x \).
+The \emph{input action} \( \Ain{x}{y} \) is receiving the base value \( y \) via \( x \).
+The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
+
+We will use the notation \( \names{\alpha} \) to denote the set of names that occur in the action \( \alpha \).
+
+The transition relation is then defined by the following rules:
+\begin{mathpar}
+  \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
+  \and
+  \inferrule[In]{ }{\transition{\Pin{x}{l}{P}}{\Ain{x}{a}}{\subst{P}{a}{l}}}
+  \and
+  \inferrule[Par-L]{\transition{P}{\alpha}{P'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
+  \and
+  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
+  \and
+  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{a}}{P'} \\ \transition{Q}{\Ain{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Comm-R]{\transition{P}{\Ain{x}{a}}{P'} \\ \transition{Q}{\Aoutf{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ x \notin \names{\alpha}}{\transition{\Pres{x}{P}}{\alpha}{\Pres{x}{P'}}}
+  \and
+  \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
+\end{mathpar}
+Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
+On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable name.
+
+As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
+
+\subsection{Bisimilarity}
+Our notion of process equivalence relations builds on a notion of \emph{observables}.
+If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
+We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
+
+To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
+\begin{align}
+  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
+  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
+\end{align}
+
+\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is then the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
+\begin{gather}
+  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
+  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}
+\end{gather}
+We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
+Note that, since our systems have potentially infinite behaviours, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
+
+\begin{theorem}
+  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
+\end{theorem}
+
+Unfortunately, strong barbed bisimilarity is not a good notion of process equivalence, since it does not consider the environment of processes.
+For instance, \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since \( \obsout{x} \) is the only observable in both processes and they cannot perform a \( \Atau \)-action, but \( \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since the left process can perform a \( \Atau \)-action such that \( \obsout{y} \) becomes observable, whereas the right process cannot.
+
+\subsection{Contexts and congruences}
+Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
+
+A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+
+We will think of contexts as functions between processes.
+A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+
+We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+
+\subsection{Strong barbed congruence}
+A congruence is exactly what we need to make strong barbed bisimilarity consider the environment (i.e.\ the context) of processes.
+
+We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
+
+\begin{lemma}
+  \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \).
+\end{lemma}
+
+\subsection{Challenge}
+The objective of this challenge is to prove the following theorem:
+\begin{theorem}
+  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).
+\end{theorem}
+
+\paragraph{Notes about this challenge}
+Strong barbed bisimilarity is interesting by itself for this challenge, but the proof that it is an equivalence relation does not require up-to techniques, which are wide spread and interesting.
+The definition of strong barbed congruence requires introducing contexts, but this is actually nice, since they are so prevalent in the field and we want to cover as many basics as possible.
+The theorem about the strong barbed congruence requires up-to techniques, but also seems to involve introducing a structural equivalence, which is possibly not desirable.
+On the other hand, it is a nice application of structural congruence which does not ``ruin'' the semantics.
+
 \section{Challenge: Linearity}

### 2023-03-02 challenges/challenges.tex
Remove unused notations from Coinduction challenge.

@@
 
-We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
-We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
-We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
-We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
 We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-
 Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.

### 2023-03-09 challenges/challenges.tex
Add first draft of linearity challenge.

--- 2023-03-01 challenges/challenges.tex
@@
 \newcommand{\applysubst}[2]{#2#1}
+
+\newcommand{\Presd}[3]{(\bm{\nu} #1#2)~#3}
+\newcommand{\scong}[2]{#1 \equiv #2}
+
+\newcommand{\reduces}[2]{#1 \rightarrow #2}
+
+\newcommand{\Tend}{\mathbf{end}}
+\newcommand{\Tbase}{\mathbf{base}}
+\newcommand{\Tin}[1]{{?}.#1}
+\newcommand{\Tout}[1]{{!}.#1}
+
+\newcommand{\hastype}[2]{#1 : #2}
+
+\newcommand{\un}[1]{\mathbf{un}(#1)}
+\newcommand{\lin}[1]{\mathbf{lin}(#1)}
+
+\newcommand{\Cempty}{\varnothing}
+\newcommand{\Cadd}[2]{#1, #2}
+\newcommand{\Csplit}[2]{#1 \circ #2}
+\newcommand{\Cupdate}[2]{#1 + #2}
+
+\newcommand{\dual}[1]{\overline{#1}}
+
+\newcommand{\types}[2]{#1 \vdash #2}
 
@@
 \section{Challenge: Linearity}
-The setting for this challenge is a restricted \( \pi \)-calculus with a session type system.
+The idea of this challenge is to formalize a proof that requires reasoning about linearity of channels.
+Linearity is the notion that a channel must be used exactly once in a process.
+This is necessary to prove properties about session type systems.
 
-The objective of this challenge is to prove type preservation.
+The setting for this challenge is a small calculus with a session type system, the syntax and semantics of which are presented below.
+The objective of this challenge is to prove type preservation (also known as subject reduction) for the type system, i.e.\ that well-typed processes can only transition to processes which are also well-typed in the same context.
+
+\subsection{Syntax}
+We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
+The syntax is then:
+\begin{align*}
+  v,w :=&&& a \\
+  |&&& l \\
+  P,Q :=&&& \Pend \\
+  |&&& \Pout{x}{v}{P} \\
+  |&&& \Pin{x}{l}{P} \\
+  |&&& \Ppar{P}{Q} \\
+  |&&& \Presd{x}{y}{P}
+\end{align*}
+A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
+The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
+The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
+The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
+The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
+The input operator thus binds the variable name \( l \) in \( P \).
+The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+The two components can proceed independently of each other, or they can interact via shared names.
+The process \( \Presd{x}{y}{P} \) is the \emph{restriction} of the channel with endpoints named \( x \) and \( y \) to \( P \).
+Components in \( P \) can use the names \( x \) and \( y \) to interact with each other (sending on \( x \) and receiving on \( y \) or vice versa), but not with processes outside of the restriction.
+The restriction operator thus binds the names \( x \) and \( y \) in \( P \).
+Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
+Note that there is no recursion of replication in the syntax, and thus no infinite behaviours can be expressed.
+
+We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
+We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
+We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
+We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
+We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
+
+Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
+As a convention, we will identify \( \alpha \)-convertible processes.
+
+As a convention, we assume that bound names and bound variable names of any processes are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, substitutions, and sets of names or variable names.
+This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
+
+\subsection{Semantics}
+We describe the actions that the system can perform using a small step operational semantics.
+To simplify the definition of the reduction relation, we will first factor out a structural congruence relation on processes, and for this we need a few definitions.
+
+\subsubsection{Contexts and congruences}
+A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+
+We will think of contexts as functions between processes.
+A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+
+We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+
+\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
+\begin{mathpar}
+  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
+  \and
+  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
+  \and
+  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
+  \and
+  \inferrule[Sc-Res-Par]{\{x,y\} \not\subseteq \freenames{Q}}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
+  \and
+  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
+  \and
+  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
+\end{mathpar}
+
+The operational semantics is then defined as the following binary relation on processes:
+\begin{mathpar}
+  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{v}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{v}{l}}{R}})}}}
+  \and
+  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
+  \and
+  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
+  \and
+  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
+\end{mathpar}
+Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
+On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable name or a name.
+
+\subsection{Session types}
+Our process syntax allows us to write processes that are not well-formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
+As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of binding names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
+Another example is the following process, which attempts to send a variable that is not resolved at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
+
+To alleviate this issue, we introduce a \emph{session type system}, which will detect processes that are not well-formed.
+To this end, we need to first formally define what we mean by well-formed.
+We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
+A process is then \emph{well-formed} if, for each of its structurally congruenct processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
+
+Note that well-formed processes do not necessarily reduce, since e.g.\ the process \( \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y}{x_2}{\Pin{y_1}{l}{\Pend}}})}} \) is well-formed, but has no reduction.
+
+\subsubsection{Syntax}
+Our type system does not type processes directly, but instead focuses on the channels used in the process.
+The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
+\begin{align*}
+  S,T := &&& \Tend \\
+  |&&& \Tbase \\
+  |&&& \Tin{S} \\
+  |&&& \Tout{S} \\
+  \Gamma := &&& \Cempty \\
+  |&&& \Cadd{\Gamma}{\hastype{x}{S}}
+\end{align*}
+The \emph{end type}, \( \Tend \), is the type of an endpoint on which no further interaction is possible.
+The \emph{base type}, \( \Tbase \), is the type of base values.
+The \emph{input type}, \( \Tin{S} \), is the type of an endpoint which is ready to receive a value, then continue with type \( S \).
+The \emph{output type}, \( \Tout{S} \), is the type of an endpoint which is ready to send a value, then continue with type \( S \).
+
+Typing contexts gather type information about names.
+The \emph{empty type context}, \( \Cempty \), does not contain any information.
+The \emph{type context with \( x \) added}, \( \Cadd{\Gamma}{\hastype{x}{T}} \) adds the information that \( x \) has type \( T \) to the existing type context \( \Gamma \).
+All of the names added to a type context must be distinct.
+We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
+
+Since we need to determine whether endpoints are used dually to determine whether processes are well-formed, we will need to formally define the dual of a type.
+The dual of a type, \( \dual{T} \), is defined by the following equations:
+\begin{mathpar}
+  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
+  \and
+  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
+  \and
+  \inferrule{}{\dual{\Tend} = \Tend}
+\end{mathpar}
+Note that the dual function is not total, since it is not defined for base types.
+
+\subsubsection{Splitting and updating contexts}
+Our type system will maintain two invariants:
+\begin{enumerate}
+\item Endpoints are used exactly once
+\item Endpoints that are part of the same restriction have dual types
+\end{enumerate}
+The first invariant is maintained by linearly splitting type contexts when typing compositions of processes.
+The second invariant is maintained by requiring duality when typing restrictions.
+
+To keep track of linearity, we introduce two \emph{qualifiers}: \( \un{T} \) and \( \lin{T} \).
+A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
+All types are \emph{linear}, \( \lin{T} \).
+The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
+
+The \emph{linear split} of a context, \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \), is defined by the following rules:
+\begin{mathpar}
+  \inferrule[Split-Empty]{ }{\Cempty = \Csplit{\Cempty}{\Cempty}}
+  \and
+  \inferrule[Split-Un]{\Csplit{\Gamma_1}{\Gamma_2} = \Gamma \\ \un{T}}{\Cadd{\Gamma}{\hastype{x}{T}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{T}})}{(\Cadd{\Gamma_2}{\hastype{x}{T}})}}
+  \and
+  \inferrule[Split-Lin-L]{\Gamma = \Csplit{\Gamma_1}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{S}})}{\Gamma_2}}
+  \and
+  \inferrule[Split-Lin-R]{\Gamma = \Csplit{\Gamma_2}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{\Gamma_1}{(\Cadd{\Gamma_2}{\hastype{x}{S}})}}
+\end{mathpar}
+
+We will also need to update contexts with new type information in a safe way while typing input and output.
+The \emph{context update}, \( \Cupdate{\Gamma}{\hastype{x}{T}} = \Gamma' \), is defined by the following rules:
+\begin{mathpar}
+  \inferrule[Update-Name]{\hastype{x}{U} \notin \Gamma}{\Cupdate{\Gamma}{\hastype{x}{T}} = \Cadd{\Gamma}{\hastype{x}{T}}}
+  \and
+  \inferrule[Update-Un]{\un{T}}{\Cupdate{(\Cadd{\Gamma}{\hastype{x}{T}})}{\hastype{x}{T}} = (\Cadd{\Gamma}{\hastype{x}{T}})}
+\end{mathpar}
+
+\subsubsection{Typing rules}
+We have three typing judgements: one for values, one for names, and one for processes.
+The typing rules for values are as follows:
+\begin{mathpar}
+  \inferrule[T-Base]{\un{\Gamma}}{\types{\Gamma}{\hastype{a}{\Tbase}}}
+  \and
+  \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{v}{\Tbase}}}{\hastype{v}{\Tbase}}}
+\end{mathpar}
+The only typing rule for names is:
+\begin{mathpar}
+  \inferrule[T-Name]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{x}{T}}}
+\end{mathpar}
+The typing rules for processes are as follows:
+\begin{mathpar}
+  \inferrule[T-Inact]{\un{\Gamma}}{\types{\Gamma}{\Pend}}
+  \and
+  \inferrule[T-Par]{\types{\Gamma_1}{P} \\ \types{\Gamma_2}{Q}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Ppar{P}{Q}}}
+  \and
+  \inferrule[T-Res]{\types{\Cadd{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P}}{\types{\Gamma}{\Presd{x}{y}{P}}}
+  \and
+  \inferrule[T-Out]{\types{\Gamma_1}{\hastype{x}{\Tout{T}}} \\ \types{\Gamma_2}{\hastype{v}{\Tbase}} \\ \types{\Cupdate{\Gamma_3}{\hastype{x}{T}}}{P}}{\types{\Csplit{\Gamma_1}{\Csplit{\Gamma_2}{\Gamma_3}}}{\Pout{x}{v}{P}}}
+  \and
+  \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{(\Cadd{\Gamma_2}{\hastype{l}{\Tbase}})}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
+\end{mathpar}
+
+\subsection{Challenge}
+
+\begin{lemma}
+  If \( \types{\Gamma}{P} \) and \( \un{T} \) then \( \types{\Cadd{\Gamma}{\hastype{x}{T}}}{P} \).
+\end{lemma}
+
+\begin{lemma}
+  Assume \( \types{\Gamma}{P} \) and \( x \notin \freenames{P} \).
+  Then:
+  \begin{enumerate}
+  \item \( \hastype{x}{\Tout{T}} \notin \Gamma \).
+  \item \( \hastype{x}{\Tin{T}} \notin \Gamma \).
+  \item if \( \Gamma = \Cadd{\Gamma'}{\hastype{x}{T}} \) then \( \types{\Gamma'}{P} \).
+  \end{enumerate}
+\end{lemma}
+
+\begin{lemma}
+  If \( \types{\Gamma}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma}{Q} \).
+\end{lemma}
+
+\begin{lemma}
+  If \( \types{\Gamma_1}{\hastype{v}{T}} \) and \( \types{\Cadd{\Gamma_2}{\hastype{x}{T}}}{P} \) and \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then \( \types{\Gamma}{\subst{v}{x}{P}} \).
+\end{lemma}
+
+\begin{theorem}
+  If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
+\end{theorem}
+
+\begin{theorem}
+  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
+\end{theorem}
+
+\begin{theorem}
+  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
+\end{theorem}
 

### 2023-04-26 challenges.tex
rough merge of the challenges to the main file
merged the structures and changed the titles

--- 2023-03-09 challenges/challenges.tex
@@
-\documentclass[a4paper]{article}
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage[utf8]{inputenc}
-\usepackage{mathpartir}
-\usepackage{bm}
-\usepackage{graphicx}
-
-\newtheorem{lemma}{Lemma}
-\newtheorem{theorem}{Theorem}
-
-\newcommand{\Pend}{\bm{0}}
-\newcommand{\Ppar}[2]{#1 \mid #2}
-\newcommand{\Pres}[2]{(\bm{\nu} #1)~#2}
-\newcommand{\Pout}[3]{#1 ! #2 . #3}
-\newcommand{\Pin}[3]{#1 ? (#2) . #3}
-\newcommand{\Pchoice}[2]{#1 + #2}
-\newcommand{\Preplicate}[1]{{!}#1}
-
-\newcommand{\freenames}[1]{\textrm{fn}(#1)}
-\newcommand{\boundnames}[1]{\textrm{bn}(#1)}
-\newcommand{\names}[1]{\textrm{n}(#1)}
-
-\newcommand{\freevars}[1]{\textrm{fv}(#1)}
-\newcommand{\boundvars}[1]{\textrm{bv}(#1)}
-\newcommand{\vars}[1]{\textrm{vars}(#1)}
-
-\newcommand{\alphacon}[2]{#1 =_{\alpha} #2}
-\newcommand{\subst}[3]{#1\{#2/#3\}}
-
-\newcommand{\Aoutf}[2]{#1 ! #2}
-\newcommand{\Aoutb}[2]{#1 ! (#2)}
-\newcommand{\Ain}[2]{#1 ? #2}
-\newcommand{\Atau}{\tau}
-
-\newcommand{\transition}[3]{#1 \xrightarrow{#2} #3}
-
-\newcommand{\observable}[2]{#1\downarrow_{#2}}
-\newcommand{\obsin}[1]{#1?}
-\newcommand{\obsout}[1]{#1!}
-
-\newcommand\sbullet[1][.5]{\mathbin{\hbox{\scalebox{#1}{$\bullet$}}}}
-\newcommand{\sbbisim}[2]{#1 \overset{\sbullet}{\sim} #2}
-\newcommand{\nsbbisim}[2]{#1 \not\overset{\sbullet}{\sim} #2}
-
-\newcommand{\ctxhole}{[\cdot]}
-\newcommand{\applyctx}[2]{#1[#2]}
-\newcommand{\sbcong}[2]{#1 \simeq^c #2}
-
-\newcommand{\applysubst}[2]{#2#1}
-
-\newcommand{\Presd}[3]{(\bm{\nu} #1#2)~#3}
-\newcommand{\scong}[2]{#1 \equiv #2}
-
-\newcommand{\reduces}[2]{#1 \rightarrow #2}
-
-\newcommand{\Tend}{\mathbf{end}}
-\newcommand{\Tbase}{\mathbf{base}}
-\newcommand{\Tin}[1]{{?}.#1}
-\newcommand{\Tout}[1]{{!}.#1}
-
-\newcommand{\hastype}[2]{#1 : #2}
-
-\newcommand{\un}[1]{\mathbf{un}(#1)}
-\newcommand{\lin}[1]{\mathbf{lin}(#1)}
-
-\newcommand{\Cempty}{\varnothing}
-\newcommand{\Cadd}[2]{#1, #2}
-\newcommand{\Csplit}[2]{#1 \circ #2}
-\newcommand{\Cupdate}[2]{#1 + #2}
-
-\newcommand{\dual}[1]{\overline{#1}}
-
-\newcommand{\types}[2]{#1 \vdash #2}
-
-\begin{document}
-
-\section{Challenge: Scope Extrusion}
+\subsection{Challenge: Name passing and scope extrusion}
 The idea of this challenge is to formalize a proof that requires scope extrusion.
@@
 
-\subsection{Syntax}
+\subsubsection{Syntax}
 We assume the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
@@
 
-\subsection{Semantics}
+\subsubsection{Semantics}
 The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
@@
 
-\subsection{Bisimilarity}
+\subsubsection{Bisimilarity}
 Our notion of process equivalence relations builds on a notion of \emph{observables}.
@@
 
-\subsection{Challenge}
+\subsubsection{Challenge}
 The objective of this challenge is to prove the following theorem:
@@
 
-\section{Challenge: Coinduction}
+\subsection{Challenge: Coinduction and reasoning about infinite processes}
 The idea of this challenge is to formalize a proof that requires coinductive techniques.
@@
 
-\subsection{Syntax}
+\subsubsection{Syntax}
 We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
@@
 
-\subsection{Contexts and congruences}
+\subsubsection{Contexts and congruences}
 Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
@@
 
-\subsection{Strong barbed congruence}
+\subsubsection{Strong barbed congruence}
 A congruence is exactly what we need to make strong barbed bisimilarity consider the environment (i.e.\ the context) of processes.
@@
 
-\section{Challenge: Linearity}
+\subsection{Challenge: Linearity and behavioural type systems}
 The idea of this challenge is to formalize a proof that requires reasoning about linearity of channels.
@@
 
-\subsection{Semantics}
+\subsubsection{Semantics}
 We describe the actions that the system can perform using a small step operational semantics.
@@
 
-\subsection{Session types}
+\subsubsection{Session types}
 Our process syntax allows us to write processes that are not well-formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
@@
 
-\subsubsection{Syntax}
+\paragraph{Syntax}
 Our type system does not type processes directly, but instead focuses on the channels used in the process.
@@
 
-\subsubsection{Splitting and updating contexts}
+\paragraph{Splitting and updating contexts}
 Our type system will maintain two invariants:
@@
 
-\subsection{Challenge}
+\subsubsection{Challenge}
 
@@
 
-\section{Future work}
-\begin{itemize}
-\item Multiparty session types
-\item Choreography
-\item Encodings
-\item Code extraction
-\end{itemize}
-
-\end{document}
+%% Local Variables:
+%% mode: latex
+%% TeX-master: "main"
+%% End:

### 2023-04-26 main.tex
rough merge of the challenges to the main file
merged the structures and changed the titles

@@
 
+\usepackage[utf8]{inputenc}
 \usepackage{graphicx}
 \usepackage{natbib}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{mathpartir}
+\usepackage{bm}
 \usepackage{hyperref}
@@
 
-% macros (move to its own file when we have a macro).
+\input{macros.tex}
 
@@
 
-\subsection{Name passing and scope extrusion}
-
-\textbf{Problem idea}: A small proof about the untyped $\pi$-calculus.
-
-\subsection{Linearity and behavioural type systems}
-
-\textbf{Problem idea}: A small session type problem that exposes linear channels.
-
-\subsection{Coinduction and reasoning about process algebras}
-
-\textbf{Problem idea}: Something about traces, (bi)similarity, or
-subtyping under non terminating problems in behavioural types.
+\input{challenges.tex}
 
@@
 
+\section{Future work}
+\begin{itemize}
+\item Multiparty session types
+\item Choreography
+\item Encodings
+\item Code extraction
+\item Conversation types
+\item Psi-calculus
+\item Model checking for system properties
+\item More general types than session types
+\end{itemize}
+
 \section{Conclusion and related work} \label{sec:conclusion}

### 2023-04-26 main.tex
some rework

@@
 
-This section is not (at least initially) meant to be in the final
-manuscript but it is to document some reasoning for choosing
-appropriate problems. The idea is to explore three areas: scope
+ The idea is to explore three areas: scope
 extrusion in name passing calculi, linearity requirements in
@@
 
-But the exact presentation of the concrete problems should be somewhat
+But the exact presentation of the concrete problems is somewhat
 flexible. In the sense that the use of certain techniques or tools may
 require changing the presentation of the system to fit the technique.
-These changes should be allowed, and even encouraged, but the result
-should be ``morally'' the same system and there should be a
-rationale for the changes in presentation. That way we could not only
-see valid techniques but also what presentation of the rules to use
-and the rationale to chose a particular approach.
+We are more interested in the techniques that are applicable to these
+problems as a class of problems than to the exact specification we
+propose. Of course, for clarity if a solution changes the definition
+it should document in what sense the definition differs. So changes
+allowed, and even encouraged, but the result should be ``morally'' the
+same system and there should be a rationale for the changes in
+presentation. That way we not only see valid techniques but also what
+presentation of the rules to use and the rationale to chose a
+particular approach.
 

### 2023-04-26 challenges.tex
some more rework

@@
 \subsection{Challenge: Name passing and scope extrusion}
-The idea of this challenge is to formalize a proof that requires scope extrusion.
+This challenge formalises a proof that requires explicit scope extrusion.
 Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the recieving process.
@@
 \subsubsection{Syntax}
-We assume the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
+We assume the existence of a type of \emph{names} that can be compared for equality, values of which we will denote by \( x, y, \dots \).
 The syntax is then:
@@
 The restriction operator thus binds the name \( x \) in \( P \).
-Note that the scope of a restriction may change when processes interact.
+Note that the scope of a restriction may change when processes interact. Namely, a restricted name may be sent \emph{outside} of its scope.
 The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
@@
 The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which are as follows:
+The transitions are labelled by \emph{actions}, the syntax of which is as follows:
 \begin{align*}
@@
 \subsection{Challenge: Coinduction and reasoning about infinite processes}
-The idea of this challenge is to formalize a proof that requires coinductive techniques.
+This challenge formalises a proof that requires coinductive techniques.
 Coinduction is a proof technique for infinite structures, which arise in this context due to systems with behaviours that continue indefinitely.
@@
 The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
+Which, to reiterate, entails the use of coinduction due to the presence of infinite behaviour.
 
@@
 
-As a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
+Also as a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
 This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
@@
 
-We will think of contexts as functions between processes.
+We can think of contexts as functions between processes.
 A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
@@
 \subsection{Challenge: Linearity and behavioural type systems}
-The idea of this challenge is to formalize a proof that requires reasoning about linearity of channels.
+This challenge formalises a proof that requires reasoning about linearity of channels.
 Linearity is the notion that a channel must be used exactly once in a process.

### 2023-06-02 working-drafts/angle1.tex
Add early working draft outlines.

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathpartir}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{draftwatermark}

\SetWatermarkText{Draft}
\SetWatermarkScale{5}
\SetWatermarkColor[gray]{0.90}

\input{../macros.tex}

\begin{document}

\title{Working Draft: Formalization Gap Angle}

\maketitle

\begin{abstract}
  \noindent
  \begin{itemize}
  \item POPLMark made formalizations more common
  \item POPLMark does not cover the specific issues present in concurrency
  \item We propose three challenge problems concerning scope extrusion, linearity, and coinductive reasoning
  \end{itemize}
\end{abstract}

\section{Introduction}
\begin{itemize}
\item The PL community has embraced mechanization in the last decade
  \begin{itemize}
  \item POPLMark helped get the community on track for more mechanizations
  \item Since POPLMark, the number of formal proofs accompanying papers at POPL and PLDI has grown
  \end{itemize}
\item The concurrency community has not embraced mechanization
  \begin{itemize}
  \item The number of formal proofs accompanying papers at CONCUR is low and has not followed the trend at POPL/PLDI
  \end{itemize}
\item We propose a new benchmark for the specific issues present in concurrency
  \begin{itemize}
  \item A goal is to help develop momentum for more mechanizations
  \item A goal is to help develop a culture of mechanization
  \item A goal is to rally the community for collaborating on mechanizations
  \end{itemize}
\item Benchmarks cannot address every technique and field at once
  \begin{itemize}
  \item Challenges have to be simple enough to be manageable
  \item Challenges should not be encumbered by too many technical details
  \end{itemize}
\item POPLMark Reloaded extends the scope to proofs using logical relations
  \begin{itemize}
  \item Logical relations are crucial in many current developments
  \end{itemize}
\item Our benchmark extends the scope to process calculi and behavioural types
\item There are three key aspects that cause issues when mechanizing process calculi and behavioural types
  \begin{itemize}
  \item Scope extrusion in name passing systems generalises binders that makes techniques from lambda calculi inconvenient
  \item Linearity in behavioural type systems causes problems since linearity is generally not well supported
  \item Infinite behaviour modelled with coinduction causes problems since coinduction is generally not well supported
  \end{itemize}
\item Each part of our benchmark exercises the three aspects independently
  \begin{itemize}
  \item Real research projects will often need more than just one aspect
  \item We exercise each in isolation to keep solutions small
  \item We do not address how to combine techniques for the three aspects
  \end{itemize}
\item The problems we propose have been mechanized before
  \begin{itemize}
  \item A goal is to enable comparison between solutions using different techniques
  \item Our challenges are simple settings where we can compare techniques without distractions
  \end{itemize}
\end{itemize}

\section{Rationale for choosing the problems}
\begin{itemize}
\item Our challenges should cover the three aspects independently
  \begin{itemize}
  \item People can choose to implement only a single challenge
  \item People can choose to work on the challenges in any order
  \end{itemize}
\item Our challenges should be easy to understand and implement
  \begin{itemize}
  \item This is to obtain as many solutions as possible
  \item Our challenges should be small and quick to read
  \end{itemize}
\item Our challenges do not need to be met exactly
  \begin{itemize}
  \item People can choose to modify the challenge slightly to show the best aspects of their solution
  \item The spirit of each challenge is what matters
  \item Our challenges should be as technique-agnostic as possible
  \end{itemize}
\end{itemize}

\section{The challenge problems}
\begin{itemize}
\item Combining the techniques in the challenges is more of a research problem than a benchmark
\item Combining two techniques at a time can be interesting bonus challenges
\item The scope extrusion challenge is to prove that barbed bisimulation is an equivalence relation for an untyped \(\pi\)-calculus without infinite behaviour, but with non-deterministic choice
\item The coinduction challenge is to prove that strong barbed bisimulation is an equivalence relation for an untyped calculus of communicating systems with (infinite) replication of processes
\item The linearity challenge is to prove type preservation (aka.\ subject reduction) for a process calculus with a session type system
\end{itemize}

\section{Bonus challenges}
\begin{itemize}
\item Combining linearity and scope extrusion can be done by proving type preservation for the \(\pi\)-calculus with a session type system
\item Combining scope extrusion and coinduction can be done by proving that bisimulation is an equivalence relation for a simply typed \(\pi\)-calculus
\item Combining coinduction and linearity can be done by proving type preservation for CCS with a session type system
\end{itemize}

\section{The solutions}
\begin{itemize}
\item Should include a skeleton/sketch
\item Full details in an appendix
\end{itemize}

\section{Future work}
\begin{itemize}
\item Multiparty session types
\item Choreography
\item Encodings
\item Code extraction
\item Conversation types
\item Psi-calculus
\item Model checking for system properties
\item More general types than session types
\end{itemize}

\section{Conclusion and related work}

\end{document}

### 2023-06-02 working-drafts/angle2.tex
Add early working draft outlines.

--- 2023-06-02 working-drafts/angle1.tex
@@
 
-\title{Working Draft: Formalization Gap Angle}
+\title{Working Draft: Lack of Unified Approaches Angle}
 
@@
   \item POPLMark helped get the community on track for more mechanizations
-  \item Since POPLMark, the number of formal proofs accompanying papers at POPL and PLDI has grown
+  \item Since POPLMark, the number of formal proofs accompanying papers has grown
   \end{itemize}
-\item The concurrency community has not embraced mechanization
-  \begin{itemize}
-  \item The number of formal proofs accompanying papers at CONCUR is low and has not followed the trend at POPL/PLDI
-  \end{itemize}
+\item The concurrency community has not embraced mechanization to the same extent
 \item We propose a new benchmark for the specific issues present in concurrency

### 2023-06-16 extended-abstract/extended-abstract.tex
Add draft of extended abstract.

\documentclass[adraft,copyright,creativecommons]{eptcs}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathpartir}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{draftwatermark}
\usepackage{underscore}
\usepackage[T1]{fontenc}

\SetWatermarkText{Draft}
\SetWatermarkScale{2}
\SetWatermarkColor[gray]{0.90}

\input{../macros.tex}

\title{Concurrency Benchmark Challenge (Extended Abstract)}
\def\titlerunning{Concurrency Benchmark Challenge (Extended Abstract)}

\begin{document}
\maketitle

\begin{abstract}
  We propose a new collection of benchmark problems with a focus on the issues encountered when mechanizing concurrency theory.
  POPLMark sparked a flurry of work to move us closer to a world in which every paper on programming languages is accompanied by machine-checked proofs, but both POPLMark and POPLMark Reloaded were explicitly limited in scope.
  Our benchmark problems concern the mechanization of scope extrusion, linear handling of environments, and coinductive reasoning.
  Like previous benchmarks, our goal is to clarify, compare and advance the state of the art.
\end{abstract}

\section{Introduction}
The programming language community has increasingly embraced mechanizations since the influential POPLMark challenge~\cite{POPLMark}, but we are still far from the vision of a mechanized proof accompanying every paper in the field.
Since POPLMark, the number of papers with accompanying mechanized proofs has grown steadily, but many mechanizations still rely on ad hoc techniques.
The POPLMark Reloaded challenge~\cite{POPLMarkReloaded} aimed to identify and develop best practices and tool support for proofs using logical relations.
The authors of both POPLMark and POPLMark Reloaded were careful to note that their benchmarks were only a beginning, and the previous benchmarks specifically mention reasoning about concurrency using coinduction and linear environments as points for future work.

Papers in concurrency theory do often include mechanized proofs, but our experience is that choosing techniques and tools remains a challenge with few guidelines.
We propose a new collection of benchmark problems specifically designed to tackle the issues encountered when mechanizing concurrency theory.
A goal of our collection of benchmark problems is to enable comparison between techniques and push for the development of guidelines and tutorials, and our challenge problems are a simple setting in which to demonstrate and compare techniques.
Another goal is to strengthen the culture of mechanization by rallying the community to collaborate on exploring and developing new tools and techniques.

We have identified three key aspects which cause issues when mechanizing concurrency theory:
\begin{itemize}
\item Scope extrusion in name passing systems
\item Linearity in behavioural type systems
\item Infinite behaviour modelled coinductively
\end{itemize}
Each part of our benchmark exercises the three aspects independently to enable comparison and to keep the challenge problems small.
The three aspects are of course not the only ones that may cause issues in mechanizations, but they are fundamental to the field and will thus appear in most mechanizations of concurrency theory.
Actual mechanizations will additionally often require more than just one aspect, but we will not address the combination of techniques except as optional bonus challenges.

In the remainder of this paper we will discuss the design of the challenge and briefly describe the three challenge problems.
The full text of the challenge problems, including pen and paper solutions, is in the appendix.

\section{Design of the challenge}
Our challenge problems cannot address every technique and subfield at once.
We have identified three key aspects fundamental to concurrency theory mechanizations and designed our challenge problems to exercise these.
Like previous benchmark authors, we would like to stress that our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types and psi-calculi are not directly covered by our challenge problems.
Most applications will however still need the basic techniques that our challenge problems exercise.

We have designed our challenge problems to exercise the three aspects independently, and they can be solved individually and in any order.
We have designed each problem to be small and easy to understand with a basic knowledge of textbook concurrency theory.
Solutions to our challenge problems should be compared on three measures:
\begin{itemize}
\item The mechanization overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanization.
\item The adequacy of the formal statements in the mechanization, i.e.\ whether the proven theorems are easily recognizable as the theorems from the challenge.
\item The cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
\end{itemize}
Solutions to our challenges do not need to strictly follow the definitions and techniques set out in the problem text, but solutions which deviate from the original challenges will need to present more elaborate argumentation for their adequacy.

\section{The challenge problems}
In this section we will briefly describe our three challenge problems, referring to the appendix for the full text and pen and paper solution to each problem.

In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
Our scope extrusion challenge is to prove that barbed bisimulation is an equivalence relation for an untyped name passing caluclus without infinite behaviour, but with non-deterministic choice.
The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues with some approaches to binders.

Our coinduction challenge is to prove that strong barbed bisimulation is an equivalence relation for an untyped calculus of communicating systems with (infinite) replication of processes.
The proof involves reasoning coinductively about the infinite behaviours of the replication operator.

Our linearity challenge is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
Session type systems require linear handling of typing contexts, and the proof involves reasoning about the linearity of context splitting operations.

We invite anyone interested to send us solutions to the challenge problems and to publish experience reports detailing the advantages and disadvantages of their approaches.
The interested reader can find more information about the current status of the challenge at
\begin{center}
  \url{https://concurrentbenchmark.github.io/}
\end{center}

\bibliographystyle{eptcs}
\bibliography{bibliography}

\end{document}

### 2023-06-22 challenges.tex
Remove bonus challenges from challenge description.

@@
 
-\section{Bonus challenges}
-
-\subsection{Linearity and Scope Extrusion}
-The setting for this challenge is the \( \pi \)-calculus with a session type system.
-
-The objective of this challenge is to prove type preservation.
-
-\subsection{Scope Extrusion and Coinduction}
-The setting for this challenge is a simply typed \( \pi \)-calculus.
-
-The objective of this challenge is to prove bisimulation.
-
-\subsection{Coinduction and Linearity}
-The setting for this challenge is the Calculus of Communicating Systems with a session type system.
-
-The objective of this challenge is to prove type preservation.
-
 %% Local Variables:

### 2023-06-23 challenges.tex
Adjust title, abstract, parts of the main text

@@
 \subsection{Challenge: Name passing and scope extrusion}
+\label{sec:challenge:name-passing-scope-extrusion}
+
 This challenge formalises a proof that requires explicit scope extrusion.
@@
 \subsection{Challenge: Coinduction and reasoning about infinite processes}
+\label{sec:challenge:coinduction}
+
 This challenge formalises a proof that requires coinductive techniques.
@@
 \subsection{Challenge: Linearity and behavioural type systems}
+\label{sec:challenge:linearity-beh-types}
+
 This challenge formalises a proof that requires reasoning about linearity of channels.

### 2023-06-23 extended-abstract/extended-abstract.tex
Add challenge description as appendix to EA.
Adjust title, abstract, parts of the main text

@@
 
-\title{Concurrency Benchmark Challenge (Extended Abstract)}
-\def\titlerunning{Concurrency Benchmark Challenge (Extended Abstract)}
+\title{The Concurrent Calculi Formalisation Benchmark\\{\normalsize(Extended Abstract)}}
+\def\titlerunning{The Concurrent Calculi Formalisation Benchmark (Extended Abstract)}
 
@@
 \begin{abstract}
-  We propose a new collection of benchmark problems with a focus on the issues encountered when mechanizing concurrency theory.
-  POPLMark sparked a flurry of work to move us closer to a world in which every paper on programming languages is accompanied by machine-checked proofs, but both POPLMark and POPLMark Reloaded were explicitly limited in scope.
+  The POPLMark challenge has sparked a flurry of work to move us closer to a world in which every paper on programming languages is accompanied by machine-checked proofs.
+  However, POPLMark (and its follow-up POPLMark Reloaded) were explicitly limited in scope,
+  with benchmark problems that do not address concurrency.
+For this reason, we propose a new collection of benchmark problems with a focus on the issues encountered when mechanizing message-passing concurrency using process calculi.
   Our benchmark problems concern the mechanization of scope extrusion, linear handling of environments, and coinductive reasoning.
@@
 Papers in concurrency theory do often include mechanized proofs, but our experience is that choosing techniques and tools remains a challenge with few guidelines.
-We propose a new collection of benchmark problems specifically designed to tackle the issues encountered when mechanizing concurrency theory.
+We propose a new collection of benchmark problems specifically designed to tackle the issues encountered when mechanizing message-passing concurrency using process calculi.
 A goal of our collection of benchmark problems is to enable comparison between techniques and push for the development of guidelines and tutorials, and our challenge problems are a simple setting in which to demonstrate and compare techniques.
@@
 
-In the remainder of this paper we will discuss the design of the challenge and briefly describe the three challenge problems.
-The full text of the challenge problems, including pen and paper solutions, is in the appendix.
+In the rest of this paper we discuss the design of the challenge and briefly describe the three challenge problems.
+The appendix contains the full text of the challenge problems and pen and paper solutions.
 
@@
 We have designed our challenge problems to exercise the three aspects independently, and they can be solved individually and in any order.
-We have designed each problem to be small and easy to understand with a basic knowledge of textbook concurrency theory.
+We have designed each problem to be small and easy to understand with a basic knowledge of textbook concurrency theory and process calculi.
 Solutions to our challenge problems should be compared on three measures:
@@
 In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
-Our scope extrusion challenge is to prove that barbed bisimulation is an equivalence relation for an untyped name passing caluclus without infinite behaviour, but with non-deterministic choice.
+Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing caluclus without infinite behaviour, but with non-deterministic choice.
 The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues with some approaches to binders.
 
-Our coinduction challenge is to prove that strong barbed bisimulation is an equivalence relation for an untyped calculus of communicating systems with (infinite) replication of processes.
+Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed bisimulation is an equivalence relation for an untyped calculus of communicating systems with (infinite) replication of processes.
 The proof involves reasoning coinductively about the infinite behaviours of the replication operator.
 
-Our linearity challenge is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
+Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
 Session type systems require linear handling of typing contexts, and the proof involves reasoning about the linearity of context splitting operations.
@@
 We invite anyone interested to send us solutions to the challenge problems and to publish experience reports detailing the advantages and disadvantages of their approaches.
-The interested reader can find more information about the current status of the challenge at
+The interested reader can find more information about the current status of the challenge at:
 \begin{center}
@@
 
+\appendix
+
+\section{Challenges}
+\input{../challenges.tex}
+
 \end{document}

### 2023-07-10 challenges.tex
Change the order of challenges.
Comment out notes in challenges.

@@
+\subsection{Challenge: Linearity and behavioural type systems}
+\label{sec:challenge:linearity-beh-types}
+
+This challenge formalises a proof that requires reasoning about linearity of channels.
+Linearity is the notion that a channel must be used exactly once in a process.
+This is necessary to prove properties about session type systems.
+
+The setting for this challenge is a small calculus with a session type system, the syntax and semantics of which are presented below.
+The objective of this challenge is to prove type preservation (also known as subject reduction) for the type system, i.e.\ that well-typed processes can only transition to processes which are also well-typed in the same context.
+
+\subsubsection{Syntax}
+We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
+The syntax is then:
+\begin{align*}
+  v,w :=&&& a \\
+  |&&& l \\
+  P,Q :=&&& \Pend \\
+  |&&& \Pout{x}{v}{P} \\
+  |&&& \Pin{x}{l}{P} \\
+  |&&& \Ppar{P}{Q} \\
+  |&&& \Presd{x}{y}{P}
+\end{align*}
+A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
+The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
+The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
+The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
+The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
+The input operator thus binds the variable name \( l \) in \( P \).
+The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+The two components can proceed independently of each other, or they can interact via shared names.
+The process \( \Presd{x}{y}{P} \) is the \emph{restriction} of the channel with endpoints named \( x \) and \( y \) to \( P \).
+Components in \( P \) can use the names \( x \) and \( y \) to interact with each other (sending on \( x \) and receiving on \( y \) or vice versa), but not with processes outside of the restriction.
+The restriction operator thus binds the names \( x \) and \( y \) in \( P \).
+Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
+Note that there is no recursion of replication in the syntax, and thus no infinite behaviours can be expressed.
+
+We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
+We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
+We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
+We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
+We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
+
+Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
+As a convention, we will identify \( \alpha \)-convertible processes.
+
+As a convention, we assume that bound names and bound variable names of any processes are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, substitutions, and sets of names or variable names.
+This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
+
+\subsubsection{Semantics}
+We describe the actions that the system can perform using a small step operational semantics.
+To simplify the definition of the reduction relation, we will first factor out a structural congruence relation on processes, and for this we need a few definitions.
+
+\subsubsection{Contexts and congruences}
+A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+
+We will think of contexts as functions between processes.
+A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+
+We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+
+\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
+\begin{mathpar}
+  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
+  \and
+  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
+  \and
+  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
+  \and
+  \inferrule[Sc-Res-Par]{\{x,y\} \not\subseteq \freenames{Q}}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
+  \and
+  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
+  \and
+  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
+\end{mathpar}
+
+The operational semantics is then defined as the following binary relation on processes:
+\begin{mathpar}
+  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{v}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{v}{l}}{R}})}}}
+  \and
+  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
+  \and
+  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
+  \and
+  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
+\end{mathpar}
+Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
+On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable name or a name.
+
+\subsubsection{Session types}
+Our process syntax allows us to write processes that are not well-formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
+As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of binding names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
+Another example is the following process, which attempts to send a variable that is not resolved at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
+
+To alleviate this issue, we introduce a \emph{session type system}, which will detect processes that are not well-formed.
+To this end, we need to first formally define what we mean by well-formed.
+We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
+A process is then \emph{well-formed} if, for each of its structurally congruenct processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
+
+Note that well-formed processes do not necessarily reduce, since e.g.\ the process \( \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y}{x_2}{\Pin{y_1}{l}{\Pend}}})}} \) is well-formed, but has no reduction.
+
+\paragraph{Syntax}
+Our type system does not type processes directly, but instead focuses on the channels used in the process.
+The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
+\begin{align*}
+  S,T := &&& \Tend \\
+  |&&& \Tbase \\
+  |&&& \Tin{S} \\
+  |&&& \Tout{S} \\
+  \Gamma := &&& \Cempty \\
+  |&&& \Cadd{\Gamma}{\hastype{x}{S}}
+\end{align*}
+The \emph{end type}, \( \Tend \), is the type of an endpoint on which no further interaction is possible.
+The \emph{base type}, \( \Tbase \), is the type of base values.
+The \emph{input type}, \( \Tin{S} \), is the type of an endpoint which is ready to receive a value, then continue with type \( S \).
+The \emph{output type}, \( \Tout{S} \), is the type of an endpoint which is ready to send a value, then continue with type \( S \).
+
+Typing contexts gather type information about names.
+The \emph{empty type context}, \( \Cempty \), does not contain any information.
+The \emph{type context with \( x \) added}, \( \Cadd{\Gamma}{\hastype{x}{T}} \) adds the information that \( x \) has type \( T \) to the existing type context \( \Gamma \).
+All of the names added to a type context must be distinct.
+We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
+
+Since we need to determine whether endpoints are used dually to determine whether processes are well-formed, we will need to formally define the dual of a type.
+The dual of a type, \( \dual{T} \), is defined by the following equations:
+\begin{mathpar}
+  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
+  \and
+  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
+  \and
+  \inferrule{}{\dual{\Tend} = \Tend}
+\end{mathpar}
+Note that the dual function is not total, since it is not defined for base types.
+
+\paragraph{Splitting and updating contexts}
+Our type system will maintain two invariants:
+\begin{enumerate}
+\item Endpoints are used exactly once
+\item Endpoints that are part of the same restriction have dual types
+\end{enumerate}
+The first invariant is maintained by linearly splitting type contexts when typing compositions of processes.
+The second invariant is maintained by requiring duality when typing restrictions.
+
+To keep track of linearity, we introduce two \emph{qualifiers}: \( \un{T} \) and \( \lin{T} \).
+A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
+All types are \emph{linear}, \( \lin{T} \).
+The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
+
+The \emph{linear split} of a context, \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \), is defined by the following rules:
+\begin{mathpar}
+  \inferrule[Split-Empty]{ }{\Cempty = \Csplit{\Cempty}{\Cempty}}
+  \and
+  \inferrule[Split-Un]{\Csplit{\Gamma_1}{\Gamma_2} = \Gamma \\ \un{T}}{\Cadd{\Gamma}{\hastype{x}{T}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{T}})}{(\Cadd{\Gamma_2}{\hastype{x}{T}})}}
+  \and
+  \inferrule[Split-Lin-L]{\Gamma = \Csplit{\Gamma_1}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{S}})}{\Gamma_2}}
+  \and
+  \inferrule[Split-Lin-R]{\Gamma = \Csplit{\Gamma_2}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{\Gamma_1}{(\Cadd{\Gamma_2}{\hastype{x}{S}})}}
+\end{mathpar}
+
+We will also need to update contexts with new type information in a safe way while typing input and output.
+The \emph{context update}, \( \Cupdate{\Gamma}{\hastype{x}{T}} = \Gamma' \), is defined by the following rules:
+\begin{mathpar}
+  \inferrule[Update-Name]{\hastype{x}{U} \notin \Gamma}{\Cupdate{\Gamma}{\hastype{x}{T}} = \Cadd{\Gamma}{\hastype{x}{T}}}
+  \and
+  \inferrule[Update-Un]{\un{T}}{\Cupdate{(\Cadd{\Gamma}{\hastype{x}{T}})}{\hastype{x}{T}} = (\Cadd{\Gamma}{\hastype{x}{T}})}
+\end{mathpar}
+
+\subsubsection{Typing rules}
+We have three typing judgements: one for values, one for names, and one for processes.
+The typing rules for values are as follows:
+\begin{mathpar}
+  \inferrule[T-Base]{\un{\Gamma}}{\types{\Gamma}{\hastype{a}{\Tbase}}}
+  \and
+  \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{v}{\Tbase}}}{\hastype{v}{\Tbase}}}
+\end{mathpar}
+The only typing rule for names is:
+\begin{mathpar}
+  \inferrule[T-Name]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{x}{T}}}
+\end{mathpar}
+The typing rules for processes are as follows:
+\begin{mathpar}
+  \inferrule[T-Inact]{\un{\Gamma}}{\types{\Gamma}{\Pend}}
+  \and
+  \inferrule[T-Par]{\types{\Gamma_1}{P} \\ \types{\Gamma_2}{Q}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Ppar{P}{Q}}}
+  \and
+  \inferrule[T-Res]{\types{\Cadd{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P}}{\types{\Gamma}{\Presd{x}{y}{P}}}
+  \and
+  \inferrule[T-Out]{\types{\Gamma_1}{\hastype{x}{\Tout{T}}} \\ \types{\Gamma_2}{\hastype{v}{\Tbase}} \\ \types{\Cupdate{\Gamma_3}{\hastype{x}{T}}}{P}}{\types{\Csplit{\Gamma_1}{\Csplit{\Gamma_2}{\Gamma_3}}}{\Pout{x}{v}{P}}}
+  \and
+  \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{(\Cadd{\Gamma_2}{\hastype{l}{\Tbase}})}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
+\end{mathpar}
+
+\subsubsection{Challenge}
+
+\begin{lemma}
+  If \( \types{\Gamma}{P} \) and \( \un{T} \) then \( \types{\Cadd{\Gamma}{\hastype{x}{T}}}{P} \).
+\end{lemma}
+
+\begin{lemma}
+  Assume \( \types{\Gamma}{P} \) and \( x \notin \freenames{P} \).
+  Then:
+  \begin{enumerate}
+  \item \( \hastype{x}{\Tout{T}} \notin \Gamma \).
+  \item \( \hastype{x}{\Tin{T}} \notin \Gamma \).
+  \item if \( \Gamma = \Cadd{\Gamma'}{\hastype{x}{T}} \) then \( \types{\Gamma'}{P} \).
+  \end{enumerate}
+\end{lemma}
+
+\begin{lemma}
+  If \( \types{\Gamma}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma}{Q} \).
+\end{lemma}
+
+\begin{lemma}
+  If \( \types{\Gamma_1}{\hastype{v}{T}} \) and \( \types{\Cadd{\Gamma_2}{\hastype{x}{T}}}{P} \) and \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then \( \types{\Gamma}{\subst{v}{x}{P}} \).
+\end{lemma}
+
+\begin{theorem}
+  If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
+\end{theorem}
+
+\begin{theorem}
+  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
+\end{theorem}
+
+\begin{theorem}
+  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
+\end{theorem}
+
 \subsection{Challenge: Name passing and scope extrusion}
@@
 
-\paragraph{Notes about this challenge}
-Strong barbed bisimilarity is interesting by itself for this challenge, but the proof that it is an equivalence relation does not require up-to techniques, which are wide spread and interesting.
-The definition of strong barbed congruence requires introducing contexts, but this is actually nice, since they are so prevalent in the field and we want to cover as many basics as possible.
-The theorem about the strong barbed congruence requires up-to techniques, but also seems to involve introducing a structural equivalence, which is possibly not desirable.
-On the other hand, it is a nice application of structural congruence which does not ``ruin'' the semantics.
-
-\subsection{Challenge: Linearity and behavioural type systems}
-\label{sec:challenge:linearity-beh-types}
-
-This challenge formalises a proof that requires reasoning about linearity of channels.
-Linearity is the notion that a channel must be used exactly once in a process.
-This is necessary to prove properties about session type systems.
-
-The setting for this challenge is a small calculus with a session type system, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove type preservation (also known as subject reduction) for the type system, i.e.\ that well-typed processes can only transition to processes which are also well-typed in the same context.
-
-\subsubsection{Syntax}
-We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-The syntax is then:
-\begin{align*}
-  v,w :=&&& a \\
-  |&&& l \\
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{v}{P} \\
-  |&&& \Pin{x}{l}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Presd{x}{y}{P}
-\end{align*}
-A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
-The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-The input operator thus binds the variable name \( l \) in \( P \).
-The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Presd{x}{y}{P} \) is the \emph{restriction} of the channel with endpoints named \( x \) and \( y \) to \( P \).
-Components in \( P \) can use the names \( x \) and \( y \) to interact with each other (sending on \( x \) and receiving on \( y \) or vice versa), but not with processes outside of the restriction.
-The restriction operator thus binds the names \( x \) and \( y \) in \( P \).
-Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-Note that there is no recursion of replication in the syntax, and thus no infinite behaviours can be expressed.
-
-We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
-We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
-We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
-We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
-We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-As a convention, we will identify \( \alpha \)-convertible processes.
-
-As a convention, we assume that bound names and bound variable names of any processes are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, substitutions, and sets of names or variable names.
-This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-We describe the actions that the system can perform using a small step operational semantics.
-To simplify the definition of the reduction relation, we will first factor out a structural congruence relation on processes, and for this we need a few definitions.
-
-\subsubsection{Contexts and congruences}
-A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-We will think of contexts as functions between processes.
-A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
-\begin{mathpar}
-  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
-  \and
-  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
-  \and
-  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
-  \and
-  \inferrule[Sc-Res-Par]{\{x,y\} \not\subseteq \freenames{Q}}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
-  \and
-  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
-  \and
-  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
-\end{mathpar}
-
-The operational semantics is then defined as the following binary relation on processes:
-\begin{mathpar}
-  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{v}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{v}{l}}{R}})}}}
-  \and
-  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
-  \and
-  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
-  \and
-  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathpar}
-Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable name or a name.
-
-\subsubsection{Session types}
-Our process syntax allows us to write processes that are not well-formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
-As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of binding names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
-Another example is the following process, which attempts to send a variable that is not resolved at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
-
-To alleviate this issue, we introduce a \emph{session type system}, which will detect processes that are not well-formed.
-To this end, we need to first formally define what we mean by well-formed.
-We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
-A process is then \emph{well-formed} if, for each of its structurally congruenct processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
-
-Note that well-formed processes do not necessarily reduce, since e.g.\ the process \( \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y}{x_2}{\Pin{y_1}{l}{\Pend}}})}} \) is well-formed, but has no reduction.
-
-\paragraph{Syntax}
-Our type system does not type processes directly, but instead focuses on the channels used in the process.
-The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
-\begin{align*}
-  S,T := &&& \Tend \\
-  |&&& \Tbase \\
-  |&&& \Tin{S} \\
-  |&&& \Tout{S} \\
-  \Gamma := &&& \Cempty \\
-  |&&& \Cadd{\Gamma}{\hastype{x}{S}}
-\end{align*}
-The \emph{end type}, \( \Tend \), is the type of an endpoint on which no further interaction is possible.
-The \emph{base type}, \( \Tbase \), is the type of base values.
-The \emph{input type}, \( \Tin{S} \), is the type of an endpoint which is ready to receive a value, then continue with type \( S \).
-The \emph{output type}, \( \Tout{S} \), is the type of an endpoint which is ready to send a value, then continue with type \( S \).
-
-Typing contexts gather type information about names.
-The \emph{empty type context}, \( \Cempty \), does not contain any information.
-The \emph{type context with \( x \) added}, \( \Cadd{\Gamma}{\hastype{x}{T}} \) adds the information that \( x \) has type \( T \) to the existing type context \( \Gamma \).
-All of the names added to a type context must be distinct.
-We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
-
-Since we need to determine whether endpoints are used dually to determine whether processes are well-formed, we will need to formally define the dual of a type.
-The dual of a type, \( \dual{T} \), is defined by the following equations:
-\begin{mathpar}
-  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tend} = \Tend}
-\end{mathpar}
-Note that the dual function is not total, since it is not defined for base types.
-
-\paragraph{Splitting and updating contexts}
-Our type system will maintain two invariants:
-\begin{enumerate}
-\item Endpoints are used exactly once
-\item Endpoints that are part of the same restriction have dual types
-\end{enumerate}
-The first invariant is maintained by linearly splitting type contexts when typing compositions of processes.
-The second invariant is maintained by requiring duality when typing restrictions.
-
-To keep track of linearity, we introduce two \emph{qualifiers}: \( \un{T} \) and \( \lin{T} \).
-A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
-All types are \emph{linear}, \( \lin{T} \).
-The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
-
-The \emph{linear split} of a context, \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Split-Empty]{ }{\Cempty = \Csplit{\Cempty}{\Cempty}}
-  \and
-  \inferrule[Split-Un]{\Csplit{\Gamma_1}{\Gamma_2} = \Gamma \\ \un{T}}{\Cadd{\Gamma}{\hastype{x}{T}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{T}})}{(\Cadd{\Gamma_2}{\hastype{x}{T}})}}
-  \and
-  \inferrule[Split-Lin-L]{\Gamma = \Csplit{\Gamma_1}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{S}})}{\Gamma_2}}
-  \and
-  \inferrule[Split-Lin-R]{\Gamma = \Csplit{\Gamma_2}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{\Gamma_1}{(\Cadd{\Gamma_2}{\hastype{x}{S}})}}
-\end{mathpar}
-
-We will also need to update contexts with new type information in a safe way while typing input and output.
-The \emph{context update}, \( \Cupdate{\Gamma}{\hastype{x}{T}} = \Gamma' \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Update-Name]{\hastype{x}{U} \notin \Gamma}{\Cupdate{\Gamma}{\hastype{x}{T}} = \Cadd{\Gamma}{\hastype{x}{T}}}
-  \and
-  \inferrule[Update-Un]{\un{T}}{\Cupdate{(\Cadd{\Gamma}{\hastype{x}{T}})}{\hastype{x}{T}} = (\Cadd{\Gamma}{\hastype{x}{T}})}
-\end{mathpar}
-
-\subsubsection{Typing rules}
-We have three typing judgements: one for values, one for names, and one for processes.
-The typing rules for values are as follows:
-\begin{mathpar}
-  \inferrule[T-Base]{\un{\Gamma}}{\types{\Gamma}{\hastype{a}{\Tbase}}}
-  \and
-  \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{v}{\Tbase}}}{\hastype{v}{\Tbase}}}
-\end{mathpar}
-The only typing rule for names is:
-\begin{mathpar}
-  \inferrule[T-Name]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{x}{T}}}
-\end{mathpar}
-The typing rules for processes are as follows:
-\begin{mathpar}
-  \inferrule[T-Inact]{\un{\Gamma}}{\types{\Gamma}{\Pend}}
-  \and
-  \inferrule[T-Par]{\types{\Gamma_1}{P} \\ \types{\Gamma_2}{Q}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Ppar{P}{Q}}}
-  \and
-  \inferrule[T-Res]{\types{\Cadd{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P}}{\types{\Gamma}{\Presd{x}{y}{P}}}
-  \and
-  \inferrule[T-Out]{\types{\Gamma_1}{\hastype{x}{\Tout{T}}} \\ \types{\Gamma_2}{\hastype{v}{\Tbase}} \\ \types{\Cupdate{\Gamma_3}{\hastype{x}{T}}}{P}}{\types{\Csplit{\Gamma_1}{\Csplit{\Gamma_2}{\Gamma_3}}}{\Pout{x}{v}{P}}}
-  \and
-  \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{(\Cadd{\Gamma_2}{\hastype{l}{\Tbase}})}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
-\end{mathpar}
-
-\subsubsection{Challenge}
-
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \un{T} \) then \( \types{\Cadd{\Gamma}{\hastype{x}{T}}}{P} \).
-\end{lemma}
-
-\begin{lemma}
-  Assume \( \types{\Gamma}{P} \) and \( x \notin \freenames{P} \).
-  Then:
-  \begin{enumerate}
-  \item \( \hastype{x}{\Tout{T}} \notin \Gamma \).
-  \item \( \hastype{x}{\Tin{T}} \notin \Gamma \).
-  \item if \( \Gamma = \Cadd{\Gamma'}{\hastype{x}{T}} \) then \( \types{\Gamma'}{P} \).
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma}{Q} \).
-\end{lemma}
-
-\begin{lemma}
-  If \( \types{\Gamma_1}{\hastype{v}{T}} \) and \( \types{\Cadd{\Gamma_2}{\hastype{x}{T}}}{P} \) and \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then \( \types{\Gamma}{\subst{v}{x}{P}} \).
-\end{lemma}
-
-\begin{theorem}
-  If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
-\end{theorem}
-
-\begin{theorem}
-  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
-\end{theorem}
-
-\begin{theorem}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
-\end{theorem}
+%\paragraph{Notes about this challenge}
+%Strong barbed bisimilarity is interesting by itself for this challenge, but the proof that it is an equivalence relation does not require up-to techniques, which are wide spread and interesting.
+%The definition of strong barbed congruence requires introducing contexts, but this is actually nice, since they are so prevalent in the field and we want to cover as many basics as possible.
+%The theorem about the strong barbed congruence requires up-to techniques, but also seems to involve introducing a structural equivalence, which is possibly not desirable.
+%On the other hand, it is a nice application of structural congruence which does not ``ruin'' the semantics.
 

### 2023-07-10 extended-abstract/extended-abstract.tex
adding some entries to references.bib (using dblp standard style) and having the extended abstract point to it. Removed bibliograpy.bib to have just one bib file
adding a txt dir for minutes etc
Change the order of challenges.
Add authors.

@@
 \def\titlerunning{The Concurrent Calculi Formalisation Benchmark (Extended Abstract)}
+
+\author{
+     Marco Carbone \institute{IT University of Copenhagen} \email{maca@itu.dk}
+\and David Castro-Perez \institute{University of Kent} \email{D.Castro-Perez@kent.ac.uk}
+\and Francisco Ferreira \institute{Royal Holloway, University of London} \email{Francisco.FerreiraRuiz@rhul.ac.uk}    \and Lorenzo Gheri \institute{University of Oxford} \email{lorenzo.gheri@cs.ox.ac.uk}
+\and Frederik Krogsdal Jacobsen \institute{Technical University of Denmark \thanks{The work was done while visiting the University of Oxford}} \email{fkjac@dtu.dk}
+\and Alberto Momigliano \institute{Università degli Studi di Milano} \email{momigliano@di.unimi.it}
+\and Luca Padovani \institute{Università di Camerino} \email{luca.padovani@unicam.it}
+\and Alceste Scalas \institute{Technical University of Denmark} \email{alcsc@dtu.dk}
+\and Martin Vassor \institute{University of Oxford} \email{martin.vassor@cs.ox.ac.uk}
+\and Nobuko Yoshida \institute{University of Oxford} \email{nobuko.yoshida@cs.ox.ac.uk}
+}
+\def\authorrunning{Carbone, Castro-Perez, Ferreira, Gheri, Jacobsen, Momigliano, Padovani, Scalas, Vassor \& Yoshida}
 
@@
 In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
+
+Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
+Session type systems require linear handling of typing contexts, and the proof involves reasoning about the linearity of context splitting operations.
+
 Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing caluclus without infinite behaviour, but with non-deterministic choice.
@@
 The proof involves reasoning coinductively about the infinite behaviours of the replication operator.
-
-Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
-Session type systems require linear handling of typing contexts, and the proof involves reasoning about the linearity of context splitting operations.
 
@@
 \bibliographystyle{eptcs}
-\bibliography{bibliography}
+% \bibliography{bibliography}
+\bibliography{../references}
 

### 2023-07-10 extended-abstract/extended-abstract.tex
Change "extended abstract" to "talk proposal".

@@
 
-\title{The Concurrent Calculi Formalisation Benchmark\\{\normalsize(Extended Abstract)}}
-\def\titlerunning{The Concurrent Calculi Formalisation Benchmark (Extended Abstract)}
+\title{The Concurrent Calculi Formalisation Benchmark\\{\normalsize(Talk Proposal)}}
+\def\titlerunning{The Concurrent Calculi Formalisation Benchmark (Talk Proposal)}
 

### 2023-07-10 extended-abstract/extended-abstract.tex
Reorder textual appearances of challenges.
add reference to existing formalisations

@@
 For this reason, we propose a new collection of benchmark problems with a focus on the issues encountered when mechanizing message-passing concurrency using process calculi.
-  Our benchmark problems concern the mechanization of scope extrusion, linear handling of environments, and coinductive reasoning.
+  Our benchmark problems concern the mechanization of linear handling of environments, scope extrusion, and coinductive reasoning.
   Like previous benchmarks, our goal is to clarify, compare and advance the state of the art.
@@
 
-Papers in concurrency theory do often include mechanized proofs, but our experience is that choosing techniques and tools remains a challenge with few guidelines.
+Papers in concurrency theory do often include mechanized proofs (e.g.~\cite{DBLP:conf/tacas/CastroFY20, DBLP:conf/pldi/Castro-Perez0GY21,lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and sometimes even target formalisation techniques as their main subject (e.g.~\cite{DBLP:conf/tphol/Gay01, DBLP:journals/jar/BengtsonPW16, DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).
+However, our experience is that choosing techniques and tools remains a challenge with few guidelines.
 We propose a new collection of benchmark problems specifically designed to tackle the issues encountered when mechanizing message-passing concurrency using process calculi.
@@
 \begin{itemize}
+\item Linearity in behavioural type systems
 \item Scope extrusion in name passing systems
-\item Linearity in behavioural type systems
 \item Infinite behaviour modelled coinductively
@@
 
-In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
-
 Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
@@
 
+In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
 Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing caluclus without infinite behaviour, but with non-deterministic choice.

### 2023-07-11 talk-proposal/talk-proposal.tex
Rename "extended abstract" to "talk proposal".

--- 2023-07-10 extended-abstract/extended-abstract.tex
@@
 \usepackage{cleveref}
-\usepackage{draftwatermark}
+% \usepackage{draftwatermark}
 \usepackage{underscore}
@@
 
-\SetWatermarkText{Draft}
-\SetWatermarkScale{2}
-\SetWatermarkColor[gray]{0.90}
+% \SetWatermarkText{Draft}
+% \SetWatermarkScale{2}
+% \SetWatermarkColor[gray]{0.90}
 
@@
 
-Papers in concurrency theory do often include mechanized proofs (e.g.~\cite{DBLP:conf/tacas/CastroFY20, DBLP:conf/pldi/Castro-Perez0GY21,lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and sometimes even target formalisation techniques as their main subject (e.g.~\cite{DBLP:conf/tphol/Gay01, DBLP:journals/jar/BengtsonPW16, DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).
+Papers in concurrency theory do often include mechanized proofs (e.g.~\cite{DBLP:conf/tacas/CastroFY20, DBLP:conf/pldi/Castro-Perez0GY21,lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also directly discuss formalisation techniques (e.g.~\cite{DBLP:conf/tphol/Gay01, DBLP:journals/jar/BengtsonPW16, DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).
 However, our experience is that choosing techniques and tools remains a challenge with few guidelines.

### 2023-07-11 challenges.tex
Improve the presentation of the final theorems in the linearity challenge.

@@
 \subsubsection{Challenge}
+The objective of this challenge is to prove the following theorems:
+\begin{theorem}
+  If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
+\end{theorem}
 
+\begin{theorem}
+  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
+\end{theorem}
+
+\begin{corollary}
+  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
+\end{corollary}
+
+The following lemmas may be of use:
 \begin{lemma}
@@
 \end{lemma}
-
-\begin{theorem}
-  If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
-\end{theorem}
-
-\begin{theorem}
-  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
-\end{theorem}
-
-\begin{theorem}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
-\end{theorem}
 

### 2023-07-11 talk-proposal/talk-proposal.tex
Clean up talk proposal for submission.
Change "-ize" to "-ise".

--- 2023-07-10 extended-abstract/extended-abstract.tex
@@
-\documentclass[adraft,copyright,creativecommons]{eptcs}
+\documentclass[submission,copyright,creativecommons]{eptcs}
+\providecommand{\event}{ST30}
 
@@
 \usepackage{cleveref}
-\usepackage{draftwatermark}
 \usepackage{underscore}
 \usepackage[T1]{fontenc}
-
-\SetWatermarkText{Draft}
-\SetWatermarkScale{2}
-\SetWatermarkColor[gray]{0.90}
 
@@
 
+\hypersetup{
+  bookmarksnumbered,
+  pdftitle    = {\titlerunning},
+  pdfauthor   = {\authorrunning},
+  pdfsubject  = {ST30},
+  pdfkeywords = {mechanisation, process calculi, coinduction, scope extrusion, linearity}
+}
+
 \begin{document}
@@
   with benchmark problems that do not address concurrency.
-For this reason, we propose a new collection of benchmark problems with a focus on the issues encountered when mechanizing message-passing concurrency using process calculi.
-  Our benchmark problems concern the mechanization of scope extrusion, linear handling of environments, and coinductive reasoning.
+For this reason, we propose a new collection of benchmark problems with a focus on the issues encountered when mechanising message-passing concurrency using process calculi.
+  Our benchmark problems concern the mechanisation of linear handling of environments, scope extrusion, and coinductive reasoning.
   Like previous benchmarks, our goal is to clarify, compare and advance the state of the art.
@@
 \section{Introduction}
-The programming language community has increasingly embraced mechanizations since the influential POPLMark challenge~\cite{POPLMark}, but we are still far from the vision of a mechanized proof accompanying every paper in the field.
-Since POPLMark, the number of papers with accompanying mechanized proofs has grown steadily, but many mechanizations still rely on ad hoc techniques.
+The programming language community has increasingly embraced mechanisations since the influential POPLMark challenge~\cite{POPLMark}, but we are still far from the vision of a mechanised proof accompanying every paper in the field.
+Since POPLMark, the number of papers with accompanying mechanised proofs has grown steadily, but many mechanisations still rely on ad hoc techniques.
 The POPLMark Reloaded challenge~\cite{POPLMarkReloaded} aimed to identify and develop best practices and tool support for proofs using logical relations.
@@
 
-Papers in concurrency theory do often include mechanized proofs, but our experience is that choosing techniques and tools remains a challenge with few guidelines.
-We propose a new collection of benchmark problems specifically designed to tackle the issues encountered when mechanizing message-passing concurrency using process calculi.
+Papers in concurrency theory do often include mechanised proofs (e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also directly discuss formalisation techniques (e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01, DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).
+However, our experience is that choosing techniques and tools remains a challenge with few guidelines.
+We propose a new collection of benchmark problems specifically designed to tackle the issues encountered when mechanising message-passing concurrency using process calculi.
 A goal of our collection of benchmark problems is to enable comparison between techniques and push for the development of guidelines and tutorials, and our challenge problems are a simple setting in which to demonstrate and compare techniques.
-Another goal is to strengthen the culture of mechanization by rallying the community to collaborate on exploring and developing new tools and techniques.
+Another goal is to strengthen the culture of mechanisation by rallying the community to collaborate on exploring and developing new tools and techniques.
 
-We have identified three key aspects which cause issues when mechanizing concurrency theory:
+We have identified three key aspects which cause issues when mechanising concurrency theory:
 \begin{itemize}
+\item Linearity in behavioural type systems
 \item Scope extrusion in name passing systems
-\item Linearity in behavioural type systems
 \item Infinite behaviour modelled coinductively
@@
 Each part of our benchmark exercises the three aspects independently to enable comparison and to keep the challenge problems small.
-The three aspects are of course not the only ones that may cause issues in mechanizations, but they are fundamental to the field and will thus appear in most mechanizations of concurrency theory.
-Actual mechanizations will additionally often require more than just one aspect, but we will not address the combination of techniques except as optional bonus challenges.
+The three aspects are of course not the only ones that may cause issues in mechanisations, but they are fundamental to the field and will thus appear in most mechanisations of concurrency theory.
+Actual mechanisations will additionally often require more than just one aspect, but we will not address the combination of techniques except as optional bonus challenges.
 
@@
 Our challenge problems cannot address every technique and subfield at once.
-We have identified three key aspects fundamental to concurrency theory mechanizations and designed our challenge problems to exercise these.
+We have identified three key aspects fundamental to concurrency theory mechanisations and designed our challenge problems to exercise these.
 Like previous benchmark authors, we would like to stress that our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types and psi-calculi are not directly covered by our challenge problems.
@@
 \begin{itemize}
-\item The mechanization overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanization.
-\item The adequacy of the formal statements in the mechanization, i.e.\ whether the proven theorems are easily recognizable as the theorems from the challenge.
+\item The mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation.
+\item The adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge.
 \item The cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
@@
 \bibliographystyle{eptcs}
-% \bibliography{bibliography}
 \bibliography{../references}

### 2023-07-11 talk-proposal/talk-proposal.tex
Remove reference to pen and paper solutions and clarify that current challenges a drafts.
Various textual improvements.

@@
 In the rest of this paper we discuss the design of the challenge and briefly describe the three challenge problems.
-The appendix contains the full text of the challenge problems and pen and paper solutions.
+The appendix contains the full text of the current draft of the challenge problems.
 
@@
 Our challenge problems cannot address every technique and subfield at once.
-We have identified three key aspects fundamental to concurrency theory mechanisations and designed our challenge problems to exercise these.
+We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
 Like previous benchmark authors, we would like to stress that our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types and psi-calculi are not directly covered by our challenge problems.
@@
 We have designed our challenge problems to exercise the three aspects independently, and they can be solved individually and in any order.
-We have designed each problem to be small and easy to understand with a basic knowledge of textbook concurrency theory and process calculi.
+We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory and process calculi.
 Solutions to our challenge problems should be compared on three measures:
@@
 \end{itemize}
-Solutions to our challenges do not need to strictly follow the definitions and techniques set out in the problem text, but solutions which deviate from the original challenges will need to present more elaborate argumentation for their adequacy.
+Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges will need to present more elaborate argumentation for their adequacy.
 
 \section{The challenge problems}
-In this section we will briefly describe our three challenge problems, referring to the appendix for the full text and pen and paper solution to each problem.
+In this section we will briefly describe our three challenge problems, referring to the appendix for the full text of each challenge.
 
@@
 Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing caluclus without infinite behaviour, but with non-deterministic choice.
-The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues with some approaches to binders.
+The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues for some approaches to mechanising binders.
 

### 2023-07-12 talk-proposal/talk-proposal.tex
Fix a typo.

@@
 In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
-Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing caluclus without infinite behaviour, but with non-deterministic choice.
+Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing calculus without infinite behaviour, but with non-deterministic choice.
 The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues for some approaches to mechanising binders.

### 2023-07-12 talk-proposal/talk-proposal.tex
Remove a redundant sentence.

@@
 \section{Design of the challenge}
-Our challenge problems cannot address every technique and subfield at once.
 We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.

### 2023-07-12 talk-proposal/talk-proposal.tex
Clarify that "the previous benchmarks" in the first paragraph refer to POPLMark and POPLMark Reloaded.

@@
 The POPLMark Reloaded challenge~\cite{POPLMarkReloaded} aimed to identify and develop best practices and tool support for proofs using logical relations.
-The authors of both POPLMark and POPLMark Reloaded were careful to note that their benchmarks were only a beginning, and the previous benchmarks specifically mention reasoning about concurrency using coinduction and linear environments as points for future work.
+The authors of both POPLMark and POPLMark Reloaded were careful to note that their benchmarks were only a beginning, and the aforementioned benchmarks specifically mention reasoning about concurrency using coinduction and linear environments as points for future work.
 

### 2023-07-12 challenges.tex
Clarify that an inductive (or coinductive) proof principle is needed in the definition of strong barbed bisimilarity in the scope extrusion challenge.

@@
 
-Two processes \( P \) and \( Q \) are then \emph{strongly barbed bisimilar}, written \( \sbbisim{P}{Q} \), if:
+Two processes \( P \) and \( Q \) are then \emph{strong barbed bisimilar}, written \( \sbbisim{P}{Q} \), if:
 \begin{gather}
@@
 \end{gather}
-Note that, since our calculus does not have infinite behaviour, we do not specify that the relation is the largest of its kind, and the relation is thus not coinductively defined.
+Note that, since our calculus does not have infinite behaviour, there is only one relation which satisfies these criteria.
+Thus we do not specify that the relation is the largest of its kind, and the relation does not need to be coinductively defined.
+To actually prove properties about the relation, however, we will still need either an inductive (or a coinductive) proof principle for the relation.
 

### 2023-07-13 talk-proposal/talk-proposal.tex
Add type theory as a subject from which knowledge is required to understand the challenges.
Remove white space and some redundant words/sentences to bring the paper under 2 pages.

--- 2023-07-12 talk-proposal/talk-proposal.tex
@@
 \begin{abstract}
-  The POPLMark challenge has sparked a flurry of work to move us closer to a world in which every paper on programming languages is accompanied by machine-checked proofs.
+  The POPLMark challenge sparked a flurry of work on machine-checked proofs.
   However, POPLMark (and its follow-up POPLMark Reloaded) were explicitly limited in scope,
@@
 For this reason, we propose a new collection of benchmark problems with a focus on the issues encountered when mechanising message-passing concurrency using process calculi.
-  Our benchmark problems concern the mechanisation of linear handling of environments, scope extrusion, and coinductive reasoning.
-  Like previous benchmarks, our goal is to clarify, compare and advance the state of the art.
+  Our benchmark problems concern linear handling of environments, scope extrusion, and coinductive reasoning.
+  Our goal is to clarify, compare and advance the state of the art.
 \end{abstract}
@@
 Since POPLMark, the number of papers with accompanying mechanised proofs has grown steadily, but many mechanisations still rely on ad hoc techniques.
-The POPLMark Reloaded challenge~\cite{POPLMarkReloaded} aimed to identify and develop best practices and tool support for proofs using logical relations.
-The authors of both POPLMark and POPLMark Reloaded were careful to note that their benchmarks were only a beginning, and the previous benchmarks specifically mention reasoning about concurrency using coinduction and linear environments as points for future work.
-
+POPLMark Reloaded~\cite{POPLMarkReloaded} aimed to identify and develop best practices and tool support for proofs using logical relations.
+The authors of both POPLMark and POPLMark Reloaded were careful to note that their benchmarks were only a beginning, and they specifically mention reasoning about concurrency using coinduction and linear environments as points for future work.
 Papers in concurrency theory do often include mechanised proofs (e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also directly discuss formalisation techniques (e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01, DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).
@@
 We propose a new collection of benchmark problems specifically designed to tackle the issues encountered when mechanising message-passing concurrency using process calculi.
-A goal of our collection of benchmark problems is to enable comparison between techniques and push for the development of guidelines and tutorials, and our challenge problems are a simple setting in which to demonstrate and compare techniques.
+A goal of our collection of benchmark problems is to push for the development of guidelines and tutorials, and our challenge problems are a simple setting in which to demonstrate and compare techniques.
 Another goal is to strengthen the culture of mechanisation by rallying the community to collaborate on exploring and developing new tools and techniques.
-
-We have identified three key aspects which cause issues when mechanising concurrency theory:
-\begin{itemize}
-\item Linearity in behavioural type systems
-\item Scope extrusion in name passing systems
-\item Infinite behaviour modelled coinductively
-\end{itemize}
+We have identified three key aspects which cause issues when mechanising concurrency theory: linearity in behavioural type systems, scope extrusion in name passing systems and infinite behaviour modelled coinductively.
 Each part of our benchmark exercises the three aspects independently to enable comparison and to keep the challenge problems small.
 The three aspects are of course not the only ones that may cause issues in mechanisations, but they are fundamental to the field and will thus appear in most mechanisations of concurrency theory.
-Actual mechanisations will additionally often require more than just one aspect, but we will not address the combination of techniques except as optional bonus challenges.
-
+Actual mechanisations will additionally often require more than just one aspect, but we do not address the combination of techniques.
 In the rest of this paper we discuss the design of the challenge and briefly describe the three challenge problems.
-The appendix contains the full text of the current draft of the challenge problems.
 
@@
 Most applications will however still need the basic techniques that our challenge problems exercise.
-
 We have designed our challenge problems to exercise the three aspects independently, and they can be solved individually and in any order.
-We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory and process calculi.
-Solutions to our challenge problems should be compared on three measures:
-\begin{itemize}
-\item The mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation.
-\item The adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge.
-\item The cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
-\end{itemize}
+We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.
+Solutions to our challenge problems should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
 Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges will need to present more elaborate argumentation for their adequacy.
@@
 In this section we will briefly describe our three challenge problems, referring to the appendix for the full text of each challenge.
-
 Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
 Session type systems require linear handling of typing contexts, and the proof involves reasoning about the linearity of context splitting operations.
-
 In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
@@
 The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues for some approaches to mechanising binders.
-
 Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed bisimulation is an equivalence relation for an untyped calculus of communicating systems with (infinite) replication of processes.
 The proof involves reasoning coinductively about the infinite behaviours of the replication operator.
-
 We invite anyone interested to send us solutions to the challenge problems and to publish experience reports detailing the advantages and disadvantages of their approaches.

### 2023-07-13 challenges.tex
Fix overfull hbox caused by a long process description.

@@
 
-Note that well-formed processes do not necessarily reduce, since e.g.\ the process \( \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y}{x_2}{\Pin{y_1}{l}{\Pend}}})}} \) is well-formed, but has no reduction.
-
+Note that well-formed processes do not necessarily reduce, since e.g.\ the process
+\begin{equation*}
+\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
+\end{equation*}
+is well-formed, but has no reduction.
 \paragraph{Syntax}

### 2023-07-13 challenges.tex
Clarify that linearity is also necessary for other things than just session types.
Add some motivation for explicitly reasoning about scope extrusion to the challenge description in the appendix.

@@
 This is necessary to prove properties about session type systems.
+Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
 
@@
 This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
+
+Reasoning about scope extrusion explicitly can sometimes be avoided by introducing a structural congruence rule into the semantics, but doing this means we lose information about the scope when reasoning about the semantics.
+Explicitly reasoning about scope extrusion is necessary to describe e.g.\ monitors and compositions.
 

### 2023-07-19 talk-proposal-ST30/talk-proposal.tex
Reorganize repository structure.

--- 2023-07-13 talk-proposal/talk-proposal.tex
@@
 Another goal is to strengthen the culture of mechanisation by rallying the community to collaborate on exploring and developing new tools and techniques.
-We have identified three key aspects which cause issues when mechanising concurrency theory: linearity in behavioural type systems, scope extrusion in name passing systems and infinite behaviour modelled coinductively.
+We have identified three key aspects which cause issues when mechanising concurrency theory: linearity, scope extrusion in name passing systems and infinite behaviour modelled coinductively.
 Each part of our benchmark exercises the three aspects independently to enable comparison and to keep the challenge problems small.

### 2023-07-19 technical-report/main.tex
Reorganize repository structure.

--- 2023-04-26 main.tex
@@
 \usepackage{cleveref}
+\usepackage{draftwatermark}
 
-\input{macros.tex}
+\SetWatermarkText{Draft}
+\SetWatermarkScale{5}
+\SetWatermarkColor[gray]{0.90}
+
+\input{../macros.tex}
 
@@
 machine checked proofs for theoretical results in papers. The
-POPLMark~\cite{Pierce:2005} challenge was proposed for this purpose,
+POPLMark~\cite{POPLMark} challenge was proposed for this purpose,
 and to get the community on track. Since then the number of formal
@@
 new challenges are needed, work like POPLMark
-Reloaded~\cite{Pientka:2019} strives to extend the scope of the
+Reloaded~\cite{POPLMarkReloaded} strives to extend the scope of the
 original challenge to proofs using logical relations. This technique
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{images/benchmark.jpeg}
+  \includegraphics[width=0.75\textwidth]{../images/benchmark.jpeg}
   \caption{The benchmark's big picture.}
@@
 
-\input{challenges.tex}
+\input{../challenges.tex}
 
@@
 \bibliographystyle{plainnat}
-\bibliography{references}
+\bibliography{../references}
 

### 2023-07-28 paper/paper.tex
Add copy of proposal with LLNCS layout.

--- 2023-07-19 talk-proposal-ST30/talk-proposal.tex
@@
-\documentclass[submission,copyright,creativecommons]{eptcs}
-\providecommand{\event}{ST30}
-
+\documentclass[runningheads]{llncs}
 \usepackage[utf8]{inputenc}
-\usepackage{graphicx}
+\usepackage[T1]{fontenc}
 \usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage{amsthm}
+\usepackage{graphicx}
 \usepackage{mathpartir}
+\usepackage{cite}
 \usepackage{bm}
@@
 \usepackage{cleveref}
-\usepackage{underscore}
-\usepackage[T1]{fontenc}
+\usepackage{hyperref}
+\usepackage{color}
+\renewcommand\UrlFont{\color{blue}\rmfamily}
 
@@
 
-\title{The Concurrent Calculi Formalisation Benchmark\\{\normalsize(Talk Proposal)}}
-\def\titlerunning{The Concurrent Calculi Formalisation Benchmark (Talk Proposal)}
+\begin{document}
+
+\title{The Concurrent Calculi Formalisation Benchmark}
 
 \author{
-     Marco Carbone \institute{IT University of Copenhagen} \email{maca@itu.dk}
-\and David Castro-Perez \institute{University of Kent} \email{D.Castro-Perez@kent.ac.uk}
-\and Francisco Ferreira \institute{Royal Holloway, University of London} \email{Francisco.FerreiraRuiz@rhul.ac.uk}    \and Lorenzo Gheri \institute{University of Oxford} \email{lorenzo.gheri@cs.ox.ac.uk}
-\and Frederik Krogsdal Jacobsen \institute{Technical University of Denmark \thanks{The work was done while visiting the University of Oxford}} \email{fkjac@dtu.dk}
-\and Alberto Momigliano \institute{Università degli Studi di Milano} \email{momigliano@di.unimi.it}
-\and Luca Padovani \institute{Università di Camerino} \email{luca.padovani@unicam.it}
-\and Alceste Scalas \institute{Technical University of Denmark} \email{alcsc@dtu.dk}
-\and Martin Vassor \institute{University of Oxford} \email{martin.vassor@cs.ox.ac.uk}
-\and Nobuko Yoshida \institute{University of Oxford} \email{nobuko.yoshida@cs.ox.ac.uk}
-}
-\def\authorrunning{Carbone, Castro-Perez, Ferreira, Gheri, Jacobsen, Momigliano, Padovani, Scalas, Vassor \& Yoshida}
-
-\hypersetup{
-  bookmarksnumbered,
-  pdftitle    = {\titlerunning},
-  pdfauthor   = {\authorrunning},
-  pdfsubject  = {ST30},
-  pdfkeywords = {mechanisation, process calculi, coinduction, scope extrusion, linearity}
+     Marco Carbone \inst{1}\orcidID{0000-0001-9479-2632}
+\and David Castro-Perez \inst{2}\orcidID{0000-0002-6939-4189}
+\and Francisco Ferreira \inst{3}\orcidID{0000-0001-8494-7696}
+\and Lorenzo Gheri \inst{4}\orcidID{0000-0002-3191-7722}
+\and Frederik Krogsdal Jacobsen \inst{5}\orcidID{0000-0003-3651-8314}
+\and Alberto Momigliano \inst{6}\orcidID{0000-0003-0942-4777}
+\and Luca Padovani \inst{7}\orcidID{0000-0001-9097-1297}
+\and Alceste Scalas \inst{5}\orcidID{0000-0002-1153-6164}
+\and Martin Vassor \inst{4}\orcidID{0000-0002-2057-0495}
+\and Nobuko Yoshida \inst{4}\orcidID{0000-0002-3925-8557}
 }
 
-\begin{document}
+\institute{
+     IT University of Copenhagen \email{maca@itu.dk}
+\and University of Kent \email{D.Castro-Perez@kent.ac.uk}
+\and Royal Holloway, University of London \email{Francisco.FerreiraRuiz@rhul.ac.uk}
+\and University of Oxford \email{lorenzo.gheri@cs.ox.ac.uk}, \email{martin.vassor@cs.ox.ac.uk}, \email{nobuko.yoshida@cs.ox.ac.uk}
+\and Technical University of Denmark \email{fkjac@dtu.dk}, \email{alcsc@dtu.dk}
+\and Università degli Studi di Milano \email{momigliano@di.unimi.it}
+\and Università di Camerino \email{luca.padovani@unicam.it}
+}
+
+\authorrunning{M. Carbone et al.}
+
 \maketitle
@@
   Our goal is to clarify, compare and advance the state of the art.
+
+\keywords{Mechanisation \and Process calculi \and Benchmark \and Coinduction \and Scope extrusion \and Linearity}
 \end{abstract}
@@
 
-\bibliographystyle{eptcs}
+\bibliographystyle{splncs04}
 \bibliography{../references}
@@
 \appendix
-
 \section{Challenges}

### 2023-08-01 challenges.tex
some small typos fixed (hopefully)

@@
 \subsubsection{Syntax}
-We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-The syntax is then:
+
+% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
+We assume the existence of some type of \emph{base values}, represented by the symbols
+ \( a, b, \dots \), the existence of some type
+of \emph{variable names}, represented by the symbols
+\( l, m, \dots \), and the existence of some type of \emph{names},
+represented by the symbols\( x, y, \dots \).  The syntax is
+then:
 \begin{align*}
@@
 The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
+The process \( \Pout{x}{v}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
 The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
@@
 We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
-We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
+We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
 
@@
 Our process syntax allows us to write processes that are not well-formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
-As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of binding names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
+As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of restricting names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
 Another example is the following process, which attempts to send a variable that is not resolved at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
@@
 \begin{equation*}
-\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
+  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
 \end{equation*}
@@
   |&&& \Cadd{\Gamma}{\hastype{x}{S}}
+       %% AM: added
+        \\
+  |&&& \Cadd{\Gamma}{\hastype{l}{S}}
 \end{align*}
@@
 A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
-All types are \emph{linear}, \( \lin{T} \).
+All other types are \emph{linear}, \( \lin{T} \).
 The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
@@
   \and
-  \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{v}{\Tbase}}}{\hastype{v}{\Tbase}}}
+  \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{l}{\Tbase}}}{\hastype{l}{\Tbase}}}
 \end{mathpar}

### 2023-09-06 slides-ST30/template/settings.tex
Add first draft of slides for ST30.

% You might want to make the front/back page background colour the first colour in the plot cycle list.
\pgfplotscreateplotcyclelist{DTU}{%
dtured,         fill=dtured,        \\%
blue,           fill=blue,          \\%
brightgreen,    fill=brightgreen    \\%
navyblue,       fill=navyblue       \\%
yellow,         fill=yellow         \\%
orange,         fill=orange         \\%
grey,           fill=grey           \\%
red,            fill=red            \\%
green,          fill=green          \\%
purple,         fill=purple         \\%
}

% Table of contents (TOC) and numbering of headings
\setcounter{tocdepth}{1}    % Depth of table of content: sub sections will not be included in table of contents
\setcounter{secnumdepth}{2} % Depth of section numbering: sub sub sections are not numbered

### 2023-09-13 challenges.tex
adding dual presentation of session types (with intuitionistic and linear context separated). Now this is called in technical-report but it can be enabled/disabled elsewhere (file appst.tex)
- fixing subst in A3
- ednote in paper intro

- fix in dual typing rules

@@
 \( l, m, \dots \), and the existence of some type of \emph{names},
-represented by the symbols\( x, y, \dots \).  The syntax is
-then:
+represented by the symbols \( x, y, \dots \).  The syntax is
+given by the grammar:
 \begin{align*}
@@
 A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
-All other types are \emph{linear}, \( \lin{T} \).
+All types are \emph{linear}, \( \lin{T} \).
 The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
@@
 The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that barbed bisimulation for this calculus is an equivalence relation.
+The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
 
@@
 
-We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
+We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
+
+% We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
 Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.

### 2023-09-13 paper/paper.tex
- fixing subst in A3
- ednote in paper intro

- fix in dual typing rules

@@
 The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues for some approaches to mechanising binders.
-Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed bisimulation is an equivalence relation for an untyped calculus of communicating systems with (infinite) replication of processes.
+Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed bisimulation is an equivalence relation \footnote{AM: this seems wrong, see:  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).}for an untyped calculus of communicating systems with (infinite) replication of processes.
 The proof involves reasoning coinductively about the infinite behaviours of the replication operator.

### 2023-09-18 challenges.tex
added a Preliminaries section to avoid repetitions. Please checl that it mskes sense
depending on what we do, we could also put there notiond duch as barbs

@@
-\subsection{Challenge: Linearity and behavioural type systems}
+\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
+%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
+\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
+                         {\endgroup]\end{quote}}
+\long\def\ignore#1{}
+
+\subsection{Preliminaries}
+\label{sec:prelim}
+
+We list here some common notions that we will use in the following. Since the calculi under study are somewhat different, we will note in each section which will apply.
+
+We assume the existence of some type of \emph{base values}, represented by the symbols
+\( a, b, \dots \), the existence of some type
+of \emph{variable names}, represented by the symbols
+\( l, m, \dots \), and the existence of some type of \emph{names},
+represented by the symbols \( x, y, \dots \). \ednote{AM: are we assuming that names can be compared for equality everywhere or only in C2?} %
+
+The syntax of processes will include:
+the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+The two components can proceed independently of each other, or they can interact via shared names.
+The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
+
+Processes will of course include \emph{input} and \emph{output}, whose
+signature will depend on the calculus being value or name-passing. We
+use here the metavariable $c,k$ to abstract over this choice.  The
+process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
+\( c \) via \( x \), then continue as \( P \).  % The intention is that
+% the value \( v \) must be a base value when it is actually sent, and
+% this will be enforced in the semantics later on
+  The process
+\( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
+via \( x \), then continue as \( P \) with the received element
+substituted \( k \).  The input operator thus
+binds \( k \) in \( P \).
+
+The process \( \Pres{x}{P} \) is the \emph{restriction} of the name
+\( x \) to \( P \), binding \( x \) in \( P \).
+
+The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
+It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
+Replication makes it possible to express infinite behaviors.
+
+We will use the notation \( \freenames{P} \) to denote the set of
+names that occur free,
+% (i.e.\ not bound by a restriction) in \( P \),
+\( \boundnames{P} \) to denote the set of names that occur bound
+% (by a restriction)
+in \( P \) and  \( \freevars{P} \)
+to denote the set of variable names that occur free
+% (i.e.\ not bound by an input)
+in \( P \).  We will use the notation \( \boundvars{P} \)
+for the set of variable names that occur bound
+% (by an input)
+in \( P \).  We will use the notation \( \subst{P}{v}{l} \) to denote
+the process \( P \) with value \( v \) substituted for variable name
+\( l \). Similarly, \( \subst{P}{x}{y} \) to denote the process
+\( P \) with name \( x \) substituted for name \( y \).
+
+Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
+written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \)
+by a finite number of substitutions of bound variable names.  As a
+convention, we will identify \( \alpha \)-convertible processes and we
+assume that bound names and bound variable names of any processes are
+chosen to be different from the names and variable names that occur
+free in any other entities under consideration, such as processes,
+substitutions, and sets of names or variable names.  This is justified
+because any overlapping names and variable names may be
+\( \alpha \)-converted such that the assumption is satisfied.
+
+A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+
+We will think of contexts as functions between processes.
+A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+
+We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+
+\subsection{Challenge: Linearity and behavioral type systems}
 \label{sec:challenge:linearity-beh-types}
@@
 % We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-We assume the existence of some type of \emph{base values}, represented by the symbols
- \( a, b, \dots \), the existence of some type
-of \emph{variable names}, represented by the symbols
-\( l, m, \dots \), and the existence of some type of \emph{names},
-represented by the symbols \( x, y, \dots \).  The syntax is
+% We assume the existence of some type of \emph{base values}, represented by the symbols
+%  \( a, b, \dots \), the existence of some type
+% of \emph{variable names}, represented by the symbols
+% \( l, m, \dots \), and the existence of some type of \emph{names},
+% represented by the symbols \( x, y, \dots \).
+
+The syntax is
 given by the grammar:
-\begin{align*}
-  v,w :=&&& a \\
-  |&&& l \\
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{v}{P} \\
-  |&&& \Pin{x}{l}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Presd{x}{y}{P}
-\end{align*}
+\begin{metanote}
+AM:  Should we save some space like this?
+\end{metanote}
+% \begin{align*}
+%   v,w :=&&& a \\
+%   |&&& l \\
+%   P,Q :=&&& \Pend \\
+%   |&&& \Pout{x}{v}{P} \\
+%   |&&& \Pin{x}{l}{P} \\
+%   |&&& \Ppar{P}{Q} \\
+%   |&&& \Presd{x}{y}{P}
+% \end{align*}
+\[
+\begin{array}{rcl}
+  v,w & ::= & a \mid l \\
+   P,Q & ::=& \Pend \mid \Pout{x}{v}{P} \mid \Pin{x}{l}{P} \mid \Ppar{P}{Q} \mid  \Presd{x}{y}{P}
+\end{array}
+\]
 A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{v}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
+% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
+The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \), then continue as \( P \).
 The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-The input operator thus binds the variable name \( l \) in \( P \).
-The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
+The input process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
+% The input operator thus binds the variable name \( l \) in \( P \).
+% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+% The two components can proceed independently of each other, or they can interact via shared names.
 The process \( \Presd{x}{y}{P} \) is the \emph{restriction} of the channel with endpoints named \( x \) and \( y \) to \( P \).
@@
 Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-Note that there is no recursion of replication in the syntax, and thus no infinite behaviours can be expressed.
+Note that there is no recursion of replication in the syntax, and thus no infinite behaviors can be expressed.
 
-We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
-We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
-We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
-We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
-We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
+% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
+% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
+% We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
+% We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
+% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
 
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-As a convention, we will identify \( \alpha \)-convertible processes.
+% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
+% As a convention, we will identify \( \alpha \)-convertible processes.
 
-As a convention, we assume that bound names and bound variable names of any processes are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, substitutions, and sets of names or variable names.
-This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
+% As a convention, we assume that bound names and bound variable names of any processes are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, substitutions, and sets of names or variable names.
+% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
 
@@
 \subsubsection{Contexts and congruences}
-A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
 
-We will think of contexts as functions between processes.
-A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+% We will think of contexts as functions between processes.
+% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
 
-We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
 
@@
 We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
-A process is then \emph{well-formed} if, for each of its structurally congruenct processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
+A process is then \emph{well-formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
 
@@
 \subsubsection{Typing rules}
-We have three typing judgements: one for values, one for names, and one for processes.
+We have three typing judgments: one for values, one for names, and one for processes.
 The typing rules for values are as follows:
@@
 This challenge formalises a proof that requires explicit scope extrusion.
-Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the recieving process.
+Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
 This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
@@
 \subsubsection{Syntax}
-We assume the existence of a type of \emph{names} that can be compared for equality, values of which we will denote by \( x, y, \dots \).
-The syntax is then:
+In this section 
+ we assume that \emph{names}  can be compared for equality. 
+The syntax is:
 \begin{align*}
@@
 \end{align*}
-The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
+%The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
 The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
 The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
-The input operator thus binds the name \( y \) in \( P \).
-The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-The restriction operator thus binds the name \( x \) in \( P \).
+%The input operator thus binds the name \( y \) in \( P \).
+% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+% The two components can proceed independently of each other, or they can interact via shared names.
+% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
+% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
+% The restriction operator thus binds the name \( x \) in \( P \).
 Note that the scope of a restriction may change when processes interact. Namely, a restricted name may be sent \emph{outside} of its scope.
-The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-Note that there is no recursion or replication in the syntax, and thus no infinite behaviours can be expressed.
+% The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
+Note that there is no recursion or replication in the syntax, and thus no infinite behaviors can be expressed.
 
-We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
-We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
-We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
+% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
+% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
+% We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
 
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
-As a convention, we will identify \( \alpha \)-convertible processes.
+% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
+% As a convention, we will identify \( \alpha \)-convertible processes.
 
-As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
-This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
+% As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
+% This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
 
@@
 
-We will again use the notation \( \freenames{\alpha} \) to denote the set of names that occur free in the action \( \alpha \) and the notation \( \boundnames{\alpha} \) to denote the set of names that occur bound in the action \( \alpha \).
-In the free output action \( \Aoutf{x}{y} \) and the input action \( \Ain{x}{y} \), both \( x \) and \( y \) are free names.
-In the bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while \( y \) is a bound name.
-We will further use the notation \( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \) and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur in the action \( \alpha \).
+We will extend the notion of free and bound occurrences with 
+\( \freenames{\alpha} \) to denote the set of names that occur free in
+the action \( \alpha \) and  \( \boundnames{\alpha} \) to
+denote the set of names that occur bound in the action \( \alpha \).
+In the free output action \( \Aoutf{x}{y} \) and the input action
+\( \Ain{x}{y} \), both \( x \) and \( y \) are free names.  In the
+bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while
+\( y \) is a bound name.  We will further use the notation
+\( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \)
+and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur
+in the action \( \alpha \).
 
@@
 
-As a convention, we assume that bound names of any processes or actions are chosen to be different from the names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names.
-The convention has one exception, namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \), the name \( z \) (which occurs bound in \( P \) and the action \( \Aoutb{x}{z} \)) may occur free in \( Q \).
-Without this exception it would be impossible to express scope extrusion.
+We keep the convention that bound names of any processes or actions
+are chosen to be different from the names that occur free in any other
+entities under consideration, such as processes, actions,
+substitutions, and sets of names.  The convention has one exception,
+namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \),
+the name \( z \) (which occurs bound in \( P \) and the action
+\( \Aoutb{x}{z} \)) may occur free in \( Q \).  Without this exception
+it would be impossible to express scope extrusion.
 
@@
 \end{gather}
-Note that, since our calculus does not have infinite behaviour, there is only one relation which satisfies these criteria.
+Note that, since our calculus does not have infinite behavior, there is only one relation which satisfies these criteria.
 Thus we do not specify that the relation is the largest of its kind, and the relation does not need to be coinductively defined.
@@
 This challenge formalises a proof that requires coinductive techniques.
-Coinduction is a proof technique for infinite structures, which arise in this context due to systems with behaviours that continue indefinitely.
+Coinduction is a proof technique for infinite structures, which arise in this context due to systems with behaviors that continue indefinitely.
 Coinduction is the dual of induction: whereas induction is useful for proving properties of least fixed points, coinduction is useful for proving properties of greatest fixed points.
@@
 The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
-Which, to reiterate, entails the use of coinduction due to the presence of infinite behaviour.
+Which, to reiterate, entails the use of coinduction due to the presence of infinite behavior.
 
 \subsubsection{Syntax}
-We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-The syntax is then:
+% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
+The syntax differs from \ref{sec:challenge:linearity-beh-types} for including replication and having unary restriction:
 \begin{align*}
@@
 \end{align*}
-A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
-The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-The input operator thus binds the variable name \( l \) in \( P \).
-The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-The restriction operator thus binds the name \( x \) in \( P \).
-Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-Replication makes it possible to express infinite behaviours.
+% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
+% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
+% The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
+% The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
+% The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
+% The input operator thus binds the variable name \( l \) in \( P \).
+% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+% The two components can proceed independently of each other, or they can interact via shared names.
+% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
+% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
+% The restriction operator thus binds the name \( x \) in \( P \).
+% Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
+% The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
+% It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
+% Replication makes it possible to express infinite behaviours.
 
-We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
+% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
 
-% We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-As a convention, we will identify \( \alpha \)-convertible processes.
+% % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
+% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
+% As a convention, we will identify \( \alpha \)-convertible processes.
 
-Also as a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
-This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
+% Also as a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
+% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
 
@@
 
-As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
+% As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
 
@@
 We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Note that, since our systems have potentially infinite behaviours, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
+Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
 
@@
 
-\subsubsection{Contexts and congruences}
-Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
+% \subsubsection{Contexts and congruences}
+% Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
 
-A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
 
-We can think of contexts as functions between processes.
-A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+% We can think of contexts as functions between processes.
+% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
 
-We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
 

### 2023-09-19 challenges.tex
Change minor wordings in preliminaries.

@@
 
-Processes will of course include \emph{input} and \emph{output}, whose
-signature will depend on the calculus being value or name-passing. We
-use here the metavariable $c,k$ to abstract over this choice.  The
-process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
+For communication, processes include \emph{input} and \emph{output}, whose
+signature will depend on the calculus being value or name-passing.
+We use here the metavariables $c,k$ to abstract over this choice.
+The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
 \( c \) via \( x \), then continue as \( P \).  % The intention is that
@@
 % this will be enforced in the semantics later on
-  The process
-\( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
+The process \( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
 via \( x \), then continue as \( P \) with the received element
-substituted \( k \).  The input operator thus
+substituted for \( k \).  The input operator thus
 binds \( k \) in \( P \).
@@
 the process \( P \) with value \( v \) substituted for variable name
-\( l \). Similarly, \( \subst{P}{x}{y} \) to denote the process
+\( l \). Similarly, \( \subst{P}{x}{y} \) denotes the process
 \( P \) with name \( x \) substituted for name \( y \).

### 2023-09-20 challenges.tex
Change scope extrusion challenge to involve reductions instead of bisimilarity.

@@
 The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
+We will define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.
+The objective of this challenge is to prove that the two semantics are equivalent up to structural congruence.
 
 \subsubsection{Syntax}
-In this section 
- we assume that \emph{names}  can be compared for equality. 
-The syntax is:
+In this section we assume that \emph{names}  can be compared for equality.
+The syntax in this challenge will be split into two levels: processes and summations.
+The syntax of processes is given by:
 \begin{align*}
-  P,Q :=&&& \Pend \\
+  P,Q :=&&& \Ppar{P}{Q} \\
+  |&&& \Pres{x}{P} \\
+  |&&& M
+\end{align*}
+The syntax of summations is given by:
+\begin{align*}
+  M,N :=&&& \Pend \\
   |&&& \Pout{x}{y}{P} \\
   |&&& \Pin{x}{y}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Pchoice{P}{Q}
+  |&&& \Pchoice{M}{N}
 \end{align*}
+Splitting the syntax into these two layers has the effect of forcing choices to be guarded by an input or an output, which simplifies the theory.
+Allowing choices between arbitrary processes is rarely useful.
+
 %The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
@@
 
-\subsubsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
+Due to the syntactic levels, we will need to modify the definition of a context slightly by requiring that the \emph{hole} symbol \( \ctxhole \) does not appear directly as the left or right term in a choice.
+For instance, we disallow the context \( \Pchoice{\ctxhole}{\Pin{x}{y}{\Pend}} \), but allow the context \( \Pchoice{\Pout{x}{z}{\ctxhole}}{\Pin{x}{y}{\Pend}} \).
+
+\subsubsection{Reduction semantics}
+The first semantics is an operational reduction semantics, which avoids reasoning explicitly about scope extrusion by way of a structural congruence rule.
+\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
+\begin{mathpar}
+  \inferrule[Sc-Sum-Assoc]{ }{\scong{\Pchoice{M}{(\Pchoice{N}{L})}}{\Pchoice{(\Pchoice{M}{N})}{L}}}
+  \and
+  \inferrule[Sc-Sum-Comm]{ }{\scong{\Pchoice{M}{N}}{\Pchoice{N}{M}}}
+  \and
+  \inferrule[Sc-Sum-Inact]{ }{\scong{\Pchoice{M}{\Pend}}{M}}
+  \and
+  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{P}{(\Ppar{Q}{R})}}{\Ppar{(\Ppar{P}{Q})}{R}}}
+  \and
+  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
+  \and
+  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
+  \and
+  \inferrule[Sc-Res-Par]{x \notin \freenames{Q}}{\scong{\Ppar{\Pres{x}{P}}{Q}}{\Pres{x}{(\Ppar{P}{Q})}}}
+  \and
+  \inferrule[Sc-Res-Inact]{ }{\scong{\Pres{x}{\Pend}}{\Pend}}
+  \and
+  \inferrule[Sc-Res]{ }{\scong{\Pres{x}{\Pres{y}{P}}}{\Pres{y}{\Pres{x}{P}}}}
+\end{mathpar}
+
+The operational semantics is then defined as the following binary relation on processes:
+\begin{mathpar}
+  \inferrule[R-Com]{ }{\reduces{\Ppar{(\Pchoice{\Pout{x}{y}{P}}{M})}{(\Pchoice{\Pin{x}{z}{Q}}{N})}}{\Ppar{P}{\subst{Q}{y}{z}}}}
+  \and
+  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Pres{x}{P}}{\Pres{x}{Q}}}
+  \and
+  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
+  \and
+  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
+\end{mathpar}
+Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{R-Com}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
+Finally, note that rule \TirName{R-Struct} allows for applying the structural congruence both before and after the reduction.
+This ensures that the reduction relation is closed under structural congruence.
+
+\subsubsection{Transition system semantics}
+The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
 The transitions are labelled by \emph{actions}, the syntax of which is as follows:
@@
 
-\subsubsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
+\subsubsection{Challenge}
+\begin{lemma}\label{se-lemma-harmony-fact}
+  If \( \scong{P}{Q} \) and \( \transition{P}{\alpha}{P'} \), then for some \( Q' \) we have \( \transition{Q}{\alpha}{Q'} \) and \( \scong{P'}{Q'} \).
+\end{lemma}
+\begin{proof}
+  Sketch. First, show the special case when \( P \) can be rewritten to \( Q \) with a single application of an axiom of the structural congruence to some subterm of P.
+  The proof is then by induction on the number of such steps.
+\end{proof}
 
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
-\begin{align}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
+\begin{lemma}\label{se-lemma-normalized-reduction}
+  If \( \reduces{P}{Q} \) then there are \( x, y, z, z_1, \dots, z_n, R_1, N_1, R_2, N_2, \) and \( S\) such that
+  \begin{align*}
+    &\scong{P}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{(\Pchoice{\Pout{x}{y}{R_1}}{N_1})}{(\Pchoice{\Pin{x}{z}{R_2}}{N_2})})}{S})}}} \\
+    &\scong{Q}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{R_1}{\subst{R_2}{y}{z}})}{S})}}}
+  \end{align*}
+\end{lemma}
+To prove this lemma, we introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
+The first rule applied is \TirName{R-Com}. The derivation continues with an application of \TirName{R-Par}, followed by zero or more applications of \TirName{R-Res}.
+The last rule is an application of \TirName{R-Struct}.
+\begin{lemma}\label{se-lemma-normalized-derivation}
+  Every reduction has a normalized derivation.
+\end{lemma}
+\begin{proof}
+  Sketch.
+To obtain a normalized derivation from an arbitrary derivation we will need to check that rules \TirName{R-Com}, \TirName{R-Par} and \TirName{R-Res} commute with \TirName{R-Struct}, and that two applications of \TirName{R-Struct} can be combined into one.
+\end{proof}
+\begin{proof}{Of lemma \autoref{se-lemma-normalized-reduction}.}
+  Follows immediately from lemma \autoref{se-lemma-normalized-derivation} and the shape of a normalized derivation.
+\end{proof}
 
-Two processes \( P \) and \( Q \) are then \emph{strong barbed bisimilar}, written \( \sbbisim{P}{Q} \), if:
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{Q'}~\textrm{for some \( Q' \) with}~\sbbisim{P'}{Q'} \\
-  \observable{Q}{\mu}~\textrm{implies}~\observable{P}{\mu} \\
-  \transition{Q}{\Atau}{Q'}~\textrm{implies}~\transition{P}{\Atau}{P'}~\textrm{for some \( P' \) with}~\sbbisim{P'}{Q'}
-\end{gather}
-Note that, since our calculus does not have infinite behavior, there is only one relation which satisfies these criteria.
-Thus we do not specify that the relation is the largest of its kind, and the relation does not need to be coinductively defined.
-To actually prove properties about the relation, however, we will still need either an inductive (or a coinductive) proof principle for the relation.
-
-\subsubsection{Challenge}
 The objective of this challenge is to prove the following theorem:
 \begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
+  \( \reduces{P}{Q} \) if and only if \( \transition{P}{\Atau}{\scong{}{Q}} \).
 \end{theorem}
+\begin{proof}
+  We prove the two parts of the theorem separately. First, if \( \reduces{P}{Q} \), then by lemma \autoref{se-lemma-normalized-reduction}, \( \scong{P}{P'} \) with
+  \begin{equation*}
+    P' = \Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{(\Pchoice{\Pout{x}{y}{R_1}}{N_1})}{(\Pchoice{\Pin{x}{z}{R_2}}{N_2})})}{S})}}
+  \end{equation*}
+  and \( \scong{Q}{Q'} \) with
+  \begin{equation*}
+    Q' = \Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{R_1}{\subst{R_2}{y}{z}})}{S})}} \ .
+  \end{equation*}
+  We can easily check that \( \transition{P'}{\tau}{Q'} \) and so by lemma \autoref{se-lemma-harmony-fact}, \( \transition{P}{\tau}{\scong{}{\scong{Q'}{Q}}} \).
+
+  For the other direction, we first show that \( \transition{P}{\Atau}{Q} \) implies \( \reduces{P}{Q} \) from which the result follows.
+  The proof is by induction on the inference of \( \transition{P}{\Atau}{Q} \) using the following:
+  \begin{enumerate}
+  \item if \( \transition{Q}{\Aoutf{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{y}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x,y \notin \{ w_1, \dots, w_n \} \).
+  \item if \( \transition{Q}{\Aoutb{x}{z}}{Q'} \) then \( \scong{Q}{\Pres{z}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{z}{R}}{M})}{S})}}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x \notin \{ z, w_1, \dots, w_n \} \).
+  \item if \( \transition{Q}{\Ain{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pin{x}{z}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\subst{R}{y}{z}}{S})}}} \) where \( x \notin \{ w_1, \dots, w_n \} \).
+  \end{enumerate}
+\end{proof}
 

### 2023-09-22 challenges.tex
just playing around with \mid for P | Q
Split new theorem into two.

@@
 
-The objective of this challenge is to prove the following theorem:
-\begin{theorem}
-  \( \reduces{P}{Q} \) if and only if \( \transition{P}{\Atau}{\scong{}{Q}} \).
+The objective of this challenge is to prove the following theorems, which together show the equivalence between the reduction semantics and the transition system semantics up to structural congruence.
+The first of the theorems involves reasoning about scope extrusion more directly than the other, and if time does not permit proving both of the theorems, \autoref{thm:se-trans-implies-red} should be proven first.
+\begin{theorem}\label{thm:se-trans-implies-red}
+  \( \transition{P}{\Atau}{Q} \) implies \( \reduces{P}{Q} \).
 \end{theorem}
 \begin{proof}
-  We prove the two parts of the theorem separately. First, if \( \reduces{P}{Q} \), then by lemma \autoref{se-lemma-normalized-reduction}, \( \scong{P}{P'} \) with
+  The proof is by induction on the inference of \( \transition{P}{\Atau}{Q} \) using the following:
+  \begin{enumerate}
+  \item if \( \transition{Q}{\Aoutf{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{y}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x,y \notin \{ w_1, \dots, w_n \} \).
+  \item if \( \transition{Q}{\Aoutb{x}{z}}{Q'} \) then \( \scong{Q}{\Pres{z}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{z}{R}}{M})}{S})}}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x \notin \{ z, w_1, \dots, w_n \} \).
+  \item if \( \transition{Q}{\Ain{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pin{x}{z}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\subst{R}{y}{z}}{S})}}} \) where \( x \notin \{ w_1, \dots, w_n \} \).
+  \end{enumerate}
+\end{proof}
+
+\begin{theorem}
+  \( \reduces{P}{Q} \) implies the existence of a \( Q' \) such that \( \transition{P}{\Atau}{Q'} \) and \( \scong{Q}{Q'} \).
+\end{theorem}
+\begin{proof}
+  If \( \reduces{P}{Q} \), then by lemma \autoref{se-lemma-normalized-reduction}, \( \scong{P}{P'} \) with
   \begin{equation*}
@@
   \end{equation*}
-  We can easily check that \( \transition{P'}{\tau}{Q'} \) and so by lemma \autoref{se-lemma-harmony-fact}, \( \transition{P}{\tau}{\scong{}{\scong{Q'}{Q}}} \).
-
-  For the other direction, we first show that \( \transition{P}{\Atau}{Q} \) implies \( \reduces{P}{Q} \) from which the result follows.
-  The proof is by induction on the inference of \( \transition{P}{\Atau}{Q} \) using the following:
-  \begin{enumerate}
-  \item if \( \transition{Q}{\Aoutf{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{y}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x,y \notin \{ w_1, \dots, w_n \} \).
-  \item if \( \transition{Q}{\Aoutb{x}{z}}{Q'} \) then \( \scong{Q}{\Pres{z}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{z}{R}}{M})}{S})}}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x \notin \{ z, w_1, \dots, w_n \} \).
-  \item if \( \transition{Q}{\Ain{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pin{x}{z}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\subst{R}{y}{z}}{S})}}} \) where \( x \notin \{ w_1, \dots, w_n \} \).
-  \end{enumerate}
+  We can easily check that \( \transition{P'}{\tau}{Q'} \) and so by lemma \autoref{se-lemma-harmony-fact}, \( \transition{P}{\tau}{Q'} \).
 \end{proof}

### 2023-09-26 challenges.tex
Fix explanation of two-level syntax.

@@
 \end{align*}
-Splitting the syntax into these two layers has the effect of forcing choices to be guarded by an input or an output, which simplifies the theory.
-Allowing choices between arbitrary processes is rarely useful.
+Splitting the syntax into these two layers has the effect of disallowing choices directly between parallel processes.
+Allowing choices between arbitrary processes complicates the theory and is rarely useful.
 

### 2023-09-26 challenges.tex
Fix free name assumption in rule Sc-Res-Par.

--- 2023-09-20 challenges.tex
@@
          %   \end{align*}
+% \newcommand{\bmid}{\mathbf{\Vert}}
 \[
@@
   v,w & ::= & a \mid l \\
-   P,Q & ::=& \Pend \mid \Pout{x}{v}{P} \mid \Pin{x}{l}{P} \mid \Ppar{P}{Q} \mid  \Presd{x}{y}{P}
+   P,Q & ::=& \Pend \mid \Pout{x}{v}{P} \mid \Pin{x}{l}{P} \mid \PBpar{P}{Q} \mid  \Presd{x}{y}{P}
 \end{array}
@@
   \and
-  \inferrule[Sc-Res-Par]{\{x,y\} \not\subseteq \freenames{Q}}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
+  \inferrule[Sc-Res-Par]{\{x,y\} \cap \freenames{Q} = \emptyset}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
   \and
@@
 The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
-We will define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.
-The objective of this challenge is to prove that the two semantics are equivalent up to structural congruence.
+The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
 
 \subsubsection{Syntax}
-In this section we assume that \emph{names}  can be compared for equality.
-The syntax in this challenge will be split into two levels: processes and summations.
-The syntax of processes is given by:
+In this section 
+ we assume that \emph{names}  can be compared for equality. 
+The syntax is:
 \begin{align*}
-  P,Q :=&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& M
-\end{align*}
-The syntax of summations is given by:
-\begin{align*}
-  M,N :=&&& \Pend \\
+  P,Q :=&&& \Pend \\
   |&&& \Pout{x}{y}{P} \\
   |&&& \Pin{x}{y}{P} \\
-  |&&& \Pchoice{M}{N}
+  |&&& \Ppar{P}{Q} \\
+  |&&& \Pres{x}{P} \\
+  |&&& \Pchoice{P}{Q}
 \end{align*}
-Splitting the syntax into these two layers has the effect of forcing choices to be guarded by an input or an output, which simplifies the theory.
-Allowing choices between arbitrary processes is rarely useful.
-
 %The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
@@
 
-Due to the syntactic levels, we will need to modify the definition of a context slightly by requiring that the \emph{hole} symbol \( \ctxhole \) does not appear directly as the left or right term in a choice.
-For instance, we disallow the context \( \Pchoice{\ctxhole}{\Pin{x}{y}{\Pend}} \), but allow the context \( \Pchoice{\Pout{x}{z}{\ctxhole}}{\Pin{x}{y}{\Pend}} \).
-
-\subsubsection{Reduction semantics}
-The first semantics is an operational reduction semantics, which avoids reasoning explicitly about scope extrusion by way of a structural congruence rule.
-\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
-\begin{mathpar}
-  \inferrule[Sc-Sum-Assoc]{ }{\scong{\Pchoice{M}{(\Pchoice{N}{L})}}{\Pchoice{(\Pchoice{M}{N})}{L}}}
-  \and
-  \inferrule[Sc-Sum-Comm]{ }{\scong{\Pchoice{M}{N}}{\Pchoice{N}{M}}}
-  \and
-  \inferrule[Sc-Sum-Inact]{ }{\scong{\Pchoice{M}{\Pend}}{M}}
-  \and
-  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{P}{(\Ppar{Q}{R})}}{\Ppar{(\Ppar{P}{Q})}{R}}}
-  \and
-  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
-  \and
-  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
-  \and
-  \inferrule[Sc-Res-Par]{x \notin \freenames{Q}}{\scong{\Ppar{\Pres{x}{P}}{Q}}{\Pres{x}{(\Ppar{P}{Q})}}}
-  \and
-  \inferrule[Sc-Res-Inact]{ }{\scong{\Pres{x}{\Pend}}{\Pend}}
-  \and
-  \inferrule[Sc-Res]{ }{\scong{\Pres{x}{\Pres{y}{P}}}{\Pres{y}{\Pres{x}{P}}}}
-\end{mathpar}
-
-The operational semantics is then defined as the following binary relation on processes:
-\begin{mathpar}
-  \inferrule[R-Com]{ }{\reduces{\Ppar{(\Pchoice{\Pout{x}{y}{P}}{M})}{(\Pchoice{\Pin{x}{z}{Q}}{N})}}{\Ppar{P}{\subst{Q}{y}{z}}}}
-  \and
-  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Pres{x}{P}}{\Pres{x}{Q}}}
-  \and
-  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
-  \and
-  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathpar}
-Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{R-Com}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
-Finally, note that rule \TirName{R-Struct} allows for applying the structural congruence both before and after the reduction.
-This ensures that the reduction relation is closed under structural congruence.
-
-\subsubsection{Transition system semantics}
-The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
+\subsubsection{Semantics}
+The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
 The transitions are labelled by \emph{actions}, the syntax of which is as follows:
@@
 
+\subsubsection{Bisimilarity}
+Our notion of process equivalence relations builds on a notion of \emph{observables}.
+If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
+We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
+
+To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
+\begin{align}
+  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
+  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
+\end{align}
+
+Two processes \( P \) and \( Q \) are then \emph{strong barbed bisimilar}, written \( \sbbisim{P}{Q} \), if:
+\begin{gather}
+  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
+  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{Q'}~\textrm{for some \( Q' \) with}~\sbbisim{P'}{Q'} \\
+  \observable{Q}{\mu}~\textrm{implies}~\observable{P}{\mu} \\
+  \transition{Q}{\Atau}{Q'}~\textrm{implies}~\transition{P}{\Atau}{P'}~\textrm{for some \( P' \) with}~\sbbisim{P'}{Q'}
+\end{gather}
+Note that, since our calculus does not have infinite behavior, there is only one relation which satisfies these criteria.
+Thus we do not specify that the relation is the largest of its kind, and the relation does not need to be coinductively defined.
+To actually prove properties about the relation, however, we will still need either an inductive (or a coinductive) proof principle for the relation.
+
 \subsubsection{Challenge}
-\begin{lemma}\label{se-lemma-harmony-fact}
-  If \( \scong{P}{Q} \) and \( \transition{P}{\alpha}{P'} \), then for some \( Q' \) we have \( \transition{Q}{\alpha}{Q'} \) and \( \scong{P'}{Q'} \).
-\end{lemma}
-\begin{proof}
-  Sketch. First, show the special case when \( P \) can be rewritten to \( Q \) with a single application of an axiom of the structural congruence to some subterm of P.
-  The proof is then by induction on the number of such steps.
-\end{proof}
-
-\begin{lemma}\label{se-lemma-normalized-reduction}
-  If \( \reduces{P}{Q} \) then there are \( x, y, z, z_1, \dots, z_n, R_1, N_1, R_2, N_2, \) and \( S\) such that
-  \begin{align*}
-    &\scong{P}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{(\Pchoice{\Pout{x}{y}{R_1}}{N_1})}{(\Pchoice{\Pin{x}{z}{R_2}}{N_2})})}{S})}}} \\
-    &\scong{Q}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{R_1}{\subst{R_2}{y}{z}})}{S})}}}
-  \end{align*}
-\end{lemma}
-To prove this lemma, we introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
-The first rule applied is \TirName{R-Com}. The derivation continues with an application of \TirName{R-Par}, followed by zero or more applications of \TirName{R-Res}.
-The last rule is an application of \TirName{R-Struct}.
-\begin{lemma}\label{se-lemma-normalized-derivation}
-  Every reduction has a normalized derivation.
-\end{lemma}
-\begin{proof}
-  Sketch.
-To obtain a normalized derivation from an arbitrary derivation we will need to check that rules \TirName{R-Com}, \TirName{R-Par} and \TirName{R-Res} commute with \TirName{R-Struct}, and that two applications of \TirName{R-Struct} can be combined into one.
-\end{proof}
-\begin{proof}{Of lemma \autoref{se-lemma-normalized-reduction}.}
-  Follows immediately from lemma \autoref{se-lemma-normalized-derivation} and the shape of a normalized derivation.
-\end{proof}
-
 The objective of this challenge is to prove the following theorem:
 \begin{theorem}
-  \( \reduces{P}{Q} \) if and only if \( \transition{P}{\Atau}{\scong{}{Q}} \).
+  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
 \end{theorem}
-\begin{proof}
-  We prove the two parts of the theorem separately. First, if \( \reduces{P}{Q} \), then by lemma \autoref{se-lemma-normalized-reduction}, \( \scong{P}{P'} \) with
-  \begin{equation*}
-    P' = \Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{(\Pchoice{\Pout{x}{y}{R_1}}{N_1})}{(\Pchoice{\Pin{x}{z}{R_2}}{N_2})})}{S})}}
-  \end{equation*}
-  and \( \scong{Q}{Q'} \) with
-  \begin{equation*}
-    Q' = \Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{R_1}{\subst{R_2}{y}{z}})}{S})}} \ .
-  \end{equation*}
-  We can easily check that \( \transition{P'}{\tau}{Q'} \) and so by lemma \autoref{se-lemma-harmony-fact}, \( \transition{P}{\tau}{\scong{}{\scong{Q'}{Q}}} \).
-
-  For the other direction, we first show that \( \transition{P}{\Atau}{Q} \) implies \( \reduces{P}{Q} \) from which the result follows.
-  The proof is by induction on the inference of \( \transition{P}{\Atau}{Q} \) using the following:
-  \begin{enumerate}
-  \item if \( \transition{Q}{\Aoutf{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{y}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x,y \notin \{ w_1, \dots, w_n \} \).
-  \item if \( \transition{Q}{\Aoutb{x}{z}}{Q'} \) then \( \scong{Q}{\Pres{z}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{z}{R}}{M})}{S})}}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x \notin \{ z, w_1, \dots, w_n \} \).
-  \item if \( \transition{Q}{\Ain{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pin{x}{z}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\subst{R}{y}{z}}{S})}}} \) where \( x \notin \{ w_1, \dots, w_n \} \).
-  \end{enumerate}
-\end{proof}
 

### 2023-09-27 challenges/challenges.tex
Add a few more examples of potential topics.
Split challenges.tex into multiple files.

--- 2023-03-02 challenges/challenges.tex
@@
-\documentclass[a4paper]{article}
+\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
+%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
+\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
+                         {\endgroup]\end{quote}}
+\long\def\ignore#1{}
 
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage[utf8]{inputenc}
-\usepackage{mathpartir}
-\usepackage{bm}
-\usepackage{graphicx}
-
-\newtheorem{lemma}{Lemma}
-\newtheorem{theorem}{Theorem}
-
-\newcommand{\Pend}{\bm{0}}
-\newcommand{\Ppar}[2]{#1 \mid #2}
-\newcommand{\Pres}[2]{(\bm{\nu} #1)~#2}
-\newcommand{\Pout}[3]{#1 ! #2 . #3}
-\newcommand{\Pin}[3]{#1 ? (#2) . #3}
-\newcommand{\Pchoice}[2]{#1 + #2}
-\newcommand{\Preplicate}[1]{{!}#1}
-
-\newcommand{\freenames}[1]{\textrm{fn}(#1)}
-\newcommand{\boundnames}[1]{\textrm{bn}(#1)}
-\newcommand{\names}[1]{\textrm{n}(#1)}
-
-\newcommand{\freevars}[1]{\textrm{fv}(#1)}
-\newcommand{\boundvars}[1]{\textrm{bv}(#1)}
-\newcommand{\vars}[1]{\textrm{vars}(#1)}
-
-\newcommand{\alphacon}[2]{#1 =_{\alpha} #2}
-\newcommand{\subst}[3]{#1\{#2/#3\}}
-
-\newcommand{\Aoutf}[2]{#1 ! #2}
-\newcommand{\Aoutb}[2]{#1 ! (#2)}
-\newcommand{\Ain}[2]{#1 ? #2}
-\newcommand{\Atau}{\tau}
-
-\newcommand{\transition}[3]{#1 \xrightarrow{#2} #3}
-
-\newcommand{\observable}[2]{#1\downarrow_{#2}}
-\newcommand{\obsin}[1]{#1?}
-\newcommand{\obsout}[1]{#1!}
-
-\newcommand\sbullet[1][.5]{\mathbin{\hbox{\scalebox{#1}{$\bullet$}}}}
-\newcommand{\sbbisim}[2]{#1 \overset{\sbullet}{\sim} #2}
-\newcommand{\nsbbisim}[2]{#1 \not\overset{\sbullet}{\sim} #2}
-
-\newcommand{\ctxhole}{[\cdot]}
-\newcommand{\applyctx}[2]{#1[#2]}
-\newcommand{\sbcong}[2]{#1 \simeq^c #2}
-
-\newcommand{\applysubst}[2]{#2#1}
-
-\begin{document}
-
-\section{Challenge: Scope Extrusion}
-The idea of this challenge is to formalize a proof that requires scope extrusion.
-Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the recieving process.
-This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
-
-The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that barbed bisimulation for this calculus is an equivalence relation.
-
-\subsection{Syntax}
-We assume the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-The syntax is then:
-\begin{align*}
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{y}{P} \\
-  |&&& \Pin{x}{y}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Pchoice{P}{Q}
-\end{align*}
-The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
-The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
-The input operator thus binds the name \( y \) in \( P \).
-The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-The restriction operator thus binds the name \( x \) in \( P \).
-Note that the scope of a restriction may change when processes interact.
-The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-Note that there is no recursion or replication in the syntax, and thus no infinite behaviours can be expressed.
-
-We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
-We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
-We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
-
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
-As a convention, we will identify \( \alpha \)-convertible processes.
-
-As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
-This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which are as follows:
-\begin{align*}
-  \alpha := &&& \Aoutf{x}{y} \\
-  |&&& \Ain{x}{y} \\
-  |&&& \Aoutb{x}{y} \\
-  |&&& \Atau
-\end{align*}
-The \emph{free output action} \( \Aoutf{x}{y} \) is sending the name \( y \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the name \( y \) via \( x \).
-The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
-
-We will again use the notation \( \freenames{\alpha} \) to denote the set of names that occur free in the action \( \alpha \) and the notation \( \boundnames{\alpha} \) to denote the set of names that occur bound in the action \( \alpha \).
-In the free output action \( \Aoutf{x}{y} \) and the input action \( \Ain{x}{y} \), both \( x \) and \( y \) are free names.
-In the bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while \( y \) is a bound name.
-We will further use the notation \( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \) and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur in the action \( \alpha \).
-
-The transition relation is then defined by the following rules:
-\begin{mathpar}
-  \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{z}{P}}{\Ain{x}{y}}{\subst{P}{y}{z}}}
-  \and
-  \inferrule[Sum-L]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{P}{Q}}{\alpha}{P'}}
-  \and
-  \inferrule[Sum-R]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{Q}{P}}{\alpha}{P'}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'} \\ \boundnames{\alpha} \cap \freenames{Q} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'} \\ \boundnames{\alpha} \cap \freenames{P} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{y}}{P'} \\ \transition{Q}{\Ain{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{y}}{P'} \\ \transition{Q}{\Aoutf{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Close-L]{\transition{P}{\Aoutb{x}{z}}{P'} \\ \transition{Q}{\Ain{x}{z}}{Q'} \\ z \notin \freenames{Q}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Open]{\transition{P}{\Aoutf{x}{z}}{P'} \\ z \neq x}{\transition{\Pres{z}{P}}{\Aoutb{x}{z}}{P'}}
-  \and
-  \inferrule[Close-R]{\transition{P}{\Ain{x}{z}}{P'} \\ \transition{Q}{\Aoutb{x}{z}}{Q'} \\ z \notin \freenames{P}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
-\end{mathpar}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
-
-As a convention, we assume that bound names of any processes or actions are chosen to be different from the names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names.
-The convention has one exception, namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \), the name \( z \) (which occurs bound in \( P \) and the action \( \Aoutb{x}{z} \)) may occur free in \( Q \).
-Without this exception it would be impossible to express scope extrusion.
-
-\subsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
-
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
-\begin{align}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
-
-Two processes \( P \) and \( Q \) are then \emph{strongly barbed bisimilar}, written \( \sbbisim{P}{Q} \), if:
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{Q'}~\textrm{for some \( Q' \) with}~\sbbisim{P'}{Q'} \\
-  \observable{Q}{\mu}~\textrm{implies}~\observable{P}{\mu} \\
-  \transition{Q}{\Atau}{Q'}~\textrm{implies}~\transition{P}{\Atau}{P'}~\textrm{for some \( P' \) with}~\sbbisim{P'}{Q'}
-\end{gather}
-Note that, since our calculus does not have infinite behaviour, we do not specify that the relation is the largest of its kind, and the relation is thus not coinductively defined.
-
-\subsection{Challenge}
-The objective of this challenge is to prove the following theorem:
-\begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-
-\section{Challenge: Coinduction}
-The idea of this challenge is to formalize a proof that requires coinductive techniques.
-Coinduction is a proof technique for infinite structures, which arise in this context due to systems with behaviours that continue indefinitely.
-Coinduction is the dual of induction: whereas induction is useful for proving properties of least fixed points, coinduction is useful for proving properties of greatest fixed points.
-
-The setting for this challenge is an untyped calculus of communicating systems with replication of processes, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
-
-\subsection{Syntax}
-We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-The syntax is then:
-\begin{align*}
-  v,w :=&&& a \\
-  |&&& l \\
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{v}{P} \\
-  |&&& \Pin{x}{l}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Preplicate{P}
-\end{align*}
-A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
-The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-The input operator thus binds the variable name \( l \) in \( P \).
-The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-The restriction operator thus binds the name \( x \) in \( P \).
-Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-Replication makes it possible to express infinite behaviours.
-
-We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-As a convention, we will identify \( \alpha \)-convertible processes.
-
-As a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
-This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which are as follows:
-\begin{align*}
-  \alpha := &&& \Aoutf{x}{a} \\
-  |&&& \Ain{x}{a} \\
-  |&&& \Atau
-\end{align*}
-The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value \( a \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the base value \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
-
-We will use the notation \( \names{\alpha} \) to denote the set of names that occur in the action \( \alpha \).
-
-The transition relation is then defined by the following rules:
-\begin{mathpar}
-  \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{l}{P}}{\Ain{x}{a}}{\subst{P}{a}{l}}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{a}}{P'} \\ \transition{Q}{\Ain{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{a}}{P'} \\ \transition{Q}{\Aoutf{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ x \notin \names{\alpha}}{\transition{\Pres{x}{P}}{\alpha}{\Pres{x}{P'}}}
-  \and
-  \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
-\end{mathpar}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable name.
-
-As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
-
-\subsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
-
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
-\begin{align}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
-
-\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is then the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}
-\end{gather}
-We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Note that, since our systems have potentially infinite behaviours, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
-
-\begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-
-Unfortunately, strong barbed bisimilarity is not a good notion of process equivalence, since it does not consider the environment of processes.
-For instance, \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since \( \obsout{x} \) is the only observable in both processes and they cannot perform a \( \Atau \)-action, but \( \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since the left process can perform a \( \Atau \)-action such that \( \obsout{y} \) becomes observable, whereas the right process cannot.
-
-\subsection{Contexts and congruences}
-Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
-
-A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-We will think of contexts as functions between processes.
-A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsection{Strong barbed congruence}
-A congruence is exactly what we need to make strong barbed bisimilarity consider the environment (i.e.\ the context) of processes.
-
-We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
-
-\begin{lemma}
-  \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \).
-\end{lemma}
-
-\subsection{Challenge}
-The objective of this challenge is to prove the following theorem:
-\begin{theorem}
-  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).
-\end{theorem}
-
-\paragraph{Notes about this challenge}
-Strong barbed bisimilarity is interesting by itself for this challenge, but the proof that it is an equivalence relation does not require up-to techniques, which are wide spread and interesting.
-The definition of strong barbed congruence requires introducing contexts, but this is actually nice, since they are so prevalent in the field and we want to cover as many basics as possible.
-The theorem about the strong barbed congruence requires up-to techniques, but also seems to involve introducing a structural equivalence, which is possibly not desirable.
-On the other hand, it is a nice application of structural congruence which does not ``ruin'' the semantics.
-
-\section{Challenge: Linearity}
-The setting for this challenge is a restricted \( \pi \)-calculus with a session type system.
-
-The objective of this challenge is to prove type preservation.
-
-\section{Bonus challenges}
-
-\subsection{Linearity and Scope Extrusion}
-The setting for this challenge is the \( \pi \)-calculus with a session type system.
-
-The objective of this challenge is to prove type preservation.
-
-\subsection{Scope Extrusion and Coinduction}
-The setting for this challenge is a simply typed \( \pi \)-calculus.
-
-The objective of this challenge is to prove bisimulation.
-
-\subsection{Coinduction and Linearity}
-The setting for this challenge is the Calculus of Communicating Systems with a session type system.
-
-The objective of this challenge is to prove type preservation.
-
-\section{Future work}
-\begin{itemize}
-\item Multiparty session types
-\item Choreography
-\item Encodings
-\item Code extraction
-\end{itemize}
-
-\end{document}
+\input{../challenges/preliminaries.tex}
+\input{../challenges/linearity.tex}
+\input{../challenges/scope-extrusion.tex}
+\input{../challenges/coinduction.tex}

### 2023-09-27 challenges/coinduction.tex
Split challenges.tex into multiple files.

--- 2023-09-26 challenges.tex
@@
-\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
-%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
-\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
-                         {\endgroup]\end{quote}}
-\long\def\ignore#1{}
-
-\subsection{Preliminaries}
-\label{sec:prelim}
-
-We list here some common notions that we will use in the following. Since the calculi under study are somewhat different, we will note in each section which will apply.
-
-We assume the existence of some type of \emph{base values}, represented by the symbols
-\( a, b, \dots \), the existence of some type
-of \emph{variable names}, represented by the symbols
-\( l, m, \dots \), and the existence of some type of \emph{names},
-represented by the symbols \( x, y, \dots \). \ednote{AM: are we assuming that names can be compared for equality everywhere or only in C2?} %
-
-The syntax of processes will include:
-the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-
-For communication, processes include \emph{input} and \emph{output}, whose
-signature will depend on the calculus being value or name-passing.
-We use here the metavariables $c,k$ to abstract over this choice.
-The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
-\( c \) via \( x \), then continue as \( P \).  % The intention is that
-% the value \( v \) must be a base value when it is actually sent, and
-% this will be enforced in the semantics later on
-The process \( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
-via \( x \), then continue as \( P \) with the received element
-substituted for \( k \).  The input operator thus
-binds \( k \) in \( P \).
-
-The process \( \Pres{x}{P} \) is the \emph{restriction} of the name
-\( x \) to \( P \), binding \( x \) in \( P \).
-
-The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-Replication makes it possible to express infinite behaviors.
-
-We will use the notation \( \freenames{P} \) to denote the set of
-names that occur free,
-% (i.e.\ not bound by a restriction) in \( P \),
-\( \boundnames{P} \) to denote the set of names that occur bound
-% (by a restriction)
-in \( P \) and  \( \freevars{P} \)
-to denote the set of variable names that occur free
-% (i.e.\ not bound by an input)
-in \( P \).  We will use the notation \( \boundvars{P} \)
-for the set of variable names that occur bound
-% (by an input)
-in \( P \).  We will use the notation \( \subst{P}{v}{l} \) to denote
-the process \( P \) with value \( v \) substituted for variable name
-\( l \). Similarly, \( \subst{P}{x}{y} \) denotes the process
-\( P \) with name \( x \) substituted for name \( y \).
-
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
-written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \)
-by a finite number of substitutions of bound variable names.  As a
-convention, we will identify \( \alpha \)-convertible processes and we
-assume that bound names and bound variable names of any processes are
-chosen to be different from the names and variable names that occur
-free in any other entities under consideration, such as processes,
-substitutions, and sets of names or variable names.  This is justified
-because any overlapping names and variable names may be
-\( \alpha \)-converted such that the assumption is satisfied.
-
-A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-We will think of contexts as functions between processes.
-A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsection{Challenge: Linearity and behavioral type systems}
-\label{sec:challenge:linearity-beh-types}
-
-This challenge formalises a proof that requires reasoning about linearity of channels.
-Linearity is the notion that a channel must be used exactly once in a process.
-This is necessary to prove properties about session type systems.
-Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
-
-The setting for this challenge is a small calculus with a session type system, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove type preservation (also known as subject reduction) for the type system, i.e.\ that well-typed processes can only transition to processes which are also well-typed in the same context.
-
-\subsubsection{Syntax}
-
-% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-% We assume the existence of some type of \emph{base values}, represented by the symbols
-%  \( a, b, \dots \), the existence of some type
-% of \emph{variable names}, represented by the symbols
-% \( l, m, \dots \), and the existence of some type of \emph{names},
-% represented by the symbols \( x, y, \dots \).
-
-The syntax is
-given by the grammar:
-\begin{metanote}
-AM:  Should we save some space like this?
-\end{metanote}
-% \begin{align*}
-%   v,w :=&&& a \\
-%   |&&& l \\
-%   P,Q :=&&& \Pend \\
-%   |&&& \Pout{x}{v}{P} \\
-%   |&&& \Pin{x}{l}{P} \\
-%   |&&& \Ppar{P}{Q} \\
-%   |&&& \Presd{x}{y}{P}
-         %   \end{align*}
-% \newcommand{\bmid}{\mathbf{\Vert}}
-\[
-\begin{array}{rcl}
-  v,w & ::= & a \mid l \\
-   P,Q & ::=& \Pend \mid \Pout{x}{v}{P} \mid \Pin{x}{l}{P} \mid \PBpar{P}{Q} \mid  \Presd{x}{y}{P}
-\end{array}
-\]
-A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \), then continue as \( P \).
-The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-% The input operator thus binds the variable name \( l \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Presd{x}{y}{P} \) is the \emph{restriction} of the channel with endpoints named \( x \) and \( y \) to \( P \).
-Components in \( P \) can use the names \( x \) and \( y \) to interact with each other (sending on \( x \) and receiving on \( y \) or vice versa), but not with processes outside of the restriction.
-The restriction operator thus binds the names \( x \) and \( y \) in \( P \).
-Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-Note that there is no recursion of replication in the syntax, and thus no infinite behaviors can be expressed.
-
-% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
-% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
-% We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
-% We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
-
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% As a convention, we assume that bound names and bound variable names of any processes are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, substitutions, and sets of names or variable names.
-% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-We describe the actions that the system can perform using a small step operational semantics.
-To simplify the definition of the reduction relation, we will first factor out a structural congruence relation on processes, and for this we need a few definitions.
-
-\subsubsection{Contexts and congruences}
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We will think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
-\begin{mathpar}
-  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
-  \and
-  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
-  \and
-  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
-  \and
-  \inferrule[Sc-Res-Par]{\{x,y\} \cap \freenames{Q} = \emptyset}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
-  \and
-  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
-  \and
-  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
-\end{mathpar}
-
-The operational semantics is then defined as the following binary relation on processes:
-\begin{mathpar}
-  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{v}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{v}{l}}{R}})}}}
-  \and
-  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
-  \and
-  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
-  \and
-  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathpar}
-Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable name or a name.
-
-\subsubsection{Session types}
-Our process syntax allows us to write processes that are not well-formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
-As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of restricting names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
-Another example is the following process, which attempts to send a variable that is not resolved at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
-
-To alleviate this issue, we introduce a \emph{session type system}, which will detect processes that are not well-formed.
-To this end, we need to first formally define what we mean by well-formed.
-We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
-A process is then \emph{well-formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
-
-Note that well-formed processes do not necessarily reduce, since e.g.\ the process
-\begin{equation*}
-  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
-\end{equation*}
-is well-formed, but has no reduction.
-\paragraph{Syntax}
-Our type system does not type processes directly, but instead focuses on the channels used in the process.
-The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
-\begin{align*}
-  S,T := &&& \Tend \\
-  |&&& \Tbase \\
-  |&&& \Tin{S} \\
-  |&&& \Tout{S} \\
-  \Gamma := &&& \Cempty \\
-  |&&& \Cadd{\Gamma}{\hastype{x}{S}}
-       %% AM: added
-        \\
-  |&&& \Cadd{\Gamma}{\hastype{l}{S}}
-\end{align*}
-The \emph{end type}, \( \Tend \), is the type of an endpoint on which no further interaction is possible.
-The \emph{base type}, \( \Tbase \), is the type of base values.
-The \emph{input type}, \( \Tin{S} \), is the type of an endpoint which is ready to receive a value, then continue with type \( S \).
-The \emph{output type}, \( \Tout{S} \), is the type of an endpoint which is ready to send a value, then continue with type \( S \).
-
-Typing contexts gather type information about names.
-The \emph{empty type context}, \( \Cempty \), does not contain any information.
-The \emph{type context with \( x \) added}, \( \Cadd{\Gamma}{\hastype{x}{T}} \) adds the information that \( x \) has type \( T \) to the existing type context \( \Gamma \).
-All of the names added to a type context must be distinct.
-We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
-
-Since we need to determine whether endpoints are used dually to determine whether processes are well-formed, we will need to formally define the dual of a type.
-The dual of a type, \( \dual{T} \), is defined by the following equations:
-\begin{mathpar}
-  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tend} = \Tend}
-\end{mathpar}
-Note that the dual function is not total, since it is not defined for base types.
-
-\paragraph{Splitting and updating contexts}
-Our type system will maintain two invariants:
-\begin{enumerate}
-\item Endpoints are used exactly once
-\item Endpoints that are part of the same restriction have dual types
-\end{enumerate}
-The first invariant is maintained by linearly splitting type contexts when typing compositions of processes.
-The second invariant is maintained by requiring duality when typing restrictions.
-
-To keep track of linearity, we introduce two \emph{qualifiers}: \( \un{T} \) and \( \lin{T} \).
-A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
-All types are \emph{linear}, \( \lin{T} \).
-The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
-
-The \emph{linear split} of a context, \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Split-Empty]{ }{\Cempty = \Csplit{\Cempty}{\Cempty}}
-  \and
-  \inferrule[Split-Un]{\Csplit{\Gamma_1}{\Gamma_2} = \Gamma \\ \un{T}}{\Cadd{\Gamma}{\hastype{x}{T}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{T}})}{(\Cadd{\Gamma_2}{\hastype{x}{T}})}}
-  \and
-  \inferrule[Split-Lin-L]{\Gamma = \Csplit{\Gamma_1}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{S}})}{\Gamma_2}}
-  \and
-  \inferrule[Split-Lin-R]{\Gamma = \Csplit{\Gamma_2}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{\Gamma_1}{(\Cadd{\Gamma_2}{\hastype{x}{S}})}}
-\end{mathpar}
-
-We will also need to update contexts with new type information in a safe way while typing input and output.
-The \emph{context update}, \( \Cupdate{\Gamma}{\hastype{x}{T}} = \Gamma' \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Update-Name]{\hastype{x}{U} \notin \Gamma}{\Cupdate{\Gamma}{\hastype{x}{T}} = \Cadd{\Gamma}{\hastype{x}{T}}}
-  \and
-  \inferrule[Update-Un]{\un{T}}{\Cupdate{(\Cadd{\Gamma}{\hastype{x}{T}})}{\hastype{x}{T}} = (\Cadd{\Gamma}{\hastype{x}{T}})}
-\end{mathpar}
-
-\subsubsection{Typing rules}
-We have three typing judgments: one for values, one for names, and one for processes.
-The typing rules for values are as follows:
-\begin{mathpar}
-  \inferrule[T-Base]{\un{\Gamma}}{\types{\Gamma}{\hastype{a}{\Tbase}}}
-  \and
-  \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{l}{\Tbase}}}{\hastype{l}{\Tbase}}}
-\end{mathpar}
-The only typing rule for names is:
-\begin{mathpar}
-  \inferrule[T-Name]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{x}{T}}}
-\end{mathpar}
-The typing rules for processes are as follows:
-\begin{mathpar}
-  \inferrule[T-Inact]{\un{\Gamma}}{\types{\Gamma}{\Pend}}
-  \and
-  \inferrule[T-Par]{\types{\Gamma_1}{P} \\ \types{\Gamma_2}{Q}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Ppar{P}{Q}}}
-  \and
-  \inferrule[T-Res]{\types{\Cadd{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P}}{\types{\Gamma}{\Presd{x}{y}{P}}}
-  \and
-  \inferrule[T-Out]{\types{\Gamma_1}{\hastype{x}{\Tout{T}}} \\ \types{\Gamma_2}{\hastype{v}{\Tbase}} \\ \types{\Cupdate{\Gamma_3}{\hastype{x}{T}}}{P}}{\types{\Csplit{\Gamma_1}{\Csplit{\Gamma_2}{\Gamma_3}}}{\Pout{x}{v}{P}}}
-  \and
-  \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{(\Cadd{\Gamma_2}{\hastype{l}{\Tbase}})}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
-\end{mathpar}
-
-\subsubsection{Challenge}
-The objective of this challenge is to prove the following theorems:
-\begin{theorem}
-  If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
-\end{theorem}
-
-\begin{theorem}
-  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
-\end{theorem}
-
-\begin{corollary}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
-\end{corollary}
-
-The following lemmas may be of use:
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \un{T} \) then \( \types{\Cadd{\Gamma}{\hastype{x}{T}}}{P} \).
-\end{lemma}
-
-\begin{lemma}
-  Assume \( \types{\Gamma}{P} \) and \( x \notin \freenames{P} \).
-  Then:
-  \begin{enumerate}
-  \item \( \hastype{x}{\Tout{T}} \notin \Gamma \).
-  \item \( \hastype{x}{\Tin{T}} \notin \Gamma \).
-  \item if \( \Gamma = \Cadd{\Gamma'}{\hastype{x}{T}} \) then \( \types{\Gamma'}{P} \).
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma}{Q} \).
-\end{lemma}
-
-\begin{lemma}
-  If \( \types{\Gamma_1}{\hastype{v}{T}} \) and \( \types{\Cadd{\Gamma_2}{\hastype{x}{T}}}{P} \) and \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then \( \types{\Gamma}{\subst{v}{x}{P}} \).
-\end{lemma}
-
-\subsection{Challenge: Name passing and scope extrusion}
-\label{sec:challenge:name-passing-scope-extrusion}
-
-This challenge formalises a proof that requires explicit scope extrusion.
-Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
-This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
-
-Reasoning about scope extrusion explicitly can sometimes be avoided by introducing a structural congruence rule into the semantics, but doing this means we lose information about the scope when reasoning about the semantics.
-Explicitly reasoning about scope extrusion is necessary to describe e.g.\ monitors and compositions.
-
-The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
-
-\subsubsection{Syntax}
-In this section 
- we assume that \emph{names}  can be compared for equality. 
-The syntax is:
-\begin{align*}
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{y}{P} \\
-  |&&& \Pin{x}{y}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Pchoice{P}{Q}
-\end{align*}
-%The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
-The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
-%The input operator thus binds the name \( y \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-% The restriction operator thus binds the name \( x \) in \( P \).
-Note that the scope of a restriction may change when processes interact. Namely, a restricted name may be sent \emph{outside} of its scope.
-% The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-Note that there is no recursion or replication in the syntax, and thus no infinite behaviors can be expressed.
-
-% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
-% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
-% We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
-
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
-% This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which is as follows:
-\begin{align*}
-  \alpha := &&& \Aoutf{x}{y} \\
-  |&&& \Ain{x}{y} \\
-  |&&& \Aoutb{x}{y} \\
-  |&&& \Atau
-\end{align*}
-The \emph{free output action} \( \Aoutf{x}{y} \) is sending the name \( y \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the name \( y \) via \( x \).
-The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
-
-We will extend the notion of free and bound occurrences with 
-\( \freenames{\alpha} \) to denote the set of names that occur free in
-the action \( \alpha \) and  \( \boundnames{\alpha} \) to
-denote the set of names that occur bound in the action \( \alpha \).
-In the free output action \( \Aoutf{x}{y} \) and the input action
-\( \Ain{x}{y} \), both \( x \) and \( y \) are free names.  In the
-bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while
-\( y \) is a bound name.  We will further use the notation
-\( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \)
-and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur
-in the action \( \alpha \).
-
-The transition relation is then defined by the following rules:
-\begin{mathpar}
-  \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{z}{P}}{\Ain{x}{y}}{\subst{P}{y}{z}}}
-  \and
-  \inferrule[Sum-L]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{P}{Q}}{\alpha}{P'}}
-  \and
-  \inferrule[Sum-R]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{Q}{P}}{\alpha}{P'}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'} \\ \boundnames{\alpha} \cap \freenames{Q} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'} \\ \boundnames{\alpha} \cap \freenames{P} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{y}}{P'} \\ \transition{Q}{\Ain{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{y}}{P'} \\ \transition{Q}{\Aoutf{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Close-L]{\transition{P}{\Aoutb{x}{z}}{P'} \\ \transition{Q}{\Ain{x}{z}}{Q'} \\ z \notin \freenames{Q}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Open]{\transition{P}{\Aoutf{x}{z}}{P'} \\ z \neq x}{\transition{\Pres{z}{P}}{\Aoutb{x}{z}}{P'}}
-  \and
-  \inferrule[Close-R]{\transition{P}{\Ain{x}{z}}{P'} \\ \transition{Q}{\Aoutb{x}{z}}{Q'} \\ z \notin \freenames{P}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
-\end{mathpar}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
-
-We keep the convention that bound names of any processes or actions
-are chosen to be different from the names that occur free in any other
-entities under consideration, such as processes, actions,
-substitutions, and sets of names.  The convention has one exception,
-namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \),
-the name \( z \) (which occurs bound in \( P \) and the action
-\( \Aoutb{x}{z} \)) may occur free in \( Q \).  Without this exception
-it would be impossible to express scope extrusion.
-
-\subsubsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
-
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
-\begin{align}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
-
-Two processes \( P \) and \( Q \) are then \emph{strong barbed bisimilar}, written \( \sbbisim{P}{Q} \), if:
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{Q'}~\textrm{for some \( Q' \) with}~\sbbisim{P'}{Q'} \\
-  \observable{Q}{\mu}~\textrm{implies}~\observable{P}{\mu} \\
-  \transition{Q}{\Atau}{Q'}~\textrm{implies}~\transition{P}{\Atau}{P'}~\textrm{for some \( P' \) with}~\sbbisim{P'}{Q'}
-\end{gather}
-Note that, since our calculus does not have infinite behavior, there is only one relation which satisfies these criteria.
-Thus we do not specify that the relation is the largest of its kind, and the relation does not need to be coinductively defined.
-To actually prove properties about the relation, however, we will still need either an inductive (or a coinductive) proof principle for the relation.
-
-\subsubsection{Challenge}
-The objective of this challenge is to prove the following theorem:
-\begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-
 \subsection{Challenge: Coinduction and reasoning about infinite processes}
@@
 \end{theorem}
-
-%\paragraph{Notes about this challenge}
-%Strong barbed bisimilarity is interesting by itself for this challenge, but the proof that it is an equivalence relation does not require up-to techniques, which are wide spread and interesting.
-%The definition of strong barbed congruence requires introducing contexts, but this is actually nice, since they are so prevalent in the field and we want to cover as many basics as possible.
-%The theorem about the strong barbed congruence requires up-to techniques, but also seems to involve introducing a structural equivalence, which is possibly not desirable.
-%On the other hand, it is a nice application of structural congruence which does not ``ruin'' the semantics.
-
-%% Local Variables:
-%% mode: latex
-%% TeX-master: "main"
-%% End:

### 2023-09-27 challenges/linearity.tex
Split challenges.tex into multiple files.

--- 2023-09-26 challenges.tex
@@
-\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
-%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
-\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
-                         {\endgroup]\end{quote}}
-\long\def\ignore#1{}
-
-\subsection{Preliminaries}
-\label{sec:prelim}
-
-We list here some common notions that we will use in the following. Since the calculi under study are somewhat different, we will note in each section which will apply.
-
-We assume the existence of some type of \emph{base values}, represented by the symbols
-\( a, b, \dots \), the existence of some type
-of \emph{variable names}, represented by the symbols
-\( l, m, \dots \), and the existence of some type of \emph{names},
-represented by the symbols \( x, y, \dots \). \ednote{AM: are we assuming that names can be compared for equality everywhere or only in C2?} %
-
-The syntax of processes will include:
-the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-
-For communication, processes include \emph{input} and \emph{output}, whose
-signature will depend on the calculus being value or name-passing.
-We use here the metavariables $c,k$ to abstract over this choice.
-The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
-\( c \) via \( x \), then continue as \( P \).  % The intention is that
-% the value \( v \) must be a base value when it is actually sent, and
-% this will be enforced in the semantics later on
-The process \( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
-via \( x \), then continue as \( P \) with the received element
-substituted for \( k \).  The input operator thus
-binds \( k \) in \( P \).
-
-The process \( \Pres{x}{P} \) is the \emph{restriction} of the name
-\( x \) to \( P \), binding \( x \) in \( P \).
-
-The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-Replication makes it possible to express infinite behaviors.
-
-We will use the notation \( \freenames{P} \) to denote the set of
-names that occur free,
-% (i.e.\ not bound by a restriction) in \( P \),
-\( \boundnames{P} \) to denote the set of names that occur bound
-% (by a restriction)
-in \( P \) and  \( \freevars{P} \)
-to denote the set of variable names that occur free
-% (i.e.\ not bound by an input)
-in \( P \).  We will use the notation \( \boundvars{P} \)
-for the set of variable names that occur bound
-% (by an input)
-in \( P \).  We will use the notation \( \subst{P}{v}{l} \) to denote
-the process \( P \) with value \( v \) substituted for variable name
-\( l \). Similarly, \( \subst{P}{x}{y} \) denotes the process
-\( P \) with name \( x \) substituted for name \( y \).
-
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
-written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \)
-by a finite number of substitutions of bound variable names.  As a
-convention, we will identify \( \alpha \)-convertible processes and we
-assume that bound names and bound variable names of any processes are
-chosen to be different from the names and variable names that occur
-free in any other entities under consideration, such as processes,
-substitutions, and sets of names or variable names.  This is justified
-because any overlapping names and variable names may be
-\( \alpha \)-converted such that the assumption is satisfied.
-
-A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-We will think of contexts as functions between processes.
-A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
 \subsection{Challenge: Linearity and behavioral type systems}
@@
 \end{lemma}
-
-\subsection{Challenge: Name passing and scope extrusion}
-\label{sec:challenge:name-passing-scope-extrusion}
-
-This challenge formalises a proof that requires explicit scope extrusion.
-Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
-This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
-
-Reasoning about scope extrusion explicitly can sometimes be avoided by introducing a structural congruence rule into the semantics, but doing this means we lose information about the scope when reasoning about the semantics.
-Explicitly reasoning about scope extrusion is necessary to describe e.g.\ monitors and compositions.
-
-The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
-
-\subsubsection{Syntax}
-In this section 
- we assume that \emph{names}  can be compared for equality. 
-The syntax is:
-\begin{align*}
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{y}{P} \\
-  |&&& \Pin{x}{y}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Pchoice{P}{Q}
-\end{align*}
-%The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
-The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
-%The input operator thus binds the name \( y \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-% The restriction operator thus binds the name \( x \) in \( P \).
-Note that the scope of a restriction may change when processes interact. Namely, a restricted name may be sent \emph{outside} of its scope.
-% The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-Note that there is no recursion or replication in the syntax, and thus no infinite behaviors can be expressed.
-
-% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
-% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
-% We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
-
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
-% This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which is as follows:
-\begin{align*}
-  \alpha := &&& \Aoutf{x}{y} \\
-  |&&& \Ain{x}{y} \\
-  |&&& \Aoutb{x}{y} \\
-  |&&& \Atau
-\end{align*}
-The \emph{free output action} \( \Aoutf{x}{y} \) is sending the name \( y \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the name \( y \) via \( x \).
-The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
-
-We will extend the notion of free and bound occurrences with 
-\( \freenames{\alpha} \) to denote the set of names that occur free in
-the action \( \alpha \) and  \( \boundnames{\alpha} \) to
-denote the set of names that occur bound in the action \( \alpha \).
-In the free output action \( \Aoutf{x}{y} \) and the input action
-\( \Ain{x}{y} \), both \( x \) and \( y \) are free names.  In the
-bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while
-\( y \) is a bound name.  We will further use the notation
-\( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \)
-and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur
-in the action \( \alpha \).
-
-The transition relation is then defined by the following rules:
-\begin{mathpar}
-  \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{z}{P}}{\Ain{x}{y}}{\subst{P}{y}{z}}}
-  \and
-  \inferrule[Sum-L]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{P}{Q}}{\alpha}{P'}}
-  \and
-  \inferrule[Sum-R]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{Q}{P}}{\alpha}{P'}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'} \\ \boundnames{\alpha} \cap \freenames{Q} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'} \\ \boundnames{\alpha} \cap \freenames{P} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{y}}{P'} \\ \transition{Q}{\Ain{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{y}}{P'} \\ \transition{Q}{\Aoutf{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Close-L]{\transition{P}{\Aoutb{x}{z}}{P'} \\ \transition{Q}{\Ain{x}{z}}{Q'} \\ z \notin \freenames{Q}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Open]{\transition{P}{\Aoutf{x}{z}}{P'} \\ z \neq x}{\transition{\Pres{z}{P}}{\Aoutb{x}{z}}{P'}}
-  \and
-  \inferrule[Close-R]{\transition{P}{\Ain{x}{z}}{P'} \\ \transition{Q}{\Aoutb{x}{z}}{Q'} \\ z \notin \freenames{P}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
-\end{mathpar}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
-
-We keep the convention that bound names of any processes or actions
-are chosen to be different from the names that occur free in any other
-entities under consideration, such as processes, actions,
-substitutions, and sets of names.  The convention has one exception,
-namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \),
-the name \( z \) (which occurs bound in \( P \) and the action
-\( \Aoutb{x}{z} \)) may occur free in \( Q \).  Without this exception
-it would be impossible to express scope extrusion.
-
-\subsubsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
-
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
-\begin{align}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
-
-Two processes \( P \) and \( Q \) are then \emph{strong barbed bisimilar}, written \( \sbbisim{P}{Q} \), if:
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{Q'}~\textrm{for some \( Q' \) with}~\sbbisim{P'}{Q'} \\
-  \observable{Q}{\mu}~\textrm{implies}~\observable{P}{\mu} \\
-  \transition{Q}{\Atau}{Q'}~\textrm{implies}~\transition{P}{\Atau}{P'}~\textrm{for some \( P' \) with}~\sbbisim{P'}{Q'}
-\end{gather}
-Note that, since our calculus does not have infinite behavior, there is only one relation which satisfies these criteria.
-Thus we do not specify that the relation is the largest of its kind, and the relation does not need to be coinductively defined.
-To actually prove properties about the relation, however, we will still need either an inductive (or a coinductive) proof principle for the relation.
-
-\subsubsection{Challenge}
-The objective of this challenge is to prove the following theorem:
-\begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-
-\subsection{Challenge: Coinduction and reasoning about infinite processes}
-\label{sec:challenge:coinduction}
-
-This challenge formalises a proof that requires coinductive techniques.
-Coinduction is a proof technique for infinite structures, which arise in this context due to systems with behaviors that continue indefinitely.
-Coinduction is the dual of induction: whereas induction is useful for proving properties of least fixed points, coinduction is useful for proving properties of greatest fixed points.
-
-The setting for this challenge is an untyped calculus of communicating systems with replication of processes, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
-Which, to reiterate, entails the use of coinduction due to the presence of infinite behavior.
-
-\subsubsection{Syntax}
-% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-The syntax differs from \ref{sec:challenge:linearity-beh-types} for including replication and having unary restriction:
-\begin{align*}
-  v,w :=&&& a \\
-  |&&& l \\
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{v}{P} \\
-  |&&& \Pin{x}{l}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Preplicate{P}
-\end{align*}
-% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-% The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
-% The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-% The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-% The input operator thus binds the variable name \( l \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-% The restriction operator thus binds the name \( x \) in \( P \).
-% Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-% The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-% It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-% Replication makes it possible to express infinite behaviours.
-
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
-
-% % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% Also as a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
-% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which are as follows:
-\begin{align*}
-  \alpha := &&& \Aoutf{x}{a} \\
-  |&&& \Ain{x}{a} \\
-  |&&& \Atau
-\end{align*}
-The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value \( a \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the base value \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
-
-We will use the notation \( \names{\alpha} \) to denote the set of names that occur in the action \( \alpha \).
-
-The transition relation is then defined by the following rules:
-\begin{mathpar}
-  \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{l}{P}}{\Ain{x}{a}}{\subst{P}{a}{l}}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{a}}{P'} \\ \transition{Q}{\Ain{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{a}}{P'} \\ \transition{Q}{\Aoutf{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ x \notin \names{\alpha}}{\transition{\Pres{x}{P}}{\alpha}{\Pres{x}{P'}}}
-  \and
-  \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
-\end{mathpar}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable name.
-
-% As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
-
-\subsubsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
-
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
-\begin{align}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
-
-\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is then the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}
-\end{gather}
-We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
-
-\begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-
-Unfortunately, strong barbed bisimilarity is not a good notion of process equivalence, since it does not consider the environment of processes.
-For instance, \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since \( \obsout{x} \) is the only observable in both processes and they cannot perform a \( \Atau \)-action, but \( \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since the left process can perform a \( \Atau \)-action such that \( \obsout{y} \) becomes observable, whereas the right process cannot.
-
-% \subsubsection{Contexts and congruences}
-% Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
-
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We can think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsubsection{Strong barbed congruence}
-A congruence is exactly what we need to make strong barbed bisimilarity consider the environment (i.e.\ the context) of processes.
-
-We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
-
-\begin{lemma}
-  \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \).
-\end{lemma}
-
-\subsubsection{Challenge}
-The objective of this challenge is to prove the following theorem:
-\begin{theorem}
-  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).
-\end{theorem}
-
-%\paragraph{Notes about this challenge}
-%Strong barbed bisimilarity is interesting by itself for this challenge, but the proof that it is an equivalence relation does not require up-to techniques, which are wide spread and interesting.
-%The definition of strong barbed congruence requires introducing contexts, but this is actually nice, since they are so prevalent in the field and we want to cover as many basics as possible.
-%The theorem about the strong barbed congruence requires up-to techniques, but also seems to involve introducing a structural equivalence, which is possibly not desirable.
-%On the other hand, it is a nice application of structural congruence which does not ``ruin'' the semantics.
-
-%% Local Variables:
-%% mode: latex
-%% TeX-master: "main"
-%% End:

### 2023-09-27 challenges/preliminaries.tex
Split challenges.tex into multiple files.

--- 2023-09-26 challenges.tex
@@
-\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
-%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
-\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
-                         {\endgroup]\end{quote}}
-\long\def\ignore#1{}
-
 \subsection{Preliminaries}
@@
 We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsection{Challenge: Linearity and behavioral type systems}
-\label{sec:challenge:linearity-beh-types}
-
-This challenge formalises a proof that requires reasoning about linearity of channels.
-Linearity is the notion that a channel must be used exactly once in a process.
-This is necessary to prove properties about session type systems.
-Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
-
-The setting for this challenge is a small calculus with a session type system, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove type preservation (also known as subject reduction) for the type system, i.e.\ that well-typed processes can only transition to processes which are also well-typed in the same context.
-
-\subsubsection{Syntax}
-
-% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-% We assume the existence of some type of \emph{base values}, represented by the symbols
-%  \( a, b, \dots \), the existence of some type
-% of \emph{variable names}, represented by the symbols
-% \( l, m, \dots \), and the existence of some type of \emph{names},
-% represented by the symbols \( x, y, \dots \).
-
-The syntax is
-given by the grammar:
-\begin{metanote}
-AM:  Should we save some space like this?
-\end{metanote}
-% \begin{align*}
-%   v,w :=&&& a \\
-%   |&&& l \\
-%   P,Q :=&&& \Pend \\
-%   |&&& \Pout{x}{v}{P} \\
-%   |&&& \Pin{x}{l}{P} \\
-%   |&&& \Ppar{P}{Q} \\
-%   |&&& \Presd{x}{y}{P}
-         %   \end{align*}
-% \newcommand{\bmid}{\mathbf{\Vert}}
-\[
-\begin{array}{rcl}
-  v,w & ::= & a \mid l \\
-   P,Q & ::=& \Pend \mid \Pout{x}{v}{P} \mid \Pin{x}{l}{P} \mid \PBpar{P}{Q} \mid  \Presd{x}{y}{P}
-\end{array}
-\]
-A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \), then continue as \( P \).
-The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-% The input operator thus binds the variable name \( l \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Presd{x}{y}{P} \) is the \emph{restriction} of the channel with endpoints named \( x \) and \( y \) to \( P \).
-Components in \( P \) can use the names \( x \) and \( y \) to interact with each other (sending on \( x \) and receiving on \( y \) or vice versa), but not with processes outside of the restriction.
-The restriction operator thus binds the names \( x \) and \( y \) in \( P \).
-Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-Note that there is no recursion of replication in the syntax, and thus no infinite behaviors can be expressed.
-
-% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
-% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
-% We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
-% We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
-
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% As a convention, we assume that bound names and bound variable names of any processes are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, substitutions, and sets of names or variable names.
-% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-We describe the actions that the system can perform using a small step operational semantics.
-To simplify the definition of the reduction relation, we will first factor out a structural congruence relation on processes, and for this we need a few definitions.
-
-\subsubsection{Contexts and congruences}
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We will think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
-\begin{mathpar}
-  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
-  \and
-  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
-  \and
-  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
-  \and
-  \inferrule[Sc-Res-Par]{\{x,y\} \cap \freenames{Q} = \emptyset}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
-  \and
-  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
-  \and
-  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
-\end{mathpar}
-
-The operational semantics is then defined as the following binary relation on processes:
-\begin{mathpar}
-  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{v}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{v}{l}}{R}})}}}
-  \and
-  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
-  \and
-  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
-  \and
-  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathpar}
-Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable name or a name.
-
-\subsubsection{Session types}
-Our process syntax allows us to write processes that are not well-formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
-As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of restricting names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
-Another example is the following process, which attempts to send a variable that is not resolved at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
-
-To alleviate this issue, we introduce a \emph{session type system}, which will detect processes that are not well-formed.
-To this end, we need to first formally define what we mean by well-formed.
-We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
-A process is then \emph{well-formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
-
-Note that well-formed processes do not necessarily reduce, since e.g.\ the process
-\begin{equation*}
-  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
-\end{equation*}
-is well-formed, but has no reduction.
-\paragraph{Syntax}
-Our type system does not type processes directly, but instead focuses on the channels used in the process.
-The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
-\begin{align*}
-  S,T := &&& \Tend \\
-  |&&& \Tbase \\
-  |&&& \Tin{S} \\
-  |&&& \Tout{S} \\
-  \Gamma := &&& \Cempty \\
-  |&&& \Cadd{\Gamma}{\hastype{x}{S}}
-       %% AM: added
-        \\
-  |&&& \Cadd{\Gamma}{\hastype{l}{S}}
-\end{align*}
-The \emph{end type}, \( \Tend \), is the type of an endpoint on which no further interaction is possible.
-The \emph{base type}, \( \Tbase \), is the type of base values.
-The \emph{input type}, \( \Tin{S} \), is the type of an endpoint which is ready to receive a value, then continue with type \( S \).
-The \emph{output type}, \( \Tout{S} \), is the type of an endpoint which is ready to send a value, then continue with type \( S \).
-
-Typing contexts gather type information about names.
-The \emph{empty type context}, \( \Cempty \), does not contain any information.
-The \emph{type context with \( x \) added}, \( \Cadd{\Gamma}{\hastype{x}{T}} \) adds the information that \( x \) has type \( T \) to the existing type context \( \Gamma \).
-All of the names added to a type context must be distinct.
-We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
-
-Since we need to determine whether endpoints are used dually to determine whether processes are well-formed, we will need to formally define the dual of a type.
-The dual of a type, \( \dual{T} \), is defined by the following equations:
-\begin{mathpar}
-  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tend} = \Tend}
-\end{mathpar}
-Note that the dual function is not total, since it is not defined for base types.
-
-\paragraph{Splitting and updating contexts}
-Our type system will maintain two invariants:
-\begin{enumerate}
-\item Endpoints are used exactly once
-\item Endpoints that are part of the same restriction have dual types
-\end{enumerate}
-The first invariant is maintained by linearly splitting type contexts when typing compositions of processes.
-The second invariant is maintained by requiring duality when typing restrictions.
-
-To keep track of linearity, we introduce two \emph{qualifiers}: \( \un{T} \) and \( \lin{T} \).
-A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
-All types are \emph{linear}, \( \lin{T} \).
-The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
-
-The \emph{linear split} of a context, \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Split-Empty]{ }{\Cempty = \Csplit{\Cempty}{\Cempty}}
-  \and
-  \inferrule[Split-Un]{\Csplit{\Gamma_1}{\Gamma_2} = \Gamma \\ \un{T}}{\Cadd{\Gamma}{\hastype{x}{T}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{T}})}{(\Cadd{\Gamma_2}{\hastype{x}{T}})}}
-  \and
-  \inferrule[Split-Lin-L]{\Gamma = \Csplit{\Gamma_1}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{S}})}{\Gamma_2}}
-  \and
-  \inferrule[Split-Lin-R]{\Gamma = \Csplit{\Gamma_2}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{\Gamma_1}{(\Cadd{\Gamma_2}{\hastype{x}{S}})}}
-\end{mathpar}
-
-We will also need to update contexts with new type information in a safe way while typing input and output.
-The \emph{context update}, \( \Cupdate{\Gamma}{\hastype{x}{T}} = \Gamma' \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Update-Name]{\hastype{x}{U} \notin \Gamma}{\Cupdate{\Gamma}{\hastype{x}{T}} = \Cadd{\Gamma}{\hastype{x}{T}}}
-  \and
-  \inferrule[Update-Un]{\un{T}}{\Cupdate{(\Cadd{\Gamma}{\hastype{x}{T}})}{\hastype{x}{T}} = (\Cadd{\Gamma}{\hastype{x}{T}})}
-\end{mathpar}
-
-\subsubsection{Typing rules}
-We have three typing judgments: one for values, one for names, and one for processes.
-The typing rules for values are as follows:
-\begin{mathpar}
-  \inferrule[T-Base]{\un{\Gamma}}{\types{\Gamma}{\hastype{a}{\Tbase}}}
-  \and
-  \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{l}{\Tbase}}}{\hastype{l}{\Tbase}}}
-\end{mathpar}
-The only typing rule for names is:
-\begin{mathpar}
-  \inferrule[T-Name]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{x}{T}}}
-\end{mathpar}
-The typing rules for processes are as follows:
-\begin{mathpar}
-  \inferrule[T-Inact]{\un{\Gamma}}{\types{\Gamma}{\Pend}}
-  \and
-  \inferrule[T-Par]{\types{\Gamma_1}{P} \\ \types{\Gamma_2}{Q}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Ppar{P}{Q}}}
-  \and
-  \inferrule[T-Res]{\types{\Cadd{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P}}{\types{\Gamma}{\Presd{x}{y}{P}}}
-  \and
-  \inferrule[T-Out]{\types{\Gamma_1}{\hastype{x}{\Tout{T}}} \\ \types{\Gamma_2}{\hastype{v}{\Tbase}} \\ \types{\Cupdate{\Gamma_3}{\hastype{x}{T}}}{P}}{\types{\Csplit{\Gamma_1}{\Csplit{\Gamma_2}{\Gamma_3}}}{\Pout{x}{v}{P}}}
-  \and
-  \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{(\Cadd{\Gamma_2}{\hastype{l}{\Tbase}})}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
-\end{mathpar}
-
-\subsubsection{Challenge}
-The objective of this challenge is to prove the following theorems:
-\begin{theorem}
-  If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
-\end{theorem}
-
-\begin{theorem}
-  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
-\end{theorem}
-
-\begin{corollary}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
-\end{corollary}
-
-The following lemmas may be of use:
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \un{T} \) then \( \types{\Cadd{\Gamma}{\hastype{x}{T}}}{P} \).
-\end{lemma}
-
-\begin{lemma}
-  Assume \( \types{\Gamma}{P} \) and \( x \notin \freenames{P} \).
-  Then:
-  \begin{enumerate}
-  \item \( \hastype{x}{\Tout{T}} \notin \Gamma \).
-  \item \( \hastype{x}{\Tin{T}} \notin \Gamma \).
-  \item if \( \Gamma = \Cadd{\Gamma'}{\hastype{x}{T}} \) then \( \types{\Gamma'}{P} \).
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma}{Q} \).
-\end{lemma}
-
-\begin{lemma}
-  If \( \types{\Gamma_1}{\hastype{v}{T}} \) and \( \types{\Cadd{\Gamma_2}{\hastype{x}{T}}}{P} \) and \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then \( \types{\Gamma}{\subst{v}{x}{P}} \).
-\end{lemma}
-
-\subsection{Challenge: Name passing and scope extrusion}
-\label{sec:challenge:name-passing-scope-extrusion}
-
-This challenge formalises a proof that requires explicit scope extrusion.
-Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
-This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
-
-Reasoning about scope extrusion explicitly can sometimes be avoided by introducing a structural congruence rule into the semantics, but doing this means we lose information about the scope when reasoning about the semantics.
-Explicitly reasoning about scope extrusion is necessary to describe e.g.\ monitors and compositions.
-
-The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
-
-\subsubsection{Syntax}
-In this section 
- we assume that \emph{names}  can be compared for equality. 
-The syntax is:
-\begin{align*}
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{y}{P} \\
-  |&&& \Pin{x}{y}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Pchoice{P}{Q}
-\end{align*}
-%The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
-The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
-%The input operator thus binds the name \( y \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-% The restriction operator thus binds the name \( x \) in \( P \).
-Note that the scope of a restriction may change when processes interact. Namely, a restricted name may be sent \emph{outside} of its scope.
-% The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-Note that there is no recursion or replication in the syntax, and thus no infinite behaviors can be expressed.
-
-% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
-% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
-% We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
-
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
-% This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which is as follows:
-\begin{align*}
-  \alpha := &&& \Aoutf{x}{y} \\
-  |&&& \Ain{x}{y} \\
-  |&&& \Aoutb{x}{y} \\
-  |&&& \Atau
-\end{align*}
-The \emph{free output action} \( \Aoutf{x}{y} \) is sending the name \( y \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the name \( y \) via \( x \).
-The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
-
-We will extend the notion of free and bound occurrences with 
-\( \freenames{\alpha} \) to denote the set of names that occur free in
-the action \( \alpha \) and  \( \boundnames{\alpha} \) to
-denote the set of names that occur bound in the action \( \alpha \).
-In the free output action \( \Aoutf{x}{y} \) and the input action
-\( \Ain{x}{y} \), both \( x \) and \( y \) are free names.  In the
-bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while
-\( y \) is a bound name.  We will further use the notation
-\( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \)
-and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur
-in the action \( \alpha \).
-
-The transition relation is then defined by the following rules:
-\begin{mathpar}
-  \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{z}{P}}{\Ain{x}{y}}{\subst{P}{y}{z}}}
-  \and
-  \inferrule[Sum-L]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{P}{Q}}{\alpha}{P'}}
-  \and
-  \inferrule[Sum-R]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{Q}{P}}{\alpha}{P'}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'} \\ \boundnames{\alpha} \cap \freenames{Q} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'} \\ \boundnames{\alpha} \cap \freenames{P} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{y}}{P'} \\ \transition{Q}{\Ain{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{y}}{P'} \\ \transition{Q}{\Aoutf{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Close-L]{\transition{P}{\Aoutb{x}{z}}{P'} \\ \transition{Q}{\Ain{x}{z}}{Q'} \\ z \notin \freenames{Q}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Open]{\transition{P}{\Aoutf{x}{z}}{P'} \\ z \neq x}{\transition{\Pres{z}{P}}{\Aoutb{x}{z}}{P'}}
-  \and
-  \inferrule[Close-R]{\transition{P}{\Ain{x}{z}}{P'} \\ \transition{Q}{\Aoutb{x}{z}}{Q'} \\ z \notin \freenames{P}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
-\end{mathpar}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
-
-We keep the convention that bound names of any processes or actions
-are chosen to be different from the names that occur free in any other
-entities under consideration, such as processes, actions,
-substitutions, and sets of names.  The convention has one exception,
-namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \),
-the name \( z \) (which occurs bound in \( P \) and the action
-\( \Aoutb{x}{z} \)) may occur free in \( Q \).  Without this exception
-it would be impossible to express scope extrusion.
-
-\subsubsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
-
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
-\begin{align}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
-
-Two processes \( P \) and \( Q \) are then \emph{strong barbed bisimilar}, written \( \sbbisim{P}{Q} \), if:
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{Q'}~\textrm{for some \( Q' \) with}~\sbbisim{P'}{Q'} \\
-  \observable{Q}{\mu}~\textrm{implies}~\observable{P}{\mu} \\
-  \transition{Q}{\Atau}{Q'}~\textrm{implies}~\transition{P}{\Atau}{P'}~\textrm{for some \( P' \) with}~\sbbisim{P'}{Q'}
-\end{gather}
-Note that, since our calculus does not have infinite behavior, there is only one relation which satisfies these criteria.
-Thus we do not specify that the relation is the largest of its kind, and the relation does not need to be coinductively defined.
-To actually prove properties about the relation, however, we will still need either an inductive (or a coinductive) proof principle for the relation.
-
-\subsubsection{Challenge}
-The objective of this challenge is to prove the following theorem:
-\begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-
-\subsection{Challenge: Coinduction and reasoning about infinite processes}
-\label{sec:challenge:coinduction}
-
-This challenge formalises a proof that requires coinductive techniques.
-Coinduction is a proof technique for infinite structures, which arise in this context due to systems with behaviors that continue indefinitely.
-Coinduction is the dual of induction: whereas induction is useful for proving properties of least fixed points, coinduction is useful for proving properties of greatest fixed points.
-
-The setting for this challenge is an untyped calculus of communicating systems with replication of processes, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
-Which, to reiterate, entails the use of coinduction due to the presence of infinite behavior.
-
-\subsubsection{Syntax}
-% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-The syntax differs from \ref{sec:challenge:linearity-beh-types} for including replication and having unary restriction:
-\begin{align*}
-  v,w :=&&& a \\
-  |&&& l \\
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{v}{P} \\
-  |&&& \Pin{x}{l}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Preplicate{P}
-\end{align*}
-% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-% The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
-% The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-% The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-% The input operator thus binds the variable name \( l \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-% The restriction operator thus binds the name \( x \) in \( P \).
-% Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-% The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-% It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-% Replication makes it possible to express infinite behaviours.
-
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
-
-% % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% Also as a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
-% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which are as follows:
-\begin{align*}
-  \alpha := &&& \Aoutf{x}{a} \\
-  |&&& \Ain{x}{a} \\
-  |&&& \Atau
-\end{align*}
-The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value \( a \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the base value \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
-
-We will use the notation \( \names{\alpha} \) to denote the set of names that occur in the action \( \alpha \).
-
-The transition relation is then defined by the following rules:
-\begin{mathpar}
-  \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{l}{P}}{\Ain{x}{a}}{\subst{P}{a}{l}}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{a}}{P'} \\ \transition{Q}{\Ain{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{a}}{P'} \\ \transition{Q}{\Aoutf{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ x \notin \names{\alpha}}{\transition{\Pres{x}{P}}{\alpha}{\Pres{x}{P'}}}
-  \and
-  \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
-\end{mathpar}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable name.
-
-% As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
-
-\subsubsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
-
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
-\begin{align}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
-
-\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is then the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}
-\end{gather}
-We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
-
-\begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-
-Unfortunately, strong barbed bisimilarity is not a good notion of process equivalence, since it does not consider the environment of processes.
-For instance, \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since \( \obsout{x} \) is the only observable in both processes and they cannot perform a \( \Atau \)-action, but \( \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since the left process can perform a \( \Atau \)-action such that \( \obsout{y} \) becomes observable, whereas the right process cannot.
-
-% \subsubsection{Contexts and congruences}
-% Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
-
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We can think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsubsection{Strong barbed congruence}
-A congruence is exactly what we need to make strong barbed bisimilarity consider the environment (i.e.\ the context) of processes.
-
-We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
-
-\begin{lemma}
-  \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \).
-\end{lemma}
-
-\subsubsection{Challenge}
-The objective of this challenge is to prove the following theorem:
-\begin{theorem}
-  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).
-\end{theorem}
-
-%\paragraph{Notes about this challenge}
-%Strong barbed bisimilarity is interesting by itself for this challenge, but the proof that it is an equivalence relation does not require up-to techniques, which are wide spread and interesting.
-%The definition of strong barbed congruence requires introducing contexts, but this is actually nice, since they are so prevalent in the field and we want to cover as many basics as possible.
-%The theorem about the strong barbed congruence requires up-to techniques, but also seems to involve introducing a structural equivalence, which is possibly not desirable.
-%On the other hand, it is a nice application of structural congruence which does not ``ruin'' the semantics.
-
-%% Local Variables:
-%% mode: latex
-%% TeX-master: "main"
-%% End:

### 2023-09-27 challenges/scope-extrusion.tex
Split challenges.tex into multiple files.

--- 2023-09-26 challenges.tex
@@
-\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
-%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
-\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
-                         {\endgroup]\end{quote}}
-\long\def\ignore#1{}
-
-\subsection{Preliminaries}
-\label{sec:prelim}
-
-We list here some common notions that we will use in the following. Since the calculi under study are somewhat different, we will note in each section which will apply.
-
-We assume the existence of some type of \emph{base values}, represented by the symbols
-\( a, b, \dots \), the existence of some type
-of \emph{variable names}, represented by the symbols
-\( l, m, \dots \), and the existence of some type of \emph{names},
-represented by the symbols \( x, y, \dots \). \ednote{AM: are we assuming that names can be compared for equality everywhere or only in C2?} %
-
-The syntax of processes will include:
-the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-
-For communication, processes include \emph{input} and \emph{output}, whose
-signature will depend on the calculus being value or name-passing.
-We use here the metavariables $c,k$ to abstract over this choice.
-The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
-\( c \) via \( x \), then continue as \( P \).  % The intention is that
-% the value \( v \) must be a base value when it is actually sent, and
-% this will be enforced in the semantics later on
-The process \( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
-via \( x \), then continue as \( P \) with the received element
-substituted for \( k \).  The input operator thus
-binds \( k \) in \( P \).
-
-The process \( \Pres{x}{P} \) is the \emph{restriction} of the name
-\( x \) to \( P \), binding \( x \) in \( P \).
-
-The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-Replication makes it possible to express infinite behaviors.
-
-We will use the notation \( \freenames{P} \) to denote the set of
-names that occur free,
-% (i.e.\ not bound by a restriction) in \( P \),
-\( \boundnames{P} \) to denote the set of names that occur bound
-% (by a restriction)
-in \( P \) and  \( \freevars{P} \)
-to denote the set of variable names that occur free
-% (i.e.\ not bound by an input)
-in \( P \).  We will use the notation \( \boundvars{P} \)
-for the set of variable names that occur bound
-% (by an input)
-in \( P \).  We will use the notation \( \subst{P}{v}{l} \) to denote
-the process \( P \) with value \( v \) substituted for variable name
-\( l \). Similarly, \( \subst{P}{x}{y} \) denotes the process
-\( P \) with name \( x \) substituted for name \( y \).
-
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
-written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \)
-by a finite number of substitutions of bound variable names.  As a
-convention, we will identify \( \alpha \)-convertible processes and we
-assume that bound names and bound variable names of any processes are
-chosen to be different from the names and variable names that occur
-free in any other entities under consideration, such as processes,
-substitutions, and sets of names or variable names.  This is justified
-because any overlapping names and variable names may be
-\( \alpha \)-converted such that the assumption is satisfied.
-
-A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-We will think of contexts as functions between processes.
-A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsection{Challenge: Linearity and behavioral type systems}
-\label{sec:challenge:linearity-beh-types}
-
-This challenge formalises a proof that requires reasoning about linearity of channels.
-Linearity is the notion that a channel must be used exactly once in a process.
-This is necessary to prove properties about session type systems.
-Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
-
-The setting for this challenge is a small calculus with a session type system, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove type preservation (also known as subject reduction) for the type system, i.e.\ that well-typed processes can only transition to processes which are also well-typed in the same context.
-
-\subsubsection{Syntax}
-
-% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-% We assume the existence of some type of \emph{base values}, represented by the symbols
-%  \( a, b, \dots \), the existence of some type
-% of \emph{variable names}, represented by the symbols
-% \( l, m, \dots \), and the existence of some type of \emph{names},
-% represented by the symbols \( x, y, \dots \).
-
-The syntax is
-given by the grammar:
-\begin{metanote}
-AM:  Should we save some space like this?
-\end{metanote}
-% \begin{align*}
-%   v,w :=&&& a \\
-%   |&&& l \\
-%   P,Q :=&&& \Pend \\
-%   |&&& \Pout{x}{v}{P} \\
-%   |&&& \Pin{x}{l}{P} \\
-%   |&&& \Ppar{P}{Q} \\
-%   |&&& \Presd{x}{y}{P}
-% \end{align*}
-\[
-\begin{array}{rcl}
-  v,w & ::= & a \mid l \\
-   P,Q & ::=& \Pend \mid \Pout{x}{v}{P} \mid \Pin{x}{l}{P} \mid \Ppar{P}{Q} \mid  \Presd{x}{y}{P}
-\end{array}
-\]
-A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \), then continue as \( P \).
-The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-% The input operator thus binds the variable name \( l \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Presd{x}{y}{P} \) is the \emph{restriction} of the channel with endpoints named \( x \) and \( y \) to \( P \).
-Components in \( P \) can use the names \( x \) and \( y \) to interact with each other (sending on \( x \) and receiving on \( y \) or vice versa), but not with processes outside of the restriction.
-The restriction operator thus binds the names \( x \) and \( y \) in \( P \).
-Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-Note that there is no recursion of replication in the syntax, and thus no infinite behaviors can be expressed.
-
-% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
-% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
-% We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
-% We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
-
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% As a convention, we assume that bound names and bound variable names of any processes are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, substitutions, and sets of names or variable names.
-% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-We describe the actions that the system can perform using a small step operational semantics.
-To simplify the definition of the reduction relation, we will first factor out a structural congruence relation on processes, and for this we need a few definitions.
-
-\subsubsection{Contexts and congruences}
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We will think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
-\begin{mathpar}
-  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
-  \and
-  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
-  \and
-  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
-  \and
-  \inferrule[Sc-Res-Par]{\{x,y\} \not\subseteq \freenames{Q}}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
-  \and
-  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
-  \and
-  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
-\end{mathpar}
-
-The operational semantics is then defined as the following binary relation on processes:
-\begin{mathpar}
-  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{v}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{v}{l}}{R}})}}}
-  \and
-  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
-  \and
-  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
-  \and
-  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathpar}
-Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable name or a name.
-
-\subsubsection{Session types}
-Our process syntax allows us to write processes that are not well-formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
-As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of restricting names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
-Another example is the following process, which attempts to send a variable that is not resolved at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
-
-To alleviate this issue, we introduce a \emph{session type system}, which will detect processes that are not well-formed.
-To this end, we need to first formally define what we mean by well-formed.
-We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
-A process is then \emph{well-formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
-
-Note that well-formed processes do not necessarily reduce, since e.g.\ the process
-\begin{equation*}
-  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
-\end{equation*}
-is well-formed, but has no reduction.
-\paragraph{Syntax}
-Our type system does not type processes directly, but instead focuses on the channels used in the process.
-The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
-\begin{align*}
-  S,T := &&& \Tend \\
-  |&&& \Tbase \\
-  |&&& \Tin{S} \\
-  |&&& \Tout{S} \\
-  \Gamma := &&& \Cempty \\
-  |&&& \Cadd{\Gamma}{\hastype{x}{S}}
-       %% AM: added
-        \\
-  |&&& \Cadd{\Gamma}{\hastype{l}{S}}
-\end{align*}
-The \emph{end type}, \( \Tend \), is the type of an endpoint on which no further interaction is possible.
-The \emph{base type}, \( \Tbase \), is the type of base values.
-The \emph{input type}, \( \Tin{S} \), is the type of an endpoint which is ready to receive a value, then continue with type \( S \).
-The \emph{output type}, \( \Tout{S} \), is the type of an endpoint which is ready to send a value, then continue with type \( S \).
-
-Typing contexts gather type information about names.
-The \emph{empty type context}, \( \Cempty \), does not contain any information.
-The \emph{type context with \( x \) added}, \( \Cadd{\Gamma}{\hastype{x}{T}} \) adds the information that \( x \) has type \( T \) to the existing type context \( \Gamma \).
-All of the names added to a type context must be distinct.
-We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
-
-Since we need to determine whether endpoints are used dually to determine whether processes are well-formed, we will need to formally define the dual of a type.
-The dual of a type, \( \dual{T} \), is defined by the following equations:
-\begin{mathpar}
-  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tend} = \Tend}
-\end{mathpar}
-Note that the dual function is not total, since it is not defined for base types.
-
-\paragraph{Splitting and updating contexts}
-Our type system will maintain two invariants:
-\begin{enumerate}
-\item Endpoints are used exactly once
-\item Endpoints that are part of the same restriction have dual types
-\end{enumerate}
-The first invariant is maintained by linearly splitting type contexts when typing compositions of processes.
-The second invariant is maintained by requiring duality when typing restrictions.
-
-To keep track of linearity, we introduce two \emph{qualifiers}: \( \un{T} \) and \( \lin{T} \).
-A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
-All types are \emph{linear}, \( \lin{T} \).
-The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
-
-The \emph{linear split} of a context, \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Split-Empty]{ }{\Cempty = \Csplit{\Cempty}{\Cempty}}
-  \and
-  \inferrule[Split-Un]{\Csplit{\Gamma_1}{\Gamma_2} = \Gamma \\ \un{T}}{\Cadd{\Gamma}{\hastype{x}{T}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{T}})}{(\Cadd{\Gamma_2}{\hastype{x}{T}})}}
-  \and
-  \inferrule[Split-Lin-L]{\Gamma = \Csplit{\Gamma_1}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{S}})}{\Gamma_2}}
-  \and
-  \inferrule[Split-Lin-R]{\Gamma = \Csplit{\Gamma_2}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{\Gamma_1}{(\Cadd{\Gamma_2}{\hastype{x}{S}})}}
-\end{mathpar}
-
-We will also need to update contexts with new type information in a safe way while typing input and output.
-The \emph{context update}, \( \Cupdate{\Gamma}{\hastype{x}{T}} = \Gamma' \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Update-Name]{\hastype{x}{U} \notin \Gamma}{\Cupdate{\Gamma}{\hastype{x}{T}} = \Cadd{\Gamma}{\hastype{x}{T}}}
-  \and
-  \inferrule[Update-Un]{\un{T}}{\Cupdate{(\Cadd{\Gamma}{\hastype{x}{T}})}{\hastype{x}{T}} = (\Cadd{\Gamma}{\hastype{x}{T}})}
-\end{mathpar}
-
-\subsubsection{Typing rules}
-We have three typing judgments: one for values, one for names, and one for processes.
-The typing rules for values are as follows:
-\begin{mathpar}
-  \inferrule[T-Base]{\un{\Gamma}}{\types{\Gamma}{\hastype{a}{\Tbase}}}
-  \and
-  \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{l}{\Tbase}}}{\hastype{l}{\Tbase}}}
-\end{mathpar}
-The only typing rule for names is:
-\begin{mathpar}
-  \inferrule[T-Name]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{x}{T}}}
-\end{mathpar}
-The typing rules for processes are as follows:
-\begin{mathpar}
-  \inferrule[T-Inact]{\un{\Gamma}}{\types{\Gamma}{\Pend}}
-  \and
-  \inferrule[T-Par]{\types{\Gamma_1}{P} \\ \types{\Gamma_2}{Q}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Ppar{P}{Q}}}
-  \and
-  \inferrule[T-Res]{\types{\Cadd{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P}}{\types{\Gamma}{\Presd{x}{y}{P}}}
-  \and
-  \inferrule[T-Out]{\types{\Gamma_1}{\hastype{x}{\Tout{T}}} \\ \types{\Gamma_2}{\hastype{v}{\Tbase}} \\ \types{\Cupdate{\Gamma_3}{\hastype{x}{T}}}{P}}{\types{\Csplit{\Gamma_1}{\Csplit{\Gamma_2}{\Gamma_3}}}{\Pout{x}{v}{P}}}
-  \and
-  \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{(\Cadd{\Gamma_2}{\hastype{l}{\Tbase}})}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
-\end{mathpar}
-
-\subsubsection{Challenge}
-The objective of this challenge is to prove the following theorems:
-\begin{theorem}
-  If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
-\end{theorem}
-
-\begin{theorem}
-  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
-\end{theorem}
-
-\begin{corollary}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
-\end{corollary}
-
-The following lemmas may be of use:
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \un{T} \) then \( \types{\Cadd{\Gamma}{\hastype{x}{T}}}{P} \).
-\end{lemma}
-
-\begin{lemma}
-  Assume \( \types{\Gamma}{P} \) and \( x \notin \freenames{P} \).
-  Then:
-  \begin{enumerate}
-  \item \( \hastype{x}{\Tout{T}} \notin \Gamma \).
-  \item \( \hastype{x}{\Tin{T}} \notin \Gamma \).
-  \item if \( \Gamma = \Cadd{\Gamma'}{\hastype{x}{T}} \) then \( \types{\Gamma'}{P} \).
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma}{Q} \).
-\end{lemma}
-
-\begin{lemma}
-  If \( \types{\Gamma_1}{\hastype{v}{T}} \) and \( \types{\Cadd{\Gamma_2}{\hastype{x}{T}}}{P} \) and \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then \( \types{\Gamma}{\subst{v}{x}{P}} \).
-\end{lemma}
-
 \subsection{Challenge: Name passing and scope extrusion}
@@
 \end{proof}
-
-\subsection{Challenge: Coinduction and reasoning about infinite processes}
-\label{sec:challenge:coinduction}
-
-This challenge formalises a proof that requires coinductive techniques.
-Coinduction is a proof technique for infinite structures, which arise in this context due to systems with behaviors that continue indefinitely.
-Coinduction is the dual of induction: whereas induction is useful for proving properties of least fixed points, coinduction is useful for proving properties of greatest fixed points.
-
-The setting for this challenge is an untyped calculus of communicating systems with replication of processes, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
-Which, to reiterate, entails the use of coinduction due to the presence of infinite behavior.
-
-\subsubsection{Syntax}
-% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-The syntax differs from \ref{sec:challenge:linearity-beh-types} for including replication and having unary restriction:
-\begin{align*}
-  v,w :=&&& a \\
-  |&&& l \\
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{v}{P} \\
-  |&&& \Pin{x}{l}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Preplicate{P}
-\end{align*}
-% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-% The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
-% The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-% The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-% The input operator thus binds the variable name \( l \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-% The restriction operator thus binds the name \( x \) in \( P \).
-% Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-% The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-% It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-% Replication makes it possible to express infinite behaviours.
-
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
-
-% % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% Also as a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
-% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which are as follows:
-\begin{align*}
-  \alpha := &&& \Aoutf{x}{a} \\
-  |&&& \Ain{x}{a} \\
-  |&&& \Atau
-\end{align*}
-The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value \( a \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the base value \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
-
-We will use the notation \( \names{\alpha} \) to denote the set of names that occur in the action \( \alpha \).
-
-The transition relation is then defined by the following rules:
-\begin{mathpar}
-  \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{l}{P}}{\Ain{x}{a}}{\subst{P}{a}{l}}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{a}}{P'} \\ \transition{Q}{\Ain{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{a}}{P'} \\ \transition{Q}{\Aoutf{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ x \notin \names{\alpha}}{\transition{\Pres{x}{P}}{\alpha}{\Pres{x}{P'}}}
-  \and
-  \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
-\end{mathpar}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable name.
-
-% As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
-
-\subsubsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
-
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
-\begin{align}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
-
-\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is then the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}
-\end{gather}
-We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
-
-\begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-
-Unfortunately, strong barbed bisimilarity is not a good notion of process equivalence, since it does not consider the environment of processes.
-For instance, \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since \( \obsout{x} \) is the only observable in both processes and they cannot perform a \( \Atau \)-action, but \( \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since the left process can perform a \( \Atau \)-action such that \( \obsout{y} \) becomes observable, whereas the right process cannot.
-
-% \subsubsection{Contexts and congruences}
-% Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
-
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We can think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsubsection{Strong barbed congruence}
-A congruence is exactly what we need to make strong barbed bisimilarity consider the environment (i.e.\ the context) of processes.
-
-We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
-
-\begin{lemma}
-  \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \).
-\end{lemma}
-
-\subsubsection{Challenge}
-The objective of this challenge is to prove the following theorem:
-\begin{theorem}
-  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).
-\end{theorem}
-
-%\paragraph{Notes about this challenge}
-%Strong barbed bisimilarity is interesting by itself for this challenge, but the proof that it is an equivalence relation does not require up-to techniques, which are wide spread and interesting.
-%The definition of strong barbed congruence requires introducing contexts, but this is actually nice, since they are so prevalent in the field and we want to cover as many basics as possible.
-%The theorem about the strong barbed congruence requires up-to techniques, but also seems to involve introducing a structural equivalence, which is possibly not desirable.
-%On the other hand, it is a nice application of structural congruence which does not ``ruin'' the semantics.
-
-%% Local Variables:
-%% mode: latex
-%% TeX-master: "main"
-%% End:

### 2023-09-27 challenges/linearity.tex
removed update op from Vasco's style ctx, since it's pointless with our fragment

@@
 All of the names added to a type context must be distinct.
-We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
+Note that
+the order in which information is added to a type context thus does not matter.
+% We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
 
@@
 
-We will also need to update contexts with new type information in a safe way while typing input and output.
-The \emph{context update}, \( \Cupdate{\Gamma}{\hastype{x}{T}} = \Gamma' \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Update-Name]{\hastype{x}{U} \notin \Gamma}{\Cupdate{\Gamma}{\hastype{x}{T}} = \Cadd{\Gamma}{\hastype{x}{T}}}
-  \and
-  \inferrule[Update-Un]{\un{T}}{\Cupdate{(\Cadd{\Gamma}{\hastype{x}{T}})}{\hastype{x}{T}} = (\Cadd{\Gamma}{\hastype{x}{T}})}
-\end{mathpar}
+% We will also need to update contexts with new type information in a safe way while typing input and output.
+% The \emph{context update}, \( \Cupdate{\Gamma}{\hastype{x}{T}} = \Gamma' \), is defined by the following rules:
+% \begin{mathpar}
+%   \inferrule[Update-Name]{\hastype{x}{U} \notin \Gamma}{\Cupdate{\Gamma}{\hastype{x}{T}} = \Cadd{\Gamma}{\hastype{x}{T}}}
+%   \and
+%   \inferrule[Update-Un]{\un{T}}{\Cupdate{(\Cadd{\Gamma}{\hastype{x}{T}})}{\hastype{x}{T}} = (\Cadd{\Gamma}{\hastype{x}{T}})}
+% \end{mathpar}
 
@@
   \and
-  \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{(\Cadd{\Gamma_2}{\hastype{l}{\Tbase}})}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
+  \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{\Cadd{\Gamma_2}{\hastype{l}{\Tbase}}}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
 \end{mathpar}

### 2023-09-28 challenges/coinduction.tex

@@
 
-This challenge formalises a proof that requires coinductive techniques.
-Coinduction is a proof technique for infinite structures, which arise in this context due to systems with behaviors that continue indefinitely.
-Coinduction is the dual of induction: whereas induction is useful for proving properties of least fixed points, coinduction is useful for proving properties of greatest fixed points.
+\marco{Why reasoning is only about infinite processes? Would it be
+  unreasonable to shorten the title to {\em Coinduction and infinite
+    processes}?}
 
-The setting for this challenge is an untyped calculus of communicating systems with replication of processes, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed bisimulation for this calculus is an equivalence relation.
-Which, to reiterate, entails the use of coinduction due to the presence of infinite behavior.
+This challenge is a about the mechanisation of proofs
+% formalises a proof
+that require coinduction.
+% coinductive techniques.
+%
+% Coinduction is a proof technique for infinite structures,
+Coinduction is a technique for defining and proving properties of
+infinite structures,
+%
+which, in the context of the concurrent benchmark, arise
+% in this context
+due to systems with behaviors that continue indefinitely.
+%
+Coinduction is the dual of induction: whereas induction is useful for
+proving properties of least fixed points, coinduction is useful for
+proving properties of greatest fixed points.
+%
+\marco{Should we be more specific here and give a deeper explanation
+  of what coinduction is? If not, then why do we even say that it's
+  the dual of induction?}
+%
 
-\subsubsection{Syntax}
-% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-The syntax differs from \ref{sec:challenge:linearity-beh-types} for including replication and having unary restriction:
-\begin{align*}
-  v,w :=&&& a \\
-  |&&& l \\
-  P,Q :=&&& \Pend \\
-  |&&& \Pout{x}{v}{P} \\
-  |&&& \Pin{x}{l}{P} \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& \Preplicate{P}
-\end{align*}
-% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-% The process \( \Pout{x}{a}{P} \) is an \emph{output}, which can send the value \( v \) via \( x \), then continue as \( P \).
-% The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-% The process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-% The input operator thus binds the variable name \( l \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-% The restriction operator thus binds the name \( x \) in \( P \).
-% Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-% The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-% It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-% Replication makes it possible to express infinite behaviours.
+For this challenge, we adopt a fragment of the untyped $\pi$-calculus
+that includes process replication.
+% The setting for this challenge is an untyped calculus of
+% communicating systems with replication of processes, the syntax and
+% semantics of which are presented below.
+The objective of this challenge is to prove that strong barbed
+bisimilarity
+% for this calculus
+is an equivalence relation. Which, to reiterate, entails the use of
+coinduction due to the presence of infinite behaviour.
 
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
+\paragraph{Syntax.}
+% We assume the existence of some type of \emph{base values}, values
+% of which we will denote by \( a, b, \dots \), the existence of some
+% type of \emph{variable names}, values of which we will denote by
+% \( l, m, \dots \), and the existence of some type of \emph{names},
+% values of which we will denote by \( x, y, \dots \).
+The syntax of the calculus is identical to that of
+Section~\ref{sec:challenge:linearity-beh-types} (challenge on
+linearity), except from restriction, now unary, and the addition of
+replication:
+% for including replication and having unary restriction:
+\begin{displaymath}
+  \begin{array}{rcl}
+    v,w & ::= & a \quad\mid\quad l \\
+    P,Q & ::=& \Pend
+               \quad \mid\quad \Pout{x}{v}{P}
+               \quad \mid\quad \Pin{x}{l}{P}
+               \quad \mid\quad \Ppar{P}{Q}
+               \quad \mid\quad \Pres{x}{}{P}
+               \quad \mid\quad !P
+  \end{array}
+\end{displaymath}
+% \begin{align*}
+%   v,w :=&&& a \\
+%   |&&& l \\
+%   P,Q :=&&& \Pend \\
+%   |&&& \Pout{x}{v}{P} \\
+%   |&&& \Pin{x}{l}{P} \\
+%   |&&& \Ppar{P}{Q} \\
+%   |&&& \Pres{x}{P} \\
+%   |&&& \Preplicate{P}
+% \end{align*}
+% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a
+% variable name \( l \).  The process \( \Pend \) is \emph{inaction}:
+% a process which can do nothing.  The process \( \Pout{x}{a}{P} \) is
+% an \emph{output}, which can send the value \( v \) via \( x \), then
+% continue as \( P \).  The intention is that the value \( v \) must
+% be a base value when it is actually sent, and this will be enforced
+% in the semantics later on.  The process \( \Pin{x}{l}{P} \) is an
+% \emph{input}, which can receive a base value via \( x \), then
+% continue as \( P \) with the received value substituted for the
+% variable name \( l \).  The input operator thus binds the variable
+% name \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
+% \emph{composition} of process \( P \) and process \( Q \).  The two
+% components can proceed independently of each other, or they can
+% interact via shared names.  The process \( \Pres{x}{P} \) is the
+% \emph{restriction} of the name \( x \) to \( P \).  Components in
+% \( P \) can use the name \( x \) to interact with each other, but
+% not with processes outside of the restriction.  The restriction
+% operator thus binds the name \( x \) in \( P \).  Note that the
+% scope of a restriction may not change when processes interact, since
+% it is only possible to send and receive values, and not names.  The
+% process \( \Preplicate{P} \) is the \emph{replication} of the
+% process \( P \).  It can be thought of as the infinite composition
+% \( \Ppar{P}{\Ppar{P}{\cdots}} \).  Replication makes it possible to
+% express infinite behaviours.
+
+% We will use the notation \( \subst{P}{v}{l} \) to denote the process
+% \( P \) with value \( v \) substituted for variable name \( l \).
 
 % % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
+% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
+% written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from
+% \( P \) by a finite number of substitutions of bound variable names.
+% As a convention, we will identify \( \alpha \)-convertible
+% processes.
 
-% Also as a convention, we assume that the bound names and bound variable names occurring in any collection of processes are chosen to be different from the free names and free variable names occurring in those processes and from the names and variable names occurring in any substitutions applied to the processes.
-% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
+% Also as a convention, we assume that the bound names and bound
+% variable names occurring in any collection of processes are chosen
+% to be different from the free names and free variable names
+% occurring in those processes and from the names and variable names
+% occurring in any substitutions applied to the processes.  This is
+% justified because any overlapping names and variable names may be
+% \( \alpha \)-converted such that the assumption is satisfied.
 
-\subsubsection{Semantics}
-The semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which are as follows:
+\paragraph{Semantics.} For this challenge, we give an LTS semantics.
+The semantics of the system describe the actions that the system can
+perform by defining a labelled transition relation on processes.  The
+transitions are labelled by \emph{actions}, the syntax of which are as
+follows:
 \begin{align*}
@@
 \end{align*}
-The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value \( a \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the base value \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
+The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value
+\( a \) via \( x \).  The \emph{input action} \( \Ain{x}{y} \) is
+receiving the base value \( y \) via \( x \).  The \emph{internal
+  action} \( \Atau \) is performing some unobservable action, e.g.\
+internal communication.
 
@@
 \end{theorem}
+\begin{proof}
+  We need to prove that $\sbbisim{}{}$ is reflexive, symmetric, and
+  commutative.
+  \begin{itemize}
+  \item Reflexivity is straightforward. For any $P$, we need to show
+    that $P\sbbisim{}{} P$. Condition from equation~\ref{} follows
+    trivially ($P$ has the same barbs as itself). Similarly, the
+    second condition in equation~\ref{} follows trivially.
+  \item
+  \item
+  \end{itemize}
 
-Unfortunately, strong barbed bisimilarity is not a good notion of process equivalence, since it does not consider the environment of processes.
-For instance, \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since \( \obsout{x} \) is the only observable in both processes and they cannot perform a \( \Atau \)-action, but \( \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}} \) since the left process can perform a \( \Atau \)-action such that \( \obsout{y} \) becomes observable, whereas the right process cannot.
+\end{proof}
+
+Unfortunately, strong barbed bisimilarity is not a good notion of
+process equivalence, since it does not consider the environment of
+processes.  For instance,
+\( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \)
+since \( \obsout{x} \) is the only observable in both processes and
+they cannot perform a \( \Atau \)-action, but
+\(
+\nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}}
+\) since the left process can perform a \( \Atau \)-action such that
+\( \obsout{y} \) becomes observable, whereas the right process cannot.
 

### 2023-09-29 challenges/scope-extrusion.tex
Remove choice/sums from scope extrusion challenge.

@@
 In this section we assume that \emph{names}  can be compared for equality.
-The syntax in this challenge will be split into two levels: processes and summations.
 The syntax of processes is given by:
 \begin{align*}
-  P,Q :=&&& \Ppar{P}{Q} \\
-  |&&& \Pres{x}{P} \\
-  |&&& M
-\end{align*}
-The syntax of summations is given by:
-\begin{align*}
-  M,N :=&&& \Pend \\
+  P,Q := &&& \Pend \\
+  |&&& \Ppar{P}{Q} \\
   |&&& \Pout{x}{y}{P} \\
   |&&& \Pin{x}{y}{P} \\
-  |&&& \Pchoice{M}{N}
+  |&&& \Pres{x}{P} \\
 \end{align*}
-Splitting the syntax into these two layers has the effect of disallowing choices directly between parallel processes.
-Allowing choices between arbitrary processes complicates the theory and is rarely useful.
 
@@
 
-Due to the syntactic levels, we will need to modify the definition of a context slightly by requiring that the \emph{hole} symbol \( \ctxhole \) does not appear directly as the left or right term in a choice.
-For instance, we disallow the context \( \Pchoice{\ctxhole}{\Pin{x}{y}{\Pend}} \), but allow the context \( \Pchoice{\Pout{x}{z}{\ctxhole}}{\Pin{x}{y}{\Pend}} \).
-
 \subsubsection{Reduction semantics}
@@
 \begin{mathpar}
-  \inferrule[Sc-Sum-Assoc]{ }{\scong{\Pchoice{M}{(\Pchoice{N}{L})}}{\Pchoice{(\Pchoice{M}{N})}{L}}}
-  \and
-  \inferrule[Sc-Sum-Comm]{ }{\scong{\Pchoice{M}{N}}{\Pchoice{N}{M}}}
-  \and
-  \inferrule[Sc-Sum-Inact]{ }{\scong{\Pchoice{M}{\Pend}}{M}}
-  \and
   \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{P}{(\Ppar{Q}{R})}}{\Ppar{(\Ppar{P}{Q})}{R}}}
@@
 \begin{mathpar}
-  \inferrule[R-Com]{ }{\reduces{\Ppar{(\Pchoice{\Pout{x}{y}{P}}{M})}{(\Pchoice{\Pin{x}{z}{Q}}{N})}}{\Ppar{P}{\subst{Q}{y}{z}}}}
+  \inferrule[R-Com]{ }{\reduces{\Ppar{\Pout{x}{y}{P}}{\Pin{x}{z}{Q}}}{\Ppar{P}{\subst{Q}{y}{z}}}}
   \and
@@
   \and
-  \inferrule[Sum-L]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{P}{Q}}{\alpha}{P'}}
-  \and
-  \inferrule[Sum-R]{\transition{P}{\alpha}{P'}}{\transition{\Pchoice{Q}{P}}{\alpha}{P'}}
-  \and
   \inferrule[Par-L]{\transition{P}{\alpha}{P'} \\ \boundnames{\alpha} \cap \freenames{Q} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
@@
 \begin{lemma}\label{se-lemma-normalized-reduction}
-  If \( \reduces{P}{Q} \) then there are \( x, y, z, z_1, \dots, z_n, R_1, N_1, R_2, N_2, \) and \( S\) such that
+  If \( \reduces{P}{Q} \) then there are \( x, y, z, z_1, \dots, z_n, R_1, R_2, \) and \( S\) such that
   \begin{align*}
-    &\scong{P}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{(\Pchoice{\Pout{x}{y}{R_1}}{N_1})}{(\Pchoice{\Pin{x}{z}{R_2}}{N_2})})}{S})}}} \\
+    &\scong{P}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{\Pout{x}{y}{R_1}}{\Pin{x}{z}{R_2}})}{S})}}} \\
     &\scong{Q}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{R_1}{\subst{R_2}{y}{z}})}{S})}}}
@@
   \begin{enumerate}
-  \item if \( \transition{Q}{\Aoutf{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{y}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x,y \notin \{ w_1, \dots, w_n \} \).
-  \item if \( \transition{Q}{\Aoutb{x}{z}}{Q'} \) then \( \scong{Q}{\Pres{z}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pout{x}{z}{R}}{M})}{S})}}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x \notin \{ z, w_1, \dots, w_n \} \).
-  \item if \( \transition{Q}{\Ain{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{(\Pchoice{\Pin{x}{z}{R}}{M})}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\subst{R}{y}{z}}{S})}}} \) where \( x \notin \{ w_1, \dots, w_n \} \).
+  \item if \( \transition{Q}{\Aoutf{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\Pout{x}{y}{R}}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x,y \notin \{ w_1, \dots, w_n \} \).
+  \item if \( \transition{Q}{\Aoutb{x}{z}}{Q'} \) then \( \scong{Q}{\Pres{z}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\Pout{x}{z}{R}}{S})}}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x \notin \{ z, w_1, \dots, w_n \} \).
+  \item if \( \transition{Q}{\Ain{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\Pin{x}{z}{R}}{S})}}} \) and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\subst{R}{y}{z}}{S})}}} \) where \( x \notin \{ w_1, \dots, w_n \} \).
   \end{enumerate}
@@
   \begin{equation*}
-    P' = \Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{(\Pchoice{\Pout{x}{y}{R_1}}{N_1})}{(\Pchoice{\Pin{x}{z}{R_2}}{N_2})})}{S})}}
+    P' = \Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{\Pout{x}{y}{R_1}}{\Pin{x}{z}{R_2}})}{S})}}
   \end{equation*}

### 2023-10-03 challenges/scope-extrusion.tex
Change \autoref to \cref to fix duplicate "lemma" text.

@@
 \end{proof}
-\begin{proof}{Of lemma \autoref{se-lemma-normalized-reduction}.}
-  Follows immediately from lemma \autoref{se-lemma-normalized-derivation} and the shape of a normalized derivation.
+\begin{proof}{Of \cref{se-lemma-normalized-reduction}.}
+  Follows immediately from \cref{se-lemma-normalized-derivation} and the shape of a normalized derivation.
 \end{proof}
@@
 The objective of this challenge is to prove the following theorems, which together show the equivalence between the reduction semantics and the transition system semantics up to structural congruence.
-The first of the theorems involves reasoning about scope extrusion more directly than the other, and if time does not permit proving both of the theorems, \autoref{thm:se-trans-implies-red} should be proven first.
+The first of the theorems involves reasoning about scope extrusion more directly than the other, and if time does not permit proving both of the theorems, \cref{thm:se-trans-implies-red} should be proven first.
 \begin{theorem}\label{thm:se-trans-implies-red}
@@
 \begin{proof}
-  If \( \reduces{P}{Q} \), then by lemma \autoref{se-lemma-normalized-reduction}, \( \scong{P}{P'} \) with
+  If \( \reduces{P}{Q} \), then by \cref{se-lemma-normalized-reduction}, \( \scong{P}{P'} \) with
   \begin{equation*}
@@
   \end{equation*}
-  We can easily check that \( \transition{P'}{\tau}{Q'} \) and so by lemma \autoref{se-lemma-harmony-fact}, \( \transition{P}{\tau}{Q'} \).
+  We can easily check that \( \transition{P'}{\tau}{Q'} \) and so by \cref{se-lemma-harmony-fact}, \( \transition{P}{\tau}{Q'} \).
 \end{proof}

### 2023-10-03 challenges/scope-extrusion.tex
Improve formatting of proof sketches.

@@
 \end{lemma}
-\begin{proof}
-  Sketch. First, show the special case when \( P \) can be rewritten to \( Q \) with a single application of an axiom of the structural congruence to some subterm of P.
+\begin{proof}[Sketch]
+  First, show the special case when \( P \) can be rewritten to \( Q \) with a single application of an axiom of the structural congruence to some subterm of P.
   The proof is then by induction on the number of such steps.
@@
 \end{lemma}
-\begin{proof}
-  Sketch.
+\begin{proof}[Sketch]
   To obtain a normalized derivation from an arbitrary derivation we will need to check that rules \TirName{R-Com}, \TirName{R-Par} and \TirName{R-Res} commute with \TirName{R-Struct}, and that two applications of \TirName{R-Struct} can be combined into one.
 \end{proof}
-\begin{proof}{Of \cref{se-lemma-normalized-reduction}.}
+\begin{proof}[Of \cref{se-lemma-normalized-reduction}]
   Follows immediately from \cref{se-lemma-normalized-derivation} and the shape of a normalized derivation.
@@
 \end{theorem}
-\begin{proof}
+\begin{proof}[Sketch]
   The proof is by induction on the inference of \( \transition{P}{\Atau}{Q} \) using the following:

### 2023-10-03 challenges/scope-extrusion.tex
Add a bit more explanation to scope extrusion challenge.

@@
 Note that there is no recursion or replication in the syntax, and thus no infinite behaviors can be expressed.
+This simplifies the theory and is orthogonal to the concept of scope extrusion.
 
@@
 \begin{proof}[Sketch]
-  The proof is by induction on the inference of \( \transition{P}{\Atau}{Q} \) using the following:
+  The proof is by induction on the inference of \( \transition{P}{\Atau}{Q} \) using the following lemmas:
   \begin{enumerate}

### 2023-10-03 challenges/scope-extrusion.tex
Change order of lemmas in scope extrusion.

@@
 \end{proof}
+For the proof of the challenge theorem, we will introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
+The first rule applied is \TirName{R-Com}. The derivation continues with an application of \TirName{R-Par}, followed by zero or more applications of \TirName{R-Res}.
+The last rule is an application of \TirName{R-Struct}.
+\begin{lemma}\label{se-lemma-normalized-derivation}
+  Every reduction has a normalized derivation.
+\end{lemma}
+\begin{proof}[Sketch]
+  To obtain a normalized derivation from an arbitrary derivation we will need to check that rules \TirName{R-Com}, \TirName{R-Par} and \TirName{R-Res} commute with \TirName{R-Struct}, and that two applications of \TirName{R-Struct} can be combined into one.
+\end{proof}
 
@@
 \end{lemma}
-To prove this lemma, we introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
-The first rule applied is \TirName{R-Com}. The derivation continues with an application of \TirName{R-Par}, followed by zero or more applications of \TirName{R-Res}.
-The last rule is an application of \TirName{R-Struct}.
-\begin{lemma}\label{se-lemma-normalized-derivation}
-  Every reduction has a normalized derivation.
-\end{lemma}
-\begin{proof}[Sketch]
-  To obtain a normalized derivation from an arbitrary derivation we will need to check that rules \TirName{R-Com}, \TirName{R-Par} and \TirName{R-Res} commute with \TirName{R-Struct}, and that two applications of \TirName{R-Struct} can be combined into one.
-\end{proof}
-\begin{proof}[Of \cref{se-lemma-normalized-reduction}]
+\begin{proof}
   Follows immediately from \cref{se-lemma-normalized-derivation} and the shape of a normalized derivation.

### 2023-10-03 paper/paper.tex
Split challenges.tex into multiple files.
Add Daniel as an author to the paper
Change Lorenzo's affiliation to Liverpool.

@@
 \and Alceste Scalas \inst{5}\orcidID{0000-0002-1153-6164}
-\and Martin Vassor \inst{4}\orcidID{0000-0002-2057-0495}
-\and Nobuko Yoshida \inst{4}\orcidID{0000-0002-3925-8557}
+\and Martin Vassor \inst{8}\orcidID{0000-0002-2057-0495}
+\and Nobuko Yoshida \inst{8}\orcidID{0000-0002-3925-8557}
+\and Daniel Zackon \inst{9}\orcidID{0000-0000-0000-0000}
 }
@@
 \and Royal Holloway, University of London \email{Francisco.FerreiraRuiz@rhul.ac.uk}
-\and University of Oxford \email{lorenzo.gheri@cs.ox.ac.uk}, \email{martin.vassor@cs.ox.ac.uk}, \email{nobuko.yoshida@cs.ox.ac.uk}
+\and University of Liverpool \email{Lorenzo.Gheri@liverpool.ac.uk}
 \and Technical University of Denmark \email{fkjac@dtu.dk}, \email{alcsc@dtu.dk}
@@
 \and Università di Camerino \email{luca.padovani@unicam.it}
+\and University of Oxford \email{martin.vassor@cs.ox.ac.uk}, \email{nobuko.yoshida@cs.ox.ac.uk}
+\and McGill University \email{daniel.zackon@mail.mcgill.ca}
 }
@@
 \section{Challenges}
-\input{../challenges.tex}
+\input{../challenges/challenges.tex}
 

### 2023-10-04 paper/paper.tex
Update Daniel's ORCID.
Move paper bibliography after appendix to suit LLNCS guidelines.
Add addresses to affiliations.

@@
 \and Nobuko Yoshida \inst{8}\orcidID{0000-0002-3925-8557}
-\and Daniel Zackon \inst{9}\orcidID{0000-0000-0000-0000}
+\and Daniel Zackon \inst{9}\orcidID{0009-0008-6153-2955}
 }
@@
 \institute{
-     IT University of Copenhagen \email{maca@itu.dk}
-\and University of Kent \email{D.Castro-Perez@kent.ac.uk}
-\and Royal Holloway, University of London \email{Francisco.FerreiraRuiz@rhul.ac.uk}
-\and University of Liverpool \email{Lorenzo.Gheri@liverpool.ac.uk}
-\and Technical University of Denmark \email{fkjac@dtu.dk}, \email{alcsc@dtu.dk}
-\and Università degli Studi di Milano \email{momigliano@di.unimi.it}
-\and Università di Camerino \email{luca.padovani@unicam.it}
-\and University of Oxford \email{martin.vassor@cs.ox.ac.uk}, \email{nobuko.yoshida@cs.ox.ac.uk}
-\and McGill University \email{daniel.zackon@mail.mcgill.ca}
+     IT University of Copenhagen, Copenhagen, Denmark \email{maca@itu.dk}
+\and University of Kent, Canterbury, United Kingdom \email{D.Castro-Perez@kent.ac.uk}
+\and Royal Holloway, University of London, Egham, United Kingdom \email{Francisco.FerreiraRuiz@rhul.ac.uk}
+\and University of Liverpool, Liverpool, United Kingdom \email{Lorenzo.Gheri@liverpool.ac.uk}
+\and Technical University of Denmark, Kgs. Lyngby, Denmark \email{fkjac@dtu.dk}, \email{alcsc@dtu.dk}
+\and Università degli Studi di Milano, Milan, Italy \email{momigliano@di.unimi.it}
+\and Università di Camerino, Camerino, Italy \email{luca.padovani@unicam.it}
+\and University of Oxford, Oxford, United Kingdom \email{martin.vassor@cs.ox.ac.uk}, \email{nobuko.yoshida@cs.ox.ac.uk}
+\and McGill University, Montreal, Canada \email{daniel.zackon@mail.mcgill.ca}
 }
@@
 
-\bibliographystyle{splncs04}
-\bibliography{../references}
-
 \appendix
@@
 
+\bibliographystyle{splncs04}
+\bibliography{../references}
+
 \end{document}

### 2023-10-04 paper/paper.tex
small update of the abstract

@@
 \begin{abstract}
-  The POPLMark challenge sparked a flurry of work on machine-checked proofs.
-  However, POPLMark (and its follow-up POPLMark Reloaded) were explicitly limited in scope,
-  with benchmark problems that do not address concurrency.
-For this reason, we propose a new collection of benchmark problems with a focus on the issues encountered when mechanising message-passing concurrency using process calculi.
-  Our benchmark problems concern linear handling of environments, scope extrusion, and coinductive reasoning.
-  Our goal is to clarify, compare and advance the state of the art.
+  The POPLMark challenge sparked a flurry of work on machine-checked
+  proofs. However, POPLMark (and its follow-up POPLMark Reloaded) were
+  purposedly limited in scope, with benchmark problems that do not
+  address concurrency. For this reason, we propose a new collection of
+  benchmark problems with a focus on the issues encountered when
+  mechanising message-passing concurrency using process calculi. Our
+  benchmark problems concern linear handling of environments, scope
+  extrusion, and coinductive reasoning. Our goal is to clarify,
+  compare and advance the state of the art.
 
@@
 \section{Introduction}
+
 The programming language community has increasingly embraced mechanisations since the influential POPLMark challenge~\cite{POPLMark}, but we are still far from the vision of a mechanised proof accompanying every paper in the field.

### 2023-10-04 paper/paper.tex
First paragraph of the introduction. It is not much, but it is honest work.

@@
 
-The programming language community has increasingly embraced mechanisations since the influential POPLMark challenge~\cite{POPLMark}, but we are still far from the vision of a mechanised proof accompanying every paper in the field.
-Since POPLMark, the number of papers with accompanying mechanised proofs has grown steadily, but many mechanisations still rely on ad hoc techniques.
-POPLMark Reloaded~\cite{POPLMarkReloaded} aimed to identify and develop best practices and tool support for proofs using logical relations.
-The authors of both POPLMark and POPLMark Reloaded were careful to note that their benchmarks were only a beginning, and they specifically mention reasoning about concurrency using coinduction and linear environments as points for future work.
-Papers in concurrency theory do often include mechanised proofs (e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also directly discuss formalisation techniques (e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01, DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).
-However, our experience is that choosing techniques and tools remains a challenge with few guidelines.
-We propose a new collection of benchmark problems specifically designed to tackle the issues encountered when mechanising message-passing concurrency using process calculi.
-A goal of our collection of benchmark problems is to push for the development of guidelines and tutorials, and our challenge problems are a simple setting in which to demonstrate and compare techniques.
-Another goal is to strengthen the culture of mechanisation by rallying the community to collaborate on exploring and developing new tools and techniques.
-We have identified three key aspects which cause issues when mechanising concurrency theory: linearity, scope extrusion in name passing systems and infinite behaviour modelled coinductively.
-Each part of our benchmark exercises the three aspects independently to enable comparison and to keep the challenge problems small.
-The three aspects are of course not the only ones that may cause issues in mechanisations, but they are fundamental to the field and will thus appear in most mechanisations of concurrency theory.
-Actual mechanisations will additionally often require more than just one aspect, but we do not address the combination of techniques.
-In the rest of this paper we discuss the design of the challenge and briefly describe the three challenge problems.
+The introduction of the influential POPLMark challenge~\cite{POPLMark}
+spearheaded a shift to including mechanised proofs accompanying
+published papers. POPLMark's abstract starts with an important
+question ``How close are we to a world where every paper on program-
+ming languages is accompanied by an electronic appendix with machine-
+checked proofs?''. Much progress has been done in that respect,
+notably the POPLMark reloaded~\cite{POPLMarkReloaded} challenge aims
+to identify and develop best practices and tool support for proofs
+using logical relations.
+
+The authors of both POPLMark and POPLMark Reloaded
+were careful to note that their benchmarks were only a beginning, and
+they specifically mention reasoning about concurrency using
+coinduction and linear environments as points for future work. Papers
+in concurrency theory do often include mechanised proofs
+(e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
+  DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also
+directly discuss formalisation techniques
+(e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
+  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}). However,
+our experience is that choosing techniques and tools remains a
+challenge with few guidelines. We propose a new collection of
+benchmark problems specifically designed to tackle the issues
+encountered when mechanising message-passing concurrency using process
+calculi. A goal of our collection of benchmark problems is to push for
+the development of guidelines and tutorials, and our challenge
+problems are a simple setting in which to demonstrate and compare
+techniques. Another goal is to strengthen the culture of mechanisation
+by rallying the community to collaborate on exploring and developing
+new tools and techniques. We have identified three key aspects which
+cause issues when mechanising concurrency theory: linearity, scope
+extrusion in name passing systems and infinite behaviour modelled
+coinductively. Each part of our benchmark exercises the three aspects
+independently to enable comparison and to keep the challenge problems
+small. The three aspects are of course not the only ones that may
+cause issues in mechanisations, but they are fundamental to the field
+and will thus appear in most mechanisations of concurrency theory.
+Actual mechanisations will additionally often require more than just
+one aspect, but we do not address the combination of techniques. In
+the rest of this paper we discuss the design of the challenge and
+briefly describe the three challenge problems.
 

### 2023-10-05 paper/paper.tex
Minor fixes in introduction.

@@
 published papers. POPLMark's abstract starts with an important
-question ``How close are we to a world where every paper on program-
+question: ``How close are we to a world where every paper on program-
 ming languages is accompanied by an electronic appendix with machine-
 checked proofs?''. Much progress has been done in that respect,
-notably the POPLMark reloaded~\cite{POPLMarkReloaded} challenge aims
+notably the POPLMark Reloaded~\cite{POPLMarkReloaded} challenge aims
 to identify and develop best practices and tool support for proofs
@@
 extrusion in name passing systems and infinite behaviour modelled
-coinductively. Each part of our benchmark exercises the three aspects
+coinductively. Each part of our benchmark exercises one of the three aspects
 independently to enable comparison and to keep the challenge problems
@@
 and will thus appear in most mechanisations of concurrency theory.
-Actual mechanisations will additionally often require more than just
+Mechanisations of actual research will additionally often require more than just
 one aspect, but we do not address the combination of techniques. In

### 2023-10-05 paper/paper.tex
Add more introduction.

@@
 Mechanisations of actual research will additionally often require more than just
-one aspect, but we do not address the combination of techniques. In
-the rest of this paper we discuss the design of the challenge and
-briefly describe the three challenge problems.
+one aspect, but we do not address the combination of techniques.
+
+Large scale formalisation of concurrency theory is already possible.
+(Paragraphs about related work in this vein)
+
+However, it is our experience that tools and techniques are chosen in a very ad hoc fashion.
+Like previous benchmark authors, we thus seek to answer a number of questions:
+\begin{enumerate}
+\item What is the current state of the art in mechanising process calculi?
+\item What techniques and best practices can be recommended when starting formalisation projects within concurrency theory?
+\item What improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
+\end{enumerate}
+
+Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate and evaluate the many tools and techniques available.
+Our problems are drawn from the basic metatheory of process calculi and include as little syntax and as few definitions as possible to obtain an environment in which to express theorems that require interesting proof techniques.
+The theorems themselves are all well-known results, which suits our aim of exploring mechanisation techniques instead of new theory.
+The small scale of the challenge problems also means that they are also accessible to mechanisation experts who may not be familiar with concurrency theory.
+While the challenge problems themselves are thus not of scientific interest, we expect that the best practices and mechanisations developed by solving the challenge problems will also be useful for future research.
+
+We have begun collecting solutions to our challenge problems on our website.
+In the long term, we expect to use the website for promoting best practices, tutorials and guidelines derived from solutions to our challenge problems.
+The interested reader can find more information about the current status at:
+\begin{center}
+  \url{https://concurrentbenchmark.github.io/}
+\end{center}
+We encourage anyone interested to try the challenge problems using their favorite tools and techniques, and to send us their solutions.
+
+In the rest of this paper we discuss the overall design of the benchmark, describe our three challenge problems, and discuss possible extensions and further work.
 
@@
 We invite anyone interested to send us solutions to the challenge problems and to publish experience reports detailing the advantages and disadvantages of their approaches.
-The interested reader can find more information about the current status of the challenge at:
-\begin{center}
-  \url{https://concurrentbenchmark.github.io/}
-\end{center}
 

### 2023-10-05 paper/paper.tex
Add a bit more strucure to the paper.

@@
 
-\section{Design of the challenge}
+\section{Design of the benchmark}
+
+\subsection{Evaluation criteria}
+
 We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
@@
 In this section we will briefly describe our three challenge problems, referring to the appendix for the full text of each challenge.
+
+\subsection{Linearity}
 Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
@@
 In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
+
+\subsection{Scope extrusion}
 Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing calculus without infinite behaviour, but with non-deterministic choice.
 The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues for some approaches to mechanising binders.
+
+\subsection{Coinduction}
 Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed bisimulation is an equivalence relation \footnote{AM: this seems wrong, see:  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).}for an untyped calculus of communicating systems with (infinite) replication of processes.
@@
 We invite anyone interested to send us solutions to the challenge problems and to publish experience reports detailing the advantages and disadvantages of their approaches.
+
+\section{Going beyond the challenge problems}
+
+\section{Conclusion}
+
+\subsubsection*{Acknowledgements}
 

### 2023-10-05 paper/paper.tex
Add a list of topics for design discussion in the paper.

@@
 \section{Design of the benchmark}
+\subsection{Topics we discussed during the meetings}
+- Linearity type system dual vs. Vasco's?
+- Only linear channels or also unrestricted channels?
+- Pi-calculus or CCS for linearity?
+- Choice in general
+  - Choice can be encoded, so is not necessary
+  - Complicates the theory
+  - Add as ``future work''
+- Choice in scope extrusion
+- Extension categories: other topics and small technical differences
+- Why not unified syntax?
+  - Choice and coinduction does not mix
+- Different notions of bisimulation in the coinduction challenge
+- Linearity: the context is polymorphic
+- Coinduction: why is congruence interesting?
+- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
+- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
+- A goal is to build a culture of formalising things in the community
+- Linearity is generally interesting
+- Encoding techniques are not different from any other field, so there's nothing ``special'' about the formalisations here
+- We don't care about the content of messages in coinduction, only synchronization
+- Why explicitly bind two endpoints in linearity? To make the duality/polarity connection obvious
+- Linearity: Do we care about avoiding deadlocks? No, this is stronger than safety as in well-formedness
+- Asynchrony? Not necessary, this is a it more advanced
+- ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics
 

### 2023-10-05 paper/paper.tex
Add another note and fix a whitespace error.

@@
 published papers. POPLMark's abstract starts with an important
-question: ``How close are we to a world where every paper on program-
-ming languages is accompanied by an electronic appendix with machine-
+question: ``How close are we to a world where every paper on programming languages is accompanied by an electronic appendix with machine-
 checked proofs?''. Much progress has been done in that respect,
@@
 - Asynchrony? Not necessary, this is a it more advanced
+- Why do we keep the restrictions around in the LTS semantics in scope extrusion?
 - ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics

### 2023-10-06 challenges/linearity.tex
fixing the statement of the subst lemma. Not sure about the proof ...
a couple of metanotes

@@
 The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
+
 \begin{align*}
@@
         \\
-  |&&& \Cadd{\Gamma}{\hastype{l}{S}}
+  |&&& \Cadd{\Gamma}{\hastype{l}{\TBase}} %% changed, AM
 \end{align*}
@@
 
-\paragraph{Splitting and updating contexts}
+\paragraph{Splitting contexts}
 Our type system will maintain two invariants:
@@
 \end{lemma}
-
+\begin{metanote}
+  AM: the statement of substitution was wrong: now it is consistent with the op sem, but I need to check the (I assume now simpler) proof.
+Note that $T$ must be \textbf{base} and \( \un{\Gamma_i} \) must also hold.
+\end{metanote}
 \begin{lemma}
-  If \( \types{\Gamma_1}{\hastype{v}{T}} \) and \( \types{\Cadd{\Gamma_2}{\hastype{x}{T}}}{P} \) and \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then \( \types{\Gamma}{\subst{v}{x}{P}} \).
+  If \( \types{\Gamma_1}{\hastype{v}{T}} \) and
+  \( \types{\Cadd{\Gamma_2}{\hastype{l}{T}}}{P} \) and
+  \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then
+  \( \types{\Gamma}{\subst{P}{l}{v}} \).
 \end{lemma}

### 2023-10-09 paper/paper.tex
Move discussion after challenge problem description in the paper.

@@
 
-In the rest of this paper we discuss the overall design of the benchmark, describe our three challenge problems, and discuss possible extensions and further work.
+In the next section we will describe our three challenge problems, before discussing our design choices in \cref{sec:design-discussion}.
+We will then discuss possible extensions and further topics in \cref{sec:going-beyond} before concluding in \cref{sec:conclusion}.
+The full challenge problems, including proofs, are available in \cref{app:challenges}.
 
-\section{Design of the benchmark}
+\section{The challenge problems}
+In this section we will briefly describe our three challenge problems, referring to the appendix for the full text of each challenge.
+
+\subsection{Linearity}
+Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
+Session type systems require linear handling of typing contexts, and the proof involves reasoning about the linearity of context splitting operations.
+In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
+
+\subsection{Scope extrusion}
+Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing calculus without infinite behaviour, but with non-deterministic choice.
+The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues for some approaches to mechanising binders.
+
+\subsection{Coinduction}
+Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed bisimulation is an equivalence relation \footnote{AM: this seems wrong, see:  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).}for an untyped calculus of communicating systems with (infinite) replication of processes.
+The proof involves reasoning coinductively about the infinite behaviours of the replication operator.
+We invite anyone interested to send us solutions to the challenge problems and to publish experience reports detailing the advantages and disadvantages of their approaches.
+
+\section{Design of the benchmark}\label{sec:design-discussion}
 \subsection{Topics we discussed during the meetings}
@@
 \subsection{Evaluation criteria}
-
 We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
@@
 
-\section{The challenge problems}
-In this section we will briefly describe our three challenge problems, referring to the appendix for the full text of each challenge.
+\section{Going beyond the challenge problems}\label{sec:going-beyond}
 
-\subsection{Linearity}
-Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
-Session type systems require linear handling of typing contexts, and the proof involves reasoning about the linearity of context splitting operations.
-In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
-
-\subsection{Scope extrusion}
-Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing calculus without infinite behaviour, but with non-deterministic choice.
-The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues for some approaches to mechanising binders.
-
-\subsection{Coinduction}
-Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed bisimulation is an equivalence relation \footnote{AM: this seems wrong, see:  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).}for an untyped calculus of communicating systems with (infinite) replication of processes.
-The proof involves reasoning coinductively about the infinite behaviours of the replication operator.
-We invite anyone interested to send us solutions to the challenge problems and to publish experience reports detailing the advantages and disadvantages of their approaches.
-
-\section{Going beyond the challenge problems}
-
-\section{Conclusion}
+\section{Conclusion}\label{sec:conclusion}
 
@@
 \appendix
-\section{Challenges}
+\section{Challenges}\label{app:challenges}
 \input{../challenges/challenges.tex}

### 2023-10-09 paper/paper.tex
Francisco: a pass over the introduction

@@
 The introduction of the influential POPLMark challenge~\cite{POPLMark}
-spearheaded a shift to including mechanised proofs accompanying
-published papers. POPLMark's abstract starts with an important
-question: ``How close are we to a world where every paper on programming languages is accompanied by an electronic appendix with machine-
-checked proofs?''. Much progress has been done in that respect,
-notably the POPLMark Reloaded~\cite{POPLMarkReloaded} challenge aims
-to identify and develop best practices and tool support for proofs
-using logical relations.
+spearheaded a shift to publications that include mechanised proofs.
+POPLMark's abstract starts with an important question: ``How close are
+we to a world where every paper on programming languages is
+accompanied by an electronic appendix with machine- checked proofs?''.
+Much progress has been done in that respect, proof assistants,
+libraries, and best practices evolved from this challenge.
+Furthermore, extensions such as the POPLMark
+Reloaded~\cite{POPLMarkReloaded} challenge aim to identify and develop
+best practices and tool support for proofs using logical relations.
 
-The authors of both POPLMark and POPLMark Reloaded
-were careful to note that their benchmarks were only a beginning, and
-they specifically mention reasoning about concurrency using
-coinduction and linear environments as points for future work. Papers
-in concurrency theory do often include mechanised proofs
+The authors of both POPLMark and POPLMark Reloaded both explicitly
+note that their benchmarks were only a beginning. Furthermore, they
+specifically mention reasoning about concurrency using coinduction and
+linear environments as points for future work. While papers in
+concurrency theory do often include mechanised proofs
 (e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
@@
 (e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
-  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}). However,
-our experience is that choosing techniques and tools remains a
-challenge with few guidelines. We propose a new collection of
-benchmark problems specifically designed to tackle the issues
-encountered when mechanising message-passing concurrency using process
-calculi. A goal of our collection of benchmark problems is to push for
-the development of guidelines and tutorials, and our challenge
-problems are a simple setting in which to demonstrate and compare
-techniques. Another goal is to strengthen the culture of mechanisation
-by rallying the community to collaborate on exploring and developing
-new tools and techniques. We have identified three key aspects which
-cause issues when mechanising concurrency theory: linearity, scope
-extrusion in name passing systems and infinite behaviour modelled
-coinductively. Each part of our benchmark exercises one of the three aspects
-independently to enable comparison and to keep the challenge problems
-small. The three aspects are of course not the only ones that may
-cause issues in mechanisations, but they are fundamental to the field
-and will thus appear in most mechanisations of concurrency theory.
-Mechanisations of actual research will additionally often require more than just
-one aspect, but we do not address the combination of techniques.
+  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}), our
+experience is that choosing techniques and tools remains a challenge
+with few guidelines. Thus, in order to further increase the reach of
+POPLMark challenge, we propose a new collection of benchmark problems
+specifically designed to tackle the issues encountered when
+mechanising message-passing concurrency using process calculi.
+
+A goal of our collection of benchmark problems is to push for the
+development of guidelines and tutorials, and our challenge problems
+are a simple setting in which to demonstrate and compare techniques,
+libraries, and proof assistant features. Another goal is to strengthen
+the culture of mechanisation by rallying the community to collaborate
+on exploring and developing new tools and techniques. We have
+identified three key aspects which cause issues when mechanising
+concurrency theory: linearity, scope extrusion in name passing systems
+and infinite behaviour modelled coinductively. Each part of our
+benchmark exercises one of the three aspects independently to enable
+comparison and to keep the challenge problems small. The three aspects
+are of course not the only ones that may cause issues in
+mechanisations, but they are fundamental to the field and will thus
+appear in most mechanisations of concurrency theory. Mechanisations of
+actual research will additionally often require more than just one
+aspect, but we do not address the combination of techniques.
 
 Large scale formalisation of concurrency theory is already possible.
-(Paragraphs about related work in this vein)
-
-However, it is our experience that tools and techniques are chosen in a very ad hoc fashion.
-Like previous benchmark authors, we thus seek to answer a number of questions:
+%(Paragraphs about related work in this vein)
+However, it is our experience that tools and techniques are chosen in
+a very ad hoc fashion. Moreover, it is our opinion that the community
+will benefit from a set of solutions exploring different ways of
+mechanising typical problems in the field. Like previous benchmark
+authors, we thus seek to answer a number of questions:
 \begin{enumerate}
@@
 
-Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate and evaluate the many tools and techniques available.
+Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate, evaluate, and compare the many tools and techniques available.
 Our problems are drawn from the basic metatheory of process calculi and include as little syntax and as few definitions as possible to obtain an environment in which to express theorems that require interesting proof techniques.

### 2023-10-09 paper/paper.tex
Fix outdated description of coinduction challenge.

@@
 \subsection{Coinduction}
-Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed bisimulation is an equivalence relation \footnote{AM: this seems wrong, see:  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).}for an untyped calculus of communicating systems with (infinite) replication of processes.
+Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed congruence acts like a congruence for an untyped calculus of communicating systems with (infinite) replication of processes.
 The proof involves reasoning coinductively about the infinite behaviours of the replication operator.

### 2023-10-09 paper/paper.tex
Sort discussion points by challenge.

@@
 \documentclass[runningheads]{llncs}
+\synctex=1
 \usepackage[utf8]{inputenc}
@@
 \renewcommand\UrlFont{\color{blue}\rmfamily}
+
+\usepackage[status=draft]{fixme}
+\fxsetup{theme=color}
 
@@
 \begin{abstract}
-  The POPLMark challenge sparked a flurry of work on machine-checked
-  proofs. However, POPLMark (and its follow-up POPLMark Reloaded) were
-  purposedly limited in scope, with benchmark problems that do not
-  address concurrency. For this reason, we propose a new collection of
-  benchmark problems with a focus on the issues encountered when
-  mechanising message-passing concurrency using process calculi. Our
-  benchmark problems concern linear handling of environments, scope
-  extrusion, and coinductive reasoning. Our goal is to clarify,
-  compare and advance the state of the art.
+  The POPLMark challenge and its follow-up POPLMark Reloaded sparked a flurry of
+  work on machine-checked proofs, and fostered the adoption of proof
+  mechanisation in programming language research.  However, both POPLMark and
+  POPLMark Reloaded were purposely limited in scope, with benchmark problems
+  that do not address concurrency-related issues.
+  %
+  For this reason, we propose a new collection of benchmark problems with a
+  focus on the challenges and issues that typically arise when mechanising
+  message-passing concurrency using process calculi.  Our benchmark problems
+  address three key topics: linearity, scope extrusion, and coinductive
+  reasoning.  The goal of this new challenge is to clarify, compare, and advance
+  the state of the art, fostering the adoption of proof mechanisation in future
+  research on message-passing concurrency.
 
@@
 
-The introduction of the influential POPLMark challenge~\cite{POPLMark}
-spearheaded a shift to publications that include mechanised proofs.
-POPLMark's abstract starts with an important question: ``How close are
-we to a world where every paper on programming languages is
-accompanied by an electronic appendix with machine- checked proofs?''.
-Much progress has been done in that respect, proof assistants,
-libraries, and best practices evolved from this challenge.
-Furthermore, extensions such as the POPLMark
-Reloaded~\cite{POPLMarkReloaded} challenge aim to identify and develop
+The abstract of the influential POPLMark challenge~\cite{POPLMark} starts with
+an important question: \emph{``How close are we to a world where every paper on
+programming languages is accompanied by an electronic appendix with
+machine-checked proofs?''} The introduction of the POPLMark
+challenge spearheaded a shift towards publications that include
+mechanised proofs, fostering the progress of proof assistants,
+libraries, and best practices.  Later on, the follow-up POPLMark
+Reloaded challenge~\cite{POPLMarkReloaded} has encouraged the development of
 best practices and tool support for proofs using logical relations.
 
-The authors of both POPLMark and POPLMark Reloaded both explicitly
+The authors of both POPLMark and POPLMark Reloaded explicitly
 note that their benchmarks were only a beginning.  Furthermore, they
 specifically mention reasoning about concurrency using coinduction and
-linear environments as points for future work. While papers in
-concurrency theory do often include mechanised proofs
+linear environments as points for future work.  In this spirit, in order to
+further increase the reach of the POPLMark challenges, we propose a new
+collection of benchmark problems specifically designed to address typical issues
+that arise when mechanising message-passing concurrency using process calculi.
+
+Large scale mechanisation of concurrency theory is already possible,
+\fxwarning{Citations?}
+%(Paragraphs about related work in this vein)
+and several papers in the realm of process calculi do include mechanised proofs
 (e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
-  DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also
-directly discuss formalisation techniques
+  DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or
+discuss proof mechanisation techniques
 (e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
-  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}), our
-experience is that choosing techniques and tools remains a challenge
-with few guidelines. Thus, in order to further increase the reach of
-POPLMark challenge, we propose a new collection of benchmark problems
-specifically designed to tackle the issues encountered when
-mechanising message-passing concurrency using process calculi.
+  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).  However, our
+experience is that choosing the appropriate proof mechanisation techniques and
+tools remains a significant challenge with few guidelines.  This often leads
+researchers towards trial-and-error, attempting sub-optimal and ad-hoc
+solutions; this increases the overall effort required by proof mechanisation,
+and hampers the reuse of techniques and results.
 
-A goal of our collection of benchmark problems is to push for the
-development of guidelines and tutorials, and our challenge problems
-are a simple setting in which to demonstrate and compare techniques,
-libraries, and proof assistant features. Another goal is to strengthen
-the culture of mechanisation by rallying the community to collaborate
-on exploring and developing new tools and techniques. We have
-identified three key aspects which cause issues when mechanising
-concurrency theory: linearity, scope extrusion in name passing systems
-and infinite behaviour modelled coinductively. Each part of our
-benchmark exercises one of the three aspects independently to enable
-comparison and to keep the challenge problems small. The three aspects
-are of course not the only ones that may cause issues in
-mechanisations, but they are fundamental to the field and will thus
-appear in most mechanisations of concurrency theory. Mechanisations of
-actual research will additionally often require more than just one
-aspect, but we do not address the combination of techniques.
-
-Large scale formalisation of concurrency theory is already possible.
-%(Paragraphs about related work in this vein)
-However, it is our experience that tools and techniques are chosen in
-a very ad hoc fashion. Moreover, it is our opinion that the community
-will benefit from a set of solutions exploring different ways of
-mechanising typical problems in the field. Like previous benchmark
-authors, we thus seek to answer a number of questions:
+Therefore, it is our opinion that the research community in message-passing
+concurrency (and more specifically, in process calculi) will benefit from a set
+of solutions exploring different ways of mechanising typical problems in the
+field. Like the authors of the previous POPLMark challenges, we thus seek to
+answer a number of questions:
 \begin{enumerate}
@@
 
-Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate, evaluate, and compare the many tools and techniques available.
-Our problems are drawn from the basic metatheory of process calculi and include as little syntax and as few definitions as possible to obtain an environment in which to express theorems that require interesting proof techniques.
-The theorems themselves are all well-known results, which suits our aim of exploring mechanisation techniques instead of new theory.
-The small scale of the challenge problems also means that they are also accessible to mechanisation experts who may not be familiar with concurrency theory.
-While the challenge problems themselves are thus not of scientific interest, we expect that the best practices and mechanisations developed by solving the challenge problems will also be useful for future research.
+\subsection*{Benchmark design goals and considerations}%
 
-We have begun collecting solutions to our challenge problems on our website.
-In the long term, we expect to use the website for promoting best practices, tutorials and guidelines derived from solutions to our challenge problems.
-The interested reader can find more information about the current status at:
+We have identified three key aspects which typically cause difficulties when
+mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion} in
+name-passing systems, and \emph{coinductive reasoning} for infinite behaviours.
+Our main goals are:
+
+\begin{enumerate}
+\item Enable the comparison of different proof mechanisation approaches, making
+  the challenges accessible to mechanisation experts who may not be familiar
+  with concurrency theory. To this end, we strive to keep the challenge problems
+  orthogonal: our collection of benchmark problems is designed to exercise one
+  of the aforementioned three aspects independently from the others, and include
+  as little syntax and as few definitions as possible.
+
+\item Encourage the development of guidelines and tutorials demonstrating and
+  comparing the many available proof mechanisation techniques, libraries, and
+  proof assistant features.  Therefore, our challenge problems strive for
+  minimality, and adopt a simple and uniform notation.
+
+\item Prioritise the exploration of mechanisation techniques that may be useful
+  and reusable for future research. To this end, our problems are drawn from the
+  basic metatheory of process calculi (instead of requiring the development of
+  new theory), and focus on well-known theorems and results that involve
+  interesting proof techniques.
+\end{enumerate}
+
+We also see the goals above as a stepping stone towards strengthening the
+culture of mechanisation, by rallying the community to collaborate on exploring
+and developing new tools and techniques.
+
+The three aspects we selected (linearity, scope extrusion, coinduction) are of
+course not the only ones that may cause issues in mechanisations --- but they
+are fundamental to the field of concurrency theory, and they emerge in most
+mechanisations.  Moreover, the mechanisations of novel research results often
+requires addressing more than just one of these aspects at once --- and we see
+the combination of techniques as a next step beyond the scope of this challenge
+(we discuss the topic in \cref{sec:going-beyond}).
+
+% Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate, evaluate, and compare the many tools and techniques available.
+% While the challenge problems themselves are thus not of scientific interest, we expect that the best practices and mechanisations developed by solving the challenge problems will also be useful for future research.
+
+We have begun collecting solutions to our challenge problems on our website:
+%
 \begin{center}
@@
 \end{center}
-We encourage anyone interested to try the challenge problems using their favorite tools and techniques, and to send us their solutions.
+%
+In the longer term, we expect to use the website for promoting best practices,
+tutorials and guidelines derived from solutions to our challenge problems.
+%The interested reader can find more information about the current status at:
+We encourage anyone interested to try the challenge problems using their
+favourite tools and techniques, and to send us their solutions.
 
-In the next section we will describe our three challenge problems, before discussing our design choices in \cref{sec:design-discussion}.
-We will then discuss possible extensions and further topics in \cref{sec:going-beyond} before concluding in \cref{sec:conclusion}.
+\paragraph{Outline of this paper.}
+In \cref{sec:challenge-problems} we describe in more detail the three key
+aspects of our challenge problems.  We discuss our design choices in
+\cref{sec:design-discussion}, and possible extensions and further topics in
+\cref{sec:going-beyond}.  We conclude in \cref{sec:conclusion}.
 The full challenge problems, including proofs, are available in \cref{app:challenges}.
 
-\section{The challenge problems}
+\section{The challenge problems}\label{sec:challenge-problems}
 In this section we will briefly describe our three challenge problems, referring to the appendix for the full text of each challenge.
@@
 \section{Design of the benchmark}\label{sec:design-discussion}
-\subsection{Topics we discussed during the meetings}
-- Linearity type system dual vs. Vasco's?
-- Only linear channels or also unrestricted channels?
-- Pi-calculus or CCS for linearity?
+\subsection{General}
+- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
 - Choice in general
@@
   - Add as ``future work''
-- Choice in scope extrusion
 - Extension categories: other topics and small technical differences
@@
   - Choice and coinduction does not mix
-- Different notions of bisimulation in the coinduction challenge
+- A goal is to build a culture of formalising things in the community
+- Encoding techniques are not different from any other field, so there's nothing ``special'' about the formalisations here
+
+\subsection{Linearity}
+- Linearity type system dual vs. Vasco's?
+- Only linear channels or also unrestricted channels?
+- Pi-calculus or CCS for linearity?
 - Linearity: the context is polymorphic
-- Coinduction: why is congruence interesting?
-- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
-- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
-- A goal is to build a culture of formalising things in the community
 - Linearity is generally interesting
-- Encoding techniques are not different from any other field, so there's nothing ``special'' about the formalisations here
-- We don't care about the content of messages in coinduction, only synchronization
 - Why explicitly bind two endpoints in linearity? To make the duality/polarity connection obvious
@@
 - Asynchrony? Not necessary, this is a it more advanced
+
+\subsection{Scope extrusion}
+- Choice in scope extrusion
+- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
 - Why do we keep the restrictions around in the LTS semantics in scope extrusion?
+
+\subsection{Coinduction}
+- Different notions of bisimulation in the coinduction challenge
+- Coinduction: why is congruence interesting?
+- We don't care about the content of messages in coinduction, only synchronization
 - ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics
@@
 
+As with the challenges that came before, this one is intended to be
+useful on its own, but not to cover every problem that may be faced
+when mechanising systems based on message passing calculi. The current
+challenge can be extended in two dimensions: the first, about taking
+the existing challenges further. And the second, about devising
+challenges not covered by the current benchmarks.
+
+The proposed challenges try to minimise the overlap of the techniques
+they exercise, however, most systems will combine some or all aspects
+of the challenges. An interesting second round of formalisations could
+entail combining the features of the challenges. In the following
+table we consider the following three additions: First the addition of
+choice to the type level and the process level. Second, the addition
+of recursion (and recursion types for typed systems). And finally, the
+addition of channel delegation to those calculi that do not already
+support it.
+
+The following table shows which extension should be suitable (that we
+mark with $+$) for each challenge:
+
+\vspace{.5em}
+
+\begin{center}\small
+  \begin{tabular}{|r|c|c|c|}
+    \hline
+    & Linearity & \shortstack{Scope \\ extrusion} & Coinduction \\
+    \hline
+    Choice & $+$ & $+$  & $+$ \\
+    \hline
+    \shortstack{Recursive \\ types} & $+$  & $+$ & \\
+    \hline
+    Delegation & $+$ & & $+$ \\
+    \hline
+  \end{tabular}
+\end{center}
+
+\vspace{.5em}
+
+The other dimension to extend these calculi corresponds to systems
+that are not covered by our challenges. Extending in this dimension
+means proposing new challenges to cover for different features of
+message passing calculi, or to cover different features of mechanised
+proof.
+
+Some interesting aspects of message passing calculi to explore in
+further challenges could be: Multiparty session types, and
+choreographies because their definitions and meta-theory have elements
+that we do not cover here, however the elements that we do cover will
+appear in these problems. Also, while we explore a form of barbed
+bisimilarity one could easily design a challenge exploring different
+notions of bisimilarity (barbed, or weak, etc.), and trace
+equivalence.
+
+On the side of proof assistants, the current challenge proposes
+mechanising the proofs of theorems. An interesting venue to explore
+would be to take advantage of other aspects of proof assistants. For
+example, proof assistants are often able to produce certified code (by
+using code extraction, or compiling and running their definitions).
+Additionally, while having effective automation in our current
+challenge is possible, a challenge could be designed with the explicit
+objective of automating aspects of the proofs. And finally, a
+challenge could propose the integration with other formal reasoning
+tools, namely model checkers. These tools are in common use in the
+field. However, there combining automated proofs with proof assistants
+could be a great path for attempting larger proofs with less effort.
+
+Ultimately, the current challenge can be extended in several
+worthwhile directions, and we look forward to a future when they
+indeed are extended. Moreover, we see the current challenge as setting
+the foundation for those future extensions. It is our expectation that
+the solutions of those hypothetical challenges will build on solutions
+to this challenge.
+
 \section{Conclusion}\label{sec:conclusion}

### 2023-10-10 challenges/coinduction.tex
Fix outdated description of coinduction challenge.
project proposal deadline tomorrow at 12, I will be full time on this from Wednesday afternoon

@@
 
-\paragraph{Semantics.} For this challenge, we give an LTS semantics.
-The semantics of the system describe the actions that the system can
+\paragraph{Semantics.} We give an LTS semantics for this challenge. As
+an alternative, it is also possible to adopt a reduction semantics
+which features less rules but needs a structural congruence. The
+semantics of the system describes the actions that the system can
 perform by defining a labelled transition relation on processes.  The
@@
 follows:
+% \begin{align*}
+%   \alpha := &&& \Aoutf{x}{a} \\
+%   |&&& \Ain{x}{a} \\
+%   |&&& \Atau
+% \end{align*}
 \begin{align*}
-  \alpha := &&& \Aoutf{x}{a} \\
-  |&&& \Ain{x}{a} \\
-  |&&& \Atau
+  \alpha\ :=\  \Aoutf{x}{a} \ \mid\ \Ain{x}{a} \ \mid\ \Atau
 \end{align*}
@@
 \begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}
+  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu}\label{eq:bisim1} \\
+  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}\label{eq:bisim2}
 \end{gather}
@@
 \begin{proof}
-  We need to prove that $\sbbisim{}{}$ is reflexive, symmetric, and
-  commutative.
+  % We need to prove that $\sbbisim{}{}$ is reflexive, symmetric, and
+  % commutative.
+  We prove the three properties separately: 
   \begin{itemize}
-  \item Reflexivity is straightforward. For any $P$, we need to show
-    that $P\sbbisim{}{} P$. Condition from equation~\ref{} follows
-    trivially ($P$ has the same barbs as itself). Similarly, the
-    second condition in equation~\ref{} follows trivially.
-  \item
-  \item
+  \item Reflexivity is straightforward: for any $P$, we need to show
+    that $P\sbbisim{}{} P$. Both conditions in
+    equations~\ref{eq:bisim1} and \ref{eq:bisim2} follow trivially.
+  \item Symmetry is similar to the previous case: for both
+    $P\sbbisim{}{} Q$ and $Q\sbbisim{}{} P$ the same conditions apply.
+  \item For transitivity, we need to prove that if $P\sbbisim{}{} Q$
+    and $Q\sbbisim{}{} R$ then $P\sbbisim{}{} R$. If
+    $\observable{P}{\mu}$ then, by $P\sbbisim{}{} Q$,
+    $\observable{Q}{\mu}$. Hence, by $Q\sbbisim{}{} R$,
+    $\observable{R}{\mu}$. Moreover, if $\transition{P}{\Atau}{P'}$
+    there exists $Q'$ such that $\transition{Q}{\Atau}{Q'}$ and
+    $P'\sbbisim{}{}Q'$. Also, $\transition{R}{\Atau}{R'}$ with
+    $Q'\sbbisim{}{}R'$. Coinductively, $P'\sbbisim{}{}R'$, hence
+    $P\sbbisim{}{}R$.
   \end{itemize}

### 2023-10-10 paper/paper.tex

--- 2023-10-09 paper/paper.tex
@@
 
+As with the challenges that came before, this challenge is intended to
+have useful solutions, but not to cover every problem that may be
+faced when mechanised systems based on message passing calculi. Going
+beyond the current challenges could be in one of two dimensions: the
+first dimension about making harder challenges starting from the
+current ones, and a second one about devising challenges not covered
+by the current benchmarks.
+
+The proposed challenges are orthogonal, however, in general systems
+will combine some or all aspects of the challenges. An interesting
+second round of formalisations could entail combining the features of
+the challenges. In the following table we consider the following three
+additions: First the addition of choice to the type level and the
+process level. Second, the addition of recursion (and recursion types
+for typed systems). And finally, the addition of channel delegation
+for those calculi that do not already support it.
+
+The following table shows which extension should be suitable (denoted
+by a $+$) for each challenge:
+
+\vspace{.5em}
+
+\begin{tabular}{|r|c|c|c|}
+  \hline
+  & Linearity & \shortstack{Scope \\ extrusion} & Coinduction \\
+  \hline
+  Choice & $+$ & $+$  & $+$ \\
+  \hline
+  \shortstack{Recursive \\ types} & $+$  & $+$ & \\
+  \hline
+  Delegation & $+$ & & $+$ \\
+  \hline
+\end{tabular}
+
+\vspace{.5em}
+
+Note that these additions, make the challenges less orthogonal by
+combining some of their features. This is to be expected, in the sense
+that the proposed challenges should form the building blocks of more
+complex systems.
+
+The other dimension to extend these calculi corresponds to systems
+that are not covered by our challenges\dots
+% Multiparty session types, and choreographies.
+
+% different notions of bisimilarity (barbed, or weak, etc.), trace equivalence.
+
+% Exercising different features of proof assistants like: code
+% extraction and/or code automation.
+
+% Interaction between formalised proofs and model checkers.
+
+% And any other
+
 \section{Conclusion}\label{sec:conclusion}

### 2023-10-10 challenges/linearity.tex
fixing a latex typo
More importantly:

by the dual typing rules, all types in Γ are base, so we could just
collect the variables, w/o the types. Similarly we just need to check
that v is well-formed. In fact, we may not need Γ at all.
- separating sr from safety
- preparing the fule to insert dual rules

@@
 
-The syntax is
-given by the grammar:
-\begin{metanote}
-AM:  Should we save some space like this?
-\end{metanote}
+The syntax is given by the grammar:
 % \begin{align*}
@@
 We describe the actions that the system can perform using a small step operational semantics.
-To simplify the definition of the reduction relation, we will first factor out a structural congruence relation on processes, and for this we need a few definitions.
-
-\subsubsection{Contexts and congruences}
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We will think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
+To simplify the definition of the reduction relation, we will  factor out a \emph{structural congruence} relation on processes, namely  ithe smallest congruence relation that satisfies the following axioms:
 \begin{mathpar}
@@
 \begin{mathpar}
-  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{v}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{v}{l}}{R}})}}}
+  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{a}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{a}{l}}{R}})}}}
   \and
@@
 
+\begin{metanote}
+  AM: the above paragraph is repeated in the coinduction challenge and should be removed there
+\end{metanote}
 \subsubsection{Session types}
@@
 To alleviate this issue, we introduce a \emph{session type system}, which will detect processes that are not well-formed.
-To this end, we need to first formally define what we mean by well-formed.
-We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
-A process is then \emph{well-formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
 
-Note that well-formed processes do not necessarily reduce, since e.g.\ the process
-\begin{equation*}
-  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
-\end{equation*}
-is well-formed, but has no reduction.
+\begin{metanote}
+  AM: I've moved the discussion of well-formedness must go after the
+  type system, since we want to separate the preservation vs
+  safety challemges
+\end{metanote}
 \paragraph{Syntax}
@@
         \\
-  |&&& \Cadd{\Gamma}{\hastype{l}{\TBase}} %% changed, AM
+  |&&& \Cadd{\Gamma}{\hastype{l}{\Tbase}} %% changed, AM
 \end{align*}
@@
 
-\paragraph{Splitting contexts}
+\subsubsection{Typing rules}
+% \paragraph{Splitting contexts}
 Our type system will maintain two invariants:
@@
 The second invariant is maintained by requiring duality when typing restrictions.
+
+\begin{metanote}
+  AM: put here the dual rules
+\end{metanote}
 
@@
 
-\subsubsection{Typing rules}
 We have three typing judgments: one for values, one for names, and one for processes.
@@
 \subsubsection{Challenge}
-The objective of this challenge is to prove the following theorems:
-\begin{theorem}
+\begin{metanote}
+AM:  I'm separating sr from safety, or whatever we call the latter
+\end{metanote}
+The objective of this challenge is to prove type preservation for our calculus with session types. We start with:
+\begin{theorem}[Subject reduction]
   If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
 \end{theorem}
-
-\begin{theorem}
-  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
-\end{theorem}
-
-\begin{corollary}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
-\end{corollary}
 
@@
 \end{lemma}
-\begin{metanote}
-  AM: the statement of substitution was wrong: now it is consistent with the op sem, but I need to check the (I assume now simpler) proof.
+% \begin{metanote}
+%   AM: the statement of substitution was wrong: now it is consistent with the op sem, but I need to check the (I assume now simpler) proof.
 Note that $T$ must be \textbf{base} and \( \un{\Gamma_i} \) must also hold.
-\end{metanote}
+% \end{metanote}
 \begin{lemma}
@@
 \end{lemma}
+
+To formulate safety, we need to formally define what we mean by well-formed.
+We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
+A process is then \emph{well-formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
+
+Note that well-formed processes do not necessarily reduce, since e.g.\ the process
+\begin{equation*}
+  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
+\end{equation*}
+is well-formed, but has no reduction.
+
+\begin{theorem}[Safety]
+  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
+\end{theorem}
+
+\begin{corollary}[Type preservation]
+  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
+\end{corollary}

### 2023-10-10 challenges/linearity.tex
dual system streamlined, most sketch of proofs added, but do not trust me

@@
 We describe the actions that the system can perform using a small step operational semantics.
-To simplify the definition of the reduction relation, we will  factor out a \emph{structural congruence} relation on processes, namely  ithe smallest congruence relation that satisfies the following axioms:
+To simplify the definition of the reduction relation, we will  factor out a \emph{structural congruence} relation on processes, namely  the smallest congruence relation that satisfies the following axioms:
 \begin{mathpar}
@@
   type system, since we want to separate the preservation vs
-  safety challemges
+  safety challenges
 \end{metanote}
@@
 The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
-
+\begin{metanote}
+  AM: take the syntax from dual file appst.tex and adapt
+\end{metanote}
 \begin{align*}
@@
 \begin{metanote}
-  AM: put here the dual rules
+  AM: put here the dual rules, remove splitting
 \end{metanote}

### 2023-10-10 paper/paper.tex
Add synctex support
Revise abstract

@@
 \begin{abstract}
-  The POPLMark challenge sparked a flurry of work on machine-checked
-  proofs. However, POPLMark (and its follow-up POPLMark Reloaded) were
-  purposedly limited in scope, with benchmark problems that do not
-  address concurrency. For this reason, we propose a new collection of
-  benchmark problems with a focus on the issues encountered when
-  mechanising message-passing concurrency using process calculi. Our
-  benchmark problems concern linear handling of environments, scope
-  extrusion, and coinductive reasoning. Our goal is to clarify,
-  compare and advance the state of the art.
+  The POPLMark challenge and its follow-up POPLMark Reloaded sparked a flurry of
+  work on machine-checked proofs, and forstered the adoption of proof
+  mechanisation in programming language research.  However, both POPLMark and
+  POPLMark Reloaded were purposely limited in scope, with benchmark problems
+  that do not address concurrency-related issues.
+  %
+  For this reason, we propose a new collection of benchmark problems with a
+  focus on the challenges and issues that typically arise when mechanising
+  message-passing concurrency using process calculi.  Our benchmark problems
+  address three key topics: linearity, scope extrusion, and coinductive
+  reasoning.  The goal of this new challenge is to clarify, compare, and advance
+  the state of the art, fostering the adoption of proof mechanisation in future
+  research on message-passing concurrency.
 

### 2023-10-10 paper/paper.tex
Revise introduction + add package "fixme"

--- 2023-10-09 paper/paper.tex
@@
 \section{Design of the benchmark}\label{sec:design-discussion}
-\subsection{General}
-- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
+\subsection{Topics we discussed during the meetings}
+- Linearity type system dual vs. Vasco's?
+- Only linear channels or also unrestricted channels?
+- Pi-calculus or CCS for linearity?
 - Choice in general
@@
   - Add as ``future work''
+- Choice in scope extrusion
 - Extension categories: other topics and small technical differences
@@
   - Choice and coinduction does not mix
+- Different notions of bisimulation in the coinduction challenge
+- Linearity: the context is polymorphic
+- Coinduction: why is congruence interesting?
+- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
+- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
 - A goal is to build a culture of formalising things in the community
+- Linearity is generally interesting
 - Encoding techniques are not different from any other field, so there's nothing ``special'' about the formalisations here
-
-\subsection{Linearity}
-- Linearity type system dual vs. Vasco's?
-- Only linear channels or also unrestricted channels?
-- Pi-calculus or CCS for linearity?
-- Linearity: the context is polymorphic
-- Linearity is generally interesting
+- We don't care about the content of messages in coinduction, only synchronization
 - Why explicitly bind two endpoints in linearity? To make the duality/polarity connection obvious
@@
 - Asynchrony? Not necessary, this is a it more advanced
-
-\subsection{Scope extrusion}
-- Choice in scope extrusion
-- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
 - Why do we keep the restrictions around in the LTS semantics in scope extrusion?
-
-\subsection{Coinduction}
-- Different notions of bisimulation in the coinduction challenge
-- Coinduction: why is congruence interesting?
-- We don't care about the content of messages in coinduction, only synchronization
 - ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics
@@
 
-As with the challenges that came before, this one is intended to be
-useful on its own, but not to cover every problem that may be faced
-when mechanising systems based on message passing calculi. The current
-challenge can be extended in two dimensions: the first, about taking
-the existing challenges further. And the second, about devising
-challenges not covered by the current benchmarks.
+As with the challenges that came before, this challenge is intended to
+have useful solutions, but not to cover every problem that may be
+faced when mechanised systems based on message passing calculi. Going
+beyond the current challenges could be in one of two dimensions: the
+first dimension about making harder challenges starting from the
+current ones, and a second one about devising challenges not covered
+by the current benchmarks.
 
-The proposed challenges try to minimise the overlap of the techniques
-they exercise, however, most systems will combine some or all aspects
-of the challenges. An interesting second round of formalisations could
-entail combining the features of the challenges. In the following
-table we consider the following three additions: First the addition of
-choice to the type level and the process level. Second, the addition
-of recursion (and recursion types for typed systems). And finally, the
-addition of channel delegation to those calculi that do not already
-support it.
+The proposed challenges are orthogonal, however, in general systems
+will combine some or all aspects of the challenges. An interesting
+second round of formalisations could entail combining the features of
+the challenges. In the following table we consider the following three
+additions: First the addition of choice to the type level and the
+process level. Second, the addition of recursion (and recursion types
+for typed systems). And finally, the addition of channel delegation
+for those calculi that do not already support it.
 
-The following table shows which extension should be suitable (that we
-mark with $+$) for each challenge:
+The following table shows which extension should be suitable (denoted
+by a $+$) for each challenge:
 
@@
 
-\begin{center}\small
 \begin{tabular}{|r|c|c|c|}
@@
 \end{tabular}
-\end{center}
 
@@
 
+Note that these additions, make the challenges less orthogonal by
+combining some of their features. This is to be expected, in the sense
+that the proposed challenges should form the building blocks of more
+complex systems.
+
 The other dimension to extend these calculi corresponds to systems
-that are not covered by our challenges. Extending in this dimension
-means proposing new challenges to cover for different features of
-message passing calculi, or to cover different features of mechanised
-proof.
+that are not covered by our challenges\dots
+% Multiparty session types, and choreographies.
 
-Some interesting aspects of message passing calculi to explore in
-further challenges could be: Multiparty session types, and
-choreographies because their definitions and meta-theory have elements
-that we do not cover here, however the elements that we do cover will
-appear in these problems. Also, while we explore a form of barbed
-bisimilarity one could easily design a challenge exploring different
-notions of bisimilarity (barbed, or weak, etc.), and trace
-equivalence.
+% different notions of bisimilarity (barbed, or weak, etc.), trace equivalence.
 
-On the side of proof assistants, the current challenge proposes
-mechanising the proofs of theorems. An interesting venue to explore
-would be to take advantage of other aspects of proof assistants. For
-example, proof assistants are often able to produce certified code (by
-using code extraction, or compiling and running their definitions).
-Additionally, while having effective automation in our current
-challenge is possible, a challenge could be designed with the explicit
-objective of automating aspects of the proofs. And finally, a
-challenge could propose the integration with other formal reasoning
-tools, namely model checkers. These tools are in common use in the
-field. However, there combining automated proofs with proof assistants
-could be a great path for attempting larger proofs with less effort.
+% Exercising different features of proof assistants like: code
+% extraction and/or code automation.
 
-Ultimately, the current challenge can be extended in several
-worthwhile directions, and we look forward to a future when they
-indeed are extended. Moreover, we see the current challenge as setting
-the foundation for those future extensions. It is our expectation that
-the solutions of those hypothetical challenges will build on solutions
-to this challenge.
+% Interaction between formalised proofs and model checkers.
+
+% And any other
 

### 2023-10-10 paper/paper.tex

--- 2023-10-10 paper/paper.tex
@@
 
-As with the challenges that came before, this challenge is intended to
-have useful solutions, but not to cover every problem that may be
-faced when mechanised systems based on message passing calculi. Going
-beyond the current challenges could be in one of two dimensions: the
-first dimension about making harder challenges starting from the
-current ones, and a second one about devising challenges not covered
-by the current benchmarks.
+As with the challenges that came before, this one is intended to be
+useful on its own, but not to cover every problem that may be faced
+when mechanising systems based on message passing calculi. The current
+challenge can be extended in two dimensions: the first, about taking
+the existing challenges further. And the second, about devising
+challenges not covered by the current benchmarks.
 
-The proposed challenges are orthogonal, however, in general systems
-will combine some or all aspects of the challenges. An interesting
-second round of formalisations could entail combining the features of
-the challenges. In the following table we consider the following three
-additions: First the addition of choice to the type level and the
-process level. Second, the addition of recursion (and recursion types
-for typed systems). And finally, the addition of channel delegation
-for those calculi that do not already support it.
+The proposed challenges try to minimise the overlap of the techniques
+they exercise, however, most systems will combine some or all aspects
+of the challenges. An interesting second round of formalisations could
+entail combining the features of the challenges. In the following
+table we consider the following three additions: First the addition of
+choice to the type level and the process level. Second, the addition
+of recursion (and recursion types for typed systems). And finally, the
+addition of channel delegation to those calculi that do not already
+support it.
 
-The following table shows which extension should be suitable (denoted
-by a $+$) for each challenge:
+The following table shows which extension should be suitable (that we
+mark with $+$) for each challenge:
 
@@
 
+\begin{center}\small
   \begin{tabular}{|r|c|c|c|}
@@
   \end{tabular}
+\end{center}
 
@@
 
-Note that these additions, make the challenges less orthogonal by
-combining some of their features. This is to be expected, in the sense
-that the proposed challenges should form the building blocks of more
-complex systems.
+The other dimension to extend these calculi corresponds to systems
+that are not covered by our challenges. Extending in this dimension
+means proposing new challenges to cover for different features of
+message passing calculi, or to cover different features of mechanised
+proof.
 
-The other dimension to extend these calculi corresponds to systems
-that are not covered by our challenges\dots
-% Multiparty session types, and choreographies.
+Some interesting aspects of message passing calculi to explore in
+further challenges could be: Multiparty session types, and
+choreographies because their definitions and meta-theory have elements
+that we do not cover here, however the elements that we do cover will
+appear in these problems. Also, while we explore a form of barbed
+bisimilarity one could easily design a challenge exploring different
+notions of bisimilarity (barbed, or weak, etc.), and trace
+equivalence.
 
-% different notions of bisimilarity (barbed, or weak, etc.), trace equivalence.
+On the side of proof assistants, the current challenge proposes
+mechanising the proofs of theorems. An interesting venue to explore
+would be to take advantage of other aspects of proof assistants. For
+example, proof assistants are often able to produce certified code (by
+using code extraction, or compiling and running their definitions).
+Additionally, while having effective automation in our current
+challenge is possible, a challenge could be designed with the explicit
+objective of automating aspects of the proofs. And finally, a
+challenge could propose the integration with other formal reasoning
+tools, namely model checkers. These tools are in common use in the
+field. However, there combining automated proofs with proof assistants
+could be a great path for attempting larger proofs with less effort.
 
-% Exercising different features of proof assistants like: code
-% extraction and/or code automation.
-
-% Interaction between formalised proofs and model checkers.
-
-% And any other
+Ultimately, the current challenge can be extended in several
+worthwhile directions, and we look forward to a future when they
+indeed are extended. Moreover, we see the current challenge as setting
+the foundation for those future extensions. It is our expectation that
+the solutions of those hypothetical challenges will build on solutions
+to this challenge.
 

### 2023-10-11 paper/paper.tex
Add some points to the design discussion.

@@
 \section{Design of the benchmark}\label{sec:design-discussion}
-\subsection{Topics we discussed during the meetings}
+In this section, we will describe the considerations we considered during the design of the benchmark challenges.
+We will start with some general remarks, then describe the individual design considerations for each challenge problem.
+
+While our challenges require reasoning about binders, which was the main topic of the original POPLMark, they also require other techniques.
+Our challenges also require reasoning about logical relations which goes above the reasoning involved in POPLMark Reloaded.
+
+We have decided not to include the choice construct in our challenge problems.
+Choice constructs complicate the theory, and our aim is to keep the challenges focused.
+For the same reason, each challenge problem only includes the fragment necessary to express the theorems we are interested in with regards to the challenge.
+We leave extensions of the syntax for a future benchmark challenge.
+
+Our challenge problems do not include reasoning about encodings, since we do not believe that encodings require any different reasoning techniques in a concurrency setting than in any other setting.
+
+\subsection{Linearity}
 - Linearity type system dual vs. Vasco's?
-- Only linear channels or also unrestricted channels?
-- Pi-calculus or CCS for linearity?
-- Choice in general
-  - Choice can be encoded, so is not necessary
-  - Complicates the theory
-  - Add as ``future work''
-- Choice in scope extrusion
-- Extension categories: other topics and small technical differences
-- Why not unified syntax?
-  - Choice and coinduction does not mix
-- Different notions of bisimulation in the coinduction challenge
 - Linearity: the context is polymorphic
+- Asynchrony? Not necessary, that is a bit more advanced
+
+For our challenge on linear reasoning we have chosen a safety theorem for a session type system since this allows us to involve linear reasoning without many definitions.
+For the same reason we have chosen not to include delegation of channels in the syntax and to use a reduction semantics instead of a transition system semantics.
+
+Linearity is of course interesting in many other subfields of programming languages, and we have thus chosen to base our challenge on a  session type system primarily due to the common use of session types in concurrency theory.
+
+Inspired by Vasconcelos~\cite{Vasconcelos2012}, we use a syntax where restriction binds two variables.
+We have chosen this presentation to make the duality of names obvious in the type system.
+
+The notion of well-formedness we have chosen for our challenge problem is quite basic, and does not for instance consider deadlocked processes.
+While more sophisticated notions of well-formedness are also interesting, the proofs of these properties are more complicated and would thus make the challenge problem more difficult and less focused on the linear aspects of the system.
+
+\subsection{Scope extrusion}
+- Why do we keep the restrictions around in the LTS semantics in scope extrusion?
+
+Our scope extrusion challenge is the challenge most closely related to the original POPLMark challenge since it concerns the properties of restrictions.
+It is not clear that the techniques used to solve the POPLMark challenges about binders can be easily extended to reasoning about scope extrusion.
+Notably, some techniques\footnote{which?} solve the POPLMark challenge by forbidding free variables, which means that the encoding of the semantics that cover scope extrusion are not obvious.
+
+\subsection{Coinduction}
 - Coinduction: why is congruence interesting?
-- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
-- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
-- A goal is to build a culture of formalising things in the community
-- Linearity is generally interesting
-- Encoding techniques are not different from any other field, so there's nothing ``special'' about the formalisations here
-- We don't care about the content of messages in coinduction, only synchronization
-- Why explicitly bind two endpoints in linearity? To make the duality/polarity connection obvious
-- Linearity: Do we care about avoiding deadlocks? No, this is stronger than safety as in well-formedness
-- Asynchrony? Not necessary, this is a it more advanced
-- Why do we keep the restrictions around in the LTS semantics in scope extrusion?
 - ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics
+
+Our coinduction challenge concerns strong barbed bisimulation and congruence.
+Though weak barbed congruence is a more common behavioural equivalence for \(\pi\)-terms, we avoid having to abstract over the number of internal transitions when considering the strong notions of equivalence, thus simplifying the theory.
+
+We have chosen not to include delegation in the coinduction challenge since it is orthogonal to our primary aim of exploring coinductive proof techniques.
 

### 2023-10-11 paper/paper.tex
grammar
rephrasing in the going beyond section

@@
 As with the challenges that came before, this one is intended to be
-useful on its own, but not to cover every problem that may be faced
-when mechanising systems based on message passing calculi. The current
-challenge can be extended in two dimensions: the first, about taking
-the existing challenges further. And the second, about devising
-challenges not covered by the current benchmarks.
+useful on its own, but not to exhaustively cover every possible
+problem in the area. The current challenge can be extended in two
+dimensions: the first is taking the existing challenges further.
+And the second is devising new challenges not covered by the current
+benchmarks.
 
@@
 tools, namely model checkers. These tools are in common use in the
-field. However, there combining automated proofs with proof assistants
-could be a great path for attempting larger proofs with less effort.
+field. However, combining automated proofs with proof assistants
+offers the potential to ease the path for attempting larger proofs
+with less effort.
 

### 2023-10-11 paper/paper.tex
Fix a few typos.

@@
 researchers towards trial-and-error, attempting sub-optimal and ad-hoc
-solutions; this increases the overall effort required by proof mechanisation,
+solutions; this increases the overall effort required for proof mechanisation,
 and hampers the reuse of techniques and results.
@@
 are fundamental to the field of concurrency theory, and they emerge in most
-mechanisations.  Moreover, the mechanisations of novel research results often
+mechanisations.  Moreover, the mechanisation of novel research results often
 requires addressing more than just one of these aspects at once --- and we see

### 2023-10-11 paper/paper.tex
Fix a few more small mistakes.

@@
 Session type systems require linear handling of typing contexts, and the proof involves reasoning about the linearity of context splitting operations.
-In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
 
 \subsection{Scope extrusion}
-Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove that barbed bisimulation is an equivalence relation for an untyped name passing calculus without infinite behaviour, but with non-deterministic choice.
+In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
+Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove the equivalence between reduction semantics and labelled transition system semantics up to structural congruence.
 The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues for some approaches to mechanising binders.
@@
 The proof involves reasoning coinductively about the infinite behaviours of the replication operator.
-We invite anyone interested to send us solutions to the challenge problems and to publish experience reports detailing the advantages and disadvantages of their approaches.
 
@@
 We will start with some general remarks, then describe the individual design considerations for each challenge problem.
+
+We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
+Like previous benchmark authors, we would like to stress that our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types and psi-calculi are not directly covered by our challenge problems.
+Most applications will however still need the basic techniques that our challenge problems exercise.
+We have designed our challenge problems to exercise the three aspects independently, and they can be solved individually and in any order.
+We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.
 
@@
 \subsection{Evaluation criteria}
-We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
-Like previous benchmark authors, we would like to stress that our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types and psi-calculi are not directly covered by our challenge problems.
-Most applications will however still need the basic techniques that our challenge problems exercise.
-We have designed our challenge problems to exercise the three aspects independently, and they can be solved individually and in any order.
-We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.
 Solutions to our challenge problems should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
@@
 
-Ultimately, the current challenge can be extended in several
+Ultimately, the current challenges can be extended in several
 worthwhile directions, and we look forward to a future when they

### 2023-10-11 paper/paper.tex
rework the design of the benchmark section

@@
 \section{Design of the benchmark}\label{sec:design-discussion}
-In this section, we will describe the considerations we considered during the design of the benchmark challenges.
-We will start with some general remarks, then describe the individual design considerations for each challenge problem.
+In this section, we outline the factors considered in designing the
+benchmark challenges. We start with some general remarks, then
+describe the individual design considerations for each challenge
+problem.
 
@@
 While our challenges require reasoning about binders, which was the main topic of the original POPLMark, they also require other techniques.
-Our challenges also require reasoning about logical relations which goes above the reasoning involved in POPLMark Reloaded.
+Our challenges also require reasoning about logical relations which goes above the reasoning involved in POPLMark Reloaded.\fxwarning{does it? as in do we define a type indexed
+  relation and then use that to prove things?}
 
-We have decided not to include the choice construct in our challenge problems.
-Choice constructs complicate the theory, and our aim is to keep the challenges focused.
-For the same reason, each challenge problem only includes the fragment necessary to express the theorems we are interested in with regards to the challenge.
-We leave extensions of the syntax for a future benchmark challenge.
+The aim of this challenge is to keep the problems focused on the
+features we want to exercise. For that reason, we have decided not to
+include the choice construct in our challenge problems. Choice
+constructs complicate the theory. For the same reason, each challenge
+problem only includes the fragment necessary to express the theorems
+we are interested in. We leave extensions of the syntax for a future
+benchmark challenge.
 
-Our challenge problems do not include reasoning about encodings, since we do not believe that encodings require any different reasoning techniques in a concurrency setting than in any other setting.
+Our challenge problems do not include reasoning about encodings. While
+this kind of reasoning is common in the field, it does not require any
+different reasoning techniques in a concurrency setting than in any
+other setting.
 

### 2023-10-11 paper/paper.tex
rework scope extrusion

@@
 
-For our challenge on linear reasoning we have chosen a safety theorem for a session type system since this allows us to involve linear reasoning without many definitions.
-For the same reason we have chosen not to include delegation of channels in the syntax and to use a reduction semantics instead of a transition system semantics.
+For our challenge on linear reasoning we have chosen a safety theorem
+for a session type system since this allows us to involve linear
+reasoning without many definitions. Accordingly, we have chosen not to
+include channel delegation and to use a reduction instead of a
+transition system semantics.
 
@@
 
-Inspired by Vasconcelos~\cite{Vasconcelos2012}, we use a syntax where restriction binds two variables.
-We have chosen this presentation to make the duality of names obvious in the type system.
+Inspired by Vasconcelos~\cite{Vasconcelos2012}, we use a syntax where
+restriction binds two variables. This presentation makes the duality
+of the two names obvious in the type system.
 
-The notion of well-formedness we have chosen for our challenge problem is quite basic, and does not for instance consider deadlocked processes.
-While more sophisticated notions of well-formedness are also interesting, the proofs of these properties are more complicated and would thus make the challenge problem more difficult and less focused on the linear aspects of the system.
+We have simple notion of well-formedness for our challenge problem,
+and does not for instance consider deadlocked processes. While more
+sophisticated notions of well-formedness are also interesting, the
+proofs of these properties are more complicated and would thus make
+the challenge problem more difficult and less focused on the linear
+aspects of the system.
 

### 2023-10-11 challenges/linearity.tex
inserted dual calculus in linearity section and other non-insignificant changes

@@
 \label{sec:challenge:linearity-beh-types}
-
+\renewcommand{\Cempty}{\cdot}
+\renewcommand{\Csplit}[2]{#1,  #2}
+\renewcommand{\Cupdate}[2]{#1, #2}
+\newcommand{\TinV}[2]{{?#1}.#2}
+\newcommand{\ToutV}[2]{{!#1}.#2}
+\newcommand{\tend}[1]{\mathbf{end}(#1)}
 This challenge formalises a proof that requires reasoning about linearity of channels.
@@
 
-The setting for this challenge is a small calculus with a session type system, the syntax and semantics of which are presented below.
-The objective of this challenge is to prove type preservation (also known as subject reduction) for the type system, i.e.\ that well-typed processes can only transition to processes which are also well-typed in the same context.
+The setting for this challenge is a small calculus with a session type
+system, the syntax and semantics of which are given below. The
+calculus is a fragment of the one presented in~\cite{Vasconcelos2012},
+formulated in the dual style of~\cite{barber96tr}.
+
+The objective of this challenge is to prove type preservation (also
+known as subject reduction) for the type system, i.e.\ that well-typed
+processes can only transition to processes which are also well-typed
+in the same context.
 
@@
 The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
-\begin{metanote}
-  AM: take the syntax from dual file appst.tex and adapt
-\end{metanote}
-\begin{align*}
-  S,T := &&& \Tend \\
-  |&&& \Tbase \\
-  |&&& \Tin{S} \\
-  |&&& \Tout{S} \\
-  \Gamma := &&& \Cempty \\
-  |&&& \Cadd{\Gamma}{\hastype{x}{S}}
-       %% AM: added
-        \\
-  |&&& \Cadd{\Gamma}{\hastype{l}{\Tbase}} %% changed, AM
-\end{align*}
+
+\[
+  \begin{array}{rcl}
+  S,T & ::= & \Tend \mid \Tbase \mid \Tin{S} \mid \Tout{S} \\
+    \Gamma &::= & \Cempty \mid \Gamma, l \\
+                    \Delta &::= & \Cempty \mid \Cadd{\Delta}{\hastype{x}{S}}
+     \end{array}
+\]
+
+% \begin{metanote}
+%   AM: take the syntax from dual file appst.tex and adapt
+% \end{metanote}
+% \begin{align*}
+%   S,T := &&& \Tend \\
+%   |&&& \Tbase \\
+%   |&&& \Tin{S} \\
+%   |&&& \Tout{S} \\
+%   \Gamma := &&& \Cempty \\
+%   |&&& \Cadd{\Gamma}{\hastype{x}{S}}
+%        %% AM: added
+%         \\
+%   |&&& \Cadd{\Gamma}{\hastype{l}{\Tbase}} %% changed, AM
+% \end{align*}
 The \emph{end type}, \( \Tend \), is the type of an endpoint on which no further interaction is possible.
@@
 
-Typing contexts gather type information about names.
-The \emph{empty type context}, \( \Cempty \), does not contain any information.
-The \emph{type context with \( x \) added}, \( \Cadd{\Gamma}{\hastype{x}{T}} \) adds the information that \( x \) has type \( T \) to the existing type context \( \Gamma \).
-All of the names added to a type context must be distinct.
-Note that
-the order in which information is added to a type context thus does not matter.
+Typing contexts gather type information about names and variables.
+We distinguish between an unrestricted context $\Gamma$ seen as a set: in our special case where we have only one base type, we only record which variable have been used.
+%
+Type information for channel names is captured in the \emph{linear}
+context $\Delta$. We use the comma as split/union, overloaded to
+singletons and $\cdot$ is the empty multiset.
+% The \emph{empty type context}, \( \Cempty \), does not contain any information.
+         %          The \emph{type context with \( x \) added}, \( \Cadd{\Gamma}{\hastype{x}{T}} \) adds the information that \( x \) has type \( T \) to the existing type context \( \Gamma \).
+We extend the Barendregt convention to contexts, so that all entries
+are distinct.  Note that the order in which information is added to a
+type context thus does not matter.
 % We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
@@
 
-\begin{metanote}
-  AM: put here the dual rules, remove splitting
-\end{metanote}
+% \begin{metanote}
+%   AM: put here the dual rules, remove splitting
+% \end{metanote}
 
-To keep track of linearity, we introduce two \emph{qualifiers}: \( \un{T} \) and \( \lin{T} \).
-A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
-All types are \emph{linear}, \( \lin{T} \).
-The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
+% To keep track of linearity, we introduce two \emph{qualifiers}: \( \un{T} \) and \( \lin{T} \).
+% A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
+% All types are \emph{linear}, \( \lin{T} \).
+% The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
 
-The \emph{linear split} of a context, \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \), is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Split-Empty]{ }{\Cempty = \Csplit{\Cempty}{\Cempty}}
-  \and
-  \inferrule[Split-Un]{\Csplit{\Gamma_1}{\Gamma_2} = \Gamma \\ \un{T}}{\Cadd{\Gamma}{\hastype{x}{T}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{T}})}{(\Cadd{\Gamma_2}{\hastype{x}{T}})}}
-  \and
-  \inferrule[Split-Lin-L]{\Gamma = \Csplit{\Gamma_1}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{S}})}{\Gamma_2}}
-  \and
-  \inferrule[Split-Lin-R]{\Gamma = \Csplit{\Gamma_2}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{\Gamma_1}{(\Cadd{\Gamma_2}{\hastype{x}{S}})}}
-\end{mathpar}
+% The \emph{linear split} of a context, \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \), is defined by the following rules:
+% \begin{mathpar}
+%   \inferrule[Split-Empty]{ }{\Cempty = \Csplit{\Cempty}{\Cempty}}
+%   \and
+%   \inferrule[Split-Un]{\Csplit{\Gamma_1}{\Gamma_2} = \Gamma \\ \un{T}}{\Cadd{\Gamma}{\hastype{x}{T}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{T}})}{(\Cadd{\Gamma_2}{\hastype{x}{T}})}}
+%   \and
+%   \inferrule[Split-Lin-L]{\Gamma = \Csplit{\Gamma_1}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{S}})}{\Gamma_2}}
+%   \and
+%   \inferrule[Split-Lin-R]{\Gamma = \Csplit{\Gamma_2}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{\Gamma_1}{(\Cadd{\Gamma_2}{\hastype{x}{S}})}}
+% \end{mathpar}
 
@@
 \begin{mathpar}
-  \inferrule[T-Base]{\un{\Gamma}}{\types{\Gamma}{\hastype{a}{\Tbase}}}
-  \and
-  \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{l}{\Tbase}}}{\hastype{l}{\Tbase}}}
+  \inferrule[T-Base]{ }{\typev{\Gamma}{\hastype{a}{\Tbase}}} \and
+  \inferrule[T-Var]{ }{\typev{\Cadd{\Gamma}l }{\hastype{l}{\Tbase}}}
 \end{mathpar}
-The only typing rule for names is:
+The typing rule for names is, where $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
 \begin{mathpar}
-  \inferrule[T-Name]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{x}{T}}}
+  \inferrule[T-Name]{{\tend\Delta }}{\types{{\Gamma; (\Delta,\hastype{x}{T})}}{\hastype{x}{T}}}
 \end{mathpar}
@@
 \begin{mathpar}
-  \inferrule[T-Inact]{\un{\Gamma}}{\types{\Gamma}{\Pend}}
+
+  \inferrule[T-Inact]{\tend\Delta }{\types{\Gamma;\Delta}{\Pend}}
   \and
-  \inferrule[T-Par]{\types{\Gamma_1}{P} \\ \types{\Gamma_2}{Q}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Ppar{P}{Q}}}
+  \inferrule[T-Par]{\types{\Gamma;\Delta_1}{P} \\ \types{\Gamma;\Delta_2}{Q}}
+  {\types{\Gamma; \Csplit{\Delta_1}{\Delta_2}}{\Ppar{P}{Q}}}
   \and
-  \inferrule[T-Res]{\types{\Cadd{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P}}{\types{\Gamma}{\Presd{x}{y}{P}}}
+  \inferrule[T-Res]{\types{\Gamma; (\Cadd{\Cadd{\Delta}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P})}{\types{\Gamma}{\Presd{x}{y}{P}}}
   \and
-  \inferrule[T-Out]{\types{\Gamma_1}{\hastype{x}{\Tout{T}}} \\ \types{\Gamma_2}{\hastype{v}{\Tbase}} \\ \types{\Cupdate{\Gamma_3}{\hastype{x}{T}}}{P}}{\types{\Csplit{\Gamma_1}{\Csplit{\Gamma_2}{\Gamma_3}}}{\Pout{x}{v}{P}}}
+    \inferrule[T-Out]{%\types{\Gamma;\Delta_1}{\hastype{x}{\Tout{T}}} \\
+      \typev{\Gamma}{\hastype{v}{\Tbase}} \\ \types{\Gamma; \Cupdate{\Delta}{\hastype{x}{T}}}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tout{T}}})}{\Pout{x}{v}{P}}}
+   %  \and
+   % \inferrule[T-Out]{\types{\Gamma;\Delta_1}{\hastype{x}{\Tout{T}}} \\
+   %   \typev{\Gamma}{\hastype{v}{\Tbase}} \\ \types{\Gamma; (\Cupdate{\Delta}{\hastype{x}{T}})}{P}}{\types{\Gamma; \Csplit{\Delta_1}{\Delta_2}}{\Pout{x}{v}{P}}}
     \and
-  \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{\Cadd{\Gamma_2}{\hastype{l}{\Tbase}}}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
+    \inferrule[T-IN]{%\types{\Gamma;\Delta_1}{\hastype{x}{\Tout{T}}} \\
+      %\typev{\Gamma}{\hastype{v}{\Tbase}} \\
+      \types{(\Gamma, \hastype l \Tbase); (\Cupdate{\Delta}{\hastype{x}{T}})}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tin{T}}})}{\Pin{x}{l}{P}}}
+  % \inferrule[T-IN]{\typev{\Gamma}{\hastype{v}{\Tbase}} \\
+  %   \types{(\Gamma, \hastype l \Tbase); \Delta}{\types{\Gamma; \Csplit {\hastype{x}{\Tout{T}}}{\Delta}}
+  %     {\Pin{x}{l}{P}}}}
+% \and
+% \inferrule[UN-L]
+% {\types{(\Gamma,l : T); \Delta}{P\{x/ l\}}}
+% {\types{\Gamma; (\Delta, x: \un{T})}{P}}
 \end{mathpar}
+
+% \begin{mathpar}
+%   \inferrule[T-Base]{\un{\Gamma}}{\types{\Gamma}{\hastype{a}{\Tbase}}}
+%   \and
+%   \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{l}{\Tbase}}}{\hastype{l}{\Tbase}}}
+% \end{mathpar}
+% The only typing rule for names is:
+% \begin{mathpar}
+%   \inferrule[T-Name]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{x}{T}}}
+% \end{mathpar}
+% The typing rules for processes are as follows:
+% \begin{mathpar}
+%   \inferrule[T-Inact]{\un{\Gamma}}{\types{\Gamma}{\Pend}}
+%   \and
+%   \inferrule[T-Par]{\types{\Gamma_1}{P} \\ \types{\Gamma_2}{Q}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Ppar{P}{Q}}}
+%   \and
+%   \inferrule[T-Res]{\types{\Cadd{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P}}{\types{\Gamma}{\Presd{x}{y}{P}}}
+%   \and
+%   \inferrule[T-Out]{\types{\Gamma_1}{\hastype{x}{\Tout{T}}} \\ \types{\Gamma_2}{\hastype{v}{\Tbase}} \\ \types{\Cupdate{\Gamma_3}{\hastype{x}{T}}}{P}}{\types{\Csplit{\Gamma_1}{\Csplit{\Gamma_2}{\Gamma_3}}}{\Pout{x}{v}{P}}}
+%   \and
+%   \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{\Cadd{\Gamma_2}{\hastype{l}{\Tbase}}}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
+% \end{mathpar}
 
@@
 \begin{theorem}[Subject reduction]
-  If \( \types{\Gamma}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma}{Q} \).
+  If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
 \end{theorem}
+\begin{proof}
+  Induction on the reduction relation, TBD.
+\end{proof}
 
 The following lemmas may be of use:
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \un{T} \) then \( \types{\Cadd{\Gamma}{\hastype{x}{T}}}{P} \).
-\end{lemma}
-
-\begin{lemma}
-  Assume \( \types{\Gamma}{P} \) and \( x \notin \freenames{P} \).
-  Then:
+\begin{lemma}[Weakening]\mbox{}
+  \label{lemma:weak}
   \begin{enumerate}
-  \item \( \hastype{x}{\Tout{T}} \notin \Gamma \).
-  \item \( \hastype{x}{\Tin{T}} \notin \Gamma \).
-  \item if \( \Gamma = \Cadd{\Gamma'}{\hastype{x}{T}} \) then \( \types{\Gamma'}{P} \).
+  \item If \( \types{\Gamma; \Delta}{P} \) then
+    \( \types{\Cadd{\Gamma}{l};\Delta}{P} \).
+      \item If \( \types{\Gamma; \Delta}{P} \) and \( \tend{\Delta} \) then
+    \( \types{\Gamma;\Cadd{\Delta}{\hastype{x}{T}}}{P} \).
 \end{enumerate}
 \end{lemma}
+\begin{metanote}
+  Linear weakening holds only if we put $\tend\Delta$ in T-NAME.
+\end{metanote}
+\begin{proof}
+  By induction on the given derivations.
+\end{proof}
+\begin{lemma}[Strengthening]\mbox{}
+  \label{lemma:strenD}
+  If \( \types{\Gamma; \Delta, z : T}{P} \) and $z\not\in fv(P)$, then \( \types{\Gamma; \Delta}{P} \).
+\end{lemma}
+\begin{proof}
+  By induction on the given derivation.
+\end{proof}
 
-\begin{lemma}
-  If \( \types{\Gamma}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma}{Q} \).
+\begin{lemma}[Substitution]\mbox{}
+  If $\types{(\Gamma, l);\Delta}{P}$ and ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
+  \( \types{\Gamma;\Delta}{\subst{P}{l}{a}} \).
 \end{lemma}
-% \begin{metanote}
-%   AM: the statement of substitution was wrong: now it is consistent with the op sem, but I need to check the (I assume now simpler) proof.
+\begin{proof}
+  By induction on the first derivation.
+\end{proof}
+\begin{lemma}[Preservation for $\equiv$]
+  If \( \types{\Gamma;\Delta}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
+\end{lemma}
+\begin{metanote}
+  AM: I'm still confused about this proof, in particular Vasco's approach to consider the axioms as symmetrization of rewrite rules
+\end{metanote}.
+\begin{proof}
+  By induction on the structure of a process context. In the base case, proceed by case analysis on the \textsl{Sc} rule applied:
+  \begin{description}
+  \item[Res:] use weakening~\ref{lemma:weak} part 2
+  \item[Comm/Assoc:] by rearranging sub-derivations noting that  order does not matter for linear context
+  \item[Inact] Left-to-right by
+    strengthening~\ref{lemma:strenD}. Vice-versa, by weakening~\ref{lemma:weak} part 2.
+  \item[Res-Par] Follows Vasco's proof by case analysis on $x : T$ being linear and applying weak and stren accordingly.
+  \item[Res] order does not matter
+  \end{description}
+  For the step case, TBD
+\end{proof}
+
+% \begin{lemma}
+%   If \( \types{\Gamma}{P} \) and \( \un{T} \) then \( \types{\Cadd{\Gamma}{\hastype{x}{T}}}{P} \).
+% \end{lemma}
+
+% \begin{lemma}
+%   Assume \( \types{\Gamma}{P} \) and \( x \notin \freenames{P} \).
+%   Then:
+%   \begin{enumerate}
+%   \item \( \hastype{x}{\Tout{T}} \notin \Gamma \).
+%   \item \( \hastype{x}{\Tin{T}} \notin \Gamma \).
+%   \item if \( \Gamma = \Cadd{\Gamma'}{\hastype{x}{T}} \) then \( \types{\Gamma'}{P} \).
+%   \end{enumerate}
+% \end{lemma}
+
+% \begin{lemma}
+%   If \( \types{\Gamma}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma}{Q} \).
+% \end{lemma}
+% % \begin{metanote}
+% %   AM: the statement of substitution was wrong: now it is consistent with the op sem, but I need to check the (I assume now simpler) proof.
 Note that $T$ must be \textbf{base} and \( \un{\Gamma_i} \) must also hold.
-% \end{metanote}
-\begin{lemma}
-  If \( \types{\Gamma_1}{\hastype{v}{T}} \) and
-  \( \types{\Cadd{\Gamma_2}{\hastype{l}{T}}}{P} \) and
-  \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then
-  \( \types{\Gamma}{\subst{P}{l}{v}} \).
-\end{lemma}
+% % \end{metanote}
+% \begin{lemma}
+%   If \( \types{\Gamma_1}{\hastype{v}{T}} \) and
+%   \( \types{\Cadd{\Gamma_2}{\hastype{l}{T}}}{P} \) and
+%   \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then
+%   \( \types{\Gamma}{\subst{P}{l}{v}} \).
+% \end{lemma}
 

### 2023-10-11 paper/paper.tex
rework scope extrusion

@@
 
-Our scope extrusion challenge is the challenge most closely related to the original POPLMark challenge since it concerns the properties of restrictions.
+This is the challenge most closely related to the original POPLMark
+challenge since it concerns the properties of binders in restrictions.
 It is not clear that the techniques used to solve the POPLMark challenges about binders can be easily extended to reasoning about scope extrusion.
-Notably, some techniques\footnote{which?} solve the POPLMark challenge by forbidding free variables, which means that the encoding of the semantics that cover scope extrusion are not obvious.
+Notably, techniques like Higher Order Abstract Syntax, or well-scoped
+De\ Bruijn indices do not have the concept of free variables. For this
+kind of system the means to encode the operational semantics are not
+obvious.
 

### 2023-10-11 paper/paper.tex
rework scope extrusion (this time for real)
comment out paragraph in the design

@@
 
-While our challenges require reasoning about binders, which was the main topic of the original POPLMark, they also require other techniques.
-Our challenges also require reasoning about logical relations which goes above the reasoning involved in POPLMark Reloaded.\fxwarning{does it? as in do we define a type indexed
-  relation and then use that to prove things?}
+% While our challenges require reasoning about binders, which was the main topic of the original POPLMark, they also require other techniques.
+% Our challenges also require reasoning about logical relations which goes above the reasoning involved in POPLMark Reloaded.
 
@@
 Notably, techniques like Higher Order Abstract Syntax, or well-scoped
-De\ Bruijn indices do not have the concept of free variables. For this
-kind of system the means to encode the operational semantics are not
-obvious.
+De\ Bruijn indices do not have the concept of free variables. It is
+not immediately obvious how to encode the operational semantics in
+this kind of system.
 
@@
 Solutions to our challenge problems should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
-Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges will need to present more elaborate argumentation for their adequacy.
+Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges need to present more elaborate argumentation for their adequacy.
 

### 2023-10-11 challenges/linearity.tex
fixing a typo in T-IN
Clean up linearity.tex.

@@
 \label{sec:challenge:linearity-beh-types}
-\renewcommand{\Cempty}{\cdot}
-\renewcommand{\Csplit}[2]{#1,  #2}
-\renewcommand{\Cupdate}[2]{#1, #2}
-\newcommand{\TinV}[2]{{?#1}.#2}
-\newcommand{\ToutV}[2]{{!#1}.#2}
-\newcommand{\tend}[1]{\mathbf{end}(#1)}
 This challenge formalises a proof that requires reasoning about linearity of channels.
@@
 \subsubsection{Syntax}
-
-% We assume the existence of some type of \emph{base values}, values of which we will denote by \( a, b, \dots \), the existence of some type of \emph{variable names}, values of which we will denote by \( l, m, \dots \), and the existence of some type of \emph{names}, values of which we will denote by \( x, y, \dots \).
-% We assume the existence of some type of \emph{base values}, represented by the symbols
-%  \( a, b, \dots \), the existence of some type
-% of \emph{variable names}, represented by the symbols
-% \( l, m, \dots \), and the existence of some type of \emph{names},
-% represented by the symbols \( x, y, \dots \).
-
 The syntax is given by the grammar:
-% \begin{align*}
-%   v,w :=&&& a \\
-%   |&&& l \\
-%   P,Q :=&&& \Pend \\
-%   |&&& \Pout{x}{v}{P} \\
-%   |&&& \Pin{x}{l}{P} \\
-%   |&&& \Ppar{P}{Q} \\
-%   |&&& \Presd{x}{y}{P}
-         %   \end{align*}
-% \newcommand{\bmid}{\mathbf{\Vert}}
 \[
@@
 A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
-% The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \), then continue as \( P \).
+
+The output process \( \Pout{x}{v}{P} \) can send the value \( v \) via \( x \), then continue as \( P \).
 The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
 The input process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
-% The input operator thus binds the variable name \( l \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
+
 The process \( \Presd{x}{y}{P} \) is the \emph{restriction} of the channel with endpoints named \( x \) and \( y \) to \( P \).
@@
 Note that there is no recursion of replication in the syntax, and thus no infinite behaviors can be expressed.
-
-% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction) in \( P \).
-% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction) in \( P \).
-% We will use the notation \( \freevars{P} \) to denote the set of variable names that occur free (i.e.\ not bound by an input) in \( P \).
-% We will use the notation \( \boundvars{P} \) to denote the set of variable names that occur bound (by an input) in \( P \).
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process \( P \) with value \( v \) substituted for variable name \( l \).
-
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound variable names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% As a convention, we assume that bound names and bound variable names of any processes are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, substitutions, and sets of names or variable names.
-% This is justified because any overlapping names and variable names may be \( \alpha \)-converted such that the assumption is satisfied.
 
@@
 
-\begin{metanote}
-  AM: the above paragraph is repeated in the coinduction challenge and should be removed there
-\end{metanote}
 \subsubsection{Session types}
@@
 As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of restricting names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
-Another example is the following process, which attempts to send a variable that is not resolved at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
+Another example is the following process, which attempts to send a variable that is not instantiated at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
 
@@
 
-\begin{metanote}
-  AM: I've moved the discussion of well-formedness must go after the
-  type system, since we want to separate the preservation vs
-  safety challenges
-\end{metanote}
 \paragraph{Syntax}
 Our type system does not type processes directly, but instead focuses on the channels used in the process.
-The syntax of \emph{session types} \( S, T \) and \emph{type contexts} \( \Gamma \) is as follows:
-
+The syntax of \emph{session types} \( S, T \), unrestricted type contexts \( \Gamma \) and linear type contexts \( \Delta \) is as follows:
 \[
@@
 \]
-
-% \begin{metanote}
-%   AM: take the syntax from dual file appst.tex and adapt
-% \end{metanote}
-% \begin{align*}
-%   S,T := &&& \Tend \\
-%   |&&& \Tbase \\
-%   |&&& \Tin{S} \\
-%   |&&& \Tout{S} \\
-%   \Gamma := &&& \Cempty \\
-%   |&&& \Cadd{\Gamma}{\hastype{x}{S}}
-%        %% AM: added
-%         \\
-%   |&&& \Cadd{\Gamma}{\hastype{l}{\Tbase}} %% changed, AM
-% \end{align*}
 The \emph{end type}, \( \Tend \), is the type of an endpoint on which no further interaction is possible.
@@
 Typing contexts gather type information about names and variables.
-We distinguish between an unrestricted context $\Gamma$ seen as a set: in our special case where we have only one base type, we only record which variable have been used.
-%
-Type information for channel names is captured in the \emph{linear}
-context $\Delta$. We use the comma as split/union, overloaded to
-singletons and $\cdot$ is the empty multiset.
-% The \emph{empty type context}, \( \Cempty \), does not contain any information.
-         %          The \emph{type context with \( x \) added}, \( \Cadd{\Gamma}{\hastype{x}{T}} \) adds the information that \( x \) has type \( T \) to the existing type context \( \Gamma \).
+We distinguish between an unrestricted context $\Gamma$ seen as a set: since  we have only one base type, we only record which variable names have been used.
+Type information for endpoints is captured in the \emph{linear}
+context \( \Delta \). We use the comma as split/union, overloaded to
+singletons, and \( \Cempty \) is the empty multiset.
 We extend the Barendregt convention to contexts, so that all entries
@@
 type context thus does not matter.
-% We treat type contexts as sets, not ordered lists, and the order in which information is added to a type context thus does not matter.
 
@@
 \end{mathpar}
-Note that the dual function is not total, since it is not defined for base types.
+Note that the dual function is not total, since it is not defined for the base type.
 
 \subsubsection{Typing rules}
-% \paragraph{Splitting contexts}
 Our type system will maintain two invariants:
@@
 
-% \begin{metanote}
-%   AM: put here the dual rules, remove splitting
-% \end{metanote}
-
-% To keep track of linearity, we introduce two \emph{qualifiers}: \( \un{T} \) and \( \lin{T} \).
-% A type is \emph{unrestricted}, \( \un{T} \), if and only if \( T = \Tbase \) or \( T = \Tend \).
-% All types are \emph{linear}, \( \lin{T} \).
-% The qualifiers extend to contexts in the following way: \( q(\Gamma) \) if and only if \( \hastype{x}{T} \in \Gamma \) implies \( q(T) \).
-
-% The \emph{linear split} of a context, \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \), is defined by the following rules:
-% \begin{mathpar}
-%   \inferrule[Split-Empty]{ }{\Cempty = \Csplit{\Cempty}{\Cempty}}
-%   \and
-%   \inferrule[Split-Un]{\Csplit{\Gamma_1}{\Gamma_2} = \Gamma \\ \un{T}}{\Cadd{\Gamma}{\hastype{x}{T}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{T}})}{(\Cadd{\Gamma_2}{\hastype{x}{T}})}}
-%   \and
-%   \inferrule[Split-Lin-L]{\Gamma = \Csplit{\Gamma_1}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{(\Cadd{\Gamma_1}{\hastype{x}{S}})}{\Gamma_2}}
-%   \and
-%   \inferrule[Split-Lin-R]{\Gamma = \Csplit{\Gamma_2}{\Gamma_2}}{\Cadd{\Gamma}{\hastype{x}{S}} = \Csplit{\Gamma_1}{(\Cadd{\Gamma_2}{\hastype{x}{S}})}}
-% \end{mathpar}
-
-% We will also need to update contexts with new type information in a safe way while typing input and output.
-% The \emph{context update}, \( \Cupdate{\Gamma}{\hastype{x}{T}} = \Gamma' \), is defined by the following rules:
-% \begin{mathpar}
-%   \inferrule[Update-Name]{\hastype{x}{U} \notin \Gamma}{\Cupdate{\Gamma}{\hastype{x}{T}} = \Cadd{\Gamma}{\hastype{x}{T}}}
-%   \and
-%   \inferrule[Update-Un]{\un{T}}{\Cupdate{(\Cadd{\Gamma}{\hastype{x}{T}})}{\hastype{x}{T}} = (\Cadd{\Gamma}{\hastype{x}{T}})}
-% \end{mathpar}
-
 We have three typing judgments: one for values, one for names, and one for processes.
@@
 \end{mathpar}
-The typing rule for names is, where $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
+The typing rule for names is as follows, where $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
 \begin{mathpar}
@@
   \and
-    \inferrule[T-Out]{%\types{\Gamma;\Delta_1}{\hastype{x}{\Tout{T}}} \\
+    \inferrule[T-Out]{
       \typev{\Gamma}{\hastype{v}{\Tbase}} \\ \types{\Gamma; \Cupdate{\Delta}{\hastype{x}{T}}}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tout{T}}})}{\Pout{x}{v}{P}}}
-   %  \and
-   % \inferrule[T-Out]{\types{\Gamma;\Delta_1}{\hastype{x}{\Tout{T}}} \\
-   %   \typev{\Gamma}{\hastype{v}{\Tbase}} \\ \types{\Gamma; (\Cupdate{\Delta}{\hastype{x}{T}})}{P}}{\types{\Gamma; \Csplit{\Delta_1}{\Delta_2}}{\Pout{x}{v}{P}}}
     \and
-    \inferrule[T-IN]{%\types{\Gamma;\Delta_1}{\hastype{x}{\Tout{T}}} \\
-      %\typev{\Gamma}{\hastype{v}{\Tbase}} \\
-      \types{(\Gamma, \hastype l \Tbase); (\Cupdate{\Delta}{\hastype{x}{T}})}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tin{T}}})}{\Pin{x}{l}{P}}}
-  % \inferrule[T-IN]{\typev{\Gamma}{\hastype{v}{\Tbase}} \\
-  %   \types{(\Gamma, \hastype l \Tbase); \Delta}{\types{\Gamma; \Csplit {\hastype{x}{\Tout{T}}}{\Delta}}
-  %     {\Pin{x}{l}{P}}}}
-% \and
-% \inferrule[UN-L]
-% {\types{(\Gamma,l : T); \Delta}{P\{x/ l\}}}
-% {\types{\Gamma; (\Delta, x: \un{T})}{P}}
+    \inferrule[T-IN]{
+      \types{(\Gamma,  l ); (\Cupdate{\Delta}{\hastype{x}{T}})}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tin{T}}})}{\Pin{x}{l}{P}}}
 \end{mathpar}
 
-% \begin{mathpar}
-%   \inferrule[T-Base]{\un{\Gamma}}{\types{\Gamma}{\hastype{a}{\Tbase}}}
-%   \and
-%   \inferrule[T-Var]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{l}{\Tbase}}}{\hastype{l}{\Tbase}}}
-% \end{mathpar}
-% The only typing rule for names is:
-% \begin{mathpar}
-%   \inferrule[T-Name]{\un{\Gamma}}{\types{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{x}{T}}}
-% \end{mathpar}
-% The typing rules for processes are as follows:
-% \begin{mathpar}
-%   \inferrule[T-Inact]{\un{\Gamma}}{\types{\Gamma}{\Pend}}
-%   \and
-%   \inferrule[T-Par]{\types{\Gamma_1}{P} \\ \types{\Gamma_2}{Q}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Ppar{P}{Q}}}
-%   \and
-%   \inferrule[T-Res]{\types{\Cadd{\Cadd{\Gamma}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P}}{\types{\Gamma}{\Presd{x}{y}{P}}}
-%   \and
-%   \inferrule[T-Out]{\types{\Gamma_1}{\hastype{x}{\Tout{T}}} \\ \types{\Gamma_2}{\hastype{v}{\Tbase}} \\ \types{\Cupdate{\Gamma_3}{\hastype{x}{T}}}{P}}{\types{\Csplit{\Gamma_1}{\Csplit{\Gamma_2}{\Gamma_3}}}{\Pout{x}{v}{P}}}
-%   \and
-%   \inferrule[T-In]{\types{\Gamma_1}{\hastype{x}{\Tin{U}}} \\ \types{\Cupdate{\Cadd{\Gamma_2}{\hastype{l}{\Tbase}}}{\hastype{x}{U}}}{P}}{\types{\Csplit{\Gamma_1}{\Gamma_2}}{\Pin{x}{l}{P}}}
-% \end{mathpar}
-
 \subsubsection{Challenge}
-\begin{metanote}
-AM:  I'm separating sr from safety, or whatever we call the latter
-\end{metanote}
-The objective of this challenge is to prove type preservation for our calculus with session types. We start with:
+The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with:
 \begin{theorem}[Subject reduction]
@@
 \end{lemma}
-\begin{metanote}
-  Linear weakening holds only if we put $\tend\Delta$ in T-NAME.
-\end{metanote}
 \begin{proof}
@@
   \label{lemma:strenD}
-  If \( \types{\Gamma; \Delta, z : T}{P} \) and $z\not\in fv(P)$, then \( \types{\Gamma; \Delta}{P} \).
+  If \( \types{\Gamma; \Delta, z : T}{P} \) and $z\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
 \end{lemma}
@@
 \end{lemma}
-\begin{metanote}
-  AM: I'm still confused about this proof, in particular Vasco's approach to consider the axioms as symmetrization of rewrite rules
-\end{metanote}.
 \begin{proof}
@@
 
-% \begin{lemma}
-%   If \( \types{\Gamma}{P} \) and \( \un{T} \) then \( \types{\Cadd{\Gamma}{\hastype{x}{T}}}{P} \).
-% \end{lemma}
-
-% \begin{lemma}
-%   Assume \( \types{\Gamma}{P} \) and \( x \notin \freenames{P} \).
-%   Then:
-%   \begin{enumerate}
-%   \item \( \hastype{x}{\Tout{T}} \notin \Gamma \).
-%   \item \( \hastype{x}{\Tin{T}} \notin \Gamma \).
-%   \item if \( \Gamma = \Cadd{\Gamma'}{\hastype{x}{T}} \) then \( \types{\Gamma'}{P} \).
-%   \end{enumerate}
-% \end{lemma}
-
-% \begin{lemma}
-%   If \( \types{\Gamma}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma}{Q} \).
-% \end{lemma}
-% % \begin{metanote}
-% %   AM: the statement of substitution was wrong: now it is consistent with the op sem, but I need to check the (I assume now simpler) proof.
-Note that $T$ must be \textbf{base} and \( \un{\Gamma_i} \) must also hold.
-% % \end{metanote}
-% \begin{lemma}
-%   If \( \types{\Gamma_1}{\hastype{v}{T}} \) and
-%   \( \types{\Cadd{\Gamma_2}{\hastype{l}{T}}}{P} \) and
-%   \( \Gamma = \Csplit{\Gamma_1}{\Gamma_2} \) then
-%   \( \types{\Gamma}{\subst{P}{l}{v}} \).
-% \end{lemma}
-
 To formulate safety, we need to formally define what we mean by well-formed.
@@
 
-\begin{theorem}[Safety]
+\begin{theorem}[Type safety]
   If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
@@
 
-\begin{corollary}[Type preservation]
+\begin{corollary}
   If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.

### 2023-10-11 challenges/linearity.tex
pass on linearity appendix

@@
 
-\subsubsection{Syntax}
-The syntax is given by the grammar:
+\subsubsection{Syntax.}
+The syntax is given by the grammar
 \[
@@
 \]
-A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
+where a \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
 
-The output process \( \Pout{x}{v}{P} \) can send the value \( v \) via \( x \), then continue as \( P \).
+The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \) and then continues as \( P \).
 The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) is an \emph{input}, which can receive a base value via \( x \), then continue as \( P \) with the received value substituted for the variable name \( l \).
+The input process \( \Pin{x}{l}{P} \) waits for a base value from \( x \) and then continues as \( P \) with the received value substituted for the variable name \( l \).
+%
+The process \( \Presd{x}{y}{P} \) represents a \emph{session} with endpoints named \( x \) and \( y \) which are bound in \( P \).
+In \( P \), the names \( x \) and \( y \) can be used to exchange messages over the session (sending on \( x \) and receiving on \( y \) or vice versa).
+%
+Finally, the process $\PBpar{P}{Q}$ represents the parallel composition of $P$ and $Q$.
+%
+Note that in this calculus channels cannot be sent in messages, therefore the topology of the communication network described by a process cannot change.
+Also, there is no recursion or replication in the syntax, hence no infinite behaviors can be expressed. In particular, we only model linear (as opposed to shared) channels.
 
-The process \( \Presd{x}{y}{P} \) is the \emph{restriction} of the channel with endpoints named \( x \) and \( y \) to \( P \).
-Components in \( P \) can use the names \( x \) and \( y \) to interact with each other (sending on \( x \) and receiving on \( y \) or vice versa), but not with processes outside of the restriction.
-The restriction operator thus binds the names \( x \) and \( y \) in \( P \).
-Note that the scope of a restriction may not change when processes interact, since it is only possible to send and receive values, and not names.
-Note that there is no recursion of replication in the syntax, and thus no infinite behaviors can be expressed.
-
-\subsubsection{Semantics}
-We describe the actions that the system can perform using a small step operational semantics.
-To simplify the definition of the reduction relation, we will  factor out a \emph{structural congruence} relation on processes, namely  the smallest congruence relation that satisfies the following axioms:
+\subsubsection{Semantics.}
+We describe the actions that the system can perform through a small step operational semantics.
+As usual, we make use of a \emph{structural congruence} relation that equates processes that we deem to be indistinguishable.
+Structural congruence is the smallest congruence relation that satisfies the following axioms:
 \begin{mathpar}
@@
 \end{mathpar}
-Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+
+Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.\luca{Unclear to me.
+Since it is a reduction semantics it's obvious that input/output actions alone cannot do anything.
+I would remove this sentence}
+%
+Note that reductions are allowed only for restricted pairs of session endpoints.
+This makes it possible to formulate subject reduction so that the typing context is exactly the same before and after each reduction.
+%
 Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
@@
 
-\subsubsection{Session types}
-Our process syntax allows us to write processes that are not well-formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
-As an example, the following process attempts to send a base value on both \( x \) and \( y\), whereas the intention of restricting names is that one of the names is used for sending and the other for receiving: \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \).
-Another example is the following process, which attempts to send a variable that is not instantiated at the time of sending: \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \).
-
-To alleviate this issue, we introduce a \emph{session type system}, which will detect processes that are not well-formed.
+\subsubsection{Session types.}
+Our process syntax allows us to write processes that are ill formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
+As an example, the process \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \) attempts to send a base value on both \( x \) and \( y\), whereas one of the names should be used for receiving in order to guarantee progress.
+Another example is the process \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \), which attempts to send a variable that is not instantiated at the time of sending.
+%
+To prevent these issues, we introduce a \emph{session type system} which
+rules out ill-formed processes.
 
@@
 \]
-The \emph{end type}, \( \Tend \), is the type of an endpoint on which no further interaction is possible.
-The \emph{base type}, \( \Tbase \), is the type of base values.
-The \emph{input type}, \( \Tin{S} \), is the type of an endpoint which is ready to receive a value, then continue with type \( S \).
-The \emph{output type}, \( \Tout{S} \), is the type of an endpoint which is ready to send a value, then continue with type \( S \).
+The \emph{end type} \( \Tend \) describes an endpoint on which no further interaction is possible.
+The \emph{base type} \( \Tbase \) describes base values.
+The \emph{input type} \( \Tin{S} \) describes endpoints used for receiving a value and then according to \( S \).
+The \emph{output type} \( \Tout{S} \) describes endpoints used for sending a value and then according to \( S \).
 
 Typing contexts gather type information about names and variables.
-We distinguish between an unrestricted context $\Gamma$ seen as a set: since  we have only one base type, we only record which variable names have been used.
-Type information for endpoints is captured in the \emph{linear}
-context \( \Delta \). We use the comma as split/union, overloaded to
-singletons, and \( \Cempty \) is the empty multiset.
-We extend the Barendregt convention to contexts, so that all entries
-are distinct.  Note that the order in which information is added to a
-type context thus does not matter.
+\emph{Unrestricted} contexts are simply sets of names since we only have one
+base type. \emph{Linear contexts} associate a type to endpoints. We use
+the comma as split/union, overloaded to singletons, and \( \Cempty \) is the
+empty multiset. We extend the Barendregt convention to contexts, so that all
+entries are distinct.  Note that the order in which information is added to a
+type context does not matter.
 
-Since we need to determine whether endpoints are used dually to determine whether processes are well-formed, we will need to formally define the dual of a type.
-The dual of a type, \( \dual{T} \), is defined by the following equations:
+Since we need to determine whether endpoints are used in complementary ways to determine whether processes are well formed, we need to formally define the dual of a type as follows:
 \begin{mathpar}
@@
 \end{mathpar}
-Note that the dual function is not total, since it is not defined for the base type.
+Note that the dual function is partial since it is undefined for the base type.
 
 \subsubsection{Typing rules}
-Our type system will maintain two invariants:
+Our type system is aimed at maintaining two invariants:
 \begin{enumerate}
-\item Endpoints are used exactly once
-\item Endpoints that are part of the same restriction have dual types
+\item No endpoint is used simultaneously by parallel processes;
+\item The two endpoints of the same session have dual types.
 \end{enumerate}
@@
 \begin{proof}
-  Induction on the reduction relation, TBD.
+  By induction on the derivation of $\reduces{P}{Q}$. TBD
 \end{proof}
@@
 
-To formulate safety, we need to formally define what we mean by well-formed.
+To formulate safety, we need to formally define what we mean by well-formed process.
 We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
-A process is then \emph{well-formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
+A process is then \emph{well formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
 
-Note that well-formed processes do not necessarily reduce, since e.g.\ the process
+Note that well-formed processes do not necessarily reduce. For example, the process
 \begin{equation*}
@@
 \end{equation*}
-is well-formed, but has no reduction.
+is well formed but also irreducible.
 
 \begin{theorem}[Type safety]
-  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
+  If \( \types{\Cempty}{P} \), then \( P \) is well formed.
 \end{theorem}
@@
 \begin{corollary}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well-formed.
+  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well formed.
 \end{corollary}

### 2023-10-11 challenges/preliminaries.tex
Remove choice/sums from scope extrusion challenge.
a couple of metanotes
fixing a typo in T-IN
fixing the syntax of substitution

@@
 The two components can proceed independently of each other, or they can interact via shared names.
-The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
+%The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
 
@@
 % (by an input)
-in \( P \).  We will use the notation \( \subst{P}{v}{l} \) to denote
-the process \( P \) with value \( v \) substituted for variable name
+in \( P \).  We will use the notation \( \subst{P}{a}{l} \) to denote
+the process \( P \) with base value \( a \) substituted for variable name
 \( l \). Similarly, \( \subst{P}{x}{y} \) denotes the process
 \( P \) with name \( x \) substituted for name \( y \).
+% \begin{metanote}
+% AM:  Should we spell out the definitions of subsitutions?
+% \end{metanote}
 

### 2023-10-12 paper/paper.tex
Revision of intro and section 3 (and section 2, minimally)

@@
 \usepackage{hyperref}
+\usepackage[inline]{enumitem}
 \usepackage{color}
@@
 
-Large scale mechanisation of concurrency theory is already possible,
-\fxwarning{Citations?}
 %(Paragraphs about related work in this vein)
-and several papers in the realm of process calculi do include mechanised proofs
+Several papers in the realm of process calculi already include mechanised proofs
 (e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
@@
 and hampers the reuse of techniques and results.
-
+%
 Therefore, it is our opinion that the research community in message-passing
@@
 of solutions exploring different ways of mechanising typical problems in the
-field. Like the authors of the previous POPLMark challenges, we thus seek to
+field.
+
+Like the authors of the previous POPLMark challenges, we seek to
 answer a number of questions:
-\begin{enumerate}
+\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
 \item What is the current state of the art in mechanising process calculi?
@@
 
-\subsection*{Benchmark design goals and considerations}%
+We have identified three key aspects which typically cause difficulties when
+mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion},
+% in name-passing systems,
+and \emph{coinductive reasoning} for infinite behaviours.
+Our benchmark problems are designed to focus on these aspects (discussed
+in more detail in \cref{sec:challenge-problems}), with three main goals:
 
-We have identified three key aspects which typically cause difficulties when
-mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion} in
-name-passing systems, and \emph{coinductive reasoning} for infinite behaviours.
-Our main goals are:
+\begin{enumerate}[label=\textbf{(G\arabic*)},leftmargin=10mm]
+\item\label{item:goal-comperison-accessibility} Enable the comparison of
+  different proof mechanisation approaches, making the challenges accessible to
+  mechanisation experts who may not be familiar with concurrency theory.
 
-\begin{enumerate}
-\item Enable the comparison of different proof mechanisation approaches, making
-  the challenges accessible to mechanisation experts who may not be familiar
-  with concurrency theory. To this end, we strive to keep the challenge problems
-  orthogonal: our collection of benchmark problems is designed to exercise one
-  of the aforementioned three aspects independently from the others, and include
-  as little syntax and as few definitions as possible.
+\item\label{item:goal-tutorials} Encourage the development of guidelines and
+  tutorials demonstrating and comparing the many available proof mechanisation
+  techniques, libraries, and proof assistant features.
 
-\item Encourage the development of guidelines and tutorials demonstrating and
-  comparing the many available proof mechanisation techniques, libraries, and
-  proof assistant features.  Therefore, our challenge problems strive for
-  minimality, and adopt a simple and uniform notation.
-
-\item Prioritise the exploration of mechanisation techniques that may be useful
-  and reusable for future research. To this end, our problems are drawn from the
-  basic metatheory of process calculi (instead of requiring the development of
-  new theory), and focus on well-known theorems and results that involve
-  interesting proof techniques.
+\item\label{item:goal-reusability} Prioritise the exploration of mechanisation
+  techniques that may be useful and reusable for future research.
 \end{enumerate}
 
-We also see the goals above as a stepping stone towards strengthening the
-culture of mechanisation, by rallying the community to collaborate on exploring
-and developing new tools and techniques.
+We also aim at strengthening the culture of mechanisation, by rallying the
+community to collaborate on exploring and developing new tools and techniques.
 
@@
 course not the only ones that may cause issues in mechanisations --- but they
-are fundamental to the field of concurrency theory, and they emerge in most
-mechanisations.  Moreover, the mechanisation of novel research results often
-requires addressing more than just one of these aspects at once --- and we see
-the combination of techniques as a next step beyond the scope of this challenge
-(we discuss the topic in \cref{sec:going-beyond}).
+are fundamental to concurrency theory, and they emerge in most mechanisations.
+The mechanisation of novel research results often requires addressing more than
+just one of these aspects at once --- and we see the combination of techniques
+as a next step beyond the scope of this challenge (we discuss this in
+\cref{sec:going-beyond}).
 
@@
 \section{The challenge problems}\label{sec:challenge-problems}
-In this section we will briefly describe our three challenge problems, referring to the appendix for the full text of each challenge.
+In this section we briefly describe the three aspects of our challenge problems, referring to the appendix for the full text of each one.
 
 \subsection{Linearity}
-Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a session type system.
-Session type systems require linear handling of typing contexts, and the proof involves reasoning about the linearity of context splitting operations.
+Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a minimal
+session typing system (without delegation of communication channels, to avoid overlapping with the scope
+extrusion challenge below).
+Session type systems require linear handling of typing contexts, and the key
+issue of this challenge is reasoning about the linearity of context splitting operations.
 
 \subsection{Scope extrusion}
-In name passing systems, scope extrusion is the expansion of the scope of a name when the name is sent outside of its original scope.
-Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove the equivalence between reduction semantics and labelled transition system semantics up to structural congruence.
-The proof involves reasoning about names that are ``in the process'' of being scope extruded, which presents issues for some approaches to mechanising binders.
+In name-passing systems, scope extrusion is the expansion of the scope of a name (i.e.~a communication channel) that is sent from one process to another.
+Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove the equivalence between the reduction semantics and the labelled transition system (LTS) semantics of a process calculus (up-to structural congruence).
+The key issue of this challenge is reasoning about names that are ``in the process'' of being scope-extruded, which often presents difficulties for the mechanisation of binders.
 
 \subsection{Coinduction}
-Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that strong barbed congruence acts like a congruence for an untyped calculus of communicating systems with (infinite) replication of processes.
-The proof involves reasoning coinductively about the infinite behaviours of the replication operator.
+Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that \emph{strong barbed bisimilarity} is a congruence for a limited calculus with (infinite) replication of processes.
+The key issue of this challenge is the coinductive reasoning about the infinite behaviours of the replication operator.
 
 \section{Design of the benchmark}\label{sec:design-discussion}
+
 In this section, we outline the factors considered in designing the
@@
 
-We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
-Like previous benchmark authors, we would like to stress that our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types and psi-calculi are not directly covered by our challenge problems.
-Most applications will however still need the basic techniques that our challenge problems exercise.
-We have designed our challenge problems to exercise the three aspects independently, and they can be solved individually and in any order.
-We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.
+% We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
+Like the previous POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types, psi-calculi, or encodings between different calculi are not directly covered.
+Still, these (and other) proof mechanisation applications will need the basic techniques that our challenge problems exercise:
+as per our design goal \ref{item:goal-reusability},
+our problems are drawn from the basic metatheory of process calculi (instead
+of requiring the development of new theory), and focus on well-known theorems
+and results that involve interesting proof techniques that may be reused in further work.
+
+To achieve our design goal \ref{item:goal-comperison-accessibility},
+we have designed our challenge problems to exercise the three aspects (linearity, scope extrusion, coinduction) independently, and they can be solved individually and in any order.
+We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.  The process calculus used in each challenge
+focuses on the features we want to exercise, and we have omitted all constructs
+(such as choices) that would complicate the mechanisation without bringing tangible
+benefits and insights.  (Re-adding the omitted constructs if of course possible, but we leave
+such extensions as a possibility for future benchmark challenge.)
+
+The minimality and uniformity of the calculi also allows us to target our
+design goal \ref{item:goal-tutorials}, i.e.~encouraging the development of
+concurrency theory proof mechanisation guidelines and tutorials.
 
@@
 
-The aim of this challenge is to keep the problems focused on the
-features we want to exercise. For that reason, we have decided not to
-include the choice construct in our challenge problems. Choice
-constructs complicate the theory. For the same reason, each challenge
-problem only includes the fragment necessary to express the theorems
-we are interested in. We leave extensions of the syntax for a future
-benchmark challenge.
-
-Our challenge problems do not include reasoning about encodings. While
-this kind of reasoning is common in the field, it does not require any
-different reasoning techniques in a concurrency setting than in any
-other setting.
+% Our challenge problems do not include reasoning about encodings. While
+% this kind of reasoning is common in the field, it does not require any
+% different reasoning techniques in a concurrency setting than in any
+% other setting.
 

### 2023-10-12 paper/paper.tex
Fix a few small mistakes.

@@
 (such as choices) that would complicate the mechanisation without bringing tangible
-benefits and insights.  (Re-adding the omitted constructs if of course possible, but we leave
-such extensions as a possibility for future benchmark challenge.)
+benefits and insights.  (Adding the omitted constructs is of course possible, but we leave
+such extensions as a possibility for a future benchmark challenge.)
 
@@
 
-Linearity is of course interesting in many other subfields of programming languages, and we have thus chosen to base our challenge on a  session type system primarily due to the common use of session types in concurrency theory.
+Linearity is of course interesting in many other subfields of programming languages, but we have chosen to base our challenge on a session type system primarily due to the common use of session types in concurrency theory.
 
 Inspired by Vasconcelos~\cite{Vasconcelos2012}, we use a syntax where
-restriction binds two variables. This presentation makes the duality
+restriction binds two names. This presentation makes the duality
 of the two names obvious in the type system.
 
-We have simple notion of well-formedness for our challenge problem,
-and does not for instance consider deadlocked processes. While more
+We have chosen a simple notion of well-formedness for our challenge problem,
+and do not for instance consider deadlocked processes. While more
 sophisticated notions of well-formedness are also interesting, the

### 2023-10-12 challenges/coinduction.tex
Fix description of coinduction challenge.

@@
 % semantics of which are presented below.
-The objective of this challenge is to prove that strong barbed
-bisimilarity
-% for this calculus
-is an equivalence relation. Which, to reiterate, entails the use of
-coinduction due to the presence of infinite behaviour.
+The objective of this challenge is to prove a context lemma for strong barbed congruence.
+The context lemma establishes that processes are strong barbed congruent if and only if the processes obtained by applying a substitution to them and composing them with an arbitrary process are strongly barbed bisimilar.
+Which, to reiterate, entails the use of coinduction due to the presence of infinite behaviour.
 

### 2023-10-12 paper/paper.tex
Fix description of coinduction challenge.

@@
 \subsection{Coinduction}
-Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove that \emph{strong barbed bisimilarity} is a congruence for a limited calculus with (infinite) replication of processes.
+Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove a context lemma involving \emph{strong barbed congruence} for a limited calculus with (infinite) replication of processes.
 The key issue of this challenge is the coinductive reasoning about the infinite behaviours of the replication operator.

### 2023-10-12 challenges/linearity.tex
Remove a superfluous note in the linearity challenge.
minor fixes to linearity, plus a metanote

@@
 
+\begin{metanote}
+ AM: do we mean type safety? If so the description is half the story
+\end{metanote}
 The objective of this challenge is to prove type preservation (also
@@
 %
-Finally, the process $\PBpar{P}{Q}$ represents the parallel composition of $P$ and $Q$.
+%Finally, the process $\PBpar{P}{Q}$ represents the parallel composition of $P$ and $Q$.
 %
@@
 We describe the actions that the system can perform through a small step operational semantics.
-As usual, we make use of a \emph{structural congruence} relation that equates processes that we deem to be indistinguishable.
+As usual, we use  a \emph{structural congruence} relation that equates processes that we deem to be indistinguishable.
 Structural congruence is the smallest congruence relation that satisfies the following axioms:
@@
 
-Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.\luca{Unclear to me.
-Since it is a reduction semantics it's obvious that input/output actions alone cannot do anything.
-I would remove this sentence}
-%
 Note that reductions are allowed only for restricted pairs of session endpoints.
@@
 \subsubsection{Session types.}
-Our process syntax allows us to write processes that are ill formed in the sense that they either use the endpoints bound by a restriction to communicate in a way that does not follow the intended duality of communication, or attempt to send something which is not a base value.
-As an example, the process \( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \) attempts to send a base value on both \( x \) and \( y\), whereas one of the names should be used for receiving in order to guarantee progress.
-Another example is the process \( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \), which attempts to send a variable that is not instantiated at the time of sending.
+Our process syntax allows us to write processes that are ill formed in
+the sense that they either use the endpoints bound by a restriction to
+communicate in a way that does not follow the intended duality, or
+attempt to send something which is not a base value.  As an example,
+the process
+\( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \)
+attempts to send a base value on both \( x \) and \( y\), whereas one
+of the names should be used for receiving in order to guarantee
+progress.  Another example is the process
+\( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \),
+which attempts to send a variable that is not instantiated at the time
+of sending.
 %
@@
 base type. \emph{Linear contexts} associate a type to endpoints. We use
-the comma as split/union, overloaded to singletons, and \( \Cempty \) is the
-empty multiset. We extend the Barendregt convention to contexts, so that all
+the comma as split/union, overloaded to singletons, and \( \Cempty \) as the
+empty context. We extend the Barendregt convention to contexts, so that all
 entries are distinct.  Note that the order in which information is added to a
@@
 \end{enumerate}
-The first invariant is maintained by linearly splitting type contexts when typing compositions of processes.
-The second invariant is maintained by requiring duality when typing restrictions.
+The first invariant is maintained by linearly splitting type contexts when typing compositions of processes, the second  by requiring duality when typing restrictions.
 
 We have three typing judgments: one for values, one for names, and one for processes.
-The typing rules for values are as follows:
+The typing rules for values are:
 \begin{mathpar}
@@
 \end{mathpar}
-The typing rule for names is as follows, where $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
+In the typing rule for names, $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
 \begin{mathpar}
@@
 
-The following lemmas may be of use:
+The following lemmas are needed:
 \begin{lemma}[Weakening]\mbox{}
@@
   \item If \( \types{\Gamma; \Delta}{P} \) then
-    \( \types{\Cadd{\Gamma}{l};\Delta}{P} \).
+    \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
       \item If \( \types{\Gamma; \Delta}{P} \) and \( \tend{\Delta} \) then
-    \( \types{\Gamma;\Cadd{\Delta}{\hastype{x}{T}}}{P} \).
+    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{T}})}{P} \).
 \end{enumerate}
@@
   \label{lemma:strenD}
-  If \( \types{\Gamma; \Delta, z : T}{P} \) and $z\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
+  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
 \end{lemma}

### 2023-10-12 challenges/linearity.tex
removing T-Name rule as folded in IN and OUT

@@
 
-We have three typing judgments: one for values, one for names, and one for processes.
+We have two typing judgments: one for values, and one for processes.
 The typing rules for values are:
@@
 \end{mathpar}
-In the typing rule for names, $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
-\begin{mathpar}
-  \inferrule[T-Name]{{\tend\Delta }}{\types{{\Gamma; (\Delta,\hastype{x}{T})}}{\hastype{x}{T}}}
-\end{mathpar}
+% In the typing rule for names, $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
+% \begin{mathpar}
+%   \inferrule[T-Name]{{\tend\Delta }}{\types{{\Gamma; (\Delta,\hastype{x}{T})}}{\hastype{x}{T}}}
+% \end{mathpar}
 The typing rules for processes are as follows:
@@
 \end{mathpar}
+Note that we do not need a judgment for typing channels, since it is already folded into the T-IN and T-OUT rules.
 

### 2023-10-12 paper/paper.tex
Add citations + move some TODOs into FIXMEs

@@
 % We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
-Like the previous POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types, psi-calculi, or encodings between different calculi are not directly covered.
+Like the previous POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
+applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17} are not directly covered.
 Still, these (and other) proof mechanisation applications will need the basic techniques that our challenge problems exercise:
@@
 we have designed our challenge problems to exercise the three aspects (linearity, scope extrusion, coinduction) independently, and they can be solved individually and in any order.
-We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.  The process calculus used in each challenge
-focuses on the features we want to exercise, and we have omitted all constructs
+We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi, and type theory.  The process calculus used in each challenge
+focuses on the features we want to exercise, and omits all constructs
 (such as choices) that would complicate the mechanisation without bringing tangible
-benefits and insights.  (Adding the omitted constructs is of course possible, but we leave
-such extensions as a possibility for a future benchmark challenge.)
+benefits and insights.  (It is of course possible to reintroduce the omitted
+constructs, but we leave this as a possibility for a future benchmark challenge.)
 
-The minimality and uniformity of the calculi also allows us to target our
-design goal \ref{item:goal-tutorials}, i.e.~encouraging the development of
-concurrency theory proof mechanisation guidelines and tutorials.
+The minimality and uniformity of the calculi also allows us to target our design
+goal \ref{item:goal-tutorials}, i.e.~encouraging the development guidelines and
+tutorials for the mechanisation of concurrency theory.
 
@@
 \subsection{Linearity}
+\fxerror{%
 - Linearity type system dual vs. Vasco's?
@@
 - Asynchrony? Not necessary, that is a bit more advanced
+}%
 
@@
 \subsection{Scope extrusion}
+\fxerror{%
 - Why do we keep the restrictions around in the LTS semantics in scope extrusion?
+}
 
@@
 \subsection{Coinduction}
+\fxerror{
 - Coinduction: why is congruence interesting?
 - ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics
+}
 

### 2023-10-12 paper/paper.tex
Update references.bib

@@
 
-As with the challenges that came before, this one is intended to be
-useful on its own, but not to exhaustively cover every possible
-problem in the area. The current challenge can be extended in two
-dimensions: the first is taking the existing challenges further.
-And the second is devising new challenges not covered by the current
-benchmarks.
+As with the POPLMark and POPLMark Reloaded challegnes, this challenge is
+intended to be useful, but it does not aim at covering every possible
+problem in the area. Our benchmark can be extended in two
+dimensions: further expanding and combining the current challenges,
+and addressing calculi and features not covered by the current challenges.
+We discuss these dimensions in the rest if this section.
 
+\paragraph{Extending and combining the current challenges.}
 The proposed challenges try to minimise the overlap of the techniques
-they exercise, however, most systems will combine some or all aspects
-of the challenges. An interesting second round of formalisations could
-entail combining the features of the challenges. In the following
-table we consider the following three additions: First the addition of
-choice to the type level and the process level. Second, the addition
+and aspects they exercise (linearity, scope extrusion, coinduction) --- but most
+research on concurrency theory combines some or all such aspects. An interesting second round of formalisations could
+entail combining the features of the challenges.
+
+In the following table we consider the following three extensions: first, the
+addition of choice to the type level and the process level. Second, the addition
 of recursion (and recursion types for typed systems). And finally, the
 addition of channel delegation to those calculi that do not already
-support it.
-
-The following table shows which extension should be suitable (that we
-mark with $+$) for each challenge:
+support it.  We mark each suitable extension with ``$+$'':
 
@@
 
-The other dimension to extend these calculi corresponds to systems
-that are not covered by our challenges. Extending in this dimension
+\paragraph{Addressing new calculi and new features.}
+Extending in this dimension
 means proposing new challenges to cover for different features of
-message passing calculi, or to cover different features of mechanised
-proof.
+message passing calculi, or to cover different proof mechanisation features.
 
-Some interesting aspects of message passing calculi to explore in
-further challenges could be: Multiparty session types, and
-choreographies because their definitions and meta-theory have elements
-that we do not cover here, however the elements that we do cover will
-appear in these problems. Also, while we explore a form of barbed
-bisimilarity one could easily design a challenge exploring different
-notions of bisimilarity (barbed, or weak, etc.), and trace
-equivalence.
+Some interesting aspects of message passing calculi to explore in further
+challenges could be multiparty session types, and choreographies because their
+definitions and meta-theory have aspects that we do not cover here\fxerror{Such
+as?} (although the aspects that we cover here do appear in these problems).
+Also, one could easily design a challenge that goes beyond our coinduction
+challenge, by exploring different notions of bisimilarity (barbed, or weak,
+etc.), and trace equivalence.
 
@@
 mechanising the proofs of theorems. An interesting venue to explore
-would be to take advantage of other aspects of proof assistants. For
+would be to take advantage of other proof assistant features. For
 example, proof assistants are often able to produce certified code (by
@@
 tools, namely model checkers. These tools are in common use in the
-field. However, combining automated proofs with proof assistants
+field --- and combining automated proofs with proof assistants
 offers the potential to ease the path for attempting larger proofs
@@
 worthwhile directions, and we look forward to a future when they
-indeed are extended. Moreover, we see the current challenge as setting
+indeed are extended. We see the current challenge as setting
 the foundation for those future extensions. It is our expectation that
@@
 
+\fxerror{TODO? Remove?}
+
 \subsubsection*{Acknowledgements}
+
+\fxerror{TODO? Remove?}
 

### 2023-10-12 challenges/coinduction.tex
Change many small things based on Dawit's comments.

@@
 If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more, and here choose to define the observables of a process as the names it can use for sending and receiving.
+We must therefore allow ourselves to observe more than just internal transitions, and here choose to define the observables of a process as the names it can use for sending and receiving.
 

### 2023-10-12 challenges/preliminaries.tex
Change many small things based on Dawit's comments.

@@
 
-We assume the existence of some type of \emph{base values}, represented by the symbols
-\( a, b, \dots \), the existence of some type
+We assume the existence of some set of \emph{base values}, represented by the symbols
+\( a, b, \dots \), the existence of some set
 of \emph{variable names}, represented by the symbols
-\( l, m, \dots \), and the existence of some type of \emph{names},
+\( l, m, \dots \), and the existence of some set of \emph{names},
 represented by the symbols \( x, y, \dots \). \ednote{AM: are we assuming that names can be compared for equality everywhere or only in C2?} %
@@
 The syntax of processes will include:
-the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{parallel composition} of process \( P \) and process \( Q \).
 The two components can proceed independently of each other, or they can interact via shared names.

### 2023-10-12 challenges/linearity.tex

--- 2023-10-12 challenges/linearity.tex
@@
   \end{enumerate}
+  \begin{metanote}
+    Note from Dawit: Does \( \Delta \) really have to be only ends to be weakened?
+I had expected that if \( P \) is typed by \( \Delta \) and \( \Delta' \) is only ends, then \( P \) is typed by \( \Cadd{\Delta}{\Delta'} \).
+  \end{metanote}
 \end{lemma}
@@
 \begin{proof}
-  By induction on the structure of a process context. In the base case, proceed by case analysis on the \textsl{Sc} rule applied:
+  By induction on the structure of a process context involved in the congruence. In the base case where the context is a hole, proceed by case analysis on the \textsl{Sc} rule applied:
   \begin{description}

### 2023-10-12 paper/paper.tex

--- 2023-10-12 paper/paper.tex
@@
 
-We have begun collecting solutions to our challenge problems on our website:
+We have begun collecting solutions to our challenges on our website:
 %
@@
 In the longer term, we expect to use the website for promoting best practices,
-tutorials and guidelines derived from solutions to our challenge problems.
+tutorials and guidelines derived from solutions to our challenges.
 %The interested reader can find more information about the current status at:
-We encourage anyone interested to try the challenge problems using their
+We encourage anyone interested to try the challenges using their
 favourite tools and techniques, and to send us their solutions.
@@
 \cref{sec:going-beyond}.  We conclude in \cref{sec:conclusion}.
-The full challenge problems, including proofs, are available in \cref{app:challenges}.
+The full challenges, including proofs, are available in \cref{app:challenges}.
 
-\section{The challenge problems}\label{sec:challenge-problems}
-In this section we briefly describe the three aspects of our challenge problems, referring to the appendix for the full text of each one.
+\section{The challenges}\label{sec:challenge-problems}
+In this section we briefly describe the three aspects of our challenges, referring to the appendix for the full text of each one.
 
@@
 benchmark challenges. We start with some general remarks, then
-describe the individual design considerations for each challenge
-problem.
+describe the individual design considerations for each challenge problem.
 
@@
 Like the previous POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types, psi-calculi, or encodings between different calculi are not directly covered.
-Still, these (and other) proof mechanisation applications will need the basic techniques that our challenge problems exercise:
+Still, these (and other) proof mechanisation applications will need the basic techniques that our challenges exercise:
 as per our design goal \ref{item:goal-reusability},
@@
 To achieve our design goal \ref{item:goal-comperison-accessibility},
-we have designed our challenge problems to exercise the three aspects (linearity, scope extrusion, coinduction) independently, and they can be solved individually and in any order.
+we have designed our challenges to exercise the three aspects (linearity, scope extrusion, coinduction) independently, and they can be solved individually and in any order.
 We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.  The process calculus used in each challenge
@@
 
-We have chosen a simple notion of well-formedness for our challenge problem,
+We have chosen a simple notion of well-formedness for our challenge,
 and do not for instance consider deadlocked processes. While more
@@
 proofs of these properties are more complicated and would thus make
-the challenge problem more difficult and less focused on the linear
+the challenge more difficult and less focused on the linear
 aspects of the system.
@@
 \subsection{Evaluation criteria}
-Solutions to our challenge problems should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
+Solutions to our challenges should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
 Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges need to present more elaborate argumentation for their adequacy.

### 2023-10-12 challenges/scope-extrusion.tex
Compact BNFs in scope extrusion challenge.

@@
 Reasoning about scope extrusion explicitly can sometimes be avoided by introducing a structural congruence rule into the semantics, but doing this means we lose information about the scope when reasoning about the semantics.
-Explicitly reasoning about scope extrusion is necessary to describe e.g.\ monitors and compositions.
+Explicitly reasoning about scope extrusion is necessary to describe e.g.\ runtime monitors and compositions of systems.
 
@@
 The syntax of processes is given by:
-\begin{align*}
-  P,Q := &&& \Pend \\
-  |&&& \Ppar{P}{Q} \\
-  |&&& \Pout{x}{y}{P} \\
-  |&&& \Pin{x}{y}{P} \\
-  |&&& \Pres{x}{P} \\
-\end{align*}
+\[
+\begin{array}{rcl}
+  P,Q & :=& \Pend \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad \Pout{x}{y}{P} \quad\mid\quad \Pin{x}{y}{P} \quad\mid\quad \Pres{x}{P}
+\end{array}
+\]
 
@@
 The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
-%The input operator thus binds the name \( y \) in \( P \).
+The input operator thus binds the name \( y \) in \( P \).
 % The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
@@
 The transitions are labelled by \emph{actions}, the syntax of which is as follows:
-\begin{align*}
-  \alpha := &&& \Aoutf{x}{y} \\
-  |&&& \Ain{x}{y} \\
-  |&&& \Aoutb{x}{y} \\
-  |&&& \Atau
-\end{align*}
+\[
+\begin{array}{rcl}
+  \alpha & :=& \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
+\end{array}
+\]
+
 The \emph{free output action} \( \Aoutf{x}{y} \) is sending the name \( y \) via \( x \).

### 2023-10-12 paper/paper.tex
fixed a comment in the further challenge section

@@
 
-Some interesting aspects of message passing calculi to explore in further
-challenges could be multiparty session types, and choreographies because their
-definitions and meta-theory have aspects that we do not cover here\fxerror{Such
-as?} (although the aspects that we cover here do appear in these problems).
-Also, one could easily design a challenge that goes beyond our coinduction
-challenge, by exploring different notions of bisimilarity (barbed, or weak,
-etc.), and trace equivalence.
+Some interesting aspects of message passing calculi to explore in
+further challenges could be multiparty session types, and
+choreographies because their definitions and meta-theory have aspects
+that we do not cover here (e.g.: representing well formed global types
+and projection as a partial function). Although, dually, some aspects
+that we cover here do appear in these problems. Also, one could easily
+design a challenge that goes beyond our coinduction challenge, by
+exploring different notions of bisimilarity (barbed, or weak, etc.),
+and trace equivalence.
 

### 2023-10-12 challenges/linearity.tex
putting quad in bnf as in coinduction.tex
Fix linearity challenge introduction.

@@
 
-\begin{metanote}
- AM: do we mean type safety? If so the description is half the story
-\end{metanote}
-The objective of this challenge is to prove type preservation (also
-known as subject reduction) for the type system, i.e.\ that well-typed
+The main objective of this challenge is to prove type preservation (also
+known as subject reduction), i.e.\ that well-typed
 processes can only transition to processes which are also well-typed
 in the same context.
+The second objective is to prove type safety, i.e.\ that well-typed processes are also well-formed in the sense that they do not use endpoints in a non-dual way.
 
@@
 \begin{array}{rcl}
-  v,w & ::= & a \mid l \\
-   P,Q & ::=& \Pend \mid \Pout{x}{v}{P} \mid \Pin{x}{l}{P} \mid \PBpar{P}{Q} \mid  \Presd{x}{y}{P}
+  v,w & ::= & a \quad\mid\quad l \\
+   P,Q & ::=& \Pend \quad\mid\quad \Pout{x}{v}{P} \quad\mid\quad \Pin{x}{l}{P} \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad  \Presd{x}{y}{P}
 \end{array}
@@
   \begin{array}{rcl}
-  S,T & ::= & \Tend \mid \Tbase \mid \Tin{S} \mid \Tout{S} \\
-    \Gamma &::= & \Cempty \mid \Gamma, l \\
-                    \Delta &::= & \Cempty \mid \Cadd{\Delta}{\hastype{x}{S}}
+  S,T & ::= & \Tend \quad\mid\quad \Tbase \quad\mid\quad \Tin{S} \quad\mid\quad \Tout{S} \\
+    \Gamma &::= & \Cempty \quad\mid\quad \Gamma, l \\
+                    \Delta &::= & \Cempty \quad\mid\quad \Cadd{\Delta}{\hastype{x}{S}}
   \end{array}

### 2023-10-12 challenges/preliminaries.tex
rework of the common section

@@
 
-We list here some common notions that we will use in the following. Since the calculi under study are somewhat different, we will note in each section which will apply.
+First, we list some common notions and conventions that we use in the challenges. Since the calculi under study are somewhat different, each section will list the changes that apply.
 
@@
 
-The syntax of processes will include:
+The syntax of processes includes:
 the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{parallel composition} of process \( P \) and process \( Q \).

### 2023-10-12 challenges/linearity.tex
comment to Dawit's comment. I'm still unhappy about the linear system. Weakening is broken in th elinear case

@@
 I had expected that if \( P \) is typed by \( \Delta \) and \( \Delta' \) is only ends, then \( P \) is typed by \( \Cadd{\Delta}{\Delta'} \).
+
+    AM:  I agree that it's weird: we should not be able to weaken a linear context, if not
+    with unrestricted types --- but this is what we need for Sc-Res-Inact to be type preserving:
+    from $\Delta\vdash 0$ we need $\Delta\vdash (\nu x y)0$. BTW, this is also a problem in Vasco's system 
   \end{metanote}
@@
   \begin{description}
-  \item[Res:] use weakening~\ref{lemma:weak} part 2
-  \item[Comm/Assoc:] by rearranging sub-derivations noting that  order does not matter for linear context
-  \item[Inact] Left-to-right by
+  \item[Par-Comm/Assoc:] by rearranging sub-derivations noting that
+    order does not matter for linear contexts
+  \item[Par-Inact] right-to-left by
     strengthening~\ref{lemma:strenD}. Vice-versa, by weakening~\ref{lemma:weak} part 2.
   \item[Res-Par] Follows Vasco's proof by case analysis on $x : T$ being linear and applying weak and stren accordingly.
+  \item[Res-Inact:] use weakening~\ref{lemma:weak} part 2
   \item[Res] order does not matter

### 2023-10-12 challenges/coinduction.tex
added some proof sketches

@@
     equations~\ref{eq:bisim1} and \ref{eq:bisim2} follow trivially.
-  \item Symmetry is similar to the previous case: for both
-    $P\sbbisim{}{} Q$ and $Q\sbbisim{}{} P$ the same conditions apply.
+  \item Symmetry follows immediately since bisimilarity is defined as
+    a {\em symmetric} relation.
   \item For transitivity, we need to prove that if $P\sbbisim{}{} Q$
-    and $Q\sbbisim{}{} R$ then $P\sbbisim{}{} R$. If
+    and $Q\sbbisim{}{} R$ then $P\sbbisim{}{} R$. In order to do so,
+    we prove that the relation
+    $\mathcal R=\{(P,R)\mid \exists Q\text{ such that }P\sbbisim{}{} Q
+    \land Q\sbbisim{}{} R\}$ is a strong barbed bisimulation. Let us
+    assume that $(P,R)\in\mathcal R$. Hence, there exists a $Q$ such
+    that $P\sbbisim{}{} Q$ and $Q\sbbisim{}{} R$. Clearly, if
     $\observable{P}{\mu}$ then, by $P\sbbisim{}{} Q$,
-    $\observable{Q}{\mu}$. Hence, by $Q\sbbisim{}{} R$,
+    $\observable{Q}{\mu}$. And, by $Q\sbbisim{}{} R$,
     $\observable{R}{\mu}$. Moreover, if $\transition{P}{\Atau}{P'}$
     there exists $Q'$ such that $\transition{Q}{\Atau}{Q'}$ and
-    $P'\sbbisim{}{}Q'$. Also, $\transition{R}{\Atau}{R'}$ with
-    $Q'\sbbisim{}{}R'$. Coinductively, $P'\sbbisim{}{}R'$, hence
-    $P\sbbisim{}{}R$.
+    $P'\sbbisim{}{}Q'$. And also, $\transition{R}{\Atau}{R'}$ with
+    $Q'\sbbisim{}{}R'$. Finally, by definition of $\mathcal R$,
+    $(P',R')\in\mathcal R$.
   \end{itemize}
@@
 \begin{lemma}
-  \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \).
+  \( \sbcong{}{} \) is the largest congruence included in
+  \( \sbbisim{}{} \).
 \end{lemma}
+\begin{proof}
+  By definition, \( \sbcong{}{} \) is a congruence and is included in
+  \( \sbbisim{}{} \). Also, if it were strictly included in another
+  relation then we would get a contradiction since it must be closed
+  under all contexts.
+\end{proof}
 
@@
 \begin{theorem}
-  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and substitution \( \sigma \), \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).
+  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and
+  substitution \( \sigma \),
+  \(
+  \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}}
+  \).
 \end{theorem}

### 2023-10-13 challenges/coinduction.tex
done with a pass for coinduction

@@
-\subsection{Challenge: Coinduction and reasoning about infinite processes}
+\subsection{Challenge: Coinduction}
+% and reasoning about infinite processes}
 \label{sec:challenge:coinduction}
 
-\marco{Why reasoning is only about infinite processes? Would it be
-  unreasonable to shorten the title to {\em Coinduction and infinite
-    processes}?}
+% \marco{Why reasoning is only about infinite processes? Would it be
+%   unreasonable to shorten the title to {\em Coinduction and infinite
+%     processes}?}
 
@@
 % Coinduction is a proof technique for infinite structures,
-Coinduction is a technique for defining and proving properties of
-infinite structures,
+This is usually connected to coinductive definitions where an infinite
+structure is defined as greatest fixed points of a recursive
+definition.
+%
+Coinduction is then a technique for defining and proving properties of
+such infinite structures,
 %
@@
 %
-\marco{Should we be more specific here and give a deeper explanation
-  of what coinduction is? If not, then why do we even say that it's
-  the dual of induction?}
+% \marco{Should we be more specific here and give a deeper explanation
+%   of what coinduction is? If not, then why do we even say that it's
+%   the dual of induction?}
 %
@@
 % semantics of which are presented below.
-The objective of this challenge is to prove a context lemma for strong barbed congruence.
-The context lemma establishes that processes are strong barbed congruent if and only if the processes obtained by applying a substitution to them and composing them with an arbitrary process are strongly barbed bisimilar.
-Which, to reiterate, entails the use of coinduction due to the presence of infinite behaviour.
+The objective of this challenge is to % prove % a context theorem for
+draw a formal connection between strong barbed congruence and strong
+barbed bisimilarity.  % I.e., The context lemma
+The result establishes that two processes are strong barbed congruent
+if and only if the processes obtained by applying a substitution to
+them and composing them in parallel with an arbitrary process are
+strongly barbed bisimilar.  Which, to reiterate, entails the use of
+coinduction due to the presence of infinite behaviour.
 
@@
 \end{displaymath}
+Note that, unlike the process calculus from
+Section~\ref{sec:challenge:linearity-beh-types}, communications occur
+over the same channel name. Moreover, since replication $!P$ allows
+for infinite copies of the process $P$, processes can dynamically
+create an infinite number of names during execution when combined with
+restriction $\Pres{x}{}{P}$.
+%
 % \begin{align*}
@@
 
-\paragraph{Semantics.} We give an LTS semantics for this challenge. As
-an alternative, it is also possible to adopt a reduction semantics
-which features less rules but needs a structural congruence. The
-semantics of the system describes the actions that the system can
-perform by defining a labelled transition relation on processes.  The
-transitions are labelled by \emph{actions}, the syntax of which are as
-follows:
+\paragraph{Semantics.} We choose to give an LTS semantics for this
+challenge. As an alternative, it is also possible to adopt a reduction
+semantics which features less rules but needs a structural
+congruence. The semantics of the system describes the actions that the
+system can perform by defining a labelled transition relation on
+processes.  The transitions are labelled by \emph{actions}, the syntax
+of which are as follows:
 % \begin{align*}
@@
 internal communication.
-
-We will use the notation \( \names{\alpha} \) to denote the set of names that occur in the action \( \alpha \).
+%
+We will use the notation \( \names{\alpha} \) to denote the set of
+names that occur in the action \( \alpha \).
 
@@
 \subsubsection{Bisimilarity}
-Our notion of process equivalence relations builds on a notion of \emph{observables}.
-If we allow ourselves only to observe internal transitions, we will relate either too few processes (in the strong case where we relate only processes with exactly the same number of internal transitions) or every process (in the weak case where we relate processes with any amount of internal transitions).
-We must therefore allow ourselves to observe more than just internal transitions, and here choose to define the observables of a process as the names it can use for sending and receiving.
+Bisimilarity is a notion of equivalence for processes and
+% .  Our notion of process equivalence relations
+builds on a notion of \emph{observable}, i.e., what can we externally
+observe from the semantics of a process and can we use that to tell
+two different process apart.  If we allow ourselves only to observe
+internal transitions, i.e., observe that a process is internally
+performing a step of computation, we will relate either too few
+processes (in the strong case where we relate only processes with
+exactly the same number of internal transitions) or every process (in
+the weak case where we relate processes with any amount of internal
+transitions).  We must therefore allow ourselves to observe more than
+just internal transitions, and here choose to define the observables
+of a process as the names it can use for sending and receiving.
 
-To this end, we define the \emph{observability predicate} \( \observable{P}{\mu} \) by the following rules:
+To this end, we define the \emph{observability predicate}
+\( \observable{P}{\mu} \) as follows:
 \begin{align}
@@
 
-Unfortunately, strong barbed bisimilarity is not a good notion of
-process equivalence, since it does not consider the environment of
-processes.  For instance,
-\( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \)
-since \( \obsout{x} \) is the only observable in both processes and
-they cannot perform a \( \Atau \)-action, but
-\(
-\nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}}
-\) since the left process can perform a \( \Atau \)-action such that
+Unfortunately, strong barbed bisimilarity is not a good process
+equivalence since it is not a congruence. This is because we wish to
+be able to substitute a process with an equivalent one in any context
+it may appear.
+% equivalent processes to be since it does not consider the
+% environment of processes.
+For instance, the processes $\Pout{x}{a}{\Pout{y}{b}{\Pend}}$ and
+$\Pout{x}{a}{\Pend}$ are strong barbed bisimilar, i.e.,
+$\sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}}$. This
+is because \( \obsout{x} \) is the only observable in both processes
+and they cannot perform a \( \Atau \)-action. However, if composed in
+parallel with the process $\Pin{x}{l}{\Pend}$, the relation will no
+longer hold. In fact,
+$\nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}\Ppar{}{\Pin{x}{l}{\Pend}}}$
+%
+because the left process can perform a \( \Atau \)-action such that
 \( \obsout{y} \) becomes observable, whereas the right process cannot.
+
+% For instance,
+% \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \)
+% since \( \obsout{x} \) is the only observable in both processes and
+% they cannot perform a \( \Atau \)-action, but
+% \(
+% \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}}
+% \) since the left process can perform a \( \Atau \)-action such that
+% \( \obsout{y} \) becomes observable, whereas the right process cannot.
 
@@
 \subsubsection{Strong barbed congruence}
-A congruence is exactly what we need to make strong barbed bisimilarity consider the environment (i.e.\ the context) of processes.
+In order to make sure that strong barbed bisimilarity detects cases
+like the one above, we need to restrict it so that it becomes a
+congruence,
+i.e., % we need to make strong barbed bisimilarity consider
+it has to consider the environment in which processes may be placed.
+% (i.e.\ the context) of processes.
 
@@
 \subsubsection{Challenge}
-The objective of this challenge is to prove the following theorem:
+The objective of this challenge is to prove a theorem that shows that
+making strong barbed bisimilarity sensitive to substitution and
+parallel composition is equivalent to strong barbed
+congruence. Formally,
 \begin{theorem}

### 2023-10-13 paper/paper.tex

@@
 
-In the following table we consider the following three extensions: first, the
+In \cref{tab:extensions} we consider the following three extensions: first, the
 addition of choice to the type level and the process level. Second, the addition
@@
 addition of channel delegation to those calculi that do not already
-support it.  We mark each suitable extension with ``$+$'':
+support it.
 
-\vspace{.5em}
-
-\begin{center}\small
+\begin{table}
+  \caption{Possible extensions of the challenges. Each suitable extension is marked ``+''. \label{tab:extensions}}
+  \begin{center}
     \begin{tabular}{|r|c|c|c|}
       \hline
-    & Linearity & \shortstack{Scope \\ extrusion} & Coinduction \\
+      & Linearity & Scope extrusion & Coinduction \\
       \hline
@@
       \hline
-    \shortstack{Recursive \\ types} & $+$  & $+$ & \\
+      Recursive types & $+$  & $+$ & \\
       \hline
@@
   \end{center}
-
-\vspace{.5em}
+\end{table}
 

### 2023-10-13 challenges/coinduction.tex
Fix a few small issues.

@@
 challenge. As an alternative, it is also possible to adopt a reduction
-semantics which features less rules but needs a structural
+semantics which features fewer rules but needs a structural
 congruence. The semantics of the system describes the actions that the
@@
 % .  Our notion of process equivalence relations
-builds on a notion of \emph{observable}, i.e., what can we externally
-observe from the semantics of a process and can we use that to tell
-two different process apart.  If we allow ourselves only to observe
+builds on a notion of \emph{observables}, i.e., what we can externally
+observe from the semantics of a process. If we allow ourselves only to observe
 internal transitions, i.e., observe that a process is internally
@@
 \subsubsection{Strong barbed congruence}
-In order to make sure that strong barbed bisimilarity detects cases
-like the one above, we need to restrict it so that it becomes a
-congruence,
+In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,
 i.e., % we need to make strong barbed bisimilarity consider
-it has to consider the environment in which processes may be placed.
+we have to consider the environment in which processes may be placed.
 % (i.e.\ the context) of processes.

### 2023-10-13 challenges/preliminaries.tex
Remove mention of equality on names since it is obvious and confusing.
Fix a few small issues.
Add definition of applying arbitrary substitution.

@@
 \( l, m, \dots \), and the existence of some set of \emph{names},
-represented by the symbols \( x, y, \dots \). \ednote{AM: are we assuming that names can be compared for equality everywhere or only in C2?} %
+represented by the symbols \( x, y, \dots \).
+We assume that all of these sets are infinite and that their elements can be compared for equality.
 
@@
 \( P \) with name \( x \) substituted for name \( y \).
-% \begin{metanote}
-% AM:  Should we spell out the definitions of subsitutions?
-% \end{metanote}
+We will use the notation \( \applysubst{\sigma}{P} \) to denote the process \( P \) with a finite number of arbitrary substitutions applied to it.
 

### 2023-10-13 challenges/linearity.tex
fixing the statement of weakening and the proof of cong preservation in linearity

@@
       \item If \( \types{\Gamma; \Delta}{P} \) and \( \tend{\Delta} \) then
-    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{T}})}{P} \).
+    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
   \end{enumerate}
-  \begin{metanote}
-    Note from Dawit: Does \( \Delta \) really have to be only ends to be weakened?
+  % \begin{metanote}
+  %   Note from Dawit: Does \( \Delta \) really have to be only ends to be weakened?
 I had expected that if \( P \) is typed by \( \Delta \) and \( \Delta' \) is only ends, then \( P \) is typed by \( \Cadd{\Delta}{\Delta'} \).
 
-    AM:  I agree that it's weird: we should not be able to weaken a linear context, if not
-    with unrestricted types --- but this is what we need for Sc-Res-Inact to be type preserving:
-    from $\Delta\vdash 0$ we need $\Delta\vdash (\nu x y)0$. BTW, this is also a problem in Vasco's system 
-  \end{metanote}
+  %   AM:  I agree that it's weird: we should not be able to weaken a linear context, if not
+  %   with unrestricted types --- but this is what we need for Sc-Res-Inact to be type preserving:
+  %   from $\Delta\vdash 0$ we need $\Delta\vdash (\nu x y)0$. BTW, this is also a problem in Vasco's system 
+  % \end{metanote}
 \end{lemma}
@@
   \item[Par-Inact] right-to-left by
-    strengthening~\ref{lemma:strenD}. Vice-versa, by weakening~\ref{lemma:weak} part 2.
+    strengthening~\ref{lemma:strenD}. Vice-versa, by picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
   \item[Res-Par] Follows Vasco's proof by case analysis on $x : T$ being linear and applying weak and stren accordingly.

### 2023-10-13 paper/paper.tex
Fix small issues.

@@
 \begin{abstract}
-  The POPLMark challenge and its follow-up POPLMark Reloaded sparked a flurry of
+  The POPLMark challenge and its follow-up POPLMark Re\-loaded sparked a flurry of
   work on machine-checked proofs, and fostered the adoption of proof
@@
 
-Like the authors of the previous POPLMark challenges, we seek to
+Like the authors of the POPLMark challenges, we seek to
 answer a number of questions:
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
-\item What is the current state of the art in mechanising process calculi?
-\item What techniques and best practices can be recommended when starting formalisation projects within concurrency theory?
-\item What improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
+\item\label{item:rq1} What is the current state of the art in mechanising process calculi?
+\item\label{item:rq2} What techniques and best practices can be recommended when starting formalisation projects within concurrency theory?
+\item\label{item:rq3} What improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
 \end{enumerate}
@@
 \subsection{Evaluation criteria}
+The idea of our benchmark is to obtain evidence towards answering questions \cref{item:rq1,item:rq2,item:rq3}, and we are thus interested in not only the solutions, but in the experience of solving the challenges with the chosen approach.
 Solutions to our challenges should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.

### 2023-10-13 paper/paper.tex
Clean up paper.tex.

@@
 
-%(Paragraphs about related work in this vein)
 Several papers in the realm of process calculi already include mechanised proofs
@@
 mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion},
-% in name-passing systems,
 and \emph{coinductive reasoning} for infinite behaviours.
@@
 tutorials and guidelines derived from solutions to our challenges.
-%The interested reader can find more information about the current status at:
 We encourage anyone interested to try the challenges using their
@@
 \cref{sec:design-discussion}, and possible extensions and further topics in
-\cref{sec:going-beyond}.  We conclude in \cref{sec:conclusion}.
+\cref{sec:going-beyond}.
 The full challenges, including proofs, are available in \cref{app:challenges}.
@@
 \subsection{Linearity}
-\fxerror{%
-- Linearity type system dual vs. Vasco's?
-- Linearity: the context is polymorphic
-- Asynchrony? Not necessary, that is a bit more advanced
-}%
-
 For our challenge on linear reasoning we have chosen a safety theorem
@@
 \subsection{Scope extrusion}
-\fxerror{%
-- Why do we keep the restrictions around in the LTS semantics in scope extrusion?
-}
-
 This is the challenge most closely related to the original POPLMark
@@
 \subsection{Coinduction}
-\fxerror{
-- Coinduction: why is congruence interesting?
-- ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics
-}
+%\fxerror{
+% ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics?
+%}
 
@@
 
-\section{Conclusion}\label{sec:conclusion}
-
-\fxerror{TODO? Remove?}
-
-\subsubsection*{Acknowledgements}
-
-\fxerror{TODO? Remove?}
-
 \appendix

### 2023-10-13 paper/paper.tex
Anonymize paper.

@@
 \title{The Concurrent Calculi Formalisation Benchmark}
-
-\author{
-     Marco Carbone \inst{1}\orcidID{0000-0001-9479-2632}
-\and David Castro-Perez \inst{2}\orcidID{0000-0002-6939-4189}
-\and Francisco Ferreira \inst{3}\orcidID{0000-0001-8494-7696}
-\and Lorenzo Gheri \inst{4}\orcidID{0000-0002-3191-7722}
-\and Frederik Krogsdal Jacobsen \inst{5}\orcidID{0000-0003-3651-8314}
-\and Alberto Momigliano \inst{6}\orcidID{0000-0003-0942-4777}
-\and Luca Padovani \inst{7}\orcidID{0000-0001-9097-1297}
-\and Alceste Scalas \inst{5}\orcidID{0000-0002-1153-6164}
-\and Martin Vassor \inst{8}\orcidID{0000-0002-2057-0495}
-\and Nobuko Yoshida \inst{8}\orcidID{0000-0002-3925-8557}
-\and Daniel Zackon \inst{9}\orcidID{0009-0008-6153-2955}
-}
-
-\institute{
-     IT University of Copenhagen, Copenhagen, Denmark \email{maca@itu.dk}
-\and University of Kent, Canterbury, United Kingdom \email{D.Castro-Perez@kent.ac.uk}
-\and Royal Holloway, University of London, Egham, United Kingdom \email{Francisco.FerreiraRuiz@rhul.ac.uk}
-\and University of Liverpool, Liverpool, United Kingdom \email{Lorenzo.Gheri@liverpool.ac.uk}
-\and Technical University of Denmark, Kgs. Lyngby, Denmark \email{fkjac@dtu.dk}, \email{alcsc@dtu.dk}
-\and Università degli Studi di Milano, Milan, Italy \email{momigliano@di.unimi.it}
-\and Università di Camerino, Camerino, Italy \email{luca.padovani@unicam.it}
-\and University of Oxford, Oxford, United Kingdom \email{martin.vassor@cs.ox.ac.uk}, \email{nobuko.yoshida@cs.ox.ac.uk}
-\and McGill University, Montreal, Canada \email{daniel.zackon@mail.mcgill.ca}
-}
-
-\authorrunning{M. Carbone et al.}
 

### 2023-10-13 paper/paper.tex
Move challenges into paper text and delete appendix.

@@
 \section{Introduction}
-
 The abstract of the influential POPLMark challenge~\cite{POPLMark} starts with
@@
 field.
-
 Like the authors of the POPLMark challenges, we seek to
@@
 \end{enumerate}
-
 We have identified three key aspects which typically cause difficulties when
@@
 \end{enumerate}
-
 We also aim at strengthening the culture of mechanisation, by rallying the
@@
 \paragraph{Outline of this paper.}
-In \cref{sec:challenge-problems} we describe in more detail the three key
-aspects of our challenge problems.  We discuss our design choices in
-\cref{sec:design-discussion}, and possible extensions and further topics in
-\cref{sec:going-beyond}.
-The full challenges, including proofs, are available in \cref{app:challenges}.
+In \cref{sec:challenge-problems} we describe our three challenges.
+We discuss our design choices in \cref{sec:design-discussion}, and possible extensions and further topics in \cref{sec:going-beyond}.
 
 \section{The challenges}\label{sec:challenge-problems}
-In this section we briefly describe the three aspects of our challenges, referring to the appendix for the full text of each one.
+\label{sec:prelim}
+First, we list some common notions and conventions that we use in the challenges. Since the calculi under study are somewhat different, each section will list the changes that apply.
 
-\subsection{Linearity}
-Our \textbf{linearity challenge} (Appendix~\ref{sec:challenge:linearity-beh-types}) is to prove type preservation (also known as subject reduction) for a process calculus with a minimal
-session typing system (without delegation of communication channels, to avoid overlapping with the scope
-extrusion challenge below).
-Session type systems require linear handling of typing contexts, and the key
+We assume the existence of some set of \emph{base values}, represented by the symbols
+\( a, b, \dots \), the existence of some set
+of \emph{variable names}, represented by the symbols
+\( l, m, \dots \), and the existence of some set of \emph{names},
+represented by the symbols \( x, y, \dots \).
+We assume that all of these sets are infinite and that their elements can be compared for equality.
+
+The syntax of processes includes:
+the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{parallel composition} of process \( P \) and process \( Q \).
+The two components can proceed independently of each other, or they can interact via shared names.
+%The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
+
+For communication, processes include \emph{input} and \emph{output}, whose
+signature will depend on the calculus being value or name-passing.
+We use here the metavariables $c,k$ to abstract over this choice.
+The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
+\( c \) via \( x \), then continue as \( P \).  % The intention is that
+% the value \( v \) must be a base value when it is actually sent, and
+% this will be enforced in the semantics later on
+The process \( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
+via \( x \), then continue as \( P \) with the received element
+substituted for \( k \).  The input operator thus
+binds \( k \) in \( P \).
+
+The process \( \Pres{x}{P} \) is the \emph{restriction} of the name
+\( x \) to \( P \), binding \( x \) in \( P \).
+
+The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
+It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
+Replication makes it possible to express infinite behaviors.
+
+We will use the notation \( \freenames{P} \) to denote the set of
+names that occur free,
+% (i.e.\ not bound by a restriction) in \( P \),
+\( \boundnames{P} \) to denote the set of names that occur bound
+% (by a restriction)
+in \( P \) and  \( \freevars{P} \)
+to denote the set of variable names that occur free
+% (i.e.\ not bound by an input)
+in \( P \).  We will use the notation \( \boundvars{P} \)
+for the set of variable names that occur bound
+% (by an input)
+in \( P \).  We will use the notation \( \subst{P}{a}{l} \) to denote
+the process \( P \) with base value \( a \) substituted for variable name
+\( l \). Similarly, \( \subst{P}{x}{y} \) denotes the process
+\( P \) with name \( x \) substituted for name \( y \).
+We will use the notation \( \applysubst{\sigma}{P} \) to denote the process \( P \) with a finite number of arbitrary substitutions applied to it.
+
+Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
+written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \)
+by a finite number of substitutions of bound variable names.  As a
+convention, we will identify \( \alpha \)-convertible processes and we
+assume that bound names and bound variable names of any processes are
+chosen to be different from the names and variable names that occur
+free in any other entities under consideration, such as processes,
+substitutions, and sets of names or variable names.  This is justified
+because any overlapping names and variable names may be
+\( \alpha \)-converted such that the assumption is satisfied.
+
+A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+
+We will think of contexts as functions between processes.
+A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+
+We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+
+\subsection{Challenge: Linearity and behavioral type systems}
+\label{sec:challenge:linearity-beh-types}
+This challenge formalises a proof that requires reasoning about linearity of channels.
+Linearity is the notion that a channel must be used exactly once in a process.
+This is necessary to prove properties about session type systems, and the key
 issue of this challenge is reasoning about the linearity of context splitting operations.
+Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
 
-\subsection{Scope extrusion}
-In name-passing systems, scope extrusion is the expansion of the scope of a name (i.e.~a communication channel) that is sent from one process to another.
-Our \textbf{scope extrusion challenge} (Appendix~\ref{sec:challenge:name-passing-scope-extrusion}) is to prove the equivalence between the reduction semantics and the labelled transition system (LTS) semantics of a process calculus (up-to structural congruence).
+The setting for this challenge is a small calculus with a session type
+system, the syntax and semantics of which are given below. The
+calculus is a fragment of the one presented in~\cite{Vasconcelos2012},
+formulated in the dual style of~\cite{barber96tr}.
+
+The main objective of this challenge is to prove type preservation (also
+known as subject reduction), i.e.\ that well-typed
+processes can only transition to processes which are also well-typed
+in the same context.
+The second objective is to prove type safety, i.e.\ that well-typed processes are also well-formed in the sense that they do not use endpoints in a non-dual way.
+
+\subsubsection{Syntax.}
+The syntax is given by the grammar
+\[
+\begin{array}{rcl}
+  v,w & ::= & a \quad\mid\quad l \\
+   P,Q & ::=& \Pend \quad\mid\quad \Pout{x}{v}{P} \quad\mid\quad \Pin{x}{l}{P} \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad  \Presd{x}{y}{P}
+\end{array}
+\]
+where a \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
+
+The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \) and then continues as \( P \).
+The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
+The input process \( \Pin{x}{l}{P} \) waits for a base value from \( x \) and then continues as \( P \) with the received value substituted for the variable name \( l \).
+%
+The process \( \Presd{x}{y}{P} \) represents a \emph{session} with endpoints named \( x \) and \( y \) which are bound in \( P \).
+In \( P \), the names \( x \) and \( y \) can be used to exchange messages over the session (sending on \( x \) and receiving on \( y \) or vice versa).
+%
+%Finally, the process $\PBpar{P}{Q}$ represents the parallel composition of $P$ and $Q$.
+%
+Note that in this calculus channels cannot be sent in messages, therefore the topology of the communication network described by a process cannot change.
+Also, there is no recursion or replication in the syntax, hence no infinite behaviors can be expressed. In particular, we only model linear (as opposed to shared) channels.
+
+\subsubsection{Semantics.}
+We describe the actions that the system can perform through a small step operational semantics.
+As usual, we use  a \emph{structural congruence} relation that equates processes that we deem to be indistinguishable.
+Structural congruence is the smallest congruence relation that satisfies the following axioms:
+\begin{mathpar}
+  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
+  \and
+  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
+  \and
+  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
+  \and
+  \inferrule[Sc-Res-Par]{\{x,y\} \cap \freenames{Q} = \emptyset}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
+  \and
+  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
+  \and
+  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
+\end{mathpar}
+
+The operational semantics is then defined as the following binary relation on processes:
+\begin{mathpar}
+  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{a}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{a}{l}}{R}})}}}
+  \and
+  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
+  \and
+  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
+  \and
+  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
+\end{mathpar}
+
+Note that reductions are allowed only for restricted pairs of session endpoints.
+This makes it possible to formulate subject reduction so that the typing context is exactly the same before and after each reduction.
+%
+Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
+On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable name or a name.
+
+\subsubsection{Session types.}
+Our process syntax allows us to write processes that are ill formed in
+the sense that they either use the endpoints bound by a restriction to
+communicate in a way that does not follow the intended duality, or
+attempt to send something which is not a base value.  As an example,
+the process
+\( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \)
+attempts to send a base value on both \( x \) and \( y\), whereas one
+of the names should be used for receiving in order to guarantee
+progress.  Another example is the process
+\( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \),
+which attempts to send a variable that is not instantiated at the time
+of sending.
+%
+To prevent these issues, we introduce a \emph{session type system} which
+rules out ill-formed processes.
+
+\paragraph{Syntax}
+Our type system does not type processes directly, but instead focuses on the channels used in the process.
+The syntax of \emph{session types} \( S, T \), unrestricted type contexts \( \Gamma \) and linear type contexts \( \Delta \) is as follows:
+\[
+  \begin{array}{rcl}
+  S,T & ::= & \Tend \quad\mid\quad \Tbase \quad\mid\quad \Tin{S} \quad\mid\quad \Tout{S} \\
+    \Gamma &::= & \Cempty \quad\mid\quad \Gamma, l \\
+                    \Delta &::= & \Cempty \quad\mid\quad \Cadd{\Delta}{\hastype{x}{S}}
+  \end{array}
+\]
+The \emph{end type} \( \Tend \) describes an endpoint on which no further interaction is possible.
+The \emph{base type} \( \Tbase \) describes base values.
+The \emph{input type} \( \Tin{S} \) describes endpoints used for receiving a value and then according to \( S \).
+The \emph{output type} \( \Tout{S} \) describes endpoints used for sending a value and then according to \( S \).
+
+Typing contexts gather type information about names and variables.
+\emph{Unrestricted} contexts are simply sets of names since we only have one
+base type. \emph{Linear contexts} associate a type to endpoints. We use
+the comma as split/union, overloaded to singletons, and \( \Cempty \) as the
+empty context. We extend the Barendregt convention to contexts, so that all
+entries are distinct.  Note that the order in which information is added to a
+type context does not matter.
+
+Since we need to determine whether endpoints are used in complementary ways to determine whether processes are well formed, we need to formally define the dual of a type as follows:
+\begin{mathpar}
+  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
+  \and
+  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
+  \and
+  \inferrule{}{\dual{\Tend} = \Tend}
+\end{mathpar}
+Note that the dual function is partial since it is undefined for the base type.
+
+\subsubsection{Typing rules}
+Our type system is aimed at maintaining two invariants:
+\begin{enumerate}
+\item No endpoint is used simultaneously by parallel processes;
+\item The two endpoints of the same session have dual types.
+\end{enumerate}
+The first invariant is maintained by linearly splitting type contexts when typing compositions of processes, the second by requiring duality when typing restrictions.
+
+We have two typing judgments: one for values, and one for processes.
+The typing rules for values are:
+\begin{mathpar}
+  \inferrule[T-Base]{ }{\typev{\Gamma}{\hastype{a}{\Tbase}}} \and
+  \inferrule[T-Var]{ }{\typev{\Cadd{\Gamma}l }{\hastype{l}{\Tbase}}}
+\end{mathpar}
+% In the typing rule for names, $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
+% \begin{mathpar}
+%   \inferrule[T-Name]{{\tend\Delta }}{\types{{\Gamma; (\Delta,\hastype{x}{T})}}{\hastype{x}{T}}}
+% \end{mathpar}
+The typing rules for processes are as follows:
+\begin{mathpar}
+
+  \inferrule[T-Inact]{\tend\Delta }{\types{\Gamma;\Delta}{\Pend}}
+  \and
+  \inferrule[T-Par]{\types{\Gamma;\Delta_1}{P} \\ \types{\Gamma;\Delta_2}{Q}}
+  {\types{\Gamma; \Csplit{\Delta_1}{\Delta_2}}{\Ppar{P}{Q}}}
+  \and
+  \inferrule[T-Res]{\types{\Gamma; (\Cadd{\Cadd{\Delta}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P})}{\types{\Gamma}{\Presd{x}{y}{P}}}
+  \and
+    \inferrule[T-Out]{
+      \typev{\Gamma}{\hastype{v}{\Tbase}} \\ \types{\Gamma; \Cupdate{\Delta}{\hastype{x}{T}}}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tout{T}}})}{\Pout{x}{v}{P}}}
+    \and
+    \inferrule[T-IN]{
+      \types{(\Gamma,  l ); (\Cupdate{\Delta}{\hastype{x}{T}})}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tin{T}}})}{\Pin{x}{l}{P}}}
+\end{mathpar}
+Note that we do not need a judgment for typing channels, since it is already folded into the \TirName{T-In} and \TirName{T-Out} rules.
+
+\subsubsection{Challenge}
+The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with:
+\begin{theorem}[Subject reduction]
+  If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
+\end{theorem}
+
+The following lemmas are needed:
+\begin{lemma}[Weakening]\mbox{}
+  \label{lemma:weak}
+  \begin{enumerate}
+  \item If \( \types{\Gamma; \Delta}{P} \) then
+    \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
+      \item If \( \types{\Gamma; \Delta}{P} \) and \( \tend{\Delta} \) then
+    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
+  \end{enumerate}
+  % \begin{metanote}
+  %   Note from Dawit: Does \( \Delta \) really have to be only ends to be weakened?
+I had expected that if \( P \) is typed by \( \Delta \) and \( \Delta' \) is only ends, then \( P \) is typed by \( \Cadd{\Delta}{\Delta'} \).
+
+  %   AM:  I agree that it's weird: we should not be able to weaken a linear context, if not
+  %   with unrestricted types --- but this is what we need for Sc-Res-Inact to be type preserving:
+  %   from $\Delta\vdash 0$ we need $\Delta\vdash (\nu x y)0$. BTW, this is also a problem in Vasco's system 
+  % \end{metanote}
+\end{lemma}
+
+\begin{lemma}[Strengthening]\mbox{}
+  \label{lemma:strenD}
+  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
+\end{lemma}
+
+\begin{lemma}[Substitution]\mbox{}
+  If $\types{(\Gamma, l);\Delta}{P}$ and ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
+  \( \types{\Gamma;\Delta}{\subst{P}{l}{a}} \).
+\end{lemma}
+
+\begin{lemma}[Preservation for $\equiv$]
+  If \( \types{\Gamma;\Delta}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
+\end{lemma}
+
+To formulate safety, we need to formally define what we mean by well-formed process.
+We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
+A process is then \emph{well formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
+
+Note that well-formed processes do not necessarily reduce. For example, the process
+\begin{equation*}
+  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
+\end{equation*}
+is well formed but also irreducible.
+
+\begin{theorem}[Type safety]
+  If \( \types{\Cempty}{P} \), then \( P \) is well formed.
+\end{theorem}
+
+\begin{corollary}
+  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well formed.
+\end{corollary}
+
+\subsection{Challenge: Name passing and scope extrusion}
+\label{sec:challenge:name-passing-scope-extrusion}
+
+This challenge formalises a proof that requires explicit scope extrusion.
+Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
+This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication
 The key issue of this challenge is reasoning about names that are ``in the process'' of being scope-extruded, which often presents difficulties for the mechanisation of binders.
 
-\subsection{Coinduction}
-Our \textbf{coinduction challenge} (Appendix~\ref{sec:challenge:coinduction}) is to prove a context lemma involving \emph{strong barbed congruence} for a limited calculus with (infinite) replication of processes.
+Reasoning about scope extrusion explicitly can sometimes be avoided by introducing a structural congruence rule into the semantics, but doing this means we lose information about the scope when reasoning about the semantics.
+Explicitly reasoning about scope extrusion is necessary to describe e.g.\ runtime monitors and compositions of systems.
+
+The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
+We will define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.
+The objective of this challenge is to prove that the two semantics are equivalent up to structural congruence.
+
+\subsubsection{Syntax}
+The syntax of processes is given by:
+\[
+\begin{array}{rcl}
+  P,Q & :=& \Pend \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad \Pout{x}{y}{P} \quad\mid\quad \Pin{x}{y}{P} \quad\mid\quad \Pres{x}{P}
+\end{array}
+\]
+
+%The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
+The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
+The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
+The input operator thus binds the name \( y \) in \( P \).
+% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+% The two components can proceed independently of each other, or they can interact via shared names.
+% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
+% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
+% The restriction operator thus binds the name \( x \) in \( P \).
+Note that the scope of a restriction may change when processes interact. Namely, a restricted name may be sent \emph{outside} of its scope.
+% The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
+Note that there is no recursion or replication in the syntax, and thus no infinite behaviors can be expressed.
+This simplifies the theory and is orthogonal to the concept of scope extrusion.
+
+% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
+% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
+% We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
+
+% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
+% As a convention, we will identify \( \alpha \)-convertible processes.
+
+% As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
+% This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
+
+\subsubsection{Reduction semantics}
+The first semantics is an operational reduction semantics, which avoids reasoning explicitly about scope extrusion by way of a structural congruence rule.
+\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
+\begin{mathpar}
+  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{P}{(\Ppar{Q}{R})}}{\Ppar{(\Ppar{P}{Q})}{R}}}
+  \and
+  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
+  \and
+  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
+  \and
+  \inferrule[Sc-Res-Par]{x \notin \freenames{Q}}{\scong{\Ppar{\Pres{x}{P}}{Q}}{\Pres{x}{(\Ppar{P}{Q})}}}
+  \and
+  \inferrule[Sc-Res-Inact]{ }{\scong{\Pres{x}{\Pend}}{\Pend}}
+  \and
+  \inferrule[Sc-Res]{ }{\scong{\Pres{x}{\Pres{y}{P}}}{\Pres{y}{\Pres{x}{P}}}}
+\end{mathpar}
+
+The operational semantics is then defined as the following binary relation on processes:
+\begin{mathpar}
+  \inferrule[R-Com]{ }{\reduces{\Ppar{\Pout{x}{y}{P}}{\Pin{x}{z}{Q}}}{\Ppar{P}{\subst{Q}{y}{z}}}}
+  \and
+  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Pres{x}{P}}{\Pres{x}{Q}}}
+  \and
+  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
+  \and
+  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
+\end{mathpar}
+Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{R-Com}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
+Finally, note that rule \TirName{R-Struct} allows for applying the structural congruence both before and after the reduction.
+This ensures that the reduction relation is closed under structural congruence.
+
+\subsubsection{Transition system semantics}
+The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
+The transitions are labelled by \emph{actions}, the syntax of which is as follows:
+\[
+\begin{array}{rcl}
+  \alpha & :=& \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
+\end{array}
+\]
+
+The \emph{free output action} \( \Aoutf{x}{y} \) is sending the name \( y \) via \( x \).
+The \emph{input action} \( \Ain{x}{y} \) is receiving the name \( y \) via \( x \).
+The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
+The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
+
+We will extend the notion of free and bound occurrences with 
+\( \freenames{\alpha} \) to denote the set of names that occur free in
+the action \( \alpha \) and  \( \boundnames{\alpha} \) to
+denote the set of names that occur bound in the action \( \alpha \).
+In the free output action \( \Aoutf{x}{y} \) and the input action
+\( \Ain{x}{y} \), both \( x \) and \( y \) are free names.  In the
+bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while
+\( y \) is a bound name.  We will further use the notation
+\( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \)
+and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur
+in the action \( \alpha \).
+
+The transition relation is then defined by the following rules:
+\begin{mathpar}
+  \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
+  \and
+  \inferrule[In]{ }{\transition{\Pin{x}{z}{P}}{\Ain{x}{y}}{\subst{P}{y}{z}}}
+  \and
+  \inferrule[Par-L]{\transition{P}{\alpha}{P'} \\ \boundnames{\alpha} \cap \freenames{Q} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
+  \and
+  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'} \\ \boundnames{\alpha} \cap \freenames{P} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
+  \and
+  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{y}}{P'} \\ \transition{Q}{\Ain{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Comm-R]{\transition{P}{\Ain{x}{y}}{P'} \\ \transition{Q}{\Aoutf{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Close-L]{\transition{P}{\Aoutb{x}{z}}{P'} \\ \transition{Q}{\Ain{x}{z}}{Q'} \\ z \notin \freenames{Q}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
+  \and
+  \inferrule[Open]{\transition{P}{\Aoutf{x}{z}}{P'} \\ z \neq x}{\transition{\Pres{z}{P}}{\Aoutb{x}{z}}{P'}}
+  \and
+  \inferrule[Close-R]{\transition{P}{\Ain{x}{z}}{P'} \\ \transition{Q}{\Aoutb{x}{z}}{Q'} \\ z \notin \freenames{P}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
+  \and
+  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
+\end{mathpar}
+Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{In}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
+
+We keep the convention that bound names of any processes or actions
+are chosen to be different from the names that occur free in any other
+entities under consideration, such as processes, actions,
+substitutions, and sets of names.  The convention has one exception,
+namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \),
+the name \( z \) (which occurs bound in \( P \) and the action
+\( \Aoutb{x}{z} \)) may occur free in \( Q \).  Without this exception
+it would be impossible to express scope extrusion.
+
+\subsubsection{Challenge}
+\begin{lemma}\label{se-lemma-harmony-fact}
+  If \( \scong{P}{Q} \) and \( \transition{P}{\alpha}{P'} \), then for some \( Q' \) we have \( \transition{Q}{\alpha}{Q'} \) and \( \scong{P'}{Q'} \).
+\end{lemma}
+
+For the proof of the challenge theorem, we will introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
+The first rule applied is \TirName{R-Com}. The derivation continues with an application of \TirName{R-Par}, followed by zero or more applications of \TirName{R-Res}.
+The last rule is an application of \TirName{R-Struct}.
+\begin{lemma}\label{se-lemma-normalized-derivation}
+  Every reduction has a normalized derivation.
+\end{lemma}
+
+\begin{lemma}\label{se-lemma-normalized-reduction}
+  If \( \reduces{P}{Q} \) then there are \( x, y, z, z_1, \dots, z_n, R_1, R_2, \) and \( S\) such that
+  \begin{align*}
+    &\scong{P}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{\Pout{x}{y}{R_1}}{\Pin{x}{z}{R_2}})}{S})}}} \\
+    &\scong{Q}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{R_1}{\subst{R_2}{y}{z}})}{S})}}}
+  \end{align*}
+\end{lemma}
+
+The objective of this challenge is to prove the following theorems, which together show the equivalence between the reduction semantics and the transition system semantics up to structural congruence.
+The first of the theorems involves reasoning about scope extrusion more directly than the other, and if time does not permit proving both of the theorems, \cref{thm:se-trans-implies-red} should be proven first.
+\begin{theorem}\label{thm:se-trans-implies-red}
+  \( \transition{P}{\Atau}{Q} \) implies \( \reduces{P}{Q} \).
+\end{theorem}
+
+\begin{theorem}
+  \( \reduces{P}{Q} \) implies the existence of a \( Q' \) such that \( \transition{P}{\Atau}{Q'} \) and \( \scong{Q}{Q'} \).
+\end{theorem}
+
+\subsection{Challenge: Coinduction and infinite }
+\label{sec:challenge:coinduction}
+This challenge is a about the mechanisation of proofs
+% formalises a proof
+that require coinduction.
+% coinductive techniques.
+%
+% Coinduction is a proof technique for infinite structures,
+This is usually connected to coinductive definitions where an infinite
+structure is defined as greatest fixed points of a recursive
+definition.
+%
+Coinduction is then a technique for defining and proving properties of
+such infinite structures,
+%
+which, in the context of the concurrent benchmark, arise
+% in this context
+due to systems with behaviors that continue indefinitely.
+%
+Coinduction is the dual of induction: whereas induction is useful for
+proving properties of least fixed points, coinduction is useful for
+proving properties of greatest fixed points.
+%
+% \marco{Should we be more specific here and give a deeper explanation
+%   of what coinduction is? If not, then why do we even say that it's
+%   the dual of induction?}
+%
+
+For this challenge, we adopt a fragment of the untyped $\pi$-calculus
+that includes process replication.
+% The setting for this challenge is an untyped calculus of
+% communicating systems with replication of processes, the syntax and
+% semantics of which are presented below.
+The objective of this challenge is to % prove % a context theorem for
+draw a formal connection between strong barbed congruence and strong
+barbed bisimilarity.  % I.e., The context lemma
+The result establishes that two processes are strong barbed congruent
+if and only if the processes obtained by applying a substitution to
+them and composing them in parallel with an arbitrary process are
+strongly barbed bisimilar.
 The key issue of this challenge is the coinductive reasoning about the infinite behaviours of the replication operator.
+
+\paragraph{Syntax.}
+% We assume the existence of some type of \emph{base values}, values
+% of which we will denote by \( a, b, \dots \), the existence of some
+% type of \emph{variable names}, values of which we will denote by
+% \( l, m, \dots \), and the existence of some type of \emph{names},
+% values of which we will denote by \( x, y, \dots \).
+The syntax of the calculus is identical to that of
+Section~\ref{sec:challenge:linearity-beh-types} (challenge on
+linearity), except from restriction, now unary, and the addition of
+replication:
+% for including replication and having unary restriction:
+\begin{displaymath}
+  \begin{array}{rcl}
+    v,w & ::= & a \quad\mid\quad l \\
+    P,Q & ::=& \Pend
+               \quad \mid\quad \Pout{x}{v}{P}
+               \quad \mid\quad \Pin{x}{l}{P}
+               \quad \mid\quad \Ppar{P}{Q}
+               \quad \mid\quad \Pres{x}{}{P}
+               \quad \mid\quad !P
+  \end{array}
+\end{displaymath}
+Note that, unlike the process calculus from
+Section~\ref{sec:challenge:linearity-beh-types}, communications occur
+over the same channel name. Moreover, since replication $!P$ allows
+for infinite copies of the process $P$, processes can dynamically
+create an infinite number of names during execution when combined with
+restriction $\Pres{x}{}{P}$.
+%
+% \begin{align*}
+%   v,w :=&&& a \\
+%   |&&& l \\
+%   P,Q :=&&& \Pend \\
+%   |&&& \Pout{x}{v}{P} \\
+%   |&&& \Pin{x}{l}{P} \\
+%   |&&& \Ppar{P}{Q} \\
+%   |&&& \Pres{x}{P} \\
+%   |&&& \Preplicate{P}
+% \end{align*}
+% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a
+% variable name \( l \).  The process \( \Pend \) is \emph{inaction}:
+% a process which can do nothing.  The process \( \Pout{x}{a}{P} \) is
+% an \emph{output}, which can send the value \( v \) via \( x \), then
+% continue as \( P \).  The intention is that the value \( v \) must
+% be a base value when it is actually sent, and this will be enforced
+% in the semantics later on.  The process \( \Pin{x}{l}{P} \) is an
+% \emph{input}, which can receive a base value via \( x \), then
+% continue as \( P \) with the received value substituted for the
+% variable name \( l \).  The input operator thus binds the variable
+% name \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
+% \emph{composition} of process \( P \) and process \( Q \).  The two
+% components can proceed independently of each other, or they can
+% interact via shared names.  The process \( \Pres{x}{P} \) is the
+% \emph{restriction} of the name \( x \) to \( P \).  Components in
+% \( P \) can use the name \( x \) to interact with each other, but
+% not with processes outside of the restriction.  The restriction
+% operator thus binds the name \( x \) in \( P \).  Note that the
+% scope of a restriction may not change when processes interact, since
+% it is only possible to send and receive values, and not names.  The
+% process \( \Preplicate{P} \) is the \emph{replication} of the
+% process \( P \).  It can be thought of as the infinite composition
+% \( \Ppar{P}{\Ppar{P}{\cdots}} \).  Replication makes it possible to
+% express infinite behaviours.
+
+% We will use the notation \( \subst{P}{v}{l} \) to denote the process
+% \( P \) with value \( v \) substituted for variable name \( l \).
+
+% % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
+% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
+% written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from
+% \( P \) by a finite number of substitutions of bound variable names.
+% As a convention, we will identify \( \alpha \)-convertible
+% processes.
+
+% Also as a convention, we assume that the bound names and bound
+% variable names occurring in any collection of processes are chosen
+% to be different from the free names and free variable names
+% occurring in those processes and from the names and variable names
+% occurring in any substitutions applied to the processes.  This is
+% justified because any overlapping names and variable names may be
+% \( \alpha \)-converted such that the assumption is satisfied.
+
+\paragraph{Semantics.} We choose to give an LTS semantics for this
+challenge. As an alternative, it is also possible to adopt a reduction
+semantics which features fewer rules but needs a structural
+congruence. The semantics of the system describes the actions that the
+system can perform by defining a labelled transition relation on
+processes.  The transitions are labelled by \emph{actions}, the syntax
+of which are as follows:
+% \begin{align*}
+%   \alpha := &&& \Aoutf{x}{a} \\
+%   |&&& \Ain{x}{a} \\
+%   |&&& \Atau
+% \end{align*}
+\begin{align*}
+  \alpha\ :=\  \Aoutf{x}{a} \ \mid\ \Ain{x}{a} \ \mid\ \Atau
+\end{align*}
+The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value
+\( a \) via \( x \).  The \emph{input action} \( \Ain{x}{y} \) is
+receiving the base value \( y \) via \( x \).  The \emph{internal
+  action} \( \Atau \) is performing some unobservable action, e.g.\
+internal communication.
+%
+We will use the notation \( \names{\alpha} \) to denote the set of
+names that occur in the action \( \alpha \).
+
+The transition relation is then defined by the following rules:
+\begin{mathpar}
+  \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
+  \and
+  \inferrule[In]{ }{\transition{\Pin{x}{l}{P}}{\Ain{x}{a}}{\subst{P}{a}{l}}}
+  \and
+  \inferrule[Par-L]{\transition{P}{\alpha}{P'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
+  \and
+  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
+  \and
+  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{a}}{P'} \\ \transition{Q}{\Ain{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Comm-R]{\transition{P}{\Ain{x}{a}}{P'} \\ \transition{Q}{\Aoutf{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ x \notin \names{\alpha}}{\transition{\Pres{x}{P}}{\alpha}{\Pres{x}{P'}}}
+  \and
+  \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
+\end{mathpar}
+Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
+On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable name.
+
+% As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
+
+\subsubsection{Bisimilarity}
+Bisimilarity is a notion of equivalence for processes and
+% .  Our notion of process equivalence relations
+builds on a notion of \emph{observables}, i.e., what we can externally
+observe from the semantics of a process. If we allow ourselves only to observe
+internal transitions, i.e., observe that a process is internally
+performing a step of computation, we will relate either too few
+processes (in the strong case where we relate only processes with
+exactly the same number of internal transitions) or every process (in
+the weak case where we relate processes with any amount of internal
+transitions).  We must therefore allow ourselves to observe more than
+just internal transitions, and here choose to define the observables
+of a process as the names it can use for sending and receiving.
+
+To this end, we define the \emph{observability predicate}
+\( \observable{P}{\mu} \) as follows:
+\begin{align}
+  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
+  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
+\end{align}
+
+\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is then the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
+\begin{gather}
+  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu}\label{eq:bisim1} \\
+  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}\label{eq:bisim2}
+\end{gather}
+We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
+Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
+
+\begin{theorem}
+  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
+\end{theorem}
+
+Unfortunately, strong barbed bisimilarity is not a good process
+equivalence since it is not a congruence. This is because we wish to
+be able to substitute a process with an equivalent one in any context
+it may appear.
+% equivalent processes to be since it does not consider the
+% environment of processes.
+For instance, the processes $\Pout{x}{a}{\Pout{y}{b}{\Pend}}$ and
+$\Pout{x}{a}{\Pend}$ are strong barbed bisimilar, i.e.,
+$\sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}}$. This
+is because \( \obsout{x} \) is the only observable in both processes
+and they cannot perform a \( \Atau \)-action. However, if composed in
+parallel with the process $\Pin{x}{l}{\Pend}$, the relation will no
+longer hold. In fact,
+$\nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}\Ppar{}{\Pin{x}{l}{\Pend}}}$
+%
+because the left process can perform a \( \Atau \)-action such that
+\( \obsout{y} \) becomes observable, whereas the right process cannot.
+
+% For instance,
+% \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \)
+% since \( \obsout{x} \) is the only observable in both processes and
+% they cannot perform a \( \Atau \)-action, but
+% \(
+% \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}}
+% \) since the left process can perform a \( \Atau \)-action such that
+% \( \obsout{y} \) becomes observable, whereas the right process cannot.
+
+% \subsubsection{Contexts and congruences}
+% Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
+
+% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+
+% We can think of contexts as functions between processes.
+% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+
+% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+
+\subsubsection{Strong barbed congruence}
+In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,
+i.e., % we need to make strong barbed bisimilarity consider
+we have to consider the environment in which processes may be placed.
+% (i.e.\ the context) of processes.
+
+We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
+
+\begin{lemma}
+  \( \sbcong{}{} \) is the largest congruence included in
+  \( \sbbisim{}{} \).
+\end{lemma}
+
+\subsubsection{Challenge}
+The objective of this challenge is to prove a theorem that shows that
+making strong barbed bisimilarity sensitive to substitution and
+parallel composition is equivalent to strong barbed
+congruence. Formally,
+\begin{theorem}
+  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and
+  substitution \( \sigma \),
+  \(
+  \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}}
+  \).
+\end{theorem}
 
@@
 
-\begin{table}
+\begin{table}[btp]
   \caption{Possible extensions of the challenges. Each suitable extension is marked ``+''. \label{tab:extensions}}
@@
 
-\appendix
-\section{Challenges}\label{app:challenges}
-\input{../challenges/challenges.tex}
-
 \bibliographystyle{splncs04}

### 2023-10-13 paper/paper.tex
Mercilessly cut to fit 15 pages.

@@
 in the action \( \alpha \).
-
 The transition relation is then defined by the following rules:
@@
 names that occur in the action \( \alpha \).
-
 The transition relation is then defined by the following rules:
@@
 of a process as the names it can use for sending and receiving.
-
 To this end, we define the \emph{observability predicate}
@@
 describe the individual design considerations for each challenge problem.
-
 % We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
@@
 and results that involve interesting proof techniques that may be reused in further work.
-
 To achieve our design goal \ref{item:goal-comperison-accessibility},
@@
 constructs, but we leave this as a possibility for a future benchmark challenge.)
-
 The minimality and uniformity of the calculi also allows us to target our design
@@
 
-\subsection{Linearity}
+\paragraph{Linearity}
 For our challenge on linear reasoning we have chosen a safety theorem
@@
 transition system semantics.
-
 Linearity is of course interesting in many other subfields of programming languages, but we have chosen to base our challenge on a session type system primarily due to the common use of session types in concurrency theory.
-
 Inspired by Vasconcelos~\cite{Vasconcelos2012}, we use a syntax where
@@
 of the two names obvious in the type system.
-
 We have chosen a simple notion of well-formedness for our challenge,
@@
 the challenge more difficult and less focused on the linear
-aspects of the system.
-
-\subsection{Scope extrusion}
+aspects.
+
+\paragraph{Scope extrusion}
 This is the challenge most closely related to the original POPLMark
@@
 
-\subsection{Coinduction}
+\paragraph{Coinduction}
 %\fxerror{
@@
 Though weak barbed congruence is a more common behavioural equivalence for \(\pi\)-terms, we avoid having to abstract over the number of internal transitions when considering the strong notions of equivalence, thus simplifying the theory.
-
 We have chosen not to include delegation in the coinduction challenge since it is orthogonal to our primary aim of exploring coinductive proof techniques.
@@
 
-As with the POPLMark and POPLMark Reloaded challegnes, this challenge is
-intended to be useful, but it does not aim at covering every possible
+As with the POPLMark and POPLMark Reloaded challegnes, this challenge does not aim to cover every possible
 problem in the area. Our benchmark can be extended in two
 dimensions: further expanding and combining the current challenges,
-and addressing calculi and features not covered by the current challenges.
-We discuss these dimensions in the rest if this section.
+and addressing features not covered by the current challenges.
 
@@
 entail combining the features of the challenges.
-
 In \cref{tab:extensions} we consider the following three extensions: first, the
@@
   \caption{Possible extensions of the challenges. Each suitable extension is marked ``+''. \label{tab:extensions}}
-  \begin{center}
+  \begin{center}\small
     \begin{tabular}{|r|c|c|c|}
@@
 message passing calculi, or to cover different proof mechanisation features.
-
 Some interesting aspects of message passing calculi to explore in
@@
 and trace equivalence.
-
 On the side of proof assistants, the current challenge proposes
@@
 field --- and combining automated proofs with proof assistants
-offers the potential to ease the path for attempting larger proofs
-with less effort.
-
+offers the potential to ease the path for attempting larger proofs.
 Ultimately, the current challenges can be extended in several

### 2023-10-13 paper/paper.tex
Fix title of coinduction challenge.
updated weakening
fixed some text in second challenge

@@
     \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
-      \item If \( \types{\Gamma; \Delta}{P} \) and \( \tend{\Delta} \) then
+      \item If \( \types{\Gamma; \Delta}{P} \) % and \( \tend{\Delta} \) 
+        then
     \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
@@
 
-\subsubsection{Challenge}
+Before getting to our challenge, we prove some intermediate results. 
 \begin{lemma}\label{se-lemma-harmony-fact}
@@
 
+\subsubsection{Challenge}
 The objective of this challenge is to prove the following theorems, which together show the equivalence between the reduction semantics and the transition system semantics up to structural congruence.
@@
 
-\subsection{Challenge: Coinduction and infinite }
+\subsection{Challenge: Coinduction and infinite processes}
 \label{sec:challenge:coinduction}

### 2023-10-13 challenges/linearity.tex
Fix formatting of rule names in linearity challenge.
draft of well-formedness proof

@@
 \end{mathpar}
-Note that we do not need a judgment for typing channels, since it is already folded into the T-IN and T-OUT rules.
+Note that we do not need a judgment for typing channels, since it is already folded into the \TirName{T-In} and \TirName{T-Out} rules.
 
@@
 
-To formulate safety, we need to formally define what we mean by well-formed process.
-We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
-A process is then \emph{well formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
+To formulate safety, we need to formally define what we mean by
+well-formed process.  We say that a process \( P \) is \emph{prefixed
+  at variable \( x \)} if \( P\equiv\Pout{x}{v}{P'} \) or
+\( P\equiv\Pin{x}{l}{P'} \) for some $P'$.
+%
+A process $P$ is then \emph{well formed} if, for every $P_1$, $P_2$,
+and $R$ such that
+\( P\equiv \Presd{x_1}{y_1}{\dots
+  \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}} \), with
+\( n \geq 0 \), it holds that, if \( P_1 \) is prefixed at \( x_1 \)
+and \( P_2 \) is prefixed at \( y_1 \) (or viceversa), then
+\( \Ppar{P_1}{P_2}
+\equiv\Ppar{\Pout{x_1}{a}{P_1'}}{\Pin{y_1}{l}{P_2'}} \), for some
+$P_1'$ and $P_2'$.
 
@@
 \end{theorem}
+\begin{proof}
+  We proceed by induction on the definition of $\types{}{}$. The key
+  cases are that for parallel composition and that for name
+  restriction. In the case of parallel composition, linearity (and
+  context splitting) ensure that every process in parallel will have
+  different channel names. Then, the restriction rule enforces
+  duality, therefore an output will be matched by an input.
+\end{proof}
 

### 2023-10-13 challenges/linearity.tex
add proofs

@@
   \and
-  \inferrule[T-Res]{\types{\Gamma; (\Cadd{\Cadd{\Delta}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P})}{\types{\Gamma}{\Presd{x}{y}{P}}}
+  \inferrule[T-Res]{\types{\Gamma; (\Cadd{\Cadd{\Delta}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P})}{\types{\Gamma;\Delta}{\Presd{x}{y}{P}}}
   \and
@@
 \begin{proof}
-  By induction on the given derivation.
+  By induction on $P$, for each case inverting $\types{\Gamma; (\Delta, x : T)}{P}$ to the uniquely applicable rule.
+  \begin{description}
+    \item[$P = \Pend$] By inversion on the typing we have $\tend{\Delta}$. Apply T-INACT.
+    \item[$P = \Ppar{P_0}{P_1}$] By inversion on the typing we have  $ \Delta, x : T = \Delta_0 , \Delta_1$.
+      By cases on which environment $x$ is in, apply IH on that environment. 
+     \item[$P = \Presd{y}{z}{P}$] We may assume $x \notin \{y,z\}$. By inversion, $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From the assumptions present we can apply IH.\\
+
+       The remaining case are analogous
+
+  \end{description}
 \end{proof}
@@
   If $\types{(\Gamma, l);\Delta}{P}$ and ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
-  \( \types{\Gamma;\Delta}{\subst{P}{l}{a}} \).
+  \( \types{\Gamma;\Delta}{\subst{P}{a}{l}} \).
 \end{lemma}
 \begin{proof}
-  By induction on the first derivation.
+  By induction on $P$ inverting  $\types{(\Gamma, l);\Delta}{P}$ in each case.
+  \begin{description}
+    \item[$P = \Pend$] Immediate after inverting typing.
+    \item[$P = \Ppar{P_0}{P_1}$] By inversion on the typing we get derivations for $P_0$ and $P_1$. Apply IH on both.
+     \item[$P = \Presd{x}{y}{P}$] Immediate after inverting typing.
+     \item[$P = \Pout{x}{v}{P}$] Invert typing. we have  $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
+       We must show $\typev{\Gamma}{\hastype{v}{\Tbase}}$ to build the conclusion with IH and T-OUT. Go by cases on the struture of $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
+     \item[$P = \Pin{x}{v}{P}$] Immediate after inverting typing.
+
+  \end{description}
 \end{proof}

### 2023-10-13 challenges/linearity.tex
adding detailed proof of SR

--- 2023-10-13 challenges/linearity.tex
@@
 \subsubsection{Challenge}
-The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with:
-\begin{theorem}[Subject reduction]
-  If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
-\end{theorem}
-\begin{proof}
-  By induction on the derivation of $\reduces{P}{Q}$. TBD
-\end{proof}
+The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with some lemmas:
 
-The following lemmas are needed:
 \begin{lemma}[Weakening]\mbox{}
@@
   \end{enumerate}
-  % \begin{metanote}
-  %   Note from Dawit: Does \( \Delta \) really have to be only ends to be weakened?
-I had expected that if \( P \) is typed by \( \Delta \) and \( \Delta' \) is only ends, then \( P \) is typed by \( \Cadd{\Delta}{\Delta'} \).
-
-  %   AM:  I agree that it's weird: we should not be able to weaken a linear context, if not
-  %   with unrestricted types --- but this is what we need for Sc-Res-Inact to be type preserving:
-  %   from $\Delta\vdash 0$ we need $\Delta\vdash (\nu x y)0$. BTW, this is also a problem in Vasco's system 
-  % \end{metanote}
 \end{lemma}
@@
 \begin{lemma}[Substitution]\mbox{}
+  \label{le:subst}
   If $\types{(\Gamma, l);\Delta}{P}$ and ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
@@
 \begin{lemma}[Preservation for $\equiv$]
+  \label{le:presequiv}
   If \( \types{\Gamma;\Delta}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
@@
   For the step case, TBD
+\end{proof}
+
+\begin{theorem}[Subject reduction]
+  If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
+\end{theorem}
+\begin{proof}
+  By induction on the derivation of $\reduces{P}{Q}$. The cases R-PAR
+  and R-RES follows immediately by IH. Case R-STRUCT appeals twice to
+  preservation of $\equiv$ (Lemma~\ref{le:presequiv}) and to the
+  IH. For R-COM, suppose that T-RES introduces in $\Delta$ the
+  assumptions $x : !.U, y : ?\dual U$.  Building the only tree for the
+  hypothesis, we know that $\Delta = \Delta_1, \Delta_2, \Delta_3$
+  where $\Gamma;\Delta_3 \vdash R$. We also have
+  $ \Gamma;\Delta_1 , x : U \vdash P$, ${\cal D}_{2} $ a proof of
+  $\Gamma,l;\Delta_2, y : \dual U \vdash Q$ and $\cal V $ is a proof of
+  $\Gamma \vdash a : \Tbase$.  From ${\cal D}_{2}$ and $\cal V$ we use the
+  substitution Lemma~\ref{le:subst} to obtain $\Gamma;\Delta_2 , y : \dual U \vdash
+  Q[a/l]$. We then conclude the proof with rules T-PAR, T-PAR and
+  T-RES.
 \end{proof}

### 2023-10-13 paper/paper.tex
Readd the appendix to make it easier to work on proofs.

@@
 
-Note that well-formed processes do not necessarily reduce. For example, the process
+Note that well-formed processes do not necessarily reduce. For example, the process:
 \begin{equation*}
@@
 \paragraph{Scope extrusion}
-This is the challenge most closely related to the original POPLMark
+This is the challenge most closely related to the POPLMark
 challenge since it concerns the properties of binders in restrictions.
-It is not clear that the techniques used to solve the POPLMark challenges about binders can be easily extended to reasoning about scope extrusion.
+It is not clear that the techniques used to solve the original challenges about binders can be easily extended to reasoning about scope extrusion.
 Notably, techniques like Higher Order Abstract Syntax, or well-scoped
@@
 
+\appendix
+\section{Challenges}\label{app:challenges}
+\input{../challenges/challenges.tex}
+
 \bibliographystyle{splncs04}

### 2023-10-14 challenges/linearity.tex
updated proof of type safety

@@
 \begin{theorem}[Type safety]
-  If \( \types{\Cempty}{P} \), then \( P \) is well formed.
+  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
 \end{theorem}
 \begin{proof}
-  We proceed by induction on the definition of $\types{}{}$. The key
-  cases are that for parallel composition and that for name
-  restriction. In the case of parallel composition, linearity (and
-  context splitting) ensure that every process in parallel will have
-  different channel names. Then, the restriction rule enforces
-  duality, therefore an output will be matched by an input.
+  In order to prove that $P$ is well-formed, let us consider any
+  process of the form
+  $\Presd{x_1}{y_1}{\dots
+    \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}}$ that is
+  structurally congruent to $P$. Moreover, assume that $P_1$ is
+  prefixed at some $x_1$ and $P_2$ is prefixed at some $x_2$ such that
+  $P_1\equiv\Pout{x}{v}{P_1'}$ (the opposite case proceeds
+  similarly). We need to show that $P_2\equiv\Pin{x}{v}{P_2'}$. This
+  can be easily done by contradiction. In fact, if
+  $P_2\equiv\Pout{x}{v}{P_2'}$ then the typing rule for restriction
+  would be violated since the type of $x_1$ and $y_1$ cannot be dual.
+  % 
+  % We proceed by induction on the definition of $\types{}{}$. The key
+  % cases are that for parallel composition and that for name
+  % restriction. In the case of parallel composition, linearity (and
+  % context splitting) ensure that every process in parallel will have
+  % different channel names. Then, the restriction rule enforces
+  % duality, therefore an output will be matched by an input.
 \end{proof}

### 2023-10-14 paper/paper.tex
Many small adjustments, reformatting, typos

@@
 \usepackage{amssymb}
+\usepackage{mathtools}
 \usepackage{graphicx}
@@
 \label{sec:prelim}
-First, we list some common notions and conventions that we use in the challenges. Since the calculi under study are somewhat different, each section will list the changes that apply.
+First, we list some common notions and conventions that we use in the
+challenges. Since the calculi under study are somewhat different, each section
+lists the changes that apply.
 
@@
 \( l, m, \dots \), and the existence of some set of \emph{names},
-represented by the symbols \( x, y, \dots \).
-We assume that all of these sets are infinite and that their elements can be compared for equality.
+represented by the symbols \( x, y, \dots \).%
+\footnote{%
+  Unlike the standard $\pi$-calculus, we distinguish variables from names to
+  better control the expressiveness of the calculi under study, and the scope of
+  the corresponding challenges: the key distinction is that names can be sent
+  and received (thus realising channel delegation), whereas variables can only
+  be substituted with base values.%
+} %
+\fxerror{Double-check the footnote!}%
+We assume that all of these sets are infinite and that their elements can be
+compared for equality.
 
@@
 For communication, processes include \emph{input} and \emph{output}, whose
-signature will depend on the calculus being value or name-passing.
-We use here the metavariables $c,k$ to abstract over this choice.
+signature depends on the calculus being value-passing or name-passing.
+We use here the metavariables $c,k$ to abstract over this choice
+--- i.e.~$c$ may be either a value or a variable or a name, whereas $k$ may
+be a variable or a name.
 The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
@@
 It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-Replication makes it possible to express infinite behaviors.
-
-We will use the notation \( \freenames{P} \) to denote the set of
+Replication makes it possible to express infinite behaviours.
+
+We use the notation \( \freenames{P} \) to denote the set of
 names that occur free,
@@
 % (i.e.\ not bound by an input)
-in \( P \).  We will use the notation \( \boundvars{P} \)
+in \( P \).  We use the notation \( \boundvars{P} \)
 for the set of variable names that occur bound
 % (by an input)
-in \( P \).  We will use the notation \( \subst{P}{a}{l} \) to denote
+in \( P \).  We use the notation \( \subst{P}{a}{l} \) to denote
 the process \( P \) with base value \( a \) substituted for variable name
@@
 \( P \) with name \( x \) substituted for name \( y \).
-We will use the notation \( \applysubst{\sigma}{P} \) to denote the process \( P \) with a finite number of arbitrary substitutions applied to it.
+We use the notation \( \applysubst{\sigma}{P} \) to denote the process \( P \) with a finite number of arbitrary substitutions applied to it.
 
@@
 by a finite number of substitutions of bound variable names.  As a
-convention, we will identify \( \alpha \)-convertible processes and we
+convention, we identify \( \alpha \)-convertible processes and we
 assume that bound names and bound variable names of any processes are
@@
 As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-We will think of contexts as functions between processes.
-A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+%
+A context acts as a function between processes:
+a context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+The replacement is literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
 
@@
 \[
-\begin{array}{rcl}
-  v,w & ::= & a \quad\mid\quad l \\
-   P,Q & ::=& \Pend \quad\mid\quad \Pout{x}{v}{P} \quad\mid\quad \Pin{x}{l}{P} \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad  \Presd{x}{y}{P}
+\begin{array}{r@{\qquad}c@{\qquad}l}
+  v,w & \Coloneqq & a \quad\mid\quad l \\
+   P,Q & \Coloneqq & \Pend \quad\mid\quad \Pout{x}{v}{P} \quad\mid\quad \Pin{x}{l}{P} \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad  \Presd{x}{y}{P}
 \end{array}
@@
 The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \) and then continues as \( P \).
-The intention is that the value \( v \) must be a base value when it is actually sent, and this will be enforced in the semantics later on.
+The intention is that the value \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
 The input process \( \Pin{x}{l}{P} \) waits for a base value from \( x \) and then continues as \( P \) with the received value substituted for the variable name \( l \).
@@
 
-The operational semantics is then defined as the following binary relation on processes:
+The operational semantics is defined as the following relation on processes:
 \begin{mathpar}
@@
 
-\paragraph{Syntax}
+\subsubsection{Syntax.}
 Our type system does not type processes directly, but instead focuses on the channels used in the process.
@@
 \[
-  \begin{array}{rcl}
-  S,T & ::= & \Tend \quad\mid\quad \Tbase \quad\mid\quad \Tin{S} \quad\mid\quad \Tout{S} \\
-    \Gamma &::= & \Cempty \quad\mid\quad \Gamma, l \\
+  \begin{array}{r@{\qquad}c@{\qquad}l}
+  S,T & \Coloneqq & \Tend \quad\mid\quad \Tbase \quad\mid\quad \Tin{S} \quad\mid\quad \Tout{S} \\
+    \Gamma & \Coloneqq & \Cempty \quad\mid\quad \Gamma, l \\
                     \Delta &::= & \Cempty \quad\mid\quad \Cadd{\Delta}{\hastype{x}{S}}
@@
 
-\subsubsection{Typing rules}
+\subsubsection{Typing rules.}
 Our type system is aimed at maintaining two invariants:
@@
 
-\subsubsection{Challenge}
+\subsubsection{Challenge.}
 The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with:
@@
 The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
-We will define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.
+We define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.
 The objective of this challenge is to prove that the two semantics are equivalent up to structural congruence.
 
-\subsubsection{Syntax}
+\subsubsection{Syntax.}
 The syntax of processes is given by:
 \[
-\begin{array}{rcl}
-  P,Q & :=& \Pend \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad \Pout{x}{y}{P} \quad\mid\quad \Pin{x}{y}{P} \quad\mid\quad \Pres{x}{P}
+\begin{array}{r@{\qquad}c@{\qquad}l}
+  P,Q & \Coloneqq & \Pend \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad \Pout{x}{y}{P} \quad\mid\quad \Pin{x}{y}{P} \quad\mid\quad \Pres{x}{P}
 \end{array}
@@
 
-\subsubsection{Reduction semantics}
+\subsubsection{Reduction semantics.}
 The first semantics is an operational reduction semantics, which avoids reasoning explicitly about scope extrusion by way of a structural congruence rule.
 \emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
+
 \begin{mathpar}
@@
 Note also that due to rule \TirName{R-Com}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
-Finally, note that rule \TirName{R-Struct} allows for applying the structural congruence both before and after the reduction.
-This ensures that the reduction relation is closed under structural congruence.
-
-\subsubsection{Transition system semantics}
+Finally, note that rule \TirName{R-Struct} allows for applying the structural congruence both before and after the reduction:
+this makes the reduction relation closed under structural congruence.
+
+\subsubsection{Transition system semantics.}
 The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which is as follows:
+The transitions are labelled by \emph{actions}, the syntax of which is:
 \[
-\begin{array}{rcl}
-  \alpha & :=& \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
+\begin{array}{r@{\qquad}c@{\qquad}l}
+  \alpha & \Coloneqq & \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
 \end{array}
@@
 
-We will extend the notion of free and bound occurrences with
+We extend the notion of free and bound occurrences with
 \( \freenames{\alpha} \) to denote the set of names that occur free in
@@
 bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while
-\( y \) is a bound name.  We will further use the notation
+\( y \) is a bound name.  We also use the notation
 \( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \)
@@
 
-For the proof of the challenge theorem, we will introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
+For the proof of the challenge theorem, we introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
 The first rule applied is \TirName{R-Com}. The derivation continues with an application of \TirName{R-Par}, followed by zero or more applications of \TirName{R-Res}.
@@
 
-\subsubsection{Challenge}
+\subsubsection{Challenge.}
 The objective of this challenge is to prove the following theorems, which together show the equivalence between the reduction semantics and the transition system semantics up to structural congruence.
@@
 
-\paragraph{Syntax.}
+\subsubsection{Syntax.}
 % We assume the existence of some type of \emph{base values}, values
@@
 \begin{displaymath}
-  \begin{array}{rcl}
-    v,w & ::= & a \quad\mid\quad l \\
-    P,Q & ::=& \Pend
+  \begin{array}{r@{\qquad}c@{\qquad}l}
+    v,w & \Coloneqq & a \quad\mid\quad l \\
+    P,Q & \Coloneqq & \Pend
                \quad \mid\quad \Pout{x}{v}{P}
@@
 
-\paragraph{Semantics.} We choose to give an LTS semantics for this
+\subsubsection{Semantics.} We choose to give an LTS semantics for this
 challenge. As an alternative, it is also possible to adopt a reduction
@@
 processes.  The transitions are labelled by \emph{actions}, the syntax
-of which are as follows:
+of which is as follows:
 % \begin{align*}
@@
 \begin{align*}
-  \alpha\ :=\  \Aoutf{x}{a} \ \mid\ \Ain{x}{a} \ \mid\ \Atau
+  \alpha\ \ \Coloneqq\ \ \Aoutf{x}{a} \ \mid\ \Ain{x}{a} \ \mid\ \Atau
 \end{align*}
@@
 %
-We will use the notation \( \names{\alpha} \) to denote the set of
+We use the notation \( \names{\alpha} \) to denote the set of
 names that occur in the action \( \alpha \).
@@
 
-\subsubsection{Bisimilarity}
+\subsubsection{Bisimilarity.}
 Bisimilarity is a notion of equivalence for processes and
@@
 builds on a notion of \emph{observables}, i.e., what we can externally
-observe from the semantics of a process. If we allow ourselves only to observe
-internal transitions, i.e., observe that a process is internally
-performing a step of computation, we will relate either too few
+observe from the semantics of a process. If we allowed ourselves only to observe
+internal transitions (i.e., observe that a process is internally
+performing a step of computation) we would relate either too few
 processes (in the strong case where we relate only processes with
@@
 transitions).  We must therefore allow ourselves to observe more than
-just internal transitions, and here choose to define the observables
+just internal transitions, and here we choose to define the observables
 of a process as the names it can use for sending and receiving.
@@
 \( \observable{P}{\mu} \) as follows:
-\begin{align}
+\begin{align*}
   \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
   \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
+\end{align*}
 
 \emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is then the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu}\label{eq:bisim1} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}\label{eq:bisim2}
-\end{gather}
+\begin{gather*}
+  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\%\label{eq:bisim1} \\
+  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'} %\label{eq:bisim2}
+\end{gather*}
 We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
@@
 Unfortunately, strong barbed bisimilarity is not a good process
-equivalence since it is not a congruence. This is because we wish to
-be able to substitute a process with an equivalent one in any context
-it may appear.
+equivalence since it is not a congruence, hence it does not allow for
+substituting a process with an equivalent one in any context.
 % equivalent processes to be since it does not consider the
@@
 is because \( \obsout{x} \) is the only observable in both processes
-and they cannot perform a \( \Atau \)-action. However, if composed in
-parallel with the process $\Pin{x}{l}{\Pend}$, the relation will no
-longer hold. In fact,
+and they cannot perform a \( \Atau \)-action. However, in the context 
+$C = \Ppar{\ctxhole}{\Pin{x}{l}{\Pend}}$, the relation no
+longer holds: in fact,
 $\nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}\Ppar{}{\Pin{x}{l}{\Pend}}}$
@@
 
-\subsubsection{Strong barbed congruence}
+\subsubsection{Strong barbed congruence.}
 In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,
@@
 
-We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
+% We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying
+We say that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
 
@@
 
-\subsubsection{Challenge}
+\subsubsection{Challenge.}
 The objective of this challenge is to prove a theorem that shows that
@@
 benchmark challenges. We start with some general remarks, then
-describe the individual design considerations for each challenge problem.
+describe the individual design considerations for each challenge problem,
+and the evaluation criteria.
 % We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
+
+\subsubsection{General design remarks.}
 Like the previous POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
-applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17} are not directly covered.
-Still, these (and other) proof mechanisation applications will need the basic techniques that our challenge problems exercise:
+applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.
+Still, these (and other) proof mechanisation applications need the basic techniques that our challenge problems exercise:
 as per our design goal \ref{item:goal-reusability},
-our problems are drawn from the basic metatheory of process calculi (instead
-of requiring the development of new theory), and focus on well-known theorems
+our problems are drawn from the basic metatheory of process calculi (without requiring the development of new theory), and focus on well-known theorems
 and results that involve interesting proof techniques that may be reused in further work.
 To achieve our design goal \ref{item:goal-comperison-accessibility},
-we have designed our challenge problems to exercise the three aspects (linearity, scope extrusion, coinduction) independently, and they can be solved individually and in any order.
-We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi, and type theory.  The process calculus used in each challenge
+we have designed our challenge problems to exercise the three aspects (linearity, scope extrusion, coinduction) independently, and they can be solved individually and in any order;
+each problem should be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi, and type theory.  The process calculus used in each challenge
 focuses on the features we want to exercise, and omits all constructs
 (such as choices) that would complicate the mechanisation without bringing tangible
-benefits and insights.  (It is of course possible to reintroduce the omitted
-constructs, but we leave this as a possibility for a future benchmark challenge.)
-The minimality and uniformity of the calculi also allows us to target our design
-goal \ref{item:goal-tutorials}, i.e.~encouraging the development guidelines and
+benefits and insights.\footnote{It is of course possible to reintroduce the omitted
+constructs, but we leave this as a possibility for a future benchmark challenge.}
+The minimality and uniformity of the calculi also allows us to target the design
+goal \ref{item:goal-tutorials}, i.e.~encouraging the creation of guidelines and
 tutorials for the mechanisation of concurrency theory.
@@
 
-\paragraph{Linearity}
+\subsubsection{Linearity.}
 For our challenge on linear reasoning we have chosen a safety theorem
@@
 
-\paragraph{Scope extrusion}
+\subsubsection{Scope extrusion.}
 This is the challenge most closely related to the POPLMark
@@
 
-\paragraph{Coinduction}
+\subsubsection{Coinduction.}
 %\fxerror{
@@
 Our coinduction challenge concerns strong barbed bisimulation and congruence.
-Though weak barbed congruence is a more common behavioural equivalence for \(\pi\)-terms, we avoid having to abstract over the number of internal transitions when considering the strong notions of equivalence, thus simplifying the theory.
+Though weak barbed congruence is a more common behavioural equivalence for \(\pi\)-terms, here we prefer strong equivalences to avoid the need to abstract over the number of internal transitions, % when considering the strong notions of equivalence,
+thus simplifying the theory.
 We have chosen not to include delegation in the coinduction challenge since it is orthogonal to our primary aim of exploring coinductive proof techniques.
 
-\subsection{Evaluation criteria}
-The idea of our benchmark is to obtain evidence towards answering questions \cref{item:rq1,item:rq2,item:rq3}, and we are thus interested in not only the solutions, but in the experience of solving the challenges with the chosen approach.
-Solutions to our challenges should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
+\subsubsection{Evaluation criteria.}
+The idea of our benchmark is to obtain evidence towards answering questions \cref{item:rq1,item:rq2,item:rq3}: therefore, we are interested in not only the solutions, but in the experience of solving the challenges with the chosen approach.
+Solutions to our challenges should be compared on three measures:
+\begin{itemize}
+\item the mechanisation overhead, i.e.~the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; \item the adequacy of the formal statements in the mechanisation, i.e.~whether the proven theorems are easily recognisable as the theorems from the challenge; \item the cost of entry for the tools and techniques employed, i.e.~the difficulty of learning to use the techniques.
+\end{itemize}
 Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges need to present more elaborate argumentation for their adequacy.
 
-\section{Going beyond the challenge problems}\label{sec:going-beyond}
+\section{Going beyond the challenge problems.}\label{sec:going-beyond}
 
@@
 
+\vspace{-1mm}%
 \paragraph{Extending and combining the current challenges.}
@@
 and aspects they exercise (linearity, scope extrusion, coinduction) --- but most
-research on concurrency theory combines some or all such aspects. An interesting second round of formalisations could
+research on concurrency theory combines some or all such aspects. A second round of formalisations could
 entail combining the features of the challenges.
-In \cref{tab:extensions} we consider the following three extensions: first, the
-addition of choice to the type level and the process level. Second, the addition
-of recursion (and recursion types for typed systems). And finally, the
+In \cref{tab:extensions} we consider the following three extensions:
+\emph{(1)} the addition of choice to the type level and the process level,
+\emph{(2)} the addition of recursion (and recursive types for typed systems),
+and \emph{(3)} the
 addition of channel delegation to those calculi that do not already
@@
   \begin{center}\small
+    \vspace{-4mm}%
     \begin{tabular}{|r|c|c|c|}
@@
     \end{tabular}
+    \vspace{-3mm}%
   \end{center}
@@
 
+\vspace{-1mm}%
 \paragraph{Addressing new calculi and new features.}
 Extending in this dimension
-means proposing new challenges to cover for different features of
-message passing calculi, or to cover different proof mechanisation features.
+means proposing new challenges covering different features of
+process calculi, or different proof mechanisation features.
 Some interesting aspects of message passing calculi to explore in
-further challenges could be multiparty session types, and
-choreographies because their definitions and meta-theory have aspects
-that we do not cover here (e.g.: representing well formed global types
-and projection as a partial function). Although, dually, some aspects
-that we cover here do appear in these problems. Also, one could easily
+further challenges could be multiparty session types and
+choreographies, as their meta-theory includes aspects
+that we do not cover here (e.g.: well-formed global types,
+projection as a partial function) --- although, dually, some aspects
+that we cover here do appear in their theory. Also, one could easily
 design a challenge that goes beyond our coinduction challenge, by
-exploring different notions of bisimilarity (barbed, or weak, etc.),
+exploring different notions of bisimilarity (barbed, weak, etc.),
 and trace equivalence.
-On the side of proof assistants, the current challenge proposes
-mechanising the proofs of theorems. An interesting venue to explore
-would be to take advantage of other proof assistant features. For
-example, proof assistants are often able to produce certified code (by
-using code extraction, or compiling and running their definitions).
-Additionally, while having effective automation in our current
-challenge is possible, a challenge could be designed with the explicit
-objective of automating aspects of the proofs. And finally, a
-challenge could propose the integration with other formal reasoning
-tools, namely model checkers. These tools are in common use in the
-field --- and combining automated proofs with proof assistants
-offers the potential to ease the path for attempting larger proofs.
+
+On the side of proof assistants, the current challenge proposes mechanising the
+proofs of theorems. An interesting venue to explore would be to take advantage
+of other proof assistant features. For example, proof assistants can often
+produce certified code (via code extraction, or compiling and running their
+definitions). Additionally, while our current challenge allow for proof
+automation, a more specific challenge could be designed with the objective of
+automating aspects of the proofs. Finally, a challenge could propose the
+integration with other formal reasoning tools, namely model checkers. These
+tools are in common use in the field --- and combining automated proofs with
+proof assistants offers the potential to ease the path towards larger proofs.
+
 Ultimately, the current challenges can be extended in several
@@
 the foundation for those future extensions. It is our expectation that
-the solutions of those hypothetical challenges will build on solutions
+the solutions of those future challenges will build on solutions
 to this challenge.

### 2023-10-14 challenges/linearity.tex

@@
   \begin{description}
-    \item[$P = \Pend$] By inversion on the typing we have $\tend{\Delta}$. Apply T-INACT.
+    \item[$P = \Pend$] By inversion on the typing we have $\tend{\Delta}$. Apply \TirName{T-Inact}.
     \item[$P = \Ppar{P_0}{P_1}$] By inversion on the typing we have  $ \Delta, x : T = \Delta_0 , \Delta_1$.
       By cases on which environment $x$ is in, apply IH on that environment. 
-     \item[$P = \Presd{y}{z}{P}$] We may assume $x \notin \{y,z\}$. By inversion, $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From the assumptions present we can apply IH.\\
-
-       The remaining case are analogous
-
+    \item[$P = \Presd{y}{z}{P}$] We may assume $x \notin \{y,z\}$. By inversion, $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From the assumptions present we can apply IH.
   \end{description}
+  The remaining cases are analogous.
 \end{proof}
@@
 \begin{proof}
-  By induction on $P$ inverting  $\types{(\Gamma, l);\Delta}{P}$ in each case.
+  By induction on $P$, inverting  $\types{(\Gamma, l);\Delta}{P}$ in each case.
   \begin{description}
@@
      \item[$P = \Presd{x}{y}{P}$] Immediate after inverting typing.
-     \item[$P = \Pout{x}{v}{P}$] Invert typing. we have  $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
-       We must show $\typev{\Gamma}{\hastype{v}{\Tbase}}$ to build the conclusion with IH and T-OUT. Go by cases on the struture of $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
+     \item[$P = \Pout{x}{v}{P}$] Invert typing. We have $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
+       We must show $\typev{\Gamma}{\hastype{v}{\Tbase}}$ to build the conclusion with IH and \TirName{T-Out}. Go by cases on the structure of $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
      \item[$P = \Pin{x}{v}{P}$] Immediate after inverting typing.
@@
 \begin{proof}
-  By induction on the structure of a process context involved in the congruence. In the base case where the context is a hole, proceed by case analysis on the \textsl{Sc} rule applied:
+  By induction on the structure of a process context involved in the congruence. In the base case where the context is a hole, proceed by case analysis on the \TirName{Sc} rule applied:
   \begin{description}
@@
 \begin{proof}
-  By induction on the derivation of $\reduces{P}{Q}$. The cases R-PAR
-  and R-RES follows immediately by IH. Case R-STRUCT appeals twice to
+  By induction on the derivation of $\reduces{P}{Q}$. The cases \TirName{R-Par} and \TirName{R-Res} follow immediately by IH. Case \TirName{R-Struct} appeals twice to
   preservation of $\equiv$ (Lemma~\ref{le:presequiv}) and to the
-  IH. For R-COM, suppose that T-RES introduces in $\Delta$ the
-  assumptions $x : !.U, y : ?\dual U$.  Building the only tree for the
+  IH. For \TirName{R-Com}, suppose that \TirName{T-Res} introduces in $\Delta$ the
+  assumptions $\Cadd{\hastype{x}{\Tout{U}}}{\hastype{y}{\Tin{\dual{U}}}}$.  Building the only tree for the
   hypothesis, we know that $\Delta = \Delta_1, \Delta_2, \Delta_3$
-  where $\Gamma;\Delta_3 \vdash R$. We also have
-  $ \Gamma;\Delta_1 , x : U \vdash P$, ${\cal D}_{2} $ a proof of
-  $\Gamma,l;\Delta_2, y : \dual U \vdash Q$ and $\cal V $ is a proof of
-  $\Gamma \vdash a : \Tbase$.  From ${\cal D}_{2}$ and $\cal V$ we use the
-  substitution Lemma~\ref{le:subst} to obtain $\Gamma;\Delta_2 , y : \dual U \vdash
-  Q[a/l]$. We then conclude the proof with rules T-PAR, T-PAR and
-  T-RES.
+  where $\types{\Gamma;\Delta_3}{R}$. We also have
+  $\types{\Gamma;(\Cadd{\Delta_1}{\hastype{x}{U}})}{P}$, ${\cal D}_{2} $ a proof of
+  $\types{\Gamma,l;(\Cadd{\Delta_2}{\hastype{y}{\dual{U}}})}{Q}$ and $\cal V $ a proof of
+  $\typev{\Gamma}{\hastype{a}{\Tbase}}$.  From ${\cal D}_{2}$ and $\cal V$ we use the
+  substitution \cref{le:subst} to obtain $\types{\Gamma;\Cadd{\Delta_2}{\hastype{y}{\dual{U}}}}{\subst{Q}{a}{l}}$. We then conclude the proof with rules \TirName{T-Par} (twice) and \TirName{T-Res}.
 \end{proof}

### 2023-10-14 challenges/linearity.tex
updated some proofs

@@
   \label{lemma:strenD}
-  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
+  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and
+  $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
 \end{lemma}
 \begin{proof}
-  By induction on $P$, for each case inverting $\types{\Gamma; (\Delta, x : T)}{P}$ to the uniquely applicable rule.
+  By induction on the last applied rule of the proof of
+  $\types{\Gamma; (\Delta, x : T)}{P}$.
+  % , for each case inverting $\types{\Gamma; (\Delta, x : T)}{P}$ to
+  % the uniquely applicable rule.
   \begin{description}
-    \item[$P = \Pend$] By inversion on the typing we have $\tend{\Delta}$. Apply \TirName{T-Inact}.
-    \item[$P = \Ppar{P_0}{P_1}$] By inversion on the typing we have  $ \Delta, x : T = \Delta_0 , \Delta_1$.
-      By cases on which environment $x$ is in, apply IH on that environment. 
-    \item[$P = \Presd{y}{z}{P}$] We may assume $x \notin \{y,z\}$. By inversion, $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From the assumptions present we can apply IH.
+  \item[\TirName{T-End}] Since $\tend{\Delta}$, we can just reapply
+    the rule without $x:T$.
+    % [$P = \Pend$] we Apply \TirName{T-Inact}.
+
+  \item[\TirName{T-Par}] In this case, we have that
+    $ \Delta, x : T = \Delta_0 , \Delta_1$.  By cases on which
+    environment $x$ is in, we just apply the induction hypothesis on
+    that environment.
+    %
+    %
+    % [$P = \Ppar{P_0}{P_1}$] By inversion on the typing we have  $ \Delta, x : T = \Delta_0 , \Delta_1$.
+    %   By cases on which environment $x$ is in, apply IH on that environment. 
+  \item[\TirName{T-Res}] Without loss of generality, we assume that
+    $x \notin \{y,z\}$. Since
+    $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From
+    the assumptions present we can apply the induction hypothesis.
+    % inversion,
+    % $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From
+    % the assumptions present we can apply IH.
+    %
+    % [$P = \Presd{y}{z}{P}$] We may assume $x \notin \{y,z\}$. By
+    % inversion,
+    % $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From
+    % the assumptions present we can apply IH.
   \end{description}
@@
     \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}}$ that is
-  structurally congruent to $P$. Moreover, assume that $P_1$ is
-  prefixed at some $x_1$ and $P_2$ is prefixed at some $x_2$ such that
-  $P_1\equiv\Pout{x}{v}{P_1'}$ (the opposite case proceeds
-  similarly). We need to show that $P_2\equiv\Pin{x}{v}{P_2'}$. This
-  can be easily done by contradiction. In fact, if
-  $P_2\equiv\Pout{x}{v}{P_2'}$ then the typing rule for restriction
-  would be violated since the type of $x_1$ and $y_1$ cannot be dual.
+  structurally congruent to $P$. Clearly, by Lemma~\ref{le:presequiv},
+  well-typedness must be preserved by structural congruence. Moreover,
+  assume that $P_1$ is prefixed at $x_1$ and $P_2$ is prefixed at
+  $y_1$ such that $P_1\equiv\Pout{x_1}{v}{P_1'}$ (the opposite case
+  proceeds similarly). We need to show that
+  $P_2\equiv\Pin{y_1}{v}{P_2'}$. This can be easily done by
+  contradiction. In fact, if $P_2\equiv\Pout{x}{v}{P_2'}$ then the
+  typing rule for restriction would be violated since the type of
+  $x_1$ and $y_1$ cannot be dual.
   % 

### 2023-10-14 challenges/linearity.tex

@@
   IH. For \TirName{R-Com}, suppose that \TirName{T-Res} introduces in $\Delta$ the
-  assumptions $\Cadd{\hastype{x}{\Tout{U}}}{\hastype{y}{\Tin{\dual{U}}}}$.  Building the only tree for the
+  assumptions $\Cadd{\hastype{x}{\Tout{U}}}{\hastype{y}{\Tin{\dual{U}}}}$.  Building the only derivation for the
   hypothesis, we know that $\Delta = \Delta_1, \Delta_2, \Delta_3$
@@
   assume that $P_1$ is prefixed at $x_1$ and $P_2$ is prefixed at
-  $y_1$ such that $P_1\equiv\Pout{x_1}{v}{P_1'}$ (the opposite case
+  $y_1$ such that $\scong{P_1}{\Pout{x_1}{v}{P_1'}}$ (the opposite case
   proceeds similarly). We need to show that
-  $P_2\equiv\Pin{y_1}{v}{P_2'}$. This can be easily done by
-  contradiction. In fact, if $P_2\equiv\Pout{x}{v}{P_2'}$ then the
+  $\scong{P_2}{\Pin{y_1}{v}{P_2'}}$. This can be easily done by
+  contradiction. In fact, if $\scong{P_2}{\Pout{x}{v}{P_2'}}$ then the
   typing rule for restriction would be violated since the type of

### 2023-10-14 challenges/linearity.tex
minor fixes to subject reduction proof. Preservation of cong still in need of attention

@@
     \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
-      \item If \( \types{\Gamma; \Delta}{P} \) and \( \tend{\Delta} \) then
+  \item If \( \types{\Gamma; \Delta}{P} \) % and \( \tend{\Delta} \)
+    then
     \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
@@
 \begin{proof}
-  By induction on the last applied rule of the proof of
+  By induction on the derivation of 
+%%  last applied rule of the proof of --AM changed for uniformity
   $\types{\Gamma; (\Delta, x : T)}{P}$.
@@
   \begin{description}
-  \item[Par-Comm/Assoc:] by rearranging sub-derivations noting that
+  \item[\TirName {Par-Comm/Assoc}] By rearranging sub-derivations noting that
     order does not matter for linear contexts
-  \item[Par-Inact] right-to-left by
-    strengthening~\ref{lemma:strenD}. Vice-versa, by picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
-  \item[Res-Par] Follows Vasco's proof by case analysis on $x : T$ being linear and applying weak and stren accordingly.
-  \item[Res-Inact:] use weakening~\ref{lemma:weak} part 2
-  \item[Res] order does not matter
+  \item[\TirName{Par-Inact}]
+    % right-to-left by strengthening~\ref{lemma:strenD}. Vice-versa,
+    By picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
+  \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear and applying weak and strenghening accordingly.
+  \item[\TirName{Res-Inact:}] use weakening~\ref{lemma:weak} part 2
+  \item[\TirName{Res}] Noting that order does not matter
   \end{description}

### 2023-10-14 challenges/coinduction.tex
First draft of coinduction main theorem proof.

@@
 \end{theorem}
+\begin{proof}
+  \fxerror{This direction of the proof uses structural congruence, which we haven't defined plus some extra lemmas about it.
+The other direction is the interesting case anyway, so might as well scrap this direction of the theorem?}
+  First, assume \( \sbcong{P}{Q} \).
+  Suppose that \( \sigma \) substitutes \( y_1, \dots, y_n \) by \( z_1, \dots, z_n \).
+  For a fixed \( R \) and some fresh name \( x \), we then define
+  \begin{equation*}
+    C = \Ppar{\Ppar{\Pin{x}{y_1}{\dots \Pin{x}{y_n}{\Pend}}}{\Pout{x}{z_1}{\dots \Pout{x}{z_n}{\ctxhole}}}}{R}
+  \end{equation*}
+  and from the assumption have \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \).
+  \( \applyctx{C}{P} \) reduces to \( \Ppar{\applysubst{\sigma}{P}}{R} \) in \( n \) steps and, up to \( \scong{}{} \), the only process \( S \) such that \( \applyctx{C}{Q} \) reduces to \( S \) and not \( \observable{S}{\obsout{x}} \) in \( n \) steps is \( \Ppar{\applysubst{\sigma}{Q}}{R} \).
+  From this we conclude that \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \).
+
+  For the other direction, it suffices to show that if \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \) for any \( R \) and \( \sigma \), then \( \sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}} \) for any \( C \), \( R \) and \( \sigma \), since \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \).
+  We proceed by induction on \( C \).
+  \begin{description}
+  \item[\( C = \Pin{x}{z}{C'} \)] Let \( \mathcal{S} = \{ (\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}, \Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}) \mid \text{$R$ and $\sigma$ arbitrary} \} \cup \sbbisim{}{} \).
+    We can easily check that \( \mathcal{S} \) is a strong barbed bisimulation, noting that \( \sbbisim{}{} \) is preserved by restriction and is contained in \( \mathcal{S} \).
+  \item[\( C = \Ppar{C'}{S} \)] Then by the induction hypothesis,
+    \begin{equation*}
+      \sbbisim{\sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C'}{P}}}{(\Ppar{\applysubst{\sigma}{S}}{R})}}}{\sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C'}{Q}}}{(\Ppar{\applysubst{\sigma}{S}}{R})}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}}}
+    \end{equation*}
+    for any \( R \) and \( \sigma \).
+  \item[\( C = \Pres{z}{C'} \)] Then by the induction hypothesis we have \( \sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C'}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C'}{Q}}}{R}} \) for any \( R \) and \( \sigma \).
+    Without loss of generality, we assume that \( z \notin \freenames{R} \cup \names{\sigma} \).
+    Then, using that \( \sbbisim{}{} \) is preserved by restriction, we have
+    \begin{equation*}
+      \sbbisim{\sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Pres{z}{(\Ppar{\applysubst{\sigma}{\applyctx{C'}{P}}}{R})}}}{\sbbisim{\Pres{z}{(\Ppar{\applysubst{\sigma}{\applyctx{C'}{Q}}}{R})}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}}}
+    \end{equation*}
+  \item[\( C = \Preplicate{C'} \)] TBD (this is the one that uses up-to techniques)
+  \end{description}
+  The remaining cases are similar.
+\end{proof}

### 2023-10-15 challenges/linearity.tex
done with Lemmas and Theorem 2

@@
   \item[\TirName{T-Res}] Without loss of generality, we assume that
-    $x \notin \{y,z\}$. Since
-    $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From
-    the assumptions present we can apply the induction hypothesis.
+    $x \notin \{y,z\}$, for $P=\Presd{y}{z}{P'}$. Since
+    $\types{\Gamma;\;\Delta, y : T_0, z : \dual{T_0}, x : T}{P'}$, by
+    induction hypothesis, we have
+    $\types{\Gamma;\;\Delta, y : T_0, z : \dual{T_0}}{P'}$. Applying
+    again \TirName{T-Res}, we have
+    $\types{\Gamma;\;\Delta}{\Presd yz{P'}}$.
     % inversion,
@@
   \label{le:subst}
-  If $\types{(\Gamma, l);\Delta}{P}$ and ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
+  If $\types{(\Gamma, l);\Delta}{P}$ and
+  ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
   \( \types{\Gamma;\Delta}{\subst{P}{a}{l}} \).
@@
 \begin{proof}
-  By induction on $P$, inverting  $\types{(\Gamma, l);\Delta}{P}$ in each case.
+  By induction on the derivation of $\types{(\Gamma, l);\Delta}{P}$.
   \begin{description}
-    \item[$P = \Pend$] Immediate after inverting typing.
-    \item[$P = \Ppar{P_0}{P_1}$] By inversion on the typing we get derivations for $P_0$ and $P_1$. Apply IH on both.
-     \item[$P = \Presd{x}{y}{P}$] Immediate after inverting typing.
-     \item[$P = \Pout{x}{v}{P}$] Invert typing. We have $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
-       We must show $\typev{\Gamma}{\hastype{v}{\Tbase}}$ to build the conclusion with IH and \TirName{T-Out}. Go by cases on the structure of $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
-     \item[$P = \Pin{x}{v}{P}$] Immediate after inverting typing.
+  \item[\TirName{T-End}] Immediate since $\tend{\Delta}$.
 
+  \item[\TirName{T-Par}] For $P = \Ppar{P_0}{P_1}$, we apply the
+    induction hypothesis on the derivations for $P_0$ and $P_1$.
+
+  \item[\TirName{T-Res}] Immediate on the premise of the rule.
+
+  \item[\TirName{T-Out}] Let $P = \Pout{x}{v}{P'}$. We have that
+    $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.  We must show
+    $\typev{\Gamma}{\hastype{v}{\Tbase}}$ to build the conclusion with
+    the induction hypothesis and \TirName{T-Out}. We proceed by cases
+    on the structure of $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
+    % \item[$P = \Pin{x}{v}{P}$] Immediate after inverting typing.
   \end{description}
+  The remaining cases are analogous.
+
 \end{proof}
@@
   assume that $P_1$ is prefixed at $x_1$ and $P_2$ is prefixed at
-  $y_1$ such that $\scong{P_1}{\Pout{x_1}{v}{P_1'}}$ (the opposite case
-  proceeds similarly). We need to show that
-  $\scong{P_2}{\Pin{y_1}{v}{P_2'}}$. This can be easily done by
-  contradiction. In fact, if $\scong{P_2}{\Pout{x}{v}{P_2'}}$ then the
-  typing rule for restriction would be violated since the type of
+  $y_1$ such that $\scong{P_1}{\Pout{x_1}{v}{P_1'}}$ (the opposite
+  case proceeds similarly). We need to show that
+  $\scong{P_2}{\Pin{y_1}{l}{P_2'}}$. This can be easily done by
+  contradiction. In fact, if $\scong{P_2}{\Pout{y_1}{v}{P_2'}}$ then
+  the typing rule for restriction would be violated since the type of
   $x_1$ and $y_1$ cannot be dual.

### 2023-10-15 challenges/linearity.tex
Small typographical fixes in linearity challenge.

@@
   \label{lemma:strenD}
-  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and
+  If \( \types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P} \) and
   $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
@@
 %%  last applied rule of the proof of --AM changed for uniformity
-  $\types{\Gamma; (\Delta, x : T)}{P}$.
+  $\types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P}$.
   % , for each case inverting $\types{\Gamma; (\Delta, x : T)}{P}$ to
@@
   \item[\TirName{T-End}] Since $\tend{\Delta}$, we can just reapply
-    the rule without $x:T$.
+    the rule without $\hastype{x}{T}$.
     % [$P = \Pend$] we Apply \TirName{T-Inact}.
@@
   \item[\TirName{T-Par}] In this case, we have that
-    $ \Delta, x : T = \Delta_0 , \Delta_1$.  By cases on which
-    environment $x$ is in, we just apply the induction hypothesis on
-    that environment.
+    $\Cadd{\Delta}{\hastype{x}{T}} = \Csplit{\Delta_0}{\Delta_1}$.  By cases on which
+    context $x$ is in, we just apply the induction hypothesis on
+    that context.
     %
@@
     $x \notin \{y,z\}$, for $P=\Presd{y}{z}{P'}$. Since
-    $\types{\Gamma;\;\Delta, y : T_0, z : \dual{T_0}, x : T}{P'}$, by
+    $\types{\Gamma;(\Cadd{\Cadd{\Cadd{\Delta}{\hastype{y}{T_0}}}{\hastype{z}{\dual{T_0}}}}{\hastype{x}{T}})}{P'}$, by
     induction hypothesis, we have
-    $\types{\Gamma;\;\Delta, y : T_0, z : \dual{T_0}}{P'}$. Applying
+    $\types{\Gamma;(\Cadd{\Cadd{\Delta}{\hastype{y}{T_0}}}{\hastype{z}{\dual{T_0}}})}{P'}$. Applying
     again \TirName{T-Res}, we have
-    $\types{\Gamma;\;\Delta}{\Presd yz{P'}}$.
+    $\types{\Gamma;\Delta}{\Presd{y}{z}{P'}}$.
     % inversion,
@@
   \label{le:subst}
-  If $\types{(\Gamma, l);\Delta}{P}$ and
+  If $\types{(\Cadd{\Gamma}{l});\Delta}{P}$ and
   ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
@@
 \begin{proof}
-  By induction on the derivation of $\types{(\Gamma, l);\Delta}{P}$.
+  By induction on the derivation of $\types{(\Cadd{\Gamma}{l});\Delta}{P}$.
   \begin{description}
@@
   \item[\TirName{T-Out}] Let $P = \Pout{x}{v}{P'}$. We have that
-    $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.  We must show
+    $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.  We must show
     $\typev{\Gamma}{\hastype{v}{\Tbase}}$ to build the conclusion with
     the induction hypothesis and \TirName{T-Out}. We proceed by cases
-    on the structure of $\typev{(\Gamma,l)}{\hastype{v}{\Tbase}}$.
+    on the structure of $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.
     % \item[$P = \Pin{x}{v}{P}$] Immediate after inverting typing.
@@
 \( n \geq 0 \), it holds that, if \( P_1 \) is prefixed at \( x_1 \)
-and \( P_2 \) is prefixed at \( y_1 \) (or viceversa), then
+and \( P_2 \) is prefixed at \( y_1 \) (or vice versa), then
 \( \Ppar{P_1}{P_2}

### 2023-10-15 challenges/coinduction.tex
Add proof of coinduction challenge.

--- 2023-10-13 challenges/coinduction.tex
@@
 The result establishes that two processes are strong barbed congruent
-if and only if the processes obtained by applying a substitution to
+if the processes obtained by applying a finite number of substitutions to
 them and composing them in parallel with an arbitrary process are
@@
 making strong barbed bisimilarity sensitive to substitution and
-parallel composition is equivalent to strong barbed
-congruence. Formally,
+parallel composition is enough to show strong barbed
+congruence.
+To prove the theorem, we will use an \emph{up-to technique}.
+A relation \( \mathcal{S} \) is called a \emph{strong barbed bisimulation up to \( \sbbisim{}{} \)} if, whenever \( (P,Q) \in \mathcal{S} \), the following conditions hold:
+\begin{enumerate}
+\item \( \observable{P}{\mu} \) if and only if \( \observable{Q}{\mu} \).
+\item if \( \transition{P}{\Atau}{P'} \) then \( \transition{Q}{\Atau}{Q'} \) for some \( Q' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
+\item if \( \transition{Q}{\Atau}{Q'} \) then \( \transition{P}{\Atau}{P'} \) for some \( P' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
+\end{enumerate}
+These relations are useful owing to the following lemma:
+\begin{lemma}\label{lemma:up-to}
+  If \( \mathcal{S} \) is a strong barbed bisimulation up to \( \sbbisim{}{} \), \( (P,Q) \in \mathcal{S} \) implies \( \sbbisim{P}{Q} \).
+\end{lemma}
+\begin{proof}
+  We check that \( \uptosbb{}{\mathcal{S}}{} \) is a strong barbed bisimulation and is thus included in \( \sbbisim{}{} \).
+\end{proof}
+
+We are now ready to prove the theorem:
 \begin{theorem}
-  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and
+  \( \sbcong{P}{Q} \) if, for any process \( R \) and
   substitution \( \sigma \),
@@
 \end{theorem}
+\begin{proof}
+  Since \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \), it suffices to show that if \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \) for any \( R \) and \( \sigma \), then \( \sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}} \) for any \( C \), \( R \) and \( \sigma \).
+  We proceed by induction on \( C \).
+  \begin{description}
+  \item[\( C = \Pin{x}{z}{C'} \)] Let \( \mathcal{S} = \{ (\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}, \Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}) \mid \text{$R$ and $\sigma$ arbitrary} \}~\cup \sbbisim{}{} \).
+    We can easily check that \( \mathcal{S} \) is a strong barbed bisimulation, noting that \( \sbbisim{}{} \) is preserved by restriction and is contained in \( \mathcal{S} \).
+  \item[\( C = \Ppar{C'}{S} \)] Then by the induction hypothesis,
+    \begin{equation*}
+      \sbbisim{\sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C'}{P}}}{(\Ppar{\applysubst{\sigma}{S}}{R})}}}{\sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C'}{Q}}}{(\Ppar{\applysubst{\sigma}{S}}{R})}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}}}
+    \end{equation*}
+    for any \( R \) and \( \sigma \).
+  \item[\( C = \Pres{z}{C'} \)] Then by the induction hypothesis we have \( \sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C'}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C'}{Q}}}{R}} \) for any \( R \) and \( \sigma \).
+    Without loss of generality, we assume that \( z \notin \freenames{R} \cup \names{\sigma} \).
+    Then, using that \( \sbbisim{}{} \) is preserved by restriction, we have
+    \begin{equation*}
+      \sbbisim{\sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Pres{z}{(\Ppar{\applysubst{\sigma}{\applyctx{C'}{P}}}{R})}}}{\sbbisim{\Pres{z}{(\Ppar{\applysubst{\sigma}{\applyctx{C'}{Q}}}{R})}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}}}
+    \end{equation*}
+  \item[\( C = \Preplicate{C'} \)] Let \( \mathcal{S} = \{ (\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}, \Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}) \mid \text{$R$ and $\sigma$ arbitrary} \} \).
+    Using \cref{lemma:up-to}, it suffices to show that \( \mathcal{S} \) is a strong barbed bisimulation up to \( \sbbisim{}{} \).
+    To this end, let
+    \begin{align*}
+      A &= \applysubst{\sigma}{\applyctx{C'}{P}}, & A' &= \applysubst{\sigma}{\applyctx{C}{P}} \\
+      B &= \applysubst{\sigma}{\applyctx{C'}{Q}}, & B' &= \applysubst{\sigma}{\applyctx{C}{Q}} ,
+    \end{align*}
+    noting that \( A' = \Preplicate{A} \) and \( B' = \Preplicate{B} \).
+
+    Suppose \( \transition{\Ppar{R}{A'}}{\Atau}{S} \) for some \( S \).
+    Then we can show by a case analysis on the derivation of this transition that there exists a \( T \) such that \( \transition{\Ppar{R}{(\Ppar{A}{A})}}{\Atau}{T} \) and \( \sbbisim{S}{\Ppar{T}{A'}} \).
+    Using the induction hypothesis twice, we note that \( \sbbisim{\Ppar{R}{B'}}{\Ppar{R}{\Ppar{(\Ppar{A}{A})}{B'}}} \).
+    Since by rule \TirName{Par-L}, \( \transition{\Ppar{\Ppar{R}{(\Ppar{A}{A})}}{B'}}{\Atau}{\Ppar{T}{B'}} \), there must thus exist a \( U \) such that \( \transition{\Ppar{R}{B'}}{\Atau}{U} \), \( \sbbisim{U}{\Ppar{T}{B'}} \) and \( \uptosbb{S}{\mathcal{S}}{U} \) as required.
+    The proof for \( \transition{\Ppar{R}{B'}}{\Atau}{S} \) is analogous.
+  \end{description}
+  The remaining cases are similar.
+\end{proof}

### 2023-10-15 paper/paper.tex
Add proof of coinduction challenge.

@@
 making strong barbed bisimilarity sensitive to substitution and
-parallel composition is equivalent to strong barbed
+parallel composition is enough to show strong barbed
 congruence. Formally,
 \begin{theorem}
-  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and
+  \( \sbcong{P}{Q} \) if for any process \( R \) and
   substitution \( \sigma \),
@@
 
-\section{Going beyond the challenge problems.}\label{sec:going-beyond}
+\section{Going beyond the challenge problems}\label{sec:going-beyond}
 

### 2023-10-15 challenges/linearity.tex
first draft of complete proof of preservation of \scong

@@
 \end{proof}
-\begin{lemma}[Preservation for $\equiv$]
-  \label{le:presequiv}
-  If \( \types{\Gamma;\Delta}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
+
+\newcommand{\sconga}[2]{#1 \stackrel{a}{\equiv} #2}
+To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms.
+
+\begin{lemma}[Preservation for $\stackrel{a}{\equiv}$]
+  \label{le:presequiva}
+  If \( \sconga{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
 \end{lemma}
 \begin{proof}
-  By induction on the structure of a process context involved in the congruence. In the base case where the context is a hole, proceed by case analysis on the \TirName{Sc} rule applied:
+  By case analysis on the \TirName{Sc} rule applied:
   \begin{description}
@@
   \item[\TirName{Par-Inact}]
-    % right-to-left by strengthening~\ref{lemma:strenD}. Vice-versa,
-    By picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
-  \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear and applying weak and strenghening accordingly.
-  \item[\TirName{Res-Inact:}] use weakening~\ref{lemma:weak} part 2
+   Right-to-left by strengthening~\ref{lemma:strenD}. Vice-versa,
+    by picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
+  \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear or $\tend$ and applying weakening and strengthening accordingly.
+  \item[\TirName{Res-Inact:}] Use weakening~\ref{lemma:weak} part 2
   \item[\TirName{Res}] Noting that order does not matter
   \end{description}
-  For the step case, TBD
+\end{proof}
+
+Now, following Sangiorgi \& Walker define as the compatible equivalence relation induced by $\sconga\cdot\cdot$, namely add to reflexivity, symmetry  and transitivity the following:
+ \begin{mathpar}
+   \inferrule[Cong]{\sconga P Q}{\scong {C[P]}{C[Q]}}
+ \end{mathpar}
+
+\begin{lemma}[Preservation for $\equiv$]
+  \label{le:presequiv}
+    If \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
+\end{lemma}
+\begin{proof}
+  By induction on the structure of the derivation of \( \scong{P}{Q} \), with an inner induction of the structure of a process context.
+  \begin{description}
+  \item[\TirName {Refl}] Immediate
+      \item[\TirName {Sym}] By IH.
+      \item[\TirName {Trans}]  By two appeal to the IH
+          \item[\TirName {Cong}] By induction on the structure of $C$.
+If the context is a hole, apply Lemma~\ref{le:presequiva}.
+In the step case, apply the IH: for example if $C$ has the form ${\Ppar C R}$, noting that ${\PBpar C R}[P]$ is equal to ${\Ppar {C[P]} R}$ we have \( \types{\Gamma;\Delta}{{\PBpar C R}[P]} \)  iff \( \types{\Gamma;\Delta}{{\PBpar C R}[Q]} \) by rule T-Par and the IH.
+  \end{description}
 \end{proof}

### 2023-10-16 challenges/linearity.tex
Backport my changes into challenges/.

@@
 This challenge formalises a proof that requires reasoning about linearity of channels.
-Linearity is the notion that a channel must be used exactly once in a process.
-This is necessary to prove properties about session type systems.
+Linearity is the notion that a channel must be used exactly once in a process
+This is necessary to prove properties about session type systems, and the key
+issue of this challenge is reasoning about the linearity of context splitting operations.
 Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
@@
     by picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
-  \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear or $\tend$ and applying weakening and strengthening accordingly.
-  \item[\TirName{Res-Inact:}] Use weakening~\ref{lemma:weak} part 2
+  \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear or $\Tend$ and applying weakening and strengthening accordingly.
+  \item[\TirName{Res-Inact:}] Use weakening %~\ref{lemma:weak} part 2
   \item[\TirName{Res}] Noting that order does not matter
@@
 
-Now, following Sangiorgi \& Walker define as the compatible equivalence relation induced by $\sconga\cdot\cdot$, namely add to reflexivity, symmetry  and transitivity the following:
+Now, following Sangiorgi \& Walker~\cite{picalcbook}, we give rules for the compatible
+equivalence relation induced by $\sconga\cdot\cdot$, which we still
+write as $\scong\cdot\cdot$: namely, add to reflexivity, symmetry and
+transitivity the following condition:
  \begin{mathpar}
@@
       \item[\TirName {Trans}]  By two appeal to the IH
-          \item[\TirName {Cong}] By induction on the structure of $C$.
-If the context is a hole, apply Lemma~\ref{le:presequiva}.
-In the step case, apply the IH: for example if $C$ has the form ${\Ppar C R}$, noting that ${\PBpar C R}[P]$ is equal to ${\Ppar {C[P]} R}$ we have \( \types{\Gamma;\Delta}{{\PBpar C R}[P]} \)  iff \( \types{\Gamma;\Delta}{{\PBpar C R}[Q]} \) by rule T-Par and the IH.
+      \item[\TirName {Cong}] By induction on the structure of $C$. If
+        the context is a hole, apply Lemma~\ref{le:presequiva}. In the
+        step case, apply the IH: for example if $C$ has the form
+        ${\Ppar {C'} R}$, noting that ${\PBpar{ C'} R}[P]$ is equal to
+        ${\Ppar {C'[P]} R}$ we have
+        \( \types{\Gamma;\Delta}{{\PBpar {C'} R}[P]} \) iff
+        \( \types{\Gamma;\Delta}{{\PBpar {C'} R}[Q]} \) by rule T-Par and
+        the IH.
   \end{description}

### 2023-10-16 challenges/scope-extrusion.tex
Remove mention of equality on names since it is obvious and confusing.
Backport my changes into challenges/.

@@
 This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
+The key issue of this challenge is reasoning about names that are ``in the process'' of being scope-extruded, which often presents difficulties for the mechanisation of binders.
 
@@
 \subsubsection{Syntax}
-In this section we assume that \emph{names}  can be compared for equality.
 The syntax of processes is given by:

### 2023-10-16 challenges/linearity.tex
minor fixes to Lemma 13

@@
 This challenge formalises a proof that requires reasoning about linearity of channels.
-Linearity is the notion that a channel must be used exactly once in a process
-This is necessary to prove properties about session type systems, and the key
-issue of this challenge is reasoning about the linearity of context splitting operations.
+Linearity is the notion that a channel must be used exactly once in a process.
+This is necessary to prove properties about session type systems.
 Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
@@
 
-Now, following Sangiorgi \& Walker~\cite{picalcbook}, we give rules for the compatible
+Now, following Sangiorgi \& Walker, we give rules for the compatible
 equivalence relation induced by $\sconga\cdot\cdot$, which we still

### 2023-10-16 challenges/coinduction.tex
Backport my changes into challenges/.
Backport Alceste's changes into challenges/.

@@
-\subsection{Challenge: Coinduction}
+\subsection{Challenge: Coinduction and infinite processes}
 % and reasoning about infinite processes}
@@
 them and composing them in parallel with an arbitrary process are
-strongly barbed bisimilar.  Which, to reiterate, entails the use of
-coinduction due to the presence of infinite behaviour.
+strongly barbed bisimilar.
+The key issue of this challenge is the coinductive reasoning about the infinite behaviours of the replication operator.
 
-\paragraph{Syntax.}
+\subsubsection{Syntax.}
 % We assume the existence of some type of \emph{base values}, values
@@
 \begin{align*}
-  \alpha\ :=\  \Aoutf{x}{a} \ \mid\ \Ain{x}{a} \ \mid\ \Atau
+  \alpha\ \Coloneqq\ \ \Aoutf{x}{a} \ \mid\ \Ain{x}{a} \ \mid\ \Atau
 \end{align*}
@@
 \( \observable{P}{\mu} \) as follows:
-\begin{align}
+\begin{align*}
   \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
   \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align}
+\end{align*}
 
@@
 is because \( \obsout{x} \) is the only observable in both processes
-and they cannot perform a \( \Atau \)-action. However, if composed in
-parallel with the process $\Pin{x}{l}{\Pend}$, the relation will no
-longer hold. In fact,
+and they cannot perform a \( \Atau \)-action. However, in the context
+$C = \Ppar{\ctxhole}{\Pin{x}{l}{\Pend}}$, the relation no
+longer holds: in fact,
 $\nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}\Ppar{}{\Pin{x}{l}{\Pend}}}$

### 2023-10-16 challenges/linearity.tex
Backport Alceste's changes into challenges/.

--- 2023-10-16 challenges/linearity.tex
@@
 
-\subsubsection{Challenge}
+\subsubsection{Challenge.}
 The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with some lemmas:

### 2023-10-16 challenges/preliminaries.tex
Backport Alceste's changes into challenges/.

@@
 
-First, we list some common notions and conventions that we use in the challenges. Since the calculi under study are somewhat different, each section will list the changes that apply.
+First, we list some common notions and conventions that we use in the challenges. Since the calculi under study are somewhat different, each section lists the changes that apply.
 
@@
 \( l, m, \dots \), and the existence of some set of \emph{names},
-represented by the symbols \( x, y, \dots \).
+represented by the symbols \( x, y, \dots \).%
+\footnote{%
+  Unlike the standard $\pi$-calculus, we distinguish variables from names to
+  better control the expressiveness of the calculi under study, and the scope of
+  the corresponding challenges: the key distinction is that names can be sent
+  and received (thus realising channel delegation), whereas variables can only
+  be substituted with base values.%
+} %
+\fxerror{Double-check the footnote!}%
+We assume that all of these sets are infinite and that their elements can be
+compared for equality.
 We assume that all of these sets are infinite and that their elements can be compared for equality.
@@
 For communication, processes include \emph{input} and \emph{output}, whose
-signature will depend on the calculus being value or name-passing.
-We use here the metavariables $c,k$ to abstract over this choice.
+signature signature depends on the calculus being value-passing or name-passing.
+We use here the metavariables $c,k$ to abstract over this choice
+--- i.e.~$c$ may be either a value or a variable or a name, whereas $k$ may
+be a variable or a name.
 The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
@@
 It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-Replication makes it possible to express infinite behaviors.
+Replication makes it possible to express infinite behaviours.
 
-We will use the notation \( \freenames{P} \) to denote the set of
+We use the notation \( \freenames{P} \) to denote the set of
 names that occur free,

### 2023-10-16 challenges/scope-extrusion.tex
Backport Alceste's changes into challenges/.

@@
 Note also that due to rule \TirName{R-Com}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
-Finally, note that rule \TirName{R-Struct} allows for applying the structural congruence both before and after the reduction.
-This ensures that the reduction relation is closed under structural congruence.
+Finally, note that rule \TirName{R-Struct} allows for applying the structural congruence both before and after the reduction:
+this makes the reduction relation closed under structural congruence.
 
-\subsubsection{Transition system semantics}
+\subsubsection{Transition system semantics.}
 The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which is as follows:
+The transitions are labelled by \emph{actions}, the syntax of which is:
 \[
-\begin{array}{rcl}
-  \alpha & :=& \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
+\begin{array}{r@{\qquad}c@{\qquad}l}
+  \alpha & \Coloneqq & \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
 \end{array}
@@
 
-We will extend the notion of free and bound occurrences with 
+We extend the notion of free and bound occurrences with 
 \( \freenames{\alpha} \) to denote the set of names that occur free in
@@
 
-\subsubsection{Challenge}
+\subsubsection{Challenge.}
 \begin{lemma}\label{se-lemma-harmony-fact}
@@
 \end{proof}
-For the proof of the challenge theorem, we will introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
+For the proof of the challenge theorem, we introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
 The first rule applied is \TirName{R-Com}. The derivation continues with an application of \TirName{R-Par}, followed by zero or more applications of \TirName{R-Res}.

### 2023-10-16 challenges/preliminaries.tex
Fix footnote about difference between names and variables.

@@
   better control the expressiveness of the calculi under study, and the scope of
-  the corresponding challenges: the key distinction is that names can be sent
-  and received (thus realising channel delegation), whereas variables can only
-  be substituted with base values.%
+  the corresponding challenges: the key distinction is that names are used as communication channels (and can be sent and received in the scope extrusion challenge), whereas variables are only bound by inputs and cannot be restricted, sent nor received, in the style of value-passing CCS \cite{Milner1989}.%
 } %
-\fxerror{Double-check the footnote!}%
 We assume that all of these sets are infinite and that their elements can be

### 2023-10-16 challenges/scope-extrusion.tex
Fix footnote about difference between names and variables.

@@
 
-The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
+The setting for this challenge is a ``classic'' untyped \( \pi \)-calculus, where (unlike the calculi in the other challenges) names can be sent and received, and bound by input constructs (similarly to variable names in the other calculi).
 We define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.

### 2023-10-16 paper/paper.tex
add citation for the book
Fix footnote about difference between names and variables.

@@
 challenges. Since the calculi under study are somewhat different, each section
-lists the changes that apply.
+lists the changes that apply. The challenges are inspired by simplified versions of calculi (and proofs) from~\cite{picalcbook} and~\cite{Vasconcelos2012}.
 

### 2023-10-16 challenges/preliminaries.tex
Change all instances of "variable name" into "variable" to avoid confusion and save space.

@@
 \( a, b, \dots \), the existence of some set
-of \emph{variable names}, represented by the symbols
+of \emph{variables}, represented by the symbols
 \( l, m, \dots \), and the existence of some set of \emph{names},
@@
 in \( P \) and  \( \freevars{P} \)
-to denote the set of variable names that occur free
+to denote the set of variables that occur free
 % (i.e.\ not bound by an input)
 in \( P \).  We use the notation \( \boundvars{P} \)
-for the set of variable names that occur bound
+for the set of variables that occur bound
 % (by an input)
 in \( P \).  We use the notation \( \subst{P}{a}{l} \) to denote
-the process \( P \) with base value \( a \) substituted for variable name
+the process \( P \) with base value \( a \) substituted for variable
 \( l \). Similarly, \( \subst{P}{x}{y} \) denotes the process
@@
 written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \)
-by a finite number of substitutions of bound variable names.  As a
+by a finite number of substitutions of bound variables.  As a
 convention, we identify \( \alpha \)-convertible processes and we
-assume that bound names and bound variable names of any processes are
-chosen to be different from the names and variable names that occur
+assume that bound names and bound variables of any processes are
+chosen to be different from the names and variables that occur
 free in any other entities under consideration, such as processes,
-substitutions, and sets of names or variable names.  This is justified
-because any overlapping names and variable names may be
+substitutions, and sets of names or variables.  This is justified
+because any overlapping names and variables may be
 \( \alpha \)-converted such that the assumption is satisfied.
@@
 a context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement is literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+The replacement is literal, so names and variables that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
 

### 2023-10-16 challenges/scope-extrusion.tex
Change all instances of "variable name" into "variable" to avoid confusion and save space.

--- 2023-10-16 challenges/scope-extrusion.tex
@@
 
-The setting for this challenge is an untyped \( \pi \)-calculus, the syntax and semantics of which are presented below.
+The setting for this challenge is a ``classic'' untyped \( \pi \)-calculus, where (unlike the calculi in the other challenges) names can be sent and received, and bound by input constructs (similarly to variables in the other calculi).
 We define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.

### 2023-10-16 paper/paper.tex
Change all instances of "variable name" into "variable" to avoid confusion and save space.

@@
 \]
-where a \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).
+where a \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable \( l \).
 
@@
 The intention is that the value \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) waits for a base value from \( x \) and then continues as \( P \) with the received value substituted for the variable name \( l \).
+The input process \( \Pin{x}{l}{P} \) waits for a base value from \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
 %
@@
 Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable name or a name.
+On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable or a name.
 
@@
 
-The setting for this challenge is a ``classic'' untyped \( \pi \)-calculus, where (unlike the calculi in the other challenges) names can be sent and received, and bound by input constructs (similarly to variable names in the other calculi).
+The setting for this challenge is a ``classic'' untyped \( \pi \)-calculus, where (unlike the calculi in the other challenges) names can be sent and received, and bound by input constructs (similarly to variables in the other calculi).
 We define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.
@@
 % of which we will denote by \( a, b, \dots \), the existence of some
-% type of \emph{variable names}, values of which we will denote by
+% type of \emph{variables}, values of which we will denote by
 % \( l, m, \dots \), and the existence of some type of \emph{names},
@@
 % A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a
-% variable name \( l \).  The process \( \Pend \) is \emph{inaction}:
+% variable \( l \).  The process \( \Pend \) is \emph{inaction}:
 % a process which can do nothing.  The process \( \Pout{x}{a}{P} \) is
@@
 % continue as \( P \) with the received value substituted for the
-% variable name \( l \).  The input operator thus binds the variable
-% name \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
+% variable \( l \).  The input operator thus binds the variable
+%  \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
 % \emph{composition} of process \( P \) and process \( Q \).  The two
@@
 % We will use the notation \( \subst{P}{v}{l} \) to denote the process
-% \( P \) with value \( v \) substituted for variable name \( l \).
+% \( P \) with value \( v \) substituted for variable \( l \).
 
@@
 % written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from
-% \( P \) by a finite number of substitutions of bound variable names.
+% \( P \) by a finite number of substitutions of bound variables.
 % As a convention, we will identify \( \alpha \)-convertible
@@
 % Also as a convention, we assume that the bound names and bound
-% variable names occurring in any collection of processes are chosen
-% to be different from the free names and free variable names
-% occurring in those processes and from the names and variable names
+% variables occurring in any collection of processes are chosen
+% to be different from the free names and free variables
+% occurring in those processes and from the names and variables
 % occurring in any substitutions applied to the processes.  This is
-% justified because any overlapping names and variable names may be
+% justified because any overlapping names and variables may be
 % \( \alpha \)-converted such that the assumption is satisfied.
@@
 Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable name.
-
-% As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
+On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable.
+
+% As a convention, we assume that bound names and bound variables of any processes or actions are chosen to be different from the names and variables that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variables.
 
@@
 % A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variable names that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+% The replacement should be literal, so names and variables that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
 

### 2023-10-16 challenges/coinduction.tex
Change all instances of "variable name" into "variable" to avoid confusion and save space.
some syntax for coinduction

@@
 
-This challenge is a about the mechanisation of proofs
+This challenge is  about the mechanisation of proofs concerning infinite behaviours.
 % formalises a proof
-that require coinduction.
+%that require coinduction.
 % coinductive techniques.
@@
 This is usually connected to coinductive definitions where an infinite
-structure is defined as greatest fixed points of a recursive
+structure is defined as the greatest fixed point of a recursive
 definition.

### 2023-10-16 paper/paper.tex
some syntax for coinduction

--- 2023-10-16 paper/paper.tex
@@
 \label{sec:challenge:coinduction}
-This challenge is a about the mechanisation of proofs
+This challenge is a about the mechanisation of proofs  concerning infinite behaviours.
 % formalises a proof
-that require coinduction.
+% that require coinduction.
 % coinductive techniques.
@@
 % Coinduction is a proof technique for infinite structures,
-This is usually connected to coinductive definitions where an infinite
-structure is defined as greatest fixed points of a recursive
+This is usually connected to \emph{coinductive} definitions where an infinite
+structure is defined as the greatest fixed point of a recursive
 definition.
 %
-Coinduction is then a technique for defining and proving properties of
+Coinduction is  a technique for defining and proving properties of
 such infinite structures,
@@
                \quad \mid\quad \Pin{x}{l}{P}
-               \quad \mid\quad \Ppar{P}{Q}
+               \quad \mid\quad \PBpar{P}{Q}
                \quad \mid\quad \Pres{x}{}{P}
@@
 semantics which features fewer rules but needs a structural
-congruence. The semantics of the system describes the actions that the
-system can perform by defining a labelled transition relation on
-processes.  The transitions are labelled by \emph{actions}, the syntax
+congruence. %  The semantics of the system describes the actions that the
+% system can perform by defining a labelled transition relation on
+% processes. 
+The transitions are labelled by \emph{actions}, the syntax
 of which is as follows:
@@
 names that occur in the action \( \alpha \).
-The transition relation is then defined by the following rules:
+The transition relation is defined by the following rules:
 \begin{mathpar}
@@
 % .  Our notion of process equivalence relations
-builds on a notion of \emph{observables}, i.e., what we can externally
-observe from the semantics of a process. If we allowed ourselves only to observe
-internal transitions (i.e., observe that a process is internally
-performing a step of computation) we would relate either too few
-processes (in the strong case where we relate only processes with
-exactly the same number of internal transitions) or every process (in
-the weak case where we relate processes with any amount of internal
-transitions).  We must therefore allow ourselves to observe more than
-just internal transitions, and here we choose to define the observables
-of a process as the names it can use for sending and receiving.
+builds on the idea of \emph{observables}, i.e., what we can externally
+observe from the semantics of a process. If we allowed ourselves only
+to observe internal transitions (i.e., observe that a process is
+internally performing a step of computation) we would relate either
+too few processes (in the strong case where we relate only processes
+with exactly the same number of internal transitions) or every process
+(in the weak case where we relate processes with any amount of
+internal transitions).  As a result, we must allow ourselves to
+monitor more than just internal transitions, and we choose to describe
+a process's observables as the names it might use for sending and
+receiving.
+% We must therefore allow ourselves to observe more than
+% just internal transitions, and here we choose to define the observables
+% of a process as the names it can use for sending and receiving.
 To this end, we define the \emph{observability predicate}
@@
 
-\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is then the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
+\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
 \begin{gather*}
@@
 \begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
+  \( \sbbisim{}{} \) is an equivalence relation. %, that is, the relation is reflexive, symmetric, and transitive.
 \end{theorem}

### 2023-10-16 paper/paper.tex
just style

@@
 To achieve our design goal \ref{item:goal-comperison-accessibility},
-we have designed our challenge problems to exercise the three aspects (linearity, scope extrusion, coinduction) independently, and they can be solved individually and in any order;
-each problem should be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi, and type theory.  The process calculus used in each challenge
-focuses on the features we want to exercise, and omits all constructs
+we have drafted our challenge problems to exercise the three aspects (linearity, scope extrusion, coinduction) independently, so that they can be solved individually and in any order;
+each problem should be reasonably small and easy to understand with basic knowledge of textbook concurrency theory, process calculi, and type theory.  The process calculus used in each challenge
+focuses on the features that we want to exercise, and omits all constructs
 (such as choices) that would complicate the mechanisation without bringing tangible
@@
 \subsubsection{Linearity.}
-For our challenge on linear reasoning we have chosen a safety theorem
+For our challenge on linear reasoning we have chosen a type preservation theorem
 for a session type system since this allows us to involve linear
@@
 produce certified code (via code extraction, or compiling and running their
-definitions). Additionally, while our current challenge allow for proof
+definitions). Additionally, while our current challenge is agnostic about proof
 automation, a more specific challenge could be designed with the objective of

### 2023-10-16 challenges/coinduction.tex
more style stuff

@@
 % continue as \( P \) with the received value substituted for the
-% variable name \( l \).  The input operator thus binds the variable
-% name \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
+% variable \( l \).  The input operator thus binds the variable
+% \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
 % \emph{composition} of process \( P \) and process \( Q \).  The two
@@
 Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable name.
+On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable.
 
-% As a convention, we assume that bound names and bound variable names of any processes or actions are chosen to be different from the names and variable names that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variable names.
+% As a convention, we assume that bound names and bound variables of any processes or actions are chosen to be different from the names and variables that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variables.
 
@@
   \item Reflexivity is straightforward: for any $P$, we need to show
-    that $P\sbbisim{}{} P$. Both conditions in
-    equations~\ref{eq:bisim1} and \ref{eq:bisim2} follow trivially.
-  \item Symmetry follows immediately since bisimilarity is defined as
-    a {\em symmetric} relation.
+    that $P\sbbisim{}{} P$. Both conditions ~\ref{eq:bisim1} and \ref{eq:bisim2} follow trivially.
+  \item Symmetry follows immediately by definition.
+%    since bisimilarity is defined as a {\em symmetric} relation.
   \item For transitivity, we need to prove that if $P\sbbisim{}{} Q$

### 2023-10-16 challenges/coinduction.tex
Backport Alberto's changes to the coinduction challenge.

@@
 % Coinduction is a proof technique for infinite structures,
-This is usually connected to coinductive definitions where an infinite
+This is usually connected to \emph{coinductive} definitions where an infinite
 structure is defined as the greatest fixed point of a recursive
@@
 %
-Coinduction is then a technique for defining and proving properties of
+Coinduction is a technique for defining and proving properties of
 such infinite structures,
@@
 semantics which features fewer rules but needs a structural
-congruence. The semantics of the system describes the actions that the
-system can perform by defining a labelled transition relation on
-processes.  The transitions are labelled by \emph{actions}, the syntax
+congruence. %  The semantics of the system describes the actions that the
+% system can perform by defining a labelled transition relation on
+% processes.
+
+The transitions are labelled by \emph{actions}, the syntax
 of which is as follows:
@@
 
-The transition relation is then defined by the following rules:
+The transition relation is defined by the following rules:
 \begin{mathpar}
@@
 the weak case where we relate processes with any amount of internal
-transitions).  We must therefore allow ourselves to observe more than
-just internal transitions, and here we choose to define the observables
-of a process as the names it can use for sending and receiving.
+transitions).  As a result, we must allow ourselves to observe more than
+just internal transitions, and we choose to describe a process's observables
+as the names it might use for sending and receiving.
 
@@
 
-\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is then the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
+\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
 \begin{gather}
@@
 \begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation, that is, the relation is reflexive, symmetric, and transitive.
+  \( \sbbisim{}{} \) is an equivalence relation. %that is, the relation is reflexive, symmetric, and transitive.
 \end{theorem}

### 2023-10-16 paper/paper.tex
Move references in front of appendix to make appendix more obvious.
Backport Alberto's changes to the coinduction challenge.
Copy new material from linearity and coinduction into paper.tex.

@@
   \item If \( \types{\Gamma; \Delta}{P} \) % and \( \tend{\Delta} \)
-        then
-    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
+    then \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
   \end{enumerate}
-  % \begin{metanote}
-  %   Note from Dawit: Does \( \Delta \) really have to be only ends to be weakened?
-I had expected that if \( P \) is typed by \( \Delta \) and \( \Delta' \) is only ends, then \( P \) is typed by \( \Cadd{\Delta}{\Delta'} \).
-
-  %   AM:  I agree that it's weird: we should not be able to weaken a linear context, if not
-  %   with unrestricted types --- but this is what we need for Sc-Res-Inact to be type preserving:
-  %   from $\Delta\vdash 0$ we need $\Delta\vdash (\nu x y)0$. BTW, this is also a problem in Vasco's system
-  % \end{metanote}
 \end{lemma}
@@
   \label{lemma:strenD}
-  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
+  If \( \types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P} \) and
+  $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
 \end{lemma}
@@
 \begin{lemma}[Substitution]\mbox{}
-  If $\types{(\Gamma, l);\Delta}{P}$ and ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
-  \( \types{\Gamma;\Delta}{\subst{P}{l}{a}} \).
+  \label{le:subst}
+  If $\types{(\Cadd{\Gamma}{l});\Delta}{P}$ and
+  ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
+  \( \types{\Gamma;\Delta}{\subst{P}{a}{l}} \).
 \end{lemma}
 
+To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms.
+
+\begin{lemma}[Preservation for $\stackrel{a}{\equiv}$]
+  \label{le:presequiva}
+  If \( \sconga{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
+\end{lemma}
+
+Now, following Sangiorgi \& Walker~\cite{picalcbook}, we give rules for the compatible
+equivalence relation induced by $\sconga\cdot\cdot$, which we still
+write as $\scong\cdot\cdot$: namely, add to reflexivity, symmetry and
+transitivity the following condition:
+ \begin{mathpar}
+   \inferrule[Cong]{\sconga P Q}{\scong {C[P]}{C[Q]}}
+ \end{mathpar}
+
 \begin{lemma}[Preservation for $\equiv$]
-  If \( \types{\Gamma;\Delta}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
+  \label{le:presequiv}
+    If \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
 \end{lemma}
 
-To formulate safety, we need to formally define what we mean by well-formed process.
-We say that a process \( P \) is \emph{prefixed at variable \( x \)} if it is of the form \( \Pout{x}{v}{P} \) or \( \Pin{x}{l}{P} \).
-A process is then \emph{well formed} if, for each of its structurally congruent processes of the form \( \Presd{x_1}{y_1}{\dots \Presd{x_n}{y_n}{(\Ppar{\Ppar{P}{Q}}{R})}} \), with \( n \geq 0 \), it holds that, if \( P \) is prefixed at \( x_1 \) and \( Q \) is prefixed at \( y_1 \), then \( \Ppar{P}{Q} \) is of the form \( \Ppar{\Pout{x_1}{a}{P'}}{\Pin{y_1}{l}{Q'}} \).
-
-Note that well-formed processes do not necessarily reduce. For example, the process:
+To formulate safety, we need to formally define what we mean by
+well-formed process.  We say that a process \( P \) is \emph{prefixed
+  at variable \( x \)} if \( P\equiv\Pout{x}{v}{P'} \) or
+\( P\equiv\Pin{x}{l}{P'} \) for some $P'$.
+%
+A process $P$ is then \emph{well formed} if, for every $P_1$, $P_2$,
+and $R$ such that
+\( P\equiv \Presd{x_1}{y_1}{\dots
+  \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}} \), with
+\( n \geq 0 \), it holds that, if \( P_1 \) is prefixed at \( x_1 \)
+and \( P_2 \) is prefixed at \( y_1 \) (or vice versa), then
+\( \Ppar{P_1}{P_2}
+\equiv\Ppar{\Pout{x_1}{a}{P_1'}}{\Pin{y_1}{l}{P_2'}} \), for some
+$P_1'$ and $P_2'$.
+
+Note that well-formed processes do not necessarily reduce. For example, the process
 \begin{equation*}
@@
 \label{sec:challenge:coinduction}
-This challenge is a about the mechanisation of proofs  concerning infinite behaviours.
+This challenge is about the mechanisation of proofs concerning infinite behaviours.
 % formalises a proof
@@
 parallel composition is enough to show strong barbed
-congruence. Formally,
+congruence.
+To prove the theorem, we will use an \emph{up-to technique}.
+A relation \( \mathcal{S} \) is called a \emph{strong barbed bisimulation up to \( \sbbisim{}{} \)} if, whenever \( (P,Q) \in \mathcal{S} \), the following conditions hold:
+\begin{enumerate}
+\item \( \observable{P}{\mu} \) if and only if \( \observable{Q}{\mu} \).
+\item if \( \transition{P}{\Atau}{P'} \) then \( \transition{Q}{\Atau}{Q'} \) for some \( Q' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
+\item if \( \transition{Q}{\Atau}{Q'} \) then \( \transition{P}{\Atau}{P'} \) for some \( P' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
+\end{enumerate}
+These relations are useful owing to the following lemma:
+\begin{lemma}\label{lemma:up-to}
+  If \( \mathcal{S} \) is a strong barbed bisimulation up to \( \sbbisim{}{} \), \( (P,Q) \in \mathcal{S} \) implies \( \sbbisim{P}{Q} \).
+\end{lemma}
+
+We are now ready to prove the theorem:
 \begin{theorem}
@@
 
+\bibliographystyle{splncs04}
+\bibliography{../references}
+
 \appendix
@@
 
-\bibliographystyle{splncs04}
-\bibliography{../references}
-
 \end{document}

### 2023-10-17 paper/paper.tex
Typo

@@
 Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
-This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication
+This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
 The key issue of this challenge is reasoning about names that are ``in the process'' of being scope-extruded, which often presents difficulties for the mechanisation of binders.

### 2023-10-17 paper/paper.tex
Edit lightly for space.

@@
 \paragraph{Outline of this paper.}
-In \cref{sec:challenge-problems} we describe our three challenges.
+In \cref{sec:challenge-problems} we describe our three challenges (proofs of the theorems are available in \cref{app:challenges}).
 We discuss our design choices in \cref{sec:design-discussion}, and possible extensions and further topics in \cref{sec:going-beyond}.
@@
 The operational semantics is defined as the following relation on processes:
-\begin{mathpar}
+\begin{mathparpagebreakable}
   \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{a}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{a}{l}}{R}})}}}
@@
   \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathpar}
+\end{mathparpagebreakable}
 
@@
 Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable or a name.
+Since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable or a name.
 
@@
 The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
+The \emph{internal action} \( \Atau \) is performing internal communication.
 
@@
 The transition relation is then defined by the following rules:
-\begin{mathpar}
+\begin{mathparpagebreakable}
   \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
@@
   \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
-\end{mathpar}
+\end{mathparpagebreakable}
 Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
@@
 \label{sec:challenge:coinduction}
-This challenge is about the mechanisation of proofs concerning infinite behaviours.
+This challenge is about the mechanisation of proofs concerning processes with infinite behaviours.
 % formalises a proof
@@
 %
-Coinduction is a technique for defining and proving properties of
-such infinite structures,
-%
-which, in the context of the concurrent benchmark, arise
-% in this context
-due to systems with behaviors that continue indefinitely.
-%
-Coinduction is the dual of induction: whereas induction is useful for
-proving properties of least fixed points, coinduction is useful for
-proving properties of greatest fixed points.
+Coinduction is a technique for proving properties of
+such infinite structures.
+%
+%Coinduction is the dual of induction: whereas induction is useful for
+%proving properties of least fixed points, coinduction is useful for
+%proving properties of greatest fixed points.
 %
@@
 % values of which we will denote by \( x, y, \dots \).
-The syntax of the calculus is identical to that of
-Section~\ref{sec:challenge:linearity-beh-types} (challenge on
-linearity), except from restriction, now unary, and the addition of
-replication:
+The syntax of processes is given by:
 % for including replication and having unary restriction:
@@
 \end{displaymath}
-Note that, unlike the process calculus from
-Section~\ref{sec:challenge:linearity-beh-types}, communications occur
-over the same channel name. Moreover, since replication $!P$ allows
-for infinite copies of the process $P$, processes can dynamically
-create an infinite number of names during execution when combined with
-restriction $\Pres{x}{}{P}$.
+The output process \( \Pout{x}{v}{P} \) sends the value \( v \) on channel \( x \) and continues as \( P \).
+The intention is that \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
+The input process \( \Pin{x}{l}{P} \) waits for a base value from channel \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
+Since replication allows for infinite copies of the process $P$, processes can dynamically
+create an infinite number of names during execution.
+
 %
@@
 \subsubsection{Semantics.} We choose to give an LTS semantics for this
-challenge. As an alternative, it is also possible to adopt a reduction
-semantics which features fewer rules but needs a structural
-congruence. %  The semantics of the system describes the actions that the
+challenge. %As an alternative, it is also possible to adopt a reduction
+%semantics which features fewer rules but needs a structural
+%congruence. %  The semantics of the system describes the actions that the
 % system can perform by defining a labelled transition relation on
@@
 receiving the base value \( y \) via \( x \).  The \emph{internal
-  action} \( \Atau \) is performing some unobservable action, e.g.\
-internal communication.
+  action} \( \Atau \) is performing internal communication.
 %
@@
 The transition relation is defined by the following rules:
-\begin{mathpar}
+\begin{mathparpagebreakable}
   \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
@@
   \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
-\end{mathpar}
+\end{mathparpagebreakable}
 Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
@@
 congruence.
-To prove the theorem, we will use an \emph{up-to technique}.
+To prove the theorem, we will use an \emph{up-to technique}, utilizing the following definition and lemma.
 A relation \( \mathcal{S} \) is called a \emph{strong barbed bisimulation up to \( \sbbisim{}{} \)} if, whenever \( (P,Q) \in \mathcal{S} \), the following conditions hold:
@@
 \end{enumerate}
-These relations are useful owing to the following lemma:
 \begin{lemma}\label{lemma:up-to}
@@
 
-We are now ready to prove the theorem:
 \begin{theorem}
@@
 \subsubsection{General design remarks.}
-Like the previous POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
+Like the POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
 applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.
-Still, these (and other) proof mechanisation applications need the basic techniques that our challenge problems exercise:
+Still, these (and other) applications need the basic techniques that our challenge problems exercise:
 as per our design goal \ref{item:goal-reusability},
@@
 (such as choices) that would complicate the mechanisation without bringing tangible
-benefits and insights.\footnote{It is of course possible to reintroduce the omitted
-constructs, but we leave this as a possibility for a future benchmark challenge.}
+benefits and insights.
 The minimality and uniformity of the calculi also allows us to target the design
-goal \ref{item:goal-tutorials}, i.e.~encouraging the creation of guidelines and
-tutorials for the mechanisation of concurrency theory.
+goal \ref{item:goal-tutorials}.
 
@@
 We have chosen a simple notion of well-formedness for our challenge,
-and do not for instance consider deadlocked processes. While more
-sophisticated notions of well-formedness are also interesting, the
-proofs of these properties are more complicated and would thus make
+and do not for instance consider deadlocks. While more
+sophisticated notions of well-formedness are interesting, the
+proofs of these properties are more complicated and would make
 the challenge more difficult and less focused on the linear
@@
 entail combining the features of the challenges.
-In \cref{tab:extensions} we consider the following three extensions:
+In \cref{table:extensions} we consider the following three extensions:
 \emph{(1)} the addition of choice to the type level and the process level,
@@
 
+\clearpage
 \appendix

### 2023-10-17 paper/paper.tex
Fix broken reference.
furhter comment about the book and paper used as sources

@@
 First, we list some common notions and conventions that we use in the
-challenges. Since the calculi under study are somewhat different, each section
-lists the changes that apply. The challenges are inspired by simplified versions of calculi (and proofs) from~\cite{picalcbook} and~\cite{Vasconcelos2012}.
+challenges. Since the calculi under study are somewhat different, each
+section lists the changes that apply. The challenges are inspired by
+simplified versions of calculi (and proofs) from chapters 1 and 2
+of~\cite{picalcbook} and~\cite{Vasconcelos2012} where the idea is to
+use sources that can be considered standard and close to current practice.
 
@@
 entail combining the features of the challenges.
-In \cref{table:extensions} we consider the following three extensions:
+In \cref{tab:extensions} we consider the following three extensions:
 \emph{(1)} the addition of choice to the type level and the process level,

### 2023-10-17 paper/paper.tex
reworded the HOAS and free variables comment

@@
 It is not clear that the techniques used to solve the original challenges about binders can be easily extended to reasoning about scope extrusion.
-Notably, techniques like Higher Order Abstract Syntax, or well-scoped
-De\ Bruijn indices do not have the concept of free variables. It is
-not immediately obvious how to encode the operational semantics in
-this kind of system.
+For example, techniques like Higher Order Abstract Syntax, or well-scoped
+De\ Bruijn indices do not have the concept of free variables and thus encoding this challenge with them requires some attention.
 

### 2023-10-17 paper/paper.tex
merge

--- 2023-10-17 paper/paper.tex
@@
 \fxsetup{theme=color}
+% \setlength{\textwidth}{12.4cm}
+\setlength{\textheight}{19.9cm}
 
@@
 \paragraph{Outline of this paper.}
-In \cref{sec:challenge-problems} we describe our three challenges (proofs of the theorems are available in \cref{app:challenges}).
+In \cref{sec:challenge-problems} we describe our three challenges.
 We discuss our design choices in \cref{sec:design-discussion}, and possible extensions and further topics in \cref{sec:going-beyond}.
@@
 The operational semantics is defined as the following relation on processes:
-\begin{mathparpagebreakable}
+\begin{mathpar}
   \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{a}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{a}{l}}{R}})}}}
@@
   \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathparpagebreakable}
+\end{mathpar}
 
@@
 Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
-Since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable or a name.
+On the other hand, since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable or a name.
 
@@
 Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
-This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
+This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication
 The key issue of this challenge is reasoning about names that are ``in the process'' of being scope-extruded, which often presents difficulties for the mechanisation of binders.
@@
 The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing internal communication.
+The \emph{internal action} \( \Atau \) is performing some unobservable action, e.g.\ internal communication.
 
@@
 The transition relation is then defined by the following rules:
-\begin{mathparpagebreakable}
+\begin{mathpar}
   \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
@@
   \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
-\end{mathparpagebreakable}
+\end{mathpar}
 Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
@@
 \label{sec:challenge:coinduction}
-This challenge is about the mechanisation of proofs concerning processes with infinite behaviours.
+This challenge is about the mechanisation of proofs concerning infinite behaviours.
 % formalises a proof
@@
 %
-Coinduction is a technique for proving properties of
-such infinite structures.
-%
-%Coinduction is the dual of induction: whereas induction is useful for
-%proving properties of least fixed points, coinduction is useful for
-%proving properties of greatest fixed points.
+Coinduction is a technique for defining and proving properties of
+such infinite structures,
+%
+which, in the context of the concurrent benchmark, arise
+% in this context
+due to systems with behaviors that continue indefinitely.
+%
+Coinduction is the dual of induction: whereas induction is useful for
+proving properties of least fixed points, coinduction is useful for
+proving properties of greatest fixed points.
 %
@@
 % values of which we will denote by \( x, y, \dots \).
-The syntax of processes is given by:
+The syntax of the calculus is identical to that of
+Section~\ref{sec:challenge:linearity-beh-types} (challenge on
+linearity), except from restriction, now unary, and the addition of
+replication:
 % for including replication and having unary restriction:
@@
 \end{displaymath}
-The output process \( \Pout{x}{v}{P} \) sends the value \( v \) on channel \( x \) and continues as \( P \).
-The intention is that \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) waits for a base value from channel \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
-Since replication allows for infinite copies of the process $P$, processes can dynamically
-create an infinite number of names during execution.
-
+Note that, unlike the process calculus from
+Section~\ref{sec:challenge:linearity-beh-types}, communications occur
+over the same channel name. Moreover, since replication $!P$ allows
+for infinite copies of the process $P$, processes can dynamically
+create an infinite number of names during execution when combined with
+restriction $\Pres{x}{}{P}$.
 %
@@
 \subsubsection{Semantics.} We choose to give an LTS semantics for this
-challenge. %As an alternative, it is also possible to adopt a reduction
-%semantics which features fewer rules but needs a structural
-%congruence. %  The semantics of the system describes the actions that the
+challenge. As an alternative, it is also possible to adopt a reduction
+semantics which features fewer rules but needs a structural
+congruence. %  The semantics of the system describes the actions that the
 % system can perform by defining a labelled transition relation on
@@
 receiving the base value \( y \) via \( x \).  The \emph{internal
-  action} \( \Atau \) is performing internal communication.
+  action} \( \Atau \) is performing some unobservable action, e.g.\
+internal communication.
 %
@@
 The transition relation is defined by the following rules:
-\begin{mathparpagebreakable}
+\begin{mathpar}
   \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
@@
   \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
-\end{mathparpagebreakable}
+\end{mathpar}
 Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
@@
 congruence.
-To prove the theorem, we will use an \emph{up-to technique}, utilizing the following definition and lemma.
+To prove the theorem, we will use an \emph{up-to technique}.
 A relation \( \mathcal{S} \) is called a \emph{strong barbed bisimulation up to \( \sbbisim{}{} \)} if, whenever \( (P,Q) \in \mathcal{S} \), the following conditions hold:
@@
 \end{enumerate}
+These relations are useful owing to the following lemma:
 \begin{lemma}\label{lemma:up-to}
@@
 
+We are now ready to prove the theorem:
 \begin{theorem}
@@
 \subsubsection{General design remarks.}
-Like the POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
+Like the previous POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
 applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.
-Still, these (and other) applications need the basic techniques that our challenge problems exercise:
+Still, these (and other) proof mechanisation applications need the basic techniques that our challenge problems exercise:
 as per our design goal \ref{item:goal-reusability},
@@
 (such as choices) that would complicate the mechanisation without bringing tangible
-benefits and insights.
+benefits and insights.\footnote{It is of course possible to reintroduce the omitted
+constructs, but we leave this as a possibility for a future benchmark challenge.}
 The minimality and uniformity of the calculi also allows us to target the design
-goal \ref{item:goal-tutorials}.
+goal \ref{item:goal-tutorials}, i.e.~encouraging the creation of guidelines and
+tutorials for the mechanisation of concurrency theory.
 
@@
 We have chosen a simple notion of well-formedness for our challenge,
-and do not for instance consider deadlocks. While more
-sophisticated notions of well-formedness are interesting, the
-proofs of these properties are more complicated and would make
+and do not for instance consider deadlocked processes. While more
+sophisticated notions of well-formedness are also interesting, the
+proofs of these properties are more complicated and would thus make
 the challenge more difficult and less focused on the linear
@@
 
-\clearpage
 \appendix
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2023-10-17 challenges/coinduction.tex
Revise proof of Lemma 19

@@
 \begin{proof}
-  By definition, \( \sbcong{}{} \) is a congruence and is included in
-  \( \sbbisim{}{} \). Also, if it were strictly included in another
-  relation then we would get a contradiction since it must be closed
-  under all contexts.
+  We first prove that \( \sbcong{}{} \) is indeed a congruence, i.e.~it is an
+  equivalence relation that is preserved by all contexts.  Proving that \(
+  \sbcong{}{} \) is an equivalence is easy; to prove that \( \sbcong{}{} \)  is
+  preserved by all contexts, we show that $\forall C: \sbcong{P}{Q}$ implies
+  $\sbcong{\applyctx{C}{P}}{\applyctx{C}{Q}}$, by structural induction on the
+  context $C$.
+
+  To prove that \(\sbcong{}{}\) is the largest congruence included in
+  \(\sbbisim{}{}\), we show that for any congruence $\mathcal{S} \subseteq
+  {\sbbisim{}{}}$ we have $\mathcal{S} \subseteq {\sbcong{}{}}$. Take any $P,Q$
+  such that $P \mathrel{\mathcal{S}} Q$ (hence, $\sbbisim{P}{Q}$): since $S$ is
+  a congruence by hypothesis, this implies $\forall C: \applyctx{C}{P}
+  \mathrel{\mathcal{S}} \applyctx{C}{Q}$ (hence,
+  $\sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}}$). %
+  Therefore, by the definition of \(\sbcong{}{}\), we have $\sbcong{P}{Q}$,
+  from which we conclude $\mathcal{S} \subseteq {\sbcong{}{}}$.
 \end{proof}

### 2023-10-17 challenges/coinduction.tex
Backport editing into challenges/.

@@
 
-This challenge is  about the mechanisation of proofs concerning infinite behaviours.
+This challenge is  about the mechanisation of proofs concerning processes wih infinite behaviours.
 % formalises a proof
@@
 Coinduction is a technique for defining and proving properties of
-such infinite structures,
-%
-which, in the context of the concurrent benchmark, arise
-% in this context
-due to systems with behaviors that continue indefinitely.
-%
-Coinduction is the dual of induction: whereas induction is useful for
-proving properties of least fixed points, coinduction is useful for
-proving properties of greatest fixed points.
+such infinite structures.
 %
@@
 Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable.
+Since the rule \TirName{Out} only applies to base values, there is no way to send a variable.
 
@@
 \end{enumerate}
-These relations are useful owing to the following lemma:
+
 \begin{lemma}\label{lemma:up-to}

### 2023-10-17 challenges/linearity.tex
Move \sconga command into macros.tex.
Backport editing into challenges/.

@@
 
-\newcommand{\sconga}[2]{#1 \stackrel{a}{\equiv} #2}
 To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms.

### 2023-10-17 paper/paper.tex
Fix text introducing syntax in coinduction challenge.
Change appendix to only contain proofs.

--- 2023-10-17 paper/paper.tex
@@
 \begin{theorem}[Subject reduction]
+  \label{thm:subject-reduction}
   If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
@@
 \begin{theorem}[Type safety]
+  \label{thm:type-safety}
   If \( \types{\Cempty}{P} \), then \( P \) is well formed.
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:harmony-2}
   \( \reduces{P}{Q} \) implies the existence of a \( Q' \) such that \( \transition{P}{\Atau}{Q'} \) and \( \scong{Q}{Q'} \).
@@
 % values of which we will denote by \( x, y, \dots \).
-The syntax of processes is given by:
+The syntax of values and processes is given by:
 % for including replication and having unary restriction:
@@
 \begin{gather*}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\%\label{eq:bisim1} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'} %\label{eq:bisim2}
+  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\\label{eq:bisim1} \\
+  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'} \label{eq:bisim2}
 \end{gather*}
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:bisim-equiv}
   \( \sbbisim{}{} \) is an equivalence relation. %, that is, the relation is reflexive, symmetric, and transitive.
@@
 
-\begin{lemma}
+\begin{lemma}\label{lemma:sbcong-largest}
   \( \sbcong{}{} \) is the largest congruence included in
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:coinduction}
   \( \sbcong{P}{Q} \) if for any process \( R \) and
@@
 \appendix
-\section{Challenges}\label{app:challenges}
-\input{../challenges/challenges.tex}
+\section{Proofs}\label{app:challenges}
+\input{../proof-appendix/proofs.tex}
 

### 2023-10-17 proof-appendix/linearity.tex
Change appendix to only contain proofs.

--- 2023-10-17 challenges/linearity.tex
@@
-\subsection{Challenge: Linearity and behavioral type systems}
-\label{sec:challenge:linearity-beh-types}
-This challenge formalises a proof that requires reasoning about linearity of channels.
-Linearity is the notion that a channel must be used exactly once in a process
-This is necessary to prove properties about session type systems, and the key
-issue of this challenge is reasoning about the linearity of context splitting operations.
-Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
-
-The setting for this challenge is a small calculus with a session type
-system, the syntax and semantics of which are given below. The
-calculus is a fragment of the one presented in~\cite{Vasconcelos2012},
-formulated in the dual style of~\cite{barber96tr}.
-
-The main objective of this challenge is to prove type preservation (also
-known as subject reduction), i.e.\ that well-typed
-processes can only transition to processes which are also well-typed
-in the same context.
-The second objective is to prove type safety, i.e.\ that well-typed processes are also well-formed in the sense that they do not use endpoints in a non-dual way.
-
-\subsubsection{Syntax.}
-The syntax is given by the grammar
-\[
-\begin{array}{r@{\qquad}c@{\qquad}l}
-  v,w & \Coloneqq & a \quad\mid\quad l \\
-   P,Q & \Coloneqq & \Pend \quad\mid\quad \Pout{x}{v}{P} \quad\mid\quad \Pin{x}{l}{P} \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad  \Presd{x}{y}{P}
-\end{array}
-\]
-where a \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable \( l \).
-
-The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \) and then continues as \( P \).
-The intention is that the value \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) waits for a base value from \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
-%
-The process \( \Presd{x}{y}{P} \) represents a \emph{session} with endpoints named \( x \) and \( y \) which are bound in \( P \).
-In \( P \), the names \( x \) and \( y \) can be used to exchange messages over the session (sending on \( x \) and receiving on \( y \) or vice versa).
-%
-%Finally, the process $\PBpar{P}{Q}$ represents the parallel composition of $P$ and $Q$.
-%
-Note that in this calculus channels cannot be sent in messages, therefore the topology of the communication network described by a process cannot change.
-Also, there is no recursion or replication in the syntax, hence no infinite behaviors can be expressed. In particular, we only model linear (as opposed to shared) channels.
-
-\subsubsection{Semantics.}
-We describe the actions that the system can perform through a small step operational semantics.
-As usual, we use  a \emph{structural congruence} relation that equates processes that we deem to be indistinguishable.
-Structural congruence is the smallest congruence relation that satisfies the following axioms:
-\begin{mathpar}
-  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
-  \and
-  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
-  \and
-  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
-  \and
-  \inferrule[Sc-Res-Par]{\{x,y\} \cap \freenames{Q} = \emptyset}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
-  \and
-  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
-  \and
-  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
-\end{mathpar}
-
-The operational semantics is defined as the following relation on processes:
-\begin{mathpar}
-  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{a}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{a}{l}}{R}})}}}
-  \and
-  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
-  \and
-  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
-  \and
-  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathpar}
-
-Note that reductions are allowed only for restricted pairs of session endpoints.
-This makes it possible to formulate subject reduction so that the typing context is exactly the same before and after each reduction.
-%
-Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
-Since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable or a name.
-
-\subsubsection{Session types.}
-Our process syntax allows us to write processes that are ill formed in
-the sense that they either use the endpoints bound by a restriction to
-communicate in a way that does not follow the intended duality, or
-attempt to send something which is not a base value.  As an example,
-the process
-\( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \)
-attempts to send a base value on both \( x \) and \( y\), whereas one
-of the names should be used for receiving in order to guarantee
-progress.  Another example is the process
-\( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \),
-which attempts to send a variable that is not instantiated at the time
-of sending.
-%
-To prevent these issues, we introduce a \emph{session type system} which
-rules out ill-formed processes.
-
-\subsubsection{Syntax.}
-Our type system does not type processes directly, but instead focuses on the channels used in the process.
-The syntax of \emph{session types} \( S, T \), unrestricted type contexts \( \Gamma \) and linear type contexts \( \Delta \) is as follows:
-\[
-  \begin{array}{r@{\qquad}c@{\qquad}l}
-  S,T & \Coloneqq & \Tend \quad\mid\quad \Tbase \quad\mid\quad \Tin{S} \quad\mid\quad \Tout{S} \\
-    \Gamma & \Coloneqq & \Cempty \quad\mid\quad \Gamma, l \\
-                    \Delta &::= & \Cempty \quad\mid\quad \Cadd{\Delta}{\hastype{x}{S}}
-  \end{array}
-\]
-The \emph{end type} \( \Tend \) describes an endpoint on which no further interaction is possible.
-The \emph{base type} \( \Tbase \) describes base values.
-The \emph{input type} \( \Tin{S} \) describes endpoints used for receiving a value and then according to \( S \).
-The \emph{output type} \( \Tout{S} \) describes endpoints used for sending a value and then according to \( S \).
-
-Typing contexts gather type information about names and variables.
-\emph{Unrestricted} contexts are simply sets of names since we only have one
-base type. \emph{Linear contexts} associate a type to endpoints. We use
-the comma as split/union, overloaded to singletons, and \( \Cempty \) as the
-empty context. We extend the Barendregt convention to contexts, so that all
-entries are distinct.  Note that the order in which information is added to a
-type context does not matter.
-
-Since we need to determine whether endpoints are used in complementary ways to determine whether processes are well formed, we need to formally define the dual of a type as follows:
-\begin{mathpar}
-  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tend} = \Tend}
-\end{mathpar}
-Note that the dual function is partial since it is undefined for the base type.
-
-\subsubsection{Typing rules.}
-Our type system is aimed at maintaining two invariants:
-\begin{enumerate}
-\item No endpoint is used simultaneously by parallel processes;
-\item The two endpoints of the same session have dual types.
-\end{enumerate}
-The first invariant is maintained by linearly splitting type contexts when typing compositions of processes, the second by requiring duality when typing restrictions.
-
-We have two typing judgments: one for values, and one for processes.
-The typing rules for values are:
-\begin{mathpar}
-  \inferrule[T-Base]{ }{\typev{\Gamma}{\hastype{a}{\Tbase}}} \and
-  \inferrule[T-Var]{ }{\typev{\Cadd{\Gamma}l }{\hastype{l}{\Tbase}}}
-\end{mathpar}
-% In the typing rule for names, $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
-% \begin{mathpar}
-%   \inferrule[T-Name]{{\tend\Delta }}{\types{{\Gamma; (\Delta,\hastype{x}{T})}}{\hastype{x}{T}}}
-% \end{mathpar}
-The typing rules for processes are as follows:
-\begin{mathpar}
-
-  \inferrule[T-Inact]{\tend\Delta }{\types{\Gamma;\Delta}{\Pend}}
-  \and
-  \inferrule[T-Par]{\types{\Gamma;\Delta_1}{P} \\ \types{\Gamma;\Delta_2}{Q}}
-  {\types{\Gamma; \Csplit{\Delta_1}{\Delta_2}}{\Ppar{P}{Q}}}
-  \and
-  \inferrule[T-Res]{\types{\Gamma; (\Cadd{\Cadd{\Delta}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P})}{\types{\Gamma;\Delta}{\Presd{x}{y}{P}}}
-  \and
-    \inferrule[T-Out]{
-      \typev{\Gamma}{\hastype{v}{\Tbase}} \\ \types{\Gamma; \Cupdate{\Delta}{\hastype{x}{T}}}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tout{T}}})}{\Pout{x}{v}{P}}}
-    \and
-    \inferrule[T-IN]{
-      \types{(\Gamma,  l ); (\Cupdate{\Delta}{\hastype{x}{T}})}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tin{T}}})}{\Pin{x}{l}{P}}}
-\end{mathpar}
-Note that we do not need a judgment for typing channels, since it is already folded into the \TirName{T-In} and \TirName{T-Out} rules.
-
-\subsubsection{Challenge.}
-The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with some lemmas:
-
-\begin{lemma}[Weakening]\mbox{}
-  \label{lemma:weak}
-  \begin{enumerate}
-  \item If \( \types{\Gamma; \Delta}{P} \) then
-    \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
-  \item If \( \types{\Gamma; \Delta}{P} \) % and \( \tend{\Delta} \)
-    then
-    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
-  \end{enumerate}
-\end{lemma}
-\begin{proof}
+\subsection{Proofs: Linearity and behavioral type systems}
+\begin{proof}[of \cref{lemma:weak}]
   By induction on the given derivations.
 \end{proof}
-\begin{lemma}[Strengthening]\mbox{}
-  \label{lemma:strenD}
-  If \( \types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P} \) and
-  $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
-\end{lemma}
-\begin{proof}
+
+\begin{proof}[of \cref{lemma:strenD}]
   By induction on the derivation of
@@
 
-\begin{lemma}[Substitution]\mbox{}
-  \label{le:subst}
-  If $\types{(\Cadd{\Gamma}{l});\Delta}{P}$ and
-  ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
-  \( \types{\Gamma;\Delta}{\subst{P}{a}{l}} \).
-\end{lemma}
-\begin{proof}
+\begin{proof}[of \cref{le:subst}]
   By induction on the derivation of $\types{(\Cadd{\Gamma}{l});\Delta}{P}$.
@@
   \item[\TirName{T-Out}] Let $P = \Pout{x}{v}{P'}$. We have that
-    $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.  We must show
+    $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$. \\We must show
     $\typev{\Gamma}{\hastype{v}{\Tbase}}$ to build the conclusion with
     the induction hypothesis and \TirName{T-Out}. We proceed by cases
-    on the structure of $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.
+    on the structure of\\$\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.
     % \item[$P = \Pin{x}{v}{P}$] Immediate after inverting typing.
@@
   The remaining cases are analogous.
-
 \end{proof}
 
-To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms.
-
-\begin{lemma}[Preservation for $\stackrel{a}{\equiv}$]
-  \label{le:presequiva}
-  If \( \sconga{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
-\end{lemma}
-\begin{proof}
+\begin{proof}[of \cref{le:presequiva}]
   By case analysis on the \TirName{Sc} rule applied:
@@
   \item[\TirName {Par-Comm/Assoc}] By rearranging sub-derivations noting that
-    order does not matter for linear contexts
+    order does not matter for linear contexts.
   \item[\TirName{Par-Inact}]
-   Right-to-left by strengthening~\ref{lemma:strenD}. Vice-versa,
-    by picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
-  \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear or $\Tend$ and applying weakening and strengthening accordingly.
-  \item[\TirName{Res-Inact:}] Use weakening %~\ref{lemma:weak} part 2
-  \item[\TirName{Res}] Noting that order does not matter
+   Right-to-left by strengthening (\cref{lemma:strenD}). Vice versa,
+    by picking $T$ to be $\Tend$ and weakening (\cref{lemma:weak}(2)).
+  \item[\TirName{Res-Par}] By case analysis on $\hastype{x}{T}$ being linear or $\Tend$ and applying weakening (\cref{lemma:weak}) and strengthening (\cref{lemma:strenD}) accordingly.
+  \item[\TirName{Res-Inact:}] By weakening (\cref{lemma:weak}). %~\ref{lemma:weak} part 2
+  \item[\TirName{Res}] By rearranging sub-derivations noting that
+    order does not matter for linear contexts.
   \end{description}
@@
 
-Now, following Sangiorgi \& Walker~\cite{picalcbook}, we give rules for the compatible
-equivalence relation induced by $\sconga\cdot\cdot$, which we still
-write as $\scong\cdot\cdot$: namely, add to reflexivity, symmetry and
-transitivity the following condition:
- \begin{mathpar}
-   \inferrule[Cong]{\sconga P Q}{\scong {C[P]}{C[Q]}}
- \end{mathpar}
-
-\begin{lemma}[Preservation for $\equiv$]
-  \label{le:presequiv}
-    If \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
-\end{lemma}
-\begin{proof}
+\begin{proof}[of \cref{le:presequiv}]
   By induction on the structure of the derivation of \( \scong{P}{Q} \), with an inner induction of the structure of a process context.
   \begin{description}
-  \item[\TirName {Refl}] Immediate
-      \item[\TirName {Sym}] By IH.
-      \item[\TirName {Trans}]  By two appeal to the IH
+  \item[\TirName {Refl}] Immediate.
+      \item[\TirName {Sym}] By the induction hypothesis.
+      \item[\TirName {Trans}]  By two applications of the induction hypothesis.
       \item[\TirName {Cong}] By induction on the structure of $C$. If
-        the context is a hole, apply Lemma~\ref{le:presequiva}. In the
-        step case, apply the IH: for example if $C$ has the form
+        the context is a hole, apply \cref{le:presequiva}. In the
+        step case, apply the induction hypothesis: for example if $C$ has the form
         ${\Ppar {C'} R}$, noting that ${\PBpar{ C'} R}[P]$ is equal to
@@
         \( \types{\Gamma;\Delta}{{\PBpar {C'} R}[P]} \) iff
-        \( \types{\Gamma;\Delta}{{\PBpar {C'} R}[Q]} \) by rule T-Par and
-        the IH.
+        \( \types{\Gamma;\Delta}{{\PBpar {C'} R}[Q]} \) by rule \TirName{T-Par} and
+        the induction hypothesis.
   \end{description}
@@
 
-\begin{theorem}[Subject reduction]
-  If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
-\end{theorem}
-\begin{proof}
-  By induction on the derivation of $\reduces{P}{Q}$. The cases \TirName{R-Par} and \TirName{R-Res} follow immediately by IH. Case \TirName{R-Struct} appeals twice to
+\begin{proof}[of \cref{thm:subject-reduction}]
+  By induction on the derivation of $\reduces{P}{Q}$. The cases \TirName{R-Par} and \TirName{R-Res} follow immediately by the induction hypothesis. Case \TirName{R-Struct} appeals twice to
   preservation of $\equiv$ (Lemma~\ref{le:presequiv}) and to the
-  IH. For \TirName{R-Com}, suppose that \TirName{T-Res} introduces in $\Delta$ the
+  induction hypothesis. For \TirName{R-Com}, suppose that \TirName{T-Res} introduces in $\Delta$ the
   assumptions $\Cadd{\hastype{x}{\Tout{U}}}{\hastype{y}{\Tin{\dual{U}}}}$.  Building the only derivation for the
@@
 
-To formulate safety, we need to formally define what we mean by
-well-formed process.  We say that a process \( P \) is \emph{prefixed
-  at variable \( x \)} if \( P\equiv\Pout{x}{v}{P'} \) or
-\( P\equiv\Pin{x}{l}{P'} \) for some $P'$.
-%
-A process $P$ is then \emph{well formed} if, for every $P_1$, $P_2$,
-and $R$ such that
-\( P\equiv \Presd{x_1}{y_1}{\dots
-  \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}} \), with
-\( n \geq 0 \), it holds that, if \( P_1 \) is prefixed at \( x_1 \)
-and \( P_2 \) is prefixed at \( y_1 \) (or vice versa), then
-\( \Ppar{P_1}{P_2}
-\equiv\Ppar{\Pout{x_1}{a}{P_1'}}{\Pin{y_1}{l}{P_2'}} \), for some
-$P_1'$ and $P_2'$.
-
-Note that well-formed processes do not necessarily reduce. For example, the process
-\begin{equation*}
-  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
-\end{equation*}
-is well formed but also irreducible.
-
-\begin{theorem}[Type safety]
-  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
-\end{theorem}
-\begin{proof}
+\begin{proof}[of \cref{thm:type-safety}]
   In order to prove that $P$ is well-formed, let us consider any
@@
     \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}}$ that is
-  structurally congruent to $P$. Clearly, by Lemma~\ref{le:presequiv},
+  structurally congruent to $P$. Clearly, by \cref{le:presequiv},
   well-typedness must be preserved by structural congruence. Moreover,
@@
 \end{proof}
-
-\begin{corollary}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well formed.
-\end{corollary}

### 2023-10-17 proof-appendix/proofs.tex
Change appendix to only contain proofs.

\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
                         {\endgroup]\end{quote}}
\long\def\ignore#1{}

This appendix contains paper proofs (or in some cases, proof sketches), of the lemmas and theorems mentioned in the challenges.
The proofs are all standard, and are in most cases simplified versions of similar proofs in~\cite{picalcbook} and~\cite{Vasconcelos2012}.
These standard references may also be consulted for introductory explanation and examples of the notions discussed in the challenges and for further examples of proofs employing the proof techniques we are interested in.

\input{../proof-appendix/linearity.tex}
\input{../proof-appendix/scope-extrusion.tex}
\input{../proof-appendix/coinduction.tex}

### 2023-10-17 proof-appendix/coinduction.tex

--- 2023-10-17 challenges/coinduction.tex
@@
-\subsection{Challenge: Coinduction and infinite processes}
-% and reasoning about infinite processes}
-\label{sec:challenge:coinduction}
+\subsection{Proofs: Coinduction and infinite processes}
 
-% \marco{Why reasoning is only about infinite processes? Would it be
-%   unreasonable to shorten the title to {\em Coinduction and infinite
-%     processes}?}
-
-This challenge is  about the mechanisation of proofs concerning processes wih infinite behaviours.
-% formalises a proof
-%that require coinduction.
-% coinductive techniques.
-%
-% Coinduction is a proof technique for infinite structures,
-This is usually connected to \emph{coinductive} definitions where an infinite
-structure is defined as the greatest fixed point of a recursive
-definition.
-%
-Coinduction is a technique for defining and proving properties of
-such infinite structures.
-%
-% \marco{Should we be more specific here and give a deeper explanation
-%   of what coinduction is? If not, then why do we even say that it's
-%   the dual of induction?}
-%
-
-For this challenge, we adopt a fragment of the untyped $\pi$-calculus
-that includes process replication.
-% The setting for this challenge is an untyped calculus of
-% communicating systems with replication of processes, the syntax and
-% semantics of which are presented below.
-The objective of this challenge is to % prove % a context theorem for
-draw a formal connection between strong barbed congruence and strong
-barbed bisimilarity.  % I.e., The context lemma
-The result establishes that two processes are strong barbed congruent
-if the processes obtained by applying a finite number of substitutions to
-them and composing them in parallel with an arbitrary process are
-strongly barbed bisimilar.
-The key issue of this challenge is the coinductive reasoning about the infinite behaviours of the replication operator.
-
-\subsubsection{Syntax.}
-% We assume the existence of some type of \emph{base values}, values
-% of which we will denote by \( a, b, \dots \), the existence of some
-% type of \emph{variables}, values of which we will denote by
-% \( l, m, \dots \), and the existence of some type of \emph{names},
-% values of which we will denote by \( x, y, \dots \).
-The syntax of processes is given by:
-% for including replication and having unary restriction:
-\begin{displaymath}
-  \begin{array}{r@{\qquad}c@{\qquad}l}
-    v,w & \Coloneqq & a \quad\mid\quad l \\
-    P,Q & \Coloneqq & \Pend
-               \quad \mid\quad \Pout{x}{v}{P}
-               \quad \mid\quad \Pin{x}{l}{P}
-               \quad \mid\quad \PBpar{P}{Q}
-               \quad \mid\quad \Pres{x}{}{P}
-               \quad \mid\quad !P
-  \end{array}
-\end{displaymath}
-The output process \( \Pout{x}{v}{P} \) sends the value \( v \) on channel \( x \) and continues as \( P \).
-The intention is that \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) waits for a base value from channel \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
-Since replication allows for infinite copies of the process $P$, processes can dynamically
-create an infinite number of names during execution.
-
-%
-% \begin{align*}
-%   v,w :=&&& a \\
-%   |&&& l \\
-%   P,Q :=&&& \Pend \\
-%   |&&& \Pout{x}{v}{P} \\
-%   |&&& \Pin{x}{l}{P} \\
-%   |&&& \Ppar{P}{Q} \\
-%   |&&& \Pres{x}{P} \\
-%   |&&& \Preplicate{P}
-% \end{align*}
-% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a
-% variable \( l \).  The process \( \Pend \) is \emph{inaction}:
-% a process which can do nothing.  The process \( \Pout{x}{a}{P} \) is
-% an \emph{output}, which can send the value \( v \) via \( x \), then
-% continue as \( P \).  The intention is that the value \( v \) must
-% be a base value when it is actually sent, and this will be enforced
-% in the semantics later on.  The process \( \Pin{x}{l}{P} \) is an
-% \emph{input}, which can receive a base value via \( x \), then
-% continue as \( P \) with the received value substituted for the
-% variable \( l \).  The input operator thus binds the variable
-% \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
-% \emph{composition} of process \( P \) and process \( Q \).  The two
-% components can proceed independently of each other, or they can
-% interact via shared names.  The process \( \Pres{x}{P} \) is the
-% \emph{restriction} of the name \( x \) to \( P \).  Components in
-% \( P \) can use the name \( x \) to interact with each other, but
-% not with processes outside of the restriction.  The restriction
-% operator thus binds the name \( x \) in \( P \).  Note that the
-% scope of a restriction may not change when processes interact, since
-% it is only possible to send and receive values, and not names.  The
-% process \( \Preplicate{P} \) is the \emph{replication} of the
-% process \( P \).  It can be thought of as the infinite composition
-% \( \Ppar{P}{\Ppar{P}{\cdots}} \).  Replication makes it possible to
-% express infinite behaviours.
-
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process
-% \( P \) with value \( v \) substituted for variable \( l \).
-
-% % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
-% written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from
-% \( P \) by a finite number of substitutions of bound variables.
-% As a convention, we will identify \( \alpha \)-convertible
-% processes.
-
-% Also as a convention, we assume that the bound names and bound
-% variables occurring in any collection of processes are chosen
-% to be different from the free names and free variables
-% occurring in those processes and from the names and variables
-% occurring in any substitutions applied to the processes.  This is
-% justified because any overlapping names and variables may be
-% \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics.} We choose to give an LTS semantics for this
-challenge. %As an alternative, it is also possible to adopt a reduction
-%semantics which features fewer rules but needs a structural
-%congruence. %  The semantics of the system describes the actions that the
-% system can perform by defining a labelled transition relation on
-% processes.
-
-The transitions are labelled by \emph{actions}, the syntax
-of which is as follows:
-% \begin{align*}
-%   \alpha := &&& \Aoutf{x}{a} \\
-%   |&&& \Ain{x}{a} \\
-%   |&&& \Atau
-% \end{align*}
-\begin{align*}
-  \alpha\ \Coloneqq\ \ \Aoutf{x}{a} \ \mid\ \Ain{x}{a} \ \mid\ \Atau
-\end{align*}
-The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value
-\( a \) via \( x \).  The \emph{input action} \( \Ain{x}{y} \) is
-receiving the base value \( y \) via \( x \).  The \emph{internal
-  action} \( \Atau \) is performing internal communication.
-%
-We use the notation \( \names{\alpha} \) to denote the set of
-names that occur in the action \( \alpha \).
-
-The transition relation is defined by the following rules:
-\begin{mathpar}
-  \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{l}{P}}{\Ain{x}{a}}{\subst{P}{a}{l}}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{a}}{P'} \\ \transition{Q}{\Ain{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{a}}{P'} \\ \transition{Q}{\Aoutf{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ x \notin \names{\alpha}}{\transition{\Pres{x}{P}}{\alpha}{\Pres{x}{P'}}}
-  \and
-  \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
-\end{mathpar}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
-Since the rule \TirName{Out} only applies to base values, there is no way to send a variable.
-
-% As a convention, we assume that bound names and bound variables of any processes or actions are chosen to be different from the names and variables that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variables.
-
-\subsubsection{Bisimilarity.}
-Bisimilarity is a notion of equivalence for processes and
-% .  Our notion of process equivalence relations
-builds on a notion of \emph{observables}, i.e., what we can externally
-observe from the semantics of a process. If we allowed ourselves only to observe
-internal transitions (i.e., observe that a process is internally
-performing a step of computation) we would relate either too few
-processes (in the strong case where we relate only processes with
-exactly the same number of internal transitions) or every process (in
-the weak case where we relate processes with any amount of internal
-transitions).  As a result, we must allow ourselves to observe more than
-just internal transitions, and we choose to describe a process's observables
-as the names it might use for sending and receiving.
-
-To this end, we define the \emph{observability predicate}
-\( \observable{P}{\mu} \) as follows:
-\begin{align*}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align*}
-
-\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu}\label{eq:bisim1} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}\label{eq:bisim2}
-\end{gather}
-We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
-
-\begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation. %that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-\begin{proof}
+\begin{proof}[of \cref{thm:bisim-equiv}]
   % We need to prove that $\sbbisim{}{}$ is reflexive, symmetric, and
@@
   \end{itemize}  
-
 \end{proof}
 
-Unfortunately, strong barbed bisimilarity is not a good process
-equivalence since it is not a congruence, hence it does not allow for
-substituting a process with an equivalent one in any context.
-% equivalent processes to be since it does not consider the
-% environment of processes.
-For instance, the processes $\Pout{x}{a}{\Pout{y}{b}{\Pend}}$ and
-$\Pout{x}{a}{\Pend}$ are strong barbed bisimilar, i.e.,
-$\sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}}$. This
-is because \( \obsout{x} \) is the only observable in both processes
-and they cannot perform a \( \Atau \)-action. However, in the context
-$C = \Ppar{\ctxhole}{\Pin{x}{l}{\Pend}}$, the relation no
-longer holds: in fact,
-$\nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}\Ppar{}{\Pin{x}{l}{\Pend}}}$
-%
-because the left process can perform a \( \Atau \)-action such that
-\( \obsout{y} \) becomes observable, whereas the right process cannot.
-
-% For instance,
-% \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \)
-% since \( \obsout{x} \) is the only observable in both processes and
-% they cannot perform a \( \Atau \)-action, but
-% \(
-% \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}}
-% \) since the left process can perform a \( \Atau \)-action such that
-% \( \obsout{y} \) becomes observable, whereas the right process cannot.
-
-% \subsubsection{Contexts and congruences}
-% Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
-
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We can think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variables that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsubsection{Strong barbed congruence.}
-In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,
-i.e., % we need to make strong barbed bisimilarity consider
-we have to consider the environment in which processes may be placed.
-% (i.e.\ the context) of processes.
-
-% We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying
-We say that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
-
-\begin{lemma}
-  \( \sbcong{}{} \) is the largest congruence included in
-  \( \sbbisim{}{} \).
-\end{lemma}
-\begin{proof}
+\begin{proof}[of \cref{lemma:sbcong-largest}]
   We first prove that \( \sbcong{}{} \) is indeed a congruence, i.e.~it is an
@@
 
-\subsubsection{Challenge.}
-The objective of this challenge is to prove a theorem that shows that
-making strong barbed bisimilarity sensitive to substitution and
-parallel composition is enough to show strong barbed
-congruence.
-To prove the theorem, we will use an \emph{up-to technique}, utilizing the following definition and lemma.
-A relation \( \mathcal{S} \) is called a \emph{strong barbed bisimulation up to \( \sbbisim{}{} \)} if, whenever \( (P,Q) \in \mathcal{S} \), the following conditions hold:
-\begin{enumerate}
-\item \( \observable{P}{\mu} \) if and only if \( \observable{Q}{\mu} \).
-\item if \( \transition{P}{\Atau}{P'} \) then \( \transition{Q}{\Atau}{Q'} \) for some \( Q' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
-\item if \( \transition{Q}{\Atau}{Q'} \) then \( \transition{P}{\Atau}{P'} \) for some \( P' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
-\end{enumerate}
-
-\begin{lemma}\label{lemma:up-to}
-  If \( \mathcal{S} \) is a strong barbed bisimulation up to \( \sbbisim{}{} \), \( (P,Q) \in \mathcal{S} \) implies \( \sbbisim{P}{Q} \).
-\end{lemma}
-\begin{proof}
+\begin{proof}[of \cref{lemma:up-to}, sketch]
   We check that \( \uptosbb{}{\mathcal{S}}{} \) is a strong barbed bisimulation and is thus included in \( \sbbisim{}{} \).
@@
 
-\begin{theorem}
-  \( \sbcong{P}{Q} \) if, for any process \( R \) and
-  substitution \( \sigma \),
-  \(
-  \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}}
-  \).
-\end{theorem}
-\begin{proof}
+\begin{proof}[of \cref{thm:coinduction}]
   Since \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \), it suffices to show that if \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \) for any \( R \) and \( \sigma \), then \( \sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}} \) for any \( C \), \( R \) and \( \sigma \).

### 2023-10-17 proof-appendix/scope-extrusion.tex

\subsection{Proofs: Name passing and scope extrusion}
\begin{proof}[of \cref{se-lemma-harmony-fact}, sketch]
  First, show the special case when \( P \) can be rewritten to \( Q \) with a single application of an axiom of the structural congruence to some subterm of P.
  The proof is then by induction on the number of such steps.
\end{proof}

\begin{proof}[of \cref{se-lemma-normalized-derivation}, sketch]
  To obtain a normalized derivation from an arbitrary derivation we will need to check that rules \TirName{R-Com}, \TirName{R-Par} and \TirName{R-Res} commute with \TirName{R-Struct}, and that two applications of \TirName{R-Struct} can be combined into one.
\end{proof}

\begin{proof}[of \cref{se-lemma-normalized-reduction}]
  Follows immediately from \cref{se-lemma-normalized-derivation} and the shape of a normalized derivation.
\end{proof}

\begin{proof}[of \cref{thm:se-trans-implies-red}, sketch]
  The proof is by induction on the inference of\\\( \transition{P}{\Atau}{Q} \) using the following lemmas:
  \begin{enumerate}
  \item if \( \transition{Q}{\Aoutf{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\Pout{x}{y}{R}}{S})}}} \)\\and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x,y \notin \{ w_1, \dots, w_n \} \).
  \item if \( \transition{Q}{\Aoutb{x}{z}}{Q'} \) then \( \scong{Q}{\Pres{z}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\Pout{x}{z}{R}}{S})}}}} \)\\and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{R}{S})}}} \) where \( x \notin \{ z, w_1, \dots, w_n \} \).
  \item if \( \transition{Q}{\Ain{x}{y}}{Q'} \) then \( \scong{Q}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\Pin{x}{z}{R}}{S})}}} \)\\and \( \scong{Q'}{\Pres{w_1}{\!\dots \Pres{w_n}{(\Ppar{\subst{R}{y}{z}}{S})}}} \) where \( x \notin \{ w_1, \dots, w_n \} \).
  \end{enumerate}
\end{proof}

\begin{proof}[of \cref{thm:harmony-2}]
  If \( \reduces{P}{Q} \), then by \cref{se-lemma-normalized-reduction}, \( \scong{P}{P'} \) with
  \begin{equation*}
    P' = \Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{\Pout{x}{y}{R_1}}{\Pin{x}{z}{R_2}})}{S})}}
  \end{equation*}
  and \( \scong{Q}{Q'} \) with
  \begin{equation*}
    Q' = \Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{R_1}{\subst{R_2}{y}{z}})}{S})}} \ .
  \end{equation*}
  We can easily check that \( \transition{P'}{\tau}{Q'} \) and so by \cref{se-lemma-harmony-fact}, \( \transition{P}{\tau}{Q'} \).
\end{proof}

### 2023-10-17 paper/paper.tex
Fix a number of small things.

@@
 
-Now, following Sangiorgi \& Walker~\cite{picalcbook}, we give rules for the compatible
+Now we give rules for the compatible
 equivalence relation induced by $\sconga\cdot\cdot$, which we still
@@
 \emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
-\begin{gather*}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\\label{eq:bisim1} \\
+\begin{gather}
+  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \label{eq:bisim1}\\
   \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'} \label{eq:bisim2}
-\end{gather*}
+\end{gather}
 We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
@@
 
-As with the POPLMark and POPLMark Reloaded challegnes, this challenge does not aim to cover every possible
+As with the POPLMark and POPLMark Reloaded challenges, this challenge does not aim to cover every possible
 problem in the area. Our benchmark can be extended in two
@@
 that we do not cover here (e.g.: well-formed global types,
-projection as a partial function) --- although, dually, some aspects
+projection as a partial function) --- although some aspects
 that we cover here do appear in their theory. Also, one could easily
@@
 
-On the side of proof assistants, the current challenge proposes mechanising the
-proofs of theorems. An interesting venue to explore would be to take advantage
+On the mechanisation side, the present challenge proposes mechanising the
+proofs of theorems. An interesting avenue to explore would be to take advantage
 of other proof assistant features. For example, proof assistants can often

### 2023-10-18 paper/paper.tex
Revert "Anonymize paper."
This reverts commit ee50331f89b9b7f73136cb1a800788cfe8d62c0f.

@@
 \title{The Concurrent Calculi Formalisation Benchmark}
+
+\author{
+     Marco Carbone \inst{1}\orcidID{0000-0001-9479-2632}
+\and David Castro-Perez \inst{2}\orcidID{0000-0002-6939-4189}
+\and Francisco Ferreira \inst{3}\orcidID{0000-0001-8494-7696}
+\and Lorenzo Gheri \inst{4}\orcidID{0000-0002-3191-7722}
+\and Frederik Krogsdal Jacobsen \inst{5}\orcidID{0000-0003-3651-8314}
+\and Alberto Momigliano \inst{6}\orcidID{0000-0003-0942-4777}
+\and Luca Padovani \inst{7}\orcidID{0000-0001-9097-1297}
+\and Alceste Scalas \inst{5}\orcidID{0000-0002-1153-6164}
+\and Martin Vassor \inst{8}\orcidID{0000-0002-2057-0495}
+\and Nobuko Yoshida \inst{8}\orcidID{0000-0002-3925-8557}
+\and Daniel Zackon \inst{9}\orcidID{0009-0008-6153-2955}
+}
+
+\institute{
+     IT University of Copenhagen, Copenhagen, Denmark \email{maca@itu.dk}
+\and University of Kent, Canterbury, United Kingdom \email{D.Castro-Perez@kent.ac.uk}
+\and Royal Holloway, University of London, Egham, United Kingdom \email{Francisco.FerreiraRuiz@rhul.ac.uk}
+\and University of Liverpool, Liverpool, United Kingdom \email{Lorenzo.Gheri@liverpool.ac.uk}
+\and Technical University of Denmark, Kgs. Lyngby, Denmark \email{fkjac@dtu.dk}, \email{alcsc@dtu.dk}
+\and Università degli Studi di Milano, Milan, Italy \email{momigliano@di.unimi.it}
+\and Università di Camerino, Camerino, Italy \email{luca.padovani@unicam.it}
+\and University of Oxford, Oxford, United Kingdom \email{martin.vassor@cs.ox.ac.uk}, \email{nobuko.yoshida@cs.ox.ac.uk}
+\and McGill University, Montreal, Canada \email{daniel.zackon@mail.mcgill.ca}
+}
+
+\authorrunning{M. Carbone et al.}
 

### 2023-10-20 technical-report/main.tex
Move theorem definitions out of macros for Springer compatibility.
adding dual presentation of session types (with intuitionistic and linear context separated). Now this is called in technical-report but it can be enabled/disabled elsewhere (file appst.tex)
Split challenges.tex into multiple files.
removed appst appendix from tech report
Fix compilation error in technical report.
Backport Alceste's changes into challenges/.
update technical report

--- 2023-04-26 main.tex
@@
 \documentclass{article}
-
+\synctex=1
 \usepackage[utf8]{inputenc}
-\usepackage{graphicx}
-\usepackage{natbib}
+\usepackage[T1]{fontenc}
 \usepackage{amsmath}
@@
 \usepackage{amsthm}
+\usepackage{mathtools}
+\usepackage{graphicx}
 \usepackage{mathpartir}
+\usepackage{cite}
 \usepackage{bm}
@@
 \usepackage{cleveref}
+\usepackage{hyperref}
+\usepackage[inline]{enumitem}
+\usepackage{color}
+\renewcommand\UrlFont{\color{blue}\rmfamily}
 
-\input{macros.tex}
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}{Corollary}
+
+\usepackage[status=draft]{fixme}
+\fxsetup{theme=color}
+
+%\SetWatermarkText{Draft}
+%\SetWatermarkScale{5}
+%\SetWatermarkColor[gray]{0.90}
+
+\input{../macros.tex}
 
@@
 \title{POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types}
+\author{...}
 
@@
 
-In recent years, the programming language community has embraced
-machine checked proofs for theoretical results in papers. The
-POPLMark~\cite{Pierce:2005} challenge was proposed for this purpose,
-and to get the community on track. Since then the number of formal
-proofs accompanying papers has grown\footnote{add some POPL/PLDI/etc
-  statistics here}. However, formal proofs are not equally spread in
-the field. In this call to action, we propose to extend the key
-contribution of challenges like POPLMark is to help develop the
-necessary momentum and culture to be able to formalise results to the
-concurrent calculi community by proposing a challenge the exercises
-the key aspects of reasoning about process calculi. We propose this
-document, and a \href{https://concurrentbenchmark.github.io/}{website}
-to enable the community to collaborate on the project.
+% The abstract of the influential POPLMark challenge~\cite{POPLMark}
+% starts with an important question: \emph{``How close are we to a world
+%   where every paper on programming languages is accompanied by an
+%   electronic appendix with machine-checked proofs?''} The introduction
+% of the POPLMark challenge spearheaded a shift towards publications
+% that include mechanised proofs, fostering the progress of proof
+% assistants, libraries, and best practices. Later on, the follow-up
+% POPLMark Reloaded challenge~\cite{POPLMarkReloaded} has encouraged the
+% development of best practices and tool support for proofs using
+% logical relations.
 
-It has been a long time since the original challenge, and many changes
-have happened. Particularly, POPLMark and any challenge cannot address
-every conceivable technique. Challenges have to be simple so that the
-techniques are not shadowed by technical details and also feasible to
-implement without heroic efforts. In that vein, it is apparent that
-new challenges are needed, work like POPLMark
-Reloaded~\cite{Pientka:2019} strives to extend the scope of the
-original challenge to proofs using logical relations. This technique
-is crucial to many current developments.
+The authors of both POPLMark and POPLMark Reloaded explicitly
+note that their benchmarks were only a beginning.  Furthermore, they
+specifically mention reasoning about concurrency using coinduction and
+linear environments as points for future work.  In this spirit, in order to
+further increase the reach of the POPLMark challenges, we propose a new
+collection of benchmark problems specifically designed to address typical issues
+that arise when mechanising message-passing concurrency using process calculi.
 
-As mentioned before, in this work we address another expansion of
-scope for such challenges, going from the $\lambda$-calculus to
-process calculi such as the $\pi$-calculus and session type systems
-(and other behavioural type systems). These systems require particular
-insights to implement. So far, we have identified three key
-aspects\footnote{We want to keep the number low, but the exact nature
-  of the challenges is still open to discussion (as are most things in
-  this project).}. First, the idea of scope extrusion in name passing
-calculi. This generalises the idea of binders in ways that techniques
-that were adequate for $\lambda$-calculus binders are not so
-convenient for name passing systems. Second, session types in
-particular, and behavioural types in general often depend on linearity
-in their type systems. The formalisation of linear type systems posses
-a challenge given that it is not uniformly well supported among the
-many proof assistants. Finally, processes in concurrent systems often
-contain infinite behaviour that is typically modelled with
-coinduction. This technique is used to reason about (bi)similarity
-between processes, the equivalence of their traces, or subtyping
-between infinite trees of possible behaviours. This challenge
-concentrates on those issues.
+Several papers in the realm of process calculi already include mechanised proofs
+(e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
+  DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or
+discuss proof mechanisation techniques
+(e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
+  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).  However, our
+experience is that choosing the appropriate proof mechanisation techniques and
+tools remains a significant challenge with few guidelines.  This often leads
+researchers towards trial-and-error, attempting sub-optimal and ad-hoc
+solutions; this increases the overall effort required for proof mechanisation,
+and hampers the reuse of techniques and results.
+%
+Therefore, it is our opinion that the research community in message-passing
+concurrency (and more specifically, in process calculi) will benefit from a set
+of solutions exploring different ways of mechanising typical problems in the
+field.
+Like the authors of the POPLMark challenges, we seek to
+answer a number of questions:
+\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
+\item\label{item:rq1} What is the current state of the art in mechanising process calculi?
+\item\label{item:rq2} What techniques and best practices can be recommended when starting formalisation projects within concurrency theory?
+\item\label{item:rq3} What improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
+\end{enumerate}
+We have identified three key aspects which typically cause difficulties when
+mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion},
+and \emph{coinductive reasoning} for infinite behaviours.
+Our benchmark problems are designed to focus on these aspects (discussed
+in more detail in \cref{sec:challenge-problems}), with three main goals:
 
-In this work we propose a challenge where each part independently
-exercises one of the three areas, however it is often the case that a
-problem requires more than one of these techniques simultaneously. In
-this work, we explore them in isolation, in order to facilitate the
-implementation of smaller solutions. Concretely, we do not address the
-idea of how to combine these techniques, leaving the discussion of
-that to the future when we have collected enough entries to these
-problems.
+\begin{enumerate}[label=\textbf{(G\arabic*)},leftmargin=10mm]
+\item\label{item:goal-comperison-accessibility} Enable the comparison of
+  different proof mechanisation approaches, making the challenges accessible to
+  mechanisation experts who may not be familiar with concurrency theory.
 
-We do not claim in this work that these problems have never been
-approached and formalised before. In fact, the contrary is true. All
-of these problems have been addressed in several forms. This motivates
-this challenge to enable an easier comparison between solutions (since
-they all implement the same challenge), and in a setting that is
-\emph{simple enough to not be distracting, yet complex enough that we can
-compare strengths and weaknesses of each}.
+\item\label{item:goal-tutorials} Encourage the development of guidelines and
+  tutorials demonstrating and comparing the many available proof mechanisation
+  techniques, libraries, and proof assistant features.
 
-The rest of this paper is structured in the following way:
-\cref{sec:rationale} expands on the rationale for the selected
-challenge problems. Then \cref{sec:problems} presents the problems. We
-discuss the solutions in \cref{sec:solutions}, in particular
-\cref{sec:tutorial} walks through a tutorial solution. And finally
-\cref{sec:conclusion} presents some conclusions and it contains a call
-to action to invite the submission of new solutions.
+\item\label{item:goal-reusability} Prioritise the exploration of mechanisation
+  techniques that may be useful and reusable for future research.
+\end{enumerate}
+We also aim at strengthening the culture of mechanisation, by rallying the
+community to collaborate on exploring and developing new tools and techniques.
 
-\section{Rationale for choosing the problems}\label{sec:rationale}
+The three aspects we selected (linearity, scope extrusion, coinduction) are of
+course not the only ones that may cause issues in mechanisations --- but they
+are fundamental to concurrency theory, and they emerge in most mechanisations.
+The mechanisation of novel research results often requires addressing more than
+just one of these aspects at once --- and we see the combination of techniques
+as a next step beyond the scope of this challenge.%  (we discuss this in
+% \cref{sec:going-beyond}).
 
-This section is not (at least initially) meant to be in the final
-manuscript but it is to document some reasoning for choosing
-appropriate problems. The idea is to explore three areas: scope
-extrusion in name passing calculi, linearity requirements in
-behavioural types, and finally coinductive reasoning in trace
-equivalence, (bi)similarity, or subtyping in the presence of
-recursion.
+% Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate, evaluate, and compare the many tools and techniques available.
+% While the challenge problems themselves are thus not of scientific interest, we expect that the best practices and mechanisations developed by solving the challenge problems will also be useful for future research.
 
-The idea is to have three small problems in the aforementioned
-subjects that are independent of each other and that are easy to
-understand and to implement. This is important so we have more
-entries. They have to be small so it is not a lot to read, and
-independent so people may choose to implement only one of them, or
-work on them in any order. Moreover, the specification should only
-need to be met in spirit. The idea is to provide a solution author
-with some space to show the best aspects of their solution even if
-that means adapting the problem definition while keeping the spirit of
-the challenge.
-
-Finally, it will be important to mention in the rationale for problems that
-the key aspects are:
-\begin{itemize}
-\item Scope extrusion
-\item Linearity
-\item Coinduction
-\end{itemize}
-
-But the exact presentation of the concrete problems should be somewhat
-flexible. In the sense that the use of certain techniques or tools may
-require changing the presentation of the system to fit the technique.
-These changes should be allowed, and even encouraged, but the result
-should be ``morally'' the same system and there should be a
-rationale for the changes in presentation. That way we could not only
-see valid techniques but also what presentation of the rules to use
-and the rationale to chose a particular approach.
+We have begun collecting solutions to our challenges on our website:
+%
+\begin{center}
+  \url{https://concurrentbenchmark.github.io/}
+\end{center}
+%
+In the longer term, we expect to use the website for promoting best practices,
+tutorials and guidelines derived from solutions to our challenges.
+We encourage anyone interested to try the challenges using their
+favourite tools and techniques, and to send us their solutions.
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.75\textwidth]{images/benchmark.jpeg}
-  \caption{The benchmark's big picture.}
-  \label{fig:bigpic}
-\end{figure}
+\section{The challenges}\label{sec:challenge-problems}
+\label{sec:prelim}
+First, we list some common notions and conventions that we use in the
+challenges. Since the calculi under study are somewhat different, each
+section lists the changes that apply. The challenges are inspired by
+simplified versions of calculi (and proofs) from chapters 1 and 2
+of~\cite{picalcbook} and~\cite{Vasconcelos2012} where the idea is to
+use sources that can be considered standard and close to current practice.
 
-The big picture for the benchmark can be seen in \cref{fig:bigpic}
-where the three pillars (i.e.: concepts typically needed for process
-calculus proofs) appear in blue, with some ideas of what to prove. And
-a second step of combining the three aspects (for a total of six steps
-for the whole benchmark), and of course the $+1$-step for the
-combination of everything on the reader's next research papers.
+We assume the existence of some set of \emph{base values}, represented by the symbols
+\( a, b, \dots \), the existence of some set
+of \emph{variables}, represented by the symbols
+\( l, m, \dots \), and the existence of some set of \emph{names},
+represented by the symbols \( x, y, \dots \).%
+\footnote{%
+  Unlike the standard $\pi$-calculus, we distinguish variables from names to
+  better control the expressiveness of the calculi under study, and the scope of
+  the corresponding challenges: the key distinction is that names are used as communication channels (and can be sent and received in the scope extrusion challenge), whereas variables are only bound by inputs and cannot be restricted, sent nor received, in the style of value-passing CCS \cite{Milner1989}.%
+} %
+We assume that all of these sets are infinite and that their elements can be
+compared for equality.
 
-\input{challenges.tex}
+The syntax of processes includes:
+the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{parallel composition} of process \( P \) and process \( Q \).
+The two components can proceed independently of each other, or they can interact via shared names.
+%The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
 
-\section{The solutions}\label{sec:solutions}
+For communication, processes include \emph{input} and \emph{output}, whose
+signature depends on the calculus being value-passing or name-passing.
+We use here the metavariables $c,k$ to abstract over this choice
+--- i.e.~$c$ may be either a value or a variable or a name, whereas $k$ may
+be a variable or a name.
+The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
+\( c \) via \( x \), then continue as \( P \).  % The intention is that
+% the value \( v \) must be a base value when it is actually sent, and
+% this will be enforced in the semantics later on
+The process \( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
+via \( x \), then continue as \( P \) with the received element
+substituted for \( k \).  The input operator thus
+binds \( k \) in \( P \).
 
-\subsection{A tutorial solution on paper}\label{sec:tutorial}
+The process \( \Pres{x}{P} \) is the \emph{restriction} of the name
+\( x \) to \( P \), binding \( x \) in \( P \).
 
-Probably the skeleton here and more details on an appendix.
+The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
+It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
+Replication makes it possible to express infinite behaviours.
 
-\subsection{Flexibility in solving a challenge}
+We use the notation \( \freenames{P} \) to denote the set of
+names that occur free,
+% (i.e.\ not bound by a restriction) in \( P \),
+\( \boundnames{P} \) to denote the set of names that occur bound
+% (by a restriction)
+in \( P \) and  \( \freevars{P} \)
+to denote the set of variables that occur free
+% (i.e.\ not bound by an input)
+in \( P \).  We use the notation \( \boundvars{P} \)
+for the set of variables that occur bound
+% (by an input)
+in \( P \).  We use the notation \( \subst{P}{a}{l} \) to denote
+the process \( P \) with base value \( a \) substituted for variable
+\( l \). Similarly, \( \subst{P}{x}{y} \) denotes the process
+\( P \) with name \( x \) substituted for name \( y \).
+We use the notation \( \applysubst{\sigma}{P} \) to denote the process \( P \) with a finite number of arbitrary substitutions applied to it.
 
-\dots
+Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
+written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \)
+by a finite number of substitutions of bound variables.  As a
+convention, we identify \( \alpha \)-convertible processes and we
+assume that bound names and bound variables of any processes are
+chosen to be different from the names and variables that occur
+free in any other entities under consideration, such as processes,
+substitutions, and sets of names or variables.  This is justified
+because any overlapping names and variables may be
+\( \alpha \)-converted such that the assumption is satisfied.
 
-\section{Future work}
-\begin{itemize}
-\item Multiparty session types
-\item Choreography
-\item Encodings
-\item Code extraction
-\item Conversation types
-\item Psi-calculus
-\item Model checking for system properties
-\item More general types than session types
-\end{itemize}
+A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+%
+A context acts as a function between processes:
+a context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+The replacement is literal, so names and variables that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
 
-\section{Conclusion and related work} \label{sec:conclusion}
+We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+
+\subsection{Challenge: Linearity and behavioral type systems}
+\label{sec:challenge:linearity-beh-types}
+This challenge formalises a proof that requires reasoning about linearity of channels.
+Linearity is the notion that a channel must be used exactly once in a process.
+This is necessary to prove properties about session type systems, and the key
+issue of this challenge is reasoning about the linearity of context splitting operations.
+Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
+
+The setting for this challenge is a small calculus with a session type
+system, the syntax and semantics of which are given below. The
+calculus is a fragment of the one presented in~\cite{Vasconcelos2012},
+formulated in the dual style of~\cite{barber96tr}.
+
+The main objective of this challenge is to prove type preservation (also
+known as subject reduction), i.e.\ that well-typed
+processes can only transition to processes which are also well-typed
+in the same context.
+The second objective is to prove type safety, i.e.\ that well-typed processes are also well-formed in the sense that they do not use endpoints in a non-dual way.
+
+\subsubsection{Syntax.}
+The syntax is given by the grammar
+\[
+\begin{array}{r@{\qquad}c@{\qquad}l}
+  v,w & \Coloneqq & a \quad\mid\quad l \\
+   P,Q & \Coloneqq & \Pend \quad\mid\quad \Pout{x}{v}{P} \quad\mid\quad \Pin{x}{l}{P} \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad  \Presd{x}{y}{P}
+\end{array}
+\]
+where a \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable \( l \).
+
+The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \) and then continues as \( P \).
+The intention is that the value \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
+The input process \( \Pin{x}{l}{P} \) waits for a base value from \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
+%
+The process \( \Presd{x}{y}{P} \) represents a \emph{session} with endpoints named \( x \) and \( y \) which are bound in \( P \).
+In \( P \), the names \( x \) and \( y \) can be used to exchange messages over the session (sending on \( x \) and receiving on \( y \) or vice versa).
+%
+%Finally, the process $\PBpar{P}{Q}$ represents the parallel composition of $P$ and $Q$.
+%
+Note that in this calculus channels cannot be sent in messages, therefore the topology of the communication network described by a process cannot change.
+Also, there is no recursion or replication in the syntax, hence no infinite behaviors can be expressed. In particular, we only model linear (as opposed to shared) channels.
+
+\subsubsection{Semantics.}
+We describe the actions that the system can perform through a small step operational semantics.
+As usual, we use  a \emph{structural congruence} relation that equates processes that we deem to be indistinguishable.
+Structural congruence is the smallest congruence relation that satisfies the following axioms:
+\begin{mathpar}
+  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
+  \and
+  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
+  \and
+  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
+  \and
+  \inferrule[Sc-Res-Par]{\{x,y\} \cap \freenames{Q} = \emptyset}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
+  \and
+  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
+  \and
+  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
+\end{mathpar}
+
+The operational semantics is defined as the following relation on processes:
+\begin{mathparpagebreakable}
+  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{a}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{a}{l}}{R}})}}}
+  \and
+  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
+  \and
+  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
+  \and
+  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
+\end{mathparpagebreakable}
+
+Note that reductions are allowed only for restricted pairs of session endpoints.
+This makes it possible to formulate subject reduction so that the typing context is exactly the same before and after each reduction.
+%
+Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
+Since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable or a name.
+
+\subsubsection{Session types.}
+Our process syntax allows us to write processes that are ill formed in
+the sense that they either use the endpoints bound by a restriction to
+communicate in a way that does not follow the intended duality, or
+attempt to send something which is not a base value.  As an example,
+the process
+\( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \)
+attempts to send a base value on both \( x \) and \( y\), whereas one
+of the names should be used for receiving in order to guarantee
+progress.  Another example is the process
+\( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \),
+which attempts to send a variable that is not instantiated at the time
+of sending.
+%
+To prevent these issues, we introduce a \emph{session type system} which
+rules out ill-formed processes.
+
+\subsubsection{Syntax.}
+Our type system does not type processes directly, but instead focuses on the channels used in the process.
+The syntax of \emph{session types} \( S, T \), unrestricted type contexts \( \Gamma \) and linear type contexts \( \Delta \) is as follows:
+\[
+  \begin{array}{r@{\qquad}c@{\qquad}l}
+  S,T & \Coloneqq & \Tend \quad\mid\quad \Tbase \quad\mid\quad \Tin{S} \quad\mid\quad \Tout{S} \\
+    \Gamma & \Coloneqq & \Cempty \quad\mid\quad \Gamma, l \\
+                    \Delta &::= & \Cempty \quad\mid\quad \Cadd{\Delta}{\hastype{x}{S}}
+  \end{array}
+\]
+The \emph{end type} \( \Tend \) describes an endpoint on which no further interaction is possible.
+The \emph{base type} \( \Tbase \) describes base values.
+The \emph{input type} \( \Tin{S} \) describes endpoints used for receiving a value and then according to \( S \).
+The \emph{output type} \( \Tout{S} \) describes endpoints used for sending a value and then according to \( S \).
+
+Typing contexts gather type information about names and variables.
+\emph{Unrestricted} contexts are simply sets of names since we only have one
+base type. \emph{Linear contexts} associate a type to endpoints. We use
+the comma as split/union, overloaded to singletons, and \( \Cempty \) as the
+empty context. We extend the Barendregt convention to contexts, so that all
+entries are distinct.  Note that the order in which information is added to a
+type context does not matter.
+
+Since we need to determine whether endpoints are used in complementary ways to determine whether processes are well formed, we need to formally define the dual of a type as follows:
+\begin{mathpar}
+  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
+  \and
+  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
+  \and
+  \inferrule{}{\dual{\Tend} = \Tend}
+\end{mathpar}
+Note that the dual function is partial since it is undefined for the base type.
+
+\subsubsection{Typing rules.}
+Our type system is aimed at maintaining two invariants:
+\begin{enumerate}
+\item No endpoint is used simultaneously by parallel processes;
+\item The two endpoints of the same session have dual types.
+\end{enumerate}
+The first invariant is maintained by linearly splitting type contexts when typing compositions of processes, the second by requiring duality when typing restrictions.
+
+We have two typing judgments: one for values, and one for processes.
+The typing rules for values are:
+\begin{mathpar}
+  \inferrule[T-Base]{ }{\typev{\Gamma}{\hastype{a}{\Tbase}}} \and
+  \inferrule[T-Var]{ }{\typev{\Cadd{\Gamma}l }{\hastype{l}{\Tbase}}}
+\end{mathpar}
+% In the typing rule for names, $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
+% \begin{mathpar}
+%   \inferrule[T-Name]{{\tend\Delta }}{\types{{\Gamma; (\Delta,\hastype{x}{T})}}{\hastype{x}{T}}}
+% \end{mathpar}
+The typing rules for processes are as follows:
+\begin{mathpar}
+
+  \inferrule[T-Inact]{\tend\Delta }{\types{\Gamma;\Delta}{\Pend}}
+  \and
+  \inferrule[T-Par]{\types{\Gamma;\Delta_1}{P} \\ \types{\Gamma;\Delta_2}{Q}}
+  {\types{\Gamma; \Csplit{\Delta_1}{\Delta_2}}{\Ppar{P}{Q}}}
+  \and
+  \inferrule[T-Res]{\types{\Gamma; (\Cadd{\Cadd{\Delta}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P})}{\types{\Gamma}{\Presd{x}{y}{P}}}
+  \and
+    \inferrule[T-Out]{
+      \typev{\Gamma}{\hastype{v}{\Tbase}} \\ \types{\Gamma; \Cupdate{\Delta}{\hastype{x}{T}}}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tout{T}}})}{\Pout{x}{v}{P}}}
+    \and
+    \inferrule[T-IN]{
+      \types{(\Gamma,  l ); (\Cupdate{\Delta}{\hastype{x}{T}})}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tin{T}}})}{\Pin{x}{l}{P}}}
+\end{mathpar}
+Note that we do not need a judgment for typing channels, since it is already folded into the \TirName{T-In} and \TirName{T-Out} rules.
+
+\subsubsection{Challenge.}
+The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with:
+\begin{theorem}[Subject reduction]
+  \label{thm:subject-reduction}
+  If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
+\end{theorem}
+
+The following lemmas are needed:
+\begin{lemma}[Weakening]\mbox{}
+  \label{lemma:weak}
+  \begin{enumerate}
+  \item If \( \types{\Gamma; \Delta}{P} \) then
+    \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
+  \item If \( \types{\Gamma; \Delta}{P} \) % and \( \tend{\Delta} \)
+    then \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
+  \end{enumerate}
+\end{lemma}
+
+\begin{lemma}[Strengthening]\mbox{}
+  \label{lemma:strenD}
+  If \( \types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P} \) and
+  $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
+\end{lemma}
+
+\begin{lemma}[Substitution]\mbox{}
+  \label{le:subst}
+  If $\types{(\Cadd{\Gamma}{l});\Delta}{P}$ and
+  ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
+  \( \types{\Gamma;\Delta}{\subst{P}{a}{l}} \).
+\end{lemma}
+
+To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms.
+
+\begin{lemma}[Preservation for $\stackrel{a}{\equiv}$]
+  \label{le:presequiva}
+  If \( \sconga{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
+\end{lemma}
+
+Now we give rules for the compatible
+equivalence relation induced by $\sconga\cdot\cdot$, which we still
+write as $\scong\cdot\cdot$: namely, add to reflexivity, symmetry and
+transitivity the following condition:
+ \begin{mathpar}
+   \inferrule[Cong]{\sconga P Q}{\scong {C[P]}{C[Q]}}
+ \end{mathpar}
+
+\begin{lemma}[Preservation for $\equiv$]
+  \label{le:presequiv}
+    If \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
+\end{lemma}
+
+To formulate safety, we need to formally define what we mean by
+well-formed process.  We say that a process \( P \) is \emph{prefixed
+  at variable \( x \)} if \( P\equiv\Pout{x}{v}{P'} \) or
+\( P\equiv\Pin{x}{l}{P'} \) for some $P'$.
+%
+A process $P$ is then \emph{well formed} if, for every $P_1$, $P_2$,
+and $R$ such that
+\( P\equiv \Presd{x_1}{y_1}{\dots
+  \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}} \), with
+\( n \geq 0 \), it holds that, if \( P_1 \) is prefixed at \( x_1 \)
+and \( P_2 \) is prefixed at \( y_1 \) (or vice versa), then
+\( \Ppar{P_1}{P_2}
+\equiv\Ppar{\Pout{x_1}{a}{P_1'}}{\Pin{y_1}{l}{P_2'}} \), for some
+$P_1'$ and $P_2'$.
+
+Note that well-formed processes do not necessarily reduce. For example, the process
+\begin{equation*}
+  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
+\end{equation*}
+is well formed but also irreducible.
+
+\begin{theorem}[Type safety]
+  \label{thm:type-safety}
+  If \( \types{\Cempty}{P} \), then \( P \) is well formed.
+\end{theorem}
+
+\begin{corollary}
+  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well formed.
+\end{corollary}
+
+\subsection{Challenge: Name passing and scope extrusion}
+\label{sec:challenge:name-passing-scope-extrusion}
+
+This challenge formalises a proof that requires explicit scope extrusion.
+Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
+This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
+The key issue of this challenge is reasoning about names that are ``in the process'' of being scope-extruded, which often presents difficulties for the mechanisation of binders.
+
+Reasoning about scope extrusion explicitly can sometimes be avoided by introducing a structural congruence rule into the semantics, but doing this means we lose information about the scope when reasoning about the semantics.
+Explicitly reasoning about scope extrusion is necessary to describe e.g.\ runtime monitors and compositions of systems.
+
+The setting for this challenge is a ``classic'' untyped \( \pi \)-calculus, where (unlike the calculi in the other challenges) names can be sent and received, and bound by input constructs (similarly to variables in the other calculi).
+We define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.
+The objective of this challenge is to prove that the two semantics are equivalent up to structural congruence.
+
+\subsubsection{Syntax.}
+The syntax of processes is given by:
+\[
+\begin{array}{r@{\qquad}c@{\qquad}l}
+  P,Q & \Coloneqq & \Pend \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad \Pout{x}{y}{P} \quad\mid\quad \Pin{x}{y}{P} \quad\mid\quad \Pres{x}{P}
+\end{array}
+\]
+
+%The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
+The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
+The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
+The input operator thus binds the name \( y \) in \( P \).
+% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
+% The two components can proceed independently of each other, or they can interact via shared names.
+% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
+% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
+% The restriction operator thus binds the name \( x \) in \( P \).
+Note that the scope of a restriction may change when processes interact. Namely, a restricted name may be sent \emph{outside} of its scope.
+% The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
+Note that there is no recursion or replication in the syntax, and thus no infinite behaviors can be expressed.
+This simplifies the theory and is orthogonal to the concept of scope extrusion.
+
+% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
+% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
+% We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
+
+% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
+% As a convention, we will identify \( \alpha \)-convertible processes.
+
+% As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
+% This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
+
+\subsubsection{Reduction semantics.}
+The first semantics is an operational reduction semantics, which avoids reasoning explicitly about scope extrusion by way of a structural congruence rule.
+\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
+
+\begin{mathpar}
+  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{P}{(\Ppar{Q}{R})}}{\Ppar{(\Ppar{P}{Q})}{R}}}
+  \and
+  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
+  \and
+  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
+  \and
+  \inferrule[Sc-Res-Par]{x \notin \freenames{Q}}{\scong{\Ppar{\Pres{x}{P}}{Q}}{\Pres{x}{(\Ppar{P}{Q})}}}
+  \and
+  \inferrule[Sc-Res-Inact]{ }{\scong{\Pres{x}{\Pend}}{\Pend}}
+  \and
+  \inferrule[Sc-Res]{ }{\scong{\Pres{x}{\Pres{y}{P}}}{\Pres{y}{\Pres{x}{P}}}}
+\end{mathpar}
+
+The operational semantics is defined as the following relation on processes:
+\begin{mathpar}
+  \inferrule[R-Com]{ }{\reduces{\Ppar{\Pout{x}{y}{P}}{\Pin{x}{z}{Q}}}{\Ppar{P}{\subst{Q}{y}{z}}}}
+  \and
+  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Pres{x}{P}}{\Pres{x}{Q}}}
+  \and
+  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
+  \and
+  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
+\end{mathpar}
+Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{R-Com}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
+Finally, note that rule \TirName{R-Struct} allows for applying the structural congruence both before and after the reduction:
+this makes the reduction relation closed under structural congruence.
+
+\subsubsection{Transition system semantics.}
+The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
+The transitions are labelled by \emph{actions}, the syntax of which is:
+\[
+\begin{array}{r@{\qquad}c@{\qquad}l}
+  \alpha & \Coloneqq & \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
+\end{array}
+\]
+
+The \emph{free output action} \( \Aoutf{x}{y} \) is sending the name \( y \) via \( x \).
+The \emph{input action} \( \Ain{x}{y} \) is receiving the name \( y \) via \( x \).
+The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
+The \emph{internal action} \( \Atau \) is performing internal communication.
+
+We extend the notion of free and bound occurrences with
+\( \freenames{\alpha} \) to denote the set of names that occur free in
+the action \( \alpha \) and  \( \boundnames{\alpha} \) to
+denote the set of names that occur bound in the action \( \alpha \).
+In the free output action \( \Aoutf{x}{y} \) and the input action
+\( \Ain{x}{y} \), both \( x \) and \( y \) are free names.  In the
+bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while
+\( y \) is a bound name.  We also use the notation
+\( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \)
+and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur
+in the action \( \alpha \).
+The transition relation is then defined by the following rules:
+\begin{mathparpagebreakable}
+  \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
+  \and
+  \inferrule[In]{ }{\transition{\Pin{x}{z}{P}}{\Ain{x}{y}}{\subst{P}{y}{z}}}
+  \and
+  \inferrule[Par-L]{\transition{P}{\alpha}{P'} \\ \boundnames{\alpha} \cap \freenames{Q} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
+  \and
+  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'} \\ \boundnames{\alpha} \cap \freenames{P} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
+  \and
+  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{y}}{P'} \\ \transition{Q}{\Ain{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Comm-R]{\transition{P}{\Ain{x}{y}}{P'} \\ \transition{Q}{\Aoutf{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Close-L]{\transition{P}{\Aoutb{x}{z}}{P'} \\ \transition{Q}{\Ain{x}{z}}{Q'} \\ z \notin \freenames{Q}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
+  \and
+  \inferrule[Open]{\transition{P}{\Aoutf{x}{z}}{P'} \\ z \neq x}{\transition{\Pres{z}{P}}{\Aoutb{x}{z}}{P'}}
+  \and
+  \inferrule[Close-R]{\transition{P}{\Ain{x}{z}}{P'} \\ \transition{Q}{\Aoutb{x}{z}}{Q'} \\ z \notin \freenames{P}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
+  \and
+  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
+\end{mathparpagebreakable}
+Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{In}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
+
+We keep the convention that bound names of any processes or actions
+are chosen to be different from the names that occur free in any other
+entities under consideration, such as processes, actions,
+substitutions, and sets of names.  The convention has one exception,
+namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \),
+the name \( z \) (which occurs bound in \( P \) and the action
+\( \Aoutb{x}{z} \)) may occur free in \( Q \).  Without this exception
+it would be impossible to express scope extrusion.
+
+Before getting to our challenge, we prove some intermediate results.
+\begin{lemma}\label{se-lemma-harmony-fact}
+  If \( \scong{P}{Q} \) and \( \transition{P}{\alpha}{P'} \), then for some \( Q' \) we have \( \transition{Q}{\alpha}{Q'} \) and \( \scong{P'}{Q'} \).
+\end{lemma}
+
+For the proof of the challenge theorem, we introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
+The first rule applied is \TirName{R-Com}. The derivation continues with an application of \TirName{R-Par}, followed by zero or more applications of \TirName{R-Res}.
+The last rule is an application of \TirName{R-Struct}.
+\begin{lemma}\label{se-lemma-normalized-derivation}
+  Every reduction has a normalized derivation.
+\end{lemma}
+
+\begin{lemma}\label{se-lemma-normalized-reduction}
+  If \( \reduces{P}{Q} \) then there are \( x, y, z, z_1, \dots, z_n, R_1, R_2, \) and \( S\) such that
+  \begin{align*}
+    &\scong{P}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{\Pout{x}{y}{R_1}}{\Pin{x}{z}{R_2}})}{S})}}} \\
+    &\scong{Q}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{R_1}{\subst{R_2}{y}{z}})}{S})}}}
+  \end{align*}
+\end{lemma}
+
+\subsubsection{Challenge.}
+The objective of this challenge is to prove the following theorems, which together show the equivalence between the reduction semantics and the transition system semantics up to structural congruence.
+The first of the theorems involves reasoning about scope extrusion more directly than the other, and if time does not permit proving both of the theorems, \cref{thm:se-trans-implies-red} should be proven first.
+\begin{theorem}\label{thm:se-trans-implies-red}
+  \( \transition{P}{\Atau}{Q} \) implies \( \reduces{P}{Q} \).
+\end{theorem}
+
+\begin{theorem}\label{thm:harmony-2}
+  \( \reduces{P}{Q} \) implies the existence of a \( Q' \) such that \( \transition{P}{\Atau}{Q'} \) and \( \scong{Q}{Q'} \).
+\end{theorem}
+
+\subsection{Challenge: Coinduction and infinite processes}
+\label{sec:challenge:coinduction}
+This challenge is about the mechanisation of proofs concerning processes with infinite behaviours.
+% formalises a proof
+% that require coinduction.
+% coinductive techniques.
+%
+% Coinduction is a proof technique for infinite structures,
+This is usually connected to \emph{coinductive} definitions where an infinite
+structure is defined as the greatest fixed point of a recursive
+definition.
+%
+Coinduction is a technique for proving properties of
+such infinite structures.
+%
+%Coinduction is the dual of induction: whereas induction is useful for
+%proving properties of least fixed points, coinduction is useful for
+%proving properties of greatest fixed points.
+%
+% \marco{Should we be more specific here and give a deeper explanation
+%   of what coinduction is? If not, then why do we even say that it's
+%   the dual of induction?}
+%
+
+For this challenge, we adopt a fragment of the untyped $\pi$-calculus
+that includes process replication.
+% The setting for this challenge is an untyped calculus of
+% communicating systems with replication of processes, the syntax and
+% semantics of which are presented below.
+The objective of this challenge is to % prove % a context theorem for
+draw a formal connection between strong barbed congruence and strong
+barbed bisimilarity.  % I.e., The context lemma
+The result establishes that two processes are strong barbed congruent
+if the processes obtained by applying a finite number of substitutions to
+them and composing them in parallel with an arbitrary process are
+strongly barbed bisimilar.
+The key issue of this challenge is the coinductive reasoning about the infinite behaviours of the replication operator.
+
+\subsubsection{Syntax.}
+% We assume the existence of some type of \emph{base values}, values
+% of which we will denote by \( a, b, \dots \), the existence of some
+% type of \emph{variables}, values of which we will denote by
+% \( l, m, \dots \), and the existence of some type of \emph{names},
+% values of which we will denote by \( x, y, \dots \).
+The syntax of values and processes is given by:
+% for including replication and having unary restriction:
+\begin{displaymath}
+  \begin{array}{r@{\qquad}c@{\qquad}l}
+    v,w & \Coloneqq & a \quad\mid\quad l \\
+    P,Q & \Coloneqq & \Pend
+               \quad \mid\quad \Pout{x}{v}{P}
+               \quad \mid\quad \Pin{x}{l}{P}
+               \quad \mid\quad \PBpar{P}{Q}
+               \quad \mid\quad \Pres{x}{}{P}
+               \quad \mid\quad !P
+  \end{array}
+\end{displaymath}
+The output process \( \Pout{x}{v}{P} \) sends the value \( v \) on channel \( x \) and continues as \( P \).
+The intention is that \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
+The input process \( \Pin{x}{l}{P} \) waits for a base value from channel \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
+Since replication allows for infinite copies of the process $P$, processes can dynamically
+create an infinite number of names during execution.
+
+%
+% \begin{align*}
+%   v,w :=&&& a \\
+%   |&&& l \\
+%   P,Q :=&&& \Pend \\
+%   |&&& \Pout{x}{v}{P} \\
+%   |&&& \Pin{x}{l}{P} \\
+%   |&&& \Ppar{P}{Q} \\
+%   |&&& \Pres{x}{P} \\
+%   |&&& \Preplicate{P}
+% \end{align*}
+% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a
+% variable \( l \).  The process \( \Pend \) is \emph{inaction}:
+% a process which can do nothing.  The process \( \Pout{x}{a}{P} \) is
+% an \emph{output}, which can send the value \( v \) via \( x \), then
+% continue as \( P \).  The intention is that the value \( v \) must
+% be a base value when it is actually sent, and this will be enforced
+% in the semantics later on.  The process \( \Pin{x}{l}{P} \) is an
+% \emph{input}, which can receive a base value via \( x \), then
+% continue as \( P \) with the received value substituted for the
+% variable \( l \).  The input operator thus binds the variable
+%  \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
+% \emph{composition} of process \( P \) and process \( Q \).  The two
+% components can proceed independently of each other, or they can
+% interact via shared names.  The process \( \Pres{x}{P} \) is the
+% \emph{restriction} of the name \( x \) to \( P \).  Components in
+% \( P \) can use the name \( x \) to interact with each other, but
+% not with processes outside of the restriction.  The restriction
+% operator thus binds the name \( x \) in \( P \).  Note that the
+% scope of a restriction may not change when processes interact, since
+% it is only possible to send and receive values, and not names.  The
+% process \( \Preplicate{P} \) is the \emph{replication} of the
+% process \( P \).  It can be thought of as the infinite composition
+% \( \Ppar{P}{\Ppar{P}{\cdots}} \).  Replication makes it possible to
+% express infinite behaviours.
+
+% We will use the notation \( \subst{P}{v}{l} \) to denote the process
+% \( P \) with value \( v \) substituted for variable \( l \).
+
+% % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
+% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
+% written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from
+% \( P \) by a finite number of substitutions of bound variables.
+% As a convention, we will identify \( \alpha \)-convertible
+% processes.
+
+% Also as a convention, we assume that the bound names and bound
+% variables occurring in any collection of processes are chosen
+% to be different from the free names and free variables
+% occurring in those processes and from the names and variables
+% occurring in any substitutions applied to the processes.  This is
+% justified because any overlapping names and variables may be
+% \( \alpha \)-converted such that the assumption is satisfied.
+
+\subsubsection{Semantics.} We choose to give an LTS semantics for this
+challenge. %As an alternative, it is also possible to adopt a reduction
+%semantics which features fewer rules but needs a structural
+%congruence. %  The semantics of the system describes the actions that the
+% system can perform by defining a labelled transition relation on
+% processes.
+The transitions are labelled by \emph{actions}, the syntax
+of which is as follows:
+% \begin{align*}
+%   \alpha := &&& \Aoutf{x}{a} \\
+%   |&&& \Ain{x}{a} \\
+%   |&&& \Atau
+% \end{align*}
+\begin{align*}
+  \alpha\ \ \Coloneqq\ \ \Aoutf{x}{a} \ \mid\ \Ain{x}{a} \ \mid\ \Atau
+\end{align*}
+The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value
+\( a \) via \( x \).  The \emph{input action} \( \Ain{x}{y} \) is
+receiving the base value \( y \) via \( x \).  The \emph{internal
+  action} \( \Atau \) is performing internal communication.
+%
+We use the notation \( \names{\alpha} \) to denote the set of
+names that occur in the action \( \alpha \).
+The transition relation is defined by the following rules:
+\begin{mathparpagebreakable}
+  \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
+  \and
+  \inferrule[In]{ }{\transition{\Pin{x}{l}{P}}{\Ain{x}{a}}{\subst{P}{a}{l}}}
+  \and
+  \inferrule[Par-L]{\transition{P}{\alpha}{P'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
+  \and
+  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
+  \and
+  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{a}}{P'} \\ \transition{Q}{\Ain{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Comm-R]{\transition{P}{\Ain{x}{a}}{P'} \\ \transition{Q}{\Aoutf{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
+  \and
+  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ x \notin \names{\alpha}}{\transition{\Pres{x}{P}}{\alpha}{\Pres{x}{P'}}}
+  \and
+  \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
+\end{mathparpagebreakable}
+Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
+Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
+On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable.
+
+% As a convention, we assume that bound names and bound variables of any processes or actions are chosen to be different from the names and variables that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variables.
+
+\subsubsection{Bisimilarity.}
+Bisimilarity is a notion of equivalence for processes and
+% .  Our notion of process equivalence relations
+builds on the idea of \emph{observables}, i.e., what we can externally
+observe from the semantics of a process. If we allowed ourselves only
+to observe internal transitions (i.e., observe that a process is
+internally performing a step of computation) we would relate either
+too few processes (in the strong case where we relate only processes
+with exactly the same number of internal transitions) or every process
+(in the weak case where we relate processes with any amount of
+internal transitions).  As a result, we must allow ourselves to
+monitor more than just internal transitions, and we choose to describe
+a process's observables as the names it might use for sending and
+receiving.
+% We must therefore allow ourselves to observe more than
+% just internal transitions, and here we choose to define the observables
+% of a process as the names it can use for sending and receiving.
+To this end, we define the \emph{observability predicate}
+\( \observable{P}{\mu} \) as follows:
+\begin{align*}
+  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
+  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
+\end{align*}
+
+\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
+\begin{gather}
+  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \label{eq:bisim1}\\
+  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'} \label{eq:bisim2}
+\end{gather}
+We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
+Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
+
+\begin{theorem}\label{thm:bisim-equiv}
+  \( \sbbisim{}{} \) is an equivalence relation. %, that is, the relation is reflexive, symmetric, and transitive.
+\end{theorem}
+
+Unfortunately, strong barbed bisimilarity is not a good process
+equivalence since it is not a congruence, hence it does not allow for
+substituting a process with an equivalent one in any context.
+% equivalent processes to be since it does not consider the
+% environment of processes.
+For instance, the processes $\Pout{x}{a}{\Pout{y}{b}{\Pend}}$ and
+$\Pout{x}{a}{\Pend}$ are strong barbed bisimilar, i.e.,
+$\sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}}$. This
+is because \( \obsout{x} \) is the only observable in both processes
+and they cannot perform a \( \Atau \)-action. However, in the context
+$C = \Ppar{\ctxhole}{\Pin{x}{l}{\Pend}}$, the relation no
+longer holds: in fact,
+$\nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}\Ppar{}{\Pin{x}{l}{\Pend}}}$
+%
+because the left process can perform a \( \Atau \)-action such that
+\( \obsout{y} \) becomes observable, whereas the right process cannot.
+
+% For instance,
+% \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \)
+% since \( \obsout{x} \) is the only observable in both processes and
+% they cannot perform a \( \Atau \)-action, but
+% \(
+% \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}}
+% \) since the left process can perform a \( \Atau \)-action such that
+% \( \obsout{y} \) becomes observable, whereas the right process cannot.
+
+% \subsubsection{Contexts and congruences}
+% Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
+
+% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
+% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
+
+% We can think of contexts as functions between processes.
+% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
+% The replacement should be literal, so names and variables that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
+
+% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+
+\subsubsection{Strong barbed congruence.}
+In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,
+i.e., % we need to make strong barbed bisimilarity consider
+we have to consider the environment in which processes may be placed.
+% (i.e.\ the context) of processes.
+
+% We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying
+We say that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
+
+\begin{lemma}\label{lemma:sbcong-largest}
+  \( \sbcong{}{} \) is the largest congruence included in
+  \( \sbbisim{}{} \).
+\end{lemma}
+
+\subsubsection{Challenge.}
+The objective of this challenge is to prove a theorem that shows that
+making strong barbed bisimilarity sensitive to substitution and
+parallel composition is enough to show strong barbed
+congruence.
+To prove the theorem, we will use an \emph{up-to technique}, utilizing the following definition and lemma.
+A relation \( \mathcal{S} \) is called a \emph{strong barbed bisimulation up to \( \sbbisim{}{} \)} if, whenever \( (P,Q) \in \mathcal{S} \), the following conditions hold:
+\begin{enumerate}
+\item \( \observable{P}{\mu} \) if and only if \( \observable{Q}{\mu} \).
+\item if \( \transition{P}{\Atau}{P'} \) then \( \transition{Q}{\Atau}{Q'} \) for some \( Q' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
+\item if \( \transition{Q}{\Atau}{Q'} \) then \( \transition{P}{\Atau}{P'} \) for some \( P' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
+\end{enumerate}
+\begin{lemma}\label{lemma:up-to}
+  If \( \mathcal{S} \) is a strong barbed bisimulation up to \( \sbbisim{}{} \), \( (P,Q) \in \mathcal{S} \) implies \( \sbbisim{P}{Q} \).
+\end{lemma}
+
+\begin{theorem}\label{thm:coinduction}
+  \( \sbcong{P}{Q} \) if for any process \( R \) and
+  substitution \( \sigma \),
+  \(
+  \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}}
+  \).
+\end{theorem}
+
+\section{The proofs}
+
+\input{../proof-appendix/proofs.tex}
 
 \bibliographystyle{plainnat}
-\bibliography{references}
+\bibliography{../references}
 

### 2024-01-09 existing-mechanisations/existing-mechanisations.tex
Add list of existing mechanisations.

\subsection{Unprocessed mechanisations}
\begin{itemize}
\item \url{https://doi.org/10.4230/LIPIcs.ITP.2023.28}
\item \url{https://doi.org/10.1007/978-3-030-85315-0_8}
\item \url{https://doi.org/10.4230/LIPIcs.ITP.2022.27}
\item \url{https://doi.org/10.1007/978-3-662-54434-1_34}
\item \url{https://doi.org/10.1145/3354166.3354184}
\item \url{https://doi.org/10.1017/S0960129514000231}
\item \url{https://doi.org/10.1016/j.entcs.2007.11.010}
\item \url{https://doi.org/10.1016/S1571-0661(04)80507-3}
\item \url{https://doi.org/10.1007/BFb0028392}
\item T. F. Melham. 1994. A mechanized theory of the $\pi$-calculus in HOL. Nordic J. of Computing 1, 1 (Spring 1994), 50–76. (\url{https://dl.acm.org/doi/10.5555/640186.640190}, but paper is missing)
\item Otmane Ait-Mohamed. Vérification de l'équivalence du $\pi$-calcul dans HOL. RR-2412, INRIA. 1994 (\url{https://hal.science/inria-00074263/}, in French)
\item \url{https://doi.org/10.1017/S0956796802004653}
\item \url{https://doi.org/10.1007/3-540-45315-6_24}
\item \url{https://inria.hal.science/inria-00072970}
\item \url{https://www.lix.polytechnique.fr/Labo/Dale.Miller/lProlog/examples/pi-calculus/toc.html}
\item \url{https://doi.org/10.1007/BFb0105404}
\item \url{https://doi.org/10.1007/3-540-44755-5_16}
\item \url{https://doi.org/10.1007/3-540-44929-9_30}
\item \url{https://doi.org/10.1007/10721959_33}
\item \url{https://doi.org/10.1016/S0304-3975(00)00095-5}
\item \url{https://doi.org/10.1017/S096012951700010X}
\item \url{https://doi.org/10.1016/j.entcs.2007.11.013}
\item \url{https://doi.org/10.1145/1656242.1656248}
\item \url{https://doi.org/10.1184/R1/6587294.v1}
\item \url{https://doi.org/10.6092/issn.1972-5787/4650}
\item \url{https://doi.org/10.2168/LMCS-5(2:16)2009}
\item \url{https://doi.org/10.4204/EPTCS.203.1}
\item \url{https://doi.org/10.1007/978-3-030-45237-7_17}
\item Uma Zalakain. 2019. Type-checking session-typed $\pi$-calculus with Coq. Master's thesis, University of Glasgow. \url{https://umazalakain.info/static/Z19/msc-thesis.pdf}
\item Adam Michael Petz. 2016. A Semantics for Attestation Protocols using Session Types in Coq. Master's thesis, University of Kansas. \url{http://hdl.handle.net/1808/24137}
\item \url{https://pure.itu.dk/en/publications/representing-session-types}
\item \url{https://doi.org/10.48550/arXiv.1603.03727}
\item \url{https://doi.org/10.1145/3498684}
\item \url{https://doi.org/10.4230/LIPIcs.ITP.2021.15}
\item \url{https://doi.org/10.1007/978-3-319-22102-1_19}
\item \url{https://doi.org/10.1017/S0960129513000170}
\item \url{https://doi.org/10.1007/s10817-015-9336-2}
\item \url{https://doi.org/10.1007/978-3-540-69407-6_33}
\item \url{https://doi.org/10.1016/j.entcs.2007.08.017}
\item \url{https://doi.org/10.1145/3453483.3454041}
\item \url{https://doi.org/10.4204/EPTCS.314.3}
\item \url{https://doi.org/10.7494/csci.2017.18.3.1413}
\item \url{https://doi.org/10.1145/3371074}
\item \url{https://doi.org/10.1145/3158116}
\end{itemize}

A total of 44 papers that we currently know about.

\subsection{Template}
Abstract

\subsubsection{Approach}
\begin{itemize}
\item Proof assistant
\item 
\end{itemize}

\subsubsection{Quotes about challenges encountered}

\subsubsection{Suggestions in the paper}

### 2024-01-10 existing-mechanisations/existing-mechanisations.tex
Add table and BibTeX for existing mechanisations.

@@
 
-\subsection{Template}
-Abstract
+\subsection{Approaches}
+\begin{tabular}{l|c|c|c|c|l}
+  Reference          & Proof assistant & Linearity & Binders & Coinduction & Limitations \\\hline
+  \cite{Tirore:2023} &&&&& \\
+  \cite{Cruz-Filipe2021} &&&&& \\
+  \cite{Pohjola2022} &&&&& \\
+  \cite{Tassarotti2017} &&&&& \\
+  \cite{Thiemann2019} &&&&& \\
+  \cite{Goto2016} &&&&& \\
+  \cite{Affeldt2008} &&&&& \\
+  \cite{Scagnetto2002} &&&&& \\
+  \cite{Hirschkoff1997} &&&&& \\
+  \cite{Melham1994} &&&&& \\
+  \cite{Ait-Mohamed1994} &&&&& \\
+  \cite{Rockl2003} &&&&& \\
+  \cite{Rockl2001} &&&&& \\
+  \cite{Henry-Greard1999} &&&&& \\
+  \cite{Miller2023} &&&&& \\
+  \cite{Gordon1996} &&&&& \\
+  \cite{Gay2001} &&&&& \\
+  \cite{Despeyroux2000} &&&&& \\
+  \cite{Gillard2000} &&&&& \\
+  \cite{Honsell2001} &&&&& \\
+  \cite{Perera2018} &&&&& \\
+  \cite{Watkins2008} &&&&& \\
+  \cite{Tiu2010} &&&&& \\
+  \cite{Cervesato2007} &&&&& \\
+  \cite{Baelde2014} &&&&& \\
+  \cite{Bengtson2009} &&&&& \\
+  \cite{Orchard2016} &&&&& \\
+  \cite{Castro2020} &&&&& \\
+  \cite{Zalakain2019} &&&&& \\
+  \cite{Petz2016} &&&&& \\
+  \cite{Bock2016} &&&&& \\
+  \cite{Xi2016} &&&&& \\
+  \cite{Hirsch2022} &&&&& \\
+  \cite{Cruz-Filipe2021b} &&&&& \\
+  \cite{Maksimovic2015} &&&&& \\
+  \cite{Parrow2014} &&&&& \\
+  \cite{Bengtson2016} &&&&& \\
+  \cite{Kahsai2008} &&&&& \\
+  \cite{Bengtson2007} &&&&& \\
+  \cite{Castro-Perez2021} &&&&& \\
+  \cite{Gay2020} &&&&& \\
+  \cite{Brady2017} &&&&& \\
+  \cite{Hinrichsen2019} &&&&& \\
+  \cite{Sergey2017} &&&&&
+\end{tabular}
 
-\subsubsection{Approach}
-\begin{itemize}
-\item Proof assistant
-\item 
-\end{itemize}
+\subsection{Existing experiences and suggestions}
+In this section, we collect quotes about experiences attempting to mechanise process calculi and related systems, as well as suggestions proposed by other authors.
 
 \subsubsection{Quotes about challenges encountered}
+This section collects quotes about challenges encountered using various techniques for mechanisation.
 
-\subsubsection{Suggestions in the paper}
+\subsubsection{Quotes about effort on technical lemmas vs.\ effort on actual content}
+This section collects quotes about the amount of effort spent on technical lemmas required by the mechanisation technique versus the amount of effort spent on the theorems of interest.
+
+\subsubsection{Quotes about suggestions for improvements in proof assistant technology}
+This section collects quotes suggesting improvements in proof assistant technology and/or mechanisation techniques.

### 2024-01-15 short-paper/short-paper.tex
Add first iteration of short paper draft.

--- 2023-10-18 paper/paper.tex
@@
 \and Alceste Scalas \inst{5}\orcidID{0000-0002-1153-6164}
+\and Dawit Tirore \inst{1}\orcidID{0000-0002-1997-5161}
 \and Martin Vassor \inst{8}\orcidID{0000-0002-2057-0495}
@@
 \institute{
-     IT University of Copenhagen, Copenhagen, Denmark \email{maca@itu.dk}
+     IT University of Copenhagen, Copenhagen, Denmark \email{maca@itu.dk}, \email{dati@itu.dk}
 \and University of Kent, Canterbury, United Kingdom \email{D.Castro-Perez@kent.ac.uk}
@@
 Our benchmark problems are designed to focus on these aspects (discussed
-in more detail in \cref{sec:challenge-problems}), with three main goals:
+in detail in \cref{app:challenges}), with three main goals:
 
@@
 \paragraph{Outline of this paper.}
-In \cref{sec:challenge-problems} we describe our three challenges (proofs of the theorems are available in \cref{app:challenges}).
 We discuss our design choices in \cref{sec:design-discussion}, and possible extensions and further topics in \cref{sec:going-beyond}.
-
-\section{The challenges}\label{sec:challenge-problems}
-\label{sec:prelim}
-First, we list some common notions and conventions that we use in the
-challenges. Since the calculi under study are somewhat different, each
-section lists the changes that apply. The challenges are inspired by
-simplified versions of calculi (and proofs) from chapters 1 and 2
-of~\cite{picalcbook} and~\cite{Vasconcelos2012} where the idea is to
-use sources that can be considered standard and close to current practice.
-
-We assume the existence of some set of \emph{base values}, represented by the symbols
-\( a, b, \dots \), the existence of some set
-of \emph{variables}, represented by the symbols
-\( l, m, \dots \), and the existence of some set of \emph{names},
-represented by the symbols \( x, y, \dots \).%
-\footnote{%
-  Unlike the standard $\pi$-calculus, we distinguish variables from names to
-  better control the expressiveness of the calculi under study, and the scope of
-  the corresponding challenges: the key distinction is that names are used as communication channels (and can be sent and received in the scope extrusion challenge), whereas variables are only bound by inputs and cannot be restricted, sent nor received, in the style of value-passing CCS \cite{Milner1989}.%
-} %
-We assume that all of these sets are infinite and that their elements can be
-compared for equality.
-
-The syntax of processes includes:
-the process \( \Pend \) or \emph{inaction}, a process which can do nothing. The process \( \Ppar{P}{Q} \) is the \emph{parallel composition} of process \( P \) and process \( Q \).
-The two components can proceed independently of each other, or they can interact via shared names.
-%The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-
-For communication, processes include \emph{input} and \emph{output}, whose
-signature depends on the calculus being value-passing or name-passing.
-We use here the metavariables $c,k$ to abstract over this choice
---- i.e.~$c$ may be either a value or a variable or a name, whereas $k$ may
-be a variable or a name.
-The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
-\( c \) via \( x \), then continue as \( P \).  % The intention is that
-% the value \( v \) must be a base value when it is actually sent, and
-% this will be enforced in the semantics later on
-The process \( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
-via \( x \), then continue as \( P \) with the received element
-substituted for \( k \).  The input operator thus
-binds \( k \) in \( P \).
-
-The process \( \Pres{x}{P} \) is the \emph{restriction} of the name
-\( x \) to \( P \), binding \( x \) in \( P \).
-
-The process \( \Preplicate{P} \) is the \emph{replication} of the process \( P \).
-It can be thought of as the infinite composition \( \Ppar{P}{\Ppar{P}{\cdots}} \).
-Replication makes it possible to express infinite behaviours.
-
-We use the notation \( \freenames{P} \) to denote the set of
-names that occur free,
-% (i.e.\ not bound by a restriction) in \( P \),
-\( \boundnames{P} \) to denote the set of names that occur bound
-% (by a restriction)
-in \( P \) and  \( \freevars{P} \)
-to denote the set of variables that occur free
-% (i.e.\ not bound by an input)
-in \( P \).  We use the notation \( \boundvars{P} \)
-for the set of variables that occur bound
-% (by an input)
-in \( P \).  We use the notation \( \subst{P}{a}{l} \) to denote
-the process \( P \) with base value \( a \) substituted for variable
-\( l \). Similarly, \( \subst{P}{x}{y} \) denotes the process
-\( P \) with name \( x \) substituted for name \( y \).
-We use the notation \( \applysubst{\sigma}{P} \) to denote the process \( P \) with a finite number of arbitrary substitutions applied to it.
-
-Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
-written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \)
-by a finite number of substitutions of bound variables.  As a
-convention, we identify \( \alpha \)-convertible processes and we
-assume that bound names and bound variables of any processes are
-chosen to be different from the names and variables that occur
-free in any other entities under consideration, such as processes,
-substitutions, and sets of names or variables.  This is justified
-because any overlapping names and variables may be
-\( \alpha \)-converted such that the assumption is satisfied.
-
-A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-%
-A context acts as a function between processes:
-a context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-The replacement is literal, so names and variables that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsection{Challenge: Linearity and behavioral type systems}
-\label{sec:challenge:linearity-beh-types}
-This challenge formalises a proof that requires reasoning about linearity of channels.
-Linearity is the notion that a channel must be used exactly once in a process.
-This is necessary to prove properties about session type systems, and the key
-issue of this challenge is reasoning about the linearity of context splitting operations.
-Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
-
-The setting for this challenge is a small calculus with a session type
-system, the syntax and semantics of which are given below. The
-calculus is a fragment of the one presented in~\cite{Vasconcelos2012},
-formulated in the dual style of~\cite{barber96tr}.
-
-The main objective of this challenge is to prove type preservation (also
-known as subject reduction), i.e.\ that well-typed
-processes can only transition to processes which are also well-typed
-in the same context.
-The second objective is to prove type safety, i.e.\ that well-typed processes are also well-formed in the sense that they do not use endpoints in a non-dual way.
-
-\subsubsection{Syntax.}
-The syntax is given by the grammar
-\[
-\begin{array}{r@{\qquad}c@{\qquad}l}
-  v,w & \Coloneqq & a \quad\mid\quad l \\
-   P,Q & \Coloneqq & \Pend \quad\mid\quad \Pout{x}{v}{P} \quad\mid\quad \Pin{x}{l}{P} \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad  \Presd{x}{y}{P}
-\end{array}
-\]
-where a \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable \( l \).
-
-The output process \( \Pout{x}{v}{P} \) sends the value \( v \) via \( x \) and then continues as \( P \).
-The intention is that the value \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) waits for a base value from \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
-%
-The process \( \Presd{x}{y}{P} \) represents a \emph{session} with endpoints named \( x \) and \( y \) which are bound in \( P \).
-In \( P \), the names \( x \) and \( y \) can be used to exchange messages over the session (sending on \( x \) and receiving on \( y \) or vice versa).
-%
-%Finally, the process $\PBpar{P}{Q}$ represents the parallel composition of $P$ and $Q$.
-%
-Note that in this calculus channels cannot be sent in messages, therefore the topology of the communication network described by a process cannot change.
-Also, there is no recursion or replication in the syntax, hence no infinite behaviors can be expressed. In particular, we only model linear (as opposed to shared) channels.
-
-\subsubsection{Semantics.}
-We describe the actions that the system can perform through a small step operational semantics.
-As usual, we use  a \emph{structural congruence} relation that equates processes that we deem to be indistinguishable.
-Structural congruence is the smallest congruence relation that satisfies the following axioms:
-\begin{mathpar}
-  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
-  \and
-  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{(\Ppar{P}{Q})}{R}}{\Ppar{P}{(\Ppar{Q}{R})}}}
-  \and
-  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
-  \and
-  \inferrule[Sc-Res-Par]{\{x,y\} \cap \freenames{Q} = \emptyset}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
-  \and
-  \inferrule[Sc-Res-Inact]{ }{\scong{\Presd{x}{y}{\Pend}}{\Pend}}
-  \and
-  \inferrule[Sc-Res]{ }{\scong{\Presd{x_1}{y_1}{\Presd{x_2}{y_2}{P}}}{\Presd{x_2}{y_2}{\Presd{x_1}{y_1}{P}}}}
-\end{mathpar}
-
-The operational semantics is defined as the following relation on processes:
-\begin{mathparpagebreakable}
-  \inferrule[R-Com]{ }{\reduces{\Presd{x}{y}{(\Ppar{\Pout{x}{a}{P}}{\Ppar{\Pin{y}{l}{Q}}{R}})}}{\Presd{x}{y}{(\Ppar{P}{\Ppar{\subst{Q}{a}{l}}{R}})}}}
-  \and
-  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Presd{x}{y}{P}}{\Presd{x}{y}{Q}}}
-  \and
-  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
-  \and
-  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathparpagebreakable}
-
-Note that reductions are allowed only for restricted pairs of session endpoints.
-This makes it possible to formulate subject reduction so that the typing context is exactly the same before and after each reduction.
-%
-Note also that due to rule \TirName{R-Com}, the process \( \Pin{y}{l}{P} \) can receive \emph{any} base value.
-Since the rule \TirName{R-Com} only applies to sending base values, there is no way to send a variable or a name.
-
-\subsubsection{Session types.}
-Our process syntax allows us to write processes that are ill formed in
-the sense that they either use the endpoints bound by a restriction to
-communicate in a way that does not follow the intended duality, or
-attempt to send something which is not a base value.  As an example,
-the process
-\( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \)
-attempts to send a base value on both \( x \) and \( y\), whereas one
-of the names should be used for receiving in order to guarantee
-progress.  Another example is the process
-\( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \),
-which attempts to send a variable that is not instantiated at the time
-of sending.
-%
-To prevent these issues, we introduce a \emph{session type system} which
-rules out ill-formed processes.
-
-\subsubsection{Syntax.}
-Our type system does not type processes directly, but instead focuses on the channels used in the process.
-The syntax of \emph{session types} \( S, T \), unrestricted type contexts \( \Gamma \) and linear type contexts \( \Delta \) is as follows:
-\[
-  \begin{array}{r@{\qquad}c@{\qquad}l}
-  S,T & \Coloneqq & \Tend \quad\mid\quad \Tbase \quad\mid\quad \Tin{S} \quad\mid\quad \Tout{S} \\
-    \Gamma & \Coloneqq & \Cempty \quad\mid\quad \Gamma, l \\
-                    \Delta &::= & \Cempty \quad\mid\quad \Cadd{\Delta}{\hastype{x}{S}}
-  \end{array}
-\]
-The \emph{end type} \( \Tend \) describes an endpoint on which no further interaction is possible.
-The \emph{base type} \( \Tbase \) describes base values.
-The \emph{input type} \( \Tin{S} \) describes endpoints used for receiving a value and then according to \( S \).
-The \emph{output type} \( \Tout{S} \) describes endpoints used for sending a value and then according to \( S \).
-
-Typing contexts gather type information about names and variables.
-\emph{Unrestricted} contexts are simply sets of names since we only have one
-base type. \emph{Linear contexts} associate a type to endpoints. We use
-the comma as split/union, overloaded to singletons, and \( \Cempty \) as the
-empty context. We extend the Barendregt convention to contexts, so that all
-entries are distinct.  Note that the order in which information is added to a
-type context does not matter.
-
-Since we need to determine whether endpoints are used in complementary ways to determine whether processes are well formed, we need to formally define the dual of a type as follows:
-\begin{mathpar}
-  \inferrule{}{\dual{\Tin{S}} = \Tout{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tout{S}} = \Tin{\dual{S}}}
-  \and
-  \inferrule{}{\dual{\Tend} = \Tend}
-\end{mathpar}
-Note that the dual function is partial since it is undefined for the base type.
-
-\subsubsection{Typing rules.}
-Our type system is aimed at maintaining two invariants:
-\begin{enumerate}
-\item No endpoint is used simultaneously by parallel processes;
-\item The two endpoints of the same session have dual types.
-\end{enumerate}
-The first invariant is maintained by linearly splitting type contexts when typing compositions of processes, the second by requiring duality when typing restrictions.
-
-We have two typing judgments: one for values, and one for processes.
-The typing rules for values are:
-\begin{mathpar}
-  \inferrule[T-Base]{ }{\typev{\Gamma}{\hastype{a}{\Tbase}}} \and
-  \inferrule[T-Var]{ }{\typev{\Cadd{\Gamma}l }{\hastype{l}{\Tbase}}}
-\end{mathpar}
-% In the typing rule for names, $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
-% \begin{mathpar}
-%   \inferrule[T-Name]{{\tend\Delta }}{\types{{\Gamma; (\Delta,\hastype{x}{T})}}{\hastype{x}{T}}}
-% \end{mathpar}
-The typing rules for processes are as follows:
-\begin{mathpar}
-
-  \inferrule[T-Inact]{\tend\Delta }{\types{\Gamma;\Delta}{\Pend}}
-  \and
-  \inferrule[T-Par]{\types{\Gamma;\Delta_1}{P} \\ \types{\Gamma;\Delta_2}{Q}}
-  {\types{\Gamma; \Csplit{\Delta_1}{\Delta_2}}{\Ppar{P}{Q}}}
-  \and
-  \inferrule[T-Res]{\types{\Gamma; (\Cadd{\Cadd{\Delta}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P})}{\types{\Gamma}{\Presd{x}{y}{P}}}
-  \and
-    \inferrule[T-Out]{
-      \typev{\Gamma}{\hastype{v}{\Tbase}} \\ \types{\Gamma; \Cupdate{\Delta}{\hastype{x}{T}}}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tout{T}}})}{\Pout{x}{v}{P}}}
-    \and
-    \inferrule[T-IN]{
-      \types{(\Gamma,  l ); (\Cupdate{\Delta}{\hastype{x}{T}})}{P}}{\types{\Gamma; (\Csplit{\Delta}{\hastype{x}{\Tin{T}}})}{\Pin{x}{l}{P}}}
-\end{mathpar}
-Note that we do not need a judgment for typing channels, since it is already folded into the \TirName{T-In} and \TirName{T-Out} rules.
-
-\subsubsection{Challenge.}
-The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with:
-\begin{theorem}[Subject reduction]
-  \label{thm:subject-reduction}
-  If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
-\end{theorem}
-
-The following lemmas are needed:
-\begin{lemma}[Weakening]\mbox{}
-  \label{lemma:weak}
-  \begin{enumerate}
-  \item If \( \types{\Gamma; \Delta}{P} \) then
-    \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
-  \item If \( \types{\Gamma; \Delta}{P} \) % and \( \tend{\Delta} \)
-    then \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}[Strengthening]\mbox{}
-  \label{lemma:strenD}
-  If \( \types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P} \) and
-  $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
-\end{lemma}
-
-\begin{lemma}[Substitution]\mbox{}
-  \label{le:subst}
-  If $\types{(\Cadd{\Gamma}{l});\Delta}{P}$ and
-  ${\typev{\Gamma}{\hastype{a}{\Tbase}}}$ then
-  \( \types{\Gamma;\Delta}{\subst{P}{a}{l}} \).
-\end{lemma}
-
-To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms.
-
-\begin{lemma}[Preservation for $\stackrel{a}{\equiv}$]
-  \label{le:presequiva}
-  If \( \sconga{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
-\end{lemma}
-
-Now we give rules for the compatible
-equivalence relation induced by $\sconga\cdot\cdot$, which we still
-write as $\scong\cdot\cdot$: namely, add to reflexivity, symmetry and
-transitivity the following condition:
- \begin{mathpar}
-   \inferrule[Cong]{\sconga P Q}{\scong {C[P]}{C[Q]}}
- \end{mathpar}
-
-\begin{lemma}[Preservation for $\equiv$]
-  \label{le:presequiv}
-    If \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
-\end{lemma}
-
-To formulate safety, we need to formally define what we mean by
-well-formed process.  We say that a process \( P \) is \emph{prefixed
-  at variable \( x \)} if \( P\equiv\Pout{x}{v}{P'} \) or
-\( P\equiv\Pin{x}{l}{P'} \) for some $P'$.
-%
-A process $P$ is then \emph{well formed} if, for every $P_1$, $P_2$,
-and $R$ such that
-\( P\equiv \Presd{x_1}{y_1}{\dots
-  \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}} \), with
-\( n \geq 0 \), it holds that, if \( P_1 \) is prefixed at \( x_1 \)
-and \( P_2 \) is prefixed at \( y_1 \) (or vice versa), then
-\( \Ppar{P_1}{P_2}
-\equiv\Ppar{\Pout{x_1}{a}{P_1'}}{\Pin{y_1}{l}{P_2'}} \), for some
-$P_1'$ and $P_2'$.
-
-Note that well-formed processes do not necessarily reduce. For example, the process
-\begin{equation*}
-  \Presd{x_1}{y_1}{\Presd{x_2}{y_2}{(\Ppar{\Pout{x_1}{a}{\Pin{y_2}{l}{\Pend}}}{\Pout{y_2}{x_2}{\Pin{y_1}{l}{\Pend}}})}}
-\end{equation*}
-is well formed but also irreducible.
-
-\begin{theorem}[Type safety]
-  \label{thm:type-safety}
-  If \( \types{\Cempty}{P} \), then \( P \) is well formed.
-\end{theorem}
-
-\begin{corollary}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well formed.
-\end{corollary}
-
-\subsection{Challenge: Name passing and scope extrusion}
-\label{sec:challenge:name-passing-scope-extrusion}
-
-This challenge formalises a proof that requires explicit scope extrusion.
-Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
-This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
-The key issue of this challenge is reasoning about names that are ``in the process'' of being scope-extruded, which often presents difficulties for the mechanisation of binders.
-
-Reasoning about scope extrusion explicitly can sometimes be avoided by introducing a structural congruence rule into the semantics, but doing this means we lose information about the scope when reasoning about the semantics.
-Explicitly reasoning about scope extrusion is necessary to describe e.g.\ runtime monitors and compositions of systems.
-
-The setting for this challenge is a ``classic'' untyped \( \pi \)-calculus, where (unlike the calculi in the other challenges) names can be sent and received, and bound by input constructs (similarly to variables in the other calculi).
-We define two different semantics for our system: one that avoids explicit reasoning about scope extrusion, and one that does not.
-The objective of this challenge is to prove that the two semantics are equivalent up to structural congruence.
-
-\subsubsection{Syntax.}
-The syntax of processes is given by:
-\[
-\begin{array}{r@{\qquad}c@{\qquad}l}
-  P,Q & \Coloneqq & \Pend \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad \Pout{x}{y}{P} \quad\mid\quad \Pin{x}{y}{P} \quad\mid\quad \Pres{x}{P}
-\end{array}
-\]
-
-%The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
-The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
-The process \( \Pin{x}{y}{P} \) is an \emph{input}, which can receive a name via \( x \), then continue as \( P \) with the received name substituted for \( y \).
-The input operator thus binds the name \( y \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-% The restriction operator thus binds the name \( x \) in \( P \).
-Note that the scope of a restriction may change when processes interact. Namely, a restricted name may be sent \emph{outside} of its scope.
-% The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-Note that there is no recursion or replication in the syntax, and thus no infinite behaviors can be expressed.
-This simplifies the theory and is orthogonal to the concept of scope extrusion.
-
-% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
-% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
-% We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
-
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
-% This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Reduction semantics.}
-The first semantics is an operational reduction semantics, which avoids reasoning explicitly about scope extrusion by way of a structural congruence rule.
-\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:
-
-\begin{mathpar}
-  \inferrule[Sc-Par-Assoc]{ }{\scong{\Ppar{P}{(\Ppar{Q}{R})}}{\Ppar{(\Ppar{P}{Q})}{R}}}
-  \and
-  \inferrule[Sc-Par-Comm]{ }{\scong{\Ppar{P}{Q}}{\Ppar{Q}{P}}}
-  \and
-  \inferrule[Sc-Par-Inact]{ }{\scong{\Ppar{P}{\Pend}}{P}}
-  \and
-  \inferrule[Sc-Res-Par]{x \notin \freenames{Q}}{\scong{\Ppar{\Pres{x}{P}}{Q}}{\Pres{x}{(\Ppar{P}{Q})}}}
-  \and
-  \inferrule[Sc-Res-Inact]{ }{\scong{\Pres{x}{\Pend}}{\Pend}}
-  \and
-  \inferrule[Sc-Res]{ }{\scong{\Pres{x}{\Pres{y}{P}}}{\Pres{y}{\Pres{x}{P}}}}
-\end{mathpar}
-
-The operational semantics is defined as the following relation on processes:
-\begin{mathpar}
-  \inferrule[R-Com]{ }{\reduces{\Ppar{\Pout{x}{y}{P}}{\Pin{x}{z}{Q}}}{\Ppar{P}{\subst{Q}{y}{z}}}}
-  \and
-  \inferrule[R-Res]{\reduces{P}{Q}}{\reduces{\Pres{x}{P}}{\Pres{x}{Q}}}
-  \and
-  \inferrule[R-Par]{\reduces{P}{Q}}{\reduces{\Ppar{P}{R}}{\Ppar{Q}{R}}}
-  \and
-  \inferrule[R-Struct]{\scong{P}{P'} \\ \reduces{P'}{Q'} \\ \scong{Q}{Q'}}{\reduces{P}{Q}}
-\end{mathpar}
-Note that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{R-Com}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
-Finally, note that rule \TirName{R-Struct} allows for applying the structural congruence both before and after the reduction:
-this makes the reduction relation closed under structural congruence.
-
-\subsubsection{Transition system semantics.}
-The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
-The transitions are labelled by \emph{actions}, the syntax of which is:
-\[
-\begin{array}{r@{\qquad}c@{\qquad}l}
-  \alpha & \Coloneqq & \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
-\end{array}
-\]
-
-The \emph{free output action} \( \Aoutf{x}{y} \) is sending the name \( y \) via \( x \).
-The \emph{input action} \( \Ain{x}{y} \) is receiving the name \( y \) via \( x \).
-The \emph{bound output action} \( \Aoutb{x}{y} \) is sending a fresh name \( y \) via \( x \).
-The \emph{internal action} \( \Atau \) is performing internal communication.
-
-We extend the notion of free and bound occurrences with
-\( \freenames{\alpha} \) to denote the set of names that occur free in
-the action \( \alpha \) and  \( \boundnames{\alpha} \) to
-denote the set of names that occur bound in the action \( \alpha \).
-In the free output action \( \Aoutf{x}{y} \) and the input action
-\( \Ain{x}{y} \), both \( x \) and \( y \) are free names.  In the
-bound output action \( \Aoutb{x}{y} \), \( x \) is a free name, while
-\( y \) is a bound name.  We also use the notation
-\( \names{\alpha} \) to denote the union of \( \freenames{\alpha} \)
-and \( \boundnames{\alpha} \), i.e.\ the set of all names that occur
-in the action \( \alpha \).
-The transition relation is then defined by the following rules:
-\begin{mathparpagebreakable}
-  \inferrule[Out]{ }{\transition{\Pout{x}{y}{P}}{\Aoutf{x}{y}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{z}{P}}{\Ain{x}{y}}{\subst{P}{y}{z}}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'} \\ \boundnames{\alpha} \cap \freenames{Q} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'} \\ \boundnames{\alpha} \cap \freenames{P} = \emptyset}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{y}}{P'} \\ \transition{Q}{\Ain{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{y}}{P'} \\ \transition{Q}{\Aoutf{x}{y}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Close-L]{\transition{P}{\Aoutb{x}{z}}{P'} \\ \transition{Q}{\Ain{x}{z}}{Q'} \\ z \notin \freenames{Q}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Open]{\transition{P}{\Aoutf{x}{z}}{P'} \\ z \neq x}{\transition{\Pres{z}{P}}{\Aoutb{x}{z}}{P'}}
-  \and
-  \inferrule[Close-R]{\transition{P}{\Ain{x}{z}}{P'} \\ \transition{Q}{\Aoutb{x}{z}}{Q'} \\ z \notin \freenames{P}}{\transition{\Ppar{P}{Q}}{\tau}{\Pres{z}{\Ppar{P'}{Q'}}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ z \notin \names{\alpha}}{\transition{\Pres{z}{P}}{\alpha}{\Pres{z}{P'}}}
-\end{mathparpagebreakable}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{z}{P} \) can receive \emph{any} name.
-
-We keep the convention that bound names of any processes or actions
-are chosen to be different from the names that occur free in any other
-entities under consideration, such as processes, actions,
-substitutions, and sets of names.  The convention has one exception,
-namely that in the transition \( \transition{P}{\Aoutb{x}{z}}{Q} \),
-the name \( z \) (which occurs bound in \( P \) and the action
-\( \Aoutb{x}{z} \)) may occur free in \( Q \).  Without this exception
-it would be impossible to express scope extrusion.
-
-Before getting to our challenge, we prove some intermediate results.
-\begin{lemma}\label{se-lemma-harmony-fact}
-  If \( \scong{P}{Q} \) and \( \transition{P}{\alpha}{P'} \), then for some \( Q' \) we have \( \transition{Q}{\alpha}{Q'} \) and \( \scong{P'}{Q'} \).
-\end{lemma}
-
-For the proof of the challenge theorem, we introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.
-The first rule applied is \TirName{R-Com}. The derivation continues with an application of \TirName{R-Par}, followed by zero or more applications of \TirName{R-Res}.
-The last rule is an application of \TirName{R-Struct}.
-\begin{lemma}\label{se-lemma-normalized-derivation}
-  Every reduction has a normalized derivation.
-\end{lemma}
-
-\begin{lemma}\label{se-lemma-normalized-reduction}
-  If \( \reduces{P}{Q} \) then there are \( x, y, z, z_1, \dots, z_n, R_1, R_2, \) and \( S\) such that
-  \begin{align*}
-    &\scong{P}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{\Pout{x}{y}{R_1}}{\Pin{x}{z}{R_2}})}{S})}}} \\
-    &\scong{Q}{\Pres{z_1}{\!\dots \Pres{z_n}{(\Ppar{(\Ppar{R_1}{\subst{R_2}{y}{z}})}{S})}}}
-  \end{align*}
-\end{lemma}
-
-\subsubsection{Challenge.}
-The objective of this challenge is to prove the following theorems, which together show the equivalence between the reduction semantics and the transition system semantics up to structural congruence.
-The first of the theorems involves reasoning about scope extrusion more directly than the other, and if time does not permit proving both of the theorems, \cref{thm:se-trans-implies-red} should be proven first.
-\begin{theorem}\label{thm:se-trans-implies-red}
-  \( \transition{P}{\Atau}{Q} \) implies \( \reduces{P}{Q} \).
-\end{theorem}
-
-\begin{theorem}\label{thm:harmony-2}
-  \( \reduces{P}{Q} \) implies the existence of a \( Q' \) such that \( \transition{P}{\Atau}{Q'} \) and \( \scong{Q}{Q'} \).
-\end{theorem}
-
-\subsection{Challenge: Coinduction and infinite processes}
-\label{sec:challenge:coinduction}
-This challenge is about the mechanisation of proofs concerning processes with infinite behaviours.
-% formalises a proof
-% that require coinduction.
-% coinductive techniques.
-%
-% Coinduction is a proof technique for infinite structures,
-This is usually connected to \emph{coinductive} definitions where an infinite
-structure is defined as the greatest fixed point of a recursive
-definition.
-%
-Coinduction is a technique for proving properties of
-such infinite structures.
-%
-%Coinduction is the dual of induction: whereas induction is useful for
-%proving properties of least fixed points, coinduction is useful for
-%proving properties of greatest fixed points.
-%
-% \marco{Should we be more specific here and give a deeper explanation
-%   of what coinduction is? If not, then why do we even say that it's
-%   the dual of induction?}
-%
-
-For this challenge, we adopt a fragment of the untyped $\pi$-calculus
-that includes process replication.
-% The setting for this challenge is an untyped calculus of
-% communicating systems with replication of processes, the syntax and
-% semantics of which are presented below.
-The objective of this challenge is to % prove % a context theorem for
-draw a formal connection between strong barbed congruence and strong
-barbed bisimilarity.  % I.e., The context lemma
-The result establishes that two processes are strong barbed congruent
-if the processes obtained by applying a finite number of substitutions to
-them and composing them in parallel with an arbitrary process are
-strongly barbed bisimilar.
-The key issue of this challenge is the coinductive reasoning about the infinite behaviours of the replication operator.
-
-\subsubsection{Syntax.}
-% We assume the existence of some type of \emph{base values}, values
-% of which we will denote by \( a, b, \dots \), the existence of some
-% type of \emph{variables}, values of which we will denote by
-% \( l, m, \dots \), and the existence of some type of \emph{names},
-% values of which we will denote by \( x, y, \dots \).
-The syntax of values and processes is given by:
-% for including replication and having unary restriction:
-\begin{displaymath}
-  \begin{array}{r@{\qquad}c@{\qquad}l}
-    v,w & \Coloneqq & a \quad\mid\quad l \\
-    P,Q & \Coloneqq & \Pend
-               \quad \mid\quad \Pout{x}{v}{P}
-               \quad \mid\quad \Pin{x}{l}{P}
-               \quad \mid\quad \PBpar{P}{Q}
-               \quad \mid\quad \Pres{x}{}{P}
-               \quad \mid\quad !P
-  \end{array}
-\end{displaymath}
-The output process \( \Pout{x}{v}{P} \) sends the value \( v \) on channel \( x \) and continues as \( P \).
-The intention is that \( v \) must be a base value when it is actually sent, and this is enforced in the semantics later on.
-The input process \( \Pin{x}{l}{P} \) waits for a base value from channel \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
-Since replication allows for infinite copies of the process $P$, processes can dynamically
-create an infinite number of names during execution.
-
-%
-% \begin{align*}
-%   v,w :=&&& a \\
-%   |&&& l \\
-%   P,Q :=&&& \Pend \\
-%   |&&& \Pout{x}{v}{P} \\
-%   |&&& \Pin{x}{l}{P} \\
-%   |&&& \Ppar{P}{Q} \\
-%   |&&& \Pres{x}{P} \\
-%   |&&& \Preplicate{P}
-% \end{align*}
-% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a
-% variable \( l \).  The process \( \Pend \) is \emph{inaction}:
-% a process which can do nothing.  The process \( \Pout{x}{a}{P} \) is
-% an \emph{output}, which can send the value \( v \) via \( x \), then
-% continue as \( P \).  The intention is that the value \( v \) must
-% be a base value when it is actually sent, and this will be enforced
-% in the semantics later on.  The process \( \Pin{x}{l}{P} \) is an
-% \emph{input}, which can receive a base value via \( x \), then
-% continue as \( P \) with the received value substituted for the
-% variable \( l \).  The input operator thus binds the variable
-%  \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
-% \emph{composition} of process \( P \) and process \( Q \).  The two
-% components can proceed independently of each other, or they can
-% interact via shared names.  The process \( \Pres{x}{P} \) is the
-% \emph{restriction} of the name \( x \) to \( P \).  Components in
-% \( P \) can use the name \( x \) to interact with each other, but
-% not with processes outside of the restriction.  The restriction
-% operator thus binds the name \( x \) in \( P \).  Note that the
-% scope of a restriction may not change when processes interact, since
-% it is only possible to send and receive values, and not names.  The
-% process \( \Preplicate{P} \) is the \emph{replication} of the
-% process \( P \).  It can be thought of as the infinite composition
-% \( \Ppar{P}{\Ppar{P}{\cdots}} \).  Replication makes it possible to
-% express infinite behaviours.
-
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process
-% \( P \) with value \( v \) substituted for variable \( l \).
-
-% % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
-% written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from
-% \( P \) by a finite number of substitutions of bound variables.
-% As a convention, we will identify \( \alpha \)-convertible
-% processes.
-
-% Also as a convention, we assume that the bound names and bound
-% variables occurring in any collection of processes are chosen
-% to be different from the free names and free variables
-% occurring in those processes and from the names and variables
-% occurring in any substitutions applied to the processes.  This is
-% justified because any overlapping names and variables may be
-% \( \alpha \)-converted such that the assumption is satisfied.
-
-\subsubsection{Semantics.} We choose to give an LTS semantics for this
-challenge. %As an alternative, it is also possible to adopt a reduction
-%semantics which features fewer rules but needs a structural
-%congruence. %  The semantics of the system describes the actions that the
-% system can perform by defining a labelled transition relation on
-% processes.
-The transitions are labelled by \emph{actions}, the syntax
-of which is as follows:
-% \begin{align*}
-%   \alpha := &&& \Aoutf{x}{a} \\
-%   |&&& \Ain{x}{a} \\
-%   |&&& \Atau
-% \end{align*}
-\begin{align*}
-  \alpha\ \ \Coloneqq\ \ \Aoutf{x}{a} \ \mid\ \Ain{x}{a} \ \mid\ \Atau
-\end{align*}
-The \emph{output action} \( \Aoutf{x}{y} \) is sending the base value
-\( a \) via \( x \).  The \emph{input action} \( \Ain{x}{y} \) is
-receiving the base value \( y \) via \( x \).  The \emph{internal
-  action} \( \Atau \) is performing internal communication.
-%
-We use the notation \( \names{\alpha} \) to denote the set of
-names that occur in the action \( \alpha \).
-The transition relation is defined by the following rules:
-\begin{mathparpagebreakable}
-  \inferrule[Out]{ }{\transition{\Pout{x}{a}{P}}{\Aoutf{x}{a}}{P}}
-  \and
-  \inferrule[In]{ }{\transition{\Pin{x}{l}{P}}{\Ain{x}{a}}{\subst{P}{a}{l}}}
-  \and
-  \inferrule[Par-L]{\transition{P}{\alpha}{P'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P'}{Q}}}
-  \and
-  \inferrule[Par-R]{\transition{Q}{\alpha}{Q'}}{\transition{\Ppar{P}{Q}}{\alpha}{\Ppar{P}{Q'}}}
-  \and
-  \inferrule[Comm-L]{\transition{P}{\Aoutf{x}{a}}{P'} \\ \transition{Q}{\Ain{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Comm-R]{\transition{P}{\Ain{x}{a}}{P'} \\ \transition{Q}{\Aoutf{x}{a}}{Q'}}{\transition{\Ppar{P}{Q}}{\Atau}{\Ppar{P'}{Q'}}}
-  \and
-  \inferrule[Res]{\transition{P}{\alpha}{P'} \\ x \notin \names{\alpha}}{\transition{\Pres{x}{P}}{\alpha}{\Pres{x}{P'}}}
-  \and
-  \inferrule[Rep]{\transition{P}{\alpha}{P'}}{\transition{\Preplicate{P}}{\alpha}{\Ppar{P'}{\Preplicate{P}}}}
-\end{mathparpagebreakable}
-Note that there is no rule for inferring transitions from \( \Pend \), and that there is no rule for inferring an action of an input or output process except those that match the input/output capability.
-Note also that due to rule \TirName{In}, the process \( \Pin{x}{l}{P} \) can receive \emph{any} base value.
-On the other hand, since the rule \TirName{Out} only applies to base values, there is no way to send a variable.
-
-% As a convention, we assume that bound names and bound variables of any processes or actions are chosen to be different from the names and variables that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variables.
-
-\subsubsection{Bisimilarity.}
-Bisimilarity is a notion of equivalence for processes and
-% .  Our notion of process equivalence relations
-builds on the idea of \emph{observables}, i.e., what we can externally
-observe from the semantics of a process. If we allowed ourselves only
-to observe internal transitions (i.e., observe that a process is
-internally performing a step of computation) we would relate either
-too few processes (in the strong case where we relate only processes
-with exactly the same number of internal transitions) or every process
-(in the weak case where we relate processes with any amount of
-internal transitions).  As a result, we must allow ourselves to
-monitor more than just internal transitions, and we choose to describe
-a process's observables as the names it might use for sending and
-receiving.
-% We must therefore allow ourselves to observe more than
-% just internal transitions, and here we choose to define the observables
-% of a process as the names it can use for sending and receiving.
-To this end, we define the \emph{observability predicate}
-\( \observable{P}{\mu} \) as follows:
-\begin{align*}
-  \observable{P}{\obsin{x}}  &\quad \textrm{if \( P \) can perform an input action via \( x \).} \\
-  \observable{P}{\obsout{x}} &\quad \textrm{if \( P \) can perform an output action via \( x \).}
-\end{align*}
-
-\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
-\begin{gather}
-  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \label{eq:bisim1}\\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'} \label{eq:bisim2}
-\end{gather}
-We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
-
-\begin{theorem}\label{thm:bisim-equiv}
-  \( \sbbisim{}{} \) is an equivalence relation. %, that is, the relation is reflexive, symmetric, and transitive.
-\end{theorem}
-
-Unfortunately, strong barbed bisimilarity is not a good process
-equivalence since it is not a congruence, hence it does not allow for
-substituting a process with an equivalent one in any context.
-% equivalent processes to be since it does not consider the
-% environment of processes.
-For instance, the processes $\Pout{x}{a}{\Pout{y}{b}{\Pend}}$ and
-$\Pout{x}{a}{\Pend}$ are strong barbed bisimilar, i.e.,
-$\sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}}$. This
-is because \( \obsout{x} \) is the only observable in both processes
-and they cannot perform a \( \Atau \)-action. However, in the context
-$C = \Ppar{\ctxhole}{\Pin{x}{l}{\Pend}}$, the relation no
-longer holds: in fact,
-$\nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}\Ppar{}{\Pin{x}{l}{\Pend}}}$
-%
-because the left process can perform a \( \Atau \)-action such that
-\( \obsout{y} \) becomes observable, whereas the right process cannot.
-
-% For instance,
-% \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \)
-% since \( \obsout{x} \) is the only observable in both processes and
-% they cannot perform a \( \Atau \)-action, but
-% \(
-% \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}}
-% \) since the left process can perform a \( \Atau \)-action such that
-% \( \obsout{y} \) becomes observable, whereas the right process cannot.
-
-% \subsubsection{Contexts and congruences}
-% Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
-
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We can think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variables that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
-\subsubsection{Strong barbed congruence.}
-In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,
-i.e., % we need to make strong barbed bisimilarity consider
-we have to consider the environment in which processes may be placed.
-% (i.e.\ the context) of processes.
-
-% We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying
-We say that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).
-
-\begin{lemma}\label{lemma:sbcong-largest}
-  \( \sbcong{}{} \) is the largest congruence included in
-  \( \sbbisim{}{} \).
-\end{lemma}
-
-\subsubsection{Challenge.}
-The objective of this challenge is to prove a theorem that shows that
-making strong barbed bisimilarity sensitive to substitution and
-parallel composition is enough to show strong barbed
-congruence.
-To prove the theorem, we will use an \emph{up-to technique}, utilizing the following definition and lemma.
-A relation \( \mathcal{S} \) is called a \emph{strong barbed bisimulation up to \( \sbbisim{}{} \)} if, whenever \( (P,Q) \in \mathcal{S} \), the following conditions hold:
-\begin{enumerate}
-\item \( \observable{P}{\mu} \) if and only if \( \observable{Q}{\mu} \).
-\item if \( \transition{P}{\Atau}{P'} \) then \( \transition{Q}{\Atau}{Q'} \) for some \( Q' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
-\item if \( \transition{Q}{\Atau}{Q'} \) then \( \transition{P}{\Atau}{P'} \) for some \( P' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
-\end{enumerate}
-\begin{lemma}\label{lemma:up-to}
-  If \( \mathcal{S} \) is a strong barbed bisimulation up to \( \sbbisim{}{} \), \( (P,Q) \in \mathcal{S} \) implies \( \sbbisim{P}{Q} \).
-\end{lemma}
-
-\begin{theorem}\label{thm:coinduction}
-  \( \sbcong{P}{Q} \) if for any process \( R \) and
-  substitution \( \sigma \),
-  \(
-  \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}}
-  \).
-\end{theorem}
+The actual challenges are available in \cref{app:challenges}.
 
@@
 \appendix
-\section{Proofs}\label{app:challenges}
-\input{../proof-appendix/proofs.tex}
+\section{Challenges}\label{app:challenges}
+\input{../challenges/challenges.tex}
 
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2024-01-21 short-paper/short-paper.tex
Exactly 6 pages with motivation heading.
Copy-editing

@@
 \usepackage{hyperref}
+\usepackage{microtype}
 \usepackage[inline]{enumitem}
@@
 \and University of Oxford, Oxford, United Kingdom \email{martin.vassor@cs.ox.ac.uk}, \email{nobuko.yoshida@cs.ox.ac.uk}
-\and McGill University, Montreal, Canada \email{daniel.zackon@mail.mcgill.ca}
+\and McGill University, Montreal, Canada \email{daniel.zackon@mcgill.ca}
 }
@@
 \section{Introduction}
-The abstract of the influential POPLMark challenge~\cite{POPLMark} starts with
-an important question: \emph{``How close are we to a world where every paper on
-programming languages is accompanied by an electronic appendix with
-machine-checked proofs?''} The introduction of the POPLMark
-challenge spearheaded a shift towards publications that include
-mechanised proofs, fostering the progress of proof assistants,
-libraries, and best practices.  Later on, the follow-up POPLMark
-Reloaded challenge~\cite{POPLMarkReloaded} has encouraged the development of
-best practices and tool support for proofs using logical relations.
+By establishing a common ground for systematically evaluating mechanisation techniques, 
+the POPLMark challenge~\cite{POPLMark} accelerated progress in 
+the development of proof assistants, libraries, and best practices. In doing so, it 
+spearheaded a shift in the programming lanuage research community towards 
+publications that include mechanised proofs. 
+Its follow-up challenge, POPLMark Reloaded~\cite{POPLMarkReloaded}, 
+facilitated the development of best practices and tool support for proofs 
+using logical relations. These challenges were deliberately limited in scope, 
+while their authors called for future investigations into such areas as 
+coinduction and linear environments. In this spirit, we propose a 
+new collection of benchmark problems specifically designed to address typical 
+issues that arise when mechanising message-passing concurrency using process 
+calculi.
 
-The authors of both POPLMark and POPLMark Reloaded explicitly
-note that their benchmarks were only a beginning.  Furthermore, they
-specifically mention reasoning about concurrency using coinduction and
-linear environments as points for future work.  In this spirit, in order to
-further increase the reach of the POPLMark challenges, we propose a new
-collection of benchmark problems specifically designed to address typical issues
-that arise when mechanising message-passing concurrency using process calculi.
+Several papers in this realm already include mechanised proofs
+(e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
+  DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or discuss proof mechanisation techniques
+(e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
+  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).  Our
+experience, however, is that choosing the appropriate proof mechanisation techniques and 
+tools remains a significant challenge with a limited understanding of their 
+respective tradeoffs.  This often leads researchers toward a trial-and-error approach, 
+resulting in sub-optimal and ad-hoc solutions. In turn, this increases the overall effort 
+required for proof mechanisation, and hampers the reuse of techniques and results.
+%
+It is therefore our view that the research community in message-passing
+concurrency (and process calculi, more specifically) will benefit from a standardised set
+of challenge probems.
 
-Several papers in the realm of process calculi already include mechanised proofs
-(e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
-  DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or
-discuss proof mechanisation techniques
-(e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
-  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).  However, our
-experience is that choosing the appropriate proof mechanisation techniques and
-tools remains a significant challenge with few guidelines.  This often leads
-researchers towards trial-and-error, attempting sub-optimal and ad-hoc
-solutions; this increases the overall effort required for proof mechanisation,
-and hampers the reuse of techniques and results.
-%
-Therefore, it is our opinion that the research community in message-passing
-concurrency (and more specifically, in process calculi) will benefit from a set
-of solutions exploring different ways of mechanising typical problems in the
-field.
+To this end, our benchmark problems (discussed in detail in \cref{app:challenges}) are designed to focus on three key aspects which typically cause difficulties 
+when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, 
+and \emph{coinductive reasoning} for infinite behaviours. These are certainly not the only factors that may cause issues, but they
+are fundamental to concurrency theory, and emerge in most mechanisations.
+The mechanisation of novel research results often requires addressing more than
+just one of these aspects at once; we see the combination of techniques
+as a next step beyond the scope of this challenge (we discuss this in
+\cref{sec:going-beyond}).
+% Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate, evaluate, and compare the many tools and techniques available.
+% While the challenge problems themselves are thus not of scientific interest, we expect that the best practices and mechanisations developed by solving the challenge problems will also be useful for future research.
+
 Like the authors of the POPLMark challenges, we seek to
-answer a number of questions:
+answer several questions:
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
@@
 \end{enumerate}
-We have identified three key aspects which typically cause difficulties when
-mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion},
-and \emph{coinductive reasoning} for infinite behaviours.
-Our benchmark problems are designed to focus on these aspects (discussed
-in detail in \cref{app:challenges}), with three main goals:
+
+On the whole, this benchmark seeks to advance three main goals:
 
 \begin{enumerate}[label=\textbf{(G\arabic*)},leftmargin=10mm]
-\item\label{item:goal-comperison-accessibility} Enable the comparison of
-  different proof mechanisation approaches, making the challenges accessible to
-  mechanisation experts who may not be familiar with concurrency theory.
+\item\label{item:goal-comperison-accessibility} To enable the comparison of
+  proof mechanisation approaches, making the challenges accessible to
+  mechanisation experts who may be unfamiliar with concurrency theory;
 
-\item\label{item:goal-tutorials} Encourage the development of guidelines and
-  tutorials demonstrating and comparing the many available proof mechanisation
-  techniques, libraries, and proof assistant features.
+\item\label{item:goal-tutorials} To encourage the development of guidelines and
+  tutorials demonstrating and comparing existing proof mechanisation
+  techniques, libraries, and proof assistant features; and
 
-\item\label{item:goal-reusability} Prioritise the exploration of mechanisation
-  techniques that may be useful and reusable for future research.
+\item\label{item:goal-reusability} To prioritise the exploration of mechanisation
+  techniques that are reusable for future research.
 \end{enumerate}
@@
 community to collaborate on exploring and developing new tools and techniques.
-
-The three aspects we selected (linearity, scope extrusion, coinduction) are of
-course not the only ones that may cause issues in mechanisations --- but they
-are fundamental to concurrency theory, and they emerge in most mechanisations.
-The mechanisation of novel research results often requires addressing more than
-just one of these aspects at once --- and we see the combination of techniques
-as a next step beyond the scope of this challenge (we discuss this in
-\cref{sec:going-beyond}).
-
-% Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate, evaluate, and compare the many tools and techniques available.
-% While the challenge problems themselves are thus not of scientific interest, we expect that the best practices and mechanisations developed by solving the challenge problems will also be useful for future research.
 
@@
 We discuss our design choices in \cref{sec:design-discussion}, and possible extensions and further topics in \cref{sec:going-beyond}.
-The actual challenges are available in \cref{app:challenges}.
+The challenges themselves may be found in \cref{app:challenges}.
+
+\section{Motivation}\label{sec:motivation}
 
@@
 In this section, we outline the factors considered in designing the
-benchmark challenges. We start with some general remarks, then
+benchmark challenges. We begin with some general remarks, then
 describe the individual design considerations for each challenge problem,
@@
 applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.
-Still, these (and other) applications need the basic techniques that our challenge problems exercise:
-as per our design goal \ref{item:goal-reusability},
+Still, these (and other) applications need the basic techniques that our challenge problems emphasize:
+as per design goal \ref{item:goal-reusability},
 our problems are drawn from the basic metatheory of process calculi (without requiring the development of new theory), and focus on well-known theorems
 and results that involve interesting proof techniques that may be reused in further work.
-To achieve our design goal \ref{item:goal-comperison-accessibility},
-we have drafted our challenge problems to exercise the three aspects (linearity, scope extrusion, coinduction) independently, so that they can be solved individually and in any order;
-each problem should be reasonably small and easy to understand with basic knowledge of textbook concurrency theory, process calculi, and type theory.  The process calculus used in each challenge
-focuses on the features that we want to exercise, and omits all constructs
-(such as choices) that would complicate the mechanisation without bringing tangible
+To achieve design goal \ref{item:goal-comperison-accessibility},
+we have formulated our challenge problems to explore the three aspects (linearity, scope extrusion, coinduction) independently, so that they may be solved individually and in any order;
+each problem should be reasonably small and easily understood with basic knowledge of textbook concurrency theory, process calculi, and type theory.  The process calculus used in each challenge
+focuses on the features that we want to emphasize, and omits all constructs
+(such as choices) that would complicate the mechanisation without bringing additional tangible
 benefits and insights. 
-The minimality and uniformity of the calculi also allows us to target the design
+The minimality and uniformity of the calculi also allows us to target design
 goal \ref{item:goal-tutorials}.
@@
 For our challenge on linear reasoning we have chosen a type preservation theorem
-for a session type system since this allows us to involve linear
-reasoning without many definitions. Accordingly, we have chosen not to
+for a session type system, which allows us to introduce linear
+reasoning with few definitions. Accordingly, we have chosen not to
 include channel delegation and to use a reduction instead of a
@@
 We have chosen a simple notion of well-formedness for our challenge,
-and do not for instance consider deadlocks. While more
+and do not, for instance, consider deadlocks. While more
 sophisticated notions of well-formedness are interesting, the
-proofs of these properties are more complicated and would make
-the challenge more difficult and less focused on the linear
-aspects.
+proofs of these properties would complicate the challenge, distracting from the linear aspects.
 
@@
 challenge since it concerns the properties of binders in restrictions.
-It is not clear that the techniques used to solve the original challenges about binders can be easily extended to reasoning about scope extrusion.
-For example, techniques like Higher Order Abstract Syntax, or well-scoped
+It is not obvious that the techniques used to solve the original challenges about binders can be easily extended to reasoning about scope extrusion.
+For example, techniques like higher-order abstract syntax, or well-scoped
 De\ Bruijn indices do not have the concept of free variables and thus encoding this challenge with them requires some attention.
@@
 \subsubsection{Evaluation criteria.}
-The idea of our benchmark is to obtain evidence towards answering questions \cref{item:rq1,item:rq2,item:rq3}: therefore, we are interested in not only the solutions, but in the experience of solving the challenges with the chosen approach.
+The idea behind our benchmark is to obtain evidence towards answering questions \cref{item:rq1,item:rq2,item:rq3}.
+We are therefore interested in not only the solutions, but alsoin the experience of solving the challenges with the chosen approach.
 Solutions to our challenges should be compared on three measures:
-\begin{itemize}
-\item the mechanisation overhead, i.e.~the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; \item the adequacy of the formal statements in the mechanisation, i.e.~whether the proven theorems are easily recognisable as the theorems from the challenge; \item the cost of entry for the tools and techniques employed, i.e.~the difficulty of learning to use the techniques.
-\end{itemize}
-Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges need to present more elaborate argumentation for their adequacy.
+\begin{enumerate}
+\item Mechanisation overhead: the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation;
+\item Adequacy of the formal statements in the mechanisation: whether the proven theorems are easily recognisable as the theorems from the challenge; and
+\item Cost of entry for the tools and techniques employed: the difficulty of learning to use the techniques.
+\end{enumerate}
+Solutions to our challenges do not need to strictly follow the definitions and lemmata set out in the problem text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.
 
@@
 \paragraph{Extending and combining the current challenges.}
-The proposed challenges try to minimise the overlap of the techniques
-and aspects they exercise (linearity, scope extrusion, coinduction) --- but most
-research on concurrency theory combines some or all such aspects. A second round of formalisations could
+While the proposed challenges try to minimise the overlap of the techniques
+and aspects they explore (linearity, scope extrusion, coinduction), most
+research on concurrency theory combines some or all of these. A second round of formalisations might
 entail combining the features of the challenges.
@@
 \emph{(1)} the addition of choice to the type level and the process level,
-\emph{(2)} the addition of recursion (and recursive types for typed systems),
+\emph{(2)} the addition of recursion and (for typed systems) recursive types,
 and \emph{(3)} the
@@
 \begin{table}[btp]
-  \caption{Possible extensions of the challenges. Each suitable extension is marked ``+''. \label{tab:extensions}}
+  \caption{Possible extensions of the challenges. Each suitable extension is marked ``+''. \label{tab:extensions} \\}
   \begin{center}\small
@@
 \paragraph{Addressing new calculi and new features.}
-Extending in this dimension
+Extending in this direction
 means proposing new challenges covering different features of
 process calculi, or different proof mechanisation features.
-Some interesting aspects of message passing calculi to explore in
+Some interesting aspects of message-passing calculi to explore in
 further challenges could be multiparty session types and
 choreographies, as their meta-theory includes aspects
-that we do not cover here (e.g.: well-formed global types,
-projection as a partial function) --- although some aspects
-that we cover here do appear in their theory. Also, one could easily
+that we do not cover here (\eg: well-formed global types,
+projection as a partial function). Also, one could easily
 design a challenge that goes beyond our coinduction challenge, by
-exploring different notions of bisimilarity (barbed, weak, etc.),
+exploring different notions of bisimilarity (barbed, weak, \etc)
 and trace equivalence.
@@
 automating aspects of the proofs. Finally, a challenge could propose the
-integration with other formal reasoning tools, namely model checkers. These
-tools are in common use in the field --- and combining automated proofs with
+integration with other formal reasoning tools, such as model checkers. These
+tools are in common use in the field; combining automated proofs with
 proof assistants offers the potential to ease the path towards larger proofs.
@@
 worthwhile directions, and we look forward to a future when they
-indeed are extended. We see the current challenge as setting
+indeed are. We see the current challenge as setting
 the foundation for those future extensions. It is our expectation that
 the solutions of those future challenges will build on solutions
-to this challenge.
+to this one.
 

### 2024-01-25 short-paper/short-paper.tex
Fix some mistakes in references.

@@
 Several papers in this realm already include mechanised proofs
-(e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
+(e.g.~\cite{Castro-Perez2021,Castro2020,lmcs:9985,
   DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or discuss proof mechanisation techniques
-(e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
-  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).  Our
+(e.g.~\cite{Bengtson2016, Gay2001,
+  Thiemann2019, DBLP:conf/forte/ZalakainD21}).  Our
 experience, however, is that choosing the appropriate proof mechanisation techniques and 

### 2024-01-25 existing-mechanisations/existing-mechanisations.tex
Categorize related work by proof assistant and authors.

--- 2024-01-09 existing-mechanisations/existing-mechanisations.tex
@@
 \item \url{https://doi.org/10.1145/3158116}
+\item \url{https://doi.org/10.1007/978-3-030-78089-0_9}
 \end{itemize}
 
-A total of 44 papers that we currently know about.
+A total of 45 papers that we currently know about.
 
-\subsection{Template}
-Abstract
+\subsection{Papers written by ``us''}
+\subsubsection{Marco and Dawit}
+\cite{Tirore:2023} \url{https://doi.org/10.4230/LIPIcs.ITP.2023.28}
 
-\subsubsection{Approach}
-\begin{itemize}
-\item Proof assistant
-\item 
-\end{itemize}
+\subsubsection{Nobuko}
+\cite{Orchard2016} \url{https://doi.org/10.4204/EPTCS.203.1}
+
+\subsubsection{David, Francisco and Nobuko}
+\cite{Castro2020} \url{https://doi.org/10.1007/978-3-030-45237-7_17}
+
+\paragraph{+ Lorenzo}
+\cite{Castro-Perez2021} \url{https://doi.org/10.1145/3453483.3454041}
+
+\subsection{Approaches}
+\begin{tabular}{l|c|c|c|c|l}
+  Reference                          & Proof assistant        & Linearity & Binders & Coinduction & Limitations \\\hline
+  \cite{Tirore:2023}                 & Coq                    &           &         &             &             \\
+  \cite{Cruz-Filipe2021}             & Coq                    &           &         &             &             \\
+  \cite{Pohjola2022}                 & HOL4                   &           &         &             &             \\
+  \cite{Tassarotti2017}              & Coq                    &           &         &             &             \\
+  \cite{Thiemann2019}                & Agda                   &           &         &             &             \\
+  \cite{Goto2016}                    & Coq                    &           &         &             &             \\
+  \cite{Affeldt2008}                 & Coq                    &           &         &             &             \\
+  \cite{Scagnetto2002}               & Coq                    &           &         &             &             \\
+  \cite{Hirschkoff1997}              & Coq                    &           &         &             &             \\
+  \cite{Melham1994}                  & HOL                    &           &         &             &             \\
+  \cite{Ait-Mohamed1994}             & HOL                    &           &         &             &             \\
+  \cite{Rockl2003}                   & Isabelle               &           &         &             &             \\
+  \cite{Rockl2001}                   & Isabelle               &           &         &             &             \\
+  \cite{Henry-Greard1999}            & Coq                    &           &         &             &             \\
+  \cite{Miller2023}                  & Abella/$\lambda$Prolog &           &         &             &             \\
+  \cite{Gordon1996}                  & HOL                    &           &         &             &             \\
+  \cite{Gay2001}                     & Isabelle/HOL           &           &         &             &             \\
+  \cite{Despeyroux2000}              & Coq                    &           &         &             &             \\
+  \cite{Gillard2000}                 & Coq                    &           &         &             &             \\
+  \cite{Honsell2001}                 & Coq                    &           &         &             &             \\
+  \cite{Perera2018}                  & Agda                   &           &         &             &             \\
+  \cite{Watkins2008}                 & CLF                    &           &         &             &             \\
+  \cite{Tiu2010}                     & Bedwyr?                &           &         &             &             \\
+  \cite{Cervesato2007}               & CLF                    &           &         &             &             \\
+  \cite{Baelde2014}                  & Abella                 &           &         &             &             \\
+  \cite{Bengtson2009}                & Isabelle               &           & Nominal &             &             \\
+  \cite{Orchard2016}                 & Agda                   &           &         &             &             \\
+  \cite{Castro2020}                  & Coq                    &           &         &             &             \\
+  \cite{Zalakain2019}                & Coq                    &           &         &             &             \\
+  \cite{Petz2016}                    & Coq                    &           &         &             &             \\
+  \cite{Bock2016}                    & CLF                    &           &         &             &             \\
+  \cite{Xi2016}                      & ATS                    &           &         &             &             \\
+  \cite{Hirsch2022}                  & Coq                    &           &         &             &             \\
+  \cite{Cruz-Filipe2021b}            & Coq                    &           &         &             &             \\
+  \cite{Maksimovic2015}              & Coq                    &           &         &             &             \\
+  \cite{Parrow2014}                  & Isabelle               &           & Nominal &             &             \\
+  \cite{Bengtson2016}                & Isabelle               &           & Nominal &             &             \\
+  \cite{Kahsai2008}                  & Isabelle               &           & Nominal &             &             \\
+  \cite{Bengtson2007}                & Isabelle               &           & Nominal &             &             \\
+  \cite{Castro-Perez2021}            & Coq                    &           &         &             &             \\
+  \cite{Gay2020}                     & Agda                   &           &         &             &             \\
+  \cite{Brady2017}                   & Idris                  &           &         &             &             \\
+  \cite{Hinrichsen2019}              & Coq                    &           &         &             &             \\
+  \cite{Sergey2017}                  & Coq                    &           &         &             &             \\
+  \cite{DBLP:conf/forte/ZalakainD21} & Agda                   &           &         &             & 
+\end{tabular}
+
+\subsection{Existing experiences and suggestions}
+In this section, we collect quotes about experiences attempting to mechanise process calculi and related systems, as well as suggestions proposed by other authors.
 
 \subsubsection{Quotes about challenges encountered}
+This section collects quotes about challenges encountered using various techniques for mechanisation.
 
-\subsubsection{Suggestions in the paper}
+\subsubsection{Quotes about effort on technical lemmas vs.\ effort on actual content}
+This section collects quotes about the amount of effort spent on technical lemmas required by the mechanisation technique versus the amount of effort spent on the theorems of interest.
+
+\subsubsection{Quotes about suggestions for improvements in proof assistant technology}
+This section collects quotes suggesting improvements in proof assistant technology and/or mechanisation techniques.

### 2024-01-25 existing-mechanisations/existing-mechanisations.tex
Add references to links in list of unprocessed mechanisations.

@@
 \subsection{Unprocessed mechanisations}
-\begin{itemize}
-\item \url{https://doi.org/10.4230/LIPIcs.ITP.2023.28}
-\item \url{https://doi.org/10.1007/978-3-030-85315-0_8}
-\item \url{https://doi.org/10.4230/LIPIcs.ITP.2022.27}
-\item \url{https://doi.org/10.1007/978-3-662-54434-1_34}
-\item \url{https://doi.org/10.1145/3354166.3354184}
-\item \url{https://doi.org/10.1017/S0960129514000231}
-\item \url{https://doi.org/10.1016/j.entcs.2007.11.010}
-\item \url{https://doi.org/10.1016/S1571-0661(04)80507-3}
-\item \url{https://doi.org/10.1007/BFb0028392}
-\item T. F. Melham. 1994. A mechanized theory of the $\pi$-calculus in HOL. Nordic J. of Computing 1, 1 (Spring 1994), 50–76. (\url{https://dl.acm.org/doi/10.5555/640186.640190}, but paper is missing)
-\item Otmane Ait-Mohamed. Vérification de l'équivalence du $\pi$-calcul dans HOL. RR-2412, INRIA. 1994 (\url{https://hal.science/inria-00074263/}, in French)
-\item \url{https://doi.org/10.1017/S0956796802004653}
-\item \url{https://doi.org/10.1007/3-540-45315-6_24}
-\item \url{https://inria.hal.science/inria-00072970}
-\item \url{https://www.lix.polytechnique.fr/Labo/Dale.Miller/lProlog/examples/pi-calculus/toc.html}
-\item \url{https://doi.org/10.1007/BFb0105404}
-\item \url{https://doi.org/10.1007/3-540-44755-5_16}
-\item \url{https://doi.org/10.1007/3-540-44929-9_30}
-\item \url{https://doi.org/10.1007/10721959_33}
-\item \url{https://doi.org/10.1016/S0304-3975(00)00095-5}
-\item \url{https://doi.org/10.1017/S096012951700010X}
-\item \url{https://doi.org/10.1016/j.entcs.2007.11.013}
-\item \url{https://doi.org/10.1145/1656242.1656248}
-\item \url{https://doi.org/10.1184/R1/6587294.v1}
-\item \url{https://doi.org/10.6092/issn.1972-5787/4650}
-\item \url{https://doi.org/10.2168/LMCS-5(2:16)2009}
-\item \url{https://doi.org/10.4204/EPTCS.203.1}
-\item \url{https://doi.org/10.1007/978-3-030-45237-7_17}
-\item Uma Zalakain. 2019. Type-checking session-typed $\pi$-calculus with Coq. Master's thesis, University of Glasgow. \url{https://umazalakain.info/static/Z19/msc-thesis.pdf}
-\item Adam Michael Petz. 2016. A Semantics for Attestation Protocols using Session Types in Coq. Master's thesis, University of Kansas. \url{http://hdl.handle.net/1808/24137}
-\item \url{https://pure.itu.dk/en/publications/representing-session-types}
-\item \url{https://doi.org/10.48550/arXiv.1603.03727}
-\item \url{https://doi.org/10.1145/3498684}
-\item \url{https://doi.org/10.4230/LIPIcs.ITP.2021.15}
-\item \url{https://doi.org/10.1007/978-3-319-22102-1_19}
-\item \url{https://doi.org/10.1017/S0960129513000170}
-\item \url{https://doi.org/10.1007/s10817-015-9336-2}
-\item \url{https://doi.org/10.1007/978-3-540-69407-6_33}
-\item \url{https://doi.org/10.1016/j.entcs.2007.08.017}
-\item \url{https://doi.org/10.1145/3453483.3454041}
-\item \url{https://doi.org/10.4204/EPTCS.314.3}
-\item \url{https://doi.org/10.7494/csci.2017.18.3.1413}
-\item \url{https://doi.org/10.1145/3371074}
-\item \url{https://doi.org/10.1145/3158116}
-\item \url{https://doi.org/10.1007/978-3-030-78089-0_9}
-\end{itemize}
+\begin{description}
+\item[\cite{Tirore:2023}]                 \url{https://doi.org/10.4230/LIPIcs.ITP.2023.28}
+\item[\cite{Cruz-Filipe2021}]             \url{https://doi.org/10.1007/978-3-030-85315-0_8}
+\item[\cite{Pohjola2022}]                 \url{https://doi.org/10.4230/LIPIcs.ITP.2022.27}
+\item[\cite{Tassarotti2017}]              \url{https://doi.org/10.1007/978-3-662-54434-1_34}
+\item[\cite{Thiemann2019}]                \url{https://doi.org/10.1145/3354166.3354184}
+\item[\cite{Goto2016}]                    \url{https://doi.org/10.1017/S0960129514000231}
+\item[\cite{Affeldt2008}]                 \url{https://doi.org/10.1016/j.entcs.2007.11.010}
+\item[\cite{Scagnetto2002}]               \url{https://doi.org/10.1016/S1571-0661(04)80507-3}
+\item[\cite{Hirschkoff1997}]              \url{https://doi.org/10.1007/BFb0028392}
+\item[\cite{Melham1994}]                  T.
+F.
+Melham.
+1994.
+A mechanized theory of the $\pi$-calculus in HOL.
+Nordic J.
+of Computing 1, 1 (Spring 1994), 50–76.
+(\url{https://dl.acm.org/doi/10.5555/640186.640190}, but paper is missing)
+\item[\cite{Ait-Mohamed1994}]             Otmane Ait-Mohamed. Vérification de l'équivalence du $\pi$-calcul dans HOL. RR-2412, INRIA. 1994 (\url{https://hal.science/inria-00074263/}, in French)
+\item[\cite{Rockl2003}]                   \url{https://doi.org/10.1017/S0956796802004653}
+\item[\cite{Rockl2001}]                   \url{https://doi.org/10.1007/3-540-45315-6_24}
+\item[\cite{Henry-Greard1999}]            \url{https://inria.hal.science/inria-00072970}
+\item[\cite{Miller2023}]                  \url{https://www.lix.polytechnique.fr/Labo/Dale.Miller/lProlog/examples/pi-calculus/toc.html}
+\item[\cite{Gordon1996}]                  \url{https://doi.org/10.1007/BFb0105404}
+\item[\cite{Gay2001}]                     \url{https://doi.org/10.1007/3-540-44755-5_16}
+\item[\cite{Despeyroux2000}]              \url{https://doi.org/10.1007/3-540-44929-9_30}
+\item[\cite{Gillard2000}]                 \url{https://doi.org/10.1007/10721959_33}
+\item[\cite{Honsell2001}]                 \url{https://doi.org/10.1016/S0304-3975(00)00095-5}
+\item[\cite{Perera2018}]                  \url{https://doi.org/10.1017/S096012951700010X}
+\item[\cite{Watkins2008}]                 \url{https://doi.org/10.1016/j.entcs.2007.11.013}
+\item[\cite{Tiu2010}]                     \url{https://doi.org/10.1145/1656242.1656248}
+\item[\cite{Cervesato2007}]               \url{https://doi.org/10.1184/R1/6587294.v1}
+\item[\cite{Baelde2014}]                  \url{https://doi.org/10.6092/issn.1972-5787/4650}
+\item[\cite{Bengtson2009}]                \url{https://doi.org/10.2168/LMCS-5(2:16)2009}
+\item[\cite{Orchard2016}]                 \url{https://doi.org/10.4204/EPTCS.203.1}
+\item[\cite{Castro2020}]                  \url{https://doi.org/10.1007/978-3-030-45237-7_17}
+\item[\cite{Zalakain2019}]                Uma Zalakain.
+2019.
+Type-checking session-typed $\pi$-calculus with Coq.
+Master's thesis, University of Glasgow.
+\url{https://umazalakain.info/static/Z19/msc-thesis.pdf}
+\item[\cite{Petz2016}]                    Adam Michael Petz.
+2016.
+A Semantics for Attestation Protocols using Session Types in Coq.
+Master's thesis, University of Kansas.
+\url{http://hdl.handle.net/1808/24137}
+\item[\cite{Bock2016}]                    \url{https://pure.itu.dk/en/publications/representing-session-types}
+\item[\cite{Xi2016}]                      \url{https://doi.org/10.48550/arXiv.1603.03727}
+\item[\cite{Hirsch2022}]                  \url{https://doi.org/10.1145/3498684}
+\item[\cite{Cruz-Filipe2021b}]            \url{https://doi.org/10.4230/LIPIcs.ITP.2021.15}
+\item[\cite{Maksimovic2015}]              \url{https://doi.org/10.1007/978-3-319-22102-1_19}
+\item[\cite{Parrow2014}]                  \url{https://doi.org/10.1017/S0960129513000170}
+\item[\cite{Bengtson2016}]                \url{https://doi.org/10.1007/s10817-015-9336-2}
+\item[\cite{Kahsai2008}]                  \url{https://doi.org/10.1007/978-3-540-69407-6_33}
+\item[\cite{Bengtson2007}]                \url{https://doi.org/10.1016/j.entcs.2007.08.017}
+\item[\cite{Castro-Perez2021}]            \url{https://doi.org/10.1145/3453483.3454041}
+\item[\cite{Gay2020}]                     \url{https://doi.org/10.4204/EPTCS.314.3}
+\item[\cite{Brady2017}]                   \url{https://doi.org/10.7494/csci.2017.18.3.1413}
+\item[\cite{Hinrichsen2019}]              \url{https://doi.org/10.1145/3371074}
+\item[\cite{Sergey2017}]                  \url{https://doi.org/10.1145/3158116}
+\item[\cite{DBLP:conf/forte/ZalakainD21}] \url{https://doi.org/10.1007/978-3-030-78089-0_9}
+\end{description}
 

### 2024-01-25 paper-draft-esop/paper.tex

--- 2023-10-18 paper/paper.tex
@@
 Several papers in the realm of process calculi already include mechanised proofs
-(e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
+(e.g.~\cite{Castro-Perez2021,Castro2020,lmcs:9985,
   DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or
 discuss proof mechanisation techniques
-(e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
-  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).  However, our
+(e.g.~\cite{Bengtson2016, Gay2001,
+  Thiemann2019, DBLP:conf/forte/ZalakainD21}).  However, our
 experience is that choosing the appropriate proof mechanisation techniques and

### 2024-01-25 short-paper/short-paper.tex

--- 2024-01-21 short-paper/short-paper.tex
@@
 concurrency (and process calculi, more specifically) will benefit from a standardised set
-of challenge probems.
+of challenge problems.
 

### 2024-01-25 technical-report/main.tex
Add list of existing mechanisations.
Add table and BibTeX for existing mechanisations.

@@
 \usepackage{cleveref}
-\usepackage{hyperref}
 \usepackage[inline]{enumitem}
@@
 \title{POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types}
-\author{...}
+
+\author{}
 
@@
 
-\bibliographystyle{plainnat}
+\section{Overview of existing mechanisations}
+\input{../existing-mechanisations/existing-mechanisations.tex}
+
+\section{Future work}
+\begin{itemize}
+\item Multiparty session types
+\item Choreography
+\item Encodings
+\item Code extraction
+\item Conversation types
+\item Psi-calculus
+\item Model checking for system properties
+\item More general types than session types
+\end{itemize}
+
+\bibliographystyle{plain}
 \bibliography{../references}

### 2024-01-25 challenges/coinduction.tex
Fix text introducing syntax in coinduction challenge.
Further copy-editing

@@
 We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
+Note that, since our systems have potentially infinite behaviours, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
 

### 2024-01-25 challenges/linearity.tex
Further copy-editing

@@
-\subsection{Challenge: Linearity and behavioral type systems}
+\subsection{Challenge: Linearity and behavioural type systems}
 \label{sec:challenge:linearity-beh-types}
-This challenge formalises a proof that requires reasoning about linearity of channels.
-Linearity is the notion that a channel must be used exactly once in a process
+This challenge formalises a proof that requires reasoning about the linearity of channels.
+Linearity is the notion that a channel must be used exactly once by a process.
 This is necessary to prove properties about session type systems, and the key
 issue of this challenge is reasoning about the linearity of context splitting operations.
-Linear reasoning is also necessary to formalise e.g.\ linear and affine types for the pi-calculus and cut elimination in linear logics.
+Linear reasoning is also necessary to formalise, \eg, linear and affine types for the \(\pi\)-calculus and cut elimination in linear logics.
 
@@
 The main objective of this challenge is to prove type preservation (also
-known as subject reduction), i.e.\ that well-typed
+known as subject reduction), \ie, that well-typed
 processes can only transition to processes which are also well-typed
 in the same context.
-The second objective is to prove type safety, i.e.\ that well-typed processes are also well-formed in the sense that they do not use endpoints in a non-dual way.
+The second objective is to prove type safety, \ie, that well-typed processes are also well-formed in the sense that they do not use endpoints in a non-dual way.
 
@@
 Note that in this calculus channels cannot be sent in messages, therefore the topology of the communication network described by a process cannot change.
-Also, there is no recursion or replication in the syntax, hence no infinite behaviors can be expressed. In particular, we only model linear (as opposed to shared) channels.
+Also, there is no recursion or replication in the syntax, hence no infinite behaviours can be expressed. In particular, we only model linear (as opposed to shared) channels.
 
@@
 
-The operational semantics is defined as the following relation on processes:
+The operational semantics are defined as the following relation on processes:
 \begin{mathpar}
@@
 \begin{theorem}[Type safety]
-  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
+  If \( \types{\Gamma;\Cempty}{P} \), then \( P \) is well-formed.
 \end{theorem}
@@
 \begin{corollary}
-  If \( \types{\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well formed.
+  If \( \types{\Gamma;\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well formed.
 \end{corollary}

### 2024-01-25 challenges/scope-extrusion.tex
Backport editing into challenges/.
Further copy-editing

@@
 This challenge formalises a proof that requires explicit scope extrusion.
-Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (i.e.\ expanded) to include the receiving process.
-This e.g.\ allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
+Scope extrusion is the notion that a process can send restricted names to another process, as long as the restriction can safely be ``extruded'' (\ie, expanded) to include the receiving process.
+This, for instance, allows a process to set up a private connection by sending a restricted name to another process, then using this name for further communication.
 The key issue of this challenge is reasoning about names that are ``in the process'' of being scope-extruded, which often presents difficulties for the mechanisation of binders.
@@
 Reasoning about scope extrusion explicitly can sometimes be avoided by introducing a structural congruence rule into the semantics, but doing this means we lose information about the scope when reasoning about the semantics.
-Explicitly reasoning about scope extrusion is necessary to describe e.g.\ runtime monitors and compositions of systems.
+Explicitly reasoning about scope extrusion is necessary to describe, \eg, runtime monitors and compositions of systems.
 
@@
 % The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
-Note that there is no recursion or replication in the syntax, and thus no infinite behaviors can be expressed.
+Note that there is no recursion or replication in the syntax, and thus no infinite behaviours can be expressed.
 This simplifies the theory and is orthogonal to the concept of scope extrusion.

### 2024-01-27 existing-mechanisations/existing-mechanisations.tex
Make table of related mechanisations fit on page.
filled some entries in the table. Commented out some other which were not relevant or subsumed by previous ones

@@
 \item[\cite{Affeldt2008}]                 \url{https://doi.org/10.1016/j.entcs.2007.11.010}
-\item[\cite{Scagnetto2002}]               \url{https://doi.org/10.1016/S1571-0661(04)80507-3}
+\item[\cite{Honsell2001}]
+% \item[\cite{Scagnetto2002}]
+  \url{https://doi.org/10.1016/S1571-0661(04)80507-3} First weak HOAS
+  formalization of the full pi calculus in Coq within the ``Theory of
+  Context'', an axiomatic reflection of properties of
+  names. Congruence properties of bisimilarity mainly via Park's
+  principle.
 \item[\cite{Hirschkoff1997}]              \url{https://doi.org/10.1007/BFb0028392}
@@
 \item[\cite{Henry-Greard1999}]            \url{https://inria.hal.science/inria-00072970}
-\item[\cite{Miller2023}]                  \url{https://www.lix.polytechnique.fr/Labo/Dale.Miller/lProlog/examples/pi-calculus/toc.html}
+%\item[\cite{Miller2023}]                  \url{https://www.lix.polytechnique.fr/Labo/Dale.Miller/lProlog/examples/pi-calculus/toc.html}
 \item[\cite{Gordon1996}]                  \url{https://doi.org/10.1007/BFb0105404}
@@
 \item[\cite{Despeyroux2000}]              \url{https://doi.org/10.1007/3-540-44929-9_30}
+  weak HOAS encoding in Coq, no exotic terms. Proof of subject reduction.
 \item[\cite{Gillard2000}]                 \url{https://doi.org/10.1007/10721959_33}
@@
 \item[\cite{Perera2018}]                  \url{https://doi.org/10.1017/S096012951700010X}
-\item[\cite{Watkins2008}]                 \url{https://doi.org/10.1016/j.entcs.2007.11.013}
+% \item[\cite{Watkins2008}]                 \url{https://doi.org/10.1016/j.entcs.2007.11.013}
 \item[\cite{Tiu2010}]                     \url{https://doi.org/10.1145/1656242.1656248}
 \item[\cite{Cervesato2007}]               \url{https://doi.org/10.1184/R1/6587294.v1}
-\item[\cite{Baelde2014}]                  \url{https://doi.org/10.6092/issn.1972-5787/4650}
 \item[\cite{Bengtson2009}]                \url{https://doi.org/10.2168/LMCS-5(2:16)2009}
@@
 \item[\cite{Kahsai2008}]                  \url{https://doi.org/10.1007/978-3-540-69407-6_33}
-\item[\cite{Bengtson2007}]                \url{https://doi.org/10.1016/j.entcs.2007.08.017}
+% \item[\cite{Bengtson2007}]                \url{https://doi.org/10.1016/j.entcs.2007.08.017}
 \item[\cite{Castro-Perez2021}]            \url{https://doi.org/10.1145/3453483.3454041}
@@
 \subsection{Approaches}
+\begin{center}
+  \small
 \begin{tabular}{l|c|c|c|c|l}
@@
   \cite{Affeldt2008}                 & Coq                    &           &         &             &             \\
-  \cite{Scagnetto2002}               & Coq                    &           &         &             &             \\
+  \cite{Honsell2001}               & Coq                    &           &weak HOAS + ToC         &cofix and Park's             & issues with cofix            \\
   \cite{Hirschkoff1997}              & Coq                    &           &         &             &             \\
@@
   \cite{Henry-Greard1999}            & Coq                    &           &         &             &             \\
-  \cite{Miller2023}                  & Abella/$\lambda$Prolog &           &         &             &             \\
+  %\cite{Miller2023}                  & Abella/$\lambda$Prolog &           &         &             &             \\
   \cite{Gordon1996}                  & HOL                    &           &         &             &             \\
   \cite{Gay2001}                     & Isabelle/HOL           &           &         &             &             \\
-  \cite{Despeyroux2000}              & Coq                    &           &         &             &             \\
+  \cite{Despeyroux2000}              & Coq                    &           & weak HOAS         &             &             \\
   \cite{Gillard2000}                 & Coq                    &           &         &             &             \\
-  \cite{Honsell2001}                 & Coq                    &           &         &             &             \\
   \cite{Perera2018}                  & Agda                   &           &         &             &             \\
-  \cite{Watkins2008}                 & CLF                    &           &         &             &             \\
-  \cite{Tiu2010}                     & Bedwyr?                &           &         &             &             \\
-  \cite{Cervesato2007}               & CLF                    &           &         &             &             \\
-  \cite{Baelde2014}                  & Abella                 &           &         &             &             \\
+%  \cite{Watkins2008}                 & CLF                    &           &         &             &             \\
+  \cite{Tiu2010}                     & Abella                &           & HOAS + $\nabla$        & coDefine             &              \\
+  \cite{Cervesato2007}               & CLF                    &           & HOAS + substructural opsem       &             &   no metatheory          \\
   \cite{Bengtson2009}                & Isabelle               &           & Nominal &             &             \\
@@
   \cite{Kahsai2008}                  & Isabelle               &           & Nominal &             &             \\
-  \cite{Bengtson2007}                & Isabelle               &           & Nominal &             &             \\
+%  \cite{Bengtson2007}                & Isabelle               &           & Nominal &             &             \\
   \cite{Castro-Perez2021}            & Coq                    &           &         &             &             \\
@@
 \end{tabular}
+\end{center}
 

### 2024-01-29 short-paper/short-paper.tex
Further copy-editing

--- 2024-01-25 short-paper/short-paper.tex
@@
 the development of proof assistants, libraries, and best practices. In doing so, it 
-spearheaded a shift in the programming lanuage research community towards 
+spearheaded a shift in the programming language research community towards 
 publications that include mechanised proofs. 

### 2024-01-29 existing-mechanisations/existing-mechanisations.tex
some more entries in mech table

@@
 \item[\cite{Despeyroux2000}]              \url{https://doi.org/10.1007/3-540-44929-9_30}
-  weak HOAS encoding in Coq, no exotic terms. Proof of subject reduction.
+  weak HOAS encoding in Coq, based on \emph{concretions}, no exotic terms. Proof of subject reduction.
 \item[\cite{Gillard2000}]                 \url{https://doi.org/10.1007/10721959_33}
@@
 \begin{tabular}{l|c|c|c|c|l}
-  Reference                          & Proof assistant        & Linearity & Binders & Coinduction & Limitations \\\hline
+  Reference                          & Proof assistant        & Linearity & Binders & Coinduction & Notes/Limitations \\\hline
   \cite{Tirore:2023}                 & Coq                    &           &         &             &             \\
@@
   \cite{Tiu2010}                     & Abella                &           & HOAS + $\nabla$        & coDefine             &              \\
-  \cite{Cervesato2007}               & CLF                    &           & HOAS + substructural opsem       &             &   no metatheory          \\
+    \cite{ChaudhuriCM15}                     & Abella                &           & HOAS + $\nabla$        & bisimulation up to             &              \\
+  \cite{Cervesato2007}               & CLF                    &  substructural op.\ semantics         & HOAS       &             &   \textbf{no metatheory}          \\
   \cite{Bengtson2009}                & Isabelle               &           & Nominal &             &             \\
   \cite{Orchard2016}                 & Agda                   &           &         &             &             \\
+    \cite{CicconeP20}                 & Agda                   & context splitting, labelled types           & well typed DB        &             &             \\
   \cite{Castro2020}                  & Coq                    &           &         &             &             \\
@@
   \cite{Petz2016}                    & Coq                    &           &         &             &             \\
+    \cite{AmbalLS21}                    & Coq                    &           & DB + nameless         &             &      higher order pi       \\
   \cite{Bock2016}                    & CLF                    &           &         &             &             \\

### 2024-01-30 short-paper/short-paper.tex
small fixes, comments

@@
   focus on the challenges and issues that typically arise when mechanising
-  message-passing concurrency using process calculi.  Our benchmark problems
+  message-passing concurrency using process calculi.\luca{wondering whether it would make sense to say ``\dots using formal models of concurrent and distributed programming languages in the form of process calculi'' or something similar to stress that process calculi are not the subject of the challenges, but ``only'' an appropriate tool for this kind of endeavor}  Our benchmark problems
   address three key topics: linearity, scope extrusion, and coinductive
@@
 issues that arise when mechanising message-passing concurrency using process 
-calculi.
+calculi.\luca{see comment in abstract}
 
@@
 just one of these aspects at once; we see the combination of techniques
-as a next step beyond the scope of this challenge (we discuss this in
-\cref{sec:going-beyond}).
+as a next step beyond the scope of this challenge (we discuss this further in
+\Cref{sec:going-beyond}).
 % Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate, evaluate, and compare the many tools and techniques available.
@@
 \item\label{item:rq1} What is the current state of the art in mechanising process calculi?
-\item\label{item:rq2} What techniques and best practices can be recommended when starting formalisation projects within concurrency theory?
-\item\label{item:rq3} What improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
+\item\label{item:rq2} Which techniques and best practices can be recommended when starting formalisation projects within concurrency theory?
+\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
 \end{enumerate}
@@
 \paragraph{Outline of this paper.}
-We discuss our design choices in \cref{sec:design-discussion}, and possible extensions and further topics in \cref{sec:going-beyond}.
-The challenges themselves may be found in \cref{app:challenges}.
+We discuss our design choices in \Cref{sec:design-discussion}, and possible extensions and further topics in \Cref{sec:going-beyond}.
+The challenges themselves may be found in \Cref{app:challenges}.
 
@@
 
-\section{Design of the benchmark}\label{sec:design-discussion}
+\section{Design of the Benchmark}\label{sec:design-discussion}
 
@@
 The idea behind our benchmark is to obtain evidence towards answering questions \cref{item:rq1,item:rq2,item:rq3}.
-We are therefore interested in not only the solutions, but alsoin the experience of solving the challenges with the chosen approach.
+We are therefore interested in not only the solutions, but also in the experience of solving the challenges with the chosen approach.
 Solutions to our challenges should be compared on three measures:
@@
 entail combining the features of the challenges.
-In \cref{tab:extensions} we consider the following three extensions:
+In \Cref{tab:extensions} we consider the following three extensions:
 \emph{(1)} the addition of choice to the type level and the process level,
@@
 addition of channel delegation to those calculi that do not already
-support it.
+support it.%
+%
+\luca{I would argue that support for recursive types and processes may also
+impact the coinduction challenge. Indeed, some type systems with support for
+recursion are more easily expressed \emph{coinductively}, by allowing possibly
+infinite typing derivation. This application of coinduction is different from
+that intended in the challenge, where it's focused on expressing notions of
+process equivalence, and yet it is increasingly popular (see the related work on
+infinitary linear logics) and occasionally seen with skepticism, which would
+make a good case for formalization.}\luca{as a follow up to the previous
+comment, if there is agreement that recursive types impact the coinduction
+challenge then the relevance of Table 1 may be questionable because almost every
+extension impacts almost every challenge. In case of need we may consider
+removing the table in favor of more prose.}
 
@@
 process calculi, or different proof mechanisation features.
-Some interesting aspects of message-passing calculi to explore in
+Some interesting aspects of message-passing calculi to be explored in
 further challenges could be multiparty session types and
-choreographies, as their meta-theory includes aspects
-that we do not cover here (\eg: well-formed global types,
-projection as a partial function). Also, one could easily
+choreographies, as their meta-theory includes aspects -- \eg well-formedness conditions on global types, partiality of end-point projection function, etc. --
+that we do not cover here. Also, one could easily
 design a challenge that goes beyond our coinduction challenge, by

### 2024-01-30 short-paper/short-paper.tex
Remove outline section of the short paper.
Remove table of extensions from short paper.

@@
 
-\paragraph{Outline of this paper.}
-We discuss our design choices in \Cref{sec:design-discussion}, and possible extensions and further topics in \Cref{sec:going-beyond}.
-The challenges themselves may be found in \Cref{app:challenges}.
-
 \section{Motivation}\label{sec:motivation}
@@
 entail combining the features of the challenges.
-In \Cref{tab:extensions} we consider the following three extensions:
+We consider the following three extensions:
 \emph{(1)} the addition of choice to the type level and the process level,
@@
 infinitary linear logics) and occasionally seen with skepticism, which would
-make a good case for formalization.}\luca{as a follow up to the previous
-comment, if there is agreement that recursive types impact the coinduction
-challenge then the relevance of Table 1 may be questionable because almost every
-extension impacts almost every challenge. In case of need we may consider
-removing the table in favor of more prose.}
-
-\begin{table}[btp]
-  \caption{Possible extensions of the challenges. Each suitable extension is marked ``+''. \label{tab:extensions} \\}
-  \begin{center}\small
-    \vspace{-4mm}%
-    \begin{tabular}{|r|c|c|c|}
-      \hline
-      & Linearity & Scope extrusion & Coinduction \\
-      \hline
-      Choice & $+$ & $+$  & $+$ \\
-      \hline
-      Recursive types & $+$  & $+$ & \\
-      \hline
-      Delegation & $+$ & & $+$ \\
-      \hline
-    \end{tabular}
-    \vspace{-3mm}%
-  \end{center}
-\end{table}
+make a good case for formalization.}
 

### 2024-01-30 short-paper/lin.tex
first draft of related work abut linearity

\alb{Just a rough attempt to put some words on paper}

In mechanized meta-theory addressing linearity means choosing an
appropriate representation of a linear context. The crux of the matter
is \emph{context splitting} as showcased by the introduction rule of
the tensor. While a linear context can be seen as a multi-set, most
proof assistants have a better support for list processing.  Although
this representation is intuitive, it is overly concrete and requires
establishing dozens and dozens of intermediate lemmas that are foreign
to the mathematics of the problem, say proving type preservation. In
this space, several designs are possible: one can label occurrences of
resources constraining their usage, e.g.~\cite{CicconeP20}, or
\alb{something: ask Franscisco} as in \cite{Castro2020}. Some authors
instead impose a multiset structure over lists,
see~\cite{ChaudhuriLR19,Danielsson12}: still, the effort in developing
the infrastructure is significant. A different and interesting
approach is to bypass the problem of context splitting by adopting
% to  meta-theoretic properties
familiar ideas from linear algorithmic  type checking;
this in known as typing with \emph{leftovers}, as
exemplified in~\cite{DBLP:conf/forte/ZalakainD21}. Whatever the choice,
list-based encodings can be refined to be intrinsically-typed, if the proof assistant supports dependent types, see \cite{Thiemann2019,CicconeP20,RouvoetPKV20}.

For something completely different, one could adopt a
\emph{sub-structural} meta-logical frameworks: here it is the framework
itself that handles resource distribution and the user just has to map
her linear operations to the one offered by the framework. In fact,
the linear function space will handle all context operations
implicitly, including splitting and substitution. One such a framework
is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08}, see the encoding of
session types in~\cite{Bock2016}. Unfortunately, \emph{Clef} does not
yet support meta-theory. A compromise is the so called
\emph{two-level} approach, where one encodes a sub-structural
specification logic in a mainstream proof assistant and then use that
logic to state and prove linear properties, see for a recent
example~\cite{Felty:MSCS21}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "short-paper"
%%% End:

### 2024-01-31 short-paper/lin.tex
Minor textual fixes.

@@
 the tensor. While a linear context can be seen as a multi-set, most
-proof assistants have a better support for list processing.  Although
+proof assistants have better support for list processing.  Although
 this representation is intuitive, it is overly concrete and requires
@@
 For something completely different, one could adopt a
-\emph{sub-structural} meta-logical frameworks: here it is the framework
+\emph{sub-structural} meta-logical framework: here it is the framework
 itself that handles resource distribution and the user just has to map
-her linear operations to the one offered by the framework. In fact,
+her linear operations to the ones offered by the framework. In fact,
 the linear function space will handle all context operations
@@
 is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08}, see the encoding of
-session types in~\cite{Bock2016}. Unfortunately, \emph{Clef} does not
+session types in~\cite{Bock2016}. Unfortunately, \emph{Celf} does not
 yet support meta-theory. A compromise is the so called

### 2024-01-31 short-paper/binders.tex
Add first draft piece about binders.

The encoding of binding constructs has been extensively studied for type systems based on lambda calculi, but process calculi present unique challenges.
$\pi$-calculi typically include several different binding constructs: inputs bind a received name or value, recursive processes bind recursion variables, and restrictions bind names.
The first two act similarly to the binders known from lambda calculi, but restrictions are more challenging due to scope extrusion.

The case study \cite{AmbalLS21} compares four approaches to encoding binders in Coq as a first step towards developing tools for working with higher-order process calculi.
They found that working directly with de Bruijn indices was easiest and shortest since the existing approaches developed for lambda calculus binders do not work well with scope extrusion.
Specifically, the locally nameless representation~\cite{Chargueraud2012} cannot avoid direct manipulation of de Bruijn indices when defining the semantics of scope extrusion, cofinite quantification does not provide any benefits when working under binders, nominal approaches~\cite{Pitts2003} require explicitly giving and validating sets of free names during scope extrusion, and higher-order abstract syntax (HOAS)~\cite{Pfenning1988} requires additional axioms to work with contexts.

Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, i.e.\ induction principles on functions to work with HOAS encodings.
Some proof assistants such as Abella~\cite{Baelde2014} support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging~\cite{Momigliano2012}.

% Start of related work
\cite{Maksimovic2015} does not have name restriction, thus dodging the challenge of scope extrusion.

### 2024-01-31 short-paper/lin.tex
small changes to linearity section

@@
 In mechanized meta-theory addressing linearity means choosing an
-appropriate representation of a linear context. The crux of the matter
-is \emph{context splitting} as showcased by the introduction rule of
-the tensor. While a linear context can be seen as a multi-set, most
-proof assistants have better support for list processing.  Although
-this representation is intuitive, it is overly concrete and requires
-establishing dozens and dozens of intermediate lemmas that are foreign
-to the mathematics of the problem, say proving type preservation. In
-this space, several designs are possible: one can label occurrences of
-resources constraining their usage, e.g.~\cite{CicconeP20}, or
-\alb{something: ask Franscisco} as in \cite{Castro2020}. Some authors
-instead impose a multiset structure over lists,
-see~\cite{ChaudhuriLR19,Danielsson12}: still, the effort in developing
-the infrastructure is significant. A different and interesting
-approach is to bypass the problem of context splitting by adopting
+appropriate representation of the linear context. The crux of the matter
+is \emph{context splitting} as showcased by the typing rule for parallel composition.
+% the introduction rule of the tensor.
+While a linear context is best seen as a multi-set, most proof
+assistants have better support for list processing.  Although the
+latter representation is intuitive, it is overly concrete and may
+require establishing dozens and dozens of technical lemmas that are
+foreign to the mathematics of the problem under study, say proving
+type preservation for session types. In this space, several designs
+are possible: one can label occurrences of resources constraining
+their usage, e.g.~\cite{CicconeP20}, or \alb{something --- ask
+  Francisco} as in \cite{Castro2020}. Some authors instead impose a
+multiset structure over lists, see~\cite{ChaudhuriLR19,Danielsson12}:
+still, the effort in developing the infrastructure is significant.
+
+A different approach is to bypass the problem of context splitting by
+adopting
 % to  meta-theoretic properties
-familiar ideas from linear algorithmic  type checking;
-this in known as typing with \emph{leftovers}, as
-exemplified in~\cite{DBLP:conf/forte/ZalakainD21}. Whatever the choice,
-list-based encodings can be refined to be intrinsically-typed, if the proof assistant supports dependent types, see \cite{Thiemann2019,CicconeP20,RouvoetPKV20}.
+familiar ideas from algorithmic linear type checking: this is known as
+ \emph{typing with leftovers}, as exemplified
+in~\cite{DBLP:conf/forte/ZalakainD21}. Whatever the choice, list-based
+encodings can be refined to be intrinsically-typed, if the proof
+assistant does support dependent types, see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}.
 
@@
 \emph{sub-structural} meta-logical framework: here it is the framework
-itself that handles resource distribution and the user just has to map
-her linear operations to the ones offered by the framework. In fact,
+itself that handles resource distribution and the users just have to map
+their linear operations to the ones offered by the framework. In fact,
 the linear function space will handle all context operations
@@
 session types in~\cite{Bock2016}. Unfortunately, \emph{Celf} does not
-yet support meta-theory. A compromise is the so called
+yet fully support the verification of meta-theoretic properties. A compromise is the so called
 \emph{two-level} approach, where one encodes a sub-structural

### 2024-01-31 short-paper/binders.tex
Edits and comments

@@
-The encoding of binding constructs has been extensively studied for type systems based on lambda calculi, but process calculi present unique challenges.
+The encoding of binding constructs has been extensively studied for type systems based on $\lambda$-calculi, but process calculi present unique challenges.
 $\pi$-calculi typically include several different binding constructs: inputs bind a received name or value, recursive processes bind recursion variables, and restrictions bind names.
-The first two act similarly to the binders known from lambda calculi, but restrictions are more challenging due to scope extrusion.
+The first two act similarly to the binders known from $\lambda$-calculi, but restrictions are more challenging due to scope extrusion.
 
 The case study \cite{AmbalLS21} compares four approaches to encoding binders in Coq as a first step towards developing tools for working with higher-order process calculi.
-They found that working directly with de Bruijn indices was easiest and shortest since the existing approaches developed for lambda calculus binders do not work well with scope extrusion.
-Specifically, the locally nameless representation~\cite{Chargueraud2012} cannot avoid direct manipulation of de Bruijn indices when defining the semantics of scope extrusion, cofinite quantification does not provide any benefits when working under binders, nominal approaches~\cite{Pitts2003} require explicitly giving and validating sets of free names during scope extrusion, and higher-order abstract syntax (HOAS)~\cite{Pfenning1988} requires additional axioms to work with contexts.
+They found that working directly with de Bruijn indices was easiest and shortest since the existing approaches developed for $\lambda$-calculus binders do not work well with scope extrusion.
+Specifically, the locally nameless representation~\cite{Chargueraud2012} cannot avoid direct manipulation of de Bruijn indices when defining the semantics of scope extrusion; cofinite quantification provides no benefits when working under binders; nominal approaches~\cite{Pitts2003} require explicitly giving and validating sets of free names during scope extrusion; and higher-order abstract syntax (HOAS)~\cite{Pfenning1988} requires additional axioms to work with contexts.
 
-Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, i.e.\ induction principles on functions to work with HOAS encodings.
+Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, \ie, induction principles on functions to work with HOAS encodings.
 Some proof assistants such as Abella~\cite{Baelde2014} support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging~\cite{Momigliano2012}.
@@
 % Start of related work
-\cite{Maksimovic2015} does not have name restriction, thus dodging the challenge of scope extrusion.
+\cite{Maksimovic2015} does not have name restriction, thus dodging the challenge of scope extrusion. \dz{Not sure if relevant to mention.}

### 2024-01-31 short-paper/lin.tex
Edits and comments

@@
 
-In mechanized meta-theory addressing linearity means choosing an
+In mechanised meta-theory, addressing linearity means choosing an
 appropriate representation of the linear context. The crux of the matter
-is \emph{context splitting} as showcased by the typing rule for parallel composition.
+is \emph{context splitting}, as showcased by the typing rule for parallel composition.
 % the introduction rule of the tensor.
-While a linear context is best seen as a multi-set, most proof
-assistants have better support for list processing.  Although the
-latter representation is intuitive, it is overly concrete and may
-require establishing dozens and dozens of technical lemmas that are
-foreign to the mathematics of the problem under study, say proving
+While a linear context is perhaps best seen as a multi-set, most proof
+assistants have better support for list processing.  The
+latter representation is intuitive, but may require establishing a 
+large number of technical lemmata that are
+orthogonal to the mathematics of the problem under study, say proving
 type preservation for session types. In this space, several designs
-are possible: one can label occurrences of resources constraining
-their usage, e.g.~\cite{CicconeP20}, or \alb{something --- ask
-  Francisco} as in \cite{Castro2020}. Some authors instead impose a
-multiset structure over lists, see~\cite{ChaudhuriLR19,Danielsson12}:
-still, the effort in developing the infrastructure is significant.
+are possible: one can label occurrences of resources to constrain
+their usage (\eg~\cite{CicconeP20}) or \alb{something --- ask Francisco} as in \cite{Castro2020}. Some authors instead impose a
+multiset structure over lists (\eg~\cite{ChaudhuriLR19,Danielsson12});
+still, the effort required to develop the infrastructure is significant.
 
@@
 encodings can be refined to be intrinsically-typed, if the proof
-assistant does support dependent types, see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}.
+assistant does support dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
 
 For something completely different, one could adopt a
-\emph{sub-structural} meta-logical framework: here it is the framework
-itself that handles resource distribution and the users just have to map
+\emph{sub-structural} meta-logical framework: here the framework
+itself handles resource distribution, and users need only map
 their linear operations to the ones offered by the framework. In fact,
 the linear function space will handle all context operations
-implicitly, including splitting and substitution. One such a framework
-is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08}, see the encoding of
-session types in~\cite{Bock2016}. Unfortunately, \emph{Celf} does not
+implicitly, including splitting and substitution. One such framework
+is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of
+session types in~\cite{Bock2016}). Unfortunately, \emph{Celf} does not
 yet fully support the verification of meta-theoretic properties. A compromise is the so called
 \emph{two-level} approach, where one encodes a sub-structural
-specification logic in a mainstream proof assistant and then use that
-logic to state and prove linear properties, see for a recent
-example~\cite{Felty:MSCS21}.
+specification logic in a mainstream proof assistant and then uses that
+logic to state and prove linear properties (for a recent
+example, see~\cite{Felty:MSCS21}).
 

### 2024-01-31 short-paper/short-paper.tex
first draft of related work abut linearity
Add first draft piece about binders.
Edits and comments

@@
 \begin{abstract}
-  The POPLMark challenge and its follow-up POPLMark Re\-loaded sparked a flurry of
+  POPLMark and POPLMark Re\-loaded sparked a flurry of
   work on machine-checked proofs, and fostered the adoption of proof
-  mechanisation in programming language research. However, both POPLMark and
-  POPLMark Reloaded were purposely limited in scope, with benchmark problems
-  that do not address concurrency-related issues.
+  mechanisation in programming language research. Both challenges were purposely limited in scope, and their benchmark problems do not address concurrency-related issues.
   %
-  For this reason, we propose a new collection of benchmark problems with a
-  focus on the challenges and issues that typically arise when mechanising
-  message-passing concurrency using process calculi.\luca{wondering whether it would make sense to say ``\dots using formal models of concurrent and distributed programming languages in the form of process calculi'' or something similar to stress that process calculi are not the subject of the challenges, but ``only'' an appropriate tool for this kind of endeavor}  Our benchmark problems
+  For this reason, we propose a new collection of benchmark problems focused on the difficulties that typically arise when mechanising
+  message-passing concurrency using process calculi.\luca{wondering whether it would make sense to say ``\dots using formal models of concurrent and distributed programming languages in the form of process calculi'' or something similar to stress that process calculi are not the subject of the challenges, but ``only'' an appropriate tool for this kind of endeavor} \dz{Maybe something like, ``\ldots that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.''?} Our benchmark problems
   address three key topics: linearity, scope extrusion, and coinductive
@@
 Several papers in this realm already include mechanised proofs
-(e.g.~\cite{Castro-Perez2021,Castro2020,lmcs:9985,
+(\eg~\cite{Castro-Perez2021,Castro2020,lmcs:9985,
   DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or discuss proof mechanisation techniques
-(e.g.~\cite{Bengtson2016, Gay2001,
+(\eg~\cite{Bengtson2016, Gay2001,
   Thiemann2019, DBLP:conf/forte/ZalakainD21}).  Our
@@
 
-\section{Motivation}\label{sec:motivation}
-
 \section{Design of the Benchmark}\label{sec:design-discussion}
@@
 as per design goal \ref{item:goal-reusability},
-our problems are drawn from the basic metatheory of process calculi (without requiring the development of new theory), and focus on well-known theorems
+our problems are drawn from the basic meta-theory of process calculi (without requiring the development of new theory), and focus on well-known theorems
 and results that involve interesting proof techniques that may be reused in further work.
@@
 For example, techniques like higher-order abstract syntax, or well-scoped
-De\ Bruijn indices do not have the concept of free variables and thus encoding this challenge with them requires some attention.
+de\ Bruijn indices do not have the concept of free variables and thus encoding this challenge with them requires some attention.
 
@@
 Solutions to our challenges do not need to strictly follow the definitions and lemmata set out in the problem text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.
+
+\section{Possible approaches}\label{sec:approaches}
+
+\alb{Some introductory text to lead to the literature review. \sout{We may want to move this after the next Section, if we want to mirror the 3 challnges}}
+\dz{Agreed, I moved it for now.} \\
+\dz{Should we not also include a subsection on coinduction?} \\
+\dz{I changed the name of this section to ``Possible approaches'' from ``Motivation''. Thoughts?}
+
+\subsubsection{Linearity.}
+\input{lin}
+
+\subsubsection{Binders.}
+\input{binders}
 
@@
 further challenges could be multiparty session types and
-choreographies, as their meta-theory includes aspects -- \eg well-formedness conditions on global types, partiality of end-point projection function, etc. --
+choreographies, as their meta-theory includes aspects -- \eg, well-formedness conditions on global types, partiality of end-point projection function, etc. --
 that we do not cover here. Also, one could easily

### 2024-01-31 short-paper/short-paper.tex
Edits and comments
⠀
written paragraph on coinduction for recursion

@@
   For this reason, we propose a new collection of benchmark problems focused on the difficulties that typically arise when mechanising
-  message-passing concurrency using process calculi.\luca{wondering whether it would make sense to say ``\dots using formal models of concurrent and distributed programming languages in the form of process calculi'' or something similar to stress that process calculi are not the subject of the challenges, but ``only'' an appropriate tool for this kind of endeavor} \dz{Maybe something like, ``\ldots that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.''?} Our benchmark problems
+  message-passing concurrency using process calculi.\luca{I like Daniel's proposal below} \dz{Maybe something like, ``\ldots that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.''?} Our benchmark problems
   address three key topics: linearity, scope extrusion, and coinductive
@@
 issues that arise when mechanising message-passing concurrency using process 
-calculi.\luca{see comment in abstract}
+calculi.
 
@@
 
-\section{Possible approaches}\label{sec:approaches}
+\section{Existing approaches}\label{sec:approaches}
 
@@
 \dz{Agreed, I moved it for now.} \\
-\dz{Should we not also include a subsection on coinduction?} \\
-\dz{I changed the name of this section to ``Possible approaches'' from ``Motivation''. Thoughts?}
+\dz{I changed the name of this section to ``Existing approaches'' from ``Motivation''. Thoughts?}
 
@@
 \input{binders}
+
+\subsubsection{Coinduction.} \todo{}
 
@@
 support it.%
-%
-\luca{I would argue that support for recursive types and processes may also
-impact the coinduction challenge. Indeed, some type systems with support for
-recursion are more easily expressed \emph{coinductively}, by allowing possibly
-infinite typing derivation. This application of coinduction is different from
-that intended in the challenge, where it's focused on expressing notions of
-process equivalence, and yet it is increasingly popular (see the related work on
-infinitary linear logics) and occasionally seen with skepticism, which would
-make a good case for formalization.}
 
@@
 and trace equivalence.
+%
+\luca{Coinduction may also play a relevant role in supporting recursive
+processes and session types. Indeed, it is often the case that session type
+systems (and logics) with recursive session types (and propositions) are
+naturally expressed in \emph{infinitary form} by interpreting their typing (and
+proof) rules coinductively. I CAN ADD ONE OR TWO REFERENCES IF YOU FEEL IT'S APPROPRIATE}
 

### 2024-02-01 short-paper/short-paper.tex
comments to sec 1

@@
 \begin{abstract}
+  %% AM: implemented DZ
+% when mechanising  message-passing concurrency using process calculi.
+
   POPLMark and POPLMark Re\-loaded sparked a flurry of
@@
   %
-  For this reason, we propose a new collection of benchmark problems focused on the difficulties that typically arise when mechanising
-  message-passing concurrency using process calculi.\luca{I like Daniel's proposal below} \dz{Maybe something like, ``\ldots that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.''?} Our benchmark problems
-  address three key topics: linearity, scope extrusion, and coinductive
-  reasoning.  The goal of this new challenge is to clarify, compare, and advance
-  the state of the art, fostering the adoption of proof mechanisation in future
-  research on message-passing concurrency.
+  For this reason, we propose a new collection of benchmark problems
+  focused on the difficulties that typically arise when mechanising
+  formal models of concurrent and distributed programming languages,
+  such as process calculi.  Our benchmark problems address three key
+  topics: linearity, scope extrusion, and coinductive reasoning.  The
+  goal of this new challenge is to clarify, compare, and advance the
+  state of the art, fostering the adoption of proof mechanisation in
+  future research on message-passing concurrency.
 
@@
 Its follow-up challenge, POPLMark Reloaded~\cite{POPLMarkReloaded}, 
-facilitated the development of best practices and tool support for proofs 
+% facilitated the development of best practices and tool support
+aimed to run the same program for proofs 
 using logical relations. These challenges were deliberately limited in scope, 
@@
 
+\alb{Is there a sharp difference among those papers?}
 Several papers in this realm already include mechanised proofs
@@
 tools remains a significant challenge with a limited understanding of their 
-respective tradeoffs.  This often leads researchers toward a trial-and-error approach, 
+respective trade-offs.  This often leads researchers toward a trial-and-error approach, 
 resulting in sub-optimal and ad-hoc solutions. In turn, this increases the overall effort 
@@
 just one of these aspects at once; we see the combination of techniques
-as a next step beyond the scope of this challenge (we discuss this further in
-\Cref{sec:going-beyond}).
+as a next step beyond the scope of this challenge, as  discussed  further in
+\Cref{sec:going-beyond}.
 % Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate, evaluate, and compare the many tools and techniques available.
@@
 
-Like the authors of the POPLMark challenges, we seek to
+Similarly to the authors of the POPLMark challenges, we seek to
 answer several questions:
+\alb{We oscillate between concurrency in general and process calculi, here and before}
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
@@
 transition system semantics.
-Linearity is of course interesting in many other subfields of programming languages, but we have chosen to base our challenge on a session type system primarily due to the common use of session types in concurrency theory.
+Linearity is of course interesting in many other sub-fields of programming languages, but we have chosen to base our challenge on a session type system primarily due to the common use of session types in concurrency theory.
 Inspired by Vasconcelos~\cite{Vasconcelos2012}, we use a syntax where
@@
 
-\alb{Some introductory text to lead to the literature review. \sout{We may want to move this after the next Section, if we want to mirror the 3 challnges}}
+\alb{Some introductory text to lead to the literature review. \sout{We may want to move this after the next Section, if we want to mirror the 3 challenges}}
 \dz{Agreed, I moved it for now.} \\

### 2024-02-01 short-paper/short-paper.tex
minor edits for style

@@
 
-As with the POPLMark and POPLMark Reloaded challenges, this challenge does not aim to cover every possible
-problem in the area. Our benchmark can be extended in two
-dimensions: further expanding and combining the current challenges,
-and addressing features not covered by the current challenges.
+Similar to the POPLMark and its Reloaded counterpart, this challenge
+does not encompass all potential problems in the field. Our benchmark
+is extendable in two aspects: it can further amalgamate and broaden
+the existing challenges, and tackle aspects not yet covered by them.
+
+% As with the POPLMark and POPLMark Reloaded challenges, this challenge does not aim to cover every possible
+% problem in the area. Our benchmark can be extended in two
+% dimensions: further expanding and combining the current challenges,
+% and addressing features not covered by the current challenges.
 
@@
 choreographies, as their meta-theory includes aspects -- \eg, well-formedness conditions on global types, partiality of end-point projection function, etc. --
-that we do not cover here. Also, one could easily
+that we do not address here. Also, one could easily
 design a challenge that goes beyond our coinduction challenge, by

### 2024-02-01 short-paper/short-paper.tex
Add Alberto's comments.
Co-authored-by: Alberto Momigliano <momigliano@di.unimi.it>

@@
 Like the POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
-applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.
-Still, these (and other) applications need the basic techniques that our challenge problems emphasize:
+applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.\alb{Do we need all those citations?}  Still, these (and other) applications need the basic techniques that our challenge problems emphasize:
 as per design goal \ref{item:goal-reusability},
@@
 transition system semantics.
-Linearity is of course interesting in many other sub-fields of programming languages, but we have chosen to base our challenge on a session type system primarily due to the common use of session types in concurrency theory.
+Linearity is of course interesting in many other sub-fields of programming languages, but we have chosen to base our challenge on a session type system primarily due to the common use of session types in concurrency theory.\alb{This is weak and somewhat tautological, as a motivation}
 Inspired by Vasconcelos~\cite{Vasconcelos2012}, we use a syntax where
@@
 This is the challenge most closely related to the POPLMark
-challenge since it concerns the properties of binders in restrictions.
-It is not obvious that the techniques used to solve the original challenges about binders can be easily extended to reasoning about scope extrusion.
+challenge since it concerns the properties of binders in restrictions.\alb{I would argue also wrt input}
+It is not obvious that the techniques used to solve the original challenges about binders can be easily extended to reasoning about scope extrusion.\alb{I would say not obvious what the pros and cons are, since in fact those techniques (DB, locally nameless, HOAS, nominal) have been used.}
 For example, techniques like higher-order abstract syntax, or well-scoped
@@
 
-Our coinduction challenge concerns strong barbed bisimulation and congruence.
-Though weak barbed congruence is a more common behavioural equivalence for \(\pi\)-terms, here we prefer strong equivalences to avoid the need to abstract over the number of internal transitions, % when considering the strong notions of equivalence,
+Our coinduction challenge concerns strong barbed bisimulation and congruence.\alb{we motivate why strong vs weak, but not why barbed}
+Though weak barbed congruence is a more common behavioural equivalence for \(\pi\)-terms,\alb{?
+why terms?} here we prefer strong equivalences to avoid the need to abstract over the number of internal transitions, % when considering the strong notions of equivalence,
 thus simplifying the theory.

### 2024-02-01 short-paper/short-paper.tex
Reorganize paper to remove "existing mechanisations" section.
The idea is to separate the content of the section such that the
motivational parts for the benchmark is merged into the introduction and
the parts that motivate why we chose these challenges are merged into the
section about the design of the benchmark.

For now, I have just moved everything into the introduction to clean
up later.

@@
 \section{Introduction}
-By establishing a common ground for systematically evaluating mechanisation techniques, 
-the POPLMark challenge~\cite{POPLMark} accelerated progress in 
-the development of proof assistants, libraries, and best practices. In doing so, it 
-spearheaded a shift in the programming language research community towards 
-publications that include mechanised proofs. 
-Its follow-up challenge, POPLMark Reloaded~\cite{POPLMarkReloaded}, 
-% facilitated the development of best practices and tool support
-aimed to run the same program for proofs 
-using logical relations. These challenges were deliberately limited in scope, 
-while their authors called for future investigations into such areas as 
-coinduction and linear environments. In this spirit, we propose a 
-new collection of benchmark problems specifically designed to address typical 
-issues that arise when mechanising message-passing concurrency using process 
-calculi.
+The POPLMark challenge~\cite{POPLMark} accelerated progress in the development of proof assistants, libraries, and best practices by establishing a common ground for systematically evaluating mechanisation techniques.
+In doing so, it spearheaded a shift in the programming language research community towards publications that include mechanised proofs.
+POPLMark Reloaded~\cite{POPLMarkReloaded} proposed a similar program for proofs using logical relations.
+These challenges were deliberately limited in scope, while their authors called for future investigations into such areas as coinduction and linear environments.
+In this spirit, we propose a new collection of benchmark problems specifically designed to address typical issues that arise when mechanising message-passing concurrency using process calculi.
 
 \alb{Is there a sharp difference among those papers?}
-Several papers in this realm already include mechanised proofs
-(\eg~\cite{Castro-Perez2021,Castro2020,lmcs:9985,
-  DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or discuss proof mechanisation techniques
-(\eg~\cite{Bengtson2016, Gay2001,
-  Thiemann2019, DBLP:conf/forte/ZalakainD21}).  Our
-experience, however, is that choosing the appropriate proof mechanisation techniques and 
+Several papers in this realm already include mechanised proofs (\eg~\cite{Castro-Perez2021, Castro2020, lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or discuss proof mechanisation techniques (\eg~\cite{Bengtson2016, Gay2001, Thiemann2019, DBLP:conf/forte/ZalakainD21}).
+Our experience, however, is that choosing the appropriate proof mechanisation techniques and 
 tools remains a significant challenge with a limited understanding of their 
@@
 required for proof mechanisation, and hampers the reuse of techniques and results.
-%
+
 It is therefore our view that the research community in message-passing
@@
 of challenge problems.
+To this end, our benchmark problems (discussed in detail in \cref{app:challenges}) are designed to focus on three key aspects which typically cause difficulties when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning} for infinite behaviours.
+These are certainly not the only factors that may cause issues, but they are fundamental to concurrency theory, and emerge in most mechanisations.
 
-To this end, our benchmark problems (discussed in detail in \cref{app:challenges}) are designed to focus on three key aspects which typically cause difficulties 
-when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, 
-and \emph{coinductive reasoning} for infinite behaviours. These are certainly not the only factors that may cause issues, but they
-are fundamental to concurrency theory, and emerge in most mechanisations.
+\todo{Rewrite this section to be more coherent.}
+
+\paragraph{Linearity.}
+In mechanised meta-theory, addressing linearity means choosing an appropriate representation of the linear context.
+The crux of the matter is \emph{context splitting}, as showcased by the typing rule for parallel composition.
+While a linear context is perhaps best seen as a multi-set, most proof assistants have better support for list processing.
+The latter representation is intuitive, but may require establishing a large number of technical lemmata that are orthogonal to the mathematics of the problem under study, say proving type preservation for session types.
+In this space, several designs are possible: one can label occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}) or \alb{something --- ask Francisco} as in \cite{Castro2020}.
+Some authors instead impose a multiset structure over lists (\eg~\cite{ChaudhuriLR19,Danielsson12}); still, the effort required to develop the infrastructure is significant.
+
+A different approach is to bypass the problem of context splitting by adopting familiar ideas from algorithmic linear type checking: this is known as \emph{typing with leftovers}, as exemplified in~\cite{DBLP:conf/forte/ZalakainD21}.
+Whatever the choice, list-based encodings can be refined to be intrinsically-typed, if the proof assistant does support dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
+
+For something completely different, one could adopt a \emph{sub-structural} meta-logical framework: here the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
+In fact, the linear function space will handle all context operations implicitly, including splitting and substitution.
+One such framework is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session types in~\cite{Bock2016}).
+Unfortunately, \emph{Celf} does not yet fully support the verification of meta-theoretic properties.
+A compromise is the so called \emph{two-level} approach, where one encodes a sub-structural specification logic in a mainstream proof assistant and then uses that logic to state and prove linear properties (for a recent example, see~\cite{Felty:MSCS21}).
+
+\paragraph{Scope extrusion.}
+The encoding of binding constructs has been extensively studied for type systems based on $\lambda$-calculi, but process calculi present unique challenges.
+$\pi$-calculi typically include several different binding constructs: inputs bind a received name or value, recursive processes bind recursion variables, and restrictions bind names.
+The first two act similarly to the binders known from $\lambda$-calculi, but restrictions are more challenging due to scope extrusion.
+
+The case study \cite{AmbalLS21} compares four approaches to encoding binders in Coq as a first step towards developing tools for working with higher-order process calculi.
+They found that working directly with de Bruijn indices was easiest and shortest since the existing approaches developed for $\lambda$-calculus binders do not work well with scope extrusion.
+Specifically, the locally nameless representation~\cite{Chargueraud2012} cannot avoid direct manipulation of de Bruijn indices when defining the semantics of scope extrusion; cofinite quantification provides no benefits when working under binders; nominal approaches~\cite{Pitts2003} require explicitly giving and validating sets of free names during scope extrusion; and higher-order abstract syntax (HOAS)~\cite{Pfenning1988} requires additional axioms to work with contexts.
+
+Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, \ie, induction principles on functions to work with HOAS encodings.
+Some proof assistants such as Abella~\cite{Baelde2014} support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging~\cite{Momigliano2012}.
+
+% Start of related work
+\cite{Maksimovic2015} does not have name restriction, thus dodging the challenge of scope extrusion.
+
+\paragraph{Coinduction.} \todo{}
+
 The mechanisation of novel research results often requires addressing more than
 just one of these aspects at once; we see the combination of techniques
-as a next step beyond the scope of this challenge, as  discussed  further in
-\Cref{sec:going-beyond}.
-% Our challenge problems are designed to stimulate progress on answering these questions by providing simple environments in which to demonstrate, evaluate, and compare the many tools and techniques available.
-% While the challenge problems themselves are thus not of scientific interest, we expect that the best practices and mechanisations developed by solving the challenge problems will also be useful for future research.
+as a next step beyond the scope of this challenge, as discussed further in \Cref{sec:going-beyond}.
 
@@
 \section{Design of the Benchmark}\label{sec:design-discussion}
+\todo{Add references to support our claims and decisions in this section.}
 
@@
 Solutions to our challenges do not need to strictly follow the definitions and lemmata set out in the problem text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.
-
-\section{Existing approaches}\label{sec:approaches}
-
-\alb{Some introductory text to lead to the literature review. \sout{We may want to move this after the next Section, if we want to mirror the 3 challenges}}
-\dz{Agreed, I moved it for now.} \\
-\dz{I changed the name of this section to ``Existing approaches'' from ``Motivation''. Thoughts?}
-
-\subsubsection{Linearity.}
-\input{lin}
-
-\subsubsection{Binders.}
-\input{binders}
-
-\subsubsection{Coinduction.} \todo{}
 

### 2024-02-02 short-paper/short-paper.tex
Further reorganize paper.
I have added a comment before each section with a general outline of
the main points that I suggest could be in that section.

@@
 \begin{abstract}
-  %% AM: implemented DZ
-% when mechanising  message-passing concurrency using process calculi.
-
-  POPLMark and POPLMark Re\-loaded sparked a flurry of
-  work on machine-checked proofs, and fostered the adoption of proof
-  mechanisation in programming language research. Both challenges were purposely limited in scope, and their benchmark problems do not address concurrency-related issues.
+  POPLMark and POPLMark Reloaded sparked a flurry of work on machine-checked proofs, and fostered the adoption of proof mechanisation in programming language research.
+  Both challenges were purposely limited in scope, and their benchmark problems do not address concurrency-related issues.
   %
-  For this reason, we propose a new collection of benchmark problems
-  focused on the difficulties that typically arise when mechanising
-  formal models of concurrent and distributed programming languages,
-  such as process calculi.  Our benchmark problems address three key
-  topics: linearity, scope extrusion, and coinductive reasoning.  The
-  goal of this new challenge is to clarify, compare, and advance the
-  state of the art, fostering the adoption of proof mechanisation in
-  future research on message-passing concurrency.
+  We propose a new collection of benchmark problems focused on the difficulties that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.
+  Our benchmark problems address three key topics: linearity, scope extrusion, and coinductive reasoning.
+  The goal of this new challenge is to clarify, compare, and advance the state of the art, fostering the adoption of proof mechanisation in future research on message-passing concurrency.
 
@@
 
+% - Introduction
+%   - POPLMark accelerated progress in mechanisation tools for PL
+%     - POPLMark Reloaded tried to do the same for logical relations
+%     - But they were deliberately limited in scope to not include process calculi
+%   - We propose a new collection of benchmark problems for process calculi
+%     - Why not concurrency in general? (after this point, only mention process calculi)
+%     - Why is there a problem with process calculus mechanisations?
+%     - Our experience says there is, and we will cite other experiences below
+%   - We have noticed three key issues
+%     - They are not covered by POPLMark
+%     - There are fewer techniques than for binders
+%     - Others also mention them as problems: cite and explain references briefly for each challenge topic
+%   - Our goals are to determine the state of the art, develop best practices and tutorials, and identify improvements to proof assistant technology that would materially help the situation
 \section{Introduction}
@@
 These challenges were deliberately limited in scope, while their authors called for future investigations into such areas as coinduction and linear environments.
+
 In this spirit, we propose a new collection of benchmark problems specifically designed to address typical issues that arise when mechanising message-passing concurrency using process calculi.
+\todo{Why just process calculi and not concurrency in general?}
+While many efforts to mechanise results about process calculi have been completed (as we will discuss further below), our experience is that choosing appropriate proof mechanisation techniques and tools remains a significant challenge with a limited understanding of the trade-offs.
+This often leads researchers toward a trial-and-error approach, resulting in sub-optimal and ad-hoc solutions.
+In turn, this increases the overall effort required for proof mechanisation, and hampers the reuse of techniques and results.
 
-\alb{Is there a sharp difference among those papers?}
-Several papers in this realm already include mechanised proofs (\eg~\cite{Castro-Perez2021, Castro2020, lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or discuss proof mechanisation techniques (\eg~\cite{Bengtson2016, Gay2001, Thiemann2019, DBLP:conf/forte/ZalakainD21}).
-Our experience, however, is that choosing the appropriate proof mechanisation techniques and 
-tools remains a significant challenge with a limited understanding of their 
-respective trade-offs.  This often leads researchers toward a trial-and-error approach, 
-resulting in sub-optimal and ad-hoc solutions. In turn, this increases the overall effort 
-required for proof mechanisation, and hampers the reuse of techniques and results.
+%\alb{Is there a sharp difference among those papers?}
+%Several papers in this realm already include mechanised proofs (\eg~\cite{Castro-Perez2021, Castro2020, lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or discuss proof mechanisation techniques (\eg~\cite{Bengtson2016, Gay2001, Thiemann2019, DBLP:conf/forte/ZalakainD21}).
 
-It is therefore our view that the research community in message-passing
-concurrency (and process calculi, more specifically) will benefit from a standardised set
-of challenge problems.
+It is therefore our view that the research community in message-passing concurrency (and more specifically, process calculi) will benefit from a standardised set of challenge problems.
 To this end, our benchmark problems (discussed in detail in \cref{app:challenges}) are designed to focus on three key aspects which typically cause difficulties when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning} for infinite behaviours.
-These are certainly not the only factors that may cause issues, but they are fundamental to concurrency theory, and emerge in most mechanisations.
-
-\todo{Rewrite this section to be more coherent.}
+The mechanisation of novel research results often requires addressing more than just one of these aspects at once; we see the combination of techniques as a next step beyond the scope of this challenge, as discussed further in \Cref{sec:going-beyond}.
+These are certainly not the only factors that may cause issues, but they are fundamental to concurrency theory, and emerge in most mechanisations, as we will see presently while discussing the issues presented by the three key aspects.
 
@@
 
-The mechanisation of novel research results often requires addressing more than
-just one of these aspects at once; we see the combination of techniques
-as a next step beyond the scope of this challenge, as discussed further in \Cref{sec:going-beyond}.
-
+\subsection{Goals and how to contribute}
 Similarly to the authors of the POPLMark challenges, we seek to
@@
 \item\label{item:rq1} What is the current state of the art in mechanising process calculi?
-\item\label{item:rq2} Which techniques and best practices can be recommended when starting formalisation projects within concurrency theory?
+\item\label{item:rq2} Which techniques and best practices can be recommended when starting formalisation projects involving process calculi?
 \item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
@@
 
-On the whole, this benchmark seeks to advance three main goals:
-
+To provide a framework in which to answer these questions, our benchmark is designed to satisfy three main goals:
 \begin{enumerate}[label=\textbf{(G\arabic*)},leftmargin=10mm]
@@
 \end{enumerate}
-We also aim at strengthening the culture of mechanisation, by rallying the
-community to collaborate on exploring and developing new tools and techniques.
+We also aim at strengthening the culture of mechanisation, by rallying the community to collaborate on exploring and developing new tools and techniques.
 
@@
 %
-In the longer term, we expect to use the website for promoting best practices,
-tutorials and guidelines derived from solutions to our challenges.
-We encourage anyone interested to try the challenges using their
-favourite tools and techniques, and to send us their solutions.
+In the longer term, we expect to use the website for promoting best practices, tutorials and guidelines derived from solutions to our challenges.
+We encourage anyone interested to try the challenges using their favourite tools and techniques, and to send us their solutions.
 
+% - Design of the benchmark
+%   - We first list some general design remarks, then for each challenge
+%   - General remarks
+%     - We do not aim to be comprehensive (list other topics)
+%     - However, we believe the best practices developed could also be of use in those topics
+%     - We base our challenges on basic textbook theory with interesting proof techniques
+%     - The challenges are independent (maybe note ``failure'' of POPLMark that everyone only did the first part)
+%     - The challenges are as minimal as possible
+%   - Linearity
+%     - Type preservation for session types
+%     - The main issue is context splitting
+%     - Session types are pervasive in the field and simple to understand
+%     - Binding two names makes duality obvious
+%     - Brief overview of approaches and issues with them
+%   - Scope extrusion
+%     - Process calculi have several binders, but only restriction is different from ``normal'' ones
+%     - The main issue is scope extrusion
+%     - The pros and cons of established techniques for handling binders are different for process calculi, and it is unclear how exactly
+%     - Brief overview of approaches and issues with them
+%   - Coinduction
+%     - Strong barbed bisimulation and congruence
+%     - The main issue is the encoding of coinductive up-to techniques
+%     - Strong equivalences to simplify (we don't have to abstract over internal transitions)
+%     - Why barbed equivalences?
+%     - No delegation to simplify
+%     - Brief overview of approaches and issues with them
+%   - Evaluation criteria
+%     - We care not only about solutions, but about experiences and comparisons
+%     - Mechanisation overhead
+%     - Adequacy of the formal statements
+%     - Cost of entry for tools and techniques
+%     - Slight changes are OK if they are justified and adequate
 \section{Design of the Benchmark}\label{sec:design-discussion}
@@
 
-In this section, we outline the factors considered in designing the
-benchmark challenges. We begin with some general remarks, then
-describe the individual design considerations for each challenge problem,
-and the evaluation criteria.
-% We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
+In this section, we outline the factors considered in designing the benchmark challenges.
+We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the evaluation criteria for solutions.
 
@@
 Like the POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
-applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.\alb{Do we need all those citations?}  Still, these (and other) applications need the basic techniques that our challenge problems emphasize:
-as per design goal \ref{item:goal-reusability},
-our problems are drawn from the basic meta-theory of process calculi (without requiring the development of new theory), and focus on well-known theorems
-and results that involve interesting proof techniques that may be reused in further work.
+applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.\alb{Do we need all those citations?}
+Still, these (and other) applications need the basic techniques that our challenge problems emphasize: as per design goal \ref{item:goal-reusability}, our problems are drawn from the basic meta-theory of process calculi (without requiring the development of new theory), and focus on well-known theorems and results that involve interesting proof techniques that may be reused in further work.
 To achieve design goal \ref{item:goal-comperison-accessibility},
@@
 goal \ref{item:goal-tutorials}.
-
-% While our challenges require reasoning about binders, which was the main topic of the original POPLMark, they also require other techniques.
-% Our challenges also require reasoning about logical relations which goes above the reasoning involved in POPLMark Reloaded.
-
-% Our challenge problems do not include reasoning about encodings. While
-% this kind of reasoning is common in the field, it does not require any
-% different reasoning techniques in a concurrency setting than in any
-% other setting.
 
@@
 \subsubsection{Coinduction.}
-%\fxerror{
-% ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics?
-%}
-
 Our coinduction challenge concerns strong barbed bisimulation and congruence.\alb{we motivate why strong vs weak, but not why barbed}
 Though weak barbed congruence is a more common behavioural equivalence for \(\pi\)-terms,\alb{?
-why terms?} here we prefer strong equivalences to avoid the need to abstract over the number of internal transitions, % when considering the strong notions of equivalence,
-thus simplifying the theory.
+why terms?} here we prefer strong equivalences to avoid the need to abstract over the number of internal transitions, thus simplifying the theory.
 We have chosen not to include delegation in the coinduction challenge since it is orthogonal to our primary aim of exploring coinductive proof techniques.
@@
 
+% - Going beyond the challenge problems
+%   - We do not aim to be comprehensive, but the techniques should also be useful for other things
+%   - Our challenges are independent, but they can also be combined
+%   - Our challenges are basic textbook theory, so they can form the basis for many extensions
+%   - Our aim is to see a future where ``the basics'' are as trivial in mechanisations as they are on paper
 \section{Going beyond the challenge problems}\label{sec:going-beyond}
+Similarly to POPLMark and its Reloaded counterpart, our challenges do not encompass all issues in the field.
+Our benchmark is extendable in two aspects: one can further amalgamate and broaden the existing challenges, and tackle aspects not covered by the benchmark at all.
+All of these extensions will still need to handle at least one of the key aspects of our challenges.
 
-Similar to the POPLMark and its Reloaded counterpart, this challenge
-does not encompass all potential problems in the field. Our benchmark
-is extendable in two aspects: it can further amalgamate and broaden
-the existing challenges, and tackle aspects not yet covered by them.
-
-% As with the POPLMark and POPLMark Reloaded challenges, this challenge does not aim to cover every possible
-% problem in the area. Our benchmark can be extended in two
-% dimensions: further expanding and combining the current challenges,
-% and addressing features not covered by the current challenges.
-
-\vspace{-1mm}%
 \paragraph{Extending and combining the current challenges.}
-While the proposed challenges try to minimise the overlap of the techniques
-and aspects they explore (linearity, scope extrusion, coinduction), most
-research on concurrency theory combines some or all of these. A second round of formalisations might
-entail combining the features of the challenges.
-We consider the following three extensions:
+\todo{This paragraph is very long for how relatively little it says. Maybe remove/cut down?}
+While we have tried to minimise the overlap of the techniques and aspects our challenges explore (linearity, scope extrusion, coinduction), most research on concurrency theory combines some or all of these.
+A second round of formalisations might entail combining the features of the challenges.
+We consider the following three extensions to be obvious places to start:
 \emph{(1)} the addition of choice to the type level and the process level,
 \emph{(2)} the addition of recursion and (for typed systems) recursive types,
-and \emph{(3)} the
-addition of channel delegation to those calculi that do not already
-support it.%
+and \emph{(3)} the addition of channel delegation to those calculi that do not already support it.%
 
-\vspace{-1mm}%
 \paragraph{Addressing new calculi and new features.}
-Extending in this direction
-means proposing new challenges covering different features of
-process calculi, or different proof mechanisation features.
-Some interesting aspects of message-passing calculi to be explored in
-further challenges could be multiparty session types and
-choreographies, as their meta-theory includes aspects -- \eg, well-formedness conditions on global types, partiality of end-point projection function, etc. --
-that we do not address here. Also, one could easily
-design a challenge that goes beyond our coinduction challenge, by
-exploring different notions of bisimilarity (barbed, weak, \etc)
-and trace equivalence.
+\todo{Maybe cite some references in this section?}
+Extending in this direction means proposing new challenges covering different features of process calculi, or different proof mechanisation features.
+Some interesting aspects of message-passing calculi to be explored in further challenges could be multiparty session types and choreographies, as their meta-theory includes aspects -- \eg, well-formedness conditions on global types, partiality of end-point projection function, etc.
+-- that we do not address here.
+Also, one could easily design a challenge that goes beyond our coinduction challenge, by exploring different notions of bisimilarity (barbed, weak, \etc) and trace equivalence.
 %
@@
 
-On the mechanisation side, the present challenge proposes mechanising the
-proofs of theorems. An interesting avenue to explore would be to take advantage
-of other proof assistant features. For example, proof assistants can often
-produce certified code (via code extraction, or compiling and running their
-definitions). Additionally, while our current challenge is agnostic about proof
-automation, a more specific challenge could be designed with the objective of
-automating aspects of the proofs. Finally, a challenge could propose the
-integration with other formal reasoning tools, such as model checkers. These
-tools are in common use in the field; combining automated proofs with
-proof assistants offers the potential to ease the path towards larger proofs.
+On the mechanisation side, the present challenge proposes mechanising the proofs of theorems.
+An interesting avenue to explore would be to take advantage of other proof assistant features.
+For example, proof assistants can often produce certified code (via code extraction, or compiling and running their definitions).
+Additionally, while our current challenge is agnostic about proof automation, a more specific challenge could be designed with the objective of automating aspects of the proofs.
+Finally, a challenge could propose the integration with other formal reasoning tools, such as model checkers.
+These tools are in common use in the field; combining automated proofs with proof assistants offers the potential to ease the path towards larger proofs.
 
-Ultimately, the current challenges can be extended in several
-worthwhile directions, and we look forward to a future when they
-indeed are. We see the current challenge as setting
-the foundation for those future extensions. It is our expectation that
-the solutions of those future challenges will build on solutions
-to this one.
+Ultimately, the current challenges can be extended in several worthwhile directions, and we look forward to a future when they indeed are.
+We see the current challenge as setting the foundation for those future extensions.
+It is our hope and aim that our challenges will move us closer to a future where the key basic proof techniques are as easy to mechanise as they are to write on paper.
 

### 2024-02-02 challenges/linearity.tex
Edits and comments
Further edits

@@
 Our type system does not type processes directly, but instead focuses on the channels used in the process.
-The syntax of \emph{session types} \( S, T \), unrestricted type contexts \( \Gamma \) and linear type contexts \( \Delta \) is as follows:
+The syntax of \emph{session types} \( S, T \), unrestricted typing contexts \( \Gamma \) and linear typing contexts \( \Delta \) is as follows:
 \[
@@
 \]
-The \emph{end type} \( \Tend \) describes an endpoint on which no further interaction is possible.
+The \emph{end type} \( \Tend \) describes an endpoint over which no further interaction is possible.
 The \emph{base type} \( \Tbase \) describes base values.
@@
 the comma as split/union, overloaded to singletons, and \( \Cempty \) as the
-empty context. We extend the Barendregt convention to contexts, so that all
+empty context. We extend the Barendregt convention \todo{Source?} to contexts, so that all
 entries are distinct.  Note that the order in which information is added to a
@@
 \subsubsection{Challenge.}
-The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with some lemmas:
+The objective of this challenge is to prove subject reduction and type safety for our calculus with session types. We start with some lemmata:
 

### 2024-02-02 short-paper/short-paper.tex
Add a few comments.
Further edits

@@
 
-In this spirit, we propose a new collection of benchmark problems specifically designed to address typical issues that arise when mechanising message-passing concurrency using process calculi.
-\todo{Why just process calculi and not concurrency in general?}
-While many efforts to mechanise results about process calculi have been completed (as we will discuss further below), our experience is that choosing appropriate proof mechanisation techniques and tools remains a significant challenge with a limited understanding of the trade-offs.
+In this spirit, we propose a new collection of benchmark problems specifically designed to address typical issues that arise when mechanising formal models of concurrent and distributed programming languages in general, and process calculi in particular.
+While many efforts to mechanise results about such formalisms have been completed (as we will discuss further below), our experience is that choosing appropriate proof mechanisation techniques and tools remains a significant challenge with a limited understanding of their respective trade-offs.
 This often leads researchers toward a trial-and-error approach, resulting in sub-optimal and ad-hoc solutions.
@@
 
-It is therefore our view that the research community in message-passing concurrency (and more specifically, process calculi) will benefit from a standardised set of challenge problems.
+It is therefore our view that the research community in concurrency will benefit from a standardised set of challenge problems.
 To this end, our benchmark problems (discussed in detail in \cref{app:challenges}) are designed to focus on three key aspects which typically cause difficulties when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning} for infinite behaviours.
@@
 
+\todo{These paragraphs are still quite messy and unfocused.}
 \paragraph{Linearity.}
@@
 \section{Design of the Benchmark}\label{sec:design-discussion}
-\todo{Add references to support our claims and decisions in this section.}
+\todo{Add references to support our claims and decisions in this whole section.}
 
@@
 For our challenge on linear reasoning we have chosen a type preservation theorem
-for a session type system, which allows us to introduce linear
-reasoning with few definitions. Accordingly, we have chosen not to
-include channel delegation and to use a reduction instead of a
+for a system using session types, whose use are common in concurrency theory. This allows us to introduce linear
+reasoning with few definitions. Accordingly, we omit channel delegation and use a reduction instead of a
 transition system semantics.
-Linearity is of course interesting in many other sub-fields of programming languages, but we have chosen to base our challenge on a session type system primarily due to the common use of session types in concurrency theory.\alb{This is weak and somewhat tautological, as a motivation}
-Inspired by Vasconcelos~\cite{Vasconcelos2012}, we use a syntax where
-restriction binds two names. This presentation makes the duality
-of the two names obvious in the type system.
-We have chosen a simple notion of well-formedness for our challenge,
+Inspired by Vasconcelos~\cite{Vasconcelos2012}, we define a syntax where
+restriction binds two names, which highlights their duality within the type system.
+We have chosen a simple notion of well-formedness,
 and do not, for instance, consider deadlocks. While more
 sophisticated notions of well-formedness are interesting, the
-proofs of these properties would complicate the challenge, distracting from the linear aspects.
+proofs of these properties would distract from the core linear aspects of the challenge.
 
@@
 This is the challenge most closely related to the POPLMark
-challenge since it concerns the properties of binders in restrictions.\alb{I would argue also wrt input}
-It is not obvious that the techniques used to solve the original challenges about binders can be easily extended to reasoning about scope extrusion.\alb{I would say not obvious what the pros and cons are, since in fact those techniques (DB, locally nameless, HOAS, nominal) have been used.}
-For example, techniques like higher-order abstract syntax, or well-scoped
-de\ Bruijn indices do not have the concept of free variables and thus encoding this challenge with them requires some attention.
+challenge since it concerns the properties of binders in restrictions and input. 
+It is not obvious what the limitations and advantages are in leveraging the techniques used to solve the original challenges to reason about scope extrusion. 
+For example, techniques like higher-order abstract syntax or well-scoped
+de\ Bruijn indices lack the concept of free variables and thus encoding this challenge with them requires attention.
 
@@
 why terms?} here we prefer strong equivalences to avoid the need to abstract over the number of internal transitions, thus simplifying the theory.
-We have chosen not to include delegation in the coinduction challenge since it is orthogonal to our primary aim of exploring coinductive proof techniques.
+We exclude delegation from the coinduction challenge since it is orthogonal to our primary aim of exploring coinductive proof techniques.
 

### 2024-02-02 short-paper/short-paper.tex
minor edits and rewrites

@@
 answer several questions:
-\alb{We oscillate between concurrency in general and process calculi, here and before}
+%\alb{We oscillate between concurrency in general and process calculi, here and before}
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
@@
 Like the POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
-applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.\alb{Do we need all those citations?}
+applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.\alb{Do we need all those citations about encodings?}
 Still, these (and other) applications need the basic techniques that our challenge problems emphasize: as per design goal \ref{item:goal-reusability}, our problems are drawn from the basic meta-theory of process calculi (without requiring the development of new theory), and focus on well-known theorems and results that involve interesting proof techniques that may be reused in further work.
@@
 For our challenge on linear reasoning we have chosen a type preservation theorem
-for a system using session types, whose use are common in concurrency theory. This allows us to introduce linear
-reasoning with few definitions. Accordingly, we omit channel delegation and use a reduction instead of a
-transition system semantics.
-Inspired by Vasconcelos~\cite{Vasconcelos2012}, we define a syntax where
-restriction binds two names, which highlights their duality within the type system.
-We have chosen a simple notion of well-formedness,
-and do not, for instance, consider deadlocks. While more
-sophisticated notions of well-formedness are interesting, the
-proofs of these properties would distract from the core linear aspects of the challenge.
+for a system using session types.
+% , whose use are common in concurrency theory.
+This allows us to introduce linear reasoning with few definitions.
+With our objective focused on linearity, we forego channel delegation
+and opt for a reduction approach rather than employing a transition
+system semantics.  Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
+define a syntax where restriction binds two names,n highlighting
+their duality within the type system.  We have chosen a simple notion
+of well-formedness, and do not  consider
+deadlocks. While more sophisticated notions of well-formedness are
+interesting, the proofs of these properties would distract from the
+core linear aspects of the challenge.
 
 \subsubsection{Scope extrusion.}
-This is the challenge most closely related to the POPLMark
-challenge since it concerns the properties of binders in restrictions and input. 
-It is not obvious what the limitations and advantages are in leveraging the techniques used to solve the original challenges to reason about scope extrusion. 
-For example, techniques like higher-order abstract syntax or well-scoped
-de\ Bruijn indices lack the concept of free variables and thus encoding this challenge with them requires attention.
+This is the challenge most closely related to the POPLMark challenge
+since it concerns the properties of binders in restrictions and input.
+It is not obvious what the limitations and advantages are in
+leveraging the techniques used to solve the original challenges to
+reason about scope extrusion.  For example, techniques like
+higher-order abstract syntax or well-scoped de\ Bruijn indices lack
+the concept of free variables and thus encoding this challenge with
+them requires attention.
 
 \subsubsection{Coinduction.}
-Our coinduction challenge concerns strong barbed bisimulation and congruence.\alb{we motivate why strong vs weak, but not why barbed}
-Though weak barbed congruence is a more common behavioural equivalence for \(\pi\)-terms,\alb{?
-why terms?} here we prefer strong equivalences to avoid the need to abstract over the number of internal transitions, thus simplifying the theory.
-We exclude delegation from the coinduction challenge since it is orthogonal to our primary aim of exploring coinductive proof techniques.
+Our coinduction challenge concerns strong barbed bisimulation and
+congruence.\alb{we motivate why strong vs weak, but not why barbed}
+Though weak barbed congruence is a more common behavioural equivalence
+w.r.t.\ the \(\pi\)-calculus, here we prefer strong equivalences to avoid the
+need to abstract over the number of internal transitions, thus
+simplifying the theory.  We exclude delegation from the coinduction
+challenge since it is orthogonal to our primary aim of exploring
+coinductive proof techniques.
 
@@
 \paragraph{Extending and combining the current challenges.}
-\todo{This paragraph is very long for how relatively little it says. Maybe remove/cut down?}
-While we have tried to minimise the overlap of the techniques and aspects our challenges explore (linearity, scope extrusion, coinduction), most research on concurrency theory combines some or all of these.
-A second round of formalisations might entail combining the features of the challenges.
-We consider the following three extensions to be obvious places to start:
-\emph{(1)} the addition of choice to the type level and the process level,
-\emph{(2)} the addition of recursion and (for typed systems) recursive types,
-and \emph{(3)} the addition of channel delegation to those calculi that do not already support it.%
+% thanks, copilot
+
+We have aimed to reduce the overlap of the techniques and aspects that
+our challenges (linearity, scope extrusion, coinduction), despite
+their common use in concurrency theory research. A subsequent round
+could merge these features. We suggest starting with three extensions:
+\emph{(1)} adding choice to both type and process levels, \emph{(2)}
+introducing recursion and recursive types for typed systems, and
+\emph{(3)} incorporating channel delegation in calculi lacking it.
+%\todo{This paragraph is very long for how relatively little it says. Maybe remove/cut down?}
+
+% While we have tried to minimise the overlap of the techniques and aspects our challenges explore (linearity, scope extrusion, coinduction), most research on concurrency theory combines some or all of these.
+% A second round of formalisations might entail combining the features of the challenges.
+% We consider the following three extensions to be obvious places to start:
+% \emph{(1)} the addition of choice to the type level and the process level,
+% \emph{(2)} the addition of recursion and (for typed systems) recursive types,
+% and \emph{(3)} the addition of channel delegation to those calculi that do not already support it.%
 
 \paragraph{Addressing new calculi and new features.}
-\todo{Maybe cite some references in this section?}
+\todo{Maybe cite some references in this section? Mention the higher-order pi calculus?}
 Extending in this direction means proposing new challenges covering different features of process calculi, or different proof mechanisation features.

### 2024-02-03 short-paper/short-paper.tex
added references for infinitary proof systems/logics

@@
 %
-\luca{Coinduction may also play a relevant role in supporting recursive
-processes and session types. Indeed, it is often the case that session type
-systems (and logics) with recursive session types (and propositions) are
-naturally expressed in \emph{infinitary form} by interpreting their typing (and
-proof) rules coinductively. I CAN ADD ONE OR TWO REFERENCES IF YOU FEEL IT'S APPROPRIATE}
+Coinduction may also play a relevant role in supporting recursive processes and
+session types. Indeed, it is often the case that session type systems (and
+logics) with recursive session types (and propositions) are naturally expressed
+in \emph{infinitary form} by interpreting their typing (and proof) rules
+coinductively~\cite{BaeldeDoumaneSaurin16,DerakhshanPfenning22,HornePadovani23}.
 

### 2024-02-06 short-paper/short-paper.tex
Improve coherence in the introduction.

@@
 \section{Introduction}
-The POPLMark challenge~\cite{POPLMark} accelerated progress in the development of proof assistants, libraries, and best practices by establishing a common ground for systematically evaluating mechanisation techniques.
+The POPLMark challenge~\cite{POPLMark} accelerated the development of proof assistants, libraries, and best practices for programming language mechanisation by establishing a common ground for systematically evaluating mechanisation techniques.
 In doing so, it spearheaded a shift in the programming language research community towards publications that include mechanised proofs.
-POPLMark Reloaded~\cite{POPLMarkReloaded} proposed a similar program for proofs using logical relations.
-These challenges were deliberately limited in scope, while their authors called for future investigations into such areas as coinduction and linear environments.
+POPLMark Reloaded~\cite{POPLMarkReloaded} proposed a similar program for proofs involving logical relations.
+These challenges, and therefore the developments they spurred, were deliberately limited in scope, but their authors noted they need for future benchmarks to tackle such areas as coinduction and linear environments.
 
 In this spirit, we propose a new collection of benchmark problems specifically designed to address typical issues that arise when mechanising formal models of concurrent and distributed programming languages in general, and process calculi in particular.
-While many efforts to mechanise results about such formalisms have been completed (as we will discuss further below), our experience is that choosing appropriate proof mechanisation techniques and tools remains a significant challenge with a limited understanding of their respective trade-offs.
-This often leads researchers toward a trial-and-error approach, resulting in sub-optimal and ad-hoc solutions.
-In turn, this increases the overall effort required for proof mechanisation, and hampers the reuse of techniques and results.
+While results about such formalisms have already been mechanised (as we will discuss further below), our experience is that choosing appropriate mechanisation techniques and tools remains a significant challenge and that the trade-offs are not well understood.
+This often leads researchers toward a trial-and-error approach, resulting in sub-optimal solutions, wasted mechanisation effort, and techniques and results that are hard to reuse.
 
-%\alb{Is there a sharp difference among those papers?}
-%Several papers in this realm already include mechanised proofs (\eg~\cite{Castro-Perez2021, Castro2020, lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), or discuss proof mechanisation techniques (\eg~\cite{Bengtson2016, Gay2001, Thiemann2019, DBLP:conf/forte/ZalakainD21}).
-
-It is therefore our view that the research community in concurrency will benefit from a standardised set of challenge problems.
-To this end, our benchmark problems (discussed in detail in \cref{app:challenges}) are designed to focus on three key aspects which typically cause difficulties when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning} for infinite behaviours.
+Our benchmark problems (explained in detail in \cref{app:challenges}) are designed to focus on three key aspects which typically cause difficulties when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
 The mechanisation of novel research results often requires addressing more than just one of these aspects at once; we see the combination of techniques as a next step beyond the scope of this challenge, as discussed further in \Cref{sec:going-beyond}.
-These are certainly not the only factors that may cause issues, but they are fundamental to concurrency theory, and emerge in most mechanisations, as we will see presently while discussing the issues presented by the three key aspects.
+Our benchmark challenges are based on process calculi, as these provide a simple but realistic exhibit of the issues typically encountered when mechanising the three key aspects.
+These three aspects are certainly not the only causes of difficulties, but they are fundamental to concurrency theory, and emerge in most mechanisations, as we will see presently while discussing the issues presented by them.
 

### 2024-02-07 short-paper/short-paper.tex
start fleshing out the paragraph on coinduction

@@
 
-\paragraph{Coinduction.} \todo{}
+\paragraph{Coinduction.} \todo{add citations and explain more what the challenge does.}
+
+Coinduction is fundamental to establish properties of processes that
+may never terminate, a typical situation. We base our challenge around
+bisimulation a coinductively established property for equating the
+observed behaviour of potentially infinite processes.
+
+Coinduction is widely supported in proof assistants, Isabelle, Coq,
+and Abella seamlessly support coinductive definitions and reasoning.
+Agda and Beluga support it through the convenient abstraction
+of co-patterns. And newer entrants to the field like the Lean proof
+assistant, while not currently available, are working on adding a
+notion of coinduction to their system.
+
+As with most features, proof assistants support subtly different
+formalisms that enable different techniques and motivate this
+challenge. Some systems provide more than one abstraction to use
+coinduction. A good example of this are the PACO and interaction tree
+libraries for Coq.\todo{Francisco: add citations}.
+
+Common uses for coinduction are to define strong and weak versions of
+bisimulation, barbed notions of bisimulation, prove trace equivalence,
+etc.
+
+In this challenge we study the notion of strong barbed bisimulation
+over a fragment of the $\pi$-calculus.
 

### 2024-02-08 short-paper/short-paper.tex
Add references and clean up some wordings in coinduction paragraph.

--- 2024-02-06 short-paper/short-paper.tex
@@
 
-\paragraph{Coinduction.} \todo{}
+\paragraph{Coinduction.} \todo{explain more what the challenge does.}
+Coinduction is fundamental to establish properties of processes that
+may never terminate, which is a typical situation. We base our challenge around
+bisimulation: a coinductively established property for equating the
+observed behaviour of potentially infinite processes.
+
+Coinduction is widely supported in proof assistants.
+Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
+Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
+Newer entrants to the field like the Lean proof assistant are working on adding a notion of coinduction to their system~\cite{Avigad2019,Keizer2023}.
+
+As with most features, proof assistants support subtly different formalisms that enable different techniques and motivate this challenge.
+Some systems provide more than one abstraction to use coinduction.
+A good example of this are the parameterized coinduction~\cite{Hur2013} and interaction tree~\cite{Xia2019} libraries for Coq.
+
+Common uses for coinduction are to define strong and weak versions of
+bisimulation, barbed notions of bisimulation, prove trace equivalence,
+etc.
+
+In this challenge we study the notion of strong barbed bisimulation
+over a fragment of the $\pi$-calculus.
 

### 2024-02-09 short-paper/short-paper.tex
Rearrange related work sections.

@@
 %     - They are not covered by POPLMark
-%     - There are fewer techniques than for binders
+%     - Techniques are less settled than for binders
 %     - Others also mention them as problems: cite and explain references briefly for each challenge topic
@@
 POPLMark Reloaded~\cite{POPLMarkReloaded} proposed a similar program for proofs involving logical relations.
-These challenges, and therefore the developments they spurred, were deliberately limited in scope, but their authors noted they need for future benchmarks to tackle such areas as coinduction and linear environments.
+These challenges, and therefore the developments they spurred, were deliberately limited in scope, but their authors noted the need for future benchmarks to tackle such areas as coinduction and linear environments.
 
@@
 
-\todo{These paragraphs are still quite messy and unfocused.}
+\todo{These paragraphs are still quite messy and unfocused. They should focus on the motivation, i.e.\ mention what others have perceived as problems.}
 \paragraph{Linearity.}
@@
 The crux of the matter is \emph{context splitting}, as showcased by the typing rule for parallel composition.
-While a linear context is perhaps best seen as a multi-set, most proof assistants have better support for list processing.
-The latter representation is intuitive, but may require establishing a large number of technical lemmata that are orthogonal to the mathematics of the problem under study, say proving type preservation for session types.
-In this space, several designs are possible: one can label occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}) or \alb{something --- ask Francisco} as in \cite{Castro2020}.
-Some authors instead impose a multiset structure over lists (\eg~\cite{ChaudhuriLR19,Danielsson12}); still, the effort required to develop the infrastructure is significant.
 
-A different approach is to bypass the problem of context splitting by adopting familiar ideas from algorithmic linear type checking: this is known as \emph{typing with leftovers}, as exemplified in~\cite{DBLP:conf/forte/ZalakainD21}.
-Whatever the choice, list-based encodings can be refined to be intrinsically-typed, if the proof assistant does support dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
-
-For something completely different, one could adopt a \emph{sub-structural} meta-logical framework: here the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
-In fact, the linear function space will handle all context operations implicitly, including splitting and substitution.
-One such framework is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session types in~\cite{Bock2016}).
-Unfortunately, \emph{Celf} does not yet fully support the verification of meta-theoretic properties.
-A compromise is the so called \emph{two-level} approach, where one encodes a sub-structural specification logic in a mainstream proof assistant and then uses that logic to state and prove linear properties (for a recent example, see~\cite{Felty:MSCS21}).
+\todo{Insert related work focusing on noted issues with mechanising linearity.}
 
@@
 
-The case study \cite{AmbalLS21} compares four approaches to encoding binders in Coq as a first step towards developing tools for working with higher-order process calculi.
-They found that working directly with de Bruijn indices was easiest and shortest since the existing approaches developed for $\lambda$-calculus binders do not work well with scope extrusion.
-Specifically, the locally nameless representation~\cite{Chargueraud2012} cannot avoid direct manipulation of de Bruijn indices when defining the semantics of scope extrusion; cofinite quantification provides no benefits when working under binders; nominal approaches~\cite{Pitts2003} require explicitly giving and validating sets of free names during scope extrusion; and higher-order abstract syntax (HOAS)~\cite{Pfenning1988} requires additional axioms to work with contexts.
+\todo{Insert related work focusing on noted issues when mechanising scope extrusion.}
 
-Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, \ie, induction principles on functions to work with HOAS encodings.
-Some proof assistants such as Abella~\cite{Baelde2014} support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging~\cite{Momigliano2012}.
-
-% Start of related work
 \cite{Maksimovic2015} does not have name restriction, thus dodging the challenge of scope extrusion.
@@
 
-Coinduction is widely supported in proof assistants.
-Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
-Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
-Newer entrants to the field like the Lean proof assistant are working on adding a notion of coinduction to their system~\cite{Avigad2019,Keizer2023}.
-
-As with most features, proof assistants support subtly different formalisms that enable different techniques and motivate this challenge.
-Some systems provide more than one abstraction to use coinduction.
-A good example of this are the parameterized coinduction~\cite{Hur2013} and interaction tree~\cite{Xia2019} libraries for Coq.
-
 Common uses for coinduction are to define strong and weak versions of
@@
 over a fragment of the $\pi$-calculus.
+
+\todo{Insert related work focusing on noted issues when mechanising coinduction.}
 
@@
 \section{Design of the Benchmark}\label{sec:design-discussion}
-\todo{Add references to support our claims and decisions in this whole section.}
+\todo{Add references to support our claims and decisions in this whole section. Here is where we should mention existing approaches.}
 
@@
 \subsubsection{Linearity.}
+While a linear context is perhaps best seen as a multi-set, most proof assistants have better support for list processing.
+The latter representation is intuitive, but may require establishing a large number of technical lemmata that are orthogonal to the mathematics of the problem under study, say proving type preservation for session types.
+In this space, several designs are possible: one can label occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}) or \alb{something --- ask Francisco} as in \cite{Castro2020}.
+Some authors instead impose a multiset structure over lists (\eg~\cite{ChaudhuriLR19,Danielsson12}); still, the effort required to develop the infrastructure is significant.
+
+A different approach is to bypass the problem of context splitting by adopting familiar ideas from algorithmic linear type checking: this is known as \emph{typing with leftovers}, as exemplified in~\cite{DBLP:conf/forte/ZalakainD21}.
+Whatever the choice, list-based encodings can be refined to be intrinsically-typed, if the proof assistant does support dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
+
+For something completely different, one could adopt a \emph{sub-structural} meta-logical framework: here the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
+In fact, the linear function space will handle all context operations implicitly, including splitting and substitution.
+One such framework is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session types in~\cite{Bock2016}).
+Unfortunately, \emph{Celf} does not yet fully support the verification of meta-theoretic properties.
+A compromise is the so called \emph{two-level} approach, where one encodes a sub-structural specification logic in a mainstream proof assistant and then uses that logic to state and prove linear properties (for a recent example, see~\cite{Felty:MSCS21}).
+
+\todo{Insert discussion/related work focusing on existing techniques and how our benchmark design makes it possible to compare them.}
+
 For our challenge on linear reasoning we have chosen a type preservation theorem
@@
 since it concerns the properties of binders in restrictions and input.
-It is not obvious what the limitations and advantages are in
-leveraging the techniques used to solve the original challenges to
-reason about scope extrusion.  For example, techniques like
-higher-order abstract syntax or well-scoped de\ Bruijn indices lack
-the concept of free variables and thus encoding this challenge with
-them requires attention.
+
+The case study \cite{AmbalLS21} compares four approaches to encoding binders in Coq as a first step towards developing tools for working with higher-order process calculi.
+They found that working directly with de Bruijn indices was easiest and shortest since the existing approaches developed for $\lambda$-calculus binders do not work well with scope extrusion.
+Specifically, the locally nameless representation~\cite{Chargueraud2012} cannot avoid direct manipulation of de Bruijn indices when defining the semantics of scope extrusion; cofinite quantification provides no benefits when working under binders; nominal approaches~\cite{Pitts2003} require explicitly giving and validating sets of free names during scope extrusion; and higher-order abstract syntax (HOAS)~\cite{Pfenning1988} requires additional axioms to work with contexts.
+
+\todo{Insert discussion/related work focusing on existing techniques and how our benchmark design makes it possible to compare them.}
+
+Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, \ie, induction principles on functions to work with HOAS encodings.
+Some proof assistants such as Abella~\cite{Baelde2014} support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging~\cite{Momigliano2012}.
 
 \subsubsection{Coinduction.}
+Coinduction is widely supported in proof assistants.
+Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
+Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
+Newer entrants to the field like the Lean proof assistant are working on adding a notion of coinduction to their system~\cite{Avigad2019,Keizer2023}.
+
+As with most features, proof assistants support subtly different formalisms that enable different techniques and motivate this challenge.
+Some systems provide more than one abstraction to use coinduction.
+A good example of this are the parameterized coinduction~\cite{Hur2013} and interaction tree~\cite{Xia2019} libraries for Coq.
+
+\todo{Insert discussion/related work focusing on existing techniques and how our benchmark design makes it possible to compare them.}
+
 Our coinduction challenge concerns strong barbed bisimulation and

### 2024-02-12 short-paper/short-paper.tex
Improve some sentences and add some references about extensions.

@@
 Similarly to POPLMark and its Reloaded counterpart, our challenges do not encompass all issues in the field.
-Our benchmark is extendable in two aspects: one can further amalgamate and broaden the existing challenges, and tackle aspects not covered by the benchmark at all.
-All of these extensions will still need to handle at least one of the key aspects of our challenges.
+Our benchmark is extendable in two aspects: one can further amalgamate and broaden the existing challenges, or tackle aspects not covered by the benchmark at all.
+Any of these extensions will still need to handle at least one of the key aspects of our challenges.
+\todo{Frederik: Does splitting this section up into these two categories still make sense?
+We could save some space by not doing so, and it is not obvious to me what exactly should be on each side of the split.}
 
 \paragraph{Extending and combining the current challenges.}
-% thanks, copilot
-
-We have aimed to reduce the overlap of the techniques and aspects that
-our challenges (linearity, scope extrusion, coinduction), despite
-their common use in concurrency theory research. A subsequent round
-could merge these features. We suggest starting with three extensions:
-\emph{(1)} adding choice to both type and process levels, \emph{(2)}
-introducing recursion and recursive types for typed systems, and
+We have aimed to reduce the overlap of the aspects (linearity, scope extrusion, coinduction) that our challenges involve.
+Most concurrency theory research needs several of the aspects and often also other constructs that we have chosen not to cover for simplicity.
+We suggest three extensions that would be useful for many applications:
+\emph{(1)} adding choice constructs to both type and process levels,
+\emph{(2)} introducing recursion and recursive types for typed systems, and
 \emph{(3)} incorporating channel delegation in calculi lacking it.
-%\todo{This paragraph is very long for how relatively little it says. Maybe remove/cut down?}
-
-% While we have tried to minimise the overlap of the techniques and aspects our challenges explore (linearity, scope extrusion, coinduction), most research on concurrency theory combines some or all of these.
-% A second round of formalisations might entail combining the features of the challenges.
-% We consider the following three extensions to be obvious places to start:
-% \emph{(1)} the addition of choice to the type level and the process level,
-% \emph{(2)} the addition of recursion and (for typed systems) recursive types,
-% and \emph{(3)} the addition of channel delegation to those calculi that do not already support it.%
 
 \paragraph{Addressing new calculi and new features.}
-\todo{Maybe cite some references in this section? Mention the higher-order pi calculus?}
+\todo{Mention the higher-order pi calculus?}
 Extending in this direction means proposing new challenges covering different features of process calculi, or different proof mechanisation features.
-Some interesting aspects of message-passing calculi to be explored in further challenges could be multiparty session types and choreographies, as their meta-theory includes aspects -- \eg, well-formedness conditions on global types, partiality of end-point projection function, etc.
--- that we do not address here.
+Some interesting aspects of message-passing calculi to be explored in further challenges could be multiparty session types~\cite{10.1145/2827695} and choreographies~\cite{Carbone2013}, as their meta-theory includes aspects -- \eg well-formedness conditions on global types, partiality of end-point projection function, \etc -- that we do not address here.
 Also, one could easily design a challenge that goes beyond our coinduction challenge, by exploring different notions of bisimilarity (barbed, weak, \etc) and trace equivalence.
@@
 On the mechanisation side, the present challenge proposes mechanising the proofs of theorems.
-An interesting avenue to explore would be to take advantage of other proof assistant features.
-For example, proof assistants can often produce certified code (via code extraction, or compiling and running their definitions).
+An interesting avenue to explore would be to take advantage of other proof assistant features like extracting certified implementations.
 Additionally, while our current challenge is agnostic about proof automation, a more specific challenge could be designed with the objective of automating aspects of the proofs.

### 2024-02-12 short-paper/short-paper.tex
Shorten and improve introduction.

@@
 POPLMark Reloaded~\cite{POPLMarkReloaded} proposed a similar program for proofs involving logical relations.
-These challenges, and therefore the developments they spurred, were deliberately limited in scope, but their authors noted the need for future benchmarks to tackle such areas as coinduction and linear environments.
+These challenges, and the developments they spurred, were limited in scope, but their authors noted the need to tackle topics such as coinduction and linearity in the future.
 
@@
 
-Our benchmark problems (explained in detail in \cref{app:challenges}) are designed to focus on three key aspects which typically cause difficulties when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
-The mechanisation of novel research results often requires addressing more than just one of these aspects at once; we see the combination of techniques as a next step beyond the scope of this challenge, as discussed further in \Cref{sec:going-beyond}.
-Our benchmark challenges are based on process calculi, as these provide a simple but realistic exhibit of the issues typically encountered when mechanising the three key aspects.
-These three aspects are certainly not the only causes of difficulties, but they are fundamental to concurrency theory, and emerge in most mechanisations, as we will see presently while discussing the issues presented by them.
+Our benchmark problems (explained in detail in \cref{app:challenges}) focus on three key aspects which often cause difficulties when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
+Mechanisations must often address several of these aspects, but we see the combination of techniques as a next step beyond the scope of this challenge, as discussed further in \Cref{sec:going-beyond}.
+Our benchmark challenges are based on process calculi, as these provide a simple but realistic exhibit of typical issues encountered in mechanisation work.
+The three aspects are certainly not the only causes of difficulties, but as we will discuss presently, they emerge in most mechanisations of concurrency theory.
 

### 2024-02-14 short-paper/short-paper.tex
some coinduction

@@
 Coinduction is fundamental to establish properties of processes that
-may never terminate, which is a typical situation. We base our challenge around
-bisimulation: a coinductively established property for equating the
-observed behaviour of potentially infinite processes.
-
+may never terminate, which is a typical situation.
 Common uses for coinduction are to define strong and weak versions of
@@
 
-In this challenge we study the notion of strong barbed bisimulation
-over a fragment of the $\pi$-calculus.
+We base our challenge around
+bisimulation: a fundamental coinductively established property for equating the
+observed behaviour of potentially infinite processes. 
+In this challenge, we study the notion of strong barbed bisimulation
+(and bisimilarity)
+over a fragment of the $\pi$-calculus and,
+in particular, we focus on obtaining,
+from strong barbed bisimilarity,
+a congruence with respect to
+parallel composition and substitution.
+
+The introduction of bisimulation
+is one of the first steps, when studying the
+semantics of process calculi, both
+in classical textbooks (e.g., \cite{})
+and in almost all existent 
+mechanisations \cite{}
+(where we see a variety
+of different choices).
+
+%Common uses for coinduction are to define strong and weak versions of
+%bisimulation, barbed notions of bisimulation, prove trace equivalence,
+%etc.
+
+%In this challenge we study the notion of strong barbed bisimulation
+%over a fragment of the $\pi$-calculus.
 
@@
 
+\lore{Need to integrate the following with the above.}
+Different process-calculi
+mechanisations
+implement different techinques;
+most of the time these 
+are ad-hoc efforts to achieve a specific result.
+\lore{cite and substantiate.}
+In parallel, although most proof assistants
+offer support for coinduction in their standard libraries,
+research has mainly focused on the foundations
+of the specific system, more than on
+applicability.
+\lore{cite and substantiate.}
+
 \subsubsection{Evaluation criteria.}

### 2024-02-15 short-paper/short-paper.tex
Add some discussion of why mechanisation is valuable.

@@
 In this spirit, we propose a new collection of benchmark problems specifically designed to address typical issues that arise when mechanising formal models of concurrent and distributed programming languages in general, and process calculi in particular.
-While results about such formalisms have already been mechanised (as we will discuss further below), our experience is that choosing appropriate mechanisation techniques and tools remains a significant challenge and that the trade-offs are not well understood.
+Concurrent calculi are notably subtle: for instance it took nine years before the incorrect proof of type safety in~\cite{Honda1998} was fixed in~\cite{Yoshida2007}.
+Similarly, many results on duality, recursion, subject reduction and expressiveness in session types have turned out to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the value of machine-checkable proofs.
+While results about concurrent formalisms have already been mechanised (as we will discuss further below), our experience is that choosing appropriate mechanisation techniques and tools remains a significant challenge and that the trade-offs are not well understood.
 This often leads researchers toward a trial-and-error approach, resulting in sub-optimal solutions, wasted mechanisation effort, and techniques and results that are hard to reuse.

### 2024-02-15 short-paper/short-paper.tex
Add some various related work.

@@
 Concurrent calculi are notably subtle: for instance it took nine years before the incorrect proof of type safety in~\cite{Honda1998} was fixed in~\cite{Yoshida2007}.
-Similarly, many results on duality, recursion, subject reduction and expressiveness in session types have turned out to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the value of machine-checkable proofs.
+Similarly, many key results in peer-reviewed papers on session types have turned out to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the value of machine-checkable proofs.
 While results about concurrent formalisms have already been mechanised (as we will discuss further below), our experience is that choosing appropriate mechanisation techniques and tools remains a significant challenge and that the trade-offs are not well understood.
 This often leads researchers toward a trial-and-error approach, resulting in sub-optimal solutions, wasted mechanisation effort, and techniques and results that are hard to reuse.
+\cite{Cruz-Filipe2021b} notes that the high-level parts of mechanised proofs closely resemble the informal versions, and that the main challenge lies in getting the foundations right.
 
@@
 The crux of the matter is \emph{context splitting}, as showcased by the typing rule for parallel composition.
+
+\cite{Brady2017} notes that using algebraic effects is unwieldy when there are many communication channels and suggests an experimental approach using dependent uniqueness types.
 
@@
 of different choices).
+
+\cite{Castro-Perez2021} note that the most challenging part of their mechanisation was the convoluted definition of infinite behaviours.
 
@@
 
+\cite{DBLP:conf/forte/ZalakainD21} notes that the combination of linearity and infinite behaviour is difficult since it requires uniquely determining the resources consumed by a process.
+
 \paragraph{Addressing new calculi and new features.}
-\todo{Mention the higher-order pi calculus?}
+\todo{Mention the higher-order pi calculus? If so, also mention~\cite{Hirsch2022}}
 Extending in this direction means proposing new challenges covering different features of process calculi, or different proof mechanisation features.

### 2024-02-16 short-paper/short-paper.tex
Add some motivating related work in note form.

@@
 In this spirit, we propose a new collection of benchmark problems specifically designed to address typical issues that arise when mechanising formal models of concurrent and distributed programming languages in general, and process calculi in particular.
-Concurrent calculi are notably subtle: for instance it took nine years before the incorrect proof of type safety in~\cite{Honda1998} was fixed in~\cite{Yoshida2007}.
+Concurrent calculi are notably subtle: for instance it took nine years before the incorrect proof of type safety in the seminal paper~\cite{Honda1998} was fixed in~\cite{Yoshida2007}.
 Similarly, many key results in peer-reviewed papers on session types have turned out to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the value of machine-checkable proofs.
@@
 \cite{Brady2017} notes that using algebraic effects is unwieldy when there are many communication channels and suggests an experimental approach using dependent uniqueness types.
+\cite{Zalakain2019} notes that systems lacking linear types must simulate linearity, and that using a linearity predicate complicates congruences.
+They had to limit themselves to an asymmetric structural precongruence to be able to prove subject reduction.
 
@@
 \cite{Maksimovic2015} does not have name restriction, thus dodging the challenge of scope extrusion.
+\cite{Castro-Perez2021,Castro2020} note that popular presentations of session type systems are impossible to directly encode using intrinsically \(\alpha\)-convertible terms.
+\cite{Castro2020} suggests using several binder constructs to solve this, but notes that this requires proving many variants of the same theorems.
 
@@
 \cite{DBLP:conf/forte/ZalakainD21} notes that the combination of linearity and infinite behaviour is difficult since it requires uniquely determining the resources consumed by a process.
+\cite{Zalakain2019} notes that mixing replication and linearity seems to be very difficult without using explicit type contexts.
 

### 2024-02-17 short-paper/short-paper.tex
some rephrasing of initial paragraphs in the intro
some comments on the general organization

cleanup of duplication doi/url in bib

@@
 \section{Introduction}
-The POPLMark challenge~\cite{POPLMark} accelerated the development of proof assistants, libraries, and best practices for programming language mechanisation by establishing a common ground for systematically evaluating mechanisation techniques.
-In doing so, it spearheaded a shift in the programming language research community towards publications that include mechanised proofs.
-POPLMark Reloaded~\cite{POPLMarkReloaded} proposed a similar program for proofs involving logical relations.
-These challenges, and the developments they spurred, were limited in scope, but their authors noted the need to tackle topics such as coinduction and linearity in the future.
 
-In this spirit, we propose a new collection of benchmark problems specifically designed to address typical issues that arise when mechanising formal models of concurrent and distributed programming languages in general, and process calculi in particular.
-Concurrent calculi are notably subtle: for instance it took nine years before the incorrect proof of type safety in the seminal paper~\cite{Honda1998} was fixed in~\cite{Yoshida2007}.
-Similarly, many key results in peer-reviewed papers on session types have turned out to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the value of machine-checkable proofs.
-While results about concurrent formalisms have already been mechanised (as we will discuss further below), our experience is that choosing appropriate mechanisation techniques and tools remains a significant challenge and that the trade-offs are not well understood.
-This often leads researchers toward a trial-and-error approach, resulting in sub-optimal solutions, wasted mechanisation effort, and techniques and results that are hard to reuse.
-\cite{Cruz-Filipe2021b} notes that the high-level parts of mechanised proofs closely resemble the informal versions, and that the main challenge lies in getting the foundations right.
+The POPLMark challenge~\cite{POPLMark}
+played a pivotal role in advancing the field of proof assistants,
+libraries, and best practices for programming language
+mechanization. By providing a shared framework for systematically
+evaluating mechanization techniques, it catalyzed a significant
+shift within the programming language research community. As a
+result, scholarly publications increasingly incorporate
+mechanized proofs.
+%
+% The POPLMark challenge~\cite{POPLMark} accelerated the development of proof assistants, libraries, and best practices for programming language mechanisation by establishing a common ground for systematically evaluating mechanisation techniques.
+% In doing so, it spearheaded a shift in the programming language research community towards publications that include mechanised proofs.
+% POPLMark Reloaded~\cite{POPLMarkReloaded} proposed a similar program for proofs involving logical relations.
+% These challenges, and the developments they spurred, were limited in scope, but their authors noted the need to tackle topics such as coinduction and linearity in the future.
 
-Our benchmark problems (explained in detail in \cref{app:challenges}) focus on three key aspects which often cause difficulties when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
-Mechanisations must often address several of these aspects, but we see the combination of techniques as a next step beyond the scope of this challenge, as discussed further in \Cref{sec:going-beyond}.
-Our benchmark challenges are based on process calculi, as these provide a simple but realistic exhibit of typical issues encountered in mechanisation work.
-The three aspects are certainly not the only causes of difficulties, but as we will discuss presently, they emerge in most mechanisations of concurrency theory.
+POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a program akin to
+the original POPLMark challenge, specifically targeting proofs related
+to logical relations. While these initiatives had a narrow focus,
+their authors recognized the importance of addressing broader topics
+in the future, including coinduction and linearity.
+
+In this spirit, we propose a new collection of benchmark problems
+specifically designed to address typical issues that arise when
+mechanising formal models of concurrent and distributed programming
+languages in general, and process calculi in particular.  Concurrent
+calculi are notably subtle: for instance it took nine years before the
+incorrect proof of type safety in the seminal paper by Honda et
+al.~\cite{Honda1998} was fixed in~\cite{Yoshida2007}.  Similarly, many
+key results in peer-reviewed papers on session types have turned out
+to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the
+value of machine-checkable proofs.
+
+While results about concurrent
+formalisms have already been mechanised (as we will discuss further
+below), our experience is that choosing appropriate mechanisation
+techniques and tools remains a significant challenge and that the
+trade-offs are not well understood.  This often leads researchers
+toward a trial-and-error approach, resulting in sub-optimal solutions,
+wasted mechanisation effort, and techniques and results that are hard
+to reuse.  For example, the authors in~\cite{Cruz-Filipe2021b} note that the high-level parts of
+mechanised proofs closely resemble the informal versions, and that the
+main challenge lies in getting the foundations right.
+
+Our benchmark problems (explained in detail in \cref{app:challenges}) center around three critical aspects that frequently pose challenges in mechanizing concurrency theory:
+% focus on three key aspects which often cause difficulties when mechanising concurrency theory:
+\emph{linearity}, \emph{scope extrusion}, and \emph{coinductive
+  reasoning}.  Mechanisations must often address several of these
+aspects at the same time; however, we see the combination of
+techniques as a next step beyond the scope of this challenge, as
+discussed further in \Cref{sec:going-beyond}.  Our benchmark
+challenges are based on process calculi, as these provide a simple but
+realistic showcase of common issues encountered during mechanization endeavors
+% exhibit of typical issues encountered in mechanisation work.
+While the three highlighted aspects are not the sole sources of difficulties, they
+consistently manifest in most mechanizations of concurrency theory.
+% The three aspects are certainly not the only causes of difficulties,
+% but as we will discuss presently, they emerge in most mechanisations
+% of concurrency theory.
 
 \todo{These paragraphs are still quite messy and unfocused. They should focus on the motivation, i.e.\ mention what others have perceived as problems.}
+
+\alb{I really do not like the duplication of the discussion of the
+  three issues here and in Sec 2. I understand the desire to highlight
+  noted issues, but it does not work. The citations are unconvincing, furthermore}
 \paragraph{Linearity.}
@@
 \subsubsection{General design remarks.}
-Like the POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive:
-applications such as multiparty session types~\cite{10.1145/2827695,10.1145/3290343}, choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,DBLP:conf/ecoop/ScalasDHY17,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407} are not directly covered.\alb{Do we need all those citations about encodings?}
-Still, these (and other) applications need the basic techniques that our challenge problems emphasize: as per design goal \ref{item:goal-reusability}, our problems are drawn from the basic meta-theory of process calculi (without requiring the development of new theory), and focus on well-known theorems and results that involve interesting proof techniques that may be reused in further work.
-To achieve design goal \ref{item:goal-comperison-accessibility},
-we have formulated our challenge problems to explore the three aspects (linearity, scope extrusion, coinduction) independently, so that they may be solved individually and in any order;
-each problem should be reasonably small and easily understood with basic knowledge of textbook concurrency theory, process calculi, and type theory.  The process calculus used in each challenge
-focuses on the features that we want to emphasize, and omits all constructs
-(such as choices) that would complicate the mechanisation without bringing additional tangible
-benefits and insights. 
-The minimality and uniformity of the calculi also allows us to target design
-goal \ref{item:goal-tutorials}.
+Like the POPLMark and POPLMark Reloaded challenges, our challenge is
+not meant to be comprehensive: applications such as multiparty session
+types~\cite{10.1145/2827695,10.1145/3290343},
+choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation
+types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696},
+or encodings between different
+calculi~\cite{DBLP:journals/iandc/Gorla10,%,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,
+  DBLP:conf/ecoop/ScalasDHY17}
+  %,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407}
+are not directly covered.
+
+\alb{I've commented some citations about
+  encodings. Is it a good idea to start the section describing what we \textbf{do not do}?}
+
+Still, these (and other) applications need the basic
+techniques that our challenge problems emphasize: as per design goal
+\ref{item:goal-reusability}, our problems are drawn from the basic
+meta-theory of process calculi (without requiring the development of
+new theory), and focus on well-known theorems and results that involve
+interesting proof techniques that may be reused in further work.  To
+achieve design goal \ref{item:goal-comperison-accessibility}, we have
+formulated our challenge problems to explore the three aspects
+(linearity, scope extrusion, coinduction) independently, so that they
+may be solved individually and in any order; each problem should be
+reasonably small and easily understood with basic knowledge of
+textbook concurrency theory, process calculi, and type theory.  The
+process calculus used in each challenge focuses on the features that
+we want to emphasize, and omits all constructs (such as choices) that
+would complicate the mechanisation without bringing additional
+tangible benefits and insights.  The minimality and uniformity of the
+calculi also allows us to target design goal
+\ref{item:goal-tutorials}.
 
@@
 mechanisations
-implement different techinques;
+implement different techniques;
 most of the time these 

### 2024-02-17 short-paper/short-paper.tex
more coind

--- 2024-02-16 short-paper/short-paper.tex
@@
 Common uses for coinduction are to define strong and weak versions of
-bisimulation, barbed notions of bisimulation, prove trace equivalence,
-etc.
+bisimulation, barbed notions of bisimulation, and prove trace equivalence. %,
+%etc.
+On the other hand, coinduction in proof assistants
+is notoriously difficult: \lore{add here}
+\cite{Castro-Perez2021} note that the most challenging
+part of their mechanisation of multiparty session types
+was the convoluted definition of infinite behaviours.
 
-We base our challenge around
-bisimulation: a fundamental coinductively established property for equating the
+We base our challenge around bisimulation: a fundamental coinductively
+established property for equating the
 observed behaviour of potentially infinite processes. 
 In this challenge, we study the notion of strong barbed bisimulation
-(and bisimilarity)
+(and bisimilarity, as in \cite{picalcbook})
 over a fragment of the $\pi$-calculus and,
@@
 is one of the first steps, when studying the
-semantics of process calculi, both
-in classical textbooks (e.g., \cite{})
+(observable) behaviour of process calculi, both
+in classical textbooks (e.g., \cite{picalcbook})
 and in almost all existent 
-mechanisations \cite{}
-(where we see a variety
-of different choices).
-
-\cite{Castro-Perez2021} note that the most challenging part of their mechanisation was the convoluted definition of infinite behaviours.
+mechanisations, where we see a variety of
+different formalisation choices.
+As a significative example, we observe that,
+while many authors rely
+on the native support for coinduction offered by
+proof assistants \cite{},
+many others prefer a more set-theoretic
+approach \cite{}.
 

### 2024-02-18 short-paper/short-paper.tex
coinduction, almost there both in Intro and in Design

@@
 \paragraph{Coinduction.} \todo{explain more what the challenge does.}
+\lore{I have tried this:
+  the first paragraph motivates (``why is this difficult?''), while
+  the second talks about the challenge.}
 Coinduction is fundamental to establish properties of processes that
@@
 On the other hand, coinduction in proof assistants
-is notoriously difficult: \lore{add here}
-\cite{Castro-Perez2021} note that the most challenging
+is notoriously difficult and its representation
+is non-uniform across development.
+E.g., \cite{Castro-Perez2021} notes that the most challenging
 part of their mechanisation of multiparty session types
-was the convoluted definition of infinite behaviours.
+was the convoluted definition of infinite behaviours;
+\cite{Pohjola2022} points out that the main difficulty
+of the proofs concerning bisimulation lies
+in dealing with existential quantifiers; and
+\cite{Bengtson2016} wonders whether parametrised
+coinduction as developed in Coq \cite{Hur2013} would
+have helped their Isabelle mechanisation.
 
@@
 parallel composition and substitution.
+%
+%% The introduction of bisimulation
+%% is one of the first steps, when studying the
+%% (observable) behaviour of process calculi, both
+%% in classical textbooks (e.g., \cite{picalcbook})
+%% and in existing 
+%% mechanisations
+%% \cite{Thiemann2019,Honsell2001,Bengtson2016,Kahsai2008,Gay2020,
+%% Hirschkoff1997,Bengtson2009,Maksimovic2015}.
 
-The introduction of bisimulation
-is one of the first steps, when studying the
-(observable) behaviour of process calculi, both
-in classical textbooks (e.g., \cite{picalcbook})
-and in almost all existent 
-mechanisations, where we see a variety of
-different formalisation choices.
-As a significative example, we observe that,
-while many authors rely
-on the native support for coinduction offered by
-proof assistants \cite{},
-many others prefer a more set-theoretic
-approach \cite{}.
+%% , where we see a variety of
+%% different formalisation choices.
+%% As a significant example, we observe that,
+%% while many authors rely
+%% on the native support for coinduction offered by
+%% proof assistants
+%% \cite{Thiemann2019,Honsell2001,Bengtson2016,Kahsai2008,Gay2020},
+%% many others prefer a more ``set-theoretic''
+%% approach
+%% \cite{Hirschkoff1997,Bengtson2009,Maksimovic2015}.
 
@@
 Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
-Newer entrants to the field like the Lean proof assistant are working on adding a notion of coinduction to their system~\cite{Avigad2019,Keizer2023}.
+Newer entrants to the field like the Lean proof assistant
+are working on adding a notion of coinduction to their system~\cite{Avigad2019,Keizer2023}.
 
-As with most features, proof assistants support subtly different formalisms that enable different techniques and motivate this challenge.
+As with most features, proof assistants support
+subtly different formalisms that enable different techniques and motivate this challenge.
 Some systems provide more than one abstraction to use coinduction.
-A good example of this are the parameterized coinduction~\cite{Hur2013} and interaction tree~\cite{Xia2019} libraries for Coq.
+A good example of this are the parameterized coinduction~\cite{Hur2013}
+and interaction tree~\cite{Xia2019} libraries for Coq.
 
+\lore{Should we mention that Isabelle offers
+  the command \texttt{coinduction\_upto}? Is it related?}
 \todo{Insert discussion/related work focusing on existing techniques and how our benchmark design makes it possible to compare them.}
+\lore{I have added something below... Any advice on what more we need?}
 
@@
 congruence.\alb{we motivate why strong vs weak, but not why barbed}
+\lore{Is it not enough to quote the Sangiorgi/Walker book
+and go by authority? :) }
 Though weak barbed congruence is a more common behavioural equivalence
@@
 
-\lore{Need to integrate the following with the above.}
-Different process-calculi
-mechanisations
-implement different techniques;
-most of the time these 
-are ad-hoc efforts to achieve a specific result.
-\lore{cite and substantiate.}
-In parallel, although most proof assistants
-offer support for coinduction in their standard libraries,
-research has mainly focused on the foundations
-of the specific system, more than on
-applicability.
-\lore{cite and substantiate.}
+The introduction of strong barbed bisimulation
+is one of the first steps, when studying the
+(observable) behaviour of process calculi, both
+in classical textbooks (e.g., \cite{picalcbook})
+and in existing 
+mechanisations, where we see a variety of
+different formalisation choices,
+sometimes ad-hoc for the specific
+result or development.
+As a significant example, we observe that,
+while many authors rely
+on the native support for coinduction offered by
+proof assistants
+\cite{Honsell2001,Bengtson2016,Kahsai2008,Thiemann2019,Gay2020},
+many others prefer a more ``set-theoretic''
+approach
+\cite{Hirschkoff1997,Bengtson2009,Maksimovic2015,Pohjola2022}.
+
+%% \lore{Need to integrate the following with the above.}
+%% Different process-calculi
+%% mechanisations
+%% implement different techniques;
+%% most of the time these 
+%% are ad-hoc efforts to achieve a specific result.
+%% \lore{cite and substantiate.}
+%% In parallel, although most proof assistants
+%% offer support for coinduction in their standard libraries,
+%% research has mainly focused on the foundations
+%% of the specific system, more than on
+%% applicability.
+%% \lore{cite and substantiate.}
 

### 2024-02-18 short-paper/short-paper.tex
more comments

--- 2024-02-17 short-paper/short-paper.tex
@@
 Common uses for coinduction are to define strong and weak versions of
-bisimulation, barbed notions of bisimulation, and prove trace equivalence. %,
-%etc.
-On the other hand, coinduction in proof assistants
-is notoriously difficult: \lore{add here}
-\cite{Castro-Perez2021} note that the most challenging
-part of their mechanisation of multiparty session types
-was the convoluted definition of infinite behaviours.
+bisimulation, barbed notions of bisimulation, prove trace equivalence,
+etc.
 
-We base our challenge around bisimulation: a fundamental coinductively
-established property for equating the
+We base our challenge around
+bisimulation: a fundamental coinductively established property for equating the
 observed behaviour of potentially infinite processes. 
 In this challenge, we study the notion of strong barbed bisimulation
-(and bisimilarity, as in \cite{picalcbook})
+(and bisimilarity)
 over a fragment of the $\pi$-calculus and,
@@
 is one of the first steps, when studying the
-(observable) behaviour of process calculi, both
-in classical textbooks (e.g., \cite{picalcbook})
+semantics of process calculi, both
+in classical textbooks (e.g., \cite{})
 and in almost all existent 
-mechanisations, where we see a variety of
-different formalisation choices.
-As a significative example, we observe that,
-while many authors rely
-on the native support for coinduction offered by
-proof assistants \cite{},
-many others prefer a more set-theoretic
-approach \cite{}.
+mechanisations \cite{}
+(where we see a variety
+of different choices).
+
+\cite{Castro-Perez2021} note that the most challenging part of their mechanisation was the convoluted definition of infinite behaviours.
 
@@
 
-\subsection{Goals and how to contribute}
+\subsection{Goals}% and how to contribute}
 Similarly to the authors of the POPLMark challenges, we seek to
@@
 \subsubsection{Linearity.}
+
+\alb{I would put the challenge before the literature}
+For our challenge on linear reasoning we have chosen a type safety theorem
+for a system using session types, combining as usual type preservation with absence of errors.
+% , whose use are common in concurrency theory.
+This allows us to introduce linear reasoning with few definitions.
+With our objective focused on linearity, we forego channel delegation
+and opt for a reduction approach rather than employing a transition
+system semantics.  Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
+define a syntax where restriction binds two names, highlighting
+their duality within the type system.  We have chosen a simple notion
+of well-formedness, and do not  consider
+deadlocks. While more sophisticated notions of well-formedness are
+interesting, the proofs of these properties would distract from the
+core linear aspects of the challenge.
+
 While a linear context is perhaps best seen as a multi-set, most proof assistants have better support for list processing.
@@
 \todo{Insert discussion/related work focusing on existing techniques and how our benchmark design makes it possible to compare them.}
-
-For our challenge on linear reasoning we have chosen a type preservation theorem
-for a system using session types.
-% , whose use are common in concurrency theory.
-This allows us to introduce linear reasoning with few definitions.
-With our objective focused on linearity, we forego channel delegation
-and opt for a reduction approach rather than employing a transition
-system semantics.  Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
-define a syntax where restriction binds two names,n highlighting
-their duality within the type system.  We have chosen a simple notion
-of well-formedness, and do not  consider
-deadlocks. While more sophisticated notions of well-formedness are
-interesting, the proofs of these properties would distract from the
-core linear aspects of the challenge.
 

### 2024-02-18 short-paper/short-paper.tex
many more notes
some small rewrites to improve readability

@@
 
-In this spirit, we propose a new collection of benchmark problems
-specifically designed to address typical issues that arise when
-mechanising formal models of concurrent and distributed programming
-languages in general, and process calculi in particular.  Concurrent
-calculi are notably subtle: for instance it took nine years before the
-incorrect proof of type safety in the seminal paper by Honda et
-al.~\cite{Honda1998} was fixed in~\cite{Yoshida2007}.  Similarly, many
+In this spirit, we introduce a novel collection of benchmark problems
+specifically crafted to tackle common challenges encountered during
+the mechanization of formal models for concurrent and distributed
+programming languages. Our focus extends particularly to process
+calculi.
+%
+% In this spirit, we propose a new collection of benchmark problems
+% specifically designed to address typical issues that arise when
+% mechanising formal models of concurrent and distributed programming
+% languages in general, and process calculi in particular. 
+Concurrent calculi are notably subtle: For instance, it took nine
+years before the incorrect proof of type safety in the seminal paper
+by Honda et al.~.~\cite{Honda1998} was rectified in Yoshida \&
+Vasconcelos’s subsequent work\cite{Yoshida2007}.
+% for instance it took nine years before the incorrect proof of type
+% safety in the seminal paper by Honda et al was fixed
+% in~\cite{Yoshida2007}.
+Similarly, many
 key results in peer-reviewed papers on session types have turned out
@@
   three issues here and in Sec 2. I understand the desire to highlight
-  noted issues, but it does not work. The citations are unconvincing, furthermore}
+  noted issues, but it does not work. Some citations are unconvincing. Mostly, we are anticipating the description of the challenges. Let's unify this}
 \paragraph{Linearity.}
@@
 \paragraph{Scope extrusion.}
-The encoding of binding constructs has been extensively studied for type systems based on $\lambda$-calculi, but process calculi present unique challenges.
-$\pi$-calculi typically include several different binding constructs: inputs bind a received name or value, recursive processes bind recursion variables, and restrictions bind names.
-The first two act similarly to the binders known from $\lambda$-calculi, but restrictions are more challenging due to scope extrusion.
+The encoding of binding constructs has been extensively studied
+w.r.t.\ the $\lambda$-calculi, but process calculi present additional
+challenges.  $\pi$-calculi typically include several different binding
+constructs: inputs bind a received name or value, recursive processes
+bind recursion variables, and restrictions bind names.  The first two
+act similarly to the binders known from $\lambda$-calculi, but
+restrictions may be more challenging due to scope extrusion.
 
@@
 
+\alb{This subsection does not make a lot of sense here -- introductions do not have subsections. Let's move it in ``Design'' }
 \subsection{Goals}% and how to contribute}
@@
 answer several questions:
-%\alb{We oscillate between concurrency in general and process calculi, here and before}
+\alb{Should we stress we are interested in the \textbf{meta-theory} of process calculi, not}
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
@@
 Still, these (and other) applications need the basic
-techniques that our challenge problems emphasize: as per design goal
-\ref{item:goal-reusability}, our problems are drawn from the basic
-meta-theory of process calculi (without requiring the development of
-new theory), and focus on well-known theorems and results that involve
-interesting proof techniques that may be reused in further work.  To
+techniques that our challenge problems emphasize:
+Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
+These problems center around well-established theorems and results, showcasing intriguing proof techniques that can be leveraged in future endeavors.
+% as per design goal
+% \ref{item:goal-reusability}, our problems are drawn from the basic
+% meta-theory of process calculi (without requiring the development of
+% new theory), and focus on well-known theorems and results that involve
+% interesting proof techniques that may be reused in further work. 
+To
 achieve design goal \ref{item:goal-comperison-accessibility}, we have
@@
 (linearity, scope extrusion, coinduction) independently, so that they
-may be solved individually and in any order; each problem should be
-reasonably small and easily understood with basic knowledge of
+may be solved individually and in any order; each problem is %should be
+reasonably small and easily understandable with basic knowledge of
 textbook concurrency theory, process calculi, and type theory.  The
@@
 
+\alb{This needs some work: we do not describe the challenge at all and the citations are not the one I had in mind}
 \subsubsection{Scope extrusion.}
@@
 \subsubsection{Coinduction.}
-Coinduction is widely supported in proof assistants.
+\alb{I kinda object to ``seamlessly''. In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky. I guess Isabelle is better}
+
+Our coinduction challenge concerns strong barbed bisimulation and
+congruence.% \alb{we motivate why strong vs weak, but not why barbed}
+% \lore{Is it not enough to quote the Sangiorgi/Walker book
+% and go by authority? :)}
+Though weak barbed congruence is a more common behavioural equivalence
+w.r.t.\ the \(\pi\)-calculus, here we prefer strong equivalences to avoid the
+need to abstract over the number of internal transitions, thus
+simplifying the theory.  We exclude delegation from the coinduction
+challenge since it is orthogonal to our primary aim of exploring
+coinductive proof techniques.
+
+The introduction of strong barbed bisimulation is one of the first
+steps when studying the (observable) behaviour of process calculi,
+both in classical textbooks (e.g., \cite{picalcbook}) and in existing
+mechanisations, where we see a variety of different formalisation
+choices, sometimes ad-hoc for the specific result or development.
+
+In fact, coinduction is widely supported in proof assistants.
 Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
@@
 Newer entrants to the field like the Lean proof assistant
-are working on adding a notion of coinduction to their system~\cite{Avigad2019,Keizer2023}.
-
+are working on adding a notion of coinduction to their system~\cite{Avigad2019}%,Keizer2023}.
+%
 As with most features, proof assistants support
@@
 
-Our coinduction challenge concerns strong barbed bisimulation and
-congruence.\alb{we motivate why strong vs weak, but not why barbed}
-\lore{Is it not enough to quote the Sangiorgi/Walker book
-and go by authority? :) }
-Though weak barbed congruence is a more common behavioural equivalence
-w.r.t.\ the \(\pi\)-calculus, here we prefer strong equivalences to avoid the
-need to abstract over the number of internal transitions, thus
-simplifying the theory.  We exclude delegation from the coinduction
-challenge since it is orthogonal to our primary aim of exploring
-coinductive proof techniques.
-
-The introduction of strong barbed bisimulation
-is one of the first steps, when studying the
-(observable) behaviour of process calculi, both
-in classical textbooks (e.g., \cite{picalcbook})
-and in existing 
-mechanisations, where we see a variety of
-different formalisation choices,
-sometimes ad-hoc for the specific
-result or development.
-As a significant example, we observe that,
-while many authors rely
-on the native support for coinduction offered by
-proof assistants
-\cite{Honsell2001,Bengtson2016,Kahsai2008,Thiemann2019,Gay2020},
-many others prefer a more ``set-theoretic''
-approach
+%As a significant example, we observe that,
+While many authors rely on the native support for coinduction offered
+by proof assistants
+\cite{Honsell2001,Bengtson2016,Kahsai2008,Thiemann2019,Gay2020}, many
+others prefer a more ``set-theoretic'' approach
 \cite{Hirschkoff1997,Bengtson2009,Maksimovic2015,Pohjola2022}.
+\alb{Actually, Honsell gives up with guarded coinduction and switches to Park's half way}
 
@@
 
+\alb{This should not be a subsubsection at the same level of coinduction etc}
 \subsubsection{Evaluation criteria.}
-The idea behind our benchmark is to obtain evidence towards answering questions \cref{item:rq1,item:rq2,item:rq3}.
-We are therefore interested in not only the solutions, but also in the experience of solving the challenges with the chosen approach.
-Solutions to our challenges should be compared on three measures:
+The idea behind our benchmark is to obtain evidence towards answering
+questions \cref{item:rq1,item:rq2,item:rq3}. We are therefore
+interested not only in the solutions, but also in the experience of
+solving the challenges with the chosen approach.  Solutions to our
+challenges should be compared on three axis:
 \begin{enumerate}

### 2024-02-19 short-paper/short-paper.tex
Small edits for brevity and some spelling fixes.

@@
 libraries, and best practices for programming language
-mechanization. By providing a shared framework for systematically
-evaluating mechanization techniques, it catalyzed a significant
-shift within the programming language research community. As a
-result, scholarly publications increasingly incorporate
-mechanized proofs.
-%
-% The POPLMark challenge~\cite{POPLMark} accelerated the development of proof assistants, libraries, and best practices for programming language mechanisation by establishing a common ground for systematically evaluating mechanisation techniques.
-% In doing so, it spearheaded a shift in the programming language research community towards publications that include mechanised proofs.
-% POPLMark Reloaded~\cite{POPLMarkReloaded} proposed a similar program for proofs involving logical relations.
-% These challenges, and the developments they spurred, were limited in scope, but their authors noted the need to tackle topics such as coinduction and linearity in the future.
+mechanisation. By providing a shared framework for systematically
+evaluating mechanisation techniques, it catalyzed a significant
+shift towards publications that include mechanised proofs within the programming language research community.
 
@@
 
-In this spirit, we introduce a novel collection of benchmark problems
+In this spirit, we introduce a new collection of benchmark problems
 specifically crafted to tackle common challenges encountered during
-the mechanization of formal models for concurrent and distributed
+the mechanisation of formal models for concurrent and distributed
 programming languages. Our focus extends particularly to process
 calculi.
-%
-% In this spirit, we propose a new collection of benchmark problems
-% specifically designed to address typical issues that arise when
-% mechanising formal models of concurrent and distributed programming
-% languages in general, and process calculi in particular. 
 Concurrent calculi are notably subtle: For instance, it took nine
 years before the incorrect proof of type safety in the seminal paper
-by Honda et al.~.~\cite{Honda1998} was rectified in Yoshida \&
+by Honda et al.~\cite{Honda1998} was rectified in Yoshida \&
 Vasconcelos’s subsequent work\cite{Yoshida2007}.
-% for instance it took nine years before the incorrect proof of type
-% safety in the seminal paper by Honda et al was fixed
-% in~\cite{Yoshida2007}.
 Similarly, many
-key results in peer-reviewed papers on session types have turned out
+key results in papers on session types have turned out
 to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the
@@
 to reuse. For example, the authors in~\cite{Cruz-Filipe2021b} note that the high-level parts of
-mechanised proofs closely resemble the informal versions, and that the
+mechanised proofs closely resemble the informal ones, and that the
 main challenge lies in getting the foundations right.
 
-Our benchmark problems (explained in detail in \cref{app:challenges}) center around three critical aspects that frequently pose challenges in mechanizing concurrency theory:
-% focus on three key aspects which often cause difficulties when mechanising concurrency theory:
-\emph{linearity}, \emph{scope extrusion}, and \emph{coinductive
-  reasoning}.  Mechanisations must often address several of these
+Our benchmark problems (detailed in \cref{app:challenges}) center around three key aspects that frequently pose challenges when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
+Mechanisations must often address several of these
 aspects at the same time; however, we see the combination of
 techniques as a next step beyond the scope of this challenge, as
-discussed further in \Cref{sec:going-beyond}.  Our benchmark
-challenges are based on process calculi, as these provide a simple but
-realistic showcase of common issues encountered during mechanization endeavors
-% exhibit of typical issues encountered in mechanisation work.
-While the three highlighted aspects are not the sole sources of difficulties, they
-consistently manifest in most mechanizations of concurrency theory.
-% The three aspects are certainly not the only causes of difficulties,
-% but as we will discuss presently, they emerge in most mechanisations
-% of concurrency theory.
+discussed further in \Cref{sec:going-beyond}.
+Our benchmark challenges are based on process calculi, as these provide a simple but
+realistic showcase of common issues encountered during mechanisation efforts.
+While the three highlighted aspects are not the sole sources of difficulties, they emerge in most mechanisations of concurrency theory.
 
@@
 Common uses for coinduction are to define strong and weak versions of
-bisimulation, barbed notions of bisimulation, and prove trace equivalence. %,
-%etc.
+bisimulation, barbed notions of bisimulation, and prove trace equivalence.
 On the other hand, coinduction in proof assistants
 is notoriously difficult and its representation
-is non-uniform across development.
+is non-uniform across developments.
 E.g., \cite{Castro-Perez2021} notes that the most challenging
@@
 Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
-These problems center around well-established theorems and results, showcasing intriguing proof techniques that can be leveraged in future endeavors.
+These problems center around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in future endeavors.
 % as per design goal
@@
 solving the challenges with the chosen approach.  Solutions to our
-challenges should be compared on three axis:
+challenges should be compared on three axes:
 \begin{enumerate}

### 2024-02-19 short-paper/short-paper.tex
Reorganise the short paper.
This undoes the idea of splitting the related work in two sections,
since this never really worked out well.
This also cleans up the introduction by moving things elsewhere.

The sections about each challenge are not integrated yet, and neither
are the sections on the overall design and possible extensions.

@@
 
-\todo{These paragraphs are still quite messy and unfocused. They should focus on the motivation, i.e.\ mention what others have perceived as problems.}
+We have begun collecting solutions to our challenges on our website:
+%
+\begin{center}
+  \url{https://concurrentbenchmark.github.io/}
+\end{center}
+%
+In the longer term, we expect to use the website for promoting best practices, tutorials and guidelines derived from solutions to our challenges.
+We encourage anyone interested to try the challenges using their favourite tools and techniques, and to send us their solutions.
 
-\alb{I really do not like the duplication of the discussion of the
-  three issues here and in Sec 2. I understand the desire to highlight
-  noted issues, but it does not work. Some citations are unconvincing. Mostly, we are anticipating the description of the challenges. Let's unify this}
-\paragraph{Linearity.}
+\section{Design of the Benchmark}\label{sec:design-discussion}
+In this section, we outline the factors considered in designing the benchmark challenges.
+We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the evaluation criteria for solutions.
+
+Similarly to the authors of the POPLMark challenges, we seek to
+answer several questions:
+\alb{Should we stress we are interested in the \textbf{meta-theory} of process calculi, not}
+\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
+\item\label{item:rq1} What is the current state of the art in mechanising process calculi?
+\item\label{item:rq2} Which techniques and best practices can be recommended when starting formalisation projects involving process calculi?
+\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
+\end{enumerate}
+
+To provide a framework in which to answer these questions, our benchmark is designed to satisfy three main goals:
+\begin{enumerate}[label=\textbf{(G\arabic*)},leftmargin=10mm]
+\item\label{item:goal-comperison-accessibility} To enable the comparison of
+  proof mechanisation approaches, making the challenges accessible to
+  mechanisation experts who may be unfamiliar with concurrency theory;
+
+\item\label{item:goal-tutorials} To encourage the development of guidelines and
+  tutorials demonstrating and comparing existing proof mechanisation
+  techniques, libraries, and proof assistant features; and
+
+\item\label{item:goal-reusability} To prioritise the exploration of mechanisation
+  techniques that are reusable for future research.
+\end{enumerate}
+We also aim at strengthening the culture of mechanisation, by rallying the community to collaborate on exploring and developing new tools and techniques.
+
+\todo{Refer to the future benchmark section about other things one could do, but they still all need the basics}
+Still, these (and other) applications need the basic
+techniques that our challenge problems emphasize:
+Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
+These problems center around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in future endeavors.
+To achieve design goal \ref{item:goal-comperison-accessibility}, we have
+formulated our challenge problems to explore the three aspects
+(linearity, scope extrusion, coinduction) independently, so that they
+may be solved individually and in any order; each problem is
+reasonably small and easily understandable with basic knowledge of
+textbook concurrency theory, process calculi, and type theory.  The
+process calculus used in each challenge focuses on the features that
+we want to emphasize, and omits all constructs (such as choices) that
+would complicate the mechanisation without bringing additional
+tangible benefits and insights.  The minimality and uniformity of the
+calculi also allows us to target design goal
+\ref{item:goal-tutorials}.
+
+\subsubsection{Linearity.}
 In mechanised meta-theory, addressing linearity means choosing an appropriate representation of the linear context.
@@
 
-\todo{Insert related work focusing on noted issues with mechanising linearity.}
+\alb{We need more details about what the challenge is}
+For our challenge on linear reasoning we have chosen a type safety theorem
+for a system using session types, combining as usual type preservation with absence of errors.
+This allows us to introduce linear reasoning with few definitions.
+With our objective focused on linearity, we forego channel delegation
+and opt for a reduction approach rather than employing a transition
+system semantics.  Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
+define a syntax where restriction binds two names, highlighting
+their duality within the type system.  We have chosen a simple notion
+of well-formedness, and do not  consider
+deadlocks. While more sophisticated notions of well-formedness are
+interesting, the proofs of these properties would distract from the
+core linear aspects of the challenge.
 
-\paragraph{Scope extrusion.}
+While a linear context is perhaps best seen as a multi-set, most proof assistants have better support for list processing.
+The latter representation is intuitive, but may require establishing a large number of technical lemmata that are orthogonal to the mathematics of the problem under study, say proving type preservation for session types.
+In this space, several designs are possible: one can label occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}) or \alb{something --- ask Francisco} as in \cite{Castro2020}.
+Some authors instead impose a multiset structure over lists (\eg~\cite{ChaudhuriLR19,Danielsson12}); still, the effort required to develop the infrastructure is significant.
+
+A different approach is to bypass the problem of context splitting by adopting familiar ideas from algorithmic linear type checking: this is known as \emph{typing with leftovers}, as exemplified in~\cite{DBLP:conf/forte/ZalakainD21}.
+Whatever the choice, list-based encodings can be refined to be intrinsically-typed, if the proof assistant does support dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
+
+For something completely different, one could adopt a \emph{sub-structural} meta-logical framework: here the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
+In fact, the linear function space will handle all context operations implicitly, including splitting and substitution.
+One such framework is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session types in~\cite{Bock2016}).
+Unfortunately, \emph{Celf} does not yet fully support the verification of meta-theoretic properties.
+A compromise is the so called \emph{two-level} approach, where one encodes a sub-structural specification logic in a mainstream proof assistant and then uses that logic to state and prove linear properties (for a recent example, see~\cite{Felty:MSCS21}).
+
+\alb{This needs some work: we do not describe the challenge at all and the citations are not the one I had in mind}
+\subsubsection{Scope extrusion.}
 The encoding of binding constructs has been extensively studied
@@
 
-\todo{Insert related work focusing on noted issues when mechanising scope extrusion.}
-
 \cite{Maksimovic2015} does not have name restriction, thus dodging the challenge of scope extrusion.
@@
 
-\paragraph{Coinduction.} \todo{explain more what the challenge does.}
-\lore{I have tried this:
-  the first paragraph motivates (``why is this difficult?''), while
-  the second talks about the challenge.}
+\alb{We need more details about what the challenge is}
+This is the challenge most closely related to the POPLMark challenge
+since it concerns the properties of binders in restrictions and input.
+
+The case study \cite{AmbalLS21} compares four approaches to encoding binders in Coq as a first step towards developing tools for working with higher-order process calculi.
+They found that working directly with de Bruijn indices was easiest and shortest since the existing approaches developed for $\lambda$-calculus binders do not work well with scope extrusion.
+Specifically, the locally nameless representation~\cite{Chargueraud2012} cannot avoid direct manipulation of de Bruijn indices when defining the semantics of scope extrusion; cofinite quantification provides no benefits when working under binders; nominal approaches~\cite{Pitts2003} require explicitly giving and validating sets of free names during scope extrusion; and higher-order abstract syntax (HOAS)~\cite{Pfenning1988} requires additional axioms to work with contexts.
+
+Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, \ie, induction principles on functions to work with HOAS encodings.
+Some proof assistants such as Abella~\cite{Baelde2014} support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging~\cite{Momigliano2012}.
+
+\subsubsection{Coinduction.}
 Coinduction is fundamental to establish properties of processes that
@@
 parallel composition and substitution.
-%
-%% The introduction of bisimulation
-%% is one of the first steps, when studying the
-%% (observable) behaviour of process calculi, both
-%% in classical textbooks (e.g., \cite{picalcbook})
-%% and in existing 
-%% mechanisations
-%% \cite{Thiemann2019,Honsell2001,Bengtson2016,Kahsai2008,Gay2020,
-%% Hirschkoff1997,Bengtson2009,Maksimovic2015}.
 
-%% , where we see a variety of
-%% different formalisation choices.
-%% As a significant example, we observe that,
-%% while many authors rely
-%% on the native support for coinduction offered by
-%% proof assistants
-%% \cite{Thiemann2019,Honsell2001,Bengtson2016,Kahsai2008,Gay2020},
-%% many others prefer a more ``set-theoretic''
-%% approach
-%% \cite{Hirschkoff1997,Bengtson2009,Maksimovic2015}.
-
-%Common uses for coinduction are to define strong and weak versions of
-%bisimulation, barbed notions of bisimulation, prove trace equivalence,
-%etc.
-
-%In this challenge we study the notion of strong barbed bisimulation
-%over a fragment of the $\pi$-calculus.
-
-\todo{Insert related work focusing on noted issues when mechanising coinduction.}
-
-\alb{This subsection does not make a lot of sense here -- introductions do not have subsections. Let's move it in ``Design'' }
-\subsection{Goals}% and how to contribute}
-Similarly to the authors of the POPLMark challenges, we seek to
-answer several questions:
-\alb{Should we stress we are interested in the \textbf{meta-theory} of process calculi, not}
-\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
-\item\label{item:rq1} What is the current state of the art in mechanising process calculi?
-\item\label{item:rq2} Which techniques and best practices can be recommended when starting formalisation projects involving process calculi?
-\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
-\end{enumerate}
-
-To provide a framework in which to answer these questions, our benchmark is designed to satisfy three main goals:
-\begin{enumerate}[label=\textbf{(G\arabic*)},leftmargin=10mm]
-\item\label{item:goal-comperison-accessibility} To enable the comparison of
-  proof mechanisation approaches, making the challenges accessible to
-  mechanisation experts who may be unfamiliar with concurrency theory;
-
-\item\label{item:goal-tutorials} To encourage the development of guidelines and
-  tutorials demonstrating and comparing existing proof mechanisation
-  techniques, libraries, and proof assistant features; and
-
-\item\label{item:goal-reusability} To prioritise the exploration of mechanisation
-  techniques that are reusable for future research.
-\end{enumerate}
-We also aim at strengthening the culture of mechanisation, by rallying the community to collaborate on exploring and developing new tools and techniques.
-
-We have begun collecting solutions to our challenges on our website:
-%
-\begin{center}
-  \url{https://concurrentbenchmark.github.io/}
-\end{center}
-%
-In the longer term, we expect to use the website for promoting best practices, tutorials and guidelines derived from solutions to our challenges.
-We encourage anyone interested to try the challenges using their favourite tools and techniques, and to send us their solutions.
-
-% - Design of the benchmark
-%   - We first list some general design remarks, then for each challenge
-%   - General remarks
-%     - We do not aim to be comprehensive (list other topics)
-%     - However, we believe the best practices developed could also be of use in those topics
-%     - We base our challenges on basic textbook theory with interesting proof techniques
-%     - The challenges are independent (maybe note ``failure'' of POPLMark that everyone only did the first part)
-%     - The challenges are as minimal as possible
-%   - Linearity
-%     - Type preservation for session types
-%     - The main issue is context splitting
-%     - Session types are pervasive in the field and simple to understand
-%     - Binding two names makes duality obvious
-%     - Brief overview of approaches and issues with them
-%   - Scope extrusion
-%     - Process calculi have several binders, but only restriction is different from ``normal'' ones
-%     - The main issue is scope extrusion
-%     - The pros and cons of established techniques for handling binders are different for process calculi, and it is unclear how exactly
-%     - Brief overview of approaches and issues with them
-%   - Coinduction
-%     - Strong barbed bisimulation and congruence
-%     - The main issue is the encoding of coinductive up-to techniques
-%     - Strong equivalences to simplify (we don't have to abstract over internal transitions)
-%     - Why barbed equivalences?
-%     - No delegation to simplify
-%     - Brief overview of approaches and issues with them
-%   - Evaluation criteria
-%     - We care not only about solutions, but about experiences and comparisons
-%     - Mechanisation overhead
-%     - Adequacy of the formal statements
-%     - Cost of entry for tools and techniques
-%     - Slight changes are OK if they are justified and adequate
-\section{Design of the Benchmark}\label{sec:design-discussion}
-\todo{Add references to support our claims and decisions in this whole section. Here is where we should mention existing approaches.}
-
-In this section, we outline the factors considered in designing the benchmark challenges.
-We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the evaluation criteria for solutions.
-
-\subsubsection{General design remarks.}
-Like the POPLMark and POPLMark Reloaded challenges, our challenge is
-not meant to be comprehensive: applications such as multiparty session
-types~\cite{10.1145/2827695,10.1145/3290343},
-choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation
-types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696},
-or encodings between different
-calculi~\cite{DBLP:journals/iandc/Gorla10,%,DBLP:conf/forte/CairesP16,DBLP:journals/iandc/DardhaGS17,
-  DBLP:conf/ecoop/ScalasDHY17}
-  %,DBLP:journals/iandc/KouzapasPY19,10.1145/3479394.3479407}
-are not directly covered.
-
-\alb{I've commented some citations about
-  encodings. Is it a good idea to start the section describing what we \textbf{do not do}?}
-
-Still, these (and other) applications need the basic
-techniques that our challenge problems emphasize:
-Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
-These problems center around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in future endeavors.
-% as per design goal
-% \ref{item:goal-reusability}, our problems are drawn from the basic
-% meta-theory of process calculi (without requiring the development of
-% new theory), and focus on well-known theorems and results that involve
-% interesting proof techniques that may be reused in further work. 
-To
-achieve design goal \ref{item:goal-comperison-accessibility}, we have
-formulated our challenge problems to explore the three aspects
-(linearity, scope extrusion, coinduction) independently, so that they
-may be solved individually and in any order; each problem is %should be
-reasonably small and easily understandable with basic knowledge of
-textbook concurrency theory, process calculi, and type theory.  The
-process calculus used in each challenge focuses on the features that
-we want to emphasize, and omits all constructs (such as choices) that
-would complicate the mechanisation without bringing additional
-tangible benefits and insights.  The minimality and uniformity of the
-calculi also allows us to target design goal
-\ref{item:goal-tutorials}.
-
-\subsubsection{Linearity.}
-
-\alb{I would put the challenge before the literature}
-For our challenge on linear reasoning we have chosen a type safety theorem
-for a system using session types, combining as usual type preservation with absence of errors.
-% , whose use are common in concurrency theory.
-This allows us to introduce linear reasoning with few definitions.
-With our objective focused on linearity, we forego channel delegation
-and opt for a reduction approach rather than employing a transition
-system semantics.  Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
-define a syntax where restriction binds two names, highlighting
-their duality within the type system.  We have chosen a simple notion
-of well-formedness, and do not  consider
-deadlocks. While more sophisticated notions of well-formedness are
-interesting, the proofs of these properties would distract from the
-core linear aspects of the challenge.
-
-While a linear context is perhaps best seen as a multi-set, most proof assistants have better support for list processing.
-The latter representation is intuitive, but may require establishing a large number of technical lemmata that are orthogonal to the mathematics of the problem under study, say proving type preservation for session types.
-In this space, several designs are possible: one can label occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}) or \alb{something --- ask Francisco} as in \cite{Castro2020}.
-Some authors instead impose a multiset structure over lists (\eg~\cite{ChaudhuriLR19,Danielsson12}); still, the effort required to develop the infrastructure is significant.
-
-A different approach is to bypass the problem of context splitting by adopting familiar ideas from algorithmic linear type checking: this is known as \emph{typing with leftovers}, as exemplified in~\cite{DBLP:conf/forte/ZalakainD21}.
-Whatever the choice, list-based encodings can be refined to be intrinsically-typed, if the proof assistant does support dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
-
-For something completely different, one could adopt a \emph{sub-structural} meta-logical framework: here the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
-In fact, the linear function space will handle all context operations implicitly, including splitting and substitution.
-One such framework is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session types in~\cite{Bock2016}).
-Unfortunately, \emph{Celf} does not yet fully support the verification of meta-theoretic properties.
-A compromise is the so called \emph{two-level} approach, where one encodes a sub-structural specification logic in a mainstream proof assistant and then uses that logic to state and prove linear properties (for a recent example, see~\cite{Felty:MSCS21}).
-
-\todo{Insert discussion/related work focusing on existing techniques and how our benchmark design makes it possible to compare them.}
-
-\alb{This needs some work: we do not describe the challenge at all and the citations are not the one I had in mind}
-\subsubsection{Scope extrusion.}
-This is the challenge most closely related to the POPLMark challenge
-since it concerns the properties of binders in restrictions and input.
-
-The case study \cite{AmbalLS21} compares four approaches to encoding binders in Coq as a first step towards developing tools for working with higher-order process calculi.
-They found that working directly with de Bruijn indices was easiest and shortest since the existing approaches developed for $\lambda$-calculus binders do not work well with scope extrusion.
-Specifically, the locally nameless representation~\cite{Chargueraud2012} cannot avoid direct manipulation of de Bruijn indices when defining the semantics of scope extrusion; cofinite quantification provides no benefits when working under binders; nominal approaches~\cite{Pitts2003} require explicitly giving and validating sets of free names during scope extrusion; and higher-order abstract syntax (HOAS)~\cite{Pfenning1988} requires additional axioms to work with contexts.
-
-\todo{Insert discussion/related work focusing on existing techniques and how our benchmark design makes it possible to compare them.}
-
-Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, \ie, induction principles on functions to work with HOAS encodings.
-Some proof assistants such as Abella~\cite{Baelde2014} support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging~\cite{Momigliano2012}.
-
-\subsubsection{Coinduction.}
+\alb{We need more details about what the challenge is}
 \alb{I kinda object to ``seamlessly''. In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky. I guess Isabelle is better}
@@
   the command \texttt{coinduction\_upto}? Is it related?}
-\todo{Insert discussion/related work focusing on existing techniques and how our benchmark design makes it possible to compare them.}
-\lore{I have added something below... Any advice on what more we need?}
+\todo{Explain the the point of the congruence theorem is that it should be ``easy'' once you have the other one}
 
-%As a significant example, we observe that,
 While many authors rely on the native support for coinduction offered
@@
 
-%% \lore{Need to integrate the following with the above.}
-%% Different process-calculi
-%% mechanisations
-%% implement different techniques;
-%% most of the time these 
-%% are ad-hoc efforts to achieve a specific result.
-%% \lore{cite and substantiate.}
-%% In parallel, although most proof assistants
-%% offer support for coinduction in their standard libraries,
-%% research has mainly focused on the foundations
-%% of the specific system, more than on
-%% applicability.
-%% \lore{cite and substantiate.}
-
-\alb{This should not be a subsubsection at the same level of coinduction etc}
 \subsubsection{Evaluation criteria.}
@@
 
+Like the POPLMark and POPLMark Reloaded challenges, our challenge is
+not meant to be comprehensive: applications such as multiparty session
+types~\cite{10.1145/2827695,10.1145/3290343},
+choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation
+types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696},
+or encodings between different
+calculi~\cite{DBLP:journals/iandc/Gorla10, DBLP:conf/ecoop/ScalasDHY17} are not directly covered.\todo{Should we cite other/different encoding papers? Which ones are the ``standard'' citations?}
+
+\todo{This is a list of interesting research problems, but they are not challenges per se. Trim down.}
 On the mechanisation side, the present challenge proposes mechanising the proofs of theorems.
@@
 Additionally, while our current challenge is agnostic about proof automation, a more specific challenge could be designed with the objective of automating aspects of the proofs.
-Finally, a challenge could propose the integration with other formal reasoning tools, such as model checkers.
-These tools are in common use in the field; combining automated proofs with proof assistants offers the potential to ease the path towards larger proofs.
 

### 2024-02-19 short-paper/short-paper.tex
Brit-ify spelling.

@@
 mechanisation. By providing a shared framework for systematically
-evaluating mechanisation techniques, it catalyzed a significant
+evaluating mechanisation techniques, it catalysed a significant
 shift towards publications that include mechanised proofs within the programming language research community.
@@
 
-Our benchmark problems (detailed in \cref{app:challenges}) center around three key aspects that frequently pose challenges when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
+Our benchmark problems (detailed in \cref{app:challenges}) centre around three key aspects that frequently pose challenges when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
 Mechanisations must often address several of these
@@
 Still, these (and other) applications need the basic
-techniques that our challenge problems emphasize:
+techniques that our challenge problems emphasise:
 Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
-These problems center around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in future endeavors.
+These problems centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in future endeavours.
 To achieve design goal \ref{item:goal-comperison-accessibility}, we have
@@
 process calculus used in each challenge focuses on the features that
-we want to emphasize, and omits all constructs (such as choices) that
+we want to emphasise, and omits all constructs (such as choices) that
 would complicate the mechanisation without bringing additional
@@
 Some systems provide more than one abstraction to use coinduction.
-A good example of this are the parameterized coinduction~\cite{Hur2013}
+A good example of this are the parameterised coinduction~\cite{Hur2013}
 and interaction tree~\cite{Xia2019} libraries for Coq.

### 2024-02-19 short-paper/short-paper.tex
revision of the penultimate paragraph of intro

@@
 to reuse. For example, the authors in~\cite{Cruz-Filipe2021b} note that the high-level parts of
-mechanised proofs closely resemble the informal ones, and that the
+mechanised proofs closely resemble the informal ones, while the
 main challenge lies in getting the foundations right.
 
-Our benchmark problems (detailed in \cref{app:challenges}) centre around three key aspects that frequently pose challenges when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
-Mechanisations must often address several of these
-aspects at the same time; however, we see the combination of
-techniques as a next step beyond the scope of this challenge, as
-discussed further in \Cref{sec:going-beyond}.
-Our benchmark challenges are based on process calculi, as these provide a simple but
-realistic showcase of common issues encountered during mechanisation efforts.
-While the three highlighted aspects are not the sole sources of difficulties, they emerge in most mechanisations of concurrency theory.
+Our benchmark problems (detailed in \cref{app:challenges}) consider
+\emph{in isolation} three key aspects that frequently pose challenges
+when mechanising concurrency theory: \emph{linearity}, \emph{scope
+  extrusion}, and \emph{coinductive reasoning}.  Mechanisations must
+often address several of these aspects at the same time; however, we
+see the combination of techniques as a next
+step%beyond the scope of this challenge
+, as discussed further in \Cref{sec:going-beyond}.  Our benchmark
+challenges are based on process calculi, as these provide a simple but
+realistic showcase of common issues encountered during mechanisation
+efforts.
+% While the three highlighted aspects are not the sole sources of difficulties, they emerge in most mechanisations of concurrency theory.
 
@@
 %
-In the longer term, we expect to use the website for promoting best practices, tutorials and guidelines derived from solutions to our challenges.
-We encourage anyone interested to try the challenges using their favourite tools and techniques, and to send us their solutions.
+In the longer term, we expect to use the website for promoting best
+practices, tutorials and guidelines derived from solutions to our
+challenges.  We encourage anyone interested to try the challenges
+using their favourite tools and techniques, and to send us their
+solutions.
 

### 2024-02-19 short-paper/short-paper.tex
Small edits to introduction for brevity and clarity.

@@
 shift towards publications that include mechanised proofs within the programming language research community.
-
-POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a program akin to
-the original POPLMark challenge, specifically targeting proofs related
-to logical relations. While these initiatives had a narrow focus,
-their authors recognized the importance of addressing broader topics
-in the future, including coinduction and linearity.
+POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a similar programme targeting proofs using logical relations.
+These initiatives had a narrow focus, but their authors recognised the importance of addressing topics such as coinduction and linearity in the future.
 
@@
 the mechanisation of formal models for concurrent and distributed
-programming languages. Our focus extends particularly to process
-calculi.
+programming languages.
+We particularly focus on process calculi.
 Concurrent calculi are notably subtle: For instance, it took nine
@@
 Vasconcelos’s subsequent work\cite{Yoshida2007}.
-Similarly, many
-key results in papers on session types have turned out
-to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the
-value of machine-checkable proofs.
+Similarly, several key results in papers on session types have turned out
+to be false later on, illustrating the
+value of machine-checked proofs~\cite{Gay2020,10.1145/3290343}.
 
@@
 %
-In the longer term, we expect to use the website for promoting best
-practices, tutorials and guidelines derived from solutions to our
-challenges.  We encourage anyone interested to try the challenges
-using their favourite tools and techniques, and to send us their
-solutions.
+In the future, we expect to use the website to promote best practices and tutorials derived from solutions to our challenges.
+We encourage everyone to try the challenges using their favourite techniques, and to send us their solutions.
 

### 2024-02-19 short-paper/short-paper.tex
Edit intro part of design section to integrate design goals.

@@
 step%beyond the scope of this challenge
-, as discussed further in \Cref{sec:going-beyond}.  Our benchmark
+, as discussed further in \cref{sec:going-beyond}.  Our benchmark
 challenges are based on process calculi, as these provide a simple but
@@
 \section{Design of the Benchmark}\label{sec:design-discussion}
-In this section, we outline the factors considered in designing the benchmark challenges.
-We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the evaluation criteria for solutions.
+In this section, we outline our considerations in designing the benchmark challenges.
+We will begin with our design goals, then describe design considerations for each challenge problem, and finally suggest some evaluation criteria for solutions.
 
 Similarly to the authors of the POPLMark challenges, we seek to
-answer several questions:
-\alb{Should we stress we are interested in the \textbf{meta-theory} of process calculi, not}
+answer several questions about mechanising the meta-theory of process calculi:
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
-\item\label{item:rq1} What is the current state of the art in mechanising process calculi?
-\item\label{item:rq2} Which techniques and best practices can be recommended when starting formalisation projects involving process calculi?
+\item\label{item:rq1} What is the current state of the art in tools for mechanising the meta-theory of process calculi?
+\item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
 \item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
@@
 
-To provide a framework in which to answer these questions, our benchmark is designed to satisfy three main goals:
+To provide a framework in which to answer these questions, our benchmark is designed to satisfy three main design goals:
 \begin{enumerate}[label=\textbf{(G\arabic*)},leftmargin=10mm]
 \item\label{item:goal-comperison-accessibility} To enable the comparison of
-  proof mechanisation approaches, making the challenges accessible to
+  proof mechanisation approaches by making the challenges accessible to
   mechanisation experts who may be unfamiliar with concurrency theory;
@@
 
-\todo{Refer to the future benchmark section about other things one could do, but they still all need the basics}
-Still, these (and other) applications need the basic
-techniques that our challenge problems emphasise:
-Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
-These problems centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in future endeavours.
 To achieve design goal \ref{item:goal-comperison-accessibility}, we have
@@
 would complicate the mechanisation without bringing additional
-tangible benefits and insights.  The minimality and uniformity of the
+tangible insights.  The minimality and uniformity of the
 calculi also allows us to target design goal
 \ref{item:goal-tutorials}.
+Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
+Our challenges centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in many applications (as we will further discuss in \cref{sec:going-beyond}).
 

### 2024-02-19 short-paper/short-paper.tex
some writes to sec 3

--- 2024-02-19 short-paper/short-paper.tex
@@
 
-\section{Design of the Benchmark}\label{sec:design-discussion}
+\section{The Benchmark and its Design}\label{sec:design-discussion}
+\alb{Next paragraph needs rewriting}
 In this section, we outline the factors considered in designing the benchmark challenges.
@@
 answer several questions:
-\alb{Should we stress we are interested in the \textbf{meta-theory} of process calculi, not}
+% \alb{Should we stress we are interested in the \textbf{meta-theory} of process calculi, not}
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
@@
 \end{enumerate}
+%
+Note that the challenges revolve around the mechanization of the
+\emph{meta-theory} of process calculi, rather than interpreters or
+type checkers for the latter.
 
@@
 \paragraph{Extending and combining the current challenges.}
-We have aimed to reduce the overlap of the aspects (linearity, scope extrusion, coinduction) that our challenges involve.
-Most concurrency theory research needs several of the aspects and often also other constructs that we have chosen not to cover for simplicity.
-We suggest three extensions that would be useful for many applications:
+We have aimed to reduce the overlap of the aspects (linearity, scope
+extrusion, coinduction) that our challenges involve.  Most concurrency
+theory research needs several of those aspects and in fact there is
+anedoctical evidence of the difficulties that such a combination may induce
+in a mechanization,
+e.g.~\cite{DBLP:conf/forte/ZalakainD21}. Furthermore, one can consider
+other constructs that we have chosen not to cover for simplicity.  We
+suggest three extensions that would be useful for many applications:
 \emph{(1)} adding choice constructs to both type and process levels,
-\emph{(2)} introducing recursion and recursive types for typed systems, and
-\emph{(3)} incorporating channel delegation in calculi lacking it.
+\emph{(2)} introducing recursion and recursive types for typed
+systems, and \emph{(3)} incorporating channel delegation in calculi
+lacking it.
 
-\cite{DBLP:conf/forte/ZalakainD21} notes that the combination of linearity and infinite behaviour is difficult since it requires uniquely determining the resources consumed by a process.
-\cite{Zalakain2019} notes that mixing replication and linearity seems to be very difficult without using explicit type contexts.
+% \cite{DBLP:conf/forte/ZalakainD21} notes that the combination of linearity and infinite behaviour is difficult since it requires uniquely determining the resources consumed by a process.
+% \cite{Zalakain2019} notes that mixing replication and linearity seems to be very difficult without using explicit type contexts.
 
 \paragraph{Addressing new calculi and new features.}
-\todo{Mention the higher-order pi calculus? If so, also mention~\cite{Hirsch2022}}
-Extending in this direction means proposing new challenges covering different features of process calculi, or different proof mechanisation features.
-Some interesting aspects of message-passing calculi to be explored in further challenges could be multiparty session types~\cite{10.1145/2827695} and choreographies~\cite{Carbone2013}, as their meta-theory includes aspects -- \eg well-formedness conditions on global types, partiality of end-point projection function, \etc -- that we do not address here.
-Also, one could easily design a challenge that goes beyond our coinduction challenge, by exploring different notions of bisimilarity (barbed, weak, \etc) and trace equivalence.
+% \todo{Mention the higher-order pi calculus? If so, also
+%   mention}
+
+Like the POPLMark and POPLMark Reloaded challenges, our % challenge
+is
+not meant to be comprehensive: Extending it means proposing new
+challenges covering different features of process calculi, or
+different proof mechanisation features.  Some interesting aspects of
+message-passing calculi %to be explored in further challenges could be
+include
+multiparty session types~\cite{10.1145/2827695} and
+choreographies~\cite{Carbone2013}, as their meta-theory includes
+aspects -- \eg well-formedness conditions on global types, partiality
+of end-point projection function, \etc -- that we do not address
+here. Other revolve around higher-order calculi~\cite{Hirsch2022},
+conversation types~\cite{DBLP:journals/tcs/CairesV10},
+psi-calculi~\cite{lmcs:696}, or encodings between different
+calculi~\cite{DBLP:journals/iandc/Gorla10,
+  DBLP:conf/ecoop/ScalasDHY17}.\todo{Should we cite other/different
+  encoding papers? Which ones are the ``standard'' citations?}
+
+Also, one could easily design a challenge that goes beyond our
+coinduction challenge, by exploring different notions of process  equivalence.
 %
@@
 
-Like the POPLMark and POPLMark Reloaded challenges, our challenge is
-not meant to be comprehensive: applications such as multiparty session
-types~\cite{10.1145/2827695,10.1145/3290343},
-choreographies~\cite{DBLP:journals/jar/CruzFilipeMP23}, conversation
-types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696},
-or encodings between different
-calculi~\cite{DBLP:journals/iandc/Gorla10, DBLP:conf/ecoop/ScalasDHY17} are not directly covered.\todo{Should we cite other/different encoding papers? Which ones are the ``standard'' citations?}
-
-\todo{This is a list of interesting research problems, but they are not challenges per se. Trim down.}
-On the mechanisation side, the present challenge proposes mechanising the proofs of theorems.
-An interesting avenue to explore would be to take advantage of other proof assistant features like extracting certified implementations.
-Additionally, while our current challenge is agnostic about proof automation, a more specific challenge could be designed with the objective of automating aspects of the proofs.
+% \todo{This is a list of interesting research problems, but they are
+%   not challenges per se. Trim down.}  On the mechanisation side, the
+% present challenge proposes mechanising the proofs of theorems.
+Finally, another interesting avenue to explore would be to take
+advantage of certain proof assistant features: for example, extracting
+certified implementations (\cite{Castro-Perez2021}) or exploiting
+proof automation, such as the one offered by the \emph{Hammer}
+protocol.
+% a more specific
+% challenge could be designed with the objective of automating aspects
+% of the proofs.
 
 Ultimately, the current challenges can be extended in several worthwhile directions, and we look forward to a future when they indeed are.
-We see the current challenge as setting the foundation for those future extensions.
+% We see the current challenge as setting the foundation for those future extensions.
 It is our hope and aim that our challenges will move us closer to a future where the key basic proof techniques are as easy to mechanise as they are to write on paper.

### 2024-02-19 short-paper/short-paper.tex
linearity section revisited

@@
 
-Our benchmark problems (detailed in \cref{app:challenges}) consider
+Our benchmark problems (detailed in \cref{app:challenges} and on the website) consider
 \emph{in isolation} three key aspects that frequently pose challenges
@@
 %
-In the longer term, we expect to use the website for promoting best
-practices, tutorials and guidelines derived from solutions to our
-challenges.  We encourage anyone interested to try the challenges
-using their favourite tools and techniques, and to send us their
-solutions.
+In the future, we expect to use the website to promote best practices and tutorials derived from solutions to our challenges.
+We encourage everyone to try the challenges using their favourite techniques, and to send us their solutions.
 
-\section{The Benchmark and its Design}\label{sec:design-discussion}
+\section{The Benchmark and its design}\label{sec:design-discussion}
 \alb{Next paragraph needs rewriting}
@@
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
-\item\label{item:rq1} What is the current state of the art in mechanising process calculi?
-\item\label{item:rq2} Which techniques and best practices can be recommended when starting formalisation projects involving process calculi?
+\item\label{item:rq1} What is the current state of the art in tools for mechanising the meta-theory of process calculi?
+\item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
 \item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
@@
 would complicate the mechanisation without bringing additional
-tangible benefits and insights.  The minimality and uniformity of the
+tangible insights.  The minimality and uniformity of the
 calculi also allows us to target design goal
 \ref{item:goal-tutorials}.
+Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
+Our challenges centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in many applications (as we will further discuss in \cref{sec:going-beyond}).
 
 \subsubsection{Linearity.}
-In mechanised meta-theory, addressing linearity means choosing an appropriate representation of the linear context.
-The crux of the matter is \emph{context splitting}, as showcased by the typing rule for parallel composition.
 
-\cite{Brady2017} notes that using algebraic effects is unwieldy when there are many communication channels and suggests an experimental approach using dependent uniqueness types.
-\cite{Zalakain2019} notes that systems lacking linear types must simulate linearity, and that using a linearity predicate complicates congruences.
-They had to limit themselves to an asymmetric structural precongruence to be able to prove subject reduction.
+% \cite{Brady2017} notes that using algebraic effects is unwieldy when there are many communication channels and suggests an experimental approach using dependent uniqueness types.
+% \cite{Zalakain2019} notes that systems lacking linear types must simulate linearity, and that using a linearity predicate complicates congruences.
+% They had to limit themselves to an asymmetric structural precongruence to be able to prove subject reduction.
 
-\alb{We need more details about what the challenge is}
-For our challenge on linear reasoning we have chosen a type safety theorem
-for a system using session types, combining as usual type preservation with absence of errors.
-This allows us to introduce linear reasoning with few definitions.
-With our objective focused on linearity, we forego channel delegation
-and opt for a reduction approach rather than employing a transition
-system semantics.  Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
-define a syntax where restriction binds two names, highlighting
-their duality within the type system.  We have chosen a simple notion
-of well-formedness, and do not  consider
-deadlocks. While more sophisticated notions of well-formedness are
-interesting, the proofs of these properties would distract from the
-core linear aspects of the challenge.
+For our challenge on linear reasoning we have chosen a type safety
+theorem for a system using session types, combining as usual subject reduction with absence of errors.  This allows us to introduce
+linear reasoning with few definitions.  With this focus, we forego
+channel delegation: input and output receives/sends only values, which
+do not include channels, hence the topology of the communication
+network described by a process cannot change.  We do not allow
+recursion or replication in the syntax, hence no infinite behaviours
+can be expressed. In particular, we only model linear (as opposed to
+shared) channels. Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
+define a syntax where restriction binds two names, highlighting their
+duality within the type system. The latter enforces
+ two invariants, namely no endpoint is used simultaneously by parallel processes and
+ the two endpoints of the same session have dual types.
+The operational semantics is formulated in term of reductions rather
+than via a transition system, allowing implementors to explore
+alternative ways to encode structural congruence\footnote{\alb{Should we mention this in challenge 2?}}.
 
-While a linear context is perhaps best seen as a multi-set, most proof assistants have better support for list processing.
-The latter representation is intuitive, but may require establishing a large number of technical lemmata that are orthogonal to the mathematics of the problem under study, say proving type preservation for session types.
-In this space, several designs are possible: one can label occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}) or \alb{something --- ask Francisco} as in \cite{Castro2020}.
-Some authors instead impose a multiset structure over lists (\eg~\cite{ChaudhuriLR19,Danielsson12}); still, the effort required to develop the infrastructure is significant.
+The statement of subject reduction is the usual one, encompassing type
+preservation for structural congruence. W.r.t.\ type safety, we have
+chosen a simple notion of what an error is: well-typed processes are a
+well-formed in the sense that they do not use endpoints in a non-dual
+way.  While more sophisticated notions of well-formedness that
+prevents deadlocks are interesting, the proofs of these properties
+would distract from the core linear aspects of the challenge.
+
+In mechanised meta-theory, addressing linearity means choosing an
+appropriate representation of the linear context.  The crux of the
+matter is \emph{context splitting}, as showcased by the typing rule
+for parallel composition.  While a linear context is perhaps best seen
+as a multi-set, most proof assistants have better support for list
+processing.  The latter representation is intuitive, but may require
+establishing a large number of technical lemmata that are orthogonal
+to the mathematics of the problem under study, in our case proving
+type safety for session types.  In this space, several designs are
+possible: one can label occurrences of resources to constrain their
+usage (\eg~\cite{CicconeP20}); some authors instead impose a multiset
+structure over lists
+(\eg~\cite{Danielsson12,ChaudhuriLR19}). Alternatively, contexts can
+be implemented as finite maps (\cite{Castro2020}), whose operations
+are sensitive to a linear discipline. In all these cases, the effort
+required to develop the infrastructure is significant.
 
@@
 
+\subsubsection{Scope extrusion.}
 \alb{This needs some work: we do not describe the challenge at all and the citations are not the one I had in mind}
-\subsubsection{Scope extrusion.}
+
 The encoding of binding constructs has been extensively studied

### 2024-02-20 short-paper/short-paper.tex
Minor fixes.

@@
 
-Our benchmark problems (detailed in \cref{app:challenges} and on the website) consider
+Our benchmark problems (detailed in \cref{app:challenges} and on our website) consider
 \emph{in isolation} three key aspects that frequently pose challenges
@@
   extrusion}, and \emph{coinductive reasoning}.  Mechanisations must
-often address several of these aspects at the same time; however, we
+often address several of these aspects; however, we
 see the combination of techniques as a next
-step%beyond the scope of this challenge
-, as discussed further in \cref{sec:going-beyond}.  Our benchmark
-challenges are based on process calculi, as these provide a simple but
+step, as discussed in \cref{sec:going-beyond}.
+Our benchmark challenges are based on process calculi, as these provide a simple but
 realistic showcase of common issues encountered during mechanisation
@@
 
-\section{The Benchmark and its design}\label{sec:design-discussion}
+\section{The Benchmark and its Design}\label{sec:design-discussion}
 \alb{Next paragraph needs rewriting}
@@
 
-Similarly to the authors of the POPLMark challenges, we seek to
-answer several questions:
-% \alb{Should we stress we are interested in the \textbf{meta-theory} of process calculi, not}
+Similarly to the authors of the POPLMark challenges, we seek to answer several questions:
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
@@
 %
-Note that the challenges revolve around the mechanization of the
+\todo{Frederik: I'm not sure this note is necessary}
+Note that the challenges revolve around the mechanisation of the
 \emph{meta-theory} of process calculi, rather than interpreters or
@@
 extrusion, coinduction) that our challenges involve.  Most concurrency
-theory research needs several of those aspects and in fact there is
-anedoctical evidence of the difficulties that such a combination may induce
-in a mechanization,
+theory research needs several of these aspects and in fact there is
+anedoctal evidence of the difficulties that such a combination may induce
+in a mechanisation,
 e.g.~\cite{DBLP:conf/forte/ZalakainD21}. Furthermore, one can consider
@@
 \paragraph{Addressing new calculi and new features.}
-% \todo{Mention the higher-order pi calculus? If so, also
-%   mention}
-
-Like the POPLMark and POPLMark Reloaded challenges, our % challenge
-is
-not meant to be comprehensive: Extending it means proposing new
+Like the POPLMark and POPLMark Reloaded challenges, ours
+is not meant to be comprehensive: Extending it means proposing new
 challenges covering different features of process calculi, or
 different proof mechanisation features.  Some interesting aspects of
-message-passing calculi %to be explored in further challenges could be
-include
+message-passing calculi include
 multiparty session types~\cite{10.1145/2827695} and
@@
 calculi~\cite{DBLP:journals/iandc/Gorla10,
-  DBLP:conf/ecoop/ScalasDHY17}.\todo{Should we cite other/different
-  encoding papers? Which ones are the ``standard'' citations?}
+  DBLP:conf/ecoop/ScalasDHY17}.
+\todo{Should we cite other/different encoding papers? Which ones are the ``standard'' citations?}
 
@@
 
-% \todo{This is a list of interesting research problems, but they are
-%   not challenges per se. Trim down.}  On the mechanisation side, the
-% present challenge proposes mechanising the proofs of theorems.
 Finally, another interesting avenue to explore would be to take
 advantage of certain proof assistant features: for example, extracting
-certified implementations (\cite{Castro-Perez2021}) or exploiting
+certified implementations (as in \cite{Castro-Perez2021}) or exploiting
 proof automation, such as the one offered by the \emph{Hammer}
-protocol.
-% a more specific
-% challenge could be designed with the objective of automating aspects
-% of the proofs.
+approach.\todo{citation here?}
 
 Ultimately, the current challenges can be extended in several worthwhile directions, and we look forward to a future when they indeed are.
-% We see the current challenge as setting the foundation for those future extensions.
 It is our hope and aim that our challenges will move us closer to a future where the key basic proof techniques are as easy to mechanise as they are to write on paper.

### 2024-02-20 short-paper/short-paper.tex
references for hammer, others less revant removed

@@
 Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, \ie, induction principles on functions to work with HOAS encodings.
-Some proof assistants such as Abella~\cite{Baelde2014} support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging~\cite{Momigliano2012}.
+Some proof assistants such as Abella  support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging.
 
@@
 in \emph{infinitary form} by interpreting their typing (and proof) rules
-coinductively~\cite{BaeldeDoumaneSaurin16,DerakhshanPfenning22,HornePadovani23}.
+coinductively~\cite{%BaeldeDoumaneSaurin16,
+DerakhshanPfenning22,HornePadovani23}.
 
@@
 proof automation, such as the one offered by the \emph{Hammer}
-approach.\todo{citation here?}
+protocol~\cite{BohmeN10,CzajkaK18}.
 

### 2024-02-20 short-paper/short-paper.tex
Minor fixes in linearity section.

@@
 theorem for a system using session types, combining as usual subject reduction with absence of errors.  This allows us to introduce
-linear reasoning with few definitions.  With this focus, we forego
+linear reasoning with few definitions.  With this focus, we forgo
 channel delegation: input and output receives/sends only values, which
@@
 duality within the type system. The latter enforces
- two invariants, namely no endpoint is used simultaneously by parallel processes and
- the two endpoints of the same session have dual types.
+ two invariants, namely that no endpoint is used simultaneously by parallel processes and
+ that the two endpoints of the same session have dual types.
 The operational semantics is formulated in term of reductions rather
@@
 way.  While more sophisticated notions of well-formedness that
-prevents deadlocks are interesting, the proofs of these properties
+prevent \eg deadlocks are interesting, the proofs of those properties
 would distract from the core linear aspects of the challenge.
@@
 (\eg~\cite{Danielsson12,ChaudhuriLR19}). Alternatively, contexts can
-be implemented as finite maps (\cite{Castro2020}), whose operations
+be implemented as finite maps (as in \cite{Castro2020}), whose operations
 are sensitive to a linear discipline. In all these cases, the effort
@@
 A different approach is to bypass the problem of context splitting by adopting familiar ideas from algorithmic linear type checking: this is known as \emph{typing with leftovers}, as exemplified in~\cite{DBLP:conf/forte/ZalakainD21}.
-Whatever the choice, list-based encodings can be refined to be intrinsically-typed, if the proof assistant does support dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
+Whatever the choice, list-based encodings can be refined to be intrinsically-typed, if the proof assistant supports dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
 
-For something completely different, one could adopt a \emph{sub-structural} meta-logical framework: here the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
-In fact, the linear function space will handle all context operations implicitly, including splitting and substitution.
+A completely different approach is to adopt a \emph{sub-structural} meta-logical framework: here the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
+The linear function space will then handle all context operations implicitly, including splitting and substitution.
 One such framework is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session types in~\cite{Bock2016}).

### 2024-02-20 short-paper/short-paper.tex
Clean up the "going beyond" section a bit.

@@
 efforts.
-% While the three highlighted aspects are not the sole sources of difficulties, they emerge in most mechanisations of concurrency theory.
 
@@
 %   - Our aim is to see a future where ``the basics'' are as trivial in mechanisations as they are on paper
-\section{Going beyond the challenge problems}\label{sec:going-beyond}
-Similarly to POPLMark and its Reloaded counterpart, our challenges do not encompass all issues in the field.
-Our benchmark is extendable in two aspects: one can further amalgamate and broaden the existing challenges, or tackle aspects not covered by the benchmark at all.
-Any of these extensions will still need to handle at least one of the key aspects of our challenges.
-\todo{Frederik: Does splitting this section up into these two categories still make sense?
-We could save some space by not doing so, and it is not obvious to me what exactly should be on each side of the split.}
 
-\paragraph{Extending and combining the current challenges.}
-We have aimed to reduce the overlap of the aspects (linearity, scope
-extrusion, coinduction) that our challenges involve.  Most concurrency
-theory research needs several of these aspects and in fact there is
-anedoctal evidence of the difficulties that such a combination may induce
-in a mechanisation,
-e.g.~\cite{DBLP:conf/forte/ZalakainD21}. Furthermore, one can consider
-other constructs that we have chosen not to cover for simplicity.  We
-suggest three extensions that would be useful for many applications:
-\emph{(1)} adding choice constructs to both type and process levels,
-\emph{(2)} introducing recursion and recursive types for typed
-systems, and \emph{(3)} incorporating channel delegation in calculi
-lacking it.
+% We previously promise to discuss combinations of techniques and applications to more elaborate schemes
+\section{Going Beyond the Challenge Problems}\label{sec:going-beyond}
+Our benchmark challenges do not cover all issues in the field, but focus on the fundamental textbook aspects of linearity, scope extrusion, and coinduction.
+Many mechanisations need to combine techniques to handle several of these aspects, and some may also need to handle aspects that are not covered by our benchmark challenges at all.
 
-% \cite{DBLP:conf/forte/ZalakainD21} notes that the combination of linearity and infinite behaviour is difficult since it requires uniquely determining the resources consumed by a process.
-% \cite{Zalakain2019} notes that mixing replication and linearity seems to be very difficult without using explicit type contexts.
+Some authors have already described the difficulties that combining techniques may induce in a mechanisation, e.g.~\cite{DBLP:conf/forte/ZalakainD21}.
 
-\paragraph{Addressing new calculi and new features.}
-Like the POPLMark and POPLMark Reloaded challenges, ours
-is not meant to be comprehensive: Extending it means proposing new
-challenges covering different features of process calculi, or
-different proof mechanisation features.  Some interesting aspects of
-message-passing calculi include
-multiparty session types~\cite{10.1145/2827695} and
-choreographies~\cite{Carbone2013}, as their meta-theory includes
-aspects -- \eg well-formedness conditions on global types, partiality
-of end-point projection function, \etc -- that we do not address
-here. Other revolve around higher-order calculi~\cite{Hirsch2022},
-conversation types~\cite{DBLP:journals/tcs/CairesV10},
-psi-calculi~\cite{lmcs:696}, or encodings between different
-calculi~\cite{DBLP:journals/iandc/Gorla10,
-  DBLP:conf/ecoop/ScalasDHY17}.
+Features that are in wide use, but not covered by our benchmark, include choice constructs and recursion.
+Some interesting aspects of message-passing calculi include multiparty session types~\cite{10.1145/2827695} and choreographies~\cite{Carbone2013}, as their meta-theory includes aspects -- \eg well-formedness conditions on global types, partiality of end-point projection function, \etc -- that we do not address.
+Other areas of interest include higher-order calculi~\cite{Hirsch2022}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10, DBLP:conf/ecoop/ScalasDHY17}.
 \todo{Should we cite other/different encoding papers? Which ones are the ``standard'' citations?}
 
-Also, one could easily design a challenge that goes beyond our
-coinduction challenge, by exploring different notions of process equivalence.
-%
-Coinduction may also play a relevant role in supporting recursive processes and
-session types. Indeed, it is often the case that session type systems (and
-logics) with recursive session types (and propositions) are naturally expressed
-in \emph{infinitary form} by interpreting their typing (and proof) rules
-coinductively~\cite{%BaeldeDoumaneSaurin16,
-DerakhshanPfenning22,HornePadovani23}.
+For coinduction, one could explore different notions of process equivalence.
+
+Coinduction may also play a relevant role in supporting recursive processes and session types.
+Indeed, it is often the case that session type systems with recursive session types are naturally expressed in \emph{infinitary form} by interpreting their typing rules
+coinductively~\cite{DerakhshanPfenning22,HornePadovani23}.
 
@@
 
-Ultimately, the current challenges can be extended in several worthwhile directions, and we look forward to a future when they indeed are.
-It is our hope and aim that our challenges will move us closer to a future where the key basic proof techniques are as easy to mechanise as they are to write on paper.
+Ultimately, our challenges can be extended in several worthwhile directions, and we look forward to a future where they indeed are.
+It is our hope and aim that our challenges will move us closer to a future where the key basic proof techniques for concurrent calculi are as easy to mechanise as they are to write on paper.
 

### 2024-02-20 short-paper/short-paper.tex
first draft of scope extrusion -- needs some more work, but not today

--- 2024-02-20 short-paper/short-paper.tex
@@
 
-\subsubsection{Linearity.}
+\subsection{Linearity}
 
@@
  that the two endpoints of the same session have dual types.
-The operational semantics is formulated in term of reductions rather
-than via a transition system, allowing implementors to explore
-alternative ways to encode structural congruence\footnote{\alb{Should we mention this in challenge 2?}}.
-
-The statement of subject reduction is the usual one, encompassing type
+The operational semantics is formulated as a standard reduction relation to factor the subject reduction proof,  encompassing type
 preservation for structural congruence. W.r.t.\ type safety, we have
@@
 
-\subsubsection{Scope extrusion.}
-\alb{This needs some work: we do not describe the challenge at all and the citations are not the one I had in mind}
+\subsection{Scope extrusion}
 
-The encoding of binding constructs has been extensively studied
-w.r.t.\ the $\lambda$-calculi, but process calculi present additional
-challenges.  $\pi$-calculi typically include several different binding
-constructs: inputs bind a received name or value, recursive processes
-bind recursion variables, and restrictions bind names.  The first two
-act similarly to the binders known from $\lambda$-calculi, but
-restrictions may be more challenging due to scope extrusion.
+This challenge revolves around the formalization explicit scope
+extrusion, by which a process can send restricted names to another
+process, as long as the restriction can safely be ``extruded'' (\ie,
+expanded) to include the receiving process.  
+The setting for this challenge is a ``classic'' untyped
+\( \pi \)-calculus, where (unlike the calculi in the other challenges)
+names can be sent and received, and bound by input constructs.  We define two
+different semantics for our system:
+\begin{enumerate}
+\item A reduction system: this avoids explicit reasoning
+about scope extrusion via structural congruence, allowing implementors to explore
+alternative ways to encode the latter (\eg via context or via congruent refinements)
+\item A (early) labelled transition system
+\end{enumerate}
+The objective of this
+challenge is to prove that the two semantics are equivalent up to
+structural congruence.
 
-\cite{Maksimovic2015} does not have name restriction, thus dodging the challenge of scope extrusion.
-\cite{Castro-Perez2021,Castro2020} note that popular presentations of session type systems are impossible to directly encode using intrinsically \(\alpha\)-convertible terms.
-\cite{Castro2020} suggests using several binder constructs to solve this, but notes that this requires proving many variants of the same theorems.
+This is the challenge most closely related to the POPLMark challenge
+since it concerns the properties of binders, whose encoding has been
+extensively studied w.r.t.\ the $\lambda$-calculi. Process calculi
+present additional challenges: $\pi$-calculi typically include several
+different binding constructs: inputs bind a received name or value,
+recursive processes bind recursion variables, and restrictions bind
+names.  The first two act similarly to the binders known from
+$\lambda$-calculi, but restrictions may be more challenging due to
+scope extrusion.
 
-\alb{We need more details about what the challenge is}
-This is the challenge most closely related to the POPLMark challenge
-since it concerns the properties of binders in restrictions and input.
+Given those peculiarities, the syntax and semantics has been
+mechanized from a an early age (see~\cite{Melham94}) with many proof
+assistants and so many encoding styles that it would take too long to
+review.  They range from concrete encodings based on named syntax
+(Melham's op.cit.) to basic de Bruijn (cite) and locally nameless
+representation~\cite{Chargueraud2012} (\eg~\cite{Castro2020}).
+Nominal approaches are also common, see~\cite{Bengtson2009}, but they
+may be problematic w.r.t.~proof assistants based on dependent types.
 
-The case study \cite{AmbalLS21} compares four approaches to encoding binders in Coq as a first step towards developing tools for working with higher-order process calculi.
-They found that working directly with de Bruijn indices was easiest and shortest since the existing approaches developed for $\lambda$-calculus binders do not work well with scope extrusion.
-Specifically, the locally nameless representation~\cite{Chargueraud2012} cannot avoid direct manipulation of de Bruijn indices when defining the semantics of scope extrusion; cofinite quantification provides no benefits when working under binders; nominal approaches~\cite{Pitts2003} require explicitly giving and validating sets of free names during scope extrusion; and higher-order abstract syntax (HOAS)~\cite{Pfenning1988} requires additional axioms to work with contexts.
+\todo{Sum up some of those issues}
+% \cite{Maksimovic2015} does not have
+% name restriction, thus dodging the challenge of scope extrusion.
+\begin{quotation}
+\cite{Castro-Perez2021,Castro2020} note that popular presentations of
+session type systems are impossible to directly encode using
+intrinsically \(\alpha\)-convertible terms.  \cite{Castro2020}
+suggests using several binder constructs to solve this, but notes that
+this requires proving many variants of the same theorems.
+The case study \cite{AmbalLS21} compares four approaches to encoding
+binders in Coq as a first step towards developing tools for working
+with higher-order process calculi.  They found that working directly
+with de Bruijn indices was easiest and shortest since the existing
+approaches developed for $\lambda$-calculus binders do not work well
+with scope extrusion.  Specifically, the locally nameless
+representation~\cite{Chargueraud2012} cannot avoid direct manipulation
+of de Bruijn indices when defining the semantics of scope extrusion;
+cofinite quantification provides no benefits when working under
+binders; nominal approaches require explicitly giving
+and validating sets of free names during scope extrusion;
+\end{quotation}
 
-Additionally, many proof assistants lack support for the proof principles required to easily work with scope extrusion, \ie, induction principles on functions to work with HOAS encodings.
-Some proof assistants such as Abella  support HOAS natively, but working with open terms as required to encode scope extrusion in these systems is generally very challenging.
+Higher-order abstract syntax (HOAS) has seen extensive and successful
+use in formal reasoning~\cite{Despeyroux2000,Honsell2001,Tiu2010}. Its
+weak form aligns well with mainstream inductive proof assistants,
+simplifying the encoding of typing systems and operational
+semantics. However, when addressing more intricate concepts like
+bisimulation, HOAS necessitates extensions. These extensions may take
+the form of additional axioms~\cite{Honsell2001} or require specialized proof assistants
+like Abella, which introduces a dedicated quantifier for handling
+properties related to names~\cite{GacekMN11}.
 
-\subsubsection{Coinduction.}
+% Higher-order abstract syntax (HOAS) has been extensively and
+% successfully used~\cite{Despeyroux2000,Honsell2001,Tiu2010}, since in
+% its weak form it is compatible with mainstream inductive proof
+% assistants. While this approach disposes of most of the technicalities
+% when encoding typing systems or operational semantics, it requires
+% extensions when dealing with notions such as bisimulation, either in the
+% form of additional axioms~\cite{Honsell2001}, or of dedicated proofs
+% assistants such as Abella, which features a specific
+% quantifier~\cite{GacekMN11} to handle properties of names.
+
+\subsection{Coinduction}
 Coinduction is fundamental to establish properties of processes that

### 2024-02-20 short-paper/short-paper.tex
Add comments from Alceste.
Co-authored-by: Alceste Scalas <alcsc@dtu.dk>

@@
 \section{Introduction}
-
 The POPLMark challenge~\cite{POPLMark}
@@
 programming languages.
-We particularly focus on process calculi.
+We particularly focus on process calculi.\todo{Why? - they are PL-like but also formal models in themselves}
 Concurrent calculi are notably subtle: For instance, it took nine
@@
 % They had to limit themselves to an asymmetric structural precongruence to be able to prove subject reduction.
+
+\todo{Add an informal description of why we care about linearity here}
 
@@
 Some interesting aspects of message-passing calculi include multiparty session types~\cite{10.1145/2827695} and choreographies~\cite{Carbone2013}, as their meta-theory includes aspects -- \eg well-formedness conditions on global types, partiality of end-point projection function, \etc -- that we do not address.
-Other areas of interest include higher-order calculi~\cite{Hirsch2022}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10, DBLP:conf/ecoop/ScalasDHY17}.
-\todo{Should we cite other/different encoding papers? Which ones are the ``standard'' citations?}
+Other areas of interest include higher-order calculi~\cite{Hirsch2022}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:journals/iandc/DardhaGS17}.
 

### 2024-02-20 short-paper/short-paper.tex
Fix some minor issues and remove some comments.

@@
 \section{The Benchmark and its Design}\label{sec:design-discussion}
-\alb{Next paragraph needs rewriting}
 In this section, we outline the factors considered in designing the benchmark challenges.
@@
 \end{enumerate}
-%
-\todo{Frederik: I'm not sure this note is necessary}
-Note that the challenges revolve around the mechanisation of the
-\emph{meta-theory} of process calculi, rather than interpreters or
-type checkers for the latter.
 
@@
 \subsection{Scope extrusion}
-
-This challenge revolves around the formalization explicit scope
+This challenge revolves around the mechanisation of explicit scope
 extrusion, by which a process can send restricted names to another
@@
 Given those peculiarities, the syntax and semantics has been
-mechanized from a an early age (see~\cite{Melham94}) with many proof
+mechanised from an early age (see~\cite{Melham1994}) with many proof
 assistants and so many encoding styles that it would take too long to
@@
 bisimulation, HOAS necessitates extensions. These extensions may take
-the form of additional axioms~\cite{Honsell2001} or require specialized proof assistants
+the form of additional axioms~\cite{Honsell2001} or require specialised proof assistants
 like Abella, which introduces a dedicated quantifier for handling

### 2024-02-20 short-paper/short-paper.tex
Revise section on linearity

--- 2024-02-20 short-paper/short-paper.tex
@@
 
-\todo{Add an informal description of why we care about linearity here}
+Linear typing systems allow for tracking the use of resources in a program. In
+the case of typed process calculi (in particular, session types), linearity is
+widely used for tracking if and how a channel is used to send or receive values.
+This leads to substructural typing~\cite[Chapter 1]{Pierce2004-oq} with
+resulting mechanisation difficulties --- \eg deciding how to \emph{split the
+typing context} in a parallel composition.
 
-For our challenge on linear reasoning we have chosen a type safety
-theorem for a system using session types, combining as usual subject reduction with absence of errors.  This allows us to introduce
-linear reasoning with few definitions.  With this focus, we forgo
-channel delegation: input and output receives/sends only values, which
-do not include channels, hence the topology of the communication
-network described by a process cannot change.  We do not allow
-recursion or replication in the syntax, hence no infinite behaviours
-can be expressed. In particular, we only model linear (as opposed to
-shared) channels. Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
-define a syntax where restriction binds two names, highlighting their
-duality within the type system. The latter enforces
- two invariants, namely that no endpoint is used simultaneously by parallel processes and
- that the two endpoints of the same session have dual types.
-The operational semantics is formulated as a standard reduction relation to factor the subject reduction proof,  encompassing type
-preservation for structural congruence. W.r.t.\ type safety, we have
-chosen a simple notion of what an error is: well-typed processes are a
-well-formed in the sense that they do not use endpoints in a non-dual
-way.  While more sophisticated notions of well-formedness that
-prevent \eg deadlocks are interesting, the proofs of those properties
-would distract from the core linear aspects of the challenge.
+The goal of our challenge on linear reasoning is proving a type safety theorem
+for a process calculus with session types, by combining as usual subject
+reduction with absence of errors.  We have chosen to only model linear (as
+opposed to shared) channels. Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
+define a syntax where a restriction $(\nu ab)\ldots$ binds two dual names $a$
+and $b$ as opposite endpoints of a same channel --- and their duality is
+reflected in the type system. We model a simple notion of error: a well-typed
+processes must never use dual channel endpoints in a non-dual way (\eg by
+performing concurrent send/receive operations on a same endpoint, or two
+concurrent send operations on both endpoints).
+% The latter enforces two invariants, namely that no endpoint is
+% used simultaneously by parallel processes and that the two endpoints of the same
+% session have dual types.
+The operational semantics is formulated as a standard reduction relation; hence,
+proving subject reduction requires proving type preservation for structural
+congruence.
 
-In mechanised meta-theory, addressing linearity means choosing an
-appropriate representation of the linear context.  The crux of the
-matter is \emph{context splitting}, as showcased by the typing rule
-for parallel composition.  While a linear context is perhaps best seen
-as a multi-set, most proof assistants have better support for list
+We designed this challenge to focus on linear reasoning while minimising
+definitions and other concerns.  Therefore, we forgo channel delegation:
+send/receive operations only support values which do not include channels, hence
+the topology of the communication network described by a process cannot change.
+We do not allow recursion or replication, hence no infinite behaviours can be
+expressed. We also forgo more sophisticated notions of error-freedom (\eg
+deadlock freedom) because proving them would distract from the core linear
+aspects of the challenge.
+
+In mechanised meta-theory, addressing linearity means choosing an appropriate
+representation of the linear typing context.  While a linear context is perhaps
+best seen as a multi-set, most proof assistants have better support for list
 processing.  The latter representation is intuitive, but may require
-establishing a large number of technical lemmata that are orthogonal
-to the mathematics of the problem under study, in our case proving
-type safety for session types.  In this space, several designs are
-possible: one can label occurrences of resources to constrain their
-usage (\eg~\cite{CicconeP20}); some authors instead impose a multiset
-structure over lists
-(\eg~\cite{Danielsson12,ChaudhuriLR19}). Alternatively, contexts can
-be implemented as finite maps (as in \cite{Castro2020}), whose operations
-are sensitive to a linear discipline. In all these cases, the effort
-required to develop the infrastructure is significant.
-
-A different approach is to bypass the problem of context splitting by adopting familiar ideas from algorithmic linear type checking: this is known as \emph{typing with leftovers}, as exemplified in~\cite{DBLP:conf/forte/ZalakainD21}.
+establishing a large number of technical lemmata that are orthogonal to the
+problem under study (in our case, proving type safety for
+session types).  In this space, several designs are possible: one can label
+occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}); some
+authors instead impose a multiset structure over lists
+(\eg~\cite{Danielsson12,ChaudhuriLR19}). Alternatively, contexts can be
+implemented as finite maps (as in \cite{Castro2020}), whose operations are
+sensitive to a linear discipline. In all these cases, the effort required to
+develop the infrastructure is significant.
+%
+A different approach is to bypass the problem of context splitting by adopting familiar ideas from algorithmic linear type checking: this is known as \emph{typing with leftovers}, as shown in~\cite{DBLP:conf/forte/ZalakainD21}.
 Whatever the choice, list-based encodings can be refined to be intrinsically-typed, if the proof assistant supports dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
-
-A completely different approach is to adopt a \emph{sub-structural} meta-logical framework: here the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
+%
+A completely different approach is to adopt a \emph{substructural} meta-logical framework: the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
 The linear function space will then handle all context operations implicitly, including splitting and substitution.
-One such framework is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session types in~\cite{Bock2016}).
-Unfortunately, \emph{Celf} does not yet fully support the verification of meta-theoretic properties.
-A compromise is the so called \emph{two-level} approach, where one encodes a sub-structural specification logic in a mainstream proof assistant and then uses that logic to state and prove linear properties (for a recent example, see~\cite{Felty:MSCS21}).
+One such framework is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session types in~\cite{Bock2016});
+unfortunately, \emph{Celf} does not yet fully support the verification of meta-theoretic properties.
+A compromise is the so called \emph{two-level} approach, \ie encoding a substructural specification logic in a mainstream proof assistant and then using that logic to state and prove linear properties (for a recent example, see~\cite{Felty:MSCS21}).
 

### 2024-02-20 short-paper/short-paper.tex
minor rewrites to linearity

@@
 
-Linear typing systems allow for tracking the use of resources in a program. In
+Linear typing systems allow  tracking the use of resources in a program. In
 the case of typed process calculi (in particular, session types), linearity is
-widely used for tracking if and how a channel is used to send or receive values.
+widely used for checking if and how a channel is used to send or receive values.
 This leads to substructural typing~\cite[Chapter 1]{Pierce2004-oq} with
@@
 for a process calculus with session types, by combining as usual subject
-reduction with absence of errors.  We have chosen to only model linear (as
+reduction with absence of errors.  We have chosen  only to model linear (as
 opposed to shared) channels. Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
@@
 %
-A different approach is to bypass the problem of context splitting by adopting familiar ideas from algorithmic linear type checking: this is known as \emph{typing with leftovers}, as shown in~\cite{DBLP:conf/forte/ZalakainD21}.
-Whatever the choice, list-based encodings can be refined to be intrinsically-typed, if the proof assistant supports dependent types (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
+One alternative strategy is to bypass the problem of context splitting
+by adopting familiar ideas from algorithmic linear type checking. This
+approach, referred to as ``typing with leftovers'', is exemplified
+in~\cite{DBLP:conf/forte/ZalakainD21}.  Whatever the choice,
+list-based encodings can be refined to be intrinsically-typed, if the
+proof assistant supports dependent types
+(see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
 %
-A completely different approach is to adopt a \emph{substructural} meta-logical framework: the framework itself handles resource distribution, and users need only map their linear operations to the ones offered by the framework.
-The linear function space will then handle all context operations implicitly, including splitting and substitution.
-One such framework is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session types in~\cite{Bock2016});
-unfortunately, \emph{Celf} does not yet fully support the verification of meta-theoretic properties.
-A compromise is the so called \emph{two-level} approach, \ie encoding a substructural specification logic in a mainstream proof assistant and then using that logic to state and prove linear properties (for a recent example, see~\cite{Felty:MSCS21}).
+A radically different approach is to adopt a \emph{substructural}
+meta-logical framework: the framework itself handles resource
+distribution, and users need only map their linear operations to the
+ones offered by the framework.  The linear function space will then
+handle all context operations implicitly, including splitting and
+substitution.  The only such framework is
+\emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session
+types in~\cite{Bock2016}); unfortunately, \emph{Celf} does not yet
+fully support the verification of meta-theoretic properties.  A
+compromise is the so called \emph{two-level} approach, \ie encoding a
+substructural specification logic in a mainstream proof assistant and
+then using that logic to state and prove linear properties (for a
+recent example, see~\cite{Felty:MSCS21}).
 
@@
 extrusion, by which a process can send restricted names to another
-process, as long as the restriction can safely be ``extruded'' (\ie,
-expanded) to include the receiving process.  
-The setting for this challenge is a ``classic'' untyped
-\( \pi \)-calculus, where (unlike the calculi in the other challenges)
-names can be sent and received, and bound by input constructs.  We define two
-different semantics for our system:
+process, as long as the restriction can safely be extruded to include
+the receiving process.  The setting for this challenge is a
+``classic'' untyped \( \pi \)-calculus, where (unlike the calculi in
+the other challenges) names can be sent and received, and bound by
+input constructs.  We define two different semantics for our system:
 \begin{enumerate}

### 2024-02-21 short-paper/short-paper.tex
two small corrections to intro

@@
 We particularly focus on process calculi.\todo{Why? - they are PL-like but also formal models in themselves}
-Concurrent calculi are notably subtle: For instance, it took nine
+Concurrent calculi are notably subtle: for instance, it took nine
 years before the incorrect proof of type safety in the seminal paper
@@
 mechanised proofs closely resemble the informal ones, while the
-main challenge lies in getting the foundations right.
+main challenge lies in getting the infra-structure right.
 
@@
   extrusion}, and \emph{coinductive reasoning}.  Mechanisations must
-often address several of these aspects; however, we
+often address several of these aspects at the same time; however, we
 see the combination of techniques as a next

### 2024-02-21 short-paper/short-paper.tex
shortening the paragraph about Celf

@@
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
-\item\label{item:rq1} What is the current state of the art in tools for mechanising the meta-theory of process calculi?
+\item\label{item:rq1} What is the current state of the art in the mechanisation of  the meta-theory of process calculi?
 \item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
@@
 representation of the linear typing context.  While a linear context is perhaps
-best seen as a multi-set, most proof assistants have better support for list
+best seen as a multiset, most proof assistants have better support for list
 processing.  The latter representation is intuitive, but may require
@@
 A radically different approach is to adopt a \emph{substructural}
-meta-logical framework: the framework itself handles resource
-distribution, and users need only map their linear operations to the
-ones offered by the framework.  The linear function space will then
-handle all context operations implicitly, including splitting and
-substitution.  The only such framework is
+meta-logical framework, which handles resource
+distribution implicitly, including splitting and
+substitution: users need only map their linear operations to the
+ones offered by the framework.  The only such framework is
 \emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session

### 2024-02-21 short-paper/short-paper.tex
revisions to scope extr: still the "quotation" stuff to be fixed

@@
 \subsection{Scope extrusion}
-This challenge revolves around the mechanisation of explicit scope
+This challenge revolves around the mechanisation of % explicit
+scope
 extrusion, by which a process can send restricted names to another
@@
 \begin{enumerate}
-\item A reduction system: this avoids explicit reasoning
-about scope extrusion via structural congruence, allowing implementors to explore
-alternative ways to encode the latter (\eg via context or via congruent refinements)
-\item A (early) labelled transition system
+\item A reduction system: this avoids explicit reasoning about scope
+  extrusion by using instead structural congruence, allowing
+  implementors to explore alternative ways to encode the latter (\eg
+  via process contexts or via compatible refinement);
+\item A (early) labelled transition system.
 \end{enumerate}
@@
 
-This is the challenge most closely related to the POPLMark challenge
+This is the challenge most closely related to  POPLMark, 
 since it concerns the properties of binders, whose encoding has been
 extensively studied w.r.t.\ the $\lambda$-calculi. Process calculi
-present additional challenges: $\pi$-calculi typically include several
+present additional challenges,  typically including several
 different binding constructs: inputs bind a received name or value,
@@
 
-Given those peculiarities, the syntax and semantics has been
+Given those peculiarities, the syntax and semantics of $\pi$-calculi have been
 mechanised from an early age (see~\cite{Melham1994}) with many proof
-assistants and so many encoding styles that it would take too long to
+assistants and in so many encoding styles that it would take too long to
 review.  They range from concrete encodings based on named syntax
 (Melham's op.cit.) to basic de Bruijn (cite) and locally nameless
-representation~\cite{Chargueraud2012} (\eg~\cite{Castro2020}).
+representation%~\cite{Chargueraud2012}
+(\eg~\cite{Castro2020}).
 Nominal approaches are also common, see~\cite{Bengtson2009}, but they
@@
 Higher-order abstract syntax (HOAS) has seen extensive and successful
-use in formal reasoning~\cite{Despeyroux2000,Honsell2001,Tiu2010}. Its
+use in formal reasoning in this area~\cite{Despeyroux2000,Honsell2001,Tiu2010}. Its
 weak form aligns well with mainstream inductive proof assistants,
-simplifying the encoding of typing systems and operational
+significantly simplifying the encoding of typing systems and operational
 semantics. However, when addressing more intricate concepts like
@@
 
-\subsubsection{Evaluation criteria.}
+\subsection{Evaluation criteria.}
 The idea behind our benchmark is to obtain evidence towards answering

### 2024-02-21 short-paper/short-paper.tex
Minor edits.

@@
 mechanised proofs closely resemble the informal ones, while the
-main challenge lies in getting the infra-structure right.
+main challenge lies in getting the infrastructure right.
 
@@
 \subsection{Linearity}
-
-% \cite{Brady2017} notes that using algebraic effects is unwieldy when there are many communication channels and suggests an experimental approach using dependent uniqueness types.
-% \cite{Zalakain2019} notes that systems lacking linear types must simulate linearity, and that using a linearity predicate complicates congruences.
-% They had to limit themselves to an asymmetric structural precongruence to be able to prove subject reduction.
-
 Linear typing systems allow  tracking the use of resources in a program. In
@@
 concurrent send operations on both endpoints).
-% The latter enforces two invariants, namely that no endpoint is
-% used simultaneously by parallel processes and that the two endpoints of the same
-% session have dual types.
 The operational semantics is formulated as a standard reduction relation; hence,
@@
 review.  They range from concrete encodings based on named syntax
-(Melham's op.cit.) to basic de Bruijn (cite) and locally nameless
+\cite{Melham1994} to basic de Bruijn \todo{cite} and locally nameless
 representation%~\cite{Chargueraud2012}
@@
 
+% \cite{Brady2017} notes that using algebraic effects is unwieldy when there are many communication channels and suggests an experimental approach using dependent uniqueness types.
+% \cite{Zalakain2019} notes that systems lacking linear types must simulate linearity, and that using a linearity predicate complicates congruences.
+% They had to limit themselves to an asymmetric structural precongruence to be able to prove subject reduction.
 Some authors have already described the difficulties that combining techniques may induce in a mechanisation, e.g.~\cite{DBLP:conf/forte/ZalakainD21}.

### 2024-02-21 short-paper/short-paper.tex
Clean up the beginning of the coinduction section.

@@
 \subsection{Coinduction}
-Coinduction is fundamental to establish properties of processes that
-may never terminate, which is a typical situation.
-Common uses for coinduction are to define strong and weak versions of
-bisimulation, barbed notions of bisimulation, and prove trace equivalence.
-On the other hand, coinduction in proof assistants
-is notoriously difficult and its representation
-is non-uniform across developments.
+Process calculi often include constructs that allow processes to have infinite behaviours.
+Coinduction is a fundamental technique for defining and reasoning about infinite objects, and it can thus be used to reason about these behaviours.
+While many proof assistants support coinductive techniques, using them is notoriously difficult, and representations and proof techniques vary widely.
+The crux of our challenge is the effective use of coinductive up-to techniques.
+The intention is that it should be easy to use the bisimulation relation and its properties, but this may be difficult with some mechanisation approaches.
+
+The goal of our challenge on coinductive reasoning is proving that strong barbed bisimilarity can be turned into a congruence by making it sensitive to substitution and parallel composition.
+The setting for our challenge is an untyped \(\pi\)-calculus with replication of processes enabling infinite behaviours.
+We do not include delegation in the calculus since it is orthogonal to our primary aim of exploring coinductive proof techniques.
+We base our definition of bisimilarity on a labelled transition system semantics and an observability predicate describing the communication steps available to a process.
+The introduction of strong barbed bisimulation is one of the first steps when studying the observable behaviour of process calculi, both in classical textbooks (e.g., \cite{picalcbook}) and in existing mechanisations.
+Though weak barbed congruence is a more common behavioural equivalence, we prefer strong equivalences to simplify the theory by avoiding the need to abstract over the number of internal transitions in a trace.
+
+Coinduction is widely supported in proof assistants, but they support subtly different formalisms that enable different techniques, and some systems provide more than one abstraction to use coinduction.
+We thus see a variety of different formalisation choices for bisimulations, sometimes ad-hoc for the specific result.
+
+\todo{Reorganize the rest of this section to be more coherent}
+Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
+Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
+Newer entrants to the field like the Lean proof assistant
+are working on adding a notion of coinduction to their system~\cite{Avigad2019}.\alb{I kinda object to ``seamlessly''.
+In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky.
+I guess Isabelle is better}
+%
+A good example of this are the parameterised coinduction~\cite{Hur2013}
+and interaction tree~\cite{Xia2019} libraries for Coq.
+
 E.g., \cite{Castro-Perez2021} notes that the most challenging
@@
 
-We base our challenge around bisimulation: a fundamental coinductively
-established property for equating the
-observed behaviour of potentially infinite processes. 
-In this challenge, we study the notion of strong barbed bisimulation
-(and bisimilarity, as in \cite{picalcbook})
-over a fragment of the $\pi$-calculus and,
-in particular, we focus on obtaining,
-from strong barbed bisimilarity,
-a congruence with respect to
-parallel composition and substitution.
-
-\alb{We need more details about what the challenge is}
-\alb{I kinda object to ``seamlessly''. In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky. I guess Isabelle is better}
-
-Our coinduction challenge concerns strong barbed bisimulation and
-congruence.% \alb{we motivate why strong vs weak, but not why barbed}
-% \lore{Is it not enough to quote the Sangiorgi/Walker book
-% and go by authority? :)}
-Though weak barbed congruence is a more common behavioural equivalence
-w.r.t.\ the \(\pi\)-calculus, here we prefer strong equivalences to avoid the
-need to abstract over the number of internal transitions, thus
-simplifying the theory.  We exclude delegation from the coinduction
-challenge since it is orthogonal to our primary aim of exploring
-coinductive proof techniques.
-
-The introduction of strong barbed bisimulation is one of the first
-steps when studying the (observable) behaviour of process calculi,
-both in classical textbooks (e.g., \cite{picalcbook}) and in existing
-mechanisations, where we see a variety of different formalisation
-choices, sometimes ad-hoc for the specific result or development.
-
-In fact, coinduction is widely supported in proof assistants.
-Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
-Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
-Newer entrants to the field like the Lean proof assistant
-are working on adding a notion of coinduction to their system~\cite{Avigad2019}%,Keizer2023}.
-%
-As with most features, proof assistants support
-subtly different formalisms that enable different techniques and motivate this challenge.
-Some systems provide more than one abstraction to use coinduction.
-A good example of this are the parameterised coinduction~\cite{Hur2013}
-and interaction tree~\cite{Xia2019} libraries for Coq.
-
 \lore{Should we mention that Isabelle offers
   the command \texttt{coinduction\_upto}? Is it related?}
-\todo{Explain the the point of the congruence theorem is that it should be ``easy'' once you have the other one}
 

### 2024-02-21 short-paper/short-paper.tex
Add a comment for clarity.

@@
 \section{Going Beyond the Challenge Problems}\label{sec:going-beyond}
+\todo{Clean up this section to be more coherent.}
 Our benchmark challenges do not cover all issues in the field, but focus on the fundamental textbook aspects of linearity, scope extrusion, and coinduction.

### 2024-02-21 short-paper/short-paper.tex
Minor edits

@@
      IT University of Copenhagen, Copenhagen, Denmark \email{maca@itu.dk}, \email{dati@itu.dk}
-\and University of Kent, Canterbury, United Kingdom \email{D.Castro-Perez@kent.ac.uk}
-\and Royal Holloway, University of London, Egham, United Kingdom \email{Francisco.FerreiraRuiz@rhul.ac.uk}
-\and University of Liverpool, Liverpool, United Kingdom \email{Lorenzo.Gheri@liverpool.ac.uk}
+\and University of Kent, Canterbury, United Kingdom \email{d.castro-perez@kent.ac.uk}
+\and Royal Holloway, University of London, Egham, United Kingdom \email{francisco.ferreiraruiz@rhul.ac.uk}
+\and University of Liverpool, Liverpool, United Kingdom \email{lorenzo.gheri@liverpool.ac.uk}
 \and Technical University of Denmark, Kgs. Lyngby, Denmark \email{fkjac@dtu.dk}, \email{alcsc@dtu.dk}
@@
   The goal of this new challenge is to clarify, compare, and advance the state of the art, fostering the adoption of proof mechanisation in future research on message-passing concurrency.
+\todo{Broaden `message-passing concurrency' to `concurrency'?}
 
@@
 years before the incorrect proof of type safety in the seminal paper
-by Honda et al.~\cite{Honda1998} was rectified in Yoshida \&
+by Honda et al.~\cite{Honda1998} was rectified in Yoshida and
 Vasconcelos’s subsequent work\cite{Yoshida2007}.
@@
 In the future, we expect to use the website to promote best practices and tutorials derived from solutions to our challenges.
-We encourage everyone to try the challenges using their favourite techniques, and to send us their solutions.
+We encourage readers to try the challenges using their favourite techniques, and to send us their solutions.
 
@@
 opposed to shared) channels. Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
-define a syntax where a restriction $(\nu ab)\ldots$ binds two dual names $a$
-and $b$ as opposite endpoints of a same channel --- and their duality is
+define a syntax where a restriction $(\nu ab)$ binds two dual names $a$
+and $b$ as opposite endpoints of the same channel; their duality is
 reflected in the type system. We model a simple notion of error: a well-typed
 processes must never use dual channel endpoints in a non-dual way (\eg by
-performing concurrent send/receive operations on a same endpoint, or two
+performing concurrent send/receive operations on the same endpoint, or two
 concurrent send operations on both endpoints).
-The operational semantics is formulated as a standard reduction relation; hence,
+The operational semantics is formulated as a standard reduction relation. Thus,
 proving subject reduction requires proving type preservation for structural
@@
 We designed this challenge to focus on linear reasoning while minimising
-definitions and other concerns.  Therefore, we forgo channel delegation:
-send/receive operations only support values which do not include channels, hence
+definitions and other concerns.  We therefore forgo channel delegation:
+send/receive operations only support values that do not include channels, so
 the topology of the communication network described by a process cannot change.
-We do not allow recursion or replication, hence no infinite behaviours can be
+We do not allow recursion or replication, so infinite behaviours cannot be
 expressed. We also forgo more sophisticated notions of error-freedom (\eg
@@
 by adopting familiar ideas from algorithmic linear type checking. This
-approach, referred to as ``typing with leftovers'', is exemplified
+approach, referred to as ``typing with leftovers,'' is exemplified
 in~\cite{DBLP:conf/forte/ZalakainD21}.  Whatever the choice,
@@
   via process contexts or via compatible refinement);
-\item A (early) labelled transition system.
+\item An (early) labelled transition system.
 \end{enumerate}
@@
 since it concerns the properties of binders, whose encoding has been
-extensively studied w.r.t.\ the $\lambda$-calculi. Process calculi
+extensively studied with respect to the $\lambda$-calculi. Process calculi
 present additional challenges,  typically including several
@@
 \cite{Melham1994} to basic de Bruijn \todo{cite} and locally nameless
-representation%~\cite{Chargueraud2012}
+representation %~\cite{Chargueraud2012}
 (\eg~\cite{Castro2020}).
-Nominal approaches are also common, see~\cite{Bengtson2009}, but they
-may be problematic w.r.t.~proof assistants based on dependent types.
+Nominal approaches are also common (see~\cite{Bengtson2009}), but they
+may be problematic in proof assistants based on dependent types.
 
@@
 We base our definition of bisimilarity on a labelled transition system semantics and an observability predicate describing the communication steps available to a process.
-The introduction of strong barbed bisimulation is one of the first steps when studying the observable behaviour of process calculi, both in classical textbooks (e.g., \cite{picalcbook}) and in existing mechanisations.
+The introduction of strong barbed bisimulation is one of the first steps when studying the observable behaviour of process calculi, both in classical textbooks (\eg, \cite{picalcbook}) and in existing mechanisations.
 Though weak barbed congruence is a more common behavioural equivalence, we prefer strong equivalences to simplify the theory by avoiding the need to abstract over the number of internal transitions in a trace.
@@
 % They had to limit themselves to an asymmetric structural precongruence to be able to prove subject reduction.
-Some authors have already described the difficulties that combining techniques may induce in a mechanisation, e.g.~\cite{DBLP:conf/forte/ZalakainD21}.
+Some authors have already described the difficulties that combining techniques may induce in a mechanisation, \eg~\cite{DBLP:conf/forte/ZalakainD21}.
 
 Features that are in wide use, but not covered by our benchmark, include choice constructs and recursion.
-Some interesting aspects of message-passing calculi include multiparty session types~\cite{10.1145/2827695} and choreographies~\cite{Carbone2013}, as their meta-theory includes aspects -- \eg well-formedness conditions on global types, partiality of end-point projection function, \etc -- that we do not address.
+Some interesting aspects of message-passing calculi include multiparty session types~\cite{10.1145/2827695} and choreographies~\cite{Carbone2013}, as their meta-theory includes aspects---\eg well-formedness conditions on global types, partiality of end-point projection function, \etc---that we do not address.
 Other areas of interest include higher-order calculi~\cite{Hirsch2022}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:journals/iandc/DardhaGS17}.

### 2024-02-21 short-paper/short-paper.tex
cleanup of coinduction

--- 2024-02-21 short-paper/short-paper.tex
@@
 Higher-order abstract syntax (HOAS) has seen extensive and successful
-use in formal reasoning in this area~\cite{Despeyroux2000,Honsell2001,Tiu2010}. Its
+use in formal reasoning in this area~\cite{Despeyroux2000,Honsell2001,Tiu2010,ChaudhuriCM15}. Its
 weak form aligns well with mainstream inductive proof assistants,
@@
 \subsection{Coinduction}
-Process calculi often include constructs that allow processes to have infinite behaviours.
-Coinduction is a fundamental technique for defining and reasoning about infinite objects, and it can thus be used to reason about these behaviours.
-While many proof assistants support coinductive techniques, using them is notoriously difficult, and representations and proof techniques vary widely.
-The crux of our challenge is the effective use of coinductive up-to techniques.
-The intention is that it should be easy to use the bisimulation relation and its properties, but this may be difficult with some mechanisation approaches.
+Process calculi typically include constructs that allow processes to
+have infinite behaviours.  Coinduction serves as a fundamental method
+for the definition and analysis of infinite objects, enabling the
+examination of their behaviours.
+% Coinduction is a fundamental technique for
+% defining and reasoning about infinite objects, and it can thus be used
+% to reason about these behaviours.
 
-The goal of our challenge on coinductive reasoning is proving that strong barbed bisimilarity can be turned into a congruence by making it sensitive to substitution and parallel composition.
-The setting for our challenge is an untyped \(\pi\)-calculus with replication of processes enabling infinite behaviours.
-We do not include delegation in the calculus since it is orthogonal to our primary aim of exploring coinductive proof techniques.
-We base our definition of bisimilarity on a labelled transition system semantics and an observability predicate describing the communication steps available to a process.
-The introduction of strong barbed bisimulation is one of the first steps when studying the observable behaviour of process calculi, both in classical textbooks (e.g., \cite{picalcbook}) and in existing mechanisations.
-Though weak barbed congruence is a more common behavioural equivalence, we prefer strong equivalences to simplify the theory by avoiding the need to abstract over the number of internal transitions in a trace.
+The goal of our challenge on coinductive reasoning is proving that
+\emph{strong barbed bisimilarity} can be turned into a congruence by making
+it sensitive to substitution and parallel composition. The crux of our
+challenge is the effective use of coinductive up-to techniques.  The
+intention is that it should be easy to use the bisimulation relation
+and its properties, but this may be difficult with some mechanisation
+approaches.
 
-Coinduction is widely supported in proof assistants, but they support subtly different formalisms that enable different techniques, and some systems provide more than one abstraction to use coinduction.
-We thus see a variety of different formalisation choices for bisimulations, sometimes ad-hoc for the specific result.
+The setting for our challenge is an untyped \(\pi\)-calculus augmented with
+replication of processes in order to enable infinite behaviours.  We do not
+include delegation since it is orthogonal to our
+primary aim of exploring coinductive proof techniques.  We base our
+definition of bisimilarity on a labelled transition system semantics
+and an observability predicate describing the communication steps
+available to a process.  The introduction of strong barbed
+bisimulation is one of the first steps when studying the observable
+behaviour of process calculi, both in classical textbooks (e.g.,
+\cite{picalcbook}) and in existing mechanisations.  Though weak barbed
+congruence is a more common behavioural equivalence, we prefer strong
+equivalences to simplify the theory by avoiding the need to abstract
+over the number of internal transitions in a trace.
 
-\todo{Reorganize the rest of this section to be more coherent}
-Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
-Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
-Newer entrants to the field like the Lean proof assistant
-are working on adding a notion of coinduction to their system~\cite{Avigad2019}.\alb{I kinda object to ``seamlessly''.
+While many proof assistants support coinductive techniques, they do so
+through subtly different formalisms, enabling various
+approaches. Additionally, some systems offer multiple abstractions for
+utilizing coinduction.
+Examples are Agda's musical notation versus  co-patterns and sized types~\cite{Abel2013}.
+In Coq, guarded recursion versus a refined fixed point approach as supported by the parameterised coinduction~\cite{Hur2013}
+and interaction tree~\cite{Xia2019} libraries.
+
+When reasoning over bisimilarity many authors rely on the native
+ coinduction offered by the chosen proof assistants
+\cite{Bengtson2016,Kahsai2008,Thiemann2019,Gay2020}, others prefer a
+more ``set-theoretic'' approach
+\cite{Hirschkoff1997,Bengtson2009,Maksimovic2015,Pohjola2022}. Some
+use both and establish an internal adequacy~\cite{Honsell2001}. Few
+extend the proof assistant foundations to allow \eg reasoning about
+bisimilarity up-to~\cite{ChaudhuriCM15}.
+% While many proof assistants support
+% coinductive techniques,   but they support subtly different formalisms that enable different techniques, and some systems provide more than one abstraction to use coinduction.
+% We thus see a variety of different formalisation choices for
+% bisimulations, sometimes ad-hoc for the specific result.
+
+% \todo{Reorganize the rest of this section to be more coherent}
+% Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
+% Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
+% Newer entrants to the field like the Lean proof assistant
+% are working on adding a notion of coinduction to their system~\cite{Avigad2019}.\alb{I kinda object to ``seamlessly''.
 In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky.
@@
 %
-A good example of this are the parameterised coinduction~\cite{Hur2013}
-and interaction tree~\cite{Xia2019} libraries for Coq.
+% A good example of this are the parameterised coinduction~\cite{Hur2013}
+% and interaction tree~\cite{Xia2019} libraries for Coq.
 
-E.g., \cite{Castro-Perez2021} notes that the most challenging
-part of their mechanisation of multiparty session types
-was the convoluted definition of infinite behaviours;
-\cite{Pohjola2022} points out that the main difficulty
-of the proofs concerning bisimulation lies
-in dealing with existential quantifiers; and
-\cite{Bengtson2016} wonders whether parametrised
-coinduction as developed in Coq \cite{Hur2013} would
-have helped their Isabelle mechanisation.
+% E.g., \cite{Castro-Perez2021} notes that the most challenging
+% part of their mechanisation of multiparty session types
+% was the convoluted definition of infinite behaviours;
+% \cite{Pohjola2022} points out that the main difficulty
+% of the proofs concerning bisimulation lies
+% in dealing with existential quantifiers; and
+% \cite{Bengtson2016} wonders whether parametrised
+% coinduction as developed in Coq \cite{Hur2013} would
+% have helped their Isabelle mechanisation.
 
-\lore{Should we mention that Isabelle offers
-  the command \texttt{coinduction\_upto}? Is it related?}
+% \lore{Should we mention that Isabelle offers
+%   the command \texttt{coinduction\_upto}? Is it related?}
 
-While many authors rely on the native support for coinduction offered
-by proof assistants
-\cite{Honsell2001,Bengtson2016,Kahsai2008,Thiemann2019,Gay2020}, many
-others prefer a more ``set-theoretic'' approach
-\cite{Hirschkoff1997,Bengtson2009,Maksimovic2015,Pohjola2022}.
-\alb{Actually, Honsell gives up with guarded coinduction and switches to Park's half way}
+% While many authors rely on the native support for coinduction offered
+% by proof assistants
+% \cite{Honsell2001,Bengtson2016,Kahsai2008,Thiemann2019,Gay2020}, many
+% others prefer a more ``set-theoretic'' approach
+% \cite{Hirschkoff1997,Bengtson2009,Maksimovic2015,Pohjola2022}.
+%\alb{Actually, Honsell gives up with guarded coinduction and switches to Park's half way}
 

### 2024-02-21 challenges/linearity.tex
Add reference to Barendregt convention in linearity challenge.

@@
 the comma as split/union, overloaded to singletons, and \( \Cempty \) as the
-empty context. We extend the Barendregt convention \todo{Source?} to contexts, so that all
+empty context. We extend the Barendregt convention~\cite{Barendregt1984} to contexts, so that all
 entries are distinct.  Note that the order in which information is added to a
@@
 
-Now, following Sangiorgi \& Walker~\cite{picalcbook}, we give rules for the compatible
+Now, following Sangiorgi and Walker~\cite{picalcbook}, we give rules for the compatible
 equivalence relation induced by $\sconga\cdot\cdot$, which we still

### 2024-02-21 short-paper/short-paper.tex

@@
 \subsection{Coinduction}
+
 Process calculi typically include constructs that allow processes to

### 2024-02-22 short-paper/short-paper.tex
Minor fixes.
Delete old TeX comments.

@@
 
-% - Introduction
-%   - POPLMark accelerated progress in mechanisation tools for PL
-%     - POPLMark Reloaded tried to do the same for logical relations
-%     - But they were deliberately limited in scope to not include process calculi
-%   - We propose a new collection of benchmark problems for process calculi
-%     - Why not concurrency in general? (after this point, only mention process calculi)
-%     - Why is there a problem with process calculus mechanisations?
-%     - Our experience says there is, and we will cite other experiences below
-%   - We have noticed three key issues
-%     - They are not covered by POPLMark
-%     - Techniques are less settled than for binders
-%     - Others also mention them as problems: cite and explain references briefly for each challenge topic
-%   - Our goals are to determine the state of the art, develop best practices and tutorials, and identify improvements to proof assistant technology that would materially help the situation
 \section{Introduction}
@@
 \subsection{Scope extrusion}
-This challenge revolves around the mechanisation of % explicit
-scope
+This challenge revolves around the mechanisation of scope
 extrusion, by which a process can send restricted names to another
@@
 
-% Higher-order abstract syntax (HOAS) has been extensively and
-% successfully used~\cite{Despeyroux2000,Honsell2001,Tiu2010}, since in
-% its weak form it is compatible with mainstream inductive proof
-% assistants. While this approach disposes of most of the technicalities
-% when encoding typing systems or operational semantics, it requires
-% extensions when dealing with notions such as bisimulation, either in the
-% form of additional axioms~\cite{Honsell2001}, or of dedicated proofs
-% assistants such as Abella, which features a specific
-% quantifier~\cite{GacekMN11} to handle properties of names.
-
 \subsection{Coinduction}
@@
 examination of their behaviours.
-% Coinduction is a fundamental technique for
-% defining and reasoning about infinite objects, and it can thus be used
-% to reason about these behaviours.
 
@@
 approaches. Additionally, some systems offer multiple abstractions for
-utilizing coinduction.
+utilising coinduction.
 Examples are Agda's musical notation versus  co-patterns and sized types~\cite{Abel2013}.
@@
 bisimilarity up-to~\cite{ChaudhuriCM15}.
-% While many proof assistants support
-% coinductive techniques,   but they support subtly different formalisms that enable different techniques, and some systems provide more than one abstraction to use coinduction.
-% We thus see a variety of different formalisation choices for
-% bisimulations, sometimes ad-hoc for the specific result.
-
-% \todo{Reorganize the rest of this section to be more coherent}
-% Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
-% Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
-% Newer entrants to the field like the Lean proof assistant
-% are working on adding a notion of coinduction to their system~\cite{Avigad2019}.\alb{I kinda object to ``seamlessly''.
-In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky.
-I guess Isabelle is better}
-%
-% A good example of this are the parameterised coinduction~\cite{Hur2013}
-% and interaction tree~\cite{Xia2019} libraries for Coq.
-
-% E.g., \cite{Castro-Perez2021} notes that the most challenging
-% part of their mechanisation of multiparty session types
-% was the convoluted definition of infinite behaviours;
-% \cite{Pohjola2022} points out that the main difficulty
-% of the proofs concerning bisimulation lies
-% in dealing with existential quantifiers; and
-% \cite{Bengtson2016} wonders whether parametrised
-% coinduction as developed in Coq \cite{Hur2013} would
-% have helped their Isabelle mechanisation.
-
-% \lore{Should we mention that Isabelle offers
-%   the command \texttt{coinduction\_upto}? Is it related?}
-
-% While many authors rely on the native support for coinduction offered
-% by proof assistants
-% \cite{Honsell2001,Bengtson2016,Kahsai2008,Thiemann2019,Gay2020}, many
-% others prefer a more ``set-theoretic'' approach
-% \cite{Hirschkoff1997,Bengtson2009,Maksimovic2015,Pohjola2022}.
-%\alb{Actually, Honsell gives up with guarded coinduction and switches to Park's half way}
 

### 2024-02-22 short-paper/short-paper.tex
Minor edits in coinduction section.

@@
 While many proof assistants support coinductive techniques, they do so
-through subtly different formalisms, enabling various
-approaches. Additionally, some systems offer multiple abstractions for
-utilising coinduction.
-Examples are Agda's musical notation versus  co-patterns and sized types~\cite{Abel2013}.
-In Coq, guarded recursion versus a refined fixed point approach as supported by the parameterised coinduction~\cite{Hur2013}
+through subtly different formalisms, enabling various approaches.
+Additionally, some systems offer multiple abstractions for utilising coinduction.
+For instance, Agda offers musical notation, co-patterns and sized types~\cite{Abel2013};
+Coq offers guarded recursion and refined fixed point approaches as supported by the parameterised coinduction~\cite{Hur2013}
 and interaction tree~\cite{Xia2019} libraries.
@@
  coinduction offered by the chosen proof assistants
-\cite{Bengtson2016,Kahsai2008,Thiemann2019,Gay2020}, others prefer a
+\cite{Bengtson2016,Kahsai2008,Thiemann2019,Gay2020}, while others prefer a
 more ``set-theoretic'' approach

### 2024-02-22 short-paper/short-paper.tex
merge
Fix some minor comments.

@@
   Our benchmark problems address three key topics: linearity, scope extrusion, and coinductive reasoning.
-  The goal of this new challenge is to clarify, compare, and advance the state of the art, fostering the adoption of proof mechanisation in future research on message-passing concurrency.
-\todo{Broaden `message-passing concurrency' to `concurrency'?}
+  The goal of this new challenge is to clarify, compare, and advance the state of the art, fostering the adoption of proof mechanisation in future research on concurrency.
 
@@
 programming languages.
-We particularly focus on process calculi.\todo{Why? - they are PL-like but also formal models in themselves}
+We particularly focus on process calculi as these provide a simple but
+realistic showcase of common issues encountered during mechanisation
+efforts.
 Concurrent calculi are notably subtle: for instance, it took nine
@@
 step, as discussed in \cref{sec:going-beyond}.
-Our benchmark challenges are based on process calculi, as these provide a simple but
-realistic showcase of common issues encountered during mechanisation
-efforts.
 
@@
 review.  They range from concrete encodings based on named syntax
-\cite{Melham1994} to basic de Bruijn \todo{cite} and locally nameless
+\cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997, Perera2018} and locally nameless
 representation %~\cite{Chargueraud2012}
@@
 bisimilarity up-to~\cite{ChaudhuriCM15}.
+% While many proof assistants support
+% coinductive techniques,   but they support subtly different formalisms that enable different techniques, and some systems provide more than one abstraction to use coinduction.
+% We thus see a variety of different formalisation choices for
+% bisimulations, sometimes ad-hoc for the specific result.
+
+% \todo{Reorganize the rest of this section to be more coherent}
+% Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
+% Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
+% Newer entrants to the field like the Lean proof assistant
+% are working on adding a notion of coinduction to their system~\cite{Avigad2019}.\alb{I kinda object to ``seamlessly''.
+In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky.
+I guess Isabelle is better}
+% =======
+% Process calculi often include constructs that allow processes to have infinite behaviours.
+% Coinduction is a fundamental technique for defining and reasoning about infinite objects, and it can thus be used to reason about these behaviours.
+% While many proof assistants support coinductive techniques, using them is notoriously difficult, and representations and proof techniques vary widely.
+% The crux of our challenge is the effective use of coinductive up-to techniques.
+% The intention is that it should be easy to use the bisimulation relation and its properties, but this may be difficult with some mechanisation approaches.
+
+% The goal of our challenge on coinductive reasoning is proving that strong barbed bisimilarity can be turned into a congruence by making it sensitive to substitution and parallel composition.
+% The setting for our challenge is an untyped \(\pi\)-calculus with replication of processes enabling infinite behaviours.
+% We do not include delegation in the calculus since it is orthogonal to our primary aim of exploring coinductive proof techniques.
+% We base our definition of bisimilarity on a labelled transition system semantics and an observability predicate describing the communication steps available to a process.
+% The introduction of strong barbed bisimulation is one of the first steps when studying the observable behaviour of process calculi, both in classical textbooks (\eg, \cite{picalcbook}) and in existing mechanisations.
+% Though weak barbed congruence is a more common behavioural equivalence, we prefer strong equivalences to simplify the theory by avoiding the need to abstract over the number of internal transitions in a trace.
+
+% Coinduction is widely supported in proof assistants, but they support subtly different formalisms that enable different techniques, and some systems provide more than one abstraction to use coinduction.
+% We thus see a variety of different formalisation choices for bisimulations, sometimes ad-hoc for the specific result.
+
+% \todo{Reorganize the rest of this section to be more coherent}
+% Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
+% Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
+% Newer entrants to the field like the Lean proof assistant
+% are working on adding a notion of coinduction to their system~\cite{Avigad2019}.\alb{I kinda object to ``seamlessly''.
+In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky.
+I guess Isabelle is better}
 

### 2024-02-22 short-paper/short-paper.tex
Shorten the scope extrusion related work section.

@@
 since it concerns the properties of binders, whose encoding has been
-extensively studied with respect to the $\lambda$-calculi. Process calculi
+extensively studied with respect to $\lambda$-calculi. Process calculi
 present additional challenges, typically including several
@@
 scope extrusion.
+Scope extrusion requires reasoning about free variables, so approaches that identify \(\alpha\)-equivalent processes cannot be directly applied.
 
@@
 may be problematic in proof assistants based on dependent types.
-
-\todo{Sum up some of those issues}
-% \cite{Maksimovic2015} does not have
-% name restriction, thus dodging the challenge of scope extrusion.
-\begin{quotation}
-\cite{Castro-Perez2021,Castro2020} note that popular presentations of
-session type systems are impossible to directly encode using
-intrinsically \(\alpha\)-convertible terms.  \cite{Castro2020}
-suggests using several binder constructs to solve this, but notes that
-this requires proving many variants of the same theorems.
+Another approach is shallow embeddings~\cite{Castro-Perez2021}, which avoids the need to explicitly represent names.
 The case study \cite{AmbalLS21} compares four approaches to encoding
@@
 approaches developed for $\lambda$-calculus binders do not work well
-with scope extrusion.  Specifically, the locally nameless
-representation~\cite{Chargueraud2012} cannot avoid direct manipulation
-of de Bruijn indices when defining the semantics of scope extrusion;
-cofinite quantification provides no benefits when working under
-binders; nominal approaches require explicitly giving
-and validating sets of free names during scope extrusion;
-\end{quotation}
+with scope extrusion.
 

### 2024-02-22 short-paper/short-paper.tex
Add some comments.

@@
 by Honda et al.~\cite{Honda1998} was rectified in Yoshida and
-Vasconcelos’s subsequent work\cite{Yoshida2007}.
+Vasconcelos’s subsequent work\cite{Yoshida2007}.\todo{Marco: rewrite this to be correct}
 Similarly, several key results in papers on session types have turned out
-to be false later on, illustrating the
-value of machine-checked proofs~\cite{Gay2020,10.1145/3290343}.
+to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the
+value of machine-checked proofs.\todo{Rephrase this sentence to be more clear on the references}
 
@@
 \emph{in isolation} three key aspects that frequently pose challenges
-when mechanising concurrency theory: \emph{linearity}, \emph{scope
-  extrusion}, and \emph{coinductive reasoning}.  Mechanisations must
-often address several of these aspects at the same time; however, we
+when mechanising concurrency theory, as we will discuss in the next section: \emph{linearity}, \emph{scope
+  extrusion}, and \emph{coinductive reasoning}.
+Mechanisations must often address several of these aspects at the same time; however, we
 see the combination of techniques as a next
@@
 We designed this challenge to focus on linear reasoning while minimising
-definitions and other concerns.  We therefore forgo channel delegation:
-send/receive operations only support values that do not include channels, so
+definitions and other concerns.  We therefore forgo name passing:
+send/receive operations only support values that do not include channel names, so
 the topology of the communication network described by a process cannot change.
@@
 develop the infrastructure is significant.
+\todo{Consider adding material on linearity predicates}
 %
@@
 assistants and in so many encoding styles that it would take too long to
-review.  They range from concrete encodings based on named syntax
+review.
+\todo{However, there is no standard/consensus on what to do... elaborate}
+They range from concrete encodings based on named syntax
 \cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997, Perera2018} and locally nameless
@@
 replication of processes in order to enable infinite behaviours.  We do not
-include delegation since it is orthogonal to our
+include name passing since it is orthogonal to our
 primary aim of exploring coinductive proof techniques.  We base our

### 2024-02-22 short-paper/short-paper.tex
Fix some minor issues.

@@
 Similarly, several key results in papers on session types have turned out
-to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the
-value of machine-checked proofs.\todo{Rephrase this sentence to be more clear on the references}
+to be false later on (as described in \eg \cite{Gay2020,10.1145/3290343}), illustrating the potential
+value of machine-checked proofs.
 
@@
 review.
-\todo{However, there is no standard/consensus on what to do... elaborate}
+Despite this, almost all of these mechanisations still rely on ad-hoc solutions to accommodate scope extrusion.
 They range from concrete encodings based on named syntax
@@
 may be problematic in proof assistants based on dependent types.
-Another approach is shallow embeddings~\cite{Castro-Perez2021}, which avoids the need to explicitly represent names.
+Another approach is shallow embeddings~\cite{Castro-Perez2021}, which avoids the need to explicitly represent names.\todo{Merge the previous sentence into the HOAS section?}
 The case study \cite{AmbalLS21} compares four approaches to encoding
@@
 with de Bruijn indices was easiest and shortest since the existing
-approaches developed for $\lambda$-calculus binders do not work well
+approaches developed for $\lambda$-calculus binders did not work well
 with scope extrusion.

### 2024-02-22 short-paper/short-paper.tex
fix in intro about Kohei 09 paper

--- 2024-02-22 short-paper/short-paper.tex
@@
 the mechanisation of formal models for concurrent and distributed
-programming languages.
-We particularly focus on process calculi as these provide a simple but
-realistic showcase of common issues encountered during mechanisation
-efforts.
-Concurrent calculi are notably subtle: for instance, it took nine
-years before the incorrect proof of type safety in the seminal paper
-by Honda et al.~\cite{Honda1998} was rectified in Yoshida and
-Vasconcelos’s subsequent work\cite{Yoshida2007}.\todo{Marco: rewrite this to be correct}
-Similarly, several key results in papers on session types have turned out
-to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the
-value of machine-checked proofs.\todo{Rephrase this sentence to be more clear on the references}
+programming languages.  We particularly focus on process calculi as
+these provide a simple but realistic showcase of common issues
+encountered during mechanisation efforts.  Concurrent calculi are
+notably subtle: for instance, it took some years before the incorrect
+subject reduction proof in the original paper on session
+subtyping~\cite{GH99} was rectified in the journal extended
+version~\cite{GH05} with the use of polarities. % , and generally
+% addressed by Vasconcelos and Yoshida~\cite{Yoshida2007}.
+% it took nine years before the incorrect proof of type safety in the
+% seminal paper by Honda et al.~\cite{Honda1998} was rectified in
+% Yoshida and Vasconcelos’s subsequent
+% work\cite{Yoshida2007}.\todo{Marco: rewrite this to be correct}
+Similarly, other key results in session types papers have subsequently
+been proven incorrect~\cite{Gay2020,10.1145/3290343}, advocating the
+need for machine-checked proofs.
+% \todo{Rephrase this sentence to be more clear on the references}
 

### 2024-02-22 short-paper/short-paper.tex

--- 2024-02-22 short-paper/short-paper.tex
@@
 the mechanisation of formal models for concurrent and distributed
-programming languages.
-We particularly focus on process calculi as these provide a simple but
-realistic showcase of common issues encountered during mechanisation
-efforts.
-Concurrent calculi are notably subtle: for instance, it took nine
-years before the incorrect proof of type safety in the seminal paper
-by Honda et al.~\cite{Honda1998} was rectified in Yoshida and
-Vasconcelos’s subsequent work\cite{Yoshida2007}.\todo{Marco: rewrite this to be correct}
-Similarly, several key results in papers on session types have turned out
-to be false later on (as described in \eg \cite{Gay2020,10.1145/3290343}), illustrating the potential
-value of machine-checked proofs.
+programming languages.  We particularly focus on process calculi as
+these provide a simple but realistic showcase of common issues
+encountered during mechanisation efforts.  Concurrent calculi are
+notably subtle: for instance, it took some years before an incorrect
+subject reduction proof in the original paper on session
+subtyping~\cite{GH99} was discovered and then rectified in the journal
+extended version~\cite{GH05} with the use of
+polarities. % , and generally
+% addressed by Vasconcelos and Yoshida~\cite{Yoshida2007}.
+% it took nine years before the incorrect proof of type safety in the
+% seminal paper by Honda et al.~\cite{Honda1998} was rectified in
+% Yoshida and Vasconcelos’s subsequent
+% work\cite{Yoshida2007}.\todo{Marco: rewrite this to be correct}
+Similarly, other key results in session types papers have subsequently
+been proven incorrect~\cite{Gay2020,10.1145/3290343}, advocating the
+need for machine-checked proofs.
+% \todo{Rephrase this sentence to be more clear on the references}
+% =======
+% programming languages.
+% We particularly focus on process calculi as these provide a simple but
+% realistic showcase of common issues encountered during mechanisation
+% efforts.
+% Concurrent calculi are notably subtle: for instance, it took nine
+% years before the incorrect proof of type safety in the seminal paper
+% by Honda et al.~\cite{Honda1998} was rectified in Yoshida and
+% Vasconcelos’s subsequent work\cite{Yoshida2007}.\todo{Marco: rewrite this to be correct}
+% Similarly, several key results in papers on session types have turned out
+% to be false later on (as described in \eg \cite{Gay2020,10.1145/3290343}), illustrating the potential
+% value of machine-checked proofs.
+% >>>>>>> 1280d8dcd1c1a2b294f6f2d2cdf08e9808a9518c
 

### 2024-02-22 short-paper/short-paper.tex
Small edit.

--- 2024-02-22 short-paper/short-paper.tex
@@
 the mechanisation of formal models for concurrent and distributed
-programming languages.
-We particularly focus on process calculi as these provide a simple but
-realistic showcase of common issues encountered during mechanisation
-efforts.
-Concurrent calculi are notably subtle: for instance, it took nine
-years before the incorrect proof of type safety in the seminal paper
-by Honda et al.~\cite{Honda1998} was rectified in Yoshida and
-Vasconcelos’s subsequent work\cite{Yoshida2007}.\todo{Marco: rewrite this to be correct}
-Similarly, several key results in papers on session types have turned out
-to be false later on (as described in \eg \cite{Gay2020,10.1145/3290343}), illustrating the potential
-value of machine-checked proofs.
+programming languages.  We particularly focus on process calculi as
+these provide a simple but realistic showcase of common issues
+encountered during mechanisation efforts.  Concurrent calculi are
+notably subtle: for instance, it took some years before an incorrect
+subject reduction proof in the original paper on session
+subtyping~\cite{GH99} was discovered and then rectified in the journal
+extended version~\cite{GH05} with the use of
+polarities.
+Similarly, other key results in session types papers have subsequently
+been proven incorrect~\cite{Gay2020,10.1145/3290343}, demonstrating the
+need for machine-checked proofs.
 

### 2024-02-22 short-paper/short-paper.tex
sentence about linear predicate
changed title of sect 3, todo: conclusion

@@
 develop the infrastructure is significant.
-\todo{Consider adding material on linearity predicates}
+%\todo{Consider adding material on linearity predicates}
 %
@@
 approach, referred to as ``typing with leftovers,'' is exemplified
-in~\cite{DBLP:conf/forte/ZalakainD21}.  Whatever the choice,
+in~\cite{DBLP:conf/forte/ZalakainD21}.    Similarly, one
+could eliminate context splitting by delegating linearity checks to a
+\emph{linear predicate} defined on the process structure. These checks would
+serve as additional conditions within the typing rule,
+\eg~\cite{BP23}. Whatever the choice,
 list-based encodings can be refined to be intrinsically-typed, if the
@@
 %
+
 A radically different approach is to adopt a \emph{substructural}
@@
 recent example, see~\cite{Felty:MSCS21}). 
-
 \subsection{Scope extrusion}
@@
 % We previously promise to discuss combinations of techniques and applications to more elaborate schemes
-\section{Going Beyond the Challenge Problems}\label{sec:going-beyond}
+%\section{Going Beyond the Challenge Problems}\label{sec:going-beyond}
+\section{Future Work and Conclusion}\label{sec:going-beyond}
 \todo{Clean up this section to be more coherent.}
@@
 
+\todo{A sentence to conclude reiterating the relevance of the challenges}
 \bibliographystyle{splncs04}

### 2024-02-22 challenges/coinduction.tex
small changes to linearity section
fixed coinduction appendix, please check

@@
 
-\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
+A symmetric relation $\mathcal R$ is a {\em strong barbed
+  bisimulation} if $(P,Q)\in\mathcal R$ implies
 \begin{gather}
   \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu}\label{eq:bisim1} \\
-  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}\label{eq:bisim2}
+  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{Q'}~\textrm{and}~
+  (P',Q')\in\mathcal R\label{eq:bisim2}
 \end{gather}
-We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Note that, since our systems have potentially infinite behaviours, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.
+%
+Two processes are said to be \emph{strong barbed bisimilar}, written
+\( P\sbbisim{}{}Q \), if there exist a strong barbed bisimulation
+$\mathcal R$ such that $(P,Q)\in\mathcal R$. Note that {\em strong
+  barbed bisimilarity} \( \sbbisim{}{} \) is the largest strong barbed
+bisimulation.
+% symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
+% \begin{gather}
+%   \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu}\label{eq:bisim1} \\
+%   \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}\label{eq:bisim2}
+% \end{gather}
+% We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
+Also, since our systems have potentially infinite behaviours,
+bisimilarity cannot be defined inductively since it is the largest
+strong barbed bisimulation.
 
@@
   \item Reflexivity is straightforward: for any $P$, we need to show
-    that $P\sbbisim{}{} P$. Both conditions ~\ref{eq:bisim1} and \ref{eq:bisim2} follow trivially.
+    that $P\sbbisim{}{} P$. In order to do so, we choose the identity
+    relation and prove that it is a strong barbed
+    bisimulation. Condition~\ref{eq:bisim1} follows trivially by
+    definition. Condition~\ref{eq:bisim2} follows coinductively since
+    we must always reach identical pairs.
   \item Symmetry follows immediately by definition.

### 2024-02-23 short-paper/short-paper.tex
Merge mention of shallow embeddings for linearity into HOAS part.

@@
 may be problematic in proof assistants based on dependent types.
-Another approach is shallow embeddings~\cite{Castro-Perez2021}, which avoids the need to explicitly represent names.\todo{Merge the previous sentence into the HOAS section?}
 The case study \cite{AmbalLS21} compares four approaches to encoding
@@
 Higher-order abstract syntax (HOAS) has seen extensive and successful
-use in formal reasoning in this area~\cite{Despeyroux2000,Honsell2001,Tiu2010,ChaudhuriCM15}. Its
+use in formal reasoning in this area~\cite{Despeyroux2000,Honsell2001,Tiu2010,ChaudhuriCM15,Castro-Perez2021}. Its
 weak form aligns well with mainstream inductive proof assistants,

### 2024-02-23 challenges/coinduction.tex
Minor edits in coinduction challenge.

@@
 Two processes are said to be \emph{strong barbed bisimilar}, written
-\( P\sbbisim{}{}Q \), if there exist a strong barbed bisimulation
+\( P\sbbisim{}{}Q \), if there exists a strong barbed bisimulation
 $\mathcal R$ such that $(P,Q)\in\mathcal R$. Note that {\em strong
@@
 % We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
-Also, since our systems have potentially infinite behaviours,
+Also, since our processes have potentially infinite behaviours,
 bisimilarity cannot be defined inductively since it is the largest

### 2024-02-23 short-paper/short-paper.tex
Clean up the future work and conclusion section.

@@
 \section{Future Work and Conclusion}\label{sec:going-beyond}
-\todo{Clean up this section to be more coherent.}
-Our benchmark challenges do not cover all issues in the field, but focus on the fundamental textbook aspects of linearity, scope extrusion, and coinduction.
-Many mechanisations need to combine techniques to handle several of these aspects, and some may also need to handle aspects that are not covered by our benchmark challenges at all.
+Our benchmark challenges do not cover all issues in the field, but focus on the fundamental aspects of linearity, scope extrusion, and coinduction.
+Many mechanisations need to combine techniques to handle several of these aspects at once, and some may also need to handle aspects that are not covered by our benchmark challenges at all.
 
-% \cite{Brady2017} notes that using algebraic effects is unwieldy when there are many communication channels and suggests an experimental approach using dependent uniqueness types.
-% \cite{Zalakain2019} notes that systems lacking linear types must simulate linearity, and that using a linearity predicate complicates congruences.
-% They had to limit themselves to an asymmetric structural precongruence to be able to prove subject reduction.
-Some authors have already described the difficulties that combining techniques may induce in a mechanisation, \eg~\cite{DBLP:conf/forte/ZalakainD21}.
+%%% Combinations
+Combining techniques for mechanising the fundamental aspects covered in our benchmark is often not trivial.
+While we have chosen to focus on the aspects individually to simplify the challenges, we are of course also interested in exploring how techniques interact.
+Some authors have already described the difficulties that combining techniques may induce in a mechanisation, \eg~\cite{DBLP:conf/forte/ZalakainD21}.\todo{Add more references here?}
 
+%%% Other aspects
+Much current research on concurrent calculi includes aspects that are not covered by our benchmark challenges.
 Features that are in wide use, but not covered by our benchmark, include choice constructs and recursion.
-Some interesting aspects of message-passing calculi include multiparty session types~\cite{10.1145/2827695} and choreographies~\cite{Carbone2013}, as their meta-theory includes aspects---\eg well-formedness conditions on global types, partiality of end-point projection function, \etc---that we do not address.
-Other areas of interest include higher-order calculi~\cite{Hirsch2022}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, or encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:journals/iandc/DardhaGS17}.
+Some interesting research topics that build on the fundamental aspects in our benchmark include multiparty session types~\cite{10.1145/2827695}, choreographies~\cite{Carbone2013}, higher-order calculi~\cite{Hirsch2022}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, and encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:journals/iandc/DardhaGS17}.
+Their meta-theory includes aspects---\eg well-formedness conditions on global types, partiality of end-point projection functions, \etc---that we do not address.
 
-For coinduction, one could explore different notions of process equivalence.
+Our coinduction challenge only includes two notions of process equivalence, but many more exist in the literature.
+Coinduction may also play a role in recursive processes and session types: recursive session types can be expressed in \emph{infinitary form} by interpreting their typing rules coinductively~\cite{DerakhshanPfenning22,HornePadovani23}.
 
-Coinduction may also play a relevant role in supporting recursive processes and session types.
-Indeed, it is often the case that session type systems with recursive session types are naturally expressed in \emph{infinitary form} by interpreting their typing rules
-coinductively~\cite{DerakhshanPfenning22,HornePadovani23}.
+Unlike POPLMark, we consider \emph{animation} of calculi (as in \cite{Castro-Perez2021}) out of scope for our benchmark.
+Finally, our challenges encourage, but do not require, exploring proof automation, as offered by \eg the \emph{Hammer} protocol~\cite{BohmeN10,CzajkaK18}.
 
-Finally, another interesting avenue to explore would be to take
-advantage of certain proof assistant features: for example, extracting
-certified implementations (as in \cite{Castro-Perez2021}) or exploiting
-proof automation, such as the one offered by the \emph{Hammer}
-protocol~\cite{BohmeN10,CzajkaK18}.
-
-Ultimately, our challenges can be extended in several worthwhile directions, and we look forward to a future where they indeed are.
-It is our hope and aim that our challenges will move us closer to a future where the key basic proof techniques for concurrent calculi are as easy to mechanise as they are to write on paper.
+%%% Conclusion
+Ultimately, the fundamental aspects covered by our benchmark serve as the building blocks for most current research on concurrenct calculi.
+It is our hope and aim that exploring solutions to our challenges will move the community closer to a future where the key basic proof techniques for concurrent calculi are as easy to mechanise as they are to write on paper.
 

### 2024-02-23 challenges/linearity.tex
Further edits to draft

@@
   \begin{enumerate}
-  \item If \( \types{\Gamma; \Delta}{P} \) then
+  \item If \( \types{\Gamma; \Delta}{P} \), then
     \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
-  \item If \( \types{\Gamma; \Delta}{P} \) % and \( \tend{\Delta} \)
+  \item If \( \types{\Gamma; \Delta}{P} \), % and \( \tend{\Delta} \)
     then
@@
   \label{lemma:strenD}
-  If \( \types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P} \) and
-  $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
+  If \(\types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P}\) and
+  $x\not\in \freenames{P}$, then \(\types{\Gamma; \Delta}{P}\).
 \end{lemma}
@@
   \label{le:presequiva}
-  If \( \sconga{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
+  If \( \sconga{P}{Q} \), then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
 \end{lemma}
@@
   \label{le:presequiv}
-    If \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
+    If \( \scong{P}{Q} \), then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
 \end{lemma}
@@
 \begin{theorem}[Subject reduction]
-  If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
+  If \( \types{\Gamma;\Delta}{P} \) and \( \reduces{P}{Q} \), then \( \types{\Gamma;\Delta}{Q} \).
 \end{theorem}
@@
 \begin{corollary}
-  If \( \types{\Gamma;\Cempty}{P} \) and \( \reduces{P}{Q} \) then \( Q \) is well formed.
+  If \( \types{\Gamma;\Cempty}{P} \) and \( \reduces{P}{Q} \), then \( Q \) is well formed.
 \end{corollary}

### 2024-02-23 challenges/scope-extrusion.tex
Edits and comments
Further edits to draft

@@
 \begin{lemma}\label{se-lemma-normalized-reduction}
-  If \( \reduces{P}{Q} \) then there are \( x, y, z, z_1, \dots, z_n, R_1, R_2, \) and \( S\) such that
+  If \( \reduces{P}{Q} \), then there are \( x, y, z, z_1, \dots, z_n, R_1, R_2, \) and \( S\) such that
   \begin{align*}
@@
 \begin{proof}[Sketch]
-  The proof is by induction on the inference of \( \transition{P}{\Atau}{Q} \) using the following lemmas:
+  The proof is by induction on the inference of \( \transition{P}{\Atau}{Q} \) using the following lemmata:
   \begin{enumerate}

### 2024-02-23 short-paper/short-paper.tex
Further edits to draft

@@
 shift towards publications that include mechanised proofs within the programming language research community.
-POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a similar programme targeting proofs using logical relations.
+POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a similar programme, targeting proofs using logical relations.
 These initiatives had a narrow focus, but their authors recognised the importance of addressing topics such as coinduction and linearity in the future.
@@
 specifically crafted to tackle common challenges encountered during
-the mechanisation of formal models for concurrent and distributed
-programming languages.  We particularly focus on process calculi as
-these provide a simple but realistic showcase of common issues
-encountered during mechanisation efforts.  Concurrent calculi are
+the mechanisation of formal models of concurrent and distributed
+programming languages.  We particularly focus on process calculi, as
+they provide a simple but realistic showcase of these challenges.  Concurrent calculi are
 notably subtle: for instance, it took some years before an incorrect
 subject reduction proof in the original paper on session
-subtyping~\cite{GH99} was discovered and then rectified in the journal
-extended version~\cite{GH05} with the use of
+subtyping~\cite{GH99} was discovered and then rectified in the extended journal version~\cite{GH05} with the use of
 polarities.
@@
 below), our experience is that choosing appropriate mechanisation
-techniques and tools remains a significant challenge and that the
+techniques and tools remains a significant challenge and that their
 trade-offs are not well understood.  This often leads researchers
 toward a trial-and-error approach, resulting in sub-optimal solutions,
-wasted mechanisation effort, and techniques and results that are hard
-to reuse. For example, the authors in~\cite{Cruz-Filipe2021b} note that the high-level parts of
+wasted mechanisation efforts, and techniques and results that are hard
+to reuse. For example, the authors of~\cite{Cruz-Filipe2021b} note that the high-level parts of
 mechanised proofs closely resemble the informal ones, while the
@@
 \emph{in isolation} three key aspects that frequently pose challenges
-when mechanising concurrency theory, as we will discuss in the next section: \emph{linearity}, \emph{scope
+when mechanising concurrency theory, which we will discuss in more detail in the next section: \emph{linearity}, \emph{scope
   extrusion}, and \emph{coinductive reasoning}.
@@
 %
-In the future, we expect to use the website to promote best practices and tutorials derived from solutions to our challenges.
+In the future, we intend to use the website to promote best practices and tutorials derived from solutions to our challenges.
 We encourage readers to try the challenges using their favourite techniques, and to send us their solutions.
 
-\section{The Benchmark and its Design}\label{sec:design-discussion}
-In this section, we outline the factors considered in designing the benchmark challenges.
-We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the evaluation criteria for solutions.
+\section{Overview and design considerations}\label{sec:design-discussion}
+In this section, we outline the factors considered when designing the benchmark challenges.
+We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the criteria for evaluating solutions thereof.
 
@@
 \item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
-\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
+\item\label{item:rq3} What improvements are needed to make mechanisation tools more user-friendly with regard to the issues faced when mechanising process calculi?
 \end{enumerate}
@@
 process calculus used in each challenge focuses on the features that
-we want to emphasise, and omits all constructs (such as choices) that
+we want to emphasise, and omits all constructs %(such as choice) 
+that
 would complicate the mechanisation without bringing additional
@@
 \subsection{Linearity}
-Linear typing systems allow  tracking the use of resources in a program. In
-the case of typed process calculi (in particular, session types), linearity is
+Linear typing systems allow for the tracking of resource usage in a program. In
+the case of typed (in particular, session-typed) process calculi, linearity is
 widely used for checking if and how a channel is used to send or receive values.
-This leads to substructural typing~\cite[Chapter 1]{Pierce2004-oq} with
-resulting mechanisation difficulties --- \eg deciding how to \emph{split the
+This substructurality~\cite[Ch. 1]{Pierce2004-oq} gives rise to mechanisation difficulties: \eg deciding how to \emph{split the
 typing context} in a parallel composition.
 
-The goal of our challenge on linear reasoning is proving a type safety theorem
-for a process calculus with session types, by combining as usual subject
-reduction with absence of errors.  We have chosen  only to model linear (as
+The goal of our challenge on linear reasoning is to prove a type safety theorem
+for a process calculus with session types, by combining subject
+reduction with the absence of errors.  For simplicity we model only linear (as
 opposed to shared) channels. Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
@@
 expressed. We also forgo more sophisticated notions of error-freedom (\eg
-deadlock freedom) because proving them would distract from the core linear
+deadlock freedom) as proving them would distract from the core linear
 aspects of the challenge.
@@
 session types).  In this space, several designs are possible: one can label
-occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}); some
-authors instead impose a multiset structure over lists
+occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}), or impose a multiset structure over lists
 (\eg~\cite{Danielsson12,ChaudhuriLR19}). Alternatively, contexts can be
@@
 could eliminate context splitting by delegating linearity checks to a
-\emph{linear predicate} defined on the process structure. These checks would
-serve as additional conditions within the typing rule,
-\eg~\cite{BP23}. Whatever the choice,
+\emph{linear predicate} defined on the process structure (\eg~\cite{BP23}). These checks 
+serve as additional conditions within the typing rule. Whatever the choice,
 list-based encodings can be refined to be intrinsically-typed, if the
@@
 fully support the verification of meta-theoretic properties.  A
-compromise is the so called \emph{two-level} approach, \ie encoding a
+compromise is the \emph{two-level} approach, \ie encoding a
 substructural specification logic in a mainstream proof assistant and
@@
 recent example, see~\cite{Felty:MSCS21}).
+
 \subsection{Scope extrusion}
@@
 This is the challenge most closely related to POPLMark,
-since it concerns the properties of binders, whose encoding has been
+as it concerns the properties of binders, whose encoding has been
 extensively studied with respect to $\lambda$-calculi. Process calculi
@@
 They range from concrete encodings based on named syntax
-\cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997, Perera2018} and locally nameless
+\cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997, Perera2018} and locally-nameless
 representation %~\cite{Chargueraud2012}
@@
 with higher-order process calculi.  They found that working directly
-with de Bruijn indices was easiest and shortest since the existing
-approaches developed for $\lambda$-calculus binders did not work well
+with de~Bruijn indices was easiest since the existing
+approaches developed for $\lambda$-calculus binders worked poorly 
 with scope extrusion.
@@
 semantics. However, when addressing more intricate concepts like
-bisimulation, HOAS necessitates extensions. These extensions may take
+bisimulation, extensions to HOAS are needed. These extensions may take
 the form of additional axioms~\cite{Honsell2001} or require specialised proof assistants
@@
 Process calculi typically include constructs that allow processes to
-have infinite behaviours.  Coinduction serves as a fundamental method
+adopt infinite behaviours.  Coinduction serves as a fundamental method
 for the definition and analysis of infinite objects, enabling the
@@
 The setting for our challenge is an untyped \(\pi\)-calculus augmented with
-replication of processes in order to enable infinite behaviours.  We do not
+process replication in order to enable infinite behaviours.  We do not
 include name passing since it is orthogonal to our
@@
 use both and establish an internal adequacy~\cite{Honsell2001}. Few
-extend the proof assistant foundations to allow \eg reasoning about
+extend the proof assistant foundations to allow, \eg, reasoning about
 bisimilarity up-to~\cite{ChaudhuriCM15}.
@@
 % We base our definition of bisimilarity on a labelled transition system semantics and an observability predicate describing the communication steps available to a process.
-% The introduction of strong barbed bisimulation is one of the first steps when studying the observable behaviour of process calculi, both in classical textbooks (\eg, \cite{picalcbook}) and in existing mechanisations.
+% The introduction of strong barbed bisimulation is one of the first steps when studying the observable behaviour of process calculi, both in classical textbooks (\eg \cite{picalcbook}) and in existing mechanisations.
 % Though weak barbed congruence is a more common behavioural equivalence, we prefer strong equivalences to simplify the theory by avoiding the need to abstract over the number of internal transitions in a trace.
@@
 \subsection{Evaluation criteria.}
-The idea behind our benchmark is to obtain evidence towards answering
+The motivation behind our benchmark is to obtain evidence towards answering
 questions \cref{item:rq1,item:rq2,item:rq3}. We are therefore
@@
 \begin{enumerate}
-\item Mechanisation overhead: the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation;
+\item Mechanisation overhead: the amount of manually-written infrastructure and setup needed to express the definitions in the mechanisation;
 \item Adequacy of the formal statements in the mechanisation: whether the proven theorems are easily recognisable as the theorems from the challenge; and
@@
 \end{enumerate}
-Solutions to our challenges do not need to strictly follow the definitions and lemmata set out in the problem text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.
+Solutions to our challenges need not strictly follow the definitions and lemmata set out in the problem text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.
 
@@
 %\section{Going Beyond the Challenge Problems}\label{sec:going-beyond}
-\section{Future Work and Conclusion}\label{sec:going-beyond}
+\section{Future work and conclusions}\label{sec:going-beyond}
 Our benchmark challenges do not cover all issues in the field, but focus on the fundamental aspects of linearity, scope extrusion, and coinduction.

### 2024-02-23 challenges/coinduction.tex
Use title case for LNCS guideline compliance.

@@
-\subsection{Challenge: Coinduction and infinite processes}
+\subsection{Challenge: Coinduction and Infinite Processes}
 % and reasoning about infinite processes}
@@
 
-\subsubsection{Bisimilarity.}
+\subsubsection{Strong Barbed Bisimilarity.}
 Bisimilarity is a notion of equivalence for processes and
@@
 
-\subsubsection{Strong barbed congruence.}
+\subsubsection{Strong Barbed Congruence.}
 In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,

### 2024-02-23 challenges/linearity.tex
Use title case for LNCS guideline compliance.

@@
-\subsection{Challenge: Linearity and behavioural type systems}
+\subsection{Challenge: Linearity and Behavioural Type Systems}
 \label{sec:challenge:linearity-beh-types}
@@
 
-\subsubsection{Session types.}
+\subsubsection{Session Types.}
 Our process syntax allows us to write processes that are ill formed in
@@
 
-\subsubsection{Typing rules.}
+\subsubsection{Typing Rules.}
 Our type system is aimed at maintaining two invariants:

### 2024-02-23 challenges/scope-extrusion.tex
Use title case for LNCS guideline compliance.

@@
-\subsection{Challenge: Name passing and scope extrusion}
+\subsection{Challenge: Name Passing and Scope Extrusion}
 \label{sec:challenge:name-passing-scope-extrusion}
@@
 
-\subsubsection{Reduction semantics.}
+\subsubsection{Reduction Semantics.}
 The first semantics is an operational reduction semantics, which avoids reasoning explicitly about scope extrusion by way of a structural congruence rule.
@@
 
-\subsubsection{Transition system semantics.}
+\subsubsection{Transition System Semantics.}
 The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.

### 2024-02-23 short-paper/short-paper.tex
Use title case for LNCS guideline compliance.

@@
 
-\section{Overview and design considerations}\label{sec:design-discussion}
+\section{Overview and Design Considerations}\label{sec:design-discussion}
 In this section, we outline the factors considered when designing the benchmark challenges.
@@
 
-\subsection{Scope extrusion}
+\subsection{Scope Extrusion}
 This challenge revolves around the mechanisation of scope
@@
 
-\subsection{Evaluation criteria.}
+\subsection{Evaluation Criteria.}
 The motivation behind our benchmark is to obtain evidence towards answering
@@
 %\section{Going Beyond the Challenge Problems}\label{sec:going-beyond}
-\section{Future work and conclusions}\label{sec:going-beyond}
+\section{Future Work and Conclusions}\label{sec:going-beyond}
 Our benchmark challenges do not cover all issues in the field, but focus on the fundamental aspects of linearity, scope extrusion, and coinduction.

### 2024-02-23 short-paper/short-paper.tex
Minor edits.

@@
 wasted mechanisation efforts, and techniques and results that are hard
-to reuse. For example, the authors of~\cite{Cruz-Filipe2021b} note that the high-level parts of
+to reuse. For example, Cruz-Filipe et al.~\cite{Cruz-Filipe2021b} note that the high-level parts of
 mechanised proofs closely resemble the informal ones, while the
@@
 textbook concurrency theory, process calculi, and type theory.  The
-process calculus used in each challenge focuses on the features that
-we want to emphasise, and omits all constructs %(such as choice) 
-that
-would complicate the mechanisation without bringing additional
+process calculi used in the challenges focus on the features that
+we want to emphasise, and omit all constructs that would complicate the mechanisation without bringing additional
 tangible insights.  The minimality and uniformity of the

### 2024-02-23 short-paper/short-paper.tex
Change "problem" to "challenge".

@@
   POPLMark and POPLMark Reloaded sparked a flurry of work on machine-checked proofs, and fostered the adoption of proof mechanisation in programming language research.
-  Both challenges were purposely limited in scope, and their benchmark problems do not address concurrency-related issues.
+  Both challenges were purposely limited in scope, and their benchmark challenges do not address concurrency-related issues.
   %
-  We propose a new collection of benchmark problems focused on the difficulties that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.
-  Our benchmark problems address three key topics: linearity, scope extrusion, and coinductive reasoning.
+  We propose a new collection of benchmark challenges focused on the difficulties that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.
+  Our benchmark challenges address three key topics: linearity, scope extrusion, and coinductive reasoning.
   The goal of this new challenge is to clarify, compare, and advance the state of the art, fostering the adoption of proof mechanisation in future research on concurrency.
@@
 
-In this spirit, we introduce a new collection of benchmark problems
+In this spirit, we introduce a new collection of benchmark challenges
 specifically crafted to tackle common challenges encountered during
@@
 
-Our benchmark problems (detailed in \cref{app:challenges} and on our website) consider
+Our benchmark challenges (detailed in \cref{app:challenges} and on our website) consider
 \emph{in isolation} three key aspects that frequently pose challenges
@@
 In this section, we outline the factors considered when designing the benchmark challenges.
-We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the criteria for evaluating solutions thereof.
+We begin with some general remarks, then describe the individual design considerations for each challenge, and the criteria for evaluating solutions thereof.
 
@@
 To achieve design goal \ref{item:goal-comperison-accessibility}, we have
-formulated our challenge problems to explore the three aspects
+formulated our challenges to explore the three aspects
 (linearity, scope extrusion, coinduction) independently, so that they
-may be solved individually and in any order; each problem is
+may be solved individually and in any order; each challenge is
 reasonably small and easily understandable with basic knowledge of
@@
 \ref{item:goal-tutorials}.
-Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
+Aligned with design goal~\ref{item:goal-reusability}, our challenges are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
 Our challenges centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in many applications (as we will further discuss in \cref{sec:going-beyond}).
@@
 \end{enumerate}
-Solutions to our challenges need not strictly follow the definitions and lemmata set out in the problem text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.
+Solutions to our challenges need not strictly follow the definitions and lemmata set out in the challenge text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.
 

### 2024-02-23 challenges/coinduction.tex
Remove old comments.

@@
 \subsection{Challenge: Coinduction and Infinite Processes}
-% and reasoning about infinite processes}
 \label{sec:challenge:coinduction}
 
-% \marco{Why reasoning is only about infinite processes? Would it be
-%   unreasonable to shorten the title to {\em Coinduction and infinite
-%     processes}?}
-
 This challenge is  about the mechanisation of proofs concerning processes wih infinite behaviours.
-% formalises a proof
-%that require coinduction.
-% coinductive techniques.
-%
-% Coinduction is a proof technique for infinite structures,
 This is usually connected to \emph{coinductive} definitions where an infinite
@@
 such infinite structures.
-%
-% \marco{Should we be more specific here and give a deeper explanation
-%   of what coinduction is? If not, then why do we even say that it's
-%   the dual of induction?}
-%
 
@@
 that includes process replication.
-% The setting for this challenge is an untyped calculus of
-% communicating systems with replication of processes, the syntax and
-% semantics of which are presented below.
-The objective of this challenge is to % prove % a context theorem for
+The objective of this challenge is to
 draw a formal connection between strong barbed congruence and strong
-barbed bisimilarity.  % I.e., The context lemma
+barbed bisimilarity.
 The result establishes that two processes are strong barbed congruent
@@
 \subsubsection{Syntax.}
-% We assume the existence of some type of \emph{base values}, values
-% of which we will denote by \( a, b, \dots \), the existence of some
-% type of \emph{variables}, values of which we will denote by
-% \( l, m, \dots \), and the existence of some type of \emph{names},
-% values of which we will denote by \( x, y, \dots \).
 The syntax of values and processes is given by:
-% for including replication and having unary restriction:
 \begin{displaymath}
@@
 
-%
-% \begin{align*}
-%   v,w :=&&& a \\
-%   |&&& l \\
-%   P,Q :=&&& \Pend \\
-%   |&&& \Pout{x}{v}{P} \\
-%   |&&& \Pin{x}{l}{P} \\
-%   |&&& \Ppar{P}{Q} \\
-%   |&&& \Pres{x}{P} \\
-%   |&&& \Preplicate{P}
-% \end{align*}
-% A \emph{value} \( v, w, \dots \) is either a base value \( a \) or a
-% variable \( l \).  The process \( \Pend \) is \emph{inaction}:
-% a process which can do nothing.  The process \( \Pout{x}{a}{P} \) is
-% an \emph{output}, which can send the value \( v \) via \( x \), then
-% continue as \( P \).  The intention is that the value \( v \) must
-% be a base value when it is actually sent, and this will be enforced
-% in the semantics later on.  The process \( \Pin{x}{l}{P} \) is an
-% \emph{input}, which can receive a base value via \( x \), then
-% continue as \( P \) with the received value substituted for the
-% variable \( l \).  The input operator thus binds the variable
-% \( l \) in \( P \).  The process \( \Ppar{P}{Q} \) is the
-% \emph{composition} of process \( P \) and process \( Q \).  The two
-% components can proceed independently of each other, or they can
-% interact via shared names.  The process \( \Pres{x}{P} \) is the
-% \emph{restriction} of the name \( x \) to \( P \).  Components in
-% \( P \) can use the name \( x \) to interact with each other, but
-% not with processes outside of the restriction.  The restriction
-% operator thus binds the name \( x \) in \( P \).  Note that the
-% scope of a restriction may not change when processes interact, since
-% it is only possible to send and receive values, and not names.  The
-% process \( \Preplicate{P} \) is the \emph{replication} of the
-% process \( P \).  It can be thought of as the infinite composition
-% \( \Ppar{P}{\Ppar{P}{\cdots}} \).  Replication makes it possible to
-% express infinite behaviours.
-
-% We will use the notation \( \subst{P}{v}{l} \) to denote the process
-% \( P \) with value \( v \) substituted for variable \( l \).
-
-% % We will use the notation \( \subst{P}{v}{w} \) to denote the process \( P \) with value \( v \) substituted for value \( w \).
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible,
-% written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from
-% \( P \) by a finite number of substitutions of bound variables.
-% As a convention, we will identify \( \alpha \)-convertible
-% processes.
-
-% Also as a convention, we assume that the bound names and bound
-% variables occurring in any collection of processes are chosen
-% to be different from the free names and free variables
-% occurring in those processes and from the names and variables
-% occurring in any substitutions applied to the processes.  This is
-% justified because any overlapping names and variables may be
-% \( \alpha \)-converted such that the assumption is satisfied.
-
 \subsubsection{Semantics.} We choose to give an LTS semantics for this
-challenge. %As an alternative, it is also possible to adopt a reduction
-%semantics which features fewer rules but needs a structural
-%congruence. %  The semantics of the system describes the actions that the
-% system can perform by defining a labelled transition relation on
-% processes.
+challenge.
 
@@
 of which is as follows:
-% \begin{align*}
-%   \alpha := &&& \Aoutf{x}{a} \\
-%   |&&& \Ain{x}{a} \\
-%   |&&& \Atau
-% \end{align*}
 \begin{align*}
@@
 
-% As a convention, we assume that bound names and bound variables of any processes or actions are chosen to be different from the names and variables that occur free in any other entities under consideration, such as processes, actions, substitutions, and sets of names or variables.
-
 \subsubsection{Strong Barbed Bisimilarity.}
 Bisimilarity is a notion of equivalence for processes and
-% .  Our notion of process equivalence relations
 builds on a notion of \emph{observables}, i.e., what we can externally
@@
 bisimulation.
-% symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
-% \begin{gather}
-%   \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu}\label{eq:bisim1} \\
-%   \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'}\label{eq:bisim2}
-% \end{gather}
-% We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
 Also, since our processes have potentially infinite behaviours,
@@
 \begin{theorem}
-  \( \sbbisim{}{} \) is an equivalence relation. %that is, the relation is reflexive, symmetric, and transitive.
+  \( \sbbisim{}{} \) is an equivalence relation.
 \end{theorem}
 \begin{proof}
-  % We need to prove that $\sbbisim{}{}$ is reflexive, symmetric, and
-  % commutative.
   We prove the three properties separately: 
@@
   \item Symmetry follows immediately by definition.
-%    since bisimilarity is defined as a {\em symmetric} relation.
   \item For transitivity, we need to prove that if $P\sbbisim{}{} Q$
@@
 substituting a process with an equivalent one in any context.
-% equivalent processes to be since it does not consider the
-% environment of processes.
 For instance, the processes $\Pout{x}{a}{\Pout{y}{b}{\Pend}}$ and
@@
 
-% For instance,
-% \( \sbbisim{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pout{x}{a}{\Pend}} \)
-% since \( \obsout{x} \) is the only observable in both processes and
-% they cannot perform a \( \Atau \)-action, but
-% \(
-% \nsbbisim{\Ppar{\Pout{x}{a}{\Pout{y}{b}{\Pend}}}{\Pin{x}{l}{\Pend}}}{\Pout{x}{a}{\Pend}}
-% \) since the left process can perform a \( \Atau \)-action such that
-% \( \obsout{y} \) becomes observable, whereas the right process cannot.
-
-% \subsubsection{Contexts and congruences}
-% Before we can fix the issue with strong barbed bisimilarity, we need few more definitions.
-
-% A \emph{context} is obtained by taking a process and replacing a single occurrence of \( \Pend \) in it with the special \emph{hole} symbol \( \ctxhole \).
-% As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-
-% We can think of contexts as functions between processes.
-% A context \( C \) can be \emph{applied} to a process \( P \), written \( \applyctx{C}{P} \), by replacing the hole in C by \( P \), thus obtaining another process.
-% The replacement should be literal, so names and variables that are free in \( P \) can become bound in \( \applyctx{C}{P} \).
-
-% We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
-
 \subsubsection{Strong Barbed Congruence.}
 In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,
-i.e., % we need to make strong barbed bisimilarity consider
+i.e.,
 we have to consider the environment in which processes may be placed.
-% (i.e.\ the context) of processes.
 
-% We define \emph{strong barbed congruence}, written \( \sbcong{}{} \), by saying
 We say that two processes \( P \) and \( Q \) are \emph{strong barbed congruent}, written \( \sbcong{P}{Q} \), if \( \sbbisim{\applyctx{C}{P}}{\applyctx{C}{Q}} \) for every context \( C \).

### 2024-02-23 challenges/linearity.tex
Remove old comments.

@@
 The input process \( \Pin{x}{l}{P} \) waits for a base value from \( x \) and then continues as \( P \) with the received value substituted for the variable \( l \).
-%
 The process \( \Presd{x}{y}{P} \) represents a \emph{session} with endpoints named \( x \) and \( y \) which are bound in \( P \).
 In \( P \), the names \( x \) and \( y \) can be used to exchange messages over the session (sending on \( x \) and receiving on \( y \) or vice versa).
-%
-%Finally, the process $\PBpar{P}{Q}$ represents the parallel composition of $P$ and $Q$.
-%
 Note that in this calculus channels cannot be sent in messages, therefore the topology of the communication network described by a process cannot change.
@@
 \end{mathpar}
-% In the typing rule for names, $\tend \Phi$ denotes that all types in the range of $\Phi$ are $\Tend$:
-% \begin{mathpar}
-%   \inferrule[T-Name]{{\tend\Delta }}{\types{{\Gamma; (\Delta,\hastype{x}{T})}}{\hastype{x}{T}}}
-% \end{mathpar}
 The typing rules for processes are as follows:
@@
     \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
-  \item If \( \types{\Gamma; \Delta}{P} \), % and \( \tend{\Delta} \)
+  \item If \( \types{\Gamma; \Delta}{P} \),
     then
@@
   By induction on the derivation of
-%%  last applied rule of the proof of --AM changed for uniformity
   $\types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P}$.
-  % , for each case inverting $\types{\Gamma; (\Delta, x : T)}{P}$ to
-  % the uniquely applicable rule.
   \begin{description}
@@
     the rule without $\hastype{x}{T}$.
-    % [$P = \Pend$] we Apply \TirName{T-Inact}.
 
@@
     that context.
-    %
-    %
-    % [$P = \Ppar{P_0}{P_1}$] By inversion on the typing we have  $ \Delta, x : T = \Delta_0 , \Delta_1$.
-    %   By cases on which environment $x$ is in, apply IH on that environment.
   \item[\TirName{T-Res}] Without loss of generality, we assume that
@@
     $\types{\Gamma;\Delta}{\Presd{y}{z}{P'}}$.
-    % inversion,
-    % $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From
-    % the assumptions present we can apply IH.
-    %
-    % [$P = \Presd{y}{z}{P}$] We may assume $x \notin \{y,z\}$. By
-    % inversion,
-    % $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From
-    % the assumptions present we can apply IH.
   \end{description}
@@
     on the structure of $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.
-    % \item[$P = \Pin{x}{v}{P}$] Immediate after inverting typing.
   \end{description}
@@
   \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear or $\Tend$ and applying weakening and strengthening accordingly.
-  \item[\TirName{Res-Inact:}] Use weakening %~\ref{lemma:weak} part 2
+  \item[\TirName{Res-Inact:}] Use weakening
   \item[\TirName{Res}] Noting that order does not matter
@@
   $x_1$ and $y_1$ cannot be dual.
-  %
-  % We proceed by induction on the definition of $\types{}{}$. The key
-  % cases are that for parallel composition and that for name
-  % restriction. In the case of parallel composition, linearity (and
-  % context splitting) ensure that every process in parallel will have
-  % different channel names. Then, the restriction rule enforces
-  % duality, therefore an output will be matched by an input.
 \end{proof}

### 2024-02-23 challenges/preliminaries.tex
Remove old comments.

@@
 The two components can proceed independently of each other, or they can interact via shared names.
-%The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
 
@@
 The process \( \Pout{x}{c}{P} \) is an \emph{output}, which can send
-\( c \) via \( x \), then continue as \( P \).  % The intention is that
-% the value \( v \) must be a base value when it is actually sent, and
-% this will be enforced in the semantics later on
+\( c \) via \( x \), then continue as \( P \).
 The process \( \Pin{x}{k}{P} \) is an \emph{input}, which can receive a $c$
@@
 names that occur free,
-% (i.e.\ not bound by a restriction) in \( P \),
 \( \boundnames{P} \) to denote the set of names that occur bound
-% (by a restriction)
 in \( P \) and  \( \freevars{P} \)
 to denote the set of variables that occur free
-% (i.e.\ not bound by an input)
 in \( P \).  We use the notation \( \boundvars{P} \)
 for the set of variables that occur bound
-% (by an input)
 in \( P \).  We use the notation \( \subst{P}{a}{l} \) to denote
@@
 As a convention, we do \emph{not} identify \( \alpha \)-convertible contexts.
-%
 A context acts as a function between processes:

### 2024-02-23 short-paper/short-paper.tex
Remove LaTeX that modifies the page layout.
Remove old comments.

@@
 \usepackage{cleveref}
-\usepackage{hyperref}
 \usepackage{microtype}
@@
 \usepackage{color}
-\renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\usepackage[status=draft]{fixme}
-\fxsetup{theme=color}
-% \setlength{\textwidth}{12.4cm}
-\setlength{\textheight}{19.9cm}
 
@@
   Both challenges were purposely limited in scope, and their benchmark challenges do not address concurrency-related issues.
-  %
   We propose a new collection of benchmark challenges focused on the difficulties that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.
@@
 We have begun collecting solutions to our challenges on our website:
-%
 \begin{center}
@@
 \end{center}
-%
 In the future, we intend to use the website to promote best practices and tutorials derived from solutions to our challenges.
@@
 develop the infrastructure is significant.
-%\todo{Consider adding material on linearity predicates}
-%
 One alternative strategy is to bypass the problem of context splitting
@@
 (see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
-%
 
@@
 \cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997, Perera2018} and locally-nameless
-representation %~\cite{Chargueraud2012}
-(\eg~\cite{Castro2020}).
+representation (\eg~\cite{Castro2020}).
 Nominal approaches are also common (see~\cite{Bengtson2009}), but they
@@
 bisimilarity up-to~\cite{ChaudhuriCM15}.
-% While many proof assistants support
-% coinductive techniques,   but they support subtly different formalisms that enable different techniques, and some systems provide more than one abstraction to use coinduction.
-% We thus see a variety of different formalisation choices for
-% bisimulations, sometimes ad-hoc for the specific result.
-
-% \todo{Reorganize the rest of this section to be more coherent}
-% Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
-% Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
-% Newer entrants to the field like the Lean proof assistant
-% are working on adding a notion of coinduction to their system~\cite{Avigad2019}.\alb{I kinda object to ``seamlessly''.
-In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky.
-I guess Isabelle is better}
-% =======
-% Process calculi often include constructs that allow processes to have infinite behaviours.
-% Coinduction is a fundamental technique for defining and reasoning about infinite objects, and it can thus be used to reason about these behaviours.
-% While many proof assistants support coinductive techniques, using them is notoriously difficult, and representations and proof techniques vary widely.
-% The crux of our challenge is the effective use of coinductive up-to techniques.
-% The intention is that it should be easy to use the bisimulation relation and its properties, but this may be difficult with some mechanisation approaches.
-
-% The goal of our challenge on coinductive reasoning is proving that strong barbed bisimilarity can be turned into a congruence by making it sensitive to substitution and parallel composition.
-% The setting for our challenge is an untyped \(\pi\)-calculus with replication of processes enabling infinite behaviours.
-% We do not include delegation in the calculus since it is orthogonal to our primary aim of exploring coinductive proof techniques.
-% We base our definition of bisimilarity on a labelled transition system semantics and an observability predicate describing the communication steps available to a process.
-% The introduction of strong barbed bisimulation is one of the first steps when studying the observable behaviour of process calculi, both in classical textbooks (\eg \cite{picalcbook}) and in existing mechanisations.
-% Though weak barbed congruence is a more common behavioural equivalence, we prefer strong equivalences to simplify the theory by avoiding the need to abstract over the number of internal transitions in a trace.
-
-% Coinduction is widely supported in proof assistants, but they support subtly different formalisms that enable different techniques, and some systems provide more than one abstraction to use coinduction.
-% We thus see a variety of different formalisation choices for bisimulations, sometimes ad-hoc for the specific result.
-
-% \todo{Reorganize the rest of this section to be more coherent}
-% Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
-% Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
-% Newer entrants to the field like the Lean proof assistant
-% are working on adding a notion of coinduction to their system~\cite{Avigad2019}.\alb{I kinda object to ``seamlessly''.
-In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky.
-I guess Isabelle is better}
 
@@
 
-% - Going beyond the challenge problems
-%   - We do not aim to be comprehensive, but the techniques should also be useful for other things
-%   - Our challenges are independent, but they can also be combined
-%   - Our challenges are basic textbook theory, so they can form the basis for many extensions
-%   - Our aim is to see a future where ``the basics'' are as trivial in mechanisations as they are on paper
-
-% We previously promise to discuss combinations of techniques and applications to more elaborate schemes
-%\section{Going Beyond the Challenge Problems}\label{sec:going-beyond}
 \section{Future Work and Conclusions}\label{sec:going-beyond}
@@
 
-%%% Combinations
 Combining techniques for mechanising the fundamental aspects covered in our benchmark is often not trivial.
@@
 
-%%% Other aspects
 Much current research on concurrent calculi includes aspects that are not covered by our benchmark challenges.
@@
 
-%%% Conclusion
 Ultimately, the fundamental aspects covered by our benchmark serve as the building blocks for most current research on concurrenct calculi.

### 2024-02-23 challenges/linearity.tex
Minor formatting fixes in linearity proofs.

@@
   \item[\TirName {Par-Comm/Assoc}] By rearranging sub-derivations noting that
-    order does not matter for linear contexts
+    order does not matter for linear contexts.
   \item[\TirName{Par-Inact}]
-   Right-to-left by strengthening~\ref{lemma:strenD}. Vice-versa,
-    by picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
+   Right-to-left by Lemma~\ref{lemma:strenD}. Vice-versa,
+    by picking $T$ to be $\Tend$ and applying Lemma~\ref{lemma:weak}, part 2.
   \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear or $\Tend$ and applying weakening and strengthening accordingly.
-  \item[\TirName{Res-Inact:}] Use weakening
-  \item[\TirName{Res}] Noting that order does not matter
+  \item[\TirName{Res-Inact:}] By Lemma~\ref{lemma:weak}.
+  \item[\TirName{Res}] Noting that order does not matter.
   \end{description}
@@
       \item[\TirName {Sym}] By IH.
-      \item[\TirName {Trans}]  By two appeal to the IH
+      \item[\TirName {Trans}]  By two appeals to the IH.
       \item[\TirName {Cong}] By induction on the structure of $C$. If

### 2024-02-23 short-paper/short-paper.tex
Minor fixes

@@
 
-\keywords{Mechanisation \and Process calculi \and Benchmark \and Coinduction \and Scope extrusion \and Linearity}
+\keywords{Mechanisation \and Process calculi \and Benchmark \and Linearity \and Scope extrusion \and Coinduction}
 \end{abstract}
@@
 
-\subsection{Evaluation Criteria.}
+\subsection{Evaluation Criteria}
 The motivation behind our benchmark is to obtain evidence towards answering
@@
 
-Ultimately, the fundamental aspects covered by our benchmark serve as the building blocks for most current research on concurrenct calculi.
+Ultimately, the fundamental aspects covered by our benchmark serve as the building blocks for most current research on concurrent calculi.
 It is our hope and aim that exploring solutions to our challenges will move the community closer to a future where the key basic proof techniques for concurrent calculi are as easy to mechanise as they are to write on paper.

### 2024-02-23 short-paper/short-paper.tex
Many small edits for brevity.

@@
 In this spirit, we introduce a new collection of benchmark challenges
-specifically crafted to tackle common challenges encountered during
+crafted to tackle common challenges encountered during
 the mechanisation of formal models of concurrent and distributed
-programming languages.  We particularly focus on process calculi, as
+programming languages.  We focus on process calculi, as
 they provide a simple but realistic showcase of these challenges.  Concurrent calculi are
@@
 \end{center}
-In the future, we intend to use the website to promote best practices and tutorials derived from solutions to our challenges.
+We intend to use the website to promote best practices and tutorials derived from solutions to our challenges.
 We encourage readers to try the challenges using their favourite techniques, and to send us their solutions.
@@
 In this section, we outline the factors considered when designing the benchmark challenges.
-We begin with some general remarks, then describe the individual design considerations for each challenge, and the criteria for evaluating solutions thereof.
+We begin with some general remarks, then describe the individual design considerations for each challenge, and the criteria for evaluating solutions.
 
-Similarly to the authors of the POPLMark challenges, we seek to answer several questions:
+Similarly to the authors of POPLMark, we seek to answer several questions:
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
@@
 
-To achieve design goal \ref{item:goal-comperison-accessibility}, we have
-formulated our challenges to explore the three aspects
+To achieve design goal \ref{item:goal-comperison-accessibility}, our challenges explore the three aspects
 (linearity, scope extrusion, coinduction) independently, so that they
@@
 tangible insights.  The minimality and uniformity of the
-calculi also allows us to target design goal
+calculi allows us to target design goal
 \ref{item:goal-tutorials}.
-Aligned with design goal~\ref{item:goal-reusability}, our challenges are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
+Aligned with design goal~\ref{item:goal-reusability}, our challenges concern the fundamental meta-theory of process calculi.
 Our challenges centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in many applications (as we will further discuss in \cref{sec:going-beyond}).
@@
 concurrent send operations on both endpoints).
-The operational semantics is formulated as a standard reduction relation. Thus,
-proving subject reduction requires proving type preservation for structural
-congruence.
+The operational semantics is a standard reduction relation. Proving subject reduction thus requires proving type preservation for structural congruence.
 
@@
 establishing a large number of technical lemmata that are orthogonal to the
-problem under study (in our case, proving type safety for
-session types).  In this space, several designs are possible: one can label
+problem under study (in our case, proving type safety for session types).
+Several designs are possible: one can label
 occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}), or impose a multiset structure over lists
@@
 One alternative strategy is to bypass the problem of context splitting
-by adopting familiar ideas from algorithmic linear type checking. This
-approach, referred to as ``typing with leftovers,'' is exemplified
-in~\cite{DBLP:conf/forte/ZalakainD21}.    Similarly, one
-could eliminate context splitting by delegating linearity checks to a
-\emph{linear predicate} defined on the process structure (\eg~\cite{BP23}). These checks 
-serve as additional conditions within the typing rule. Whatever the choice,
-list-based encodings can be refined to be intrinsically-typed, if the
+by adopting ideas from algorithmic linear type checking. This
+approach, known as ``typing with leftovers,'' is exemplified
+in~\cite{DBLP:conf/forte/ZalakainD21}.
+Similarly, context splitting can be eliminated by delegating linearity checks to a
+\emph{linear predicate} defined on the process structure (\eg~\cite{BP23}).
+These checks serve as additional conditions within the typing rule.
+Whatever the choice, list-based encodings can be refined to be intrinsically-typed if the
 proof assistant supports dependent types
@@
 recursive processes bind recursion variables, and restrictions bind
-names. The first two act similarly to the binders known from
+names. The first two act similarly to the binders in
 $\lambda$-calculi, but restrictions may be more challenging due to
@@
 mechanised from an early age (see~\cite{Melham1994}) with many proof
-assistants and in so many encoding styles that it would take too long to
-review.
-Despite this, almost all of these mechanisations still rely on ad-hoc solutions to accommodate scope extrusion.
+assistants and in many encoding styles.
+Despite this, almost all of these mechanisations rely on ad-hoc solutions to encode scope extrusion.
 They range from concrete encodings based on named syntax
@@
 
-Higher-order abstract syntax (HOAS) has seen extensive and successful
+Higher-order abstract syntax (HOAS) has seen extensive
 use in formal reasoning in this area~\cite{Despeyroux2000,Honsell2001,Tiu2010,ChaudhuriCM15,Castro-Perez2021}. Its
@@
 the form of additional axioms~\cite{Honsell2001} or require specialised proof assistants
-like Abella, which introduces a dedicated quantifier for handling
+like Abella, which uses a special quantifier for handling
 properties related to names~\cite{GacekMN11}.
@@
 
-The goal of our challenge on coinductive reasoning is proving that
+The goal of our challenge on coinductive reasoning is to prove that
 \emph{strong barbed bisimilarity} can be turned into a congruence by making
@@
 process replication in order to enable infinite behaviours.  We do not
-include name passing since it is orthogonal to our
-primary aim of exploring coinductive proof techniques.  We base our
-definition of bisimilarity on a labelled transition system semantics
+include name passing since it is orthogonal to our  aim of exploring coinductive proof techniques.
+We base our definition of bisimilarity on a labelled transition system semantics
 and an observability predicate describing the communication steps
 available to a process.  The introduction of strong barbed
-bisimulation is one of the first steps when studying the observable
-behaviour of process calculi, both in classical textbooks (e.g.,
+bisimulation is one of the first steps when studying the
+behaviour of process calculi, both in textbooks (e.g.,
 \cite{picalcbook}) and in existing mechanisations.  Though weak barbed
@@
 Our benchmark challenges do not cover all issues in the field, but focus on the fundamental aspects of linearity, scope extrusion, and coinduction.
-Many mechanisations need to combine techniques to handle several of these aspects at once, and some may also need to handle aspects that are not covered by our benchmark challenges at all.
+Many mechanisations need to combine techniques to handle several of these aspects, and some may also need to handle aspects that are not covered by our benchmark.
 

### 2024-02-23 short-paper/short-paper.tex
Remove a todo that I think is done.

@@
 Ultimately, the fundamental aspects covered by our benchmark serve as the building blocks for most current research on concurrent calculi.
-It is our hope and aim that exploring solutions to our challenges will move the community closer to a future where the key basic proof techniques for concurrent calculi are as easy to mechanise as they are to write on paper.
+It is our hope and aim that exploring and comparing solutions to our challenges will move the community closer to a future where the key basic proof techniques for concurrent calculi are as easy to mechanise as they are to write on paper.
 
-\todo{A sentence to conclude reiterating the relevance of the challenges}
 \bibliographystyle{splncs04}

### 2024-02-23 short-paper/short-paper.tex
Improve abstract slightly.

@@
   POPLMark and POPLMark Reloaded sparked a flurry of work on machine-checked proofs, and fostered the adoption of proof mechanisation in programming language research.
-  Both challenges were purposely limited in scope, and their benchmark challenges do not address concurrency-related issues.
+  Both challenges were purposely limited in scope, and they do not address concurrency-related issues.
   We propose a new collection of benchmark challenges focused on the difficulties that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.
   Our benchmark challenges address three key topics: linearity, scope extrusion, and coinductive reasoning.
-  The goal of this new challenge is to clarify, compare, and advance the state of the art, fostering the adoption of proof mechanisation in future research on concurrency.
+  The goal of this new benchmark is to clarify, compare, and advance the state of the art, fostering the adoption of proof mechanisation in future research on concurrency.
 

### 2024-02-23 short-paper/short-paper.tex
some minor edits to the whole paper -- minus appendix

@@
 \begin{abstract}
-  POPLMark and POPLMark Reloaded sparked a flurry of work on machine-checked proofs, and fostered the adoption of proof mechanisation in programming language research.
-  Both challenges were purposely limited in scope, and they do not address concurrency-related issues.
-  We propose a new collection of benchmark challenges focused on the difficulties that typically arise when mechanising formal models of concurrent and distributed programming languages, such as process calculi.
-  Our benchmark challenges address three key topics: linearity, scope extrusion, and coinductive reasoning.
-  The goal of this new benchmark is to clarify, compare, and advance the state of the art, fostering the adoption of proof mechanisation in future research on concurrency.
+  POPLMark and POPLMark Reloaded sparked a flurry of work on
+  machine-checked proofs, and fostered the adoption of proof
+  mechanisation in programming language research.  Both challenges
+  were purposely limited in scope, and they do not address
+  concurrency-related issues.  We propose a new collection of
+  benchmark challenges focused on the difficulties that typically
+  arise when mechanising formal models of concurrent and distributed
+  programming languages, such as process calculi.  Our benchmark
+  challenges address three key topics: linearity, scope extrusion, and
+  coinductive reasoning.  The goal of this new benchmark is to
+  clarify, compare, and advance the state of the art, fostering the
+  adoption of proof mechanisation in future research on concurrency.
 
@@
 played a pivotal role in advancing the field of proof assistants,
-libraries, and best practices for programming language
-mechanisation. By providing a shared framework for systematically
+libraries, and best practices for the mechanisation of programming language research. By providing a shared framework for systematically
 evaluating mechanisation techniques, it catalysed a significant
@@
 POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a similar programme, targeting proofs using logical relations.
-These initiatives had a narrow focus, but their authors recognised the importance of addressing topics such as coinduction and linearity in the future.
+These initiatives had a narrow focus, and their authors recognised the importance of addressing topics such as coinduction and linearity in the future.
 
-In this spirit, we introduce a new collection of benchmark challenges
+In this spirit, we introduce a new collection of benchmarks 
 crafted to tackle common challenges encountered during
@@
 Our benchmark challenges (detailed in \cref{app:challenges} and on our website) consider
-\emph{in isolation} three key aspects that frequently pose challenges
+\emph{in isolation} three key aspects that frequently pose difficulties
 when mechanising concurrency theory, which we will discuss in more detail in the next section: \emph{linearity}, \emph{scope
@@
 
-To achieve design goal \ref{item:goal-comperison-accessibility}, our challenges explore the three aspects
+To achieve design goal~\ref{item:goal-comperison-accessibility}, our challenges explore the three aspects
 (linearity, scope extrusion, coinduction) independently, so that they
@@
 tangible insights.  The minimality and uniformity of the
-calculi allows us to target design goal
-\ref{item:goal-tutorials}.
+calculi allows us to target design goal~\ref{item:goal-tutorials}.
 Aligned with design goal~\ref{item:goal-reusability}, our challenges concern the fundamental meta-theory of process calculi.
@@
 the topology of the communication network described by a process cannot change.
-We do not allow recursion or replication, so infinite behaviours cannot be
+We do not allow recursion or replication, hence infinite behaviours cannot be
 expressed. We also forgo more sophisticated notions of error-freedom (\eg
@@
 In mechanised meta-theory, addressing linearity means choosing an appropriate
-representation of the linear typing context.  While a linear context is perhaps
+representation of a linear context.  While the latter  is perhaps
 best seen as a multiset, most proof assistants have better support for list
-processing.  The latter representation is intuitive, but may require
+processing.  This representation is intuitive, but may require
 establishing a large number of technical lemmata that are orthogonal to the
@@
 (\eg~\cite{Danielsson12,ChaudhuriLR19}). Alternatively, contexts can be
-implemented as finite maps (as in \cite{Castro2020}), whose operations are
+implemented as finite maps (as in~\cite{Castro2020}), whose operations are
 sensitive to a linear discipline. In all these cases, the effort required to
@@
 One alternative strategy is to bypass the problem of context splitting
-by adopting ideas from algorithmic linear type checking. This
+by adopting ideas from algorithmic linear type checking. One such
 approach, known as ``typing with leftovers,'' is exemplified
@@
 \emph{linear predicate} defined on the process structure (\eg~\cite{BP23}).
-These checks serve as additional conditions within the typing rule.
+These checks serve as additional conditions within the typing rules.
 Whatever the choice, list-based encodings can be refined to be intrinsically-typed if the
@@
   extrusion by using instead structural congruence, allowing
-  implementors to explore alternative ways to encode the latter (\eg
+  implementors to explore different ways to encode the latter (\eg
   via process contexts or via compatible refinement);
@@
 
-Given those peculiarities, the syntax and semantics of $\pi$-calculi have been 
-mechanised from an early age (see~\cite{Melham1994}) with many proof
-assistants and in many encoding styles.
-Despite this, almost all of these mechanisations rely on ad-hoc solutions to encode scope extrusion.
-They range from concrete encodings based on named syntax
-\cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997, Perera2018} and locally-nameless
-representation (\eg~\cite{Castro2020}).
+Given those peculiarities, the syntax and semantics of $\pi$-calculi
+have been mechanised from an early age (see~\cite{Melham1994}) with
+many proof assistants and in many encoding styles.  Despite this,
+almost all of these mechanisations rely on ad-hoc solutions to encode
+scope extrusion.  They range from concrete encodings based on named
+syntax~\cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997,
+  Perera2018} and locally-nameless representation~\cite{Castro2020}.
 Nominal approaches are also common (see~\cite{Bengtson2009}), but they
-may be problematic in proof assistants based on dependent types.
-The case study \cite{AmbalLS21} compares four approaches to encoding
-binders in Coq as a first step towards developing tools for working
-with higher-order process calculi.  They found that working directly
-with de~Bruijn indices was easiest since the existing
-approaches developed for $\lambda$-calculus binders worked poorly 
-with scope extrusion.
+may be problematic in proof assistants based on constructive type
+theories.  An overall comparison is still lacking, but the case
+study~\cite{AmbalLS21} develops four approaches to encoding binders in
+Coq % as a first step towards developing tools for working with
+in the context of higher-order process calculi.  The authors report that working directly
+with de~Bruijn indices was easiest since the existing approaches
+developed for $\lambda$-calculus binders worked poorly with scope
+extrusion.
 
@@
 bisimulation, extensions to HOAS are needed. These extensions may take
-the form of additional axioms~\cite{Honsell2001} or require specialised proof assistants
-like Abella, which uses a special quantifier for handling
+the form of additional axioms~\cite{Honsell2001} or require niche proof assistants
+such as Abella, which features a special quantifier for handling
 properties related to names~\cite{GacekMN11}.
@@
 The goal of our challenge on coinductive reasoning is to prove that
-\emph{strong barbed bisimilarity} can be turned into a congruence by making
-it sensitive to substitution and parallel composition. The crux of our
-challenge is the effective use of coinductive up-to techniques.  The
-intention is that it should be easy to use the bisimulation relation
-and its properties, but this may be difficult with some mechanisation
-approaches.
+\emph{strong barbed bisimilarity} can be turned into a congruence by
+making it sensitive to substitution and parallel composition. The crux
+of our challenge is the effective use of coinductive up-to techniques.
+The intention is that the result should be relatively easy to
+achieve once the main properties of bisimilarity are established.
+% should be easy to use the bisimulation relation and its properties,
+% but this may be difficult with some mechanisation approaches.
 
-The setting for our challenge is an untyped \(\pi\)-calculus augmented with
-process replication in order to enable infinite behaviours.  We do not
-include name passing since it is orthogonal to our  aim of exploring coinductive proof techniques.
-We base our definition of bisimilarity on a labelled transition system semantics
-and an observability predicate describing the communication steps
-available to a process.  The introduction of strong barbed
-bisimulation is one of the first steps when studying the
-behaviour of process calculi, both in textbooks (e.g.,
-\cite{picalcbook}) and in existing mechanisations.  Though weak barbed
-congruence is a more common behavioural equivalence, we prefer strong
-equivalences to simplify the theory by avoiding the need to abstract
-over the number of internal transitions in a trace.
+The setting for our challenge is an untyped \(\pi\)-calculus augmented
+with process replication in order to enable infinite behaviours.  We
+do not include name passing since it is orthogonal to our aim of
+exploring coinductive proof techniques.  We base our definition of
+bisimilarity on a labelled transition system semantics and an
+observability predicate describing the communication steps available
+to a process.  The description of strong barbed bisimulation is one of
+the first steps when studying the behaviour of process calculi, both
+in textbooks (e.g.,~\cite{picalcbook}) and in existing mechanisations.
+Though weak barbed congruence is a more common behavioural
+equivalence, we prefer strong equivalences to simplify the theory by
+avoiding the need to abstract over the number of internal transitions
+in a trace.
 
 While many proof assistants support coinductive techniques, they do so
-through subtly different formalisms, enabling various approaches.
+through different foundations and formalisms. %, enabling various approaches.
 Additionally, some systems offer multiple abstractions for utilising coinduction.
 For instance, Agda offers musical notation, co-patterns and sized types~\cite{Abel2013};
-Coq offers guarded recursion and refined fixed point approaches as supported by the parameterised coinduction~\cite{Hur2013}
+Coq features guarded recursion and refined fixed point approaches as supported by the parameterised coinduction~\cite{Hur2013}
 and interaction tree~\cite{Xia2019} libraries.
@@
 When reasoning over bisimilarity many authors rely on the native
- coinduction offered by the chosen proof assistants
-\cite{Bengtson2016,Kahsai2008,Thiemann2019,Gay2020}, while others prefer a
-more ``set-theoretic'' approach
-\cite{Hirschkoff1997,Bengtson2009,Maksimovic2015,Pohjola2022}. Some
+coinduction offered by the chosen proof
+assistant~\cite{Bengtson2016,Kahsai2008,Thiemann2019,Gay2020}, while
+others prefer a more ``set-theoretic'' approach~\cite{Hirschkoff1997,Bengtson2009,Maksimovic2015,Pohjola2022}. Some
 use both and establish an internal adequacy~\cite{Honsell2001}. Few
@@
 
-Much current research on concurrent calculi includes aspects that are not covered by our benchmark challenges.
-Features that are in wide use, but not covered by our benchmark, include choice constructs and recursion.
-Some interesting research topics that build on the fundamental aspects in our benchmark include multiparty session types~\cite{10.1145/2827695}, choreographies~\cite{Carbone2013}, higher-order calculi~\cite{Hirsch2022}, conversation types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696}, and encodings between different calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:journals/iandc/DardhaGS17}.
-Their meta-theory includes aspects---\eg well-formedness conditions on global types, partiality of end-point projection functions, \etc---that we do not address.
+Much current research on concurrent calculi includes aspects that are
+not covered by our benchmark challenges, for example constructs such as
+choice constructs and recursion.
+%  Features that are in wide
+% use, but not covered by our benchmark, include choice constructs and
+% recursion. 
+Some interesting research topics that build on the
+fundamental aspects in our benchmark include multiparty session
+types~\cite{10.1145/2827695}, choreographies~\cite{Carbone2013},
+higher-order calculi~\cite{Hirsch2022}, conversation
+types~\cite{DBLP:journals/tcs/CairesV10}, psi-calculi~\cite{lmcs:696},
+and encodings between different
+calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:journals/iandc/DardhaGS17}.
+Their meta-theory includes aspects---\eg well-formedness conditions on
+global types, partiality of end-point projection functions,
+\etc---that we do not address.
 
-Our coinduction challenge only includes two notions of process equivalence, but many more exist in the literature.
+Our coinduction challenge only addresses two notions of process equivalence, but many more exist in the literature.
 Coinduction may also play a role in recursive processes and session types: recursive session types can be expressed in \emph{infinitary form} by interpreting their typing rules coinductively~\cite{DerakhshanPfenning22,HornePadovani23}.
 
-Unlike POPLMark, we consider \emph{animation} of calculi (as in \cite{Castro-Perez2021}) out of scope for our benchmark.
+Unlike POPLMark, we consider \emph{animation} of calculi (as in~\cite{Castro-Perez2021}) out of scope for our benchmark.
 Finally, our challenges encourage, but do not require, exploring proof automation, as offered by \eg the \emph{Hammer} protocol~\cite{BohmeN10,CzajkaK18}.

### 2024-02-23 short-paper/short-paper.tex
Minor edits and remove a comment before initial submit.

@@
 shift towards publications that include mechanised proofs within the programming language research community.
-POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a similar programme, targeting proofs using logical relations.
+POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a similar programme for proofs using logical relations.
 These initiatives had a narrow focus, and their authors recognised the importance of addressing topics such as coinduction and linearity in the future.
@@
 theories.  An overall comparison is still lacking, but the case
-study~\cite{AmbalLS21} develops four approaches to encoding binders in
-Coq % as a first step towards developing tools for working with
-in the context of higher-order process calculi.  The authors report that working directly
-with de~Bruijn indices was easiest since the existing approaches
+study~\cite{AmbalLS21} explores four approaches to encoding binders in
+Coq in the context of higher-order process calculi.  The authors report that working directly
+with de~Bruijn indices was easiest since the approaches
 developed for $\lambda$-calculus binders worked poorly with scope
@@
 the first steps when studying the behaviour of process calculi, both
-in textbooks (e.g.,~\cite{picalcbook}) and in existing mechanisations.
+in textbooks (\eg~\cite{picalcbook}) and in existing mechanisations.
 Though weak barbed congruence is a more common behavioural
@@
 While we have chosen to focus on the aspects individually to simplify the challenges, we are of course also interested in exploring how techniques interact.
-Some authors have already described the difficulties that combining techniques may induce in a mechanisation, \eg~\cite{DBLP:conf/forte/ZalakainD21}.\todo{Add more references here?}
+%Some authors have already described the difficulties that combining techniques may induce in a mechanisation, \eg~\cite{DBLP:conf/forte/ZalakainD21}.\todo{Add more references here?}
 

### 2024-02-26 challenges/coinduction.tex
A few minor corrections.

@@
 
-This challenge is  about the mechanisation of proofs concerning processes wih infinite behaviours.
+This challenge is about the mechanisation of proofs concerning processes with infinite behaviours.
 This is usually connected to \emph{coinductive} definitions where an infinite
@@
 
-\subsubsection{Semantics.} We choose to give an LTS semantics for this
+\subsubsection{Semantics.} We choose to give a labelled transition system semantics for this
 challenge.

### 2024-02-26 short-paper/short-paper.tex
Small edits for brevity.

@@
 (linearity, scope extrusion, coinduction) independently, so that they
-may be solved individually and in any order; each challenge is
-reasonably small and easily understandable with basic knowledge of
+may be solved individually and in any order; each challenge is small and easily understandable with basic knowledge of
 textbook concurrency theory, process calculi, and type theory.  The
 process calculi used in the challenges focus on the features that
-we want to emphasise, and omit all constructs that would complicate the mechanisation without bringing additional
+we want to emphasise, and omit all constructs that would complicate the mechanisation without bringing
 tangible insights.  The minimality and uniformity of the
@@
 Aligned with design goal~\ref{item:goal-reusability}, our challenges concern the fundamental meta-theory of process calculi.
-Our challenges centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in many applications (as we will further discuss in \cref{sec:going-beyond}).
+Our challenges centre around well-established results, showcasing proof techniques that can be leveraged in many applications (as we will further discuss in \cref{sec:going-beyond}).
 
@@
 \item A reduction system: this avoids explicit reasoning about scope
-  extrusion by using instead structural congruence, allowing
+  extrusion by using structural congruence, allowing
   implementors to explore different ways to encode the latter (\eg
-  via process contexts or via compatible refinement);
+  via process contexts or compatible refinement);
 \item An (early) labelled transition system.
@@
 While many proof assistants support coinductive techniques, they do so
-through different foundations and formalisms. %, enabling various approaches.
-Additionally, some systems offer multiple abstractions for utilising coinduction.
+through different formalisms.
+Some systems even offer multiple abstractions for utilising coinduction.
 For instance, Agda offers musical notation, co-patterns and sized types~\cite{Abel2013};
-Coq features guarded recursion and refined fixed point approaches as supported by the parameterised coinduction~\cite{Hur2013}
+Coq features guarded recursion and refined fixed point approaches via the parameterised coinduction~\cite{Hur2013}
 and interaction tree~\cite{Xia2019} libraries.

### 2024-02-27 short-paper/short-paper.tex
Final fixes.

@@
 \input{../macros.tex}
+
+\crefname{question}{question}{questions}
 
@@
 polarities.
-Similarly, other key results in session types papers have subsequently
+Similarly, other key results in papers on session types have subsequently
 been proven incorrect~\cite{Gay2020,10.1145/3290343}, demonstrating the
@@
 We intend to use the website to promote best practices and tutorials derived from solutions to our challenges.
-We encourage readers to try the challenges using their favourite techniques, and to send us their solutions.
+We encourage readers to try the challenges using their favourite techniques, and to send us their solutions and experience reports.
 
@@
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
-\item\label{item:rq1} What is the current state of the art in the mechanisation of  the meta-theory of process calculi?
-\item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
-\item\label{item:rq3} What improvements are needed to make mechanisation tools more user-friendly with regard to the issues faced when mechanising process calculi?
+\item\label[question]{item:rq1} What is the current state of the art in the mechanisation of  the meta-theory of process calculi?
+\item\label[question]{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
+\item\label[question]{item:rq3} What improvements are needed to make mechanisation tools more user-friendly with regard to the issues faced when mechanising process calculi?
 \end{enumerate}
@@
 tangible insights.  The minimality and uniformity of the
-calculi allows us to target design goal~\ref{item:goal-tutorials}.
+calculi also allows us to target design goal~\ref{item:goal-tutorials}.
 Aligned with design goal~\ref{item:goal-reusability}, our challenges concern the fundamental meta-theory of process calculi.
@@
 opposed to shared) channels. Inspired by Vasconcelos~\cite{Vasconcelos2012}, we
-define a syntax where a restriction $(\nu ab)$ binds two dual names $a$
+define a syntax where a restriction $(\bm{\nu} ab)$ binds two dual names $a$
 and $b$ as opposite endpoints of the same channel; their duality is
-reflected in the type system. We model a simple notion of error: a well-typed
+reflected in the type system. We model a simple notion of error: well-typed
 processes must never use dual channel endpoints in a non-dual way (\eg by
 performing concurrent send/receive operations on the same endpoint, or two
-concurrent send operations on both endpoints).
+concurrent send operations on dual endpoints).
 The operational semantics is a standard reduction relation. Proving subject reduction thus requires proving type preservation for structural congruence.
@@
 representation of a linear context.  While the latter  is perhaps
-best seen as a multiset, most proof assistants have better support for list
-processing.  This representation is intuitive, but may require
+best seen as a multiset, most proof assistants have better support for lists.
+This representation is intuitive, but may require
 establishing a large number of technical lemmata that are orthogonal to the
@@
 \end{enumerate}
-The objective of this
-challenge is to prove that the two semantics are equivalent up to
-structural congruence.
+The goal of our challenge on scope extrusion is to prove that the two semantics are
+equivalent up to structural congruence.
 
@@
 The motivation behind our benchmark is to obtain evidence towards answering
-questions \cref{item:rq1,item:rq2,item:rq3}. We are therefore
+\cref{item:rq1,item:rq2,item:rq3}. We are therefore
 interested not only in the solutions, but also in the experience of
@@
 \end{enumerate}
-Solutions to our challenges need not strictly follow the definitions and lemmata set out in the challenge text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.
+Solutions to our challenges need not strictly follow the definitions and lemmata set out in the challenge text, but solutions which deviate from the original challenges should present more elaborate argumentation for their adequacy.
 
@@
 
-Combining techniques for mechanising the fundamental aspects covered in our benchmark is often not trivial.
-While we have chosen to focus on the aspects individually to simplify the challenges, we are of course also interested in exploring how techniques interact.
+Combining techniques for mechanising the fundamental aspects covered in our benchmark is non-trivial.
+While we focus on the aspects individually to simplify the challenges, we are also interested in exploring how techniques interact.
 %Some authors have already described the difficulties that combining techniques may induce in a mechanisation, \eg~\cite{DBLP:conf/forte/ZalakainD21}.\todo{Add more references here?}
@@
 calculi~\cite{DBLP:journals/iandc/Gorla10,DBLP:journals/iandc/DardhaGS17}.
-Their meta-theory includes aspects---\eg well-formedness conditions on
+The meta-theory of these topics includes aspects---\eg well-formedness conditions on
 global types, partiality of end-point projection functions,

### 2024-04-09 challenges/linearity.tex
some minor edits to the whole paper -- minus appendix
revisions following review 3 -- some commnets on still TODO in rew3.txt

@@
   \and
-  \inferrule[T-Res]{\types{\Gamma; (\Cadd{\Cadd{\Delta}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}}}{P})}{\types{\Gamma;\Delta}{\Presd{x}{y}{P}}}
+  \inferrule[T-Res]{\types{\Gamma; (\Cadd{\Cadd{\Delta}{\hastype{x}{T}}}{\hastype{y}{\dual{T}}})}{P}}{\types{\Gamma;\Delta}{\Presd{x}{y}{P}}}
   \and
@@
   \label{lemma:strenD}
-  If \(\types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P}\) and
-  $x\not\in \freenames{P}$, then \(\types{\Gamma; \Delta}{P}\).
+  \begin{enumerate}
+  \item If $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$, then
+    $\typev{\Gamma}{\hastype{v}{\Tbase}}$.
+  \item   If \(\types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P}\) and
+  $x\not\in \freenames{P}$, then \(\types{\Gamma; \Delta}{P}\)
+\end{enumerate}
 \end{lemma}
 \begin{proof}
-  By induction on the derivation of
+  \mbox{}
+  \begin{enumerate}
+  \item By immediate case analysis on the given derivation.
+  \item By induction on the derivation of
   $\types{\Gamma; (\Cadd{\Delta}{\hastype{x}{T}})}{P}$.
@@
   The remaining cases are analogous.
+\end{enumerate}
 \end{proof}
@@
 
-  \item[\TirName{T-Res}] Immediate on the premise of the rule.
+  \item[\TirName{T-Res}] Immediate by the induction hypothesis.
 
   \item[\TirName{T-Out}] Let $P = \Pout{x}{v}{P'}$. We have that
-    $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.  We must show
-    $\typev{\Gamma}{\hastype{v}{\Tbase}}$ to build the conclusion with
-    the induction hypothesis and \TirName{T-Out}. We proceed by cases
-    on the structure of $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.
+    $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.  We have 
+    $\typev{\Gamma}{\hastype{v}{\Tbase}}$ bu streghmening, so we  build the conclusion with
+    the induction hypothesis and \TirName{T-Out}.
+%    on the structure of $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.
   \end{description}
@@
 
-To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms.
+To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms of structural congruence.
 

### 2024-04-09 challenges/preliminaries.tex
revisions following review 3 -- some commnets on still TODO in rew3.txt

@@
 We assume the existence of some set of \emph{base values}, represented by the symbols
-\( a, b, \dots \), the existence of some set
+\( a, b, \dots \),  of some set
 of \emph{variables}, represented by the symbols
-\( l, m, \dots \), and the existence of some set of \emph{names},
+\( l, m, \dots \), and  of some set of \emph{names},
 represented by the symbols \( x, y, \dots \).%
@@
 compared for equality.
-We assume that all of these sets are infinite and that their elements can be compared for equality.
 

### 2024-04-09 short-paper/short-paper.tex
revisions following review 3 -- some commnets on still TODO in rew3.txt

@@
 
+% AM <-- review
 While many proof assistants support coinductive techniques, they do so
@@
 Some systems even offer multiple abstractions for utilising coinduction.
-For instance, Agda offers musical notation, co-patterns and sized types~\cite{Abel2013};
+For instance, Agda offers musical notation, co-patterns and productivity checking via sized types~\cite{Abel2013};
 Coq features guarded recursion and refined fixed point approaches via the parameterised coinduction~\cite{Hur2013}

### 2024-04-10 challenges/linearity.tex
Minor spelling fixes.
fixing the conguence thing in c2

@@
 As usual, we use  a \emph{structural congruence} relation that equates processes that we deem to be indistinguishable.
-Structural congruence is the smallest congruence relation that satisfies the following axioms:
+Structural congruence is the smallest compatible equivalence relation that satisfies the following axioms:
 \begin{mathpar}
@@
   \and
-  \inferrule[Sc-Res-Par]{\{x,y\} \cap \freenames{Q} = \emptyset}{\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
+  \inferrule[Sc-Res-Par]{ }%\{x,y\} \cap \freenames{Q} = \emptyset
+  {\scong{\Ppar{\Presd{x}{y}{P}}{Q}}{\Presd{x}{y}{(\Ppar{P}{Q})}}}
   \and
@@
 Typing contexts gather type information about names and variables.
-\emph{Unrestricted} contexts are simply sets of names since we only have one
-base type. \emph{Linear contexts} associate a type to endpoints. We use
-the comma as split/union, overloaded to singletons, and \( \Cempty \) as the
-empty context. We extend the Barendregt convention~\cite{Barendregt1984} to contexts, so that all
-entries are distinct.  Note that the order in which information is added to a
-type context does not matter.
+\emph{Unrestricted} contexts are simply sets of names since we only
+have one base type. \emph{Linear contexts} associate a type to
+endpoints. We use the comma as split/union, overloaded to singletons,
+and \( \Cempty \) as the empty context. We extend the Barendregt
+convention~\cite{Barendregt1984} to contexts, so that all entries are
+distinct.  Note that the order in which information is added to a type
+context does not matter. We denote with $\tend\Delta$ a (linear)
+context whose names have all type $\Tend$.
 
@@
     $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.  We have 
-    $\typev{\Gamma}{\hastype{v}{\Tbase}}$ bu streghmening, so we  build the conclusion with
+    $\typev{\Gamma}{\hastype{v}{\Tbase}}$ by strengthening, so we build the conclusion with
     the induction hypothesis and \TirName{T-Out}.
-%    on the structure of $\typev{(\Cadd{\Gamma}{l})}{\hastype{v}{\Tbase}}$.
   \end{description}
   The remaining cases are analogous.
-
 \end{proof}
 
-To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms of structural congruence.
+To prove that process equivalence preserves typing, it is convenient to first consider 
+ the smalleset relation closed under the six axioms of
+ structural congruence, denoted by $\sconga\cdot\cdot$:
 
@@
 
-Now, following Sangiorgi and Walker~\cite{picalcbook}, we give rules for the compatible
+Now, following Sangiorgi and Walker~\cite{picalcbook}, we  formalize the compatible
 equivalence relation induced by $\sconga\cdot\cdot$, which we still
-write as $\scong\cdot\cdot$: namely, add to reflexivity, symmetry and
-transitivity the following condition:
+write as $\scong\cdot\cdot$ as the smallest relation closed under reflexivity, symmetry, 
+transitivity and the following condition:
  \begin{mathpar}

### 2024-04-10 challenges/preliminaries.tex
fixing the conguence thing in c2

@@
 
-We say that an equivalence relation \( \mathcal{S} \) is a \emph{congruence} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).
+We say that an equivalence relation \( \mathcal{S} \) is  \emph{compatible} if \( (P,Q) \in \mathcal{S} \) implies that for any context \( C \), \( (\applyctx{C}{P}, \applyctx{C}{Q}) \in \mathcal{S} \).

### 2024-04-10 challenges/scope-extrusion.tex
Remove old comments.
fixing the conguence thing in c2

@@
 
-%The process \( \Pend \) is \emph{inaction}: a process which can do nothing.
 The process \( \Pout{x}{y}{P} \) is an \emph{output}, which can send the name \( y \) via \( x \), then continue as \( P \).
@@
 The input operator thus binds the name \( y \) in \( P \).
-% The process \( \Ppar{P}{Q} \) is the \emph{composition} of process \( P \) and process \( Q \).
-% The two components can proceed independently of each other, or they can interact via shared names.
-% The process \( \Pres{x}{P} \) is the \emph{restriction} of the name \( x \) to \( P \).
-% Components in \( P \) can use the name \( x \) to interact with each other, but not with processes outside of the restriction.
-% The restriction operator thus binds the name \( x \) in \( P \).
 Note that the scope of a restriction may change when processes interact. Namely, a restricted name may be sent \emph{outside} of its scope.
-% The process \( \Pchoice{P}{Q} \) is a non-deterministic \emph{choice} between continuing as the process \( P \) or as the process \( Q \).
 Note that there is no recursion or replication in the syntax, and thus no infinite behaviours can be expressed.
 This simplifies the theory and is orthogonal to the concept of scope extrusion.
-
-% We will use the notation \( \freenames{P} \) to denote the set of names that occur free (i.e.\ not bound by a restriction or an input) in \( P \).
-% We will use the notation \( \boundnames{P} \) to denote the set of names that occur bound (by a restriction or an input) in \( P \).
-% We will use the notation \( \subst{P}{x}{y} \) to denote the process \( P \) with \( x \) substituted for \( y \).
-
-% Two processes \( P \) and \( Q \) are \( \alpha \)-convertible, written \( \alphacon{P}{Q} \), if \( Q \) can be obtained from \( P \) by a finite number of substitutions of bound names.
-% As a convention, we will identify \( \alpha \)-convertible processes.
-
-% As a convention, we assume that the bound names occurring in any collection of processes are chosen to be different from the free names occurring in those processes and from the names occurring in any substitutions applied to the processes.
-% This is justified because any overlapping names may be \( \alpha \)-converted such that the assumption is satisfied.
 
@@
   \and
-  \inferrule[Sc-Res-Par]{x \notin \freenames{Q}}{\scong{\Ppar{\Pres{x}{P}}{Q}}{\Pres{x}{(\Ppar{P}{Q})}}}
+  \inferrule[Sc-Res-Par]{ }{\scong{\Ppar{\Pres{x}{P}}{Q}}{\Pres{x}{(\Ppar{P}{Q})}}}
+  % x \notin \freenames{Q}}
   \and
@@
 \subsubsection{Challenge.}
+As in the linearity challence,  we first consider 
+ the smalleset relation closed under the six axioms of
+ structural congruence, denoted by $\sconga\cdot\cdot$:
+\begin{lemma}\label{se-lemma-harmony-facta}
+  If \( \sconga{P}{Q} \) and \( \transition{P}{\alpha}{P'} \), then for some \( Q' \) we have \( \transition{Q}{\alpha}{Q'} \) and \( \sconga{P'}{Q'} \).
+\end{lemma}
+\begin{proof}[Sketch]
+By case analysis on the first derivation.
+\end{proof}
 \begin{lemma}\label{se-lemma-harmony-fact}
@@
 \begin{proof}[Sketch]
-  First, show the special case when \( P \) can be rewritten to \( Q \) with a single application of an axiom of the structural congruence to some subterm of P.
-  The proof is then by induction on the number of such steps.
+    By induction on the structure of the derivation of \( \scong{P}{Q} \), with an inner induction of the structure of a process context and an appeal to Lemma~\ref{se-lemma-harmony-facta}.
 \end{proof}
+
+% \begin{lemma}\label{se-lemma-harmony-fact}
+%   If \( \scong{P}{Q} \) and \( \transition{P}{\alpha}{P'} \), then for some \( Q' \) we have \( \transition{Q}{\alpha}{Q'} \) and \( \scong{P'}{Q'} \).
+% \end{lemma}
+% \begin{proof}[Sketch]
+%   First, show the special case when \( P \) can be rewritten to \( Q \) with a single application of an axiom of the structural congruence to some subterm of P.
+%   The proof is then by induction on the number of such steps.
+% \end{proof}
 For the proof of the challenge theorem, we introduce the notion of a \emph{normalized derivation} of a reduction \( \reduces{P}{Q} \), which is of the following form.

### 2024-04-10 short-paper/short-paper.tex
fixing the conguence thing in c2

--- 2024-02-27 short-paper/short-paper.tex
@@
 Some systems even offer multiple abstractions for utilising coinduction.
-For instance, Agda offers musical notation, co-patterns and sized types~\cite{Abel2013};
-Coq features guarded recursion and refined fixed point approaches via the parameterised coinduction~\cite{Hur2013}
-and interaction tree~\cite{Xia2019} libraries.
+For instance, Agda offers musical notation, co-patterns and productivity checking via sized types~\cite{Abel2013};
+Coq features guarded recursion and refined fixed point approaches via libraries such as the parameterised coinduction~\cite{Hur2013}, coinduction up-to~\cite{Pous16}
+and interaction trees~\cite{Xia2019} .
 

### 2024-04-10 challenges/linearity.tex
Fix minor textual issues.

@@
 context does not matter. We denote with $\tend\Delta$ a (linear)
-context whose names have all type $\Tend$.
+context whose names all have type $\Tend$.
 
@@
 To prove that process equivalence preserves typing, it is convenient to first consider 
- the smalleset relation closed under the six axioms of
+ the smallest relation closed under the six axioms of
  structural congruence, denoted by $\sconga\cdot\cdot$:

### 2024-04-10 challenges/scope-extrusion.tex
Fix minor textual issues.

@@
 \begin{proof}[Sketch]
-    By induction on the structure of the derivation of \( \scong{P}{Q} \), with an inner induction of the structure of a process context and an appeal to Lemma~\ref{se-lemma-harmony-facta}.
+    By induction on the structure of the derivation of \( \scong{P}{Q} \), with an inner induction on the structure of a process context and an appeal to Lemma~\ref{se-lemma-harmony-facta}.
 \end{proof}

### 2024-04-10 short-paper/short-paper.tex
Fix minor textual issues.

--- 2024-02-27 short-paper/short-paper.tex
@@
 Some systems even offer multiple abstractions for utilising coinduction.
-For instance, Agda offers musical notation, co-patterns and sized types~\cite{Abel2013};
-Coq features guarded recursion and refined fixed point approaches via the parameterised coinduction~\cite{Hur2013}
-and interaction tree~\cite{Xia2019} libraries.
+For instance, Agda offers musical notation, co-patterns and productivity checking via sized types~\cite{Abel2013};
+Coq features guarded recursion and refined fixed point approaches via libraries for \eg parameterised coinduction~\cite{Hur2013}, coinduction up-to~\cite{Pous16} and interaction trees~\cite{Xia2019}.
 

### 2024-04-15 short-paper/short-paper.tex
change
ack added
Add a full stop to ack.
Rework research questions and design goals to be more concise..

@@
 
-\title{The Concurrent Calculi Formalisation Benchmark}
+\title{The Concurrent Calculi Formalisation Benchmark\thanks{The work is partially supported by EPSRC EP/T006544/2.}}
 
@@
 \and Frederik Krogsdal Jacobsen \inst{5}\orcidID{0000-0003-3651-8314}
-\and Alberto Momigliano \inst{6}\orcidID{0000-0003-0942-4777}
+\and Alberto Momigliano\thanks{Member of the Gruppo Nazionale Calcolo Scientifico-Istituto Nazionale di Alta Matematica (GNCS-INdAM).} \inst{6}\orcidID{0000-0003-0942-4777}
 \and Luca Padovani \inst{7}\orcidID{0000-0001-9097-1297}
@@
 
-Similarly to the authors of POPLMark, we seek to answer several questions:
+Similarly to the authors of POPLMark, we seek to answer several questions about the mechanisation of the meta-theory of process calculi:
 \begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
-\item\label[question]{item:rq1} What is the current state of the art in the mechanisation of  the meta-theory of process calculi?
-\item\label[question]{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
-\item\label[question]{item:rq3} What improvements are needed to make mechanisation tools more user-friendly with regard to the issues faced when mechanising process calculi?
+\item\label[question]{item:rq1} What is the current state of the art?
+\item\label[question]{item:rq2} Which techniques and best practices can be recommended?
+\item\label[question]{item:rq3} What improvements are needed to make mechanisation tools more user-friendly?
 \end{enumerate}
@@
 
-\item\label{item:goal-reusability} To prioritise the exploration of mechanisation
-  techniques that are reusable for future research.
+\item\label{item:goal-reusability} To prioritise the exploration of reusable mechanisation techniques.
 \end{enumerate}

### 2024-04-15 short-paper/short-paper.tex
Add some more explanation of the target audiences.

@@
 may be solved individually and in any order; each challenge is small and easily understandable with basic knowledge of
-textbook concurrency theory, process calculi, and type theory.  The
-process calculi used in the challenges focus on the features that
-we want to emphasise, and omit all constructs that would complicate the mechanisation without bringing
-tangible insights.  The minimality and uniformity of the
-calculi also allows us to target design goal~\ref{item:goal-tutorials}.
+textbook concurrency theory, process calculi, and type theory.
+For mechanisation experts, our challenges should thus be accessible even without any prior understanding of process calculi.
+The process calculi used in the challenges focus on the features that we want to emphasise, and omit all constructs that would complicate the mechanisation without bringing tangible insights.
+For concurrency experts venturing into mechanisation, our challenges thus serve as good first steps.
+The minimality and uniformity of the calculi also allows us to target design goal~\ref{item:goal-tutorials}.
+For experts in both mechanisation and concurrency, our challenges serve as a framework in which to consider and share best practices and tutorials.
 Aligned with design goal~\ref{item:goal-reusability}, our challenges concern the fundamental meta-theory of process calculi.

### 2024-04-16 challenges/scope-extrusion.tex
Fix a spelling mistake.

@@
 As in the linearity challence,  we first consider 
- the smalleset relation closed under the six axioms of
+ the smallest relation closed under the six axioms of
  structural congruence, denoted by $\sconga\cdot\cdot$:

### 2024-04-17 short-paper/short-paper.tex
Move acknowledgements/disclosures into separate sections as requried by LNCS.

@@
 
-\title{The Concurrent Calculi Formalisation Benchmark\thanks{The work is partially supported by EPSRC EP/T006544/2.}}
+\title{The Concurrent Calculi Formalisation Benchmark}
 
@@
 \and Frederik Krogsdal Jacobsen \inst{5}\orcidID{0000-0003-3651-8314}
-\and Alberto Momigliano\thanks{Member of the Gruppo Nazionale Calcolo Scientifico-Istituto Nazionale di Alta Matematica (GNCS-INdAM).} \inst{6}\orcidID{0000-0003-0942-4777}
+\and Alberto Momigliano \inst{6}\orcidID{0000-0003-0942-4777}
 \and Luca Padovani \inst{7}\orcidID{0000-0001-9097-1297}
@@
 
+\begin{credits}
+\subsubsection{\ackname}
+The work is partially supported by EPSRC EP/T006544/2.
+
+\subsubsection{\discintname}
+Alberto Momigliano is a member of the Gruppo Nazionale Calcolo Scientifico-Istituto Nazionale di Alta Matematica (GNCS-INdAM).
+\end{credits}
+
 \bibliographystyle{splncs04}

### 2024-04-17 short-paper/short-paper.tex
Remove appendix for camera ready version.
Change variant spelling to more common.

@@
 
-Our benchmark challenges (detailed in \cref{app:challenges} and on our website) consider
+Our benchmark challenges (detailed on our website linked below) consider
 \emph{in isolation} three key aspects that frequently pose difficulties
@@
   extrusion by using structural congruence, allowing
-  implementors to explore different ways to encode the latter (\eg
+  implementers to explore different ways to encode the latter (\eg
   via process contexts or compatible refinement);
@@
 
-\clearpage
-\appendix
-\section{Challenges}\label{app:challenges}
-\input{../challenges/challenges.tex}
-
 \end{document}

### 2024-04-17 short-paper/short-paper.tex
minor cosmetic improvements

@@
 
-Our benchmark challenges (detailed on our website linked below) consider
-\emph{in isolation} three key aspects that frequently pose difficulties
-when mechanising concurrency theory, which we will discuss in more detail in the next section: \emph{linearity}, \emph{scope
-  extrusion}, and \emph{coinductive reasoning}.
-Mechanisations must often address several of these aspects at the same time; however, we
-see the combination of techniques as a next
-step, as discussed in \cref{sec:going-beyond}.
+Our benchmark challenges (detailed on our website linked below)
+consider \emph{in isolation} three key aspects that frequently pose
+difficulties when mechanising concurrency theory: \emph{linearity},
+\emph{scope extrusion}, and \emph{coinductive reasoning}, as we will
+discuss in more detail in the next section.  Mechanisations must often
+address several of these aspects at the same time; however, we see the
+combination of techniques as a next step, as argued in
+\cref{sec:going-beyond}.
 
@@
 \subsection{Linearity}
-Linear typing systems allow for the tracking of resource usage in a program. In
+Linear typing systems enables the tracking of resource usage in a program. In
 the case of typed (in particular, session-typed) process calculi, linearity is
@@
 use in formal reasoning in this area~\cite{Despeyroux2000,Honsell2001,Tiu2010,ChaudhuriCM15,Castro-Perez2021}. Its
-weak form aligns well with mainstream inductive proof assistants,
+weak form aligns reasonably well with mainstream inductive proof assistants,
 significantly simplifying the encoding of typing systems and operational
@@
 
-Our coinduction challenge only addresses two notions of process equivalence, but many more exist in the literature.
+Our coinduction challenge only treats two notions of process equivalence, but many more exist in the literature.
 Coinduction may also play a role in recursive processes and session types: recursive session types can be expressed in \emph{infinitary form} by interpreting their typing rules coinductively~\cite{DerakhshanPfenning22,HornePadovani23}.

### 2024-04-18 short-paper/short-paper.tex
removed duplication

@@
 not covered by our benchmark challenges, for example constructs such as
-choice constructs and recursion.
+choice and recursion.
 %  Features that are in wide

### 2024-04-18 short-paper/short-paper.tex
Update e-mail address
small typo

@@
 \and University of Oxford, Oxford, United Kingdom \email{martin.vassor@cs.ox.ac.uk}, \email{nobuko.yoshida@cs.ox.ac.uk}
-\and McGill University, Montreal, Canada \email{daniel.zackon@mcgill.ca}
+\and McGill University, Montreal, Canada \email{daniel.zackon@mail.mcgill.ca}
 }
@@
 \subsection{Linearity}
-Linear typing systems enables the tracking of resource usage in a program. In
+Linear typing systems enable the tracking of resource usage in a program. In
 the case of typed (in particular, session-typed) process calculi, linearity is

### 2024-09-29 slides/talk-proposal-ST30/talk-proposal.tex
cleanup of the repo

--- 2023-07-19 talk-proposal-ST30/talk-proposal.tex
@@
 \usepackage{amsthm}
+\usepackage{mathtools}
 \usepackage{mathpartir}
@@
 \usepackage[T1]{fontenc}
+
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}{Corollary}
 
@@
 The authors of both POPLMark and POPLMark Reloaded were careful to note that their benchmarks were only a beginning, and they specifically mention reasoning about concurrency using coinduction and linear environments as points for future work.
-Papers in concurrency theory do often include mechanised proofs (e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also directly discuss formalisation techniques (e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01, DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}).
+Papers in concurrency theory do often include mechanised proofs (e.g.~\cite{Castro-Perez2021,Castro2020,lmcs:9985, DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also directly discuss formalisation techniques (e.g.~\cite{Bengtson2016, Gay2001, Thiemann2019, DBLP:conf/forte/ZalakainD21}).
 However, our experience is that choosing techniques and tools remains a challenge with few guidelines.

### 2024-09-29 coordination24/short-paper.tex
reorganization

--- 2024-04-18 short-paper/short-paper.tex
@@
 \subsubsection{\ackname}
-The work is partially supported by EPSRC EP/T006544/2.
+The work is partially supported by the UK Engineering and Physical Sciences Research Council (EPSRC) grants EP/T006544/2, EP/V000462/1 and EP/X015955/1; Independent
+Research Fund Denmark RP-1 grant ``Hyben''; and Independent Research
+Fund Denmark RP-1 grant ``MECHANIST''.
 
